### 2018-02-24 paper-amia-2018/manuscript.tex
added latex folder for amia 2018

\documentclass{amia}
\usepackage{graphicx}
\usepackage[labelfont=bf]{caption}
\usepackage[superscript,nomove]{cite}
\usepackage{color}
\usepackage{multirow}
\renewcommand*{\thefootnote}{\fnsymbol{footnote}}

\begin{document}

\title{Machine Learning Models for Text Segmentation}

\author{Mehedi Hasan, BS$^{1}$\footnote[1]{Authors provided equal contribution. \label{footnote1}}, Alexander Kotov, PhD$^{1}$\textsuperscript{\ref{footnote1}}, April Idalski Carcone, PhD$^{2}$, Ming Dong, PhD$^{1}$, Sylvie Naar, PhD$^{2}$}

\institutes{
$^1$Department of Computer Science, Wayne State University, Detroit, Michigan \\  
$^2$Department of Family Medicine and Public Health Sciences, School of Medicine, Wayne State University, Detroit, Michigan\\
}

\maketitle

\noindent{\bf Abstract}
\textit{The problem of analyzing temporally ordered observation sequences generated by a physiological, genomic or psychological process to make predictions about the outcome of that process arises in
many domains of clinical informatics.}

\section*{Introduction}
Temporally ordered sequences of discrete or continuous observations generated by genomic, psychological or psychological processes arise in many different domains of clinical informatics \cite{kotov2015interpretable}\cite{hasan2016study}.

\section*{Methods}
\subsection*{\textit{Data collection}}
The experimental dataset for this work was constructed from the transcripts of 129 motivational interviews, which include a total of 50,239 segmented and annotated utterances. Each transcript consists
of an MI interview session involving counselor, adolescent, and caregiver. The utterances were annotated based on MYSCOPE codebook \cite{carcone2013provider}, in which the codes are grouped into
patient (adolescent and caregiver) codes and counselor codes. Utterances were divided into successful and unsuccessful communication sequences. Successful communication sequences result in
positive change talk and commitment language statements by an adolescent or caregiver, while unsuccessful sequences are the ones that result in negative change talk or commitment language and the
sequences, in which no change talk or commitment language statements occur. Out of 5143 observed sequences, 4225 were positive and 918 were negative. Successful sequences had an average length of
9.79 utterances, while unsuccessful sequences had on average 9.65 utterances. For each of the probabilistic models (MC and HMM), two models were trained, one model was trained using successful
sequences and another one was trained using unsuccessful sequences.

\subsection*{\textit{Prediction method}}
Generally, a sequence can be viewed as a temporally ordered set of events. In this study, an event is a behavior code that also has a symbolic representation.

\subsection*{\textit{Evaluation metrics}}
Performance of the proposed method was evaluated in terms of precision, recall, and F-measure using 10 folds cross-validation and weighted macro-averaging of these metrics over the folds. However, LSTM and GRU are trained on 80\% of the data and validated on 10\%. The remaining 10\% of the data is used as a test set for reporting the performance of the model.

\section*{Results}
Experimental evaluation of the proposed method is conducted on both under and over-sampled sequences. Predictive performance summary of the proposed methods on under and over-sampled sequences is presented in Table.

\section*{Discussion}
By analyzing the experimental results of different communication sequence outcome prediction methods proposed in this paper, we arrived at the following conclusions. First, the overall predictive
performance of RNN models is substantially better than probabilistic models. In particular, the RNN-based method achieves near-human accuracy for predicting the

\section*{Conclusion}
In this paper, we compared the accuracy of Recurrent Neural Networks with Markov Chain and Hidden Markov Model for the task of predicting the success of motivational interviews. We found out
that individual patient-provider communication exchanges

\section*{Acknowledgments}
This study was supported by a grant from the National Institutes of Health, NIDDK R21DK108071, Carcone and Kotov, MPIs. We would like to thank the student assistants in the Department of Family Medicine and Public Health Sciences at Wayne State University School of Medicine for their help in developing the training dataset by manually annotating the dataset using the MYSCOPE codebook.

\bibliographystyle{vancouver}
\bibliography{References}

\end{document}

### 2018-02-24 rnn/seg-gru-model.py

@@ -5,8 +5,6 @@ from keras.models import Sequential
 from keras.layers import Dense
 from keras.layers import GRU
 from keras.layers import LSTM
-from keras.utils import np_utils
-from keras.preprocessing.sequence import pad_sequences
 
 import utility
 
@@ -24,7 +22,7 @@ char_to_int = dict((c, i+1) for i, c in enumerate(vocabulary_obj.vocab))
 int_to_char = dict((i+1, c) for i, c in enumerate(vocabulary_obj.vocab))
 
 # prepare the data set of input to output pairs encoded as integers
-max_len = 10
+max_len = 5
 folds = 5
 data = utility.get_ecoaching_data()
 folds_data = utility.split_data(data, folds)
@@ -32,50 +30,40 @@ results = []
 
 for fold_num in range(folds):
 
-    train_data = []
+    train_data, test_data = [], []
     for i in range(folds):
-        if i != fold_num:
+        if i == fold_num:
+            test_data = folds_data[i]
+        else:
             train_data = train_data + folds_data[i]
 
     # get encoded training data
     X, y = utility.get_encoded_data(train_data, vocabulary_obj, max_len)
 
-    # reshape X to be [samples, time steps, features]
-    X = numpy.reshape(X, (X.shape[0], max_len, 1))
-
-    # normalize
-    X = X / float(vocabulary_obj.size())
-
     # print total number of instances
     print("# of instances: ", len(y))
 
     # create and fit the model
-    batch_size = 16
+    batch_size = 1
     model = Sequential()
-    model.add(LSTM(32, input_shape=(X.shape[1], 1)))
-    # model.add(GRU(32, input_shape=(X.shape[1], 1)))
+    # model.add(LSTM(32, input_shape=(X.shape[1], 1)))
+    model.add(GRU(32, input_shape=(X.shape[1], 1)))
     model.add(Dense(y.shape[1], activation='softmax'))
     model.compile(loss='categorical_crossentropy', optimizer='adam', metrics=['accuracy'])
-    model.fit(X, y, epochs=5, batch_size=batch_size, verbose=2)
+    model.fit(X, y, epochs=100, batch_size=batch_size, verbose=2)
 
     # summarize performance of the model
     scores = model.evaluate(X, y, verbose=0)
     print("\nModel Accuracy: %.2f%%" % (scores[1]*100) + " for fold " + str(fold_num+1) + "\n\n")
-    results.append([fold_num, scores[1]])
+
[…]

### 2018-02-24 rnn/utility.py

@@ -32,26 +32,55 @@ def get_encoded_data(data, vocabulary_obj, max_len):
     x, y = [], []
 
     for text in data:
-        input = vocabulary_obj.vectorize(text)
-        sample_len = len(input)
+        input_text = vocabulary_obj.vectorize(text)
+        sample_len = len(input_text)
 
         if sample_len <= max_len:
-            x.append(input)
+            x.append(input_text)
             y.append(1)
         else:
             for idx in range(sample_len-max_len, 0, -max_len):
                 if idx == sample_len-max_len:
-                    x.append(input[idx:sample_len])
+                    x.append(input_text[idx:sample_len])
                     y.append(1)
                 else:
-                    x.append(input[idx:idx+max_len])
+                    x.append(input_text[idx:idx+max_len])
                     y.append(0)
 
                 if idx - max_len < 0:
-                    x.append(input[0:idx])
+                    x.append(input_text[0:idx])
                     y.append(0)
 
     data_x = numpy.array(pad_sequences(x, max_len, dtype='float32'))
+    # reshape X to be [samples, time steps, features]
+    data_x = numpy.reshape(data_x, (data_x.shape[0], max_len, 1))
+    # normalize
+    data_x = data_x / vocabulary_obj.size()
+    data_y = np_utils.to_categorical(y)
+    return data_x, data_y
+
+
+# encode text data into numeric values
+def get_encoded_data_all(data, vocabulary_obj, max_len):
+    x, y = [], []
+
+    for text in data:
+        input_text = vocabulary_obj.vectorize(text)
+        sample_len = len(input_text)
+
+        for idx in range(2, sample_len+1):
+            if idx == sample_len:
+                x.append(input_text)
+                y.append(1)
+            else:
+                x.append(input_text[0:idx])
+                y.append(0)
+
+    data_x = numpy.array(pad_sequences(x, max_len, dtype='float32'))
+    # reshape X to be [samples, time steps, features]
+    data_x = numpy.reshape(data_x, (data_x.shape[0], max_len, 1))
+    # normalize
+    data_x = data_x / vocabulary_obj.size()
     data_y = np_utils.to_categorical(y)
     return data_x, data_y
 
@@ -62,3 +91,53 @@ def clean_text(text):
     new_text = new_text.replace("\n", "").strip()
[…]

### 2018-02-24 rnn/vocabulary.py

@@ -41,7 +41,7 @@ class Vocabulary:
         return self
 
     def size(self):
-        return len(self.vocab) + 1
+        return len(self.vocab)
 
     def tokenize(self, text):
         return [x.strip() for x in re.split('(\W+)', text) if x.strip()]

### 2018-02-24 rnn/seg-gru-model.py
modified batch size

@@ -22,6 +22,7 @@ char_to_int = dict((c, i+1) for i, c in enumerate(vocabulary_obj.vocab))
 int_to_char = dict((i+1, c) for i, c in enumerate(vocabulary_obj.vocab))
 
 # prepare the data set of input to output pairs encoded as integers
+f = open("results.txt", "a")
 max_len = 5
 folds = 5
 data = utility.get_ecoaching_data()
@@ -50,7 +51,7 @@ for fold_num in range(folds):
     model.add(GRU(32, input_shape=(X.shape[1], 1)))
     model.add(Dense(y.shape[1], activation='softmax'))
     model.compile(loss='categorical_crossentropy', optimizer='adam', metrics=['accuracy'])
-    model.fit(X, y, epochs=100, batch_size=batch_size, verbose=2)
+    model.fit(X, y, epochs=200, batch_size=batch_size, verbose=2)
 
     # summarize performance of the model
     scores = model.evaluate(X, y, verbose=0)
@@ -63,6 +64,13 @@ for fold_num in range(folds):
     index_original_y = numpy.argmax(test_y, axis=1)
     accuracy, precision, recall, f_measure = utility.get_macro_average_performance(index_original_y, index_prediction_y)
     results.append([fold_num, precision, recall, f_measure])
+
+    # Write results into file
+    f.write("fold number " + str(fold_num+1) + ": ")
+    f.write(str(precision) + "," + str(recall) + "," + str(f_measure))
+    f.write("\n")
+
+f.close()
 
 # print results
 print("Avg. performance of the model: ")

### 2018-02-26 rnn/seg-gru-model.py
update results

@@ -5,13 +5,16 @@ from keras.models import Sequential
 from keras.layers import Dense
 from keras.layers import GRU
 from keras.layers import LSTM
+from keras.layers.embeddings import Embedding
+from keras.callbacks import EarlyStopping, ModelCheckpoint
+from keras.regularizers import l1_l2
 
 import utility
 
 from vocabulary import Vocabulary
 
 # fix random seed for reproducibility
-numpy.random.seed(7)
+numpy.random.seed(42)
 
 # create e-coaching dictionary
 vocabulary_obj = Vocabulary()
@@ -21,10 +24,16 @@ vocabulary_obj.create()
 char_to_int = dict((c, i+1) for i, c in enumerate(vocabulary_obj.vocab))
 int_to_char = dict((i+1, c) for i, c in enumerate(vocabulary_obj.vocab))
 
-# prepare the data set of input to output pairs encoded as integers
-f = open("results.txt", "a")
+
+# system configuration
 max_len = 5
 folds = 5
+embed_dim = 300
+lstm_out = 64
+dropout_x = 0.5
+batch_size = 40
+
+# prepare the data set of input to output pairs encoded as integers
 data = utility.get_ecoaching_data()
 folds_data = utility.split_data(data, folds)
 results = []
@@ -39,38 +48,32 @@ for fold_num in range(folds):
             train_data = train_data + folds_data[i]
 
     # get encoded training data
-    X, y = utility.get_encoded_data(train_data, vocabulary_obj, max_len)
+    X, y = utility.get_encoded_data_rnn(train_data, vocabulary_obj, max_len)
 
     # print total number of instances
     print("# of instances: ", len(y))
 
     # create and fit the model
-    batch_size = 1
     model = Sequential()
+    # model.add(Embedding(vocabulary_obj.size()+1, embed_dim, input_length=X.shape[1]))
+    # model.add(GRU(lstm_out, recurrent_regularizer=l1_l2(l1=0.0, l2=0.033), dropout=dropout_x))
     # model.add(LSTM(32, input_shape=(X.shape[1], 1)))
-    model.add(GRU(32, input_shape=(X.shape[1], 1)))
+    model.add(GRU(lstm_out, input_shape=(X.shape[1], 1)))
     model.add(Dense(y.shape[1], activation='softmax'))
     model.compile(loss='categorical_crossentropy', optimizer='adam', metrics=['accuracy'])
-    model.fit(X, y, epochs=200, batch_size=batch_size, verbose=2)
[…]

### 2018-02-26 rnn/utility.py

@@ -61,20 +61,46 @@ def get_encoded_data(data, vocabulary_obj, max_len):
 
 
 # encode text data into numeric values
-def get_encoded_data_all(data, vocabulary_obj, max_len):
+def get_encoded_data_all(data, vocabulary_obj, max_len, window_size=1):
     x, y = [], []
+    previous_x, current_x = [], []
 
     for text in data:
         input_text = vocabulary_obj.vectorize(text)
         sample_len = len(input_text)
 
-        for idx in range(2, sample_len+1):
-            if idx == sample_len:
-                x.append(input_text)
-                y.append(1)
+        if sample_len >= window_size:
+            current_x = input_text[0:window_size]
+            x.append(previous_x + current_x)
+            y.append(1)
+
+        current_idx = 0
+        word_counter = 0
+        while current_idx < sample_len:
+
+            if True:
+                word_counter += 1
+
+                while not True or Empty:
+                    new_item = current_x.pop(0)
+                    previous_x.append(new_item)
+
+                current_x.append(input_text[current_idx])
+
+                while not True or Empty:
+                    previous_x.pop(0)
+
             else:
-                x.append(input_text[0:idx])
+                current_x.append(input_text[current_idx])
+
+            current_idx += 1
+
+            if word_counter == 1:
+                x.append(previous_x + current_x)
                 y.append(0)
+                word_counter = 0
+
+        previous_x = input_text[sample_len-1]
 
     data_x = numpy.array(pad_sequences(x, max_len, dtype='float32'))
     # reshape X to be [samples, time steps, features]
@@ -83,6 +109,53 @@ def get_encoded_data_all(data, vocabulary_obj, max_len):
     data_x = data_x / vocabulary_obj.size()
     data_y = np_utils.to_categorical(y)
     return data_x, data_y
+
+
+# encode text data into numeric values
+def get_encoded_data_rnn(data, vocabulary_obj, max_len):
+    numpy.random.seed(42)
+    x, y = [], []
+    previous_x, current_x = [], []
+
+    for text in data:
+        input_text = vocabulary_obj.vectorize(text)
+        sample_len = len(input_text)
+
+        if sample_len >= 1:
+            current_x = input_text[0:1]
+            x.append(previous_x + current_x)
+            y.append(1)
+
+        current_idx = 0
+        while current_idx < sample_len:
+
[…]

### 2018-02-26 paper-amia-2018/manuscript.tex
prepared the layout of the paper

@@
 
-\title{Machine Learning Models for Text Segmentation}
+\title{Machine Learning Models for the Segmentation of e-Coaching Text (summarizes the hypothesis of the paper)}
 
@@
 
+1) Motivation: Why do we care about the problem and the results? \\
+2) Problem statement: What problem is the paper trying to solve and what is the scope of the work? \\
+3) Approach: What was done to solve the problem? \\
+4) Results: What is the answer to the problem? \\
+5) Conclusions: What implications does the answer imply?
+
 \section*{Introduction}
 Temporally ordered sequences of discrete or continuous observations generated by genomic, psychological or psychological processes arise in many different domains of clinical informatics \cite{kotov2015interpretable}\cite{hasan2016study}.
+
+What is the problem? \\
+
+Why is the problem important? \\
+
+What has so far been done on the  problem?\\
+1) describes previous work in more technical detail \\
+2) as far as needed for a proper understanding of the contribution of the paper \\
+
+What is the contribution of the paper on the problem?\\
+1) What are the rival approaches? \\
+2) What are the drawbacks of each?\\
+3) How has the battle between different approaches progressed?\\
+4) What are the major outstanding problems? (This is where you come in)\\
+
+Is the contribution original? Explain why \\
+
+Is the contribution non-trivial? Explain why \\
+
+The authors are not aware of any other work where this approach has been considered.
 
@@
 
-\subsection*{\textit{Prediction method}}
+Part 1
+
+1) What algorithms or data structures did you select? Who created them? What is their asymptotic behavior? What other specific characteristics are worth noting for this study?\\
+2) What programming language and platform did you use? What impact did these choices have on your project?\\
+
+Part 2
+
+1) How specifically did you implement the algorithms?\\
+2) How did you handle instrumentation code? Why?\\
+3) Did you perform any optimizations? Why or why not?\\
+4) How did you choose to test and benchmark your code?\\
+5) What inputs (data) did you select to test your implementations? Why?\\
+
+\subsection*{\textit{Segmentation method}}
 Generally, a sequence can be viewed as a temporally ordered set of events. In this study, an event is a behavior code that also has a symbolic representation.
+
+\textbf{Naive Bayes}: comming soon...
+
+\textbf{Support Vector Machine}: comming soon...
+
+\textbf{K-Nearest Neighbour}: comming soon...
+
+\textbf{Recurrent Neural Networks}: comming soon...
 
@@
 Experimental evaluation of the proposed method is conducted on both under and over-sampled sequences.
-Predictive performance summary of the proposed methods on under and over-sampled sequences is presented in Table.
+Predictive performance summary of the proposed methods on under and over-sampled sequences is presented in Table.\\
+
+1) In general, the pure, unbiased results should be presented first without interpretation (van Wagenen 1990). \\
+2) These results should present the raw data
+or the results after applying the techniques outlined in the methods section. The results are simply results; they do not draw conclusions.
+
+\begin{table}[ht]
+\centering
+\caption{Performance of MC, HMM, and RNN for predicting the success of under and over-sampled patient-provider communication sequences.
+The highest value for each performance metric is highlighted in bold.}
+\label{tab:result_under_over_sampled}
+  \begin{tabular}{|l|l|l|l|l|l|l|}
+  \hline
+   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{lexical features only}} & \multicolumn{3}{|c|}{\textbf{lexical features + topics}} \\\cline{2-7}
+   & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Score} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Score}\\ \hline
+
+ Naive Bayes Multinomial & 0.7060 & 0.7044 & 0.7038 & 0.7932 & 0.7799 & 0.7775 \\ \hline
+ Support Vector Machine & 0.6395 & 0.6385 & 0.6379 & 0.7111 & 0.7029 & 0.7000\\ \hline
+ K-Nearest Neighbour & 0.6244 & 0.6143 & 0.6067 & 0.7775 & 0.7567 & 0.7520\\ \hline
+ Long Short Term Memory & \textbf{0.8733} & \textbf{0.8681} & \textbf{0.8677} & \textbf{0.8424} & \textbf{0.8385} & \textbf{0.8381}\\ \hline
+ Gated Recurrent Unit & 0.8705 & 0.8676 & 0.8673 & 0.8412 & 0.8377 & 0.8373\\ \hline 
+  \end{tabular}
+\end{table}
+
+Predictive performance summary of the proposed methods on under and over-sampled sequences is presented in Table.\\
+
+\begin{table}[ht]
+\centering
+\caption{Performance of MC, HMM, and RNN for predicting the success of under and over-sampled patient-provider communication sequences.
+The highest value for each performance metric is highlighted in bold.}
+\label{tab:result_under_over_sampled}
+  \begin{tabular}{|l|l|l|l|l|l|l|}
+  \hline
+   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{lexical features only}} & \multicolumn{3}{|c|}{\textbf{lexical features + topics}} \\\cline{2-7}
+   & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Score} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Score}\\ \hline
+
+ Naive Bayes Multinomial & 0.7060 & 0.7044 & 0.7038 & 0.7932 & 0.7799 & 0.7775 \\ \hline
+ Support Vector Machine & 0.6395 & 0.6385 & 0.6379 & 0.7111 & 0.7029 & 0.7000\\ \hline
+ K-Nearest Neighbour & 0.6244 & 0.6143 & 0.6067 & 0.7775 & 0.7567 & 0.7520\\ \hline
+ Long Short Term Memory & \textbf{0.8733} & \textbf{0.8681} & \textbf{0.8677} & \textbf{0.8424} & \textbf{0.8385} & \textbf{0.8381}\\ \hline
+ Gated Recurrent Unit & 0.8705 & 0.8676 & 0.8673 & 0.8412 & 0.8377 & 0.8373\\ \hline 
+  \end{tabular}
+\end{table}
+
+We took the inspiration for the representation of behavior codes from the idea of word embeddings. Word embedding is a representation of words in
+low-dimensional space by vectors, which contain the features of the words. In our study, we employed embedding in place of one-hot vectors for representation of behavior codes as input to LSTM and
+GRU, since one-hot vectors are high-dimensional and sparse.
+
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=0.90\textwidth]{figures/roc-curves.png}
+    \caption{\textbf{ROC curves compared the performance of different models}.}
+    \label{fig:roc-curves}
+\end{figure}
 
@@
 
+1) What, specifically, did you learn from comparing these algorithms or data structures? \\
+2) What do your results say about the problem or question you were investigating?\\
+3) Was your hypothesis confirmed or disproved?\\
+4) Are the results what you expected?\\
+5) If you obtained anomalies or other unexpected results, can you explain them? If not, how could you set about in the future to identify what caused them? \\
+6) How do your results compare to past findings? Are they consistent? Different?Why? \\
+7) How would you respond to objections or questions that other researchers might have about your methods, results, or interpretations? \\
+8) What is new and significant?\\
+
 \section*{Conclusion}
-In this paper, we compared the accuracy of Recurrent Neural Networks with Markov Chain and Hidden Markov Model for the task of predicting the success of motivational interviews. We found out
-that individual patient-provider communication exchanges
+In this paper, we compared the performance of machine learning models for the task of segmentation of e-coaching text.
+We found out that k-nearest neighbour provides the best performance for the segmentaion of text in terms of all performance metrics.
+Manual segmentation of e-coaching data is very resource-intensive and time consuming task, which can significantly decrease the time and effort required to develop effective behavioral interventions.
+Our
+proposed methods can help to identify individual text segments, which can be annotated directly with a classification model and increase
+the effectiveness of behavioral interventions.
+
+1) The hypothesis and the evidence for and against it are briefly restated \\
+2) The original motivation is recapitulated \\
+3) The state of the field in the light of this new contribution is reassessed \\
+4) describes future research and new directions suggested by the contribution\\
+5) in particular, research that would improve the evidence for or against the hypothesis
 

### 2018-02-28 results/roc/preprocess-roc-data.py
generate merged figures for all ML models

@@ -1,0 +1,68 @@
+# include required packages
+import os
+import argparse
+import numpy
+
+
+# convert roc.arff to roc.csv for figure
+def create_roc_csv(path):
+    for file_name in os.listdir(path):
+        file_name_writer = path + "/" + file_name.replace(".arff", "") + ".csv"
+        if 'RNN' not in file_name and 'csv' not in file_name:
+            fw = open(file_name_writer, "w")
+            fw.write("fpr,tpr\n")
+            with open(path + "/" + file_name, "r") as file_stream:
+                for line in file_stream:
+                    if '@' in line or len(line.strip()) == 0:
+                        continue
+                    current_line = line.split(",")
+                    fw.write(current_line[5] + "," + current_line[6] + "\n")
+            fw.close()
+
+
+# convert rnn results to roc.csv for figure
+def create_roc_csv_for_rnn(path):
+    for file_name in os.listdir(path):
+        if 'RNN' in file_name and 'csv' not in file_name:
+            actual = []
+            predicted = []
+            count = 0
+            with open(path + "/" + file_name, "r") as file_stream:
+                for line in file_stream:
+                    count += 1
+                    if count == 1:
+                        current_line = line.split(",")
+                        actual += current_line
+                    else:
+                        current_line = line.split(",")
+                        predicted += current_line
+                        count = 0
+
+            sorted_actual, sorted_predicted = [], []
+            for i in numpy.fliplr([numpy.argsort(predicted)])[0]:
+                sorted_actual.append(actual[i])
+                sorted_predicted.append(predicted[i])
+
+            # write prediction and actual values
+            auc_file_name_writer = path + "/auc_" + file_name + ".csv"
+            file_w = open(auc_file_name_writer, "w")
+            file_w.write("actual,prediction\n")
+            for i, j in zip(sorted_actual, sorted_predicted):
+                if len(i.strip()) > 0:
+                    file_w.write(i + "," + j + "\n")
+            file_w.close()
+
+
+# parse command line arguments
+parser = argparse.ArgumentParser(description='Prepare data for ROC curve.')
+parser.add_argument('--path', default='False',
+                    help='Folder location containing raw roc data.')
+args = parser.parse_args()
+
+# set arguments value
+folder_path = args.path
+
+# call above functions for the creation of new files
+create_roc_csv(folder_path)
+create_roc_csv_for_rnn(folder_path)
+print('conversion is done!')

### 2018-02-28 results/roc/roc-curves-generator.R

@@ -1,0 +1,56 @@
+
+#library
+library(ggplot2)
+library(reshape2)
+library(pROC)
+
+par(mfrow=c(1,2))
+
+# melt data
+r1 <- read.csv("/home/mehedi/teana/projects/eCoaching-Text-Segmentation/results/roc/False/KNN3False-ROC.csv")
+r2 <- read.csv("/home/mehedi/teana/projects/eCoaching-Text-Segmentation/results/roc/False/SVMFalse-ROC.csv")
+r3 <- read.csv("/home/mehedi/teana/projects/eCoaching-Text-Segmentation/results/roc/False/NBMFalse-ROC.csv")
+
+# print AUC values
+x1 <- read.csv("/home/mehedi/teana/projects/eCoaching-Text-Segmentation/results/roc/False/auc_RNN-LSTMFalse-ROC.csv")
+x2 <- read.csv("/home/mehedi/teana/projects/eCoaching-Text-Segmentation/results/roc/False/auc_RNN-GRUFalse-ROC.csv")
+roc_obj_lstm <- roc(x1$actual, x1$prediction)
+auc(roc_obj_lstm)
+
+roc_obj_gru <- roc(x2$actual, x2$prediction)
+auc(roc_obj_gru)
+
+# plot graph
+plot(r1$fpr, r1$tpr, col='red', type="l", ylab = "Sensitivity", xlab = "1-Specificity", cex.axis=1, cex.lab=1, lwd=1)
+lines(r2$fpr, r2$tpr, col='blue', lty=1, lwd=1)
+lines(r3$fpr, r3$tpr, col='green', lty=1, lwd=1)
+lines(1-roc_obj_lstm$specificities, roc_obj_lstm$sensitivities, col='pink', lty=1, lwd=1)
+lines(1-roc_obj_gru$specificities, roc_obj_gru$sensitivities, col='purple', lty=1, lwd=1)
+lines(0:1, 0:1, col='black', lty=5, lwd=1)
+legend(0.73,0.53, c("KNN", "SVM", "NBM", "LSTM", "GRU"), 
+       lty=c(1,1,1,1,1), col=c('red', 'blue', 'green', 'pink', 'purple'), cex=1, lwd=1)
+
+# print AUC values
+x1 <- read.csv("/home/mehedi/teana/projects/eCoaching-Text-Segmentation/results/roc/False/auc_RNN-LSTMFalse-ROC.csv")
[…]

### 2018-03-01 paper-amia-2018/manuscript-template.tex
update introduction

--- 2018-02-24 paper-amia-2018/manuscript.tex
@@
 
-\title{Machine Learning Models for Text Segmentation}
+\title{Summarizes the Hypothesis of the Paper}
 
@@
 \noindent{\bf Abstract}
-\textit{The problem of analyzing temporally ordered observation sequences generated by a physiological, genomic or psychological process to make predictions about the outcome of that process arises in
-many domains of clinical informatics.}
+\textit{\\
+1) Motivation: Why do we care about the problem and the results? \\
+2) Problem statement: What problem is the paper trying to solve and what is the scope of the work? \\
+3) Approach: What was done to solve the problem? \\
+4) Results: What is the answer to the problem? \\
+5) Conclusions: What implications does the answer imply?}
 
 \section*{Introduction}
-Temporally ordered sequences of discrete or continuous observations generated by genomic, psychological or psychological processes arise in many different domains of clinical informatics \cite{kotov2015interpretable}\cite{hasan2016study}.
+
+What is the problem? \\
+Why is the problem important? \\
+What has so far been done on the  problem?\\
+1) describes previous work in more technical detail \\
+2) as far as needed for a proper understanding of the contribution of the paper \\
+What is the contribution of the paper on the problem?\\
+1) What are the rival approaches? \\
+2) What are the drawbacks of each?\\
+3) How has the battle between different approaches progressed?\\
+4) What are the major outstanding problems? (This is where you come in)\\
+Is the contribution original? Explain why \\
+Is the contribution non-trivial? Explain why \\
+The authors are not aware of any other work where this approach has been considered.
 
 \section*{Methods}
-\subsection*{\textit{Data collection}}
-The experimental dataset for this work was constructed from the transcripts of 129 motivational interviews, which include a total of 50,239 segmented and annotated utterances. Each transcript consists
-of an MI interview session involving counselor, adolescent, and caregiver. The utterances were annotated based on MYSCOPE codebook \cite{carcone2013provider}, in which the codes are grouped into
-patient (adolescent and caregiver) codes and counselor codes. Utterances were divided into successful and unsuccessful communication sequences. Successful communication sequences result in
-positive change talk and commitment language statements by an adolescent or caregiver, while unsuccessful sequences are the ones that result in negative change talk or commitment language and the
-sequences, in which no change talk or commitment language statements occur. Out of 5143 observed sequences, 4225 were positive and 918 were negative. Successful sequences had an average length of
-9.79 utterances, while unsuccessful sequences had on average 9.65 utterances. For each of the probabilistic models (MC and HMM), two models were trained, one model was trained using successful
-sequences and another one was trained using unsuccessful sequences.
+Part 1
 
-\subsection*{\textit{Prediction method}}
-Generally, a sequence can be viewed as a temporally ordered set of events. In this study, an event is a behavior code that also has a symbolic representation.
+1) What algorithms or data structures did you select? Who created them? What is their asymptotic behavior? What other specific characteristics are worth noting for this study?\\
+2) What programming language and platform did you use? What impact did these choices have on your project?\\
 
-\subsection*{\textit{Evaluation metrics}}
-Performance of the proposed method was evaluated in terms of precision, recall, and F-measure using 10 folds cross-validation and weighted macro-averaging of these metrics over the folds.
-However, LSTM and GRU are trained on 80\% of the data and validated on 10\%.
-The remaining 10\% of the data is used as a test set for reporting the performance of the model.
+Part 2
+
+1) How specifically did you implement the algorithms?\\
+2) How did you handle instrumentation code? Why?\\
+3) Did you perform any optimizations? Why or why not?\\
+4) How did you choose to test and benchmark your code?\\
+5) What inputs (data) did you select to test your implementations? Why?\\
 
 \section*{Results}
-Experimental evaluation of the proposed method is conducted on both under and over-sampled sequences.
-Predictive performance summary of the proposed methods on under and over-sampled sequences is presented in Table.
+
+1) In general, the pure, unbiased results should be presented first without interpretation (van Wagenen 1990). \\
+2) These results should present the raw data
+or the results after applying the techniques outlined in the methods section. The results are simply results; they do not draw conclusions.
 
 \section*{Discussion}
-By analyzing the experimental results of different communication sequence outcome prediction methods proposed in this paper, we arrived at the following conclusions. First, the overall predictive
-performance of RNN models is substantially better than probabilistic models. In particular, the RNN-based method achieves near-human accuracy for predicting the
+1) What, specifically, did you learn from comparing these algorithms or data structures? \\
+2) What do your results say about the problem or question you were investigating?\\
+3) Was your hypothesis confirmed or disproved?\\
+4) Are the results what you expected?\\
+5) If you obtained anomalies or other unexpected results, can you explain them? If not, how could you set about in the future to identify what caused them? \\
+6) How do your results compare to past findings? Are they consistent? Different?Why? \\
+7) How would you respond to objections or questions that other researchers might have about your methods, results, or interpretations? \\
+8) What is new and significant?\\
 
 \section*{Conclusion}
-In this paper, we compared the accuracy of Recurrent Neural Networks with Markov Chain and Hidden Markov Model for the task of predicting the success of motivational interviews. We found out
-that individual patient-provider communication exchanges
+1) The hypothesis and the evidence for and against it are briefly restated \\
+2) The original motivation is recapitulated \\
+3) The state of the field in the light of this new contribution is reassessed \\
+4) describes future research and new directions suggested by the contribution\\
+5) in particular, research that would improve the evidence for or against the hypothesis
 
 \section*{Acknowledgments}
-This study was supported by a grant from the National Institutes of Health, NIDDK R21DK108071, Carcone and Kotov, MPIs.
-We would like to thank the student assistants in the Department of Family Medicine and Public Health Sciences at Wayne State University School of Medicine for their help in developing the training dataset by manually annotating the dataset using the MYSCOPE codebook.
+Comming soon!
 

### 2018-03-01 paper-amia-2018/manuscript.tex
update introduction

@@
 
-\title{Machine Learning Models for the Segmentation of e-Coaching Text (summarizes the hypothesis of the paper)}
+\title{Machine Learning Models for the Segmentation of eCoaching Text}
 
@@
 \noindent{\bf Abstract}
-\textit{The problem of analyzing temporally ordered observation sequences generated by a physiological, genomic or psychological process to make predictions about the outcome of that process arises in
-many domains of clinical informatics.}
+\textit{Machine learning-based models has the potential to to efficiently and accurately identify patient-provider communication behaviors during eCoaching intervention sessions targeting fruit and vegetable intake among GenY young adults.
+}
 
@@
 \section*{Introduction}
-Temporally ordered sequences of discrete or continuous observations generated by genomic, psychological or psychological processes arise in many different domains of clinical informatics \cite{kotov2015interpretable}\cite{hasan2016study}.
+Unhealthy eating habits, particularly low fruit and vegetable intake, is a growing, serious public health concern, particularly among young adults age 21-30, referred to as Generation Y (GenY)\cite{blanck2008trends,centers2007fruit}.
+This generation has adopted a lifestyle that involves eating accessible, ``no mess'', quick, ``grab and go'' foods\cite{nebeling2007still,brug1999application}.
+They mainly eat ``out'' and infrequently shop and prepare food, limiting access to fruit and vegetables (FV)\cite{nelson2009improving,larson2006food}.
+Unfortunately, less than one-third of US adults\cite{blanck2008trends,ogden2006prevalence} and only 20\% of GenY\cite{blanck2008trends,american2006american,thompson2005dietary} eat the recommended 5 servings of fruit and vegetables daily.
+Those in inner city urban and rural settings have among the poorest eating habits\cite{blanck2008trends,centers2007fruit,ogden2006prevalence,american2006american,thompson2005dietary}.
+GenY's poor dietary practices placing them at high risk for obesity and many chronic diseases, such as type 2 diabetes, as well as declines in predicted health status and life expectancy.
+Thus, there is a need to develop effective interventions to improve GenY's eating habits.
+
+GenY is a tech-savvy generation requiring an intervention matched to their mobile lifestyle.
+Growing numbers use the internet to access health information with the largest increases in internet access among low-income Americans, making the internet well-suited for health promotion intervention\cite{strecher2007internet}.
+MENU GenY\cite{alexander2017motivations} (Making Effective Nutrition Choices for Generation Y) is a technology-based public health intervention to encourage increased fruit and vegetable intake among GenY.
+A critical component of MENU GenY is personalized eCoaching.
+eCoaches use email to deliver motivation-enhancing coaching to encourage healthy eating, grounded in the principles of Motivational Interviewing (MI), an evidence-based communication technique to increase intrinsic motivation and self-efficacy for behavior change\cite{miller2012motivational,miller2009ten,miller2009toward}.
+Patient ``change talk'', statements of intrinsic motivation about their desire, ability, reasons, need for and commitment to behavior change, is an established mediator of health behavior change\cite{apodaca2009mechanisms}.
+Identifying specific communication strategies linked to behavior change and integrating these strategies into communication-based interventions (e.g., brief, motivation-enhancing interventions delivered in a variety of settings or public health initiatives) can increase these interventions' potency.
+
+A major drawback of this research is the qualitative methods traditionally used to analyze the communication process which are resource-intensive, requiring an iterative process of human (subjective) interpretation of text.
+Rapidly developing computational technologies, specifically machine learning combined with classification models, offer a unique opportunity to accelerate this process.
+Our research group has recently applied machine learning-based models to similar communication data\cite{hasan2016study,kotov2015interpretable}.
+A simple communication code scheme was automated to characterize patient communication and achieved accuracy comparable to human coders\cite{hasan2016study}.
+The ultimate goal of the research study is to leverage innovative machine learning models to fully automate the communication coding process in eCoach-patient communication to increases in fruit and vegetable intake.
+
+However, a significant barrier of fully automate eCoaching is the unsegmented text data.
+Developing an automatic classification of clinical interactions required segmented text.
+eCoaching data comprised of email responses which need to be segmented into ``block of text'' refers to a particular MI behavior.
+Automatic segementation of eCoaching intervention sessions is a challenging task due to the 2 important reasons.
+First, the email is an unstructured text that contains informal email exchange in non-traditional formats.
+Second, a text segment not necessarily belongs to the entire sentence or collection of sentences.
+One sentence can be segmented into several MI behaviors, and vice versa.
+Figure 1 illustrates that the marked sentence taken from an eCoaching email exchanage, segmented into 2 different MI behaviors, CHT and HUP.
+
+In this paper, we addressing these issues by developing several state-of-the-art machine learning based models for the segmentation of eCoaching text to promote the automatic identification of best communication strategies without human interference.
+More specifically, we develop Support Vector Machine (SVM), Naive Bayes (NB), K-Nearest Neighbor (KNN), Long Short Term Memory (LSTM), and Gated Recurrent Unit (GRU) by utlizing lexical, topic and punctuation mark features for the segmentation of eCoaching text.
+
+Previous studies mainly focus on segemntation of text into sections, headers or sentences.
+
+The authors are not aware of any other work where this approach has been considered.
+This approach will accelerate the pace of identifying effective communication strategies linked to healthy eating.
+This study can also be used to tailor interventions and develop fully automated eCoaching.
 
@@
 Is the contribution non-trivial? Explain why \\
-
-The authors are not aware of any other work where this approach has been considered.
 
@@
 \bibliographystyle{vancouver}
-\bibliography{References}
+\bibliography{references}
 

### 2018-03-01 results/roc/roc-generator.py

@@ -1,0 +1,236 @@
+# include required packages
+import numpy
+import re
+
+from keras.utils import np_utils
+from keras.preprocessing.sequence import pad_sequences
+
+
+# load e-coaching data from datafile
+def get_ecoaching_data(file_name='e-coaching_data_stem.txt'):
+    data = []
+    with open(file_name, "r") as file_stream:
+        for line in file_stream:
+            current_line = line.split(",")
+            if len(current_line) > 1:
+                data.append(clean_text(current_line[0]).strip())
+    return data
+
+
+# split data into 5 folds
+def split_data(data, folds=5):
+    fold_size = len(data)/folds
+    folds_data = []
+    for i in range(folds):
+        start, end = i*fold_size, i*fold_size + fold_size
+        folds_data.append(data[start: end])
+    return folds_data
+
+
+# encode text data into numeric values
+def get_encoded_data(data, vocabulary_obj, max_len):
+    x, y = [], []
+
+    for text in data:
+        input_text = vocabulary_obj.vectorize(text)
+        sample_len = len(input_text)
+
+        if sample_len <= max_len:
+            x.append(input_text)
+            y.append(1)
+        else:
+            for idx in range(sample_len-max_len, 0, -max_len):
+                if idx == sample_len-max_len:
+                    x.append(input_text[idx:sample_len])
+                    y.append(1)
+                else:
+                    x.append(input_text[idx:idx+max_len])
+                    y.append(0)
+
+                if idx - max_len < 0:
+                    x.append(input_text[0:idx])
+                    y.append(0)
+
+    data_x = numpy.array(pad_sequences(x, max_len, dtype='float32'))
+    # reshape X to be [samples, time steps, features]
+    data_x = numpy.reshape(data_x, (data_x.shape[0], max_len, 1))
+    # normalize
+    data_x = data_x / vocabulary_obj.size()
+    data_y = np_utils.to_categorical(y)
+    return data_x, data_y
+
+
+# encode text data into numeric values
+def get_encoded_data_all(data, vocabulary_obj, max_len, window_size=1):
+    x, y = [], []
+    previous_x, current_x = [], []
+
+    for text in data:
+        input_text = vocabulary_obj.vectorize(text)
+        sample_len = len(input_text)
+
+        if sample_len >= window_size:
+            current_x = input_text[0:window_size]
[…]

### 2018-03-01 paper-amia-2018/manuscript.tex
update Introduction and result's tables

@@
 \noindent{\bf Abstract}
-\textit{Machine learning-based models has the potential to to efficiently and accurately identify patient-provider communication behaviors during eCoaching intervention sessions targeting fruit and vegetable intake among GenY young adults.
+\textit{Machine learning-based models has the potential to to efficiently and accurately identify patient-provider communication behaviors during eCoaching intervention sessions targeting fruit and vegetable intake among young adults age 21-30.
 }
@@
 Developing an automatic classification of clinical interactions required segmented text.
-eCoaching data comprised of email responses which need to be segmented into ``block of text'' refers to a particular MI behavior.
+Nevertheless, eCoaching data comprised of email responses which need to be segmented into group of MI behavior referes to ``block of text''.
 Automatic segementation of eCoaching intervention sessions is a challenging task due to the 2 important reasons.
@@
 
-In this paper, we addressing these issues by developing several state-of-the-art machine learning based models for the segmentation of eCoaching text to promote the automatic identification of best communication strategies without human interference.
-More specifically, we develop Support Vector Machine (SVM), Naive Bayes (NB), K-Nearest Neighbor (KNN), Long Short Term Memory (LSTM), and Gated Recurrent Unit (GRU) by utlizing lexical, topic and punctuation mark features for the segmentation of eCoaching text.
+In this paper, we address this problem by developing several state-of-the-art machine learning based models for the segmentation of eCoaching text to promote the automatic identification of best communication strategies without human interference.
+More specifically, we develop Support Vector Machine (SVM), Naive Bayes (NB), K-Nearest Neighbor (KNN), Long Short Term Memory (LSTM), and Gated Recurrent Unit (GRU) by utlizing contextual, topic and punctuation mark features, to find the best model for the segmentation of eCoaching text.
 
-Previous studies mainly focus on segemntation of text into sections, headers or sentences.
-
-The authors are not aware of any other work where this approach has been considered.
-This approach will accelerate the pace of identifying effective communication strategies linked to healthy eating.
-This study can also be used to tailor interventions and develop fully automated eCoaching.
-
-What is the problem? \\
-
-Why is the problem important? \\
-
-What has so far been done on the  problem?\\
-1) describes previous work in more technical detail \\
-2) as far as needed for a proper understanding of the contribution of the paper \\
-
-What is the contribution of the paper on the problem?\\
-1) What are the rival approaches? \\
-2) What are the drawbacks of each?\\
-3) How has the battle between different approaches progressed?\\
-4) What are the major outstanding problems? (This is where you come in)\\
-
-Is the contribution original? Explain why \\
-
-Is the contribution non-trivial? Explain why \\
+Previous studies mainly focus on segemntation of text into sections and headers\cite{apostolova2009automatic,denny2009evaluation,tepper2012statistical,cho2002text} or sentence boundary detection\cite{griffis2016quantitative,kreuzthaler2015detection,treviso2016sentence} in the medical domain.
+Apostolova et al.\cite{apostolova2009automatic} applied SVM by utilizing word-vector cosine similarity metric combined with several heuristics to classify clinical report into semantic sections such as demographics, history, exam procedure, finding, impression, etc.
+After identification of each line in the document, Tepper et al.
+\cite{tepper2012statistical} trained an Maximum Entrophy models for the section classification.
+In 2009, Denny et al.\cite{denny2009evaluation} proposed a SecTag algorithm, which combined natural language processing technique, terminology-based rule, and naive bayesian score for identifying sections and headers that achieved 99\% recall with 95.6\% precision.
+On the other hand, SVM exploiting with linear kernel and recurrent convolutional neural networks with posodic, part of speech features and word embeddings, were trained by Kreuzthaler et al.\cite{kreuzthaler2015detection} and Griffis et al.\cite{griffis2016quantitative}, respectively, for the detection of sentence boundary.
+However, segmentation of clinical text, in partiicular, segmentation of MI or eCoaching text into group of MI behavior is ignored while relying on manual hand-coded approach.
+Therefore, this study introduce an innovative approach and the authors are not aware of any other work this approach has been considered for the segmentation of MI or eCoaching text into ``block of text''.
 
@@
 \centering
-\caption{Performance of MC, HMM, and RNN for predicting the success of under and over-sampled patient-provider communication sequences.
-The highest value for each performance metric is highlighted in bold.}
+\caption{Performance of NB, SVM, KNN, and RNN for identifying segmentation point in eCoaching text. The highest value for each performance metric is highlighted in bold.}
 \label{tab:result_under_over_sampled}
@@
   \hline
-   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{lexical features only}} & \multicolumn{3}{|c|}{\textbf{lexical features + topics}} \\\cline{2-7}
+   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{contextual features only}} & \multicolumn{3}{|c|}{\textbf{contextual + topics + punctuation marks}} \\\cline{2-7}
    & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Score} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Score}\\ \hline
 
- Naive Bayes Multinomial & 0.7060 & 0.7044 & 0.7038 & 0.7932 & 0.7799 & 0.7775 \\ \hline
- Support Vector Machine & 0.6395 & 0.6385 & 0.6379 & 0.7111 & 0.7029 & 0.7000\\ \hline
- K-Nearest Neighbour & 0.6244 & 0.6143 & 0.6067 & 0.7775 & 0.7567 & 0.7520\\ \hline
- Long Short Term Memory & \textbf{0.8733} & \textbf{0.8681} & \textbf{0.8677} & \textbf{0.8424} & \textbf{0.8385} & \textbf{0.8381}\\ \hline
+ Naive Bayes Multinomial & 0.594 & 0.662 & 0.626 & 0.590 & 0.666 & 0.626 \\ \hline
+ Support Vector Machine & 0.742 & \textbf{0.679} & 0.709 & 0.774 & 0.696 & 0.733\\ \hline
+ K-Nearest Neighbour & \textbf{0.808} & 0.663 & \textbf{0.728} & \textbf{0.820} & \textbf{0.742} & \textbf{0.779}\\ \hline
+ Long Short Term Memory & 0.8424 & 0.8385 & 0.8381 & 0.8424 & 0.8385 & 0.8381\\ \hline
  Gated Recurrent Unit & 0.8705 & 0.8676 & 0.8673 & 0.8412 & 0.8377 & 0.8373\\ \hline 
@@
 \centering
-\caption{Performance of MC, HMM, and RNN for predicting the success of under and over-sampled patient-provider communication sequences.
-The highest value for each performance metric is highlighted in bold.}
+\caption{Performance of NB, SVM, KNN, and RNN for identifying continuous eCoaching text. The highest value for each performance metric is highlighted in bold.}
 \label{tab:result_under_over_sampled}
@@
   \hline
-   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{lexical features only}} & \multicolumn{3}{|c|}{\textbf{lexical features + topics}} \\\cline{2-7}
+   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{contextual features only}} & \multicolumn{3}{|c|}{\textbf{contextual + topics + punctuation marks}} \\\cline{2-7}
    & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Score} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Score}\\ \hline
 
- Naive Bayes Multinomial & 0.7060 & 0.7044 & 0.7038 & 0.7932 & 0.7799 & 0.7775 \\ \hline
- Support Vector Machine & 0.6395 & 0.6385 & 0.6379 & 0.7111 & 0.7029 & 0.7000\\ \hline
- K-Nearest Neighbour & 0.6244 & 0.6143 & 0.6067 & 0.7775 & 0.7567 & 0.7520\\ \hline
- Long Short Term Memory & \textbf{0.8733} & \textbf{0.8681} & \textbf{0.8677} & \textbf{0.8424} & \textbf{0.8385} & \textbf{0.8381}\\ \hline
+ Naive Bayes Multinomial & 0.988 & 0.985 & 0.987 & 0.989 & 0.984 & 0.986 \\ \hline
+ Support Vector Machine & \textbf{0.989} & 0.992 & 0.991 & 0.990 & 0.993 & 0.991\\ \hline
+ K-Nearest Neighbour & \textbf{0.989} & \textbf{0.995} & \textbf{0.992} & \textbf{0.991} & \textbf{0.994} & \textbf{0.993}\\ \hline
+ Long Short Term Memory & 0.8424 & 0.8385 & 0.8381 & 0.8424 & 0.8385 & 0.8381\\ \hline
  Gated Recurrent Unit & 0.8705 & 0.8676 & 0.8673 & 0.8412 & 0.8377 & 0.8373\\ \hline 
@@
     \centering
-    \includegraphics[width=0.90\textwidth]{figures/roc-curves.png}
+    \includegraphics[width=1.0\textwidth]{figures/roc-curves.png}
     \caption{\textbf{ROC curves compared the performance of different models}.}
@@
 Our
-proposed methods can help to identify individual text segments, which can be annotated directly with a classification model and increase
-the effectiveness of behavioral interventions.
+proposed methods can help to identify individual text segments, which can be annotated directly with a classification model and increase the effectiveness of behavioral interventions.
+This approach will help for developing fully automated eCoaching and also accelerate the pace of identifying effective communication strategies linked to healthy eating.
 

### 2018-03-02 paper-amia-2018/manuscript.tex
fixed bug in rnn model

@@
  K-Nearest Neighbour & \textbf{0.808} & 0.663 & \textbf{0.728} & \textbf{0.820} & \textbf{0.742} & \textbf{0.779}\\ \hline
- Long Short Term Memory & 0.8424 & 0.8385 & 0.8381 & 0.8424 & 0.8385 & 0.8381\\ \hline
- Gated Recurrent Unit & 0.8705 & 0.8676 & 0.8673 & 0.8412 & 0.8377 & 0.8373\\ \hline 
+ Long Short Term Memory & 0.8424 & 0.8385 & 0.8381  & -- & -- & --\\ \hline
+ Gated Recurrent Unit & 0.8705 & 0.8676 & 0.8673 & -- & -- & --\\ \hline 
   \end{tabular}
 \end{table}
-
-Predictive performance summary of the proposed methods on under and over-sampled sequences is presented in Table.\\
 
@@
  K-Nearest Neighbour & \textbf{0.989} & \textbf{0.995} & \textbf{0.992} & \textbf{0.991} & \textbf{0.994} & \textbf{0.993}\\ \hline
- Long Short Term Memory & 0.8424 & 0.8385 & 0.8381 & 0.8424 & 0.8385 & 0.8381\\ \hline
- Gated Recurrent Unit & 0.8705 & 0.8676 & 0.8673 & 0.8412 & 0.8377 & 0.8373\\ \hline 
+ Long Short Term Memory & 0.8424 & 0.8385 & 0.8381 & -- & -- & --\\ \hline
+ Gated Recurrent Unit & 0.8705 & 0.8676 & 0.8673 & -- & -- & --\\ \hline 
+  \end{tabular}
+\end{table}
+
+Predictive performance summary of the proposed methods on under and over-sampled sequences is presented in Table.\\
+
+\begin{table}[ht]
+\centering
+\caption{Performance of NB, SVM, KNN, and RNN for the identification of segmentation and continuous eCoaching text. The highest value for each performance metric is highlighted in bold.}
+\label{tab:result_under_over_sampled}
+  \begin{tabular}{|l|l|l|l|l|l|l|}
+  \hline
+   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{contextual features only}} & \multicolumn{3}{|c|}{\textbf{contextual + topics + punctuation marks}} \\\cline{2-7}
+   & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Score} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Score}\\ \hline
+
+ Naive Bayes Multinomial & 0.975 & 0.974 & 0.975 & 0.976 & 0.974 & 0.975 \\ \hline
+ Support Vector Machine & 0.981 & 0.982 & 0.981 & 0.983 & 0.983 & 0.983\\ \hline
+ K-Nearest Neighbour & \textbf{0.983} & \textbf{0.984} & \textbf{0.983} & \textbf{0.986} & \textbf{0.986} & \textbf{0.986}\\ \hline
+ Long Short Term Memory & 0.8424 & 0.8385 & 0.8381 & -- & -- & --\\ \hline
+ Gated Recurrent Unit & 0.8705 & 0.8676 & 0.8673 & -- & -- & --\\ \hline 
   \end{tabular}
@@
 We took the inspiration for the representation of behavior codes from the idea of word embeddings. Word embedding is a representation of words in
-low-dimensional space by vectors, which contain the features of the words. In our study, we employed embedding in place of one-hot vectors for representation of behavior codes as input to LSTM and
-GRU, since one-hot vectors are high-dimensional and sparse.
+low-dimensional space by vectors, which contain the features of the words.
+In our study, we employed embedding in place of one-hot vectors for representation of behavior codes as input to LSTM and GRU, since one-hot vectors are high-dimensional and sparse.
 
@@
 \section*{Conclusion}
+Segmentation of eCoaching text is an integral part of developing an automated eCoaching intervention.
+Although several studies have done for the segmentation of clinical text into sections and sentences, none of them are used for the segmentation of text into a group of MI behavior in the setting of discourse analysis with email under the principle of motivational interviews.
 In this paper, we compared the performance of machine learning models for the task of segmentation of e-coaching text.
@@
 Manual segmentation of e-coaching data is very resource-intensive and time consuming task, which can significantly decrease the time and effort required to develop effective behavioral interventions.
-Our
-proposed methods can help to identify individual text segments, which can be annotated directly with a classification model and increase the effectiveness of behavioral interventions.
+Our proposed methods can help to identify individual text segments, which can be annotated directly with a classification model and increase the effectiveness of behavioral interventions.
 This approach will help for developing fully automated eCoaching and also accelerate the pace of identifying effective communication strategies linked to healthy eating.
-
-1) The hypothesis and the evidence for and against it are briefly restated \\
-2) The original motivation is recapitulated \\
-3) The state of the field in the light of this new contribution is reassessed \\
-4) describes future research and new directions suggested by the contribution\\
-5) in particular, research that would improve the evidence for or against the hypothesis
+As our future work, we plan to evaluate our approach on other datasets involves in discourse analysis for enhancing our proposed method.
 

### 2018-03-02 results/roc/roc-generator.py

@@ -151,7 +151,7 @@ def get_encoded_data_rnn(data, vocabulary_obj, max_len):
 # check for special character
 def is_special_char(key, vocabulary_obj):
     int_to_char = dict((i + 1, c) for i, c in enumerate(vocabulary_obj.vocab))
-    special_chars = ['exclam', 'fullstop', 'questionmark', 'comma', 'hyphen', 'clone', 'semiclone']
+    special_chars = ['exclam', 'fullstop', 'quotemark', 'questionmark', 'comma', 'hyphen', 'clone', 'semiclone']
     if int_to_char[key] in special_chars:
         return True
     else:

### 2018-03-02 rnn/seg-gru-model.py

@@ -24,7 +24,6 @@ vocabulary_obj.create()
 char_to_int = dict((c, i+1) for i, c in enumerate(vocabulary_obj.vocab))
 int_to_char = dict((i+1, c) for i, c in enumerate(vocabulary_obj.vocab))
 
-
 # system configuration
 max_len = 5
 folds = 5

### 2018-03-02 rnn/utility.py

@@ -1,6 +1,7 @@
 # include required packages
 import numpy
 import re
+import sys
 
 from keras.utils import np_utils
 from keras.preprocessing.sequence import pad_sequences
@@ -118,6 +119,7 @@ def get_encoded_data_rnn(data, vocabulary_obj, max_len):
     previous_x, current_x = [], []
 
     for text in data:
+        # print text
         input_text = vocabulary_obj.vectorize(text)
         sample_len = len(input_text)
 
@@ -126,16 +128,38 @@ def get_encoded_data_rnn(data, vocabulary_obj, max_len):
             x.append(previous_x + current_x)
             y.append(1)
 
-        current_idx = 0
+            # for idx in previous_x + current_x:
+            #     sys.stdout.write(str(vocabulary_obj.int_to_word[idx] + ', '))
+            # sys.stdout.write("1\n")
+
+            previous_x = input_text[0:1]
+            current_x = []
+
+        current_idx = 1
         while current_idx < sample_len:
 
             if is_special_char(input_text[current_idx], vocabulary_obj):
                 current_x.append(input_text[current_idx])
+                if current_idx == sample_len - 1:
+                    previous_x.append(input_text[current_idx])
             else:
                 current_x.append(input_text[current_idx])
                 x.append(previous_x + current_x)
                 y.append(0)
-                previous_x = current_x
+
+                # for idx in previous_x + current_x:
+                #     sys.stdout.write(str(vocabulary_obj.int_to_word[idx] + ', '))
+                #     pass
+                # sys.stdout.write("0\n")
+
+                if current_idx == sample_len-1 and not is_special_char(input_text[current_idx], vocabulary_obj):
+                    previous_x = [input_text[current_idx]]
+                else:
+                    if len(current_x) > 1:
+                        previous_x = [input_text[current_idx]]
+                    else:
+                        previous_x = current_x
+                current_x = []
 
             current_idx += 1
 
@@ -150,9 +174,8 @@ def get_encoded_data_rnn(data, vocabulary_obj, max_len):
 
 # check for special character
 def is_special_char(key, vocabulary_obj):
-    int_to_char = dict((i + 1, c) for i, c in enumerate(vocabulary_obj.vocab))
-    special_chars = ['exclam', 'fullstop', 'questionmark', 'comma', 'hyphen', 'clone', 'semiclone']
-    if int_to_char[key] in special_chars:
[…]

### 2018-03-02 rnn/vocabulary.py

@@ -6,6 +6,7 @@ import operator
 class Vocabulary:
     def __init__(self, vocab_path='e-coaching.vocab'):
         self.vocab = dict()
+        self.int_to_word = dict()
         self.vocab_path = vocab_path
 
     # build e-coaching dictionary
@@ -13,7 +14,7 @@ class Vocabulary:
         with codecs.open(self.vocab_path, 'r', 'UTF-8') as train_file:
             words = [x.strip().rstrip('\n') for x in train_file.readlines()]
             self.vocab = dict((c, i + 1) for i, c in enumerate(words))
-
+            self.int_to_word = dict((i+1, c) for i, c in enumerate(words))
         return self
 
     # read all the data from file and create vocabulary, also find # classes

### 2018-03-03 paper-amia-2018/manuscript.tex
update methods

@@
 One sentence can be segmented into several MI behaviors, and vice versa.
-Figure 1 illustrates that the marked sentence taken from an eCoaching email exchanage, segmented into 2 different MI behaviors, CHT and HUP.
+Figure~\ref{fig:text-segment} illustrates that the marked sentence taken from an eCoaching email exchanage, segmented into 2 different MI behaviors, CHT and HUP.
+
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=1.0\textwidth]{figures/roc-curves.png}
+    \caption{\textbf{Challenge for text segmentation}.}
+    \label{fig:text-segment}
+\end{figure}
 
@@
 \subsection*{\textit{Data collection}}
-The experimental dataset for this work was constructed from the transcripts of 129 motivational interviews, which include a total of 50,239 segmented and annotated utterances. Each transcript consists
-of an MI interview session involving counselor, adolescent, and caregiver. The utterances were annotated based on MYSCOPE codebook \cite{carcone2013provider}, in which the codes are grouped into
-patient (adolescent and caregiver) codes and counselor codes. Utterances were divided into successful and unsuccessful communication sequences. Successful communication sequences result in
-positive change talk and commitment language statements by an adolescent or caregiver, while unsuccessful sequences are the ones that result in negative change talk or commitment language and the
-sequences, in which no change talk or commitment language statements occur. Out of 5143 observed sequences, 4225 were positive and 918 were negative. Successful sequences had an average length of
-9.79 utterances, while unsuccessful sequences had on average 9.65 utterances. For each of the probabilistic models (MC and HMM), two models were trained, one model was trained using successful
-sequences and another one was trained using unsuccessful sequences.
+The experimental dataset for this work was constructed from the 49 eCoaching sessions, which include a total of 3,138 segmented and annotated MI behaviors.
+Each session consists of an MI intervention involving patient-provider communications in email.
+To filter out noise from the dataset, we removed all non-ascii characters and then applied stemming to obtain a general form of word from different word representations, such as ``eating'', ``eats'', and ``eat''.
+We formulate the text segmentation task into a binary classification, as shown in Figure~\ref{fig:classifier}.
+An email text is given as the input, it is partitioned into adjacent word pairs by sliding them.
+Each pair of them classified into either of the two categories: ``boundary'' and ``not boundary''.
+The text is segmented at position,  where an adjacent word pair classified into ``boundary'' class.
+If all pairs of word classified into ``'not boundary'', the text is treated as one about a single MI behavior.
+Totaly, we obtained 95,421 word pairs, which include 3,138 ``boundary'' and 92,283 ``not boundary'' instances.
 
-Part 1
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=1.0\textwidth]{figures/roc-curves.png}
+    \caption{\textbf{Transformation of text segmentation into text classification}.}
+    \label{fig:classifier}
+\end{figure}
 
-1) What algorithms or data structures did you select? Who created them? What is their asymptotic behavior? What other specific characteristics are worth noting for this study?\\
-2) What programming language and platform did you use? What impact did these choices have on your project?\\
+For the experiment, we utilized three type of features including word, tpoic, and punctuation mark.
+Each word represented in a binary format, where 1 indicates the appearance of the word and 0 for absence.
+Topics are considered as features since topic models are very effective\cite{kotov2015interpretable,hashimoto2016topic,lu2016modeling} to represent text documents.
+In this paper, we exploit the Labeled LDA model\cite{kotov2015interpretable} and represent each word in a vector of 2 topics, where the number of topics is experimentaly determined by the model performance.
+Punctuation mark is one of the most important feature as they indicate the boundary of a sentence, clause and phrase.
+Therefore, punctuation mark containing one of the symbols \{`.', `,', `!', `?', `:', `;', `-'\} is employed as feature.
 
-Part 2
-
-1) How specifically did you implement the algorithms?\\
-2) How did you handle instrumentation code? Why?\\
-3) Did you perform any optimizations? Why or why not?\\
-4) How did you choose to test and benchmark your code?\\
-5) What inputs (data) did you select to test your implementations? Why?\\
-
-\subsection*{\textit{Segmentation method}}
-Generally, a sequence can be viewed as a temporally ordered set of events. In this study, an event is a behavior code that also has a symbolic representation.
+\subsection*{\textit{Segmentation classifiers}}
+Several state-of-the-art classifiers, including Naive Bayes (NB)\cite{pedregosa2011scikit}, Support Vector Machine (SVM)\cite{chang2011libsvm}, K-Nearest Neighbour (KNN)\cite{pedregosa2011scikit}, two variant of Recurrent Neural Networks (RNN)\cite{bengio1993problem}: Long Short Term Memory (LSTM)\cite{hochreiter1997long} and Gated Recurrent Unit (GRU)\cite{cho2014properties}, are employed to estimate the classification performance.
 

### 2018-03-03 paper-amia-2018/manuscript.tex
update method section and fixed bug in RNN model

@@
     \includegraphics[width=1.0\textwidth]{figures/roc-curves.png}
-    \caption{\textbf{Challenge for text segmentation}.}
+    \caption{\textbf{Main challenge of text segmentation}.}
     \label{fig:text-segment}
@@
 The experimental dataset for this work was constructed from the 49 eCoaching sessions, which include a total of 3,138 segmented and annotated MI behaviors.
-Each session consists of an MI intervention involving patient-provider communications in email.
-To filter out noise from the dataset, we removed all non-ascii characters and then applied stemming to obtain a general form of word from different word representations, such as ``eating'', ``eats'', and ``eat''.
+Each session contains an MI intervention involving patient-provider communications in email.
+To filter out noise from the dataset, non-ascii characters are removed and then applied stemming to obtain a general form of word from different word representations, such as ``eating'', ``eats'', and ``eat''.
 We formulate the text segmentation task into a binary classification, as shown in Figure~\ref{fig:classifier}.
@@
 The text is segmented at position, where an adjacent word pair classified into ``boundary'' class.
-If all pairs of word classified into ``'not boundary'', the text is treated as one about a single MI behavior.
+If all pairs of word classified into ``not boundary'', the text is treated as one about a single MI behavior.
 Totaly, we obtained 95,421 word pairs, which include 3,138 ``boundary'' and 92,283 ``not boundary'' instances.
@@
 In this paper, we exploit the Labeled LDA model\cite{kotov2015interpretable} and represent each word in a vector of 2 topics, where the number of topics is experimentaly determined by the model performance.
-Punctuation mark is one of the most important feature as they indicate the boundary of a sentence, clause and phrase.
-Therefore, punctuation mark containing one of the symbols \{`.', `,', `!', `?', `:', `;', `-'\} is employed as feature.
+Punctuation mark containing one of the symbols \{`.', `,', `!', `?', `:', `;', `-'\} is also employed as feature.
+This is one of the most important feature as they indicate the boundary of a sentence, clause and phrase.
 
@@
 
-\textbf{Naive Bayes}: comming soon...
+\textbf{Naive Bayes}: this model is constructed by using the training data and estimate the prior probability of classes, and each feature given the class.
+Then, posterior probability is computed to predict the class label by applying the bayes theorem with the assumption that features are conditionally independent.
+This study utilized a specialized version of Naive Bayes called Multinomial Naive Bayes, which is best suitable for discrete features such as word.
 
-\textbf{Support Vector Machine}: comming soon...
+\textbf{Support Vector Machine}: we used this model as one of the state-of-the-art classification technique proven to perform well in text categorisation\cite{joachims1998text} for its ability to cope with very high dmensional input feature space.
+SVM finds the best hyperplane in the feature space that maximize the separation between the closest ``boundary'' and ``not boundary'' training examples.
+In this experiment, polynomial kernel is employed to train the SVM model for the segmentation of eCoaching text.
 
-\textbf{K-Nearest Neighbour}: comming soon...
+\textbf{K-Nearest Neighbour}: By this model, each training sample represented as a point in the input feature space.
+For a new test sample, Euclidean distance is calculated to find the k-nearest neighbors.
+Finally, the test sample is classified into majority class of the k-nearest neighbors.
+We experimentaly determined that best performance was achieved with k = 3 for the classification of word pairs.
 
-\textbf{Recurrent Neural Networks}: comming soon...
+\textbf{Recurrent Neural Networks}: RNN is a neural network architecture designed to capture sequential patterns present in temporal sequence such as text data.
+When we predict the ``boundary'' point, adjacent word pair will help to understand the pattern of the sequence.
+Long Short Term Memory networks usually reffered as LSTMs\cite{hochreiter1997long}, are a special type of RNN capable of handling variable size input sequence, contains internal memory.
+GRU\cite{cho2014properties} is a variant of LSTM mathematically represented by the following formula:
+
+\begin{equation}
+z_t = \sigma(W_zx_t + U_zh_{t-1} + b_z)
+\label{eq:firstgru}
+\end{equation}
+\begin{equation}
+r_t = \sigma(W_rx_t + U_rh_{t-1} + b_r)
+\label{eq:resetgru}
+\end{equation}
+\begin{equation}
+\tilde h_t = tanh(W_hx_t + r_t \odot U_hh_{t-1} + b_h) 
+\label{eq:candidategru}
+\end{equation}
+\begin{equation}
+h_t = z_t \odot h_{t-1} + (1-z_t) \odot \tilde h_t
+\label{eq:lastgru}
+\end{equation}  
+In Eq.~\ref{eq:firstgru}-\ref{eq:lastgru}, $\sigma$ corresponds to sigmoid function and $\odot$ designates an element-wise product.
+The update gate $z_t$ and reset gate $r_t$ at time step $t$ are computed by the Eq.~(\ref{eq:firstgru}) and~(\ref{eq:resetgru}), where $W_z$, $W_r$, $W_h$, $U_z$, $U_r$, $U_h$ are the weight matrices and $b_z$, $b_h$ and $b_r$ are bias vectors.
+The activation $h_t$ of the GRU at time $t$ is a linear combination of previous activation $h_{t-1}$ and the candidate activation $\tilde h_t$, which is represented by Eq.~(\ref{eq:lastgru}) and~(\ref{eq:candidategru}).
+We experimentally determined that the best performance is achieved when the number of hidden units=32, batch size=32, optimizer=adam, as well as early stopping is used based on the validation loss.
 
 \subsection*{\textit{Evaluation metrics}}
-Performance of the proposed method was evaluated in terms of precision, recall, and F-measure using 10 folds cross-validation and weighted macro-averaging of these metrics over the folds.
-However, LSTM and GRU are trained on 80\% of the data and validated on 10\%.
-The remaining 10\% of the data is used as a test set for reporting the performance of the model.
+In this experiment, standard metrics: precision, recall, and F-measure, are applied to evaluate the performance of binary classifiers\cite{aas1999text}.
+However, we didn't report accuracy as a performance metric because accuracy is highly sensitive to the prior class probabilities and does not fully describe the actual difficulty of the decision problem for unbalanced dataset.
+We conduct the experiment with 5 folds cross-validation and weighted macro-averaging of these metrics over the folds.
+All models are trained on 80\% of the word pairs and remaining 20\% of the data is used as a test set for reporting the performance of the model.
+We also estimated the area under the receiving operating characterstics curve (AUC) metric due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets\cite{hu2015kernelized}.
 
@@
 2) These results should present the raw data
-or the results after applying the techniques outlined in the methods section. The results are simply results; they do not draw conclusions.
+or the results after applying the techniques outlined in the methods section. The results are simply results; they do not draw conclusions.\\
 
@@
 This study was supported by a grant from the National Institutes of Health, NIDDK R21DK108071, Carcone and Kotov, MPIs.
-We would like to thank the student assistants in the Department of Family Medicine and Public Health Sciences at Wayne State University School of Medicine for their help in developing the training dataset by manually annotating the dataset using the MYSCOPE codebook.
+We would like to thank the student assistants in the Department of Family Medicine and Public Health Sciences at Wayne State University School of Medicine for their help in developing the training dataset by manually segmented the eCoaching text.
 

### 2018-03-03 rnn/seg-gru-model.py

@@ -25,12 +25,12 @@ char_to_int = dict((c, i+1) for i, c in enumerate(vocabulary_obj.vocab))
 int_to_char = dict((i+1, c) for i, c in enumerate(vocabulary_obj.vocab))
 
 # system configuration
-max_len = 5
+max_len = 3
 folds = 5
 embed_dim = 300
-lstm_out = 64
+lstm_out = 32
 dropout_x = 0.5
-batch_size = 40
+batch_size = 32
 
 # prepare the data set of input to output pairs encoded as integers
 data = utility.get_ecoaching_data()
@@ -56,11 +56,11 @@ for fold_num in range(folds):
     model = Sequential()
     # model.add(Embedding(vocabulary_obj.size()+1, embed_dim, input_length=X.shape[1]))
     # model.add(GRU(lstm_out, recurrent_regularizer=l1_l2(l1=0.0, l2=0.033), dropout=dropout_x))
-    model.add(LSTM(lstm_out, input_shape=(X.shape[1], 1)))
-    # model.add(GRU(lstm_out, input_shape=(X.shape[1], 1)))
+    # model.add(LSTM(lstm_out, input_shape=(X.shape[1], 1)))
+    model.add(GRU(lstm_out, input_shape=(X.shape[1], 1)))
     model.add(Dense(y.shape[1], activation='softmax'))
     model.compile(loss='categorical_crossentropy', optimizer='adam', metrics=['accuracy'])
-    model.fit(X, y, epochs=400, batch_size=batch_size, verbose=2)
+    model.fit(X, y, epochs=700, batch_size=batch_size, verbose=2)
 
     # summarize performance of the model
     scores = model.evaluate(X, y, verbose=0)

### 2018-03-03 rnn/utility.py

@@ -191,7 +191,7 @@ def clean_text(text):
 
 # evaluate model with F1, Precision and Recall
 def get_macro_average_performance(actual, predicted):
-    f = open("results.txt", "a")
+    f = open("/home/mehedi/Dropbox/results_GRU.txt", "a")
     f.write("fold number \n")
     f.write("actual: ")
     f.writelines(["%s," % item for item in actual])

### 2018-03-04 rnn/seg-gru-model.py
update figures

@@ -30,7 +30,7 @@ folds = 5
 embed_dim = 300
 lstm_out = 32
 dropout_x = 0.5
-batch_size = 32
+batch_size = 8
 
 # prepare the data set of input to output pairs encoded as integers
 data = utility.get_ecoaching_data()
@@ -60,7 +60,10 @@ for fold_num in range(folds):
     model.add(GRU(lstm_out, input_shape=(X.shape[1], 1)))
     model.add(Dense(y.shape[1], activation='softmax'))
     model.compile(loss='categorical_crossentropy', optimizer='adam', metrics=['accuracy'])
-    model.fit(X, y, epochs=700, batch_size=batch_size, verbose=2)
+    # early_stop = EarlyStopping(monitor='val_acc', min_delta=0.00005, patience=20, verbose=1, mode='auto')
+    # callbacks_list = [early_stop]
+    # model.fit(X, y, epochs=1500, callbacks=callbacks_list, batch_size=batch_size, verbose=2, validation_split=0.2)
+    model.fit(X, y, epochs=600, batch_size=batch_size, verbose=2)
 
     # summarize performance of the model
     scores = model.evaluate(X, y, verbose=0)

### 2018-03-04 paper-amia-2018/manuscript.tex
update figures
update abstract

@@
 \noindent{\bf Abstract}
-\textit{Machine learning-based models has the potential to to efficiently and accurately identify patient-provider communication behaviors during eCoaching intervention sessions targeting fruit and vegetable intake among young adults age 21-30.
-}
-
-1) Motivation: Why do we care about the problem and the results? \\
-2) Problem statement: What problem is the paper trying to solve and what is the scope of the work? \\
-3) Approach: What was done to solve the problem? \\
-4) Results: What is the answer to the problem? \\
-5) Conclusions: What implications does the answer imply?
+\textit{Poor eating habits, particularly low fruit and vegetable intake, is a growing, serious public health concern among young adults.
+An effective intervention is required to improve eating habits.
+eCoaching is an email-based intervention technique where a critical step is the segmentation of text for the automatic annotation of email exchange.
+In this study, we transformed this task into classification of detecting boundary of segmentation, and developed sevarals state-of-the-art machine learning models including Support Vector Machine, Naive Bayes, K-Nearest Neighbor (KNN), Recurrent Neural Networks by utlizing contextual, topic and punctuation mark features.
+Results indicate that KNN is the best model and achieved 0.986 F1-score in overall, 0.779 and 0.993 F1-scores for detecting boundary, not boundary, respectively.
+This study has a great implication to identify individual text segments, which can be annotated directly with a classification model, and accelerate the pace of identifying effective communication strategies linked to healthy eating.}
 
@@
 One sentence can be segmented into several MI behaviors, and vice versa.
-Figure~\ref{fig:text-segment} illustrates that the marked sentence taken from an eCoaching email exchanage, segmented into 2 different MI behaviors, CHT and HUP.
+Figure~\ref{fig:text-segment} illustrates the segmentation of an eCoaching email exchanage where first sentence segmented into 2 different MI behaviors.
+On the other hand, fourth and fifth segments contain only one and multiple sentences, respectively.
 
@@
     \centering
-    \includegraphics[width=1.0\textwidth]{figures/roc-curves.png}
-    \caption{\textbf{Main challenge of text segmentation}.}
+    \includegraphics[width=0.9\textwidth]{figures/segment-example.png}
+    \caption{\textbf{Segmentation of eCoaching text depicts the main challenges of boundary detection}.}
     \label{fig:text-segment}
@@
 We formulate the text segmentation task into a binary classification, as shown in Figure~\ref{fig:classifier}.
-An email text is given as the input, it is partitioned into adjacent word pairs by sliding them.
+An intervention session with email exchange is given as the input, it is partitioned into adjacent word pairs by sliding them.
 Each pair of them classified into either of the two categories: ``boundary'' and ``not boundary''.
@@
     \centering
-    \includegraphics[width=1.0\textwidth]{figures/roc-curves.png}
-    \caption{\textbf{Transformation of text segmentation into text classification}.}
+    \includegraphics[width=0.80\textwidth]{figures/classifier.png}
+    \caption{\textbf{Transformation of text segmentation task into text classification task}.}
     \label{fig:classifier}
@@
 
-For the experiment, we utilized three type of features including word, tpoic, and punctuation mark.
+For the experiment, we utilized three type of features including word (textual feature), tpoic, and punctuation mark.
 Each word represented in a binary format, where 1 indicates the appearance of the word and 0 for absence.
@@
 The activation $h_t$ of the GRU at time $t$ is a linear combination of previous activation $h_{t-1}$ and the candidate activation $\tilde h_t$, which is represented by Eq.~(\ref{eq:lastgru}) and~(\ref{eq:candidategru}).
-We experimentally determined that the best performance is achieved when the number of hidden units=32, batch size=32, optimizer=adam, as well as early stopping is used based on the validation loss.
+We build our RNN model with one hidden layer, output layer, and input layer which get one hot encoding of word vector as input.
+Since one-hot vector is given in the input layer, results are reported with textual features and punctuation marks only.
+We experimentally determined that the best performance is achieved when the number of hidden units = 32, batch size = 8, optimizer = adam, as well as 600 epochs is used based on the validation loss.
 
@@
 \centering
-\caption{Performance of NB, SVM, KNN, and RNN for identifying segmentation point in eCoaching text. The highest value for each performance metric is highlighted in bold.}
-\label{tab:result_under_over_sampled}
-  \begin{tabular}{|l|l|l|l|l|l|l|}
+\caption{Performance of NB, SVM, KNN, and RNN for detecting boundaries of segmentation in eCoaching text. The highest value for each performance metric is highlighted in bold.}
+\label{tab:result_boundary}
+  \begin{tabular}{|l|l|l|l|p{0.15\linewidth}|p{0.15\linewidth}|l|}
   \hline
-   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{contextual features only}} & \multicolumn{3}{|c|}{\textbf{contextual + topics + punctuation marks}} \\\cline{2-7}
+   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{contextual features only}} & \multicolumn{3}{|c|}{\textbf{contextual + punctuation marks (+ topics except RNN)}} \\\cline{2-7}
    & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Score} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Score}\\ \hline
 
- Naive Bayes Multinomial & 0.594 & 0.662 & 0.626 & 0.590 & 0.666 & 0.626 \\ \hline
- Support Vector Machine & 0.742 & \textbf{0.679} & 0.709 & 0.774 & 0.696 & 0.733\\ \hline
- K-Nearest Neighbour & \textbf{0.808} & 0.663 & \textbf{0.728} & \textbf{0.820} & \textbf{0.742} & \textbf{0.779}\\ \hline
- Long Short Term Memory & 0.8424 & 0.8385 & 0.8381  & -- & -- & --\\ \hline
- Gated Recurrent Unit & 0.8705 & 0.8676 & 0.8673 & -- & -- & --\\ \hline 
+ NB & 0.594 & 0.662 & 0.626 & 0.590 & 0.666 & 0.626 \\ \hline
+ SVM & 0.742 & \textbf{0.679} & 0.709 & 0.774 & 0.696 & 0.733\\ \hline
+ KNN & \textbf{0.808} & 0.663 & \textbf{0.728} & \textbf{0.820} & \textbf{0.742} & \textbf{0.779}\\ \hline
+ RNN-LSTM & -- & -- & -- & 0.637 & 0.414 & 0.501  \\ \hline
+ RNN-GRU & -- & -- & -- & 0.642 & 0.490 & 0.554 \\ \hline 
   \end{tabular}
@@
 \centering
-\caption{Performance of NB, SVM, KNN, and RNN for identifying continuous eCoaching text. The highest value for each performance metric is highlighted in bold.}
-\label{tab:result_under_over_sampled}
-  \begin{tabular}{|l|l|l|l|l|l|l|}
+\caption{Performance of NB, SVM, KNN, and RNN  for detecting no boundaries in eCoaching text. The highest value for each performance metric is highlighted in bold.}
+\label{tab:result_not_boundary}
+  \begin{tabular}{|l|l|l|l|p{0.15\linewidth}|p{0.15\linewidth}|l|}
   \hline
-   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{contextual features only}} & \multicolumn{3}{|c|}{\textbf{contextual + topics + punctuation marks}} \\\cline{2-7}
+   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{contextual features only}} & \multicolumn{3}{|c|}{\textbf{contextual + punctuation marks (+ topics except RNN)}} \\\cline{2-7}
    & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Score} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Score}\\ \hline
 
- Naive Bayes Multinomial & 0.988 & 0.985 & 0.987 & 0.989 & 0.984 & 0.986 \\ \hline
- Support Vector Machine & \textbf{0.989} & 0.992 & 0.991 & 0.990 & 0.993 & 0.991\\ \hline
- K-Nearest Neighbour & \textbf{0.989} & \textbf{0.995} & \textbf{0.992} & \textbf{0.991} & \textbf{0.994} & \textbf{0.993}\\ \hline
- Long Short Term Memory & 0.8424 & 0.8385 & 0.8381 & -- & -- & --\\ \hline
- Gated Recurrent Unit & 0.8705 & 0.8676 & 0.8673 & -- & -- & --\\ \hline 
+ NB & 0.988 & 0.985 & 0.987 & 0.989 & 0.984 & 0.986 \\ \hline
+ SVM & \textbf{0.989} & 0.992 & 0.991 & 0.990 & 0.993 & 0.991\\ \hline
+ KNN & \textbf{0.989} & \textbf{0.995} & \textbf{0.992} & \textbf{0.991} & \textbf{0.994} & \textbf{0.993}\\ \hline
+ RNN-LSTM & -- & -- & -- & 0.980 & 0.992 & 0.986 \\ \hline
+ RNN-GRU & -- & -- & -- & 0.983 & 0.991 & 0.987 \\ \hline 
   \end{tabular}
@@
 \centering
-\caption{Performance of NB, SVM, KNN, and RNN for the identification of segmentation and continuous eCoaching text. The highest value for each performance metric is highlighted in bold.}
-\label{tab:result_under_over_sampled}
-  \begin{tabular}{|l|l|l|l|l|l|l|}
+\caption{Weighted average performance of NB, SVM, KNN, and RNN for the segmentation of eCoaching text in detecting both ``boundary'' and ``not boundary''.
+The highest value for each performance metric is highlighted in bold.}
+\label{tab:result_weighted_avg}
+  \begin{tabular}{|l|l|l|l|p{0.15\linewidth}|p{0.15\linewidth}|l|}
   \hline
-   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{contextual features only}} & \multicolumn{3}{|c|}{\textbf{contextual + topics + punctuation marks}} \\\cline{2-7}
+   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{contextual features only)}} & \multicolumn{3}{|c|}{\textbf{contextual + punctuation marks (+ topics except RNN)}} \\\cline{2-7}
    & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Score} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Score}\\ \hline
 
- Naive Bayes Multinomial & 0.975 & 0.974 & 0.975 & 0.976 & 0.974 & 0.975 \\ \hline
- Support Vector Machine & 0.981 & 0.982 & 0.981 & 0.983 & 0.983 & 0.983\\ \hline
- K-Nearest Neighbour & \textbf{0.983} & \textbf{0.984} & \textbf{0.983} & \textbf{0.986} & \textbf{0.986} & \textbf{0.986}\\ \hline
- Long Short Term Memory & 0.8424 & 0.8385 & 0.8381 & -- & -- & --\\ \hline
- Gated Recurrent Unit & 0.8705 & 0.8676 & 0.8673 & -- & -- & --\\ \hline 
+ NB & 0.975 & 0.974 & 0.975 & 0.976 & 0.974 & 0.975 \\ \hline
+ SVM & 0.981 & 0.982 & 0.981 & 0.983 & 0.983 & 0.983\\ \hline
+ KNN & \textbf{0.983} & \textbf{0.984} & \textbf{0.983} & \textbf{0.986} & \textbf{0.986} & \textbf{0.986}\\ \hline
+ RNN-LSTM & -- & -- & -- & 0.969 & 0.973 & 0.970 \\ \hline
+ RNN-GRU & -- & -- & -- & 0.972 & 0.974 & 0.973 \\ \hline 
   \end{tabular}
@@
 
+\section*{Discussion}
+By analyzing the experimental results of different communication sequence outcome prediction methods proposed in this paper, we arrived at the following conclusions. First, the overall predictive
+performance of RNN models is substantially better than probabilistic models. In particular, the RNN-based method achieves near-human accuracy for predicting the
+
 \begin{figure}[!htb]
@@
     \includegraphics[width=1.0\textwidth]{figures/roc-curves.png}
-    \caption{\textbf{ROC curves compared the performance of different models}.}
+    \caption{\textbf{Receiver operating characteristic curves showing the performance of binary classifiers for the segmentation of eCoaching text}.}
     \label{fig:roc-curves}
 \end{figure}
-
-\section*{Discussion}
-By analyzing the experimental results of different communication sequence outcome prediction methods proposed in this paper, we arrived at the following conclusions. First, the overall predictive
-performance of RNN models is substantially better than probabilistic models. In particular, the RNN-based method achieves near-human accuracy for predicting the
 

### 2018-03-05 paper-amia-2018/manuscript.tex
update discussion

@@
 However, segmentation of clinical text, in partiicular, segmentation of MI or eCoaching text into group of MI behavior is ignored while relying on manual hand-coded approach.
-Therefore, this study introduce an innovative approach and the authors are not aware of any other work this approach has been considered for the segmentation of MI or eCoaching text into ``block of text''.
+Therefore, this study introduce a novel approach and the authors are not aware of any other work this approach has been considered for the segmentation of MI or eCoaching text into ``block of text''.
 
@@
 All models are trained on 80\% of the word pairs and remaining 20\% of the data is used as a test set for reporting the performance of the model.
-We also estimated the area under the receiving operating characterstics curve (AUC) metric due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets\cite{hu2015kernelized}.
+We also estimated the area under the receiving operating characterstics (ROC) curve\cite{kumar2011receiver} (AUC) metric due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets\cite{hu2015kernelized}.
 
 \section*{Results}
-Experimental evaluation of the proposed method is conducted on both under and over-sampled sequences.
-Predictive performance summary of the proposed methods on under and over-sampled sequences is presented in Table.\\
-
-1) In general, the pure, unbiased results should be presented first without interpretation (van Wagenen 1990). \\
-2) These results should present the raw data
-or the results after applying the techniques outlined in the methods section. The results are simply results; they do not draw conclusions.\\
+opening para: evaluated with boundary, not boudary and overall \\
 
@@
  KNN & \textbf{0.808} & 0.663 & \textbf{0.728} & \textbf{0.820} & \textbf{0.742} & \textbf{0.779}\\ \hline
- RNN-LSTM & -- & -- & -- & 0.637 & 0.414 & 0.501  \\ \hline
- RNN-GRU & -- & -- & -- & 0.642 & 0.490 & 0.554 \\ \hline 
+ LSTM & -- & -- & -- & 0.637 & 0.414 & 0.501  \\ \hline
+ GRU & -- & -- & -- & 0.642 & 0.490 & 0.554 \\ \hline 
   \end{tabular}
 \end{table}
+
+para 1:
+0) opening with referencing table
+1) KNN best in all metrics except Recall
+2) RNN shows lowest performance in terms of recall and F1 score, although GRU better than LSTM 
+3) SVM is second highest model and highest model in terms of recall when only textual features is used.
+4) NB shows lowest precision value among all and better than RNN with recall and F1-score
+5) for additional features, recall and F1 increase x, y, z, for a, b, c, respectively
+6) However, precision increase for all except NB \\
 
@@
  KNN & \textbf{0.989} & \textbf{0.995} & \textbf{0.992} & \textbf{0.991} & \textbf{0.994} & \textbf{0.993}\\ \hline
- RNN-LSTM & -- & -- & -- & 0.980 & 0.992 & 0.986 \\ \hline
- RNN-GRU & -- & -- & -- & 0.983 & 0.991 & 0.987 \\ \hline 
+ LSTM & -- & -- & -- & 0.980 & 0.992 & 0.986 \\ \hline
+ GRU & -- & -- & -- & 0.983 & 0.991 & 0.987 \\ \hline 
   \end{tabular}
@@
 
-Predictive performance summary of the proposed methods on under and over-sampled sequences is presented in Table.\\
+para 2:
+0) opening with referencing table
+2) Similar to boundary detection, KNN is best and RNN is lowest in all metrics for 'not boundary' identification
+3) Performance is very high compared to boundary detection however, perform better than random
+4) Effect of additional feature is consistent in 'not boundary' classification.
+5) for additional features, precision, recall and F1 increase x, y, z, for a, b, c, respectively \\
 
@@
  KNN & \textbf{0.983} & \textbf{0.984} & \textbf{0.983} & \textbf{0.986} & \textbf{0.986} & \textbf{0.986}\\ \hline
- RNN-LSTM & -- & -- & -- & 0.969 & 0.973 & 0.970 \\ \hline
- RNN-GRU & -- & -- & -- & 0.972 & 0.974 & 0.973 \\ \hline 
+ LSTM & -- & -- & -- & 0.969 & 0.973 & 0.970 \\ \hline
+ GRU & -- & -- & -- & 0.972 & 0.974 & 0.973 \\ \hline 
   \end{tabular}
@@
 
-We took the inspiration for the representation of behavior codes from the idea of word embeddings. Word embedding is a representation of words in
-low-dimensional space by vectors, which contain the features of the words.
-In our study, we employed embedding in place of one-hot vectors for representation of behavior codes as input to LSTM and GRU, since one-hot vectors are high-dimensional and sparse.
+para 3:
+0) opening with referencing table
+1) Overall, KNN is best and RNN is lowest in all metrics for 'not boundary' identification
+2) SVM and RNN achieved x and xx, y and yy, z and zz for precision, recall and F1, respectively.
+3) Effect of additional feature is also consistent.
+4) for additional features, precision, recall and F1 increase x, y, z, for a, b, c, respectively\\
 
 \section*{Discussion}
-By analyzing the experimental results of different communication sequence outcome prediction methods proposed in this paper, we arrived at the following conclusions. First, the overall predictive
-performance of RNN models is substantially better than probabilistic models. In particular, the RNN-based method achieves near-human accuracy for predicting the
+This study is the first large-scale efforts to evaluate the segmentation of eCoaching text.
+Experimental results indicate that KNN achieved the best results among all machine learning methods considered for this study.
+KNN achieved 0.986 F1-score in overall, 0.779 and 0.993 F1-scores for detecting ``boundary'' and ``not boundary'', respectively.
+Robust performance of KNN provides the evidence that machine learning models are capable to learn information from the email exchange.
+Although the domain of this study was intentinally quite small, we believe that our study is not limited to ecoacing domain, and it can be successfully applied to other domain as well.
+
+The additional topic and punctuation mark features, made significant improvement in performance of all maching learning methods.
+In every cases, model performs better when contextual feature is used in combination with topic and punctuation mark features.
+Thsi results also mean that segmentation performance might be improved by adding more relevant features including human insight into the problem.
 
@@
     \includegraphics[width=1.0\textwidth]{figures/roc-curves.png}
-    \caption{\textbf{Receiver operating characteristic curves showing the performance of binary classifiers for the segmentation of eCoaching text}.}
+    \caption{\textbf{Receiver operating characteristic curves showing the performance of binary classifiers for the segmentation of eCoaching text when textual features (left) and combination of textual and other features (right) are used}.}
     \label{fig:roc-curves}
@@
 
-1) What, specifically, did you learn from comparing these algorithms or data structures? \\
-2) What do your results say about the problem or question you were investigating?\\
-3) Was your hypothesis confirmed or disproved?\\
-4) Are the results what you expected?\\
-5) If you obtained anomalies or other unexpected results, can you explain them? If not, how could you set about in the future to identify what caused them? \\
-6) How do your results compare to past findings? Are they consistent? Different?Why? \\
-7) How would you respond to objections or questions that other researchers might have about your methods, results, or interpretations? \\
-8) What is new and significant?\\
+In this paper, results are reported by each class to avoid confusion about the overall model performance.
+In addition, standard metrics: precision, recall and F1-score were used to eliminate doubt about the model performance because accuracy is misleading for imbalance dataset.
+AUC values are also outlined due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets\cite{hu2015kernelized}, which was demonstrated by the ROC curves in Figure~\ref{fig:roc-curves}.
+NB shows the highest AUC values and achieved 0.978 for both cases while it provides lowest classification results except RNN.
+On the other hand, KNN and SVM exibits 0.972 and 0.835 AUCs when only textual features are used; and 0.959 and 0.844 AUCs when combination of textual, topic and punctuation mark features are used, respectively.
+Finally, LSTM and GRU demonstrate lowest AUC values among all machine learning models, achieved 0.751 and 0.759 AUCs, respectively.
+
+We observed worst results of RNN, in particular LSTM and GRU for the text segmetation.
+We believe that RNN performed poorly because it has a large set of weights which required large set of data for both class.
+It this study, we utilized 3,138 examples of boundary case which failed to achieve good results.
+While performance of RNN is poor, GRU performed better than LSTM which was already observed in other previous study\cite{chung2014empirical}.
+
+Punctuation mark plays an important role in segmentation boundary detection, and a large numbers of errors were encountered by the false positive of boundary identification.
+Similarly, additional information is the common reason of classified original segment into multiple segments.
+For example, [need help from April].
+
+Our proposed approach is a novel approach for text segmentation where previous study focus on the segmentation of text into sections, headers, and sentences.
+However, this study segmented email exchange into ``block of text'' represent a particular MI behavior.
+This work will significantly reduce the amount of resource and time required to segemnt email text manually.
+Furthermore, this paper can help to annotate each segment automatically by building a new classifer, which will accelerate the pace of finding best communication strategies to develop an effective MI intervention for healthy eating.
+
+As for future work, we plan to evaluate our methods with several similar and different datasets.
+We also plan to use combination of machine laerning and natural language approaches to improve model performance.
+For example, part-of-speech tagging and distance from boundary of the sentence might significantly enhance the classification performance.
+The limitation of this study is that eCoaching text are collected from a single medical institute; formatting, style, and email segment can be different in other settings.
 

### 2018-03-05 results/roc/preprocess-roc-data.py

@@ -26,17 +26,17 @@ def create_roc_csv_for_rnn(path):
         if 'RNN' in file_name and 'csv' not in file_name:
             actual = []
             predicted = []
-            count = 0
+            count = 1
             with open(path + "/" + file_name, "r") as file_stream:
                 for line in file_stream:
-                    count += 1
                     if count == 1:
                         current_line = line.split(",")
                         actual += current_line
+                        count = 0
                     else:
                         current_line = line.split(",")
                         predicted += current_line
-                        count = 0
+                        count = 1
 
             sorted_actual, sorted_predicted = [], []
             for i in numpy.fliplr([numpy.argsort(predicted)])[0]:

### 2018-03-05 paper-amia-2018/manuscript.tex
update results and ROC curves

@@
  KNN & \textbf{0.808} & 0.663 & \textbf{0.728} & \textbf{0.820} & \textbf{0.742} & \textbf{0.779}\\ \hline
- LSTM & -- & -- & -- & 0.637 & 0.414 & 0.501  \\ \hline
+ LSTM & -- & -- & -- & 0.619 & 0.416 & 0.497  \\ \hline
  GRU & -- & -- & -- & 0.642 & 0.490 & 0.554 \\ \hline 
@@
  KNN & \textbf{0.989} & \textbf{0.995} & \textbf{0.992} & \textbf{0.991} & \textbf{0.994} & \textbf{0.993}\\ \hline
- LSTM & -- & -- & -- & 0.980 & 0.992 & 0.986 \\ \hline
+ LSTM & -- & -- & -- & 0.981 & 0.991 & 0.986 \\ \hline
  GRU & -- & -- & -- & 0.983 & 0.991 & 0.987 \\ \hline 
@@
  KNN & \textbf{0.983} & \textbf{0.984} & \textbf{0.983} & \textbf{0.986} & \textbf{0.986} & \textbf{0.986}\\ \hline
- LSTM & -- & -- & -- & 0.969 & 0.973 & 0.970 \\ \hline
+ LSTM & -- & -- & -- & 0.969 & 0.972 & 0.970 \\ \hline
  GRU & -- & -- & -- & 0.972 & 0.974 & 0.973 \\ \hline 
@@
 NB shows the highest AUC values and achieved 0.978 for both cases while it provides lowest classification results except RNN.
-On the other hand, KNN and SVM exibits 0.972 and 0.835 AUCs when only textual features are used; and 0.959 and 0.844 AUCs when combination of textual, topic and punctuation mark features are used, respectively.
-Finally, LSTM and GRU demonstrate lowest AUC values among all machine learning models, achieved 0.751 and 0.759 AUCs, respectively.
+On the other hand, KNN and SVM exibits 0.972 and 0.835 AUCs when only textual features are used; and 0.959 and 0.844 AUCs when combination of textual, topic and punctuation mark features are used.
+Finally, RNN demonstrate lowest AUC values among all machine learning models, achieved AUC values 0.704 and 0.740 for LSTM and GRU, respectively.
 

### 2018-03-05 results/roc/roc-curves-generator.R

@@ -11,28 +11,17 @@ r1 <- read.csv("/home/mehedi/teana/projects/eCoaching-Text-Segmentation/results/
 r2 <- read.csv("/home/mehedi/teana/projects/eCoaching-Text-Segmentation/results/roc/False/SVMFalse-ROC.csv")
 r3 <- read.csv("/home/mehedi/teana/projects/eCoaching-Text-Segmentation/results/roc/False/NBMFalse-ROC.csv")
 
-# print AUC values
-x1 <- read.csv("/home/mehedi/teana/projects/eCoaching-Text-Segmentation/results/roc/False/auc_RNN-LSTMFalse-ROC.csv")
-x2 <- read.csv("/home/mehedi/teana/projects/eCoaching-Text-Segmentation/results/roc/False/auc_RNN-GRUFalse-ROC.csv")
-roc_obj_lstm <- roc(x1$actual, x1$prediction)
-auc(roc_obj_lstm)
-
-roc_obj_gru <- roc(x2$actual, x2$prediction)
-auc(roc_obj_gru)
-
 # plot graph
 plot(r1$fpr, r1$tpr, col='red', type="l", ylab = "Sensitivity", xlab = "1-Specificity", cex.axis=1, cex.lab=1, lwd=1)
 lines(r2$fpr, r2$tpr, col='blue', lty=1, lwd=1)
 lines(r3$fpr, r3$tpr, col='green', lty=1, lwd=1)
-lines(1-roc_obj_lstm$specificities, roc_obj_lstm$sensitivities, col='pink', lty=1, lwd=1)
-lines(1-roc_obj_gru$specificities, roc_obj_gru$sensitivities, col='purple', lty=1, lwd=1)
 lines(0:1, 0:1, col='black', lty=5, lwd=1)
-legend(0.73,0.53, c("KNN", "SVM", "NBM", "LSTM", "GRU"), 
-       lty=c(1,1,1,1,1), col=c('red', 'blue', 'green', 'pink', 'purple'), cex=1, lwd=1)
+legend(0.75,0.36, c("KNN", "SVM", "NB"), 
+       lty=c(1,1,1), col=c('red', 'blue', 'green'), cex=1, lwd=1)
 
 # print AUC values
-x1 <- read.csv("/home/mehedi/teana/projects/eCoaching-Text-Segmentation/results/roc/False/auc_RNN-LSTMFalse-ROC.csv")
[…]

### 2018-03-05 results/roc/roc-generator.py

@@ -1,236 +1,0 @@
-# include required packages
-import numpy
-import re
-
-from keras.utils import np_utils
-from keras.preprocessing.sequence import pad_sequences
-
-
-# load e-coaching data from datafile
-def get_ecoaching_data(file_name='e-coaching_data_stem.txt'):
-    data = []
-    with open(file_name, "r") as file_stream:
-        for line in file_stream:
-            current_line = line.split(",")
-            if len(current_line) > 1:
-                data.append(clean_text(current_line[0]).strip())
-    return data
-
-
-# split data into 5 folds
-def split_data(data, folds=5):
-    fold_size = len(data)/folds
-    folds_data = []
-    for i in range(folds):
-        start, end = i*fold_size, i*fold_size + fold_size
-        folds_data.append(data[start: end])
-    return folds_data
-
-
-# encode text data into numeric values
-def get_encoded_data(data, vocabulary_obj, max_len):
-    x, y = [], []
-
-    for text in data:
-        input_text = vocabulary_obj.vectorize(text)
-        sample_len = len(input_text)
-
-        if sample_len <= max_len:
-            x.append(input_text)
-            y.append(1)
-        else:
-            for idx in range(sample_len-max_len, 0, -max_len):
-                if idx == sample_len-max_len:
-                    x.append(input_text[idx:sample_len])
-                    y.append(1)
-                else:
-                    x.append(input_text[idx:idx+max_len])
-                    y.append(0)
-
-                if idx - max_len < 0:
-                    x.append(input_text[0:idx])
-                    y.append(0)
-
-    data_x = numpy.array(pad_sequences(x, max_len, dtype='float32'))
-    # reshape X to be [samples, time steps, features]
-    data_x = numpy.reshape(data_x, (data_x.shape[0], max_len, 1))
-    # normalize
-    data_x = data_x / vocabulary_obj.size()
-    data_y = np_utils.to_categorical(y)
-    return data_x, data_y
-
-
-# encode text data into numeric values
-def get_encoded_data_all(data, vocabulary_obj, max_len, window_size=1):
-    x, y = [], []
-    previous_x, current_x = [], []
-
-    for text in data:
-        input_text = vocabulary_obj.vectorize(text)
-        sample_len = len(input_text)
-
-        if sample_len >= window_size:
-            current_x = input_text[0:window_size]
[…]

### 2018-03-05 paper-amia-2018/manuscript.tex
update discussion

@@
   \hline
-   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{contextual features only)}} & \multicolumn{3}{|c|}{\textbf{contextual + punctuation marks (+ topics except RNN)}} \\\cline{2-7}
+   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{contextual features only}} & \multicolumn{3}{|c|}{\textbf{contextual + punctuation marks (+ topics except RNN)}} \\\cline{2-7}
    & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Score} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Score}\\ \hline
@@
 This study is the first large-scale efforts to evaluate the segmentation of eCoaching text.
-Experimental results indicate that KNN achieved the best results among all machine learning methods considered for this study.
+Experimental results indicate that KNN is the best model among all machine learning methods considered for this study.
 KNN achieved 0.986 F1-score in overall, 0.779 and 0.993 F1-scores for detecting ``boundary'' and ``not boundary'', respectively.

### 2018-03-05 paper-amia-2018/manuscript.tex
version send to april

@@
 An effective intervention is required to improve eating habits.
-eCoaching is an email-based intervention technique where a critical step is the segmentation of text for the automatic annotation of email exchange.
-In this study, we transformed this task into classification of detecting boundary of segmentation, and developed sevarals state-of-the-art machine learning models including Support Vector Machine, Naive Bayes, K-Nearest Neighbor (KNN), Recurrent Neural Networks by utlizing contextual, topic and punctuation mark features.
-Results indicate that KNN is the best model and achieved 0.986 F1-score in overall, 0.779 and 0.993 F1-scores for detecting boundary, not boundary, respectively.
+eCoaching is an email-based intervention technique where a critical step is the segmentation of text for the automatic annotation of the email exchange.
+In this study, we transformed this task into the classification of detecting the boundary of segmentation and developed several state-of-the-art machine learning models including Support Vector Machine, Naive Bayes, K-Nearest Neighbor (KNN), Recurrent Neural Networks by utilizing contextual, topic and punctuation mark features.
+Results indicate that KNN is the best model and achieved 0.986 F1-score in overall, 0.779 and 0.993 F1-scores for detecting boundary, not a boundary, respectively.
 This study has a great implication to identify individual text segments, which can be annotated directly with a classification model, and accelerate the pace of identifying effective communication strategies linked to healthy eating.}
@@
 Unfortunately, less than one-third of US adults\cite{blanck2008trends,ogden2006prevalence} and only 20\% of GenY\cite{blanck2008trends,american2006american,thompson2005dietary} eat the recommended 5 servings of fruit and vegetables daily.
-Those in inner city urban and rural settings have among the poorest eating habits\cite{blanck2008trends,centers2007fruit,ogden2006prevalence,american2006american,thompson2005dietary}.
+Those in inner-city urban and rural settings have among the poorest eating habits\cite{blanck2008trends,centers2007fruit,ogden2006prevalence,american2006american,thompson2005dietary}.
 GenY's poor dietary practices placing them at high risk for obesity and many chronic diseases, such as type 2 diabetes, as well as declines in predicted health status and life expectancy.
@@
 
-A major drawback of this research is the qualitative methods traditionally used to analyze the communication process which are resource-intensive, requiring an iterative process of human (subjective) interpretation of text.
+A major drawback of this research is the qualitative methods traditionally used to analyze the communication process which is resource-intensive, requiring an iterative process of human (subjective) interpretation of the text.
 Rapidly developing computational technologies, specifically machine learning combined with classification models, offer a unique opportunity to accelerate this process.
@@
 
-However, a significant barrier of fully automate eCoaching is the unsegmented text data.
+However, a significant barrier to fully automate eCoaching is the unsegmented text data.
 Developing an automatic classification of clinical interactions required segmented text.
-Nevertheless, eCoaching data comprised of email responses which need to be segmented into group of MI behavior referes to ``block of text''.
-Automatic segementation of eCoaching intervention sessions is a challenging task due to the 2 important reasons.
+Nevertheless, eCoaching data comprised of email responses which need to be segmented into a group of MI behavior refers to ``block of text''.
+Automatic segmentation of eCoaching intervention sessions is a challenging task due to the 2 important reasons.
 First, the email is an unstructured text that contains informal email exchange in non-traditional formats.
 Second, a text segment not necessarily belongs to the entire sentence or collection of sentences.
-One sentence can be segmented into several MI behaviors, and vice versa.
-Figure~\ref{fig:text-segment} illustrates the segmentation of an eCoaching email exchanage where first sentence segmented into 2 different MI behaviors.
+One sentence can be segmented into several MI behaviors and vice versa.
+Figure~\ref{fig:text-segment} illustrates the segmentation of an eCoaching email exchange where the first sentence segmented into 2 different MI behaviors.
 On the other hand, fourth and fifth segments contain only one and multiple sentences, respectively.
@@
 In this paper, we address this problem by developing several state-of-the-art machine learning based models for the segmentation of eCoaching text to promote the automatic identification of best communication strategies without human interference.
-More specifically, we develop Support Vector Machine (SVM), Naive Bayes (NB), K-Nearest Neighbor (KNN), Long Short Term Memory (LSTM), and Gated Recurrent Unit (GRU) by utlizing contextual, topic and punctuation mark features, to find the best model for the segmentation of eCoaching text.
+More specifically, we develop Support Vector Machine (SVM), Naive Bayes (NB), K-Nearest Neighbor (KNN), Long Short Term Memory (LSTM), and Gated Recurrent Unit (GRU) by utilizing contextual, topic and punctuation mark features, to find the best model for the segmentation of eCoaching text.
 
-Previous studies mainly focus on segemntation of text into sections and headers\cite{apostolova2009automatic,denny2009evaluation,tepper2012statistical,cho2002text} or sentence boundary detection\cite{griffis2016quantitative,kreuzthaler2015detection,treviso2016sentence} in the medical domain.
+Previous studies mainly focus on segmentation of text into sections and headers\cite{apostolova2009automatic,denny2009evaluation,tepper2012statistical,cho2002text} or sentence boundary detection\cite{griffis2016quantitative,kreuzthaler2015detection,treviso2016sentence} in the medical domain.
 Apostolova et al.\cite{apostolova2009automatic} applied SVM by utilizing word-vector cosine similarity metric combined with several heuristics to classify clinical report into semantic sections such as demographics, history, exam procedure, finding, impression, etc.
 After identification of each line in the document, Tepper et al.
-\cite{tepper2012statistical} trained an Maximum Entrophy models for the section classification.
-In 2009, Denny et al.\cite{denny2009evaluation} proposed a SecTag algorithm, which combined natural language processing technique, terminology-based rule, and naive bayesian score for identifying sections and headers that achieved 99\% recall with 95.6\% precision.
+\cite{tepper2012statistical} trained Maximum Entropy models for the section classification.
+In 2009, Denny et al.\cite{denny2009evaluation} proposed a SecTag algorithm, which combined natural language processing technique, terminology-based rule, and naive Bayesian score for identifying sections and headers that achieved 99\% recall with 95.6\% precision.
 On the other hand, SVM exploiting with linear kernel and recurrent convolutional neural networks with posodic, part of speech features and word embeddings, were trained by Kreuzthaler et al.\cite{kreuzthaler2015detection} and Griffis et al.\cite{griffis2016quantitative}, respectively, for the detection of sentence boundary.
-However, segmentation of clinical text, in partiicular, segmentation of MI or eCoaching text into group of MI behavior is ignored while relying on manual hand-coded approach.
-Therefore, this study introduce a novel approach and the authors are not aware of any other work this approach has been considered for the segmentation of MI or eCoaching text into ``block of text''.
+However, segmentation of clinical text, in particular, segmentation of MI or eCoaching text into a group of MI behavior is ignored while relying on manual hand-coded approach.
+Therefore, this study introduces a novel approach and the authors are not aware of any other work this approach has been considered for the segmentation of MI or eCoaching text into ``block of text''.
 
@@
 Each pair of them classified into either of the two categories: ``boundary'' and ``not boundary''.
-The text is segmented at position, where an adjacent word pair classified into ``boundary'' class.
+The text is segmented at the position, where an adjacent word pair classified into ``boundary'' class.
 If all pairs of word classified into ``not boundary'', the text is treated as one about a single MI behavior.
@@
 
-For the experiment, we utilized three type of features including word (textual feature), tpoic, and punctuation mark.
+For the experiment, we utilized three type of features including word (textual feature), topic, and punctuation mark.
 Each word represented in a binary format, where 1 indicates the appearance of the word and 0 for absence.
 Topics are considered as features since topic models are very effective\cite{kotov2015interpretable,hashimoto2016topic,lu2016modeling} to represent text documents.
-In this paper, we exploit the Labeled LDA model\cite{kotov2015interpretable} and represent each word in a vector of 2 topics, where the number of topics is experimentaly determined by the model performance.
+In this paper, we exploit the Labeled LDA model\cite{kotov2015interpretable} and represent each word in a vector of 2 topics, where the number of topics is experimentally determined by the model performance.
 Punctuation mark containing one of the symbols \{`.', `,', `!', `?', `:', `;', `-'\} is also employed as feature.
-This is one of the most important feature as they indicate the boundary of a sentence, clause and phrase.
+This is one of the most important features as they indicate the boundary of a sentence, clause, and phrase.
 
@@
 
-\textbf{Naive Bayes}: this model is constructed by using the training data and estimate the prior probability of classes, and each feature given the class.
-Then, posterior probability is computed to predict the class label by applying the bayes theorem with the assumption that features are conditionally independent.
+\textbf{Naive Bayes}: this model is constructed by using the training data and estimate the prior probability of classes, and each feature has given the class.
+Then, the posterior probability is computed to predict the class label by applying the Bayes theorem with the assumption that features are conditionally independent.
 This study utilized a specialized version of Naive Bayes called Multinomial Naive Bayes, which is best suitable for discrete features such as word.
 
-\textbf{Support Vector Machine}: we used this model as one of the state-of-the-art classification technique proven to perform well in text categorisation\cite{joachims1998text} for its ability to cope with very high dmensional input feature space.
-SVM finds the best hyperplane in the feature space that maximize the separation between the closest ``boundary'' and ``not boundary'' training examples.
-In this experiment, polynomial kernel is employed to train the SVM model for the segmentation of eCoaching text.
+\textbf{Support Vector Machine}: we used this model as one of the state-of-the-art classification technique proven to perform well in text categorisation\cite{joachims1998text} for its ability to cope with very high dimensional input feature space.
+SVM finds the best hyperplane in the feature space that maximizes the separation between the closest ``boundary'' and ``not boundary'' training examples.
+In this experiment, the polynomial kernel is employed to train the SVM model for the segmentation of eCoaching text.
 
@@
 Finally, the test sample is classified into majority class of the k-nearest neighbors.
-We experimentaly determined that best performance was achieved with k = 3 for the classification of word pairs.
+We experimentally determined that best performance was achieved with k = 3 for the classification of word pairs.
 
@@
 In this experiment, standard metrics: precision, recall, and F-measure, are applied to evaluate the performance of binary classifiers\cite{aas1999text}.
-However, we didn't report accuracy as a performance metric because accuracy is highly sensitive to the prior class probabilities and does not fully describe the actual difficulty of the decision problem for unbalanced dataset.
+However, we didn't report accuracy as a performance metric because accuracy is highly sensitive to the prior class probabilities and does not fully describe the actual difficulty of the decision problem for an unbalanced dataset.
 We conduct the experiment with 5 folds cross-validation and weighted macro-averaging of these metrics over the folds.
-All models are trained on 80\% of the word pairs and remaining 20\% of the data is used as a test set for reporting the performance of the model.
-We also estimated the area under the receiving operating characterstics (ROC) curve\cite{kumar2011receiver} (AUC) metric due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets\cite{hu2015kernelized}.
+All models have trained on 80\% of the word pairs and remaining 20\% of the data is used as a test set for reporting the performance of the model.
+We also estimated the area under the receiving operating characteristics (ROC) curve\cite{kumar2011receiver} (AUC) metric due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets\cite{hu2015kernelized}.
 
 \section*{Results}
-opening para: evaluated with boundary, not boudary and overall \\
+Experimental results are evaluated with weighted average, ``boundary'', and ``not boudary'' classes shown in Table~\ref{tab:result_boundary}, \ref{tab:result_not_boundary} and \ref{tab:result_weighted_avg}, respectively.\\
 
@@
 
-para 1:
-0) opening with referencing table
-1) KNN best in all metrics except Recall
-2) RNN shows lowest performance in terms of recall and F1 score, although GRU better than LSTM 
-3) SVM is second highest model and highest model in terms of recall when only textual features is used.
-4) NB shows lowest precision value among all and better than RNN with recall and F1-score
-5) for additional features, recall and F1 increase x, y, z, for a, b, c, respectively
-6) However, precision increase for all except NB \\
+As follows from Table~\ref{tab:result_boundary}, KNN performs best among all machine learning models in terms of precision and F1-score, achieved 0.808 and 0.728 when contextual features are used, and 0.820 and 0.779 when a combination of contextual topic, and punctuation mark features are used.
+However, RNN demonstrates the lowest performance among all models in terms of recall and F1-score while GRU shows 3.72\%, 17.79\% and 11.47\% higher precision, recall, and F1-score than LSTM.
+SVM is the second highest model in terms of precision and F1-score and demonstrate the highest model by achieving 0.679 recall when only textual features are used.
+On the other hand, NB exhibits lowest precision value 0.594,  among all models but performs better than RNN with recall and F1-score.
+When additional features combined with textual features, recall increases 0.6\%, 2.5\%, and 11.92\%, and F1-score increases 0\%, 3.39\%, and 7\% for NB, SVM, and KNN models, respectively.
+Nevertheless, precision increases 4.31\% and 1.49\% for SVM and KNN models, except NB.
+\\
 
@@
 
-para 2:
-0) opening with referencing table
-2) Similar to boundary detection, KNN is best and RNN is lowest in all metrics for 'not boundary' identification
-3) Performance is very high compared to boundary detection however, perform better than random
-4) Effect of additional feature is consistent in 'not boundary' classification.
-5) for additional features, precision, recall and F1 increase x, y, z, for a, b, c, respectively \\
+Table~\ref{tab:result_not_boundary} illustrates the performance of NB, SVM, KNN, and RNN models for detecting no boundaries in eCoaching text.
+Similar to boundary detection, KNN consistently outperformed over all other models and RNN displays the lowest performance in terms of all performance metrics for 'not boundary' identification.
+Moreover, performance is very high in ``not boundary'' class compared to boundary detection.
+Effect of the additional feature is also consistent with 'not boundary' classification.
+Results show that precision, recall, and F1-score increases x\%, y\%, and z\%, for NB, SVM, KNN models, respectively.
+\\
 
@@
 
-para 3:
-0) opening with referencing table
-1) Overall, KNN is best and RNN is lowest in all metrics for 'not boundary' identification
-2) SVM and RNN achieved x and xx, y and yy, z and zz for precision, recall and F1, respectively.
-3) Effect of additional feature is also consistent.
-4) for additional features, precision, recall and F1 increase x, y, z, for a, b, c, respectively\\
+Table~\ref{tab:result_weighted_avg} depicts the performance of NB, SVM, KNN, and RNN models for the segmentation of eCoaching text in detecting both ``boundary'' and ``not boundary'' classes.
+Overall, KNN performed best and RNN is the lowest among all models for 'not boundary' identification.
+SVM and RNN achieved x\% and xx\%, y\% and yy\%, z\% and zz\% for precision, recall, and F1, respectively.
+Effect of the additional feature is also consistent as above, precision, recall, and F1 increases x\%, y\%, and z\%, for NB, SVM, and KNN, respectively.\\
 
@@
 KNN achieved 0.986 F1-score in overall, 0.779 and 0.993 F1-scores for detecting ``boundary'' and ``not boundary'', respectively.
-Robust performance of KNN provides the evidence that machine learning models are capable to learn information from the email exchange.
-Although the domain of this study was intentinally quite small, we believe that our study is not limited to ecoacing domain, and it can be successfully applied to other domain as well.
+The robust performance of KNN provides the evidence that machine learning models are capable to learn information from the email exchange.
+Although the domain of this study was intentionally quite small, we believe that our study is not limited to the ecoacing domain, and it can be successfully applied to other domain as well.
 
-The additional topic and punctuation mark features, made significant improvement in performance of all maching learning methods.
-In every cases, model performs better when contextual feature is used in combination with topic and punctuation mark features.
-Thsi results also mean that segmentation performance might be improved by adding more relevant features including human insight into the problem.
+The additional topic and punctuation mark feature made a significant improvement in performance of all machine learning methods.
+In every case, the model performs better when a contextual feature is used in combination with topic and punctuation mark features.
+This results also mean that segmentation performance might be improved by adding more relevant features including human insight into the problem.
 
@@
 In this paper, results are reported by each class to avoid confusion about the overall model performance.
-In addition, standard metrics: precision, recall and F1-score were used to eliminate doubt about the model performance because accuracy is misleading for imbalance dataset.
+In addition, standard metrics: precision, recall, and F1-score were used to eliminate doubt about the model performance because accuracy is misleading for imbalance dataset.
 AUC values are also outlined due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets\cite{hu2015kernelized}, which was demonstrated by the ROC curves in Figure~\ref{fig:roc-curves}.
 NB shows the highest AUC values and achieved 0.978 for both cases while it provides lowest classification results except RNN.
-On the other hand, KNN and SVM exibits 0.972 and 0.835 AUCs when only textual features are used; and 0.959 and 0.844 AUCs when combination of textual, topic and punctuation mark features are used.
-Finally, RNN demonstrate lowest AUC values among all machine learning models, achieved AUC values 0.704 and 0.740 for LSTM and GRU, respectively.
+On the other hand, KNN and SVM exhibit 0.972 and 0.835 AUCs when only textual features are used; and 0.959 and 0.844 AUCs when a combination of textual, topic and punctuation mark features are used.
+Finally, RNN demonstrates lowest AUC values among all machine learning models, achieved AUC values 0.704 and 0.740 for LSTM and GRU, respectively.
 
-We observed worst results of RNN, in particular LSTM and GRU for the text segmetation.
-We believe that RNN performed poorly because it has a large set of weights which required large set of data for both class.
+We observed worst results of RNN, in particular, LSTM and GRU for the text segmentation.
+We believe that RNN performed poorly because it has a large set of weights which required a large set of data for both classes.
 It this study, we utilized 3,138 examples of boundary case which failed to achieve good results.
-While performance of RNN is poor, GRU performed better than LSTM which was already observed in other previous study\cite{chung2014empirical}.
+While the performance of RNN is poor, GRU performed better than LSTM which was already observed in other previous study\cite{chung2014empirical}.
 
-Punctuation mark plays an important role in segmentation boundary detection, and a large numbers of errors were encountered by the false positive of boundary identification.
-Similarly, additional information is the common reason of classified original segment into multiple segments.
+Punctuation mark plays an important role in segmentation boundary detection, and large numbers of errors were encountered by the false positive of boundary identification.
+Similarly, additional information is the common reason for classified original segment into multiple segments.
 For example, [need help from April].
 
-Our proposed approach is a novel approach for text segmentation where previous study focus on the segmentation of text into sections, headers, and sentences.
+Our proposed approach is a novel approach for text segmentation where the previous study focuses on the segmentation of text into sections, headers, and sentences.
 However, this study segmented email exchange into ``block of text'' represent a particular MI behavior.
-This work will significantly reduce the amount of resource and time required to segemnt email text manually.
-Furthermore, this paper can help to annotate each segment automatically by building a new classifer, which will accelerate the pace of finding best communication strategies to develop an effective MI intervention for healthy eating.
+This work will significantly reduce the amount of resource and time required to segment email text manually.
+Furthermore, this paper can help to annotate each segment automatically by building a new classifier, which will accelerate the pace of finding best communication strategies to develop an effective MI intervention for healthy eating.
 
 As for future work, we plan to evaluate our methods with several similar and different datasets.
-We also plan to use combination of machine laerning and natural language approaches to improve model performance.
-For example, part-of-speech tagging and distance from boundary of the sentence might significantly enhance the classification performance.
-The limitation of this study is that eCoaching text are collected from a single medical institute; formatting, style, and email segment can be different in other settings.
+We also plan to use a combination of machine learning and natural language approaches to improve model performance.
+For example, part-of-speech tagging and distance from the boundary of the sentence might significantly enhance the classification performance.
+The limitation of this study is that eCoaching text is collected from a single medical institute; formatting, style, and email segment can be different in other settings.
 
@@
 In this paper, we compared the performance of machine learning models for the task of segmentation of e-coaching text.
-We found out that k-nearest neighbour provides the best performance for the segmentaion of text in terms of all performance metrics.
-Manual segmentation of e-coaching data is very resource-intensive and time consuming task, which can significantly decrease the time and effort required to develop effective behavioral interventions.
+We found out that k-nearest neighbor provides the best performance for the segmentation of text in terms of all performance metrics.
+Manual segmentation of e-coaching data is very resource-intensive and time-consuming task, which can significantly decrease the time and effort required to develop effective behavioral interventions.
 Our proposed methods can help to identify individual text segments, which can be annotated directly with a classification model and increase the effectiveness of behavioral interventions.

### 2018-03-05 paper-amia-2018/manuscript.tex
finished first draft without proofread

@@
 In this study, we transformed this task into the classification of detecting the boundary of segmentation and developed several state-of-the-art machine learning models including Support Vector Machine, Naive Bayes, K-Nearest Neighbor (KNN), Recurrent Neural Networks by utilizing contextual, topic and punctuation mark features.
-Results indicate that KNN is the best model and achieved 0.986 F1-score in overall, 0.779 and 0.993 F1-scores for detecting boundary, not a boundary, respectively.
+Results indicate that KNN is the best model and achieved 0.986 F1-measure in overall, 0.779 and 0.993 F1-measures for detecting ``boundary'' and ``not boundary'', respectively.
 This study has a great implication to identify individual text segments, which can be annotated directly with a classification model, and accelerate the pace of identifying effective communication strategies linked to healthy eating.}
@@
 \subsection*{\textit{Evaluation metrics}}
-In this experiment, standard metrics: precision, recall, and F-measure, are applied to evaluate the performance of binary classifiers\cite{aas1999text}.
+In this experiment, standard metrics: precision, recall, and F1-measure, are applied to evaluate the performance of binary classifiers\cite{aas1999text}.
 However, we didn't report accuracy as a performance metric because accuracy is highly sensitive to the prior class probabilities and does not fully describe the actual difficulty of the decision problem for an unbalanced dataset.
@@
 \section*{Results}
-Experimental results are evaluated with weighted average, ``boundary'', and ``not boudary'' classes shown in Table~\ref{tab:result_boundary}, \ref{tab:result_not_boundary} and \ref{tab:result_weighted_avg}, respectively.\\
+Experimental results are evaluated with ``boundary'' and ``not boundary'' classes as well as their weighted average, which are shown in Table~\ref{tab:result_boundary}, \ref{tab:result_not_boundary} and \ref{tab:result_weighted_avg}, respectively.\\
 
@@
    \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{contextual features only}} & \multicolumn{3}{|c|}{\textbf{contextual + punctuation marks (+ topics except RNN)}} \\\cline{2-7}
-   & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Score} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Score}\\ \hline
+   & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-measure} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-measure}\\ \hline
 
@@
 
-As follows from Table~\ref{tab:result_boundary}, KNN performs best among all machine learning models in terms of precision and F1-score, achieved 0.808 and 0.728 when contextual features are used, and 0.820 and 0.779 when a combination of contextual topic, and punctuation mark features are used.
-However, RNN demonstrates the lowest performance among all models in terms of recall and F1-score while GRU shows 3.72\%, 17.79\% and 11.47\% higher precision, recall, and F1-score than LSTM.
-SVM is the second highest model in terms of precision and F1-score and demonstrate the highest model by achieving 0.679 recall when only textual features are used.
-On the other hand, NB exhibits lowest precision value 0.594,  among all models but performs better than RNN with recall and F1-score.
-When additional features combined with textual features, recall increases 0.6\%, 2.5\%, and 11.92\%, and F1-score increases 0\%, 3.39\%, and 7\% for NB, SVM, and KNN models, respectively.
-Nevertheless, precision increases 4.31\% and 1.49\% for SVM and KNN models, except NB.
+As follows from Table~\ref{tab:result_boundary}, KNN performs best among all machine learning models in terms of precision and F1-measure, achieved 0.808 precision with 0.728 F1-measure when contextual features are used, and 0.820 precision with 0.779 F1-measure when a combination of contextual topic, and punctuation mark features are used.
+However, RNN demonstrates the lowest performance among all models in terms of recall and F1-measure while GRU shows 3.72\%, 17.79\% and 11.47\% higher precision, recall, and F1-measure than LSTM.
+In this study, SVM appears as the second highest model in terms of precision and F1-measure, obtains highest 0.679 recall when only textual features are used.
+On the other hand, NB exhibits lowest precision value 0.594 but performs better than RNN with recall and F1-measure.
+When textual features are used in combination with topic and punctuation mark, recall increases by 0.6\%, 2.5\%, and 11.92\%, and F1-measure increases by 0\%, 3.39\%, and 7\% for NB, SVM, and KNN models, respectively.
+Nevertheless, precision increases by 4.31\% and 1.49\% for SVM and KNN methods while decreases by 0.7\% in NB.
 \\
@@
 
-Table~\ref{tab:result_not_boundary} illustrates the performance of NB, SVM, KNN, and RNN models for detecting no boundaries in eCoaching text.
-Similar to boundary detection, KNN consistently outperformed over all other models and RNN displays the lowest performance in terms of all performance metrics for 'not boundary' identification.
-Moreover, performance is very high in ``not boundary'' class compared to boundary detection.
-Effect of the additional feature is also consistent with 'not boundary' classification.
-Results show that precision, recall, and F1-score increases x\%, y\%, and z\%, for NB, SVM, KNN models, respectively.
+Table~\ref{tab:result_not_boundary} summarizes the performance of NB, SVM, KNN, and RNN models for detecting ``not boundary'' class in eCoaching text.
+Performance is remarkably high in ``not boundary'' class compared to boundary detection.
+Similar to boundary detection, KNN consistently outperforms over all other methods but obtains 22.40\%, 50.07\%, and 36.26\% higher precision, recall, and F1-measure for contextual feature and 20.85\%, 33.96\%, and 27.47\% higher precision, recall, and F1-measure for combined features.
+However, RNN demonstrates the lowest performance and exhibits 0.981 and 0.983 precision with 0.986 and 0.987 F1-measure for LSTM and GRU, respectively.
+Impact of additional features is also consistent in ``not boundary'' classification.
+Results show that F1-measure increases by 0\%, and 0.1\% for SVM and KNN models although decreases by 0.1\% for NB.
 \\
@@
 \centering
-\caption{Weighted average performance of NB, SVM, KNN, and RNN for the segmentation of eCoaching text in detecting both ``boundary'' and ``not boundary''.
+\caption{Weighted average performance of NB, SVM, KNN, and RNN for the segmentation of eCoaching text in detecting both ``boundary'' and ``not boundary'' classes.
 The highest value for each performance metric is highlighted in bold.}
@@
 
-Table~\ref{tab:result_weighted_avg} depicts the performance of NB, SVM, KNN, and RNN models for the segmentation of eCoaching text in detecting both ``boundary'' and ``not boundary'' classes.
-Overall, KNN performed best and RNN is the lowest among all models for 'not boundary' identification.
-SVM and RNN achieved x\% and xx\%, y\% and yy\%, z\% and zz\% for precision, recall, and F1, respectively.
-Effect of the additional feature is also consistent as above, precision, recall, and F1 increases x\%, y\%, and z\%, for NB, SVM, and KNN, respectively.\\
+Table~\ref{tab:result_weighted_avg} outlines the weighted average results of the experiment on the models for the segmentation of eCoaching text by classifying them into ``boundary'' and ``not boundary'' classes.
+Overall, KNN obtains the best performance with all metrics and RNN denotes the lowest performance among all methods.
+NB and SVM demonstrate moderate performace, obtain precision 0.975 and 0.981, recall 0.974 and 0.982, and F1-measure 0.975 and 0.983 when textual features are used.
+Effect of the additional features is also consistent as above, precision increases by 0.1\%, 0.2\%, and 0.3\%, recall increases by 0\%, 0.1\%, and 0.2\%, and F1-measure increases by 0\%, 0.2\%, and 0.3\%, for NB, SVM, and KNN methods, respectively, when combined features are used.\\
 
@@
 Experimental results indicate that KNN is the best model among all machine learning methods considered for this study.
-KNN achieved 0.986 F1-score in overall, 0.779 and 0.993 F1-scores for detecting ``boundary'' and ``not boundary'', respectively.
+KNN achieved 0.986 F1-measure in overall, 0.779 and 0.993 F1-measures for detecting ``boundary'' and ``not boundary'', respectively.
 The robust performance of KNN provides the evidence that machine learning models are capable to learn information from the email exchange.
@@
 In this paper, results are reported by each class to avoid confusion about the overall model performance.
-In addition, standard metrics: precision, recall, and F1-score were used to eliminate doubt about the model performance because accuracy is misleading for imbalance dataset.
+In addition, standard metrics: precision, recall, and F1-measure were used to eliminate doubt about the model performance because accuracy is misleading for imbalance dataset.
 AUC values are also outlined due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets\cite{hu2015kernelized}, which was demonstrated by the ROC curves in Figure~\ref{fig:roc-curves}.
@@
 Punctuation mark plays an important role in segmentation boundary detection, and large numbers of errors were encountered by the false positive of boundary identification.
-Similarly, additional information is the common reason for classified original segment into multiple segments.
+Similarly, additional information is the common reason for the classified original segment into multiple segments.
 For example, [need help from April].
@@
 
-As for future work, we plan to evaluate our methods with several similar and different datasets.
+As for future work, As our future work, we plan to evaluate our approach on other datasets involves in discourse analysis.
 We also plan to use a combination of machine learning and natural language approaches to improve model performance.
@@
 Manual segmentation of e-coaching data is very resource-intensive and time-consuming task, which can significantly decrease the time and effort required to develop effective behavioral interventions.
-Our proposed methods can help to identify individual text segments, which can be annotated directly with a classification model and increase the effectiveness of behavioral interventions.
+Our proposed methods can help to identify individual text segments, which can be annotated directly with a classification model.
 This approach will help for developing fully automated eCoaching and also accelerate the pace of identifying effective communication strategies linked to healthy eating.
-As our future work, we plan to evaluate our approach on other datasets involves in discourse analysis for enhancing our proposed method.
 

### 2018-03-06 paper-amia-2018/manuscript.tex
investigate different RNN models

@@
 Developing an automatic classification of clinical interactions required segmented text.
-Nevertheless, eCoaching data comprised of email responses which need to be segmented into a group of MI behavior refers to ``block of text''.
+Nevertheless, eCoaching data comprised of email responses which need to be segmented into groups of MI behaviors.
 Automatic segmentation of eCoaching intervention sessions is a challenging task due to the 2 important reasons.
@@
 On the other hand, SVM exploiting with linear kernel and recurrent convolutional neural networks with posodic, part of speech features and word embeddings, were trained by Kreuzthaler et al.\cite{kreuzthaler2015detection} and Griffis et al.\cite{griffis2016quantitative}, respectively, for the detection of sentence boundary.
-However, segmentation of clinical text, in particular, segmentation of MI or eCoaching text into a group of MI behavior is ignored while relying on manual hand-coded approach.
-Therefore, this study introduces a novel approach and the authors are not aware of any other work this approach has been considered for the segmentation of MI or eCoaching text into ``block of text''.
+However, segmentation of clinical text, in particular, segmentation of MI or eCoaching text into groups of MI behaviors, is ignored while relying on manual hand-coded approach.
+Therefore, this study introduces a novel approach and the authors are not aware of any other work this approach has been considered for the segmentation of eCoaching text.
 
@@
 
-Our proposed approach is a novel approach for text segmentation where the previous study focuses on the segmentation of text into sections, headers, and sentences.
-However, this study segmented email exchange into ``block of text'' represent a particular MI behavior.
+Our proposed approach is novel for text segmentation of eCoaching text because previous studies focus on the segmentation of text into sections, headers, and sentences.
+However, this study segmented email exchange into groups of MI behavior.
 This work will significantly reduce the amount of resource and time required to segment email text manually.
@@
 Segmentation of eCoaching text is an integral part of developing an automated eCoaching intervention.
-Although several studies have done for the segmentation of clinical text into sections and sentences, none of them are used for the segmentation of text into a group of MI behavior in the setting of discourse analysis with email under the principle of motivational interviews.
+Although several studies have done for the segmentation of clinical text into sections and sentences, none of them are used for the segmentation of text into a groups of MI behaviors in the setting of discourse analysis with email under the principle of motivational interviews.
 In this paper, we compared the performance of machine learning models for the task of segmentation of e-coaching text.

### 2018-03-06 rnn/seg-gru-model-bidirect.py

@@ -1,0 +1,83 @@
+# RNN with variable length input sequences to one word output
+import numpy
+
+from keras.models import Sequential
+from keras.layers import Dense
+from keras.layers import GRU
+from keras.layers import LSTM
+from keras.layers import Bidirectional
+from keras.layers.embeddings import Embedding
+from keras.callbacks import EarlyStopping, ModelCheckpoint
+from keras.regularizers import l1_l2
+
+import utility
+
+from vocabulary import Vocabulary
+
+# fix random seed for reproducibility
+numpy.random.seed(42)
+
+# create e-coaching dictionary
+vocabulary_obj = Vocabulary()
+vocabulary_obj.create()
+
+# create mapping of characters to integers (0-25) and the reverse
+char_to_int = dict((c, i+1) for i, c in enumerate(vocabulary_obj.vocab))
+int_to_char = dict((i+1, c) for i, c in enumerate(vocabulary_obj.vocab))
+
+# system configuration
+max_len = 100
+folds = 5
+embed_dim = 30
+lstm_out = 32
+dropout_x = 0.5
+batch_size = 64
+
+# prepare the data set of input to output pairs encoded as integers
+data = utility.get_ecoaching_data()
+folds_data = utility.split_data(data, folds)
+results = []
+
+for fold_num in range(folds):
+
+    train_data, test_data = [], []
+    for i in range(folds):
+        if i == fold_num:
+            test_data = folds_data[i]
+        else:
+            train_data = train_data + folds_data[i]
+
+    # get encoded training data
+    X, y = utility.get_encoded_long_sequence(train_data, vocabulary_obj, max_len)
+
+    # print total number of instances
+    print("# of instances: ", len(y))
+
+    # create and fit the model
+    model = Sequential()
+    model.add(Embedding(vocabulary_obj.size(), embed_dim, input_length=X.shape[1]))
+    # model.add(GRU(lstm_out, recurrent_regularizer=l1_l2(l1=0.0, l2=0.033), dropout=dropout_x))
+    # model.add(LSTM(lstm_out, input_shape=(X.shape[1], 1)))
+    model.add(Bidirectional(GRU(lstm_out)))
+    model.add(Dense(y.shape[1], activation='softmax'))
+    model.compile(loss='categorical_crossentropy', optimizer='adam', metrics=['accuracy'])
[…]

### 2018-03-06 rnn/seg-gru-model-embed.py

@@ -1,0 +1,86 @@
+# RNN with variable length input sequences to one word output
+import numpy
+
+from keras.models import Sequential
+from keras.layers import Dense
+from keras.layers import GRU
+from keras.layers import LSTM
+from keras.layers.embeddings import Embedding
+from keras.callbacks import EarlyStopping, ModelCheckpoint
+from keras.regularizers import l1_l2
+
+import utility
+
+from vocabulary import Vocabulary
+
+# fix random seed for reproducibility
+numpy.random.seed(42)
+
+# create e-coaching dictionary
+vocabulary_obj = Vocabulary()
+vocabulary_obj.create()
+
+# create mapping of characters to integers (0-25) and the reverse
+char_to_int = dict((c, i+1) for i, c in enumerate(vocabulary_obj.vocab))
+int_to_char = dict((i+1, c) for i, c in enumerate(vocabulary_obj.vocab))
+
+# system configuration
+folds = 5
+embed_dim = 30
+lstm_out = 32
+dropout_x = 0.5
+batch_size = 64
+max_len = 3
+
+# prepare the data set of input to output pairs encoded as integers
+data = utility.get_ecoaching_data()
+folds_data = utility.split_data(data, folds)
+results = []
+
+# print maximum length of the sequence
+print("maximum length of the sequence: ", max_len)
+
+for fold_num in range(folds):
+
+    train_data, test_data = [], []
+    for i in range(folds):
+        if i == fold_num:
+            test_data = folds_data[i]
+        else:
+            train_data = train_data + folds_data[i]
+
+    # get encoded training data
+    X, y = utility.get_encoded_data_rnn_embed(train_data, vocabulary_obj, max_len)
+
+    # print total number of instances
+    print("# of instances: ", len(y))
+
+    # create and fit the model
+    model = Sequential()
+    model.add(Embedding(vocabulary_obj.size(), embed_dim, input_length=X.shape[1]))
+    # model.add(GRU(lstm_out, recurrent_regularizer=l1_l2(l1=0.0, l2=0.033), dropout=dropout_x))
+    # model.add(LSTM(lstm_out, input_shape=(X.shape[1], 1)))
+    # model.add(GRU(lstm_out, input_shape=(X.shape[1], 1)))
+    model.add(GRU(lstm_out))
+    model.add(Dense(y.shape[1], activation='softmax'))
[…]

### 2018-03-06 rnn/seg-gru-model-long.py

@@ -1,0 +1,86 @@
+# RNN with variable length input sequences to one word output
+import numpy
+
+from keras.models import Sequential
+from keras.layers import Dense
+from keras.layers import GRU
+from keras.layers import LSTM
+from keras.layers.embeddings import Embedding
+from keras.callbacks import EarlyStopping, ModelCheckpoint
+from keras.regularizers import l1_l2
+
+import utility
+
+from vocabulary import Vocabulary
+
+# fix random seed for reproducibility
+numpy.random.seed(42)
+
+# create e-coaching dictionary
+vocabulary_obj = Vocabulary()
+vocabulary_obj.create()
+
+# create mapping of characters to integers (0-25) and the reverse
+char_to_int = dict((c, i+1) for i, c in enumerate(vocabulary_obj.vocab))
+int_to_char = dict((i+1, c) for i, c in enumerate(vocabulary_obj.vocab))
+
+# system configuration
+folds = 5
+embed_dim = 30
+lstm_out = 32
+dropout_x = 0.5
+batch_size = 64
+max_len = 3
+
+# prepare the data set of input to output pairs encoded as integers
+data = utility.get_ecoaching_data()
+folds_data = utility.split_data(data, folds)
+results = []
+
+# print maximum length of the sequence
+print("maximum length of the sequence: ", max_len)
+
+for fold_num in range(folds):
+
+    train_data, test_data = [], []
+    for i in range(folds):
+        if i == fold_num:
+            test_data = folds_data[i]
+        else:
+            train_data = train_data + folds_data[i]
+
+    # get encoded training data
+    X, y = utility.get_encoded_data_rnn(train_data, vocabulary_obj, max_len)
+
+    # print total number of instances
+    print("# of instances: ", len(y))
+
+    # create and fit the model
+    model = Sequential()
+    # model.add(Embedding(vocabulary_obj.size(), embed_dim, input_length=X.shape[1]))
+    # model.add(GRU(lstm_out, recurrent_regularizer=l1_l2(l1=0.0, l2=0.033), dropout=dropout_x))
+    # model.add(LSTM(lstm_out, input_shape=(X.shape[1], 1)))
+    model.add(GRU(lstm_out, input_shape=(X.shape[1], 1)))
+    # model.add(GRU(lstm_out))
+    model.add(Dense(y.shape[1], activation='softmax'))
[…]

### 2018-03-06 rnn/utility.py

@@ -62,6 +62,42 @@ def get_encoded_data(data, vocabulary_obj, max_len):
 
 
 # encode text data into numeric values
+def get_encoded_long_sequence(data, vocabulary_obj, max_len):
+    x, y = [], []
+
+    for text in data:
+        input_text = vocabulary_obj.vectorize(text)
+        sample_len = len(input_text)
+        idx = 2
+
+        while idx < sample_len:
+            if idx == sample_len-1:
+                x.append(input_text[0:sample_len])
+                y.append(1)
+
+                # for i in input_text[0:sample_len]:
+                #     sys.stdout.write(str(vocabulary_obj.int_to_word[i] + ', '))
+                # sys.stdout.write("1\n")
+            else:
+                x.append(input_text[0:idx])
+                y.append(0)
+
+                # for i in input_text[0:idx]:
+                #     sys.stdout.write(str(vocabulary_obj.int_to_word[i] + ', '))
+                # sys.stdout.write("0\n")
+
+            idx += 1
+
+    data_x = numpy.array(pad_sequences(x, max_len, padding='pre', truncating='pre', dtype='float32'))
+    # reshape X to be [samples, time steps, features]
+    # data_x = numpy.reshape(data_x, (data_x.shape[0], max_len, 1))
+    # normalize
+    data_x = data_x / vocabulary_obj.size()
+    data_y = np_utils.to_categorical(y)
+    return data_x, data_y
+
+
+# encode text data into numeric values
 def get_encoded_data_all(data, vocabulary_obj, max_len, window_size=1):
     x, y = [], []
     previous_x, current_x = [], []
@@ -172,6 +208,66 @@ def get_encoded_data_rnn(data, vocabulary_obj, max_len):
     return data_x, data_y
 
 
+# encode text data into numeric values
+def get_encoded_data_rnn_embed(data, vocabulary_obj, max_len):
+    numpy.random.seed(42)
+    x, y = [], []
+    previous_x, current_x = [], []
+
+    for text in data:
+        # print text
+        input_text = vocabulary_obj.vectorize(text)
+        sample_len = len(input_text)
+
+        if sample_len >= 1:
+            current_x = input_text[0:1]
+            x.append(previous_x + current_x)
+            y.append(1)
+
+            # for idx in previous_x + current_x:
+            #     sys.stdout.write(str(vocabulary_obj.int_to_word[idx] + ', '))
+            # sys.stdout.write("1\n")
+
+            previous_x = input_text[0:1]
+            current_x = []
+
[…]

### 2018-03-06 paper-amia-2018/manuscript.tex
fixed spelling and grammatical errors

@@
 
-\title{Machine Learning Models for the Segmentation of eCoaching Text}
+\title{Machine Learning Models for the Segmentation of E-coaching Text}
 
@@
 An effective intervention is required to improve eating habits.
-eCoaching is an email-based intervention technique where a critical step is the segmentation of text for the automatic annotation of the email exchange.
-In this study, we transformed this task into the classification of detecting the boundary of segmentation and developed several state-of-the-art machine learning models including Support Vector Machine, Naive Bayes, K-Nearest Neighbor (KNN), Recurrent Neural Networks by utilizing contextual, topic and punctuation mark features.
+E-coaching is an email-based intervention technique where a critical step is the segmentation of text for the automatic annotation of the email exchange.
+In this study, we transformed this task into a classification of detecting the boundary of segmentation and developed several state-of-the-art machine learning models including Support Vector Machine, Naive Bayes, K-Nearest Neighbor (KNN), Recurrent Neural Networks by utilizing contextual, topic and punctuation mark features.
 Results indicate that KNN is the best model and achieved 0.986 F1-measure in overall, 0.779 and 0.993 F1-measures for detecting ``boundary'' and ``not boundary'', respectively.
@@
 MENU GenY\cite{alexander2017motivations} (Making Effective Nutrition Choices for Generation Y) is a technology-based public health intervention to encourage increased fruit and vegetable intake among GenY.
-A critical component of MENU GenY is personalized eCoaching.
-eCoaches use email to deliver motivation-enhancing coaching to encourage healthy eating, grounded in the principles of Motivational Interviewing (MI), an evidence-based communication technique to increase intrinsic motivation and self-efficacy for behavior change\cite{miller2012motivational,miller2009ten,miller2009toward}.
+A critical component of MENU GenY is personalized e-coaching.
+E-coaches use email to deliver motivation-enhancing coaching to encourage healthy eating, grounded in the principles of Motivational Interviewing (MI), an evidence-based communication technique to increase intrinsic motivation and self-efficacy for behavior change\cite{miller2012motivational,miller2009ten,miller2009toward}.
 Patient ``change talk'', statements of intrinsic motivation about their desire, ability, reasons, need for and commitment to behavior change, is an established mediator of health behavior change\cite{apodaca2009mechanisms}.
@@
 
-However, a significant barrier to fully automate eCoaching is the unsegmented text data.
-Developing an automatic classification of clinical interactions required segmented text.
-Nevertheless, eCoaching data comprised of email responses which need to be segmented into groups of MI behaviors.
-Automatic segmentation of eCoaching intervention sessions is a challenging task due to the 2 important reasons.
+However, a significant barrier to fully automate e-coaching is the unsegmented email exchange.
+E-coaching data comprised of email responses which need to be segmented into groups of MI behaviors for developing an automatic classification of clinical interactions.
+However, automatic segmentation of e-coaching intervention sessions is a challenging task due to the 2 important reasons.
 First, the email is an unstructured text that contains informal email exchange in non-traditional formats.
@@
 One sentence can be segmented into several MI behaviors and vice versa.
-Figure~\ref{fig:text-segment} illustrates the segmentation of an eCoaching email exchange where the first sentence segmented into 2 different MI behaviors.
+Figure~\ref{fig:text-segment} illustrates the segmentation of an e-coaching email exchange where the first sentence segmented into 2 different MI behaviors.
 On the other hand, fourth and fifth segments contain only one and multiple sentences, respectively.
@@
     \includegraphics[width=0.9\textwidth]{figures/segment-example.png}
-    \caption{\textbf{Segmentation of eCoaching text depicts the main challenges of boundary detection}.}
+    \caption{\textbf{Segmentation of e-coaching text depicts the main challenges of boundary detection}.}
     \label{fig:text-segment}
@@
 
-In this paper, we address this problem by developing several state-of-the-art machine learning based models for the segmentation of eCoaching text to promote the automatic identification of best communication strategies without human interference.
-More specifically, we develop Support Vector Machine (SVM), Naive Bayes (NB), K-Nearest Neighbor (KNN), Long Short Term Memory (LSTM), and Gated Recurrent Unit (GRU) by utilizing contextual, topic and punctuation mark features, to find the best model for the segmentation of eCoaching text.
+In this paper, we address the text segmentation problem by developing several state-of-the-art machine learning models to promote the automatic identification of best communication strategies without human interference.
+More specifically, we developed Support Vector Machine (SVM), Naive Bayes (NB), K-Nearest Neighbor (KNN), Long Short Term Memory (LSTM), and Gated Recurrent Unit (GRU) by utilizing contextual, topic and punctuation mark features, to find the best model for the segmentation of e-coaching text.
 
@@
 In 2009, Denny et al.\cite{denny2009evaluation} proposed a SecTag algorithm, which combined natural language processing technique, terminology-based rule, and naive Bayesian score for identifying sections and headers that achieved 99\% recall with 95.6\% precision.
-On the other hand, SVM exploiting with linear kernel and recurrent convolutional neural networks with posodic, part of speech features and word embeddings, were trained by Kreuzthaler et al.\cite{kreuzthaler2015detection} and Griffis et al.\cite{griffis2016quantitative}, respectively, for the detection of sentence boundary.
-However, segmentation of clinical text, in particular, segmentation of MI or eCoaching text into groups of MI behaviors, is ignored while relying on manual hand-coded approach.
-Therefore, this study introduces a novel approach and the authors are not aware of any other work this approach has been considered for the segmentation of eCoaching text.
+On the other hand, SVM exploiting with linear kernel and recurrent convolutional neural networks with prosodic, part of speech features and word embeddings, were trained by Kreuzthaler et al.\cite{kreuzthaler2015detection} and Griffis et al.\cite{griffis2016quantitative}, respectively, for the detection of sentence boundary.
+However, segmentation of clinical text, in particular, segmentation of MI or e-coaching text into groups of MI behaviors, is ignored while relying on manual hand-coded approach.
+Therefore, this study introduces a novel approach and the authors are not aware of any other work this approach has been considered for the segmentation of e-coaching text.
 
@@
 \subsection*{\textit{Data collection}}
-The experimental dataset for this work was constructed from the 49 eCoaching sessions, which include a total of 3,138 segmented and annotated MI behaviors.
+The experimental dataset for this work was constructed from the 49 e-coaching sessions, which include a total of 3,138 segmented and annotated MI behaviors.
 Each session contains an MI intervention involving patient-provider communications in email.
@@
 \subsection*{\textit{Segmentation classifiers}}
-Several state-of-the-art classifiers, including Naive Bayes (NB)\cite{pedregosa2011scikit}, Support Vector Machine (SVM)\cite{chang2011libsvm}, K-Nearest Neighbour (KNN)\cite{pedregosa2011scikit}, two variant of Recurrent Neural Networks (RNN)\cite{bengio1993problem}: Long Short Term Memory (LSTM)\cite{hochreiter1997long} and Gated Recurrent Unit (GRU)\cite{cho2014properties}, are employed to estimate the classification performance.
+Several state-of-the-art classifiers, including Naive Bayes (NB)\cite{pedregosa2011scikit}, Support Vector Machine (SVM)\cite{chang2011libsvm}, K-Nearest Neighbor (KNN)\cite{pedregosa2011scikit}, two variant of Recurrent Neural Networks (RNN)\cite{bengio1993problem}: Long Short Term Memory (LSTM)\cite{hochreiter1997long} and Gated Recurrent Unit (GRU)\cite{cho2014properties}, are employed to estimate the classification performance.
 
@@
 SVM finds the best hyperplane in the feature space that maximizes the separation between the closest ``boundary'' and ``not boundary'' training examples.
-In this experiment, the polynomial kernel is employed to train the SVM model for the segmentation of eCoaching text.
+In this experiment, the polynomial kernel is employed to train the SVM model for the segmentation of e-coaching text.
 
-\textbf{K-Nearest Neighbour}: By this model, each training sample represented as a point in the input feature space.
+\textbf{K-Nearest Neighbor}: By this model, each training sample represented as a point in the input feature space.
 For a new test sample, Euclidean distance is calculated to find the k-nearest neighbors.
@@
 We build our RNN model with one hidden layer, output layer, and input layer which get one hot encoding of word vector as input.
-Since one-hot vector is given in the input layer, results are reported with textual features and punctuation marks only.
-We experimentally determined that the best performance is achieved when the number of hidden units = 32, batch size = 8, optimizer = adam, as well as 600 epochs is used based on the validation loss.
+Since one-hot vector is given in the input layer, results are reported with textual and punctuation mark features only.
+We experimentally determined that the best performance is achieved when the number of hidden units = 32, batch size = 8, optimizer = adam, as well as 600 epochs is used.
 
@@
 In this experiment, standard metrics: precision, recall, and F1-measure, are applied to evaluate the performance of binary classifiers\cite{aas1999text}.
-However, we didn't report accuracy as a performance metric because accuracy is highly sensitive to the prior class probabilities and does not fully describe the actual difficulty of the decision problem for an unbalanced dataset.
+However, accuracy is not reported as a performance metric because accuracy is highly sensitive to the prior class probabilities and does not fully describe the actual difficulty of the decision problem for an unbalanced dataset.
 We conduct the experiment with 5 folds cross-validation and weighted macro-averaging of these metrics over the folds.
 All models have trained on 80\% of the word pairs and remaining 20\% of the data is used as a test set for reporting the performance of the model.
-We also estimated the area under the receiving operating characteristics (ROC) curve\cite{kumar2011receiver} (AUC) metric due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets\cite{hu2015kernelized}.
+We also estimate the area under the receiving operating characteristics (ROC) curve\cite{kumar2011receiver} (AUC) metric due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets\cite{hu2015kernelized}.
 
 \section*{Results}
-Experimental results are evaluated with ``boundary'' and ``not boundary'' classes as well as their weighted average, which are shown in Table~\ref{tab:result_boundary}, \ref{tab:result_not_boundary} and \ref{tab:result_weighted_avg}, respectively.\\
+Experimental results are evaluated with ``boundary'' and ``not boundary'' classes as well as their weighted average, which are shown in Table~\ref{tab:result_boundary}, \ref{tab:result_not_boundary}, and \ref{tab:result_weighted_avg}, respectively.\\
 
@@
 \centering
-\caption{Performance of NB, SVM, KNN, and RNN for detecting boundaries of segmentation in eCoaching text. The highest value for each performance metric is highlighted in bold.}
+\caption{Performance of NB, SVM, KNN, and RNN methods for detecting segmentation boundary in e-coaching text. The highest value for each performance metric is highlighted in bold.}
 \label{tab:result_boundary}
@@
 
-As follows from Table~\ref{tab:result_boundary}, KNN performs best among all machine learning models in terms of precision and F1-measure, achieved 0.808 precision with 0.728 F1-measure when contextual features are used, and 0.820 precision with 0.779 F1-measure when a combination of contextual topic, and punctuation mark features are used.
+As follows from Table~\ref{tab:result_boundary}, KNN performs best among all machine learning models in terms of precision and F1-measure, achieved 0.808 precision with 0.728 F1-measure when contextual features are used, and 0.820 precision with 0.779 F1-measure when a combination of contextual, topic, and punctuation mark features are used.
 However, RNN demonstrates the lowest performance among all models in terms of recall and F1-measure while GRU shows 3.72\%, 17.79\% and 11.47\% higher precision, recall, and F1-measure than LSTM.
 In this study, SVM appears as the second highest model in terms of precision and F1-measure, obtains highest 0.679 recall when only textual features are used.
-On the other hand, NB exhibits lowest precision value 0.594 but performs better than RNN with recall and F1-measure.
-When textual features are used in combination with topic and punctuation mark, recall increases by 0.6\%, 2.5\%, and 11.92\%, and F1-measure increases by 0\%, 3.39\%, and 7\% for NB, SVM, and KNN models, respectively.
+On the other hand, NB exhibits lowest precision value 0.594, provides better recall and F1-measure than RNN model.
+When textual features are used in combination with topic and punctuation mark, recall increases by 0.6\%, 2.5\%, and 11.92\%; and F1-measure increases by 0\%, 3.39\%, and 7\% for NB, SVM, and KNN models, respectively.
 Nevertheless, precision increases by 4.31\% and 1.49\% for SVM and KNN methods while decreases by 0.7\% in NB.
@@
 \centering
-\caption{Performance of NB, SVM, KNN, and RNN  for detecting no boundaries in eCoaching text. The highest value for each performance metric is highlighted in bold.}
+\caption{Performance of NB, SVM, KNN, and RNN methods for the identification of ``not boundary'' class. The highest value for each performance metric is highlighted in bold.}
 \label{tab:result_not_boundary}
@@
 
-Table~\ref{tab:result_not_boundary} summarizes the performance of NB, SVM, KNN, and RNN models for detecting ``not boundary'' class in eCoaching text.
-Performance is remarkably high in ``not boundary'' class compared to boundary detection.
-Similar to boundary detection, KNN consistently outperforms over all other methods but obtains 22.40\%, 50.07\%, and 36.26\% higher precision, recall, and F1-measure for contextual feature and 20.85\%, 33.96\%, and 27.47\% higher precision, recall, and F1-measure for combined features.
+Table~\ref{tab:result_not_boundary} summarizes the performance of NB, SVM, KNN, and RNN models for detecting ``not boundary'' class in e-coaching text.
+We observed that performance is remarkably high in ``not boundary'' class compared to boundary detection.
+Similar to boundary detection, KNN consistently outperforms over all other methods but obtains 22.40\%, 50.07\%, and 36.26\% higher precision, recall, and F1-measure for contextual features; and 20.85\%, 33.96\%, and 27.47\% higher precision, recall, and F1-measure for combined features compared to boundary class.
 However, RNN demonstrates the lowest performance and exhibits 0.981 and 0.983 precision with 0.986 and 0.987 F1-measure for LSTM and GRU, respectively.
-Impact of additional features is also consistent in ``not boundary'' classification.
+Impact of additional features is also consistent with ``not boundary'' classification.
 Results show that F1-measure increases by 0\%, and 0.1\% for SVM and KNN models although decreases by 0.1\% for NB.
@@
 \centering
-\caption{Weighted average performance of NB, SVM, KNN, and RNN for the segmentation of eCoaching text in detecting both ``boundary'' and ``not boundary'' classes.
+\caption{Weighted average performance of NB, SVM, KNN, and RNN methods for the segmentation of e-coaching text in detecting both ``boundary'' and ``not boundary'' classes.
 The highest value for each performance metric is highlighted in bold.}
@@
 
-Table~\ref{tab:result_weighted_avg} outlines the weighted average results of the experiment on the models for the segmentation of eCoaching text by classifying them into ``boundary'' and ``not boundary'' classes.
+Table~\ref{tab:result_weighted_avg} outlines the weighted average results of the experiment on the models for the segmentation of e-coaching text by classifying them into ``boundary'' and ``not boundary'' classes.
 Overall, KNN obtains the best performance with all metrics and RNN denotes the lowest performance among all methods.
-NB and SVM demonstrate moderate performace, obtain precision 0.975 and 0.981, recall 0.974 and 0.982, and F1-measure 0.975 and 0.983 when textual features are used.
-Effect of the additional features is also consistent as above, precision increases by 0.1\%, 0.2\%, and 0.3\%, recall increases by 0\%, 0.1\%, and 0.2\%, and F1-measure increases by 0\%, 0.2\%, and 0.3\%, for NB, SVM, and KNN methods, respectively, when combined features are used.\\
+NB and SVM demonstrate moderate performance, obtain precision 0.975 and 0.981; recall 0.974 and 0.982; and F1-measure 0.975 and 0.983 when textual features are used.
+Influence of the additional features is also consistent as above, precision increases by 0.1\%, 0.2\%, and 0.3\%; recall increases by 0\%, 0.1\%, and 0.2\%; and F1-measure increases by 0\%, 0.2\%, and 0.3\% for NB, SVM, and KNN methods, respectively, when combined features are used.\\
 
 \section*{Discussion}
-This study is the first large-scale efforts to evaluate the segmentation of eCoaching text.
+This study is the first large-scale efforts to evaluate the segmentation of e-coaching text.
 Experimental results indicate that KNN is the best model among all machine learning methods considered for this study.
@@
 The robust performance of KNN provides the evidence that machine learning models are capable to learn information from the email exchange.
-Although the domain of this study was intentionally quite small, we believe that our study is not limited to the ecoacing domain, and it can be successfully applied to other domain as well.
+Although the domain of this study was intentionally quite small, we believe that our study is not limited to the e-coaching domain, and it can be successfully applied to other domain as well.
 
 The additional topic and punctuation mark feature made a significant improvement in performance of all machine learning methods.
-In every case, the model performs better when a contextual feature is used in combination with topic and punctuation mark features.
+Nearly all cases, the model performs better when contextual features are used in combination with topic and punctuation mark features.
 This results also mean that segmentation performance might be improved by adding more relevant features including human insight into the problem.
@@
     \includegraphics[width=1.0\textwidth]{figures/roc-curves.png}
-    \caption{\textbf{Receiver operating characteristic curves showing the performance of binary classifiers for the segmentation of eCoaching text when textual features (left) and combination of textual and other features (right) are used}.}
+    \caption{\textbf{Receiver operating characteristic curves showing the performance of binary classifiers for the segmentation of e-coaching text when textual features (left) and combination of textual and other features (right) are used}.}
     \label{fig:roc-curves}
@@
 In addition, standard metrics: precision, recall, and F1-measure were used to eliminate doubt about the model performance because accuracy is misleading for imbalance dataset.
-AUC values are also outlined due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets\cite{hu2015kernelized}, which was demonstrated by the ROC curves in Figure~\ref{fig:roc-curves}.
-NB shows the highest AUC values and achieved 0.978 for both cases while it provides lowest classification results except RNN.
+AUC values are also outlined due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets\cite{hu2015kernelized}, which are demonstrated by the ROC curves in Figure~\ref{fig:roc-curves}.
+NB shows the highest AUC values, achieved 0.978 for both cases while provides lowest classification results except RNN.
 On the other hand, KNN and SVM exhibit 0.972 and 0.835 AUCs when only textual features are used; and 0.959 and 0.844 AUCs when a combination of textual, topic and punctuation mark features are used.
@@
 We believe that RNN performed poorly because it has a large set of weights which required a large set of data for both classes.
-It this study, we utilized 3,138 examples of boundary case which failed to achieve good results.
-While the performance of RNN is poor, GRU performed better than LSTM which was already observed in other previous study\cite{chung2014empirical}.
+In this study, we utilized 3,138 examples of boundary case which failed to achieve good results.
+While the performance of RNN is poor, GRU performs better than LSTM which was observed in other previous study\cite{chung2014empirical}.
 
@@
 
-Our proposed approach is novel for text segmentation of eCoaching text because previous studies focus on the segmentation of text into sections, headers, and sentences.
-However, this study segmented email exchange into groups of MI behavior.
+Our proposed approach is novel for the segmentation of e-coaching text because previous studies mainly focus on the segmentation of text into sections, headers, and sentences in other medical domain.
+However, this study segmented email exchange into groups of MI behaviors.
 This work will significantly reduce the amount of resource and time required to segment email text manually.
@@
 
-As for future work, As our future work, we plan to evaluate our approach on other datasets involves in discourse analysis.
+As our future work, we plan to evaluate our approach on other datasets involves in discourse analysis.
 We also plan to use a combination of machine learning and natural language approaches to improve model performance.
 For example, part-of-speech tagging and distance from the boundary of the sentence might significantly enhance the classification performance.
-The limitation of this study is that eCoaching text is collected from a single medical institute; formatting, style, and email segment can be different in other settings.
+The limitation of this study is that e-coaching text is collected from a single medical institute; formatting, style, and email segment can be different in other settings.
 
 \section*{Conclusion}
-Segmentation of eCoaching text is an integral part of developing an automated eCoaching intervention.
-Although several studies have done for the segmentation of clinical text into sections and sentences, none of them are used for the segmentation of text into a groups of MI behaviors in the setting of discourse analysis with email under the principle of motivational interviews.
+Segmentation of e-coaching text is an integral part of developing an automated e-coaching intervention.
+Although several studies have done for the segmentation of clinical text into sections and sentences, none of them are used for the segmentation of text into groups of MI behaviors in the setting of discourse analysis with email under the principle of motivational interviews.
 In this paper, we compared the performance of machine learning models for the task of segmentation of e-coaching text.
 We found out that k-nearest neighbor provides the best performance for the segmentation of text in terms of all performance metrics.
-Manual segmentation of e-coaching data is very resource-intensive and time-consuming task, which can significantly decrease the time and effort required to develop effective behavioral interventions.
+Manual segmentation of e-coaching data is very resource-intensive and time-consuming task, which can significantly decrease the time and effort required to develop an effective behavioral intervention.
 Our proposed methods can help to identify individual text segments, which can be annotated directly with a classification model.
-This approach will help for developing fully automated eCoaching and also accelerate the pace of identifying effective communication strategies linked to healthy eating.
+This approach will also help for developing fully automated e-coaching and accelerate the pace of identifying effective communication strategies linked to healthy eating.
 
@@
 This study was supported by a grant from the National Institutes of Health, NIDDK R21DK108071, Carcone and Kotov, MPIs.
-We would like to thank the student assistants in the Department of Family Medicine and Public Health Sciences at Wayne State University School of Medicine for their help in developing the training dataset by manually segmented the eCoaching text.
+We would like to thank the student assistants in the Department of Family Medicine and Public Health Sciences at Wayne State University School of Medicine for their help in developing the training dataset by manually segmented the e-coaching text.
 

### 2018-03-06 paper-amia-2018/manuscript.tex
update introduction

@@
 An effective intervention is required to improve eating habits.
-E-coaching is an email-based intervention technique where a critical step is the segmentation of text for the automatic annotation of the email exchange.
-In this study, we transformed this task into a classification of detecting the boundary of segmentation and developed several state-of-the-art machine learning models including Support Vector Machine, Naive Bayes, K-Nearest Neighbor (KNN), Recurrent Neural Networks by utilizing contextual, topic and punctuation mark features.
+E-coaching is an intervention technique where eCoaches use email to deliver motivation-enhancing coaching to encourage healthy eating.
+To identify effective communication strategies of healthy eating, automatic annotation of email exchange is required where a critical step is email segmentation.
+In this study, we transformed segmentation task into a classification task and developed several state-of-the-art machine learning models including Support Vector Machine, Naive Bayes, K-Nearest Neighbor (KNN), Recurrent Neural Networks by utilizing contextual, topic and punctuation mark features.
 Results indicate that KNN is the best model and achieved 0.986 F1-measure in overall, 0.779 and 0.993 F1-measures for detecting ``boundary'' and ``not boundary'', respectively.
-This study has a great implication to identify individual text segments, which can be annotated directly with a classification model, and accelerate the pace of identifying effective communication strategies linked to healthy eating.}
+This study has a great implication to accelerate the pace of identifying effective communication strategies linked to healthy eating.}
+%identify individual text segments, which can be annotated directly with a classification model, and
 
@@
 A major drawback of this research is the qualitative methods traditionally used to analyze the communication process which is resource-intensive, requiring an iterative process of human (subjective) interpretation of the text.
-Rapidly developing computational technologies, specifically machine learning combined with classification models, offer a unique opportunity to accelerate this process.
-Our research group has recently applied machine learning-based models to similar communication data\cite{hasan2016study,kotov2015interpretable}.
+Rapidly developing computational technologies, specifically, machine learning models, offer a unique opportunity to accelerate this process.
+Our research group has recently applied machine learning-based models to MI sessions \cite{hasan2016study,kotov2015interpretable}.
 A simple communication code scheme was automated to characterize patient communication and achieved accuracy comparable to human coders\cite{hasan2016study}.
-The ultimate goal of the research study is to leverage innovative machine learning models to fully automate the communication coding process in eCoach-patient communication to increases in fruit and vegetable intake.
+However, this study utilized e-coaching data which have different format and context compared to motivational interview transcripts.
+MI session data was prepared by transcribing audio conversation which involves a counselor, an adolescent, and a caregiver.
+On the other hand, e-coaching data was directly collected from the email responses between an eCoach and young adult.
+By utilizing this e-coaching data and leveraging innovative machine learning models, the ultimate goal of this research study is to fully automate the communication coding process in eCoach-patient communication to increases in fruit and vegetable intake.
 
@@
 On the other hand, SVM exploiting with linear kernel and recurrent convolutional neural networks with prosodic, part of speech features and word embeddings, were trained by Kreuzthaler et al.\cite{kreuzthaler2015detection} and Griffis et al.\cite{griffis2016quantitative}, respectively, for the detection of sentence boundary.
-However, segmentation of clinical text, in particular, segmentation of MI or e-coaching text into groups of MI behaviors, is ignored while relying on manual hand-coded approach.
+However, segmentation of clinical conversation in the context of electronically delivered interventions, in particular, segmentation of MI or e-coaching text into groups of MI behaviors, is ignored while relying on manual hand-coded approach.
 Therefore, this study introduces a novel approach and the authors are not aware of any other work this approach has been considered for the segmentation of e-coaching text.

### 2018-03-07 paper-amia-2018/manuscript_before_1st_feedback.tex
update introduction after April's feedback

--- 2018-03-06 paper-amia-2018/manuscript.tex
@@
 \author{Mehedi Hasan, BS$^{1}$\footnote[1]{Authors provided equal contribution.
-\label{footnote1}}, Alexander Kotov, PhD$^{1}$\textsuperscript{\ref{footnote1}}, April Idalski Carcone, PhD$^{2}$, Ming Dong, PhD$^{1}$, Sylvie Naar, PhD$^{2}$}
+\label{footnote1}}, Alexander Kotov, PhD$^{1}$\textsuperscript{\ref{footnote1}}, Ming Dong, PhD$^{1}$, Sylvie Naar, PhD$^{2}$, Gwen L.
+Alexander, PhD$^{3}$, April Idalski Carcone, PhD$^{4}$}
 
@@
 $^1$Department of Computer Science, Wayne State University, Detroit, Michigan \\  
-$^2$Department of Family Medicine and Public Health Sciences, School of Medicine, Wayne State University, Detroit, Michigan\\
+$^2$Director, Center for Translational Behavioral Research, Department of Behavioral Sciences and Social Medicine, Florida State University, Tallahassee, Florida\\
+$^3$Department of Public Health Sciences, Henry Ford Health System, Detroit, Michigan\\
+$^4$Department of Family Medicine and Public Health Sciences, School of Medicine, Wayne State University, Detroit, Michigan\\
 }
@@
 Unhealthy eating habits, particularly low fruit and vegetable intake, is a growing, serious public health concern, particularly among young adults age 21-30, referred to as Generation Y (GenY)\cite{blanck2008trends,centers2007fruit}.
-This generation has adopted a lifestyle that involves eating accessible, ``no mess'', quick, ``grab and go'' foods\cite{nebeling2007still,brug1999application}.
+GenY has adopted a lifestyle that involves eating accessible, ``no mess'', quick, ``grab and go'' foods\cite{nebeling2007still,brug1999application}.
 They mainly eat ``out'' and infrequently shop and prepare food, limiting access to fruit and vegetables (FV)\cite{nelson2009improving,larson2006food}.
@@
 Those in inner-city urban and rural settings have among the poorest eating habits\cite{blanck2008trends,centers2007fruit,ogden2006prevalence,american2006american,thompson2005dietary}.
-GenY's poor dietary practices placing them at high risk for obesity and many chronic diseases, such as type 2 diabetes, as well as declines in predicted health status and life expectancy.
+GenY's poor dietary practices place them at high risk for obesity and many chronic diseases, such as type 2 diabetes, as well as declines in predicted health status and life expectancy.
 Thus, there is a need to develop effective interventions to improve GenY's eating habits.
@@
 
-A major drawback of this research is the qualitative methods traditionally used to analyze the communication process which is resource-intensive, requiring an iterative process of human (subjective) interpretation of the text.
+A major drawback will be the qualitative methods traditionally used to analyze the communication process.
+Qualitative coding of communication data is resource-intensive, requiring an iterative process of human (subjective) interpretation of the text.
 Rapidly developing computational technologies, specifically, machine learning models, offer a unique opportunity to accelerate this process.
@@
 MI session data was prepared by transcribing audio conversation which involves a counselor, an adolescent, and a caregiver.
-On the other hand, e-coaching data was directly collected from the email responses between an eCoach and young adult.
+On the other hand, e-coaching data was directly collected from the email exchange between an eCoach and young adult.
 By utilizing this e-coaching data and leveraging innovative machine learning models, the ultimate goal of this research study is to fully automate the communication coding process in eCoach-patient communication to increases in fruit and vegetable intake.
 
-However, a significant barrier to fully automate e-coaching is the unsegmented email exchange.
-E-coaching data comprised of email responses which need to be segmented into groups of MI behaviors for developing an automatic classification of clinical interactions.
-However, automatic segmentation of e-coaching intervention sessions is a challenging task due to the 2 important reasons.
+A significant barrier to fully automated text classification is the segmentation of the text data.
+During traditional qualitative coding, coders determine where to segment text, that is, where one code ends and another begins.
+E-coaching data is comprised of email responses which are unsegmented, unlike more traditional diadic clinical communication where segmentation occurs naturally due to it's conversation nature.
+Automatic segmentation of e-coaching intervention sessions is a challenging task due to the 2 important reasons.
 First, the email is an unstructured text that contains informal email exchange in non-traditional formats.
-Second, a text segment not necessarily belongs to the entire sentence or collection of sentences.
-One sentence can be segmented into several MI behaviors and vice versa.
+Second, a text segment does not necessarily belong to the entire sentence or collection of sentences.
+One sentence can be segmented into multiple MI behaviors or several sentences may represent a single MI code.
 Figure~\ref{fig:text-segment} illustrates the segmentation of an e-coaching email exchange where the first sentence segmented into 2 different MI behaviors.
@@
 We observed worst results of RNN, in particular, LSTM and GRU for the text segmentation.
-We believe that RNN performed poorly because it has a large set of weights which required a large set of data for both classes.
-In this study, we utilized 3,138 examples of boundary case which failed to achieve good results.
+We believe that RNN performed poorly because it has a large set of weights which required longer sequence with a large set of data for both classes.
+In this study, maximum length of the input sequence was 3 and we utilized 3,138 examples of boundary case which failed to achieve good results.
 While the performance of RNN is poor, GRU performs better than LSTM which was observed in other previous study\cite{chung2014empirical}.

### 2018-03-07 paper-amia-2018/manuscript.tex
update introduction

@@
 \noindent{\bf Abstract}
-\textit{Poor eating habits, particularly low fruit and vegetable intake, is a growing, serious public health concern among young adults.
-An effective intervention is required to improve eating habits.
-E-coaching is an intervention technique where eCoaches use email to deliver motivation-enhancing coaching to encourage healthy eating.
-To identify effective communication strategies of healthy eating, automatic annotation of email exchange is required where a critical step is email segmentation.
+\textit{Communication science to understand clinical process like Motivational Interviews (MI) is limited by traditional qualitative coding methods.
+Qualitative coding of communication data is very resource-intensive and time-consuming process, requires auto coding.
+This study utilized e-coaching data where email is used to deliver motivation-enhancing coaching to encourage healthy eating.
+A critical step toward automatic annotation of communication coding process is the segmentation of text data.
 In this study, we transformed segmentation task into a classification task and developed several state-of-the-art machine learning models including Support Vector Machine, Naive Bayes, K-Nearest Neighbor (KNN), Recurrent Neural Networks by utilizing contextual, topic and punctuation mark features.
-Results indicate that KNN is the best model and achieved 0.986 F1-measure in overall, 0.779 and 0.993 F1-measures for detecting ``boundary'' and ``not boundary'', respectively.
-This study has a great implication to accelerate the pace of identifying effective communication strategies linked to healthy eating.}
-%identify individual text segments, which can be annotated directly with a classification model, and
+Results indicate that KNN is the best model and achieved 0.986 F1-measure in overall, 0.779 and 0.993 F1-measures for detecting ``boundary'' and ``not boundary'' classes, respectively.
+This study has a great implication to save money and accelerate the pace of identifying effective communication strategies.}
 
 \section*{Introduction}
-Unhealthy eating habits, particularly low fruit and vegetable intake, is a growing, serious public health concern, particularly among young adults age 21-30, referred to as Generation Y (GenY)\cite{blanck2008trends,centers2007fruit}.
-This generation has adopted a lifestyle that involves eating accessible, ``no mess'', quick, ``grab and go'' foods\cite{nebeling2007still,brug1999application}.
-They mainly eat ``out'' and infrequently shop and prepare food, limiting access to fruit and vegetables (FV)\cite{nelson2009improving,larson2006food}.
-Unfortunately, less than one-third of US adults\cite{blanck2008trends,ogden2006prevalence} and only 20\% of GenY\cite{blanck2008trends,american2006american,thompson2005dietary} eat the recommended 5 servings of fruit and vegetables daily.
-Those in inner-city urban and rural settings have among the poorest eating habits\cite{blanck2008trends,centers2007fruit,ogden2006prevalence,american2006american,thompson2005dietary}.
-GenY's poor dietary practices placing them at high risk for obesity and many chronic diseases, such as type 2 diabetes, as well as declines in predicted health status and life expectancy.
-Thus, there is a need to develop effective interventions to improve GenY's eating habits.
+Communication science to understand clinical process like Motivational Interviews (MI) is limited by traditional qualitative coding methods.
+Motivational Interviewing (MI) is an evidence-based communication technique to increase intrinsic motivation and self-efficacy for behavior change\cite{miller2012motivational,miller2009ten,miller2009toward}.
+Patient ``change talk'', statements of intrinsic motivation about their desire, ability, reasons, need for and commitment to behavior change, is an established mediator of health behavior change\cite{apodaca2009mechanisms}.
+Qualitative coding of communication data has been traditionally performed manually with pre-defined codes by trained coders, which is a tedious resource-intensive task, requiring an iterative process of human (subjective) interpretation of the text.
 
-GenY is a tech-savvy generation requiring an intervention matched to their mobile lifestyle.
-Growing numbers use the internet to access health information with the largest increases in internet access among low-income Americans, making the internet well-suited for health promotion intervention\cite{strecher2007internet}.
-MENU GenY\cite{alexander2017motivations} (Making Effective Nutrition Choices for Generation Y) is a technology-based public health intervention to encourage increased fruit and vegetable intake among GenY.
-A critical component of MENU GenY is personalized e-coaching.
-E-coaches use email to deliver motivation-enhancing coaching to encourage healthy eating, grounded in the principles of Motivational Interviewing (MI), an evidence-based communication technique to increase intrinsic motivation and self-efficacy for behavior change\cite{miller2012motivational,miller2009ten,miller2009toward}.
-Patient ``change talk'', statements of intrinsic motivation about their desire, ability, reasons, need for and commitment to behavior change, is an established mediator of health behavior change\cite{apodaca2009mechanisms}.
-Identifying specific communication strategies linked to behavior change and integrating these strategies into communication-based interventions (e.g., brief, motivation-enhancing interventions delivered in a variety of settings or public health initiatives) can increase these interventions' potency.
+Rapidly developing computational technologies, specifically, machine learning models, offer a unique opportunity to accelerate this process.
+A few studies were done for the automatic annotation of clinical interactions in the setting of MI intervention.
+Lacson et al.\cite{lacson2005automatic} applied AdaBoost\cite{freund1999short} classifier for annotating the interactions in hemodialysis phone dialog into 3 categories.
+Our research group has recently applied several machine learning-based models to MI sessions \cite{hasan2016study,kotov2015interpretable}.
+A simple communication code scheme was automated to characterize patient communication and achieved accuracy comparable to human coders\cite{hasan2016study}.
 
-A major drawback of this research is the qualitative methods traditionally used to analyze the communication process which is resource-intensive, requiring an iterative process of human (subjective) interpretation of the text.
-Rapidly developing computational technologies, specifically, machine learning models, offer a unique opportunity to accelerate this process.
-Our research group has recently applied machine learning-based models to MI sessions \cite{hasan2016study,kotov2015interpretable}.
-A simple communication code scheme was automated to characterize patient communication and achieved accuracy comparable to human coders\cite{hasan2016study}.
-However, this study utilized e-coaching data which have different format and context compared to motivational interview transcripts.
-MI session data was prepared by transcribing audio conversation which involves a counselor, an adolescent, and a caregiver.
-On the other hand, e-coaching data was directly collected from the email responses between an eCoach and young adult.
-By utilizing this e-coaching data and leveraging innovative machine learning models, the ultimate goal of this research study is to fully automate the communication coding process in eCoach-patient communication to increases in fruit and vegetable intake.
+Experimental data utilized in the above MI studies were prepared by transcribing audio conversation which involves a counselor, a patient, and some cases a caregiver\cite{hasan2016study,kotov2015interpretable}.
+However, this study utilized e-coaching data which have different structure and context compared to clinical data.
+E-coaches use email to deliver motivation-enhancing coaching to encourage healthy eating, grounded in the principles of motivational interviewing.
+E-coaching data is comprised of email responses which are unsegmented, unlike more traditional dyadic clinical communication where segmentation occurs naturally due to its conversation nature.
+Discourse analysis on e-coaching text can help to understand the communication process in e-coaching text by revealing the socio-psychological characteristics of a patient\cite{siegfried1995therapeutic}.
+By utilizing e-coaching data and leveraging innovative machine learning models, the ultimate goal of this research study is to fully automate the communication coding process.
 
-However, a significant barrier to fully automate e-coaching is the unsegmented email exchange.
-E-coaching data comprised of email responses which need to be segmented into groups of MI behaviors for developing an automatic classification of clinical interactions.
-However, automatic segmentation of e-coaching intervention sessions is a challenging task due to the 2 important reasons.
+A significant barrier to fully automated the communication coding process is the segmentation of the text data.
+During traditional qualitative coding, coders determine where to segment text, that is, where one code ends and another begins.
+Automatic segmentation of e-coaching intervention sessions is a challenging task due to the 2 important reasons.
 First, the email is an unstructured text that contains informal email exchange in non-traditional formats.
-Second, a text segment not necessarily belongs to the entire sentence or collection of sentences.
-One sentence can be segmented into several MI behaviors and vice versa.
+Second, a text segment does not necessarily belong to the entire sentence or collection of sentences.
+One sentence can be segmented into multiple MI behaviors or several sentences may represent a single MI code.
 Figure~\ref{fig:text-segment} illustrates the segmentation of an e-coaching email exchange where the first sentence segmented into 2 different MI behaviors.
@@
 On the other hand, SVM exploiting with linear kernel and recurrent convolutional neural networks with prosodic, part of speech features and word embeddings, were trained by Kreuzthaler et al.\cite{kreuzthaler2015detection} and Griffis et al.\cite{griffis2016quantitative}, respectively, for the detection of sentence boundary.
-However, segmentation of clinical conversation in the context of electronically delivered interventions, in particular, segmentation of MI or e-coaching text into groups of MI behaviors, is ignored while relying on manual hand-coded approach.
+
+Recently, an online clinical intervention called MENU GenY\cite{alexander2017motivations} (Making Effective Nutrition Choices for Generation Y) was tested.
+MENU GenY is a technology-based public health intervention to encourage increased fruit and vegetable intake among young adult age 21-30, utilized personalized e-coaching data.
+The main focus of MENU GenY is to develop a better coding dictionary among GenY to improve eating habits.
+However, segmentation of clinical conversation in the context of electronically delivered interventions, in particular, segmentation of clinical interaction text into groups of MI behaviors, is still ignored while relying on manual hand-coded approach.
 Therefore, this study introduces a novel approach and the authors are not aware of any other work this approach has been considered for the segmentation of e-coaching text.
@@
 Punctuation mark plays an important role in segmentation boundary detection, and large numbers of errors were encountered by the false positive of boundary identification.
+For example, ``xxxxx xxxxxx'' was segmented after the second sentence where a punctuation mark was encountered.
 Similarly, additional information is the common reason for the classified original segment into multiple segments.
-For example, [need help from April].
+For instance, ``xxxx xxxx xxxx'' was segmented after second sentence where third sentence supports the MI code confirmed by first two sentences.
 

### 2018-03-08 paper-amia-2018/manuscript.tex
update paper; working on LSTM model

@@
 
-\title{Machine Learning Models for the Segmentation of E-coaching Text}
+\title{Machine Learning Models for the Segmentation of Clinical Exchange}
 
@@
 \subsection*{\textit{Data collection}}
-The experimental dataset for this work was constructed from the 49 e-coaching sessions, which include a total of 3,138 segmented and annotated MI behaviors.
-Each session contains an MI intervention involving patient-provider communications in email.
+The experimental dataset for this work was constructed from 49 e-coaching sessions, which include a total of 3,138 segmented and annotated MI behaviors.
+Each session represents an MI intervention delivered via email.
 To filter out noise from the dataset, non-ascii characters are removed and then applied stemming to obtain a general form of word from different word representations, such as ``eating'', ``eats'', and ``eat''.
 We formulate the text segmentation task into a binary classification, as shown in Figure~\ref{fig:classifier}.
-An intervention session with email exchange is given as the input, it is partitioned into adjacent word pairs by sliding them.
-Each pair of them classified into either of the two categories: ``boundary'' and ``not boundary''.
-The text is segmented at the position, where an adjacent word pair classified into ``boundary'' class.
-If all pairs of word classified into ``not boundary'', the text is treated as one about a single MI behavior.
+Clinical exchange is given as the input, it is partitioned as adjacent word pairs by sliding them.
+Each pair is classified into either ``boundary'' or ``not boundary'' class.
+The original text is segmented at the position, where an adjacent word pair classified into ``boundary'' class.
+If all adjacent pairs of a block of text classified into ``not boundary'', the whole block of text is then treated as one segment about a single MI behavior.
 Totaly, we obtained 95,421 word pairs, which include 3,138 ``boundary'' and 92,283 ``not boundary'' instances.
@@
 Topics are considered as features since topic models are very effective\cite{kotov2015interpretable,hashimoto2016topic,lu2016modeling} to represent text documents.
-In this paper, we exploit the Labeled LDA model\cite{kotov2015interpretable} and represent each word in a vector of 2 topics, where the number of topics is experimentally determined by the model performance.
+In this paper, we exploit a topic model named Labeled LDA\cite{ramage2009labeled} model, where topics are corresponds to labels.
+The generative process of Labeled LDA drawing the multinomial topic distributions over vocabulary for each topic.
+In our experiment, we represent each word in a vector of 2 topics, where the number of topics is experimentally determined by the model performance\cite{kotov2015interpretable}.
 Punctuation mark containing one of the symbols \{`.', `,', `!', `?', `:', `;', `-'\} is also employed as feature.
@@
 Table~\ref{tab:result_not_boundary} summarizes the performance of NB, SVM, KNN, and RNN models for detecting ``not boundary'' class in e-coaching text.
-We observed that performance is remarkably high in ``not boundary'' class compared to boundary detection.
+We observed that performance is remarkably high in ``not boundary'' class compared to boundary detection which is expected because 96.71\% instances are from ``not boundary'' class.
 Similar to boundary detection, KNN consistently outperforms over all other methods but obtains 22.40\%, 50.07\%, and 36.26\% higher precision, recall, and F1-measure for contextual features; and 20.85\%, 33.96\%, and 27.47\% higher precision, recall, and F1-measure for combined features compared to boundary class.
@@
 \section*{Discussion}
-This study is the first large-scale efforts to evaluate the segmentation of e-coaching text.
+This study is the first efforts to evaluate the automatic segmentation of e-coaching text.
 Experimental results indicate that KNN is the best model among all machine learning methods considered for this study.
 KNN achieved 0.986 F1-measure in overall, 0.779 and 0.993 F1-measures for detecting ``boundary'' and ``not boundary'', respectively.
-The robust performance of KNN provides the evidence that machine learning models are capable to learn information from the email exchange.
+The robust performance of KNN provides the evidence that machine learning models are capable to learn information from clinical exchange.
 Although the domain of this study was intentionally quite small, we believe that our study is not limited to the e-coaching domain, and it can be successfully applied to other domain as well.
@@
 Finally, RNN demonstrates lowest AUC values among all machine learning models, achieved AUC values 0.704 and 0.740 for LSTM and GRU, respectively.
+The conclusion drawn from the ROC curves, also confirmed the robustness and superiority of KNN model for the segmentation of clinical exchange.
 
 We observed worst results of RNN, in particular, LSTM and GRU for the text segmentation.
-We believe that RNN performed poorly because it has a large set of weights which required longer sequence with a large set of data for both classes.
+We believe that RNN performed poorly because it has a large set of weights which required longer sequence with a large set of data for all labels.
 In this study, maximum length of the input sequence was 3 and we utilized 3,138 examples of boundary case which failed to achieve good results.
@@
 Punctuation mark plays an important role in segmentation boundary detection, and large numbers of errors were encountered by the false positive of boundary identification.
-For example, ``xxxxx xxxxxx'' was segmented after the second sentence where a punctuation mark was encountered.
+For example, a text block ``A1 A2 A3.
+B1 B2 B3.
+C1 C2 C3 C4.'' can be segmented at position A3, B3, and C4 where a punctuation mark was encountered.
 Similarly, additional information is the common reason for the classified original segment into multiple segments.
-For instance, ``xxxx xxxx xxxx'' was segmented after second sentence where third sentence supports the MI code confirmed by first two sentences.
+For instance, the above text block can be segmented at position B3 and C4 because third sentence (C) only supports the MI code confirmed by first two sentences (A and B).
 
 Our proposed approach is novel for the segmentation of e-coaching text because previous studies mainly focus on the segmentation of text into sections, headers, and sentences in other medical domain.
-However, this study segmented email exchange into groups of MI behaviors.
-This work will significantly reduce the amount of resource and time required to segment email text manually.
-Furthermore, this paper can help to annotate each segment automatically by building a new classifier, which will accelerate the pace of finding best communication strategies to develop an effective MI intervention for healthy eating.
+However, this study segmented clinical exchange into groups of MI behaviors which will significantly reduce the amount of resource and time required to segment clinical exchange manually.
+Furthermore, these segmentation models could be integrated with auto coding classifiers to build a software package of automatic coding procedure of clinical exchange.
 
+The limitation of this study is that e-coaching text is collected from a single medical institute; formatting, style, and email segment can be different in other settings.
+Therefore, there is a need to replicate the experiments with different data sets.
 As our future work, we plan to evaluate our approach on other datasets involves in discourse analysis.
-We also plan to use a combination of machine learning and natural language approaches to improve model performance.
-For example, part-of-speech tagging and distance from the boundary of the sentence might significantly enhance the classification performance.
-The limitation of this study is that e-coaching text is collected from a single medical institute; formatting, style, and email segment can be different in other settings.
+We also plan to use more relevant features to improve model performance.
+For example, part-of-speech tagging\cite{hasan2016feedback} and distance from the begining and end of the current sentence, might significantly enhance the classification performance.
 
@@
 Segmentation of e-coaching text is an integral part of developing an automated e-coaching intervention.
-Although several studies have done for the segmentation of clinical text into sections and sentences, none of them are used for the segmentation of text into groups of MI behaviors in the setting of discourse analysis with email under the principle of motivational interviews.
+Although several studies have done in clinical interventions, they are limited by the qualitative coding of clinical interactions.
+In addition, previous studies in medical domain, mainly segmented clinical text into sections and sentences, none of them are used for the segmentation of text into groups of MI behaviors in the setting of discourse analysis with email under the principle of motivational interviews.
 In this paper, we compared the performance of machine learning models for the task of segmentation of e-coaching text.
@@
 Our proposed methods can help to identify individual text segments, which can be annotated directly with a classification model.
-This approach will also help for developing fully automated e-coaching and accelerate the pace of identifying effective communication strategies linked to healthy eating.
+This approach will also help for developing fully automated e-coaching and accelerate the pace of identifying effective communication strategies.
 
@@
 This study was supported by a grant from the National Institutes of Health, NIDDK R21DK108071, Carcone and Kotov, MPIs.
-We would like to thank the student assistants in the Department of Family Medicine and Public Health Sciences at Wayne State University School of Medicine for their help in developing the training dataset by manually segmented the e-coaching text.
+We would like to thank research staff and student assistants in the Department of Family Medicine and Public Health Sciences at Wayne State University School of Medicine for their help in preparing the training dataset.
 

### 2018-03-08 rnn/gru_sequence.py

@@ -1,0 +1,45 @@
+from random import random
+from numpy import array
+from numpy import cumsum
+from keras.models import Sequential
+from keras.layers import LSTM
+from keras.layers import Dense
+from keras.layers import TimeDistributed
+
+
+# create a sequence classification instance
+def get_sequence(n_timesteps):
+    # create a sequence of random numbers in [0,1]
+    X = array([random() for _ in range(n_timesteps)])
+    # calculate cut-off value to change class values
+    limit = n_timesteps / 4.0
+    # determine the class outcome for each item in cumulative sequence
+    y = array([0 if x < limit else 1 for x in cumsum(X)])
+    # reshape input and output data to be suitable for LSTMs
+    X = X.reshape(1, n_timesteps, 1)
+    y = y.reshape(1, n_timesteps, 1)
+    return X, y
+
+
+# define problem properties
+n_timesteps = 10
+# define LSTM
+model = Sequential()
+model.add(LSTM(20, input_shape=(n_timesteps, 1), return_sequences=True))
+model.add(TimeDistributed(Dense(1, activation='sigmoid')))
+model.compile(loss='binary_crossentropy', optimizer='adam', metrics=['acc'])
+# train LSTM
+for epoch in range(1000):
+    # generate new random sequence
+    X, y = get_sequence(n_timesteps)
+    # fit model for one epoch on this sequence
+    model.fit(X, y, epochs=1, batch_size=1, verbose=2)
+# evaluate LSTM
+X, y = get_sequence(n_timesteps)
+
+print(X)
+print(y)
+
+yhat = model.predict_classes(X, verbose=0)
+for i in range(n_timesteps):
+    print('Expected:', y[0, i], 'Predicted', yhat[0, i])

### 2018-03-08 rnn/seg-gru-model-stepwise.py

@@ -1,0 +1,117 @@
+# RNN with variable length input sequences to one word output
+import numpy
+
+from random import random
+from numpy import array
+from numpy import cumsum
+from keras.models import Sequential
+from keras.layers import LSTM
+from keras.layers import GRU
+from keras.layers import Dense
+from keras.layers import TimeDistributed
+from keras.callbacks import EarlyStopping, ModelCheckpoint, Callback
+
+import utility_stepwise
+
+from vocabulary import Vocabulary
+
+# fix random seed for reproducibility
+numpy.random.seed(42)
+
+# system configuration
+max_len = 50
+folds = 5
+lstm_out = 25
+batch_size = 1
+
+
+class ResetStatesCallback(Callback):
+    def __init__(self):
+        self.counter = 0
+
+    def on_batch_begin(self, batch, logs={}):
+        if self.counter % max_len == 0:
+            self.model.reset_states()
+            # print("Reset States..............")
+        self.counter += 1
+
+
+# create e-coaching dictionary
+vocabulary_obj = Vocabulary()
+vocabulary_obj.create()
+
+# create mapping of characters to integers (0-25) and the reverse
+char_to_int = dict((c, i+1) for i, c in enumerate(vocabulary_obj.vocab))
+int_to_char = dict((i+1, c) for i, c in enumerate(vocabulary_obj.vocab))
+
+# prepare the data set of input to output pairs encoded as integers
+data = utility_stepwise.get_ecoaching_data()
+folds_data = utility_stepwise.split_data(data, folds)
+results = []
+
+for fold_num in range(folds):
+
+    train_data, test_data = [], []
+    for i in range(folds):
+        if i == fold_num:
+            test_data = folds_data[i]
+        else:
+            train_data = train_data + folds_data[i]
+
+    # get encoded training data
+    X, y = utility_stepwise.get_encoded_sequence(train_data, vocabulary_obj, max_len)
+
+    # print total number of instances
+    print("# of instances: ", len(y))
+
+    # create and fit the model
+    model = Sequential()
+    model.add(LSTM(lstm_out, input_shape=(max_len, 1), return_sequences=True))
+    model.add(TimeDistributed(Dense(1, activation='sigmoid')))
+    model.compile(loss='binary_crossentropy', optimizer='adam', metrics=['accuracy'])
[…]

### 2018-03-08 rnn/utility_stepwise.py

@@ -1,0 +1,128 @@
+# include required packages
+import numpy
+import re
+import sys
+
+from keras.utils import np_utils
+from keras.preprocessing.sequence import pad_sequences
+
+
+# load e-coaching data from datafile
+def get_ecoaching_data(file_name='e-coaching_data_stem.txt'):
+    data = []
+    with open(file_name, "r") as file_stream:
+        for line in file_stream:
+            current_line = line.split(",")
+            if len(current_line) > 1:
+                data.append(clean_text(current_line[0]).strip())
+    return data
+
+
+# split data into 5 folds
+def split_data(data, folds=5):
+    fold_size = len(data)/folds
+    folds_data = []
+    for i in range(folds):
+        start, end = i*fold_size, i*fold_size + fold_size
+        folds_data.append(data[start: end])
+    return folds_data
+
+
+# encode text data into numeric values
+def get_encoded_sequence(data, vocabulary_obj, max_len):
+    x, y = [], []
+
+    for text in data:
+        sample_x = vocabulary_obj.vectorize(text)
+        sample_y = [0]*(len(sample_x)-1)
+        sample_y.append(1)
+
+        x.append(sample_x)
+        y.append(sample_y)
+
+    data_x = numpy.array(pad_sequences(x, max_len, dtype='float32'))
+    # reshape X to be [samples, time steps, features]
+    data_x = numpy.reshape(data_x, (data_x.shape[0], 1, max_len, 1))
+    # normalize
+    data_x = data_x / vocabulary_obj.size()
+    data_y = numpy.array(pad_sequences(y, max_len))
+    data_y = numpy.reshape(data_y, (data_y.shape[0], 1, max_len, 1))
+    return data_x, data_y
+
+
+# clean text, removed weired chars
+def clean_text(text):
+    new_text = re.sub(r'[^\x00-\x7f]', r'', text)
+    new_text = new_text.replace("\n", "").strip()
+    new_text = re.sub(r'[?|`|>|<|$|.|!|;|:|=|&|\'|\"|\-|/|)|(|\]|[|]', r'', new_text)
+    return new_text
+
+
+# evaluate model with F1, Precision and Recall
+def get_macro_average_performance(actual, predicted):
+    f = open("/home/mehedi/Dropbox/results_sequence_GRU.txt", "a")
+    f.write("fold number \n")
+    f.write("actual: ")
+    f.writelines(["%s," % item for item in actual])
+    f.write("\npredicted: ")
[…]

### 2018-03-08 rnn/var_seq_length/utility_var_sequence.py

@@ -1,0 +1,122 @@
+# include required packages
+import numpy
+import re
+import sys
+
+from keras.utils import np_utils
+from keras.preprocessing.sequence import pad_sequences
+
+
+# load e-coaching data from datafile
+def get_ecoaching_data(file_name='e-coaching_data_stem.txt'):
+    data = []
+    with open(file_name, "r") as file_stream:
+        for line in file_stream:
+            current_line = line.split(",")
+            if len(current_line) > 1:
+                data.append(clean_text(current_line[0]).strip())
+    return data
+
+
+# split data into 5 folds
+def split_data(data, folds=5):
+    fold_size = len(data)/folds
+    folds_data = []
+    for i in range(folds):
+        start, end = i*fold_size, i*fold_size + fold_size
+        folds_data.append(data[start: end])
+    return folds_data
+
+
+# encode text data into numeric values
+def get_encoded_sequence(data, vocabulary_obj):
+    x, y = [], []
+
+    for text in data:
+        sample_x = vocabulary_obj.vectorize(text)
+        sample_y = [0]*(len(sample_x)-1)
+        sample_y.append(1)
+
+        sample_x = [a / float(vocabulary_obj.size()) for a in sample_x]
+        x.append(sample_x)
+        y.append(sample_y)
+
+    return x, y
+
+
+# clean text, removed weired chars
+def clean_text(text):
+    new_text = re.sub(r'[^\x00-\x7f]', r'', text)
+    new_text = new_text.replace("\n", "").strip()
+    new_text = re.sub(r'[?|`|>|<|$|.|!|;|:|=|&|\'|\"|\-|/|)|(|\]|[|]', r'', new_text)
+    return new_text
+
+
+# evaluate model with F1, Precision and Recall
+def get_macro_average_performance(actual, predicted):
+    f = open("/home/mehedi/Dropbox/results_sequence_GRU.txt", "a")
+    f.write("fold number \n")
+    f.write("actual: ")
+    f.writelines(["%s," % item for item in actual])
+    f.write("\npredicted: ")
+    f.writelines(["%s," % item for item in predicted])
+    f.write("\n")
+
+    precision = 0.0
+    recall = 0.0
+    f_measure = 0.0
+    accuracy = 0.0
+
+    labels = [1, 0, 0, 1]
+
+    for k in [0, 2]:
+
+        tp = 0
+        fp = 0
+        tn = 0
+        fn = 0
+
+        for i in range(0, len(actual)):
[…]

### 2018-03-08 rnn/var_seq_length/var-seq-model.py

@@ -1,0 +1,115 @@
+# RNN with variable length input sequences to one word output
+import numpy
+
+from random import random
+from numpy import array
+from numpy import cumsum
+from keras.models import Sequential
+from keras.layers import LSTM
+from keras.layers import GRU
+from keras.layers import Dense
+from keras.layers import TimeDistributed
+from keras.callbacks import EarlyStopping, ModelCheckpoint, Callback
+
+import utility_var_sequence
+
+from vocabulary import Vocabulary
+
+# fix random seed for reproducibility
+numpy.random.seed(42)
+
+# system configuration
+folds = 5
+lstm_out = 10
+batch_size = 1
+max_len = 50
+
+class ResetStatesCallback(Callback):
+    def __init__(self):
+        self.counter = 0
+
+    def on_batch_begin(self, batch, logs={}):
+        if self.counter % max_len == 0:
+            self.model.reset_states()
+            # print("Reset States..............")
+        self.counter += 1
+
+# create e-coaching dictionary
+vocabulary_obj = Vocabulary()
+vocabulary_obj.create()
+
+# create mapping of characters to integers (0-25) and the reverse
+char_to_int = dict((c, i+1) for i, c in enumerate(vocabulary_obj.vocab))
+int_to_char = dict((i+1, c) for i, c in enumerate(vocabulary_obj.vocab))
+
+# prepare the data set of input to output pairs encoded as integers
+data = utility_var_sequence.get_ecoaching_data()
+folds_data = utility_var_sequence.split_data(data, folds)
+results = []
+
+for fold_num in range(folds):
+
+    train_data, test_data = [], []
+    for i in range(folds):
+        if i == fold_num:
+            test_data = folds_data[i]
+        else:
+            train_data = train_data + folds_data[i]
+
+    # get encoded training data
+    X, y = utility_var_sequence.get_encoded_sequence(train_data, vocabulary_obj)
+
+    # print total number of instances
+    print("# of instances: ", len(y))
+
+    # create and fit the model
+    model = Sequential()
+    model.add(LSTM(lstm_out, input_shape=(None, 1), return_sequences=True))
+    model.add(TimeDistributed(Dense(1, activation='sigmoid')))
+    model.compile(loss='binary_crossentropy', optimizer='adam', metrics=['accuracy'])
+
+    for i in range(len(X)):
[…]

### 2018-03-08 rnn/var_seq_length/vocabulary.py

@@ -1,0 +1,59 @@
+import re
+import codecs
+import operator
+
+
+class Vocabulary:
+    def __init__(self, vocab_path='e-coaching.vocab'):
+        self.vocab = dict()
+        self.int_to_word = dict()
+        self.vocab_path = vocab_path
+
+    # build e-coaching dictionary
+    def build(self):
+        with codecs.open(self.vocab_path, 'r', 'UTF-8') as train_file:
+            words = [x.strip().rstrip('\n') for x in train_file.readlines()]
+            self.vocab = dict((c, i + 1) for i, c in enumerate(words))
+            self.int_to_word = dict((i+1, c) for i, c in enumerate(words))
+        return self
+
+    # read all the data from file and create vocabulary, also find # classes
+    def create(self, data_file='e-coaching_data_stem.txt'):
+        word_dict = {}
+
+        with open(data_file, "r") as file_stream:
+            for line in file_stream:
+                current_line = line.strip().split(",")
+                if len(current_line) < 2:
+                    continue
+
+                words = re.split('(\W+)', self.clean_text(current_line[0]).strip())
+                for word in words:
+                    if len(word.strip()) > 0:
+                        word_dict[word] = 1
+
+        sorted_map = sorted(word_dict.items(), key=operator.itemgetter(0))
+        file_pointer = open(self.vocab_path, "w")
+        for key, val in sorted_map:
+            file_pointer.write(key + "\n")
+        file_pointer.close()
+
+        self.build()
+        return self
+
+    def size(self):
+        return len(self.vocab)
+
+    def tokenize(self, text):
+        return [x.strip() for x in re.split('(\W+)', text) if x.strip()]
+
+    def vectorize(self, text):
+        text = text.lower()
+        words = filter(lambda x: x in self.vocab, self.tokenize(text))
+        return [self.vocab[w] for w in words]
+
+    def clean_text(self, text):
+        new_text = re.sub(r'[^\x00-\x7f]', r'', text)
+        new_text = new_text.replace("\n", "").strip()
+        new_text = re.sub(r'[?|`|>|<|$|.|!|;|:|=|&|\'|\"|\-|/|)|(|\]|[|]', r'', new_text)
+        return new_text

### 2018-03-08 rnn/var_seq_length/var-seq-model.py
update LSTM

@@ -20,9 +20,10 @@ numpy.random.seed(42)
 
 # system configuration
 folds = 5
-lstm_out = 10
+lstm_out = 25
 batch_size = 1
 max_len = 50
+
 
 class ResetStatesCallback(Callback):
     def __init__(self):
@@ -31,7 +32,7 @@ class ResetStatesCallback(Callback):
     def on_batch_begin(self, batch, logs={}):
         if self.counter % max_len == 0:
             self.model.reset_states()
-            # print("Reset States..............")
+            # print("Reset States..............", max_len)
         self.counter += 1
 
 # create e-coaching dictionary
@@ -74,7 +75,7 @@ for fold_num in range(folds):
         max_len = len(X[i])
         if i % 100 == 0:
             print("Records processed: ", i)
-        model.fit(x_data, y_data, callbacks=[ResetStatesCallback()], shuffle=False, batch_size=batch_size, verbose=0)
+        model.fit(x_data, y_data, callbacks=[ResetStatesCallback()], epochs=max_len, shuffle=False, batch_size=batch_size, verbose=2)
 
     # get encoded test data
     test_x, test_y = utility_var_sequence.get_encoded_sequence(test_data, vocabulary_obj)

### 2018-03-08 paper-amia-2018/manuscript.tex
update introduction

@@
 Rapidly developing computational technologies, specifically, machine learning models, offer a unique opportunity to accelerate this process.
-A few studies were done for the automatic annotation of clinical interactions in the setting of MI intervention.
-Lacson et al.\cite{lacson2005automatic} applied AdaBoost\cite{freund1999short} classifier for annotating the interactions in hemodialysis phone dialog into 3 categories.
+Machine learning-based classification methods have been successfully applied to a variety of analytical tasks on textual data including classification\cite{nigam2000text}, sentiment analysis\cite{wang2012baselines}, and digital forensics\cite{de2002language}.
+A few studies have been done for the automatic annotation of clinical interactions in the setting of MI intervention.
+Lacson et al.\cite{lacson2005automatic} applied AdaBoost\cite{freund1999short} classifier for annotating the interactions in hemodialysis phone dialog as Clinical, Technical,
+Backchannel, and Miscellaneous categories.
 Our research group has recently applied several machine learning-based models to MI sessions \cite{hasan2016study,kotov2015interpretable}.
@@
 Experimental data utilized in the above MI studies were prepared by transcribing audio conversation which involves a counselor, a patient, and some cases a caregiver\cite{hasan2016study,kotov2015interpretable}.
-However, this study utilized e-coaching data which have different structure and context compared to clinical data.
+However, this study utilized e-coaching data which have different structure and context compared to traditional clinical data.
 E-coaches use email to deliver motivation-enhancing coaching to encourage healthy eating, grounded in the principles of motivational interviewing.
 E-coaching data is comprised of email responses which are unsegmented, unlike more traditional dyadic clinical communication where segmentation occurs naturally due to its conversation nature.
-Discourse analysis on e-coaching text can help to understand the communication process in e-coaching text by revealing the socio-psychological characteristics of a patient\cite{siegfried1995therapeutic}.
+Discourse analysis on e-coaching text can help to understand the communication process in e-coaching text by revealing the socio-psychological characteristics of a patient\cite{siegfried1995therapeutic,kalchbrenner2013recurrent,pierre2016neural}.
 By utilizing e-coaching data and leveraging innovative machine learning models, the ultimate goal of this research study is to fully automate the communication coding process.
@@
 Automatic segmentation of e-coaching intervention sessions is a challenging task due to the 2 important reasons.
-First, the email is an unstructured text that contains informal email exchange in non-traditional formats.
+First, the email is an unstructured text that contains informal email exchange as a non-traditional format.
 Second, a text segment does not necessarily belong to the entire sentence or collection of sentences.
@@
 Figure~\ref{fig:text-segment} illustrates the segmentation of an e-coaching email exchange where the first sentence segmented into 2 different MI behaviors.
-On the other hand, fourth and fifth segments contain only one and multiple sentences, respectively.
+On the other hand, fourth and fifth segments contain 1 and 3 sentences, respectively.
 
@@
 
-Recently, an online clinical intervention called MENU GenY\cite{alexander2017motivations} (Making Effective Nutrition Choices for Generation Y) was tested.
+Recently, an online clinical intervention called MENU GenY\cite{alexander2017motivations} (Making Effective Nutrition Choices for Generation Y) was proposed to test its efficacy.
 MENU GenY is a technology-based public health intervention to encourage increased fruit and vegetable intake among young adult age 21-30, utilized personalized e-coaching data.
-The main focus of MENU GenY is to develop a better coding dictionary among GenY to improve eating habits.
+The goal of MENU GenY was to develop a better coding dictionary among GenY to improve eating habits.
 However, segmentation of clinical conversation in the context of electronically delivered interventions, in particular, segmentation of clinical interaction text into groups of MI behaviors, is still ignored while relying on manual hand-coded approach.
@@
 Topics are considered as features since topic models are very effective\cite{kotov2015interpretable,hashimoto2016topic,lu2016modeling} to represent text documents.
-In this paper, we exploit a topic model named Labeled LDA\cite{ramage2009labeled} model, where topics are corresponds to labels.
+In this paper, we exploit a topic model named Labeled LDA\cite{ramage2009labeled} model, where topics correspond to labels.
 The generative process of Labeled LDA drawing the multinomial topic distributions over vocabulary for each topic.
@@
 On the other hand, NB exhibits lowest precision value 0.594, provides better recall and F1-measure than RNN model.
-When textual features are used in combination with topic and punctuation mark, recall increases by 0.6\%, 2.5\%, and 11.92\%; and F1-measure increases by 0\%, 3.39\%, and 7\% for NB, SVM, and KNN models, respectively.
+When textual features are used in combination with topic and punctuation mark features, recall increases by 0.6\%, 2.5\%, and 11.92\%; and F1-measure increases by 0\%, 3.39\%, and 7\% for NB, SVM, and KNN models, respectively.
 Nevertheless, precision increases by 4.31\% and 1.49\% for SVM and KNN methods while decreases by 0.7\% in NB.
@@
 Finally, RNN demonstrates lowest AUC values among all machine learning models, achieved AUC values 0.704 and 0.740 for LSTM and GRU, respectively.
-The conclusion drawn from the ROC curves, also confirmed the robustness and superiority of KNN model for the segmentation of clinical exchange.
+The conclusion drawn from the ROC curves also confirmed the robustness and superiority of KNN model for the segmentation of clinical exchange.
 
@@
 We believe that RNN performed poorly because it has a large set of weights which required longer sequence with a large set of data for all labels.
-In this study, maximum length of the input sequence was 3 and we utilized 3,138 examples of boundary case which failed to achieve good results.
+In this study, a maximum length of the input sequence was 3 and we utilized 3,138 examples of boundary case which failed to achieve good results.
 While the performance of RNN is poor, GRU performs better than LSTM which was observed in other previous study\cite{chung2014empirical}.
@@
 We also plan to use more relevant features to improve model performance.
-For example, part-of-speech tagging\cite{hasan2016feedback} and distance from the begining and end of the current sentence, might significantly enhance the classification performance.
+For example, part-of-speech tagging\cite{hasan2016feedback} and distance from the beginning and end of the current sentence might significantly enhance the classification performance.
 
@@
 Although several studies have done in clinical interventions, they are limited by the qualitative coding of clinical interactions.
-In addition, previous studies in medical domain, mainly segmented clinical text into sections and sentences, none of them are used for the segmentation of text into groups of MI behaviors in the setting of discourse analysis with email under the principle of motivational interviews.
+In addition, previous studies in the medical domain mainly segmented clinical text into sections and sentences, none of them are used for the segmentation of text into groups of MI behaviors in the setting of discourse analysis with email under the principle of motivational interviews.
 In this paper, we compared the performance of machine learning models for the task of segmentation of e-coaching text.

### 2018-03-08 paper-amia-2018/manuscript.tex
update RNN model results

@@
 The activation $h_t$ of the GRU at time $t$ is a linear combination of previous activation $h_{t-1}$ and the candidate activation $\tilde h_t$, which is represented by Eq.~(\ref{eq:lastgru}) and~(\ref{eq:candidategru}).
-We build our RNN model with one hidden layer, output layer, and input layer which get one hot encoding of word vector as input.
+We build our RNN model with one hidden layer, output layer, and input layer.
+One segmented text was taken as input sequence which get one hot encoding of word vector for the model input.
+A binary label was assigned as output sequence where ``boundary'' and `not boundary'' words were associated with label 1 and 0, respectively.
 Since one-hot vector is given in the input layer, results are reported with textual and punctuation mark features only.
-We experimentally determined that the best performance is achieved when the number of hidden units = 32, batch size = 8, optimizer = adam, as well as 600 epochs is used.
+We experimentally determined that the best performance is achieved when the number of hidden units = 25, batch size = 1, and optimizer = adam.
 
@@
  KNN & \textbf{0.808} & 0.663 & \textbf{0.728} & \textbf{0.820} & \textbf{0.742} & \textbf{0.779}\\ \hline
- LSTM & -- & -- & -- & 0.619 & 0.416 & 0.497  \\ \hline
- GRU & -- & -- & -- & 0.642 & 0.490 & 0.554 \\ \hline 
+ LSTM & -- & -- & -- & 0.800 & 0.646 & 0.714  \\ \hline
+ GRU & -- & -- & -- & 0.769 & 0.715 & 0.741 \\ \hline 
   \end{tabular}
@@
  SVM & \textbf{0.989} & 0.992 & 0.991 & 0.990 & 0.993 & 0.991\\ \hline
- KNN & \textbf{0.989} & \textbf{0.995} & \textbf{0.992} & \textbf{0.991} & \textbf{0.994} & \textbf{0.993}\\ \hline
- LSTM & -- & -- & -- & 0.981 & 0.991 & 0.986 \\ \hline
- GRU & -- & -- & -- & 0.983 & 0.991 & 0.987 \\ \hline 
+ KNN & \textbf{0.989} & \textbf{0.995} & \textbf{0.992} & 0.991 & \textbf{0.994} & 0.993\\ \hline
+ LSTM & -- & -- & -- & 0.993 & \textbf{0.994} & \textbf{0.994} \\ \hline
+ GRU & -- & -- & -- & \textbf{0.994} & \textbf{0.994} & \textbf{0.994} \\ \hline 
   \end{tabular}
@@
  SVM & 0.981 & 0.982 & 0.981 & 0.983 & 0.983 & 0.983\\ \hline
- KNN & \textbf{0.983} & \textbf{0.984} & \textbf{0.983} & \textbf{0.986} & \textbf{0.986} & \textbf{0.986}\\ \hline
- LSTM & -- & -- & -- & 0.969 & 0.972 & 0.970 \\ \hline
- GRU & -- & -- & -- & 0.972 & 0.974 & 0.973 \\ \hline 
+ KNN & \textbf{0.983} & \textbf{0.984} & 0.983 & \textbf{0.986} & \textbf{0.986} & \textbf{0.986}\\ \hline
+ LSTM & -- & -- & -- & \textbf{0.986} & 0.983 & 0.984 \\ \hline
+ GRU & -- & -- & -- & \textbf{0.986} & 0.985 & \textbf{0.986} \\ \hline 
   \end{tabular}
@@
 AUC values are also outlined due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets\cite{hu2015kernelized}, which are demonstrated by the ROC curves in Figure~\ref{fig:roc-curves}.
-NB shows the highest AUC values, achieved 0.978 for both cases while provides lowest classification results except RNN.
+NB shows the highest AUC values, achieved 0.978 for both cases while provides lowest classification results.
 On the other hand, KNN and SVM exhibit 0.972 and 0.835 AUCs when only textual features are used; and 0.959 and 0.844 AUCs when a combination of textual, topic and punctuation mark features are used.
-Finally, RNN demonstrates lowest AUC values among all machine learning models, achieved AUC values 0.704 and 0.740 for LSTM and GRU, respectively.
+Finally, LSTM demonstrates lowest AUC values among all machine learning models, achieved 0.82 AUC while GRU achieved 0.854 AUC.
 The conclusion drawn from the ROC curves also confirmed the robustness and superiority of KNN model for the segmentation of clinical exchange.
 
-We observed worst results of RNN, in particular, LSTM and GRU for the text segmentation.
-We believe that RNN performed poorly because it has a large set of weights which required longer sequence with a large set of data for all labels.
-In this study, a maximum length of the input sequence was 3 and we utilized 3,138 examples of boundary case which failed to achieve good results.
-While the performance of RNN is poor, GRU performs better than LSTM which was observed in other previous study\cite{chung2014empirical}.
+We observed second highest performance of RNN, in particular, LSTM and GRU for the text segmentation.
+We believe that RNN will perform better if a large set of data is utilized.
+In this study, we employed 3,138 examples of boundary case which failed to achieve best results.
+We also observed that GRU performs better than LSTM which was observed in other previous study\cite{chung2014empirical}.
 
-Punctuation mark plays an important role in segmentation boundary detection, and large numbers of errors were encountered by the false positive of boundary identification.
+Although punctuation mark plays an important role in segmentation boundary detection, and large numbers of errors were encountered by the false positive of boundary identification.
 For example, a text block ``A1 A2 A3.
 B1 B2 B3.
-C1 C2 C3 C4.'' can be segmented at position A3, B3, and C4 where a punctuation mark was encountered.
+C1 C2 C3 C4.'' can be incorrectly segmented at position A3, B3, and C4 where a punctuation mark was encountered.
 Similarly, additional information is the common reason for the classified original segment into multiple segments.
-For instance, the above text block can be segmented at position B3 and C4 because third sentence (C) only supports the MI code confirmed by first two sentences (A and B).
+For instance, the above text block can also be incorrectly segmented at position B3 and C4 because third sentence (C) only supports the MI code confirmed by first two sentences (A and B).
 

### 2018-03-08 paper-amia-2018/manuscript.tex
update results section

@@
  LSTM & -- & -- & -- & 0.800 & 0.646 & 0.714  \\ \hline
- GRU & -- & -- & -- & 0.769 & 0.715 & 0.741 \\ \hline 
+ GRU & -- & -- & -- & 0.800 & 0.715 & 0.741 \\ \hline 
   \end{tabular}
@@
 As follows from Table~\ref{tab:result_boundary}, KNN performs best among all machine learning models in terms of precision and F1-measure, achieved 0.808 precision with 0.728 F1-measure when contextual features are used, and 0.820 precision with 0.779 F1-measure when a combination of contextual, topic, and punctuation mark features are used.
-However, RNN demonstrates the lowest performance among all models in terms of recall and F1-measure while GRU shows 3.72\%, 17.79\% and 11.47\% higher precision, recall, and F1-measure than LSTM.
-In this study, SVM appears as the second highest model in terms of precision and F1-measure, obtains highest 0.679 recall when only textual features are used.
-On the other hand, NB exhibits lowest precision value 0.594, provides better recall and F1-measure than RNN model.
+However, NB demonstrates the lowest performance among all models in terms of all performance metrics.
+In this study, GRU appears as the second highest model, obtains 10.68\% higher recall with 3.78\%  higher F1-measure than LSTM.
+On the other hand, SVM exhibits highest recall 0.679 when only textual features are used.
 When textual features are used in combination with topic and punctuation mark features, recall increases by 0.6\%, 2.5\%, and 11.92\%; and F1-measure increases by 0\%, 3.39\%, and 7\% for NB, SVM, and KNN models, respectively.
@@
 We observed that performance is remarkably high in ``not boundary'' class compared to boundary detection which is expected because 96.71\% instances are from ``not boundary'' class.
-Similar to boundary detection, KNN consistently outperforms over all other methods but obtains 22.40\%, 50.07\%, and 36.26\% higher precision, recall, and F1-measure for contextual features; and 20.85\%, 33.96\%, and 27.47\% higher precision, recall, and F1-measure for combined features compared to boundary class.
-However, RNN demonstrates the lowest performance and exhibits 0.981 and 0.983 precision with 0.986 and 0.987 F1-measure for LSTM and GRU, respectively.
+KNN achieves 22.40\%, 50.07\%, and 36.26\% higher precision, recall, and F1-measure for contextual features; and 20.85\%, 33.96\%, and 27.47\% higher precision, recall, and F1-measure for combined features compared to boundary class.
+In contrast to boundary detection, RNN demonstrates the highest performance among all models.
+LSTM obtains 0.993 precision with 0.994 recall and F1-measure while LSTM exhibits 0.994 for all performance metrics.
 Impact of additional features is also consistent with ``not boundary'' classification.
@@
 Table~\ref{tab:result_weighted_avg} outlines the weighted average results of the experiment on the models for the segmentation of e-coaching text by classifying them into ``boundary'' and ``not boundary'' classes.
-Overall, KNN obtains the best performance with all metrics and RNN denotes the lowest performance among all methods.
-NB and SVM demonstrate moderate performance, obtain precision 0.975 and 0.981; recall 0.974 and 0.982; and F1-measure 0.975 and 0.983 when textual features are used.
+Overall, KNN obtains the best performance with all metrics and NB denotes the lowest performance among all methods.
+GRU demonstrates the same result as KNN for precision and F1-measure except for recall.
+SVM shows moderate performance, obtains 0.981, 0.982, and 0.983 for precision, recall, and F1-measure, respectively when textual features are used.
 Influence of the additional features is also consistent as above, precision increases by 0.1\%, 0.2\%, and 0.3\%; recall increases by 0\%, 0.1\%, and 0.2\%; and F1-measure increases by 0\%, 0.2\%, and 0.3\% for NB, SVM, and KNN methods, respectively, when combined features are used.\\
@@
 
-We observed second highest performance of RNN, in particular, LSTM and GRU for the text segmentation.
+We observed the second highest performance of RNN, in particular, LSTM and GRU for the text segmentation.
 We believe that RNN will perform better if a large set of data is utilized.

### 2018-03-08 paper-amia-2018/manuscript.tex
update results section

@@
 A binary label was assigned as output sequence where ``boundary'' and `not boundary'' words were associated with label 1 and 0, respectively.
+We reset our model state after feeding each input sequence.
 Since one-hot vector is given in the input layer, results are reported with textual and punctuation mark features only.
@@
 On the other hand, KNN and SVM exhibit 0.972 and 0.835 AUCs when only textual features are used; and 0.959 and 0.844 AUCs when a combination of textual, topic and punctuation mark features are used.
-Finally, LSTM demonstrates lowest AUC values among all machine learning models, achieved 0.82 AUC while GRU achieved 0.854 AUC.
+Finally, LSTM demonstrates lowest AUC values among all machine learning models, achieved 0.82 AUC while GRU achieves 4.15\% higher AUC than LSTM.
 The conclusion drawn from the ROC curves also confirmed the robustness and superiority of KNN model for the segmentation of clinical exchange.

### 2018-03-08 rnn/var_seq_length/var-seq-model.py

@@ -75,7 +75,7 @@ for fold_num in range(folds):
         max_len = len(X[i])
         if i % 100 == 0:
             print("Records processed: ", i)
-        model.fit(x_data, y_data, callbacks=[ResetStatesCallback()], epochs=max_len, shuffle=False, batch_size=batch_size, verbose=2)
+        model.fit(x_data, y_data, callbacks=[ResetStatesCallback()], epochs=max_len, shuffle=False, batch_size=batch_size, verbose=0)
 
     # get encoded test data
     test_x, test_y = utility_var_sequence.get_encoded_sequence(test_data, vocabulary_obj)

### 2018-03-09 paper-amia-2018/manuscript.tex
Intro updated by Alex

@@
 
-\title{Machine Learning Models for the Segmentation of Clinical Exchange}
+\title{Machine Learning Methods for Discourse Segmentation of E-Mail Based Clinical Communication}
 
 \author{Mehedi Hasan, BS$^{1}$\footnote[1]{Authors provided equal contribution.
-\label{footnote1}}, Alexander Kotov, PhD$^{1}$\textsuperscript{\ref{footnote1}}, Ming Dong, PhD$^{1}$, Sylvie Naar, PhD$^{2}$, Gwen L.
+\label{footnote1}}, Alexander Kotov, PhD$^{1}$\textsuperscript{\ref{footnote1}}, Sylvie Naar, PhD$^{2}$, Gwen L.
 Alexander, PhD$^{3}$, April Idalski Carcone, PhD$^{4}$}
@@
 $^1$Department of Computer Science, Wayne State University, Detroit, Michigan \\  
-$^2$Director, Center for Translational Behavioral Research, Department of Behavioral Sciences and Social Medicine, Florida State University, Tallahassee, Florida\\
+$^2$Center for Translational Behavioral Research, Department of Behavioral Sciences and Social Medicine, Florida State University, Tallahassee, Florida\\
 $^3$Department of Public Health Sciences, Henry Ford Health System, Detroit, Michigan\\
@@
 \textit{Communication science to understand clinical process like Motivational Interviews (MI) is limited by traditional qualitative coding methods.
-Qualitative coding of communication data is very resource-intensive and time-consuming process, requires auto coding.
+Qualitative coding of communication data is very resource-intensive and time-consuming process, which requires auto coding.
 This study utilized e-coaching data where email is used to deliver motivation-enhancing coaching to encourage healthy eating.
@@
 \section*{Introduction}
-Communication science to understand clinical process like Motivational Interviews (MI) is limited by traditional qualitative coding methods.
-Motivational Interviewing (MI) is an evidence-based communication technique to increase intrinsic motivation and self-efficacy for behavior change\cite{miller2012motivational,miller2009ten,miller2009toward}.
-Patient ``change talk'', statements of intrinsic motivation about their desire, ability, reasons, need for and commitment to behavior change, is an established mediator of health behavior change\cite{apodaca2009mechanisms}.
-Qualitative coding of communication data has been traditionally performed manually with pre-defined codes by trained coders, which is a tedious resource-intensive task, requiring an iterative process of human (subjective) interpretation of the text.
+The emergence of e-Health technologies opened up new ways to deliver a variety of behavioral interventions to any demographic group of patients in any geographical location.
+Motivational interviewing (MI), an evidence-based communication technique to increase intrinsic motivation and self-efficacy for behavior change\cite{miller2012motivational,miller2009ten,miller2009toward}, is one type of these interventions.
+MI sessions are generally aimed at eliciting ``change talk'', or statements of intrinsic motivation about patients' own desire, ability, reasons, need for and commitment to behavior change, which have been established by previous research\cite{apodaca2009mechanisms} as a reliable mediator of health behavior change.
+However, communication science approaches to understanding the efficacy of MI are inherently limited by traditional qualitative coding methods.
 
-Rapidly developing computational technologies, specifically, machine learning models, offer a unique opportunity to accelerate this process.
-Machine learning-based classification methods have been successfully applied to a variety of analytical tasks on textual data including classification\cite{nigam2000text}, sentiment analysis\cite{wang2012baselines}, and digital forensics\cite{de2002language}.
-A few studies have been done for the automatic annotation of clinical interactions in the setting of MI intervention.
-Lacson et al.\cite{lacson2005automatic} applied AdaBoost\cite{freund1999short} classifier for annotating the interactions in hemodialysis phone dialog as Clinical, Technical,
-Backchannel, and Miscellaneous categories.
-Our research group has recently applied several machine learning-based models to MI sessions \cite{hasan2016study,kotov2015interpretable}.
-A simple communication code scheme was automated to characterize patient communication and achieved accuracy comparable to human coders\cite{hasan2016study}.
+Qualitative coding of motivational interviews with pre-defined codes has been traditionally performed manually by trained annotators, which is a tedious and resource-intensive process that involves several iterations of reading, comprehension and interpretation of interview transcripts.
+Rapidly developing computational technologies, specifically, machine learning methods, offer a unique opportunity to accelerate this process.
+In particular, machine learning methods have been successfully applied to a variety of analytical tasks involving textual data, such as classification\cite{nigam2000text} and sentiment analysis\cite{wang2012baselines}.
+In our previous work, we examined the utility of machine learning methods for automated annotation \cite{hasan2016study,kotov2015interpretable} and analysis \cite{hasan2018predicting} of in-person MI sessions.
+Specifically, we demonstrated that machine learning methods can be utilized for annotation of MI transcripts according to a simple communication code scheme with the accuracy comparable to human coders\cite{hasan2016study}.
+Experimental data utilized in these studies, however, were prepared by transcribing audio conversations, which were clearly segmented into utterances by a counselor, a patient, and, in some cases, a caregiver.
 
-Experimental data utilized in the above MI studies were prepared by transcribing audio conversation which involves a counselor, a patient, and some cases a caregiver\cite{hasan2016study,kotov2015interpretable}.
-However, this study utilized e-coaching data which have different structure and context compared to traditional clinical data.
-E-coaches use email to deliver motivation-enhancing coaching to encourage healthy eating, grounded in the principles of motivational interviewing.
-E-coaching data is comprised of email responses which are unsegmented, unlike more traditional dyadic clinical communication where segmentation occurs naturally due to its conversation nature.
-Discourse analysis on e-coaching text can help to understand the communication process in e-coaching text by revealing the socio-psychological characteristics of a patient\cite{siegfried1995therapeutic,kalchbrenner2013recurrent,pierre2016neural}.
-By utilizing e-coaching data and leveraging innovative machine learning models, the ultimate goal of this research study is to fully automate the communication coding process.
+In this study, we focus on the analysis of e-Coaching sessions, MI interventions that are delivered via email and grounded in the principles of motivational interviewing.
+Specifically, the e-Coaches involved in this study used emails to communicate motivation-enhancing messages that encourage healthy eating among GenY adolescents.
+e-Coaching data is comprised of email responses, which are free-text documents, unlike more traditional dyadic clinical interviews that are naturally segmented into utterances due to their conversational nature.
 
-A significant barrier to fully automated the communication coding process is the segmentation of the text data.
-During traditional qualitative coding, coders determine where to segment text, that is, where one code ends and another begins.
-Automatic segmentation of e-coaching intervention sessions is a challenging task due to the 2 important reasons.
-First, the email is an unstructured text that contains informal email exchange as a non-traditional format.
-Second, a text segment does not necessarily belong to the entire sentence or collection of sentences.
-One sentence can be segmented into multiple MI behaviors or several sentences may represent a single MI code.
-Figure~\ref{fig:text-segment} illustrates the segmentation of an e-coaching email exchange where the first sentence segmented into 2 different MI behaviors.
-On the other hand, fourth and fifth segments contain 1 and 3 sentences, respectively.
+The unstructured nature of e-Coaching exchanges poses a unique set of challenges for their qualitative analysis.
+A significant barrier to fully automating the behavior coding process of e-Coaching emails is their segmentation into textual fragments that correspond to distinct communication behaviors.
+Automating this task is a unique and challenging problem due to the following major reasons:
+
+\begin{enumerate}
+\item Emails are unstructured text that contains informal information exchange in a non-traditional format.
+\item Discourse segments in e-Coaching emails do not necessarily correspond to sentences or collection of sentences.
+One sentence can be segmented into multiple MI behavior fragments.
+On the other hand, an MI behavior may comprise several sentences.
+\end{enumerate}
+
+Figure~\ref{fig:text-segment} illustrates a segmentation of an e-Coaching email exchange, in which the first sentence is segmented into 2 MI behavior fragments, while the fourth and fifth segments correspond to one and three sentences, respectively.
+Segmentation of e-Coaching emails corresponds to a special type of discourse analysis \cite{webber2012discourse} aimed at better understanding the effective e-Coaching communication strategies and revealing the unique socio-psychological characteristics of a patient.
 
@@
     \includegraphics[width=0.9\textwidth]{figures/segment-example.png}
-    \caption{\textbf{Segmentation of e-coaching text depicts the main challenges of boundary detection}.}
+    \caption{\textbf{Example of e-Coaching emails segmented into fragments that correspond to MI behaviors of an e-Coach and a patient}.}
     \label{fig:text-segment}
@@
 
-In this paper, we address the text segmentation problem by developing several state-of-the-art machine learning models to promote the automatic identification of best communication strategies without human interference.
-More specifically, we developed Support Vector Machine (SVM), Naive Bayes (NB), K-Nearest Neighbor (KNN), Long Short Term Memory (LSTM), and Gated Recurrent Unit (GRU) by utilizing contextual, topic and punctuation mark features, to find the best model for the segmentation of e-coaching text.
+The goal of this research study is to assess the applicability of machine learning methods for automated segmentation of e-Coaching emails into textual fragments corresponding to individual behaviors, which is the first step of the coding process of e-Coaching communications.
+In particular, we introduced contextual, topic and punctuation features and experimented with both traditional supervised machine learning methods, such as Support Vector Machine (SVM), Naive Bayes (NB) and $k$-Nearest Neighbor (KNN) classifiers, and deep learning methods, such as Long Short Term Memory (LSTM) and Gated Recurrent Unit (GRU), to find the best performing method and feature combination.
 
-Previous studies mainly focus on segmentation of text into sections and headers\cite{apostolova2009automatic,denny2009evaluation,tepper2012statistical,cho2002text} or sentence boundary detection\cite{griffis2016quantitative,kreuzthaler2015detection,treviso2016sentence} in the medical domain.
-Apostolova et al.\cite{apostolova2009automatic} applied SVM by utilizing word-vector cosine similarity metric combined with several heuristics to classify clinical report into semantic sections such as demographics, history, exam procedure, finding, impression, etc.
+Relevant previous work in the biomedical domain primarily focused on segmentation of text into sections and headers\cite{apostolova2009automatic,denny2009evaluation,tepper2012statistical,cho2002text} or sentence boundary detection\cite{griffis2016quantitative,kreuzthaler2015detection,treviso2016sentence}.
+Apostolova et al.\cite{apostolova2009automatic} applied SVM along with word-vector cosine similarity metric combined with several heuristics to segment clinical reports into sections, such as demographics, history, procedure, finding and impression.
 After identification of each line in the document, Tepper et al.
-\cite{tepper2012statistical} trained Maximum Entropy models for the section classification.
-In 2009, Denny et al.\cite{denny2009evaluation} proposed a SecTag algorithm, which combined natural language processing technique, terminology-based rule, and naive Bayesian score for identifying sections and headers that achieved 99\% recall with 95.6\% precision.
-On the other hand, SVM exploiting with linear kernel and recurrent convolutional neural networks with prosodic, part of speech features and word embeddings, were trained by Kreuzthaler et al.\cite{kreuzthaler2015detection} and Griffis et al.\cite{griffis2016quantitative}, respectively, for the detection of sentence boundary.
+\cite{tepper2012statistical} trained Maximum Entropy models for section classification.
+Denny et al.\cite{denny2009evaluation} proposed a SecTag algorithm, which combined natural language processing techniques, terminology-based rules, and Naive Bayes classifier to identify the sections and headers that achieved 99\% recall with 95.6\% precision.
+On the other hand, SVM based on prosodic and part of speech features \cite{kreuzthaler2015detection} and recurrent convolutional neural networks using word embeddings \cite{griffis2016quantitative} were utilized for detecting sentence boundaries.
+Segmentation of e-Coaching emails is different from traditional shallow discourse analysis of conversations \cite{galley2003discourse} in that the focus is on segmentation, rather than on determining the types of transitions between the utterances or assigning utterances to speakers.
 
-Recently, an online clinical intervention called MENU GenY\cite{alexander2017motivations} (Making Effective Nutrition Choices for Generation Y) was proposed to test its efficacy.
-MENU GenY is a technology-based public health intervention to encourage increased fruit and vegetable intake among young adult age 21-30, utilized personalized e-coaching data.
+Recently, an online clinical intervention called MENU GenY \cite{alexander2017motivations} (Making Effective Nutrition Choices for Generation Y) was proposed and evaluated.
+MENU GenY is a technology-based public health intervention that relies on personalized e-coaching to encourage increased fruit and vegetable intake among young adults, aged 21-30.
 The goal of MENU GenY was to develop a better coding dictionary among GenY to improve eating habits.
-However, segmentation of clinical conversation in the context of electronically delivered interventions, in particular, segmentation of clinical interaction text into groups of MI behaviors, is still ignored while relying on manual hand-coded approach.
-Therefore, this study introduces a novel approach and the authors are not aware of any other work this approach has been considered for the segmentation of e-coaching text.
+However, segmentation of clinical conversation in the context of electronically delivered interventions, in particular, segmentation of clinical interaction text into groups of MI behaviors, is still performed manually, which slows down qualitative analysis of these interventions.
+This study introduces a novel computational approach to address this problem and the authors are unaware of any other work that focused on the same problem.
 
@@
 \bibliographystyle{vancouver}
-\bibliography{references}
+\bibliography{manuscript}
 

### 2018-03-09 paper-amia-2018/manuscript.tex
merged with alex

@@
 We formulate the text segmentation task into a binary classification, as shown in Figure~\ref{fig:classifier}.
-Clinical exchange is given as the input, it is partitioned as adjacent word pairs by sliding them.
+For NB, SVM and KNN models, clinical exchange is given as the input then it is partitioned as adjacent word pairs by sliding them.
 Each pair is classified into either ``boundary'' or ``not boundary'' class.
@@
 If all adjacent pairs of a block of text classified into ``not boundary'', the whole block of text is then treated as one segment about a single MI behavior.
-Totaly, we obtained 95,421 word pairs, which include 3,138 ``boundary'' and 92,283 ``not boundary'' instances.
+Totally, we obtained 95,421 word pairs, which include 3,138 ``boundary'' and 92,283 ``not boundary'' instances.
+For RNN, a chunk of text was taken as input sequence which gets one hot encoding of word vector for the model input.
+A binary label was assigned as output sequence where ``boundary'' and `not boundary'' words were associated with label 1 and 0, respectively.
 
@@
     \centering
-    \includegraphics[width=0.80\textwidth]{figures/classifier.png}
+    \includegraphics[width=0.80\textwidth]{figures/classifier.pdf}
     \caption{\textbf{Transformation of text segmentation task into text classification task}.}
@@
 
-For the experiment, we utilized three type of features including word (textual feature), topic, and punctuation mark.
-Each word represented in a binary format, where 1 indicates the appearance of the word and 0 for absence.
-Topics are considered as features since topic models are very effective\cite{kotov2015interpretable,hashimoto2016topic,lu2016modeling} to represent text documents.
-In this paper, we exploit a topic model named Labeled LDA\cite{ramage2009labeled} model, where topics correspond to labels.
-The generative process of Labeled LDA drawing the multinomial topic distributions over vocabulary for each topic.
-In our experiment, we represent each word in a vector of 2 topics, where the number of topics is experimentally determined by the model performance\cite{kotov2015interpretable}.
+For the experiment, we utilized three type of features including word (contextual feature), punctuation mark, and topic model-based features.
+In contextual features, each word represented in a binary format, where 1 indicates the appearance of the word and 0 for absence.
+Since topic models are very effective\cite{kotov2015interpretable,hashimoto2016topic,lu2016modeling} to represent text documents, we derived features from a topic model named Labeled LDA\cite{ramage2009labeled}.
+To obtain feature values, we utilize the experimental data which is segmented as MI code or class.
+First, we derived a class-specific multinomial $p(w|c)$ per each class $c$ from class-specific topics: $p(w|c, z)$ by marginalizing over topic z, where the number of topics is experimentally determined by the model performance\cite{kotov2015interpretable}.
+
+$$p(w|c) = {\sum_{z=1}^{K}{p(w|z, c)}}$$
+
+Here, $K$ is the number of topics per class $c$. After that, word-specific class distributions $p(c|w)$ were estimated with class-specific multinomials $p(w|c)$ by using the following formula:
+
+$$p(c|w) = \frac{p(w|c)p(c)}{p(w)}$$
+
+where $p(c)$ is the prior distribution of class $c$ in the training set and $p(w)$ is a probability of word $w$ in the collection language model estimated using maximum likelihood.
+In our experiment, we represent $p(c|w)$ as the feature showing that how indicative each word $w$ for a particular class $c$.
 Punctuation mark containing one of the symbols \{`.', `,', `!', `?', `:', `;', `-'\} is also employed as feature.
@@
 We build our RNN model with one hidden layer, output layer, and input layer.
-One segmented text was taken as input sequence which get one hot encoding of word vector for the model input.
-A binary label was assigned as output sequence where ``boundary'' and `not boundary'' words were associated with label 1 and 0, respectively.
-We reset our model state after feeding each input sequence.
-Since one-hot vector is given in the input layer, results are reported with textual and punctuation mark features only.
+We reset our model state after feeding each input sequence where input was given as one-hot encoding of word vector.
+Since one-hot vector is given in the input layer, results are reported with contextual and punctuation mark features only.
 We experimentally determined that the best performance is achieved when the number of hidden units = 25, batch size = 1, and optimizer = adam.
@@
 We conduct the experiment with 5 folds cross-validation and weighted macro-averaging of these metrics over the folds.
-All models have trained on 80\% of the word pairs and remaining 20\% of the data is used as a test set for reporting the performance of the model.
+All models have trained on 80\% of the data and remaining 20\% of the data is used as a test set for reporting the performance of the model.
 We also estimate the area under the receiving operating characteristics (ROC) curve\cite{kumar2011receiver} (AUC) metric due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets\cite{hu2015kernelized}.
@@
   \hline
-   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{contextual features only}} & \multicolumn{3}{|c|}{\textbf{contextual + punctuation marks (+ topics except RNN)}} \\\cline{2-7}
+   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{contextual features only}} & \multicolumn{3}{|c|}{\textbf{contextual + punctuation marks (+topic-based except RNN)}} \\\cline{2-7}
    & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-measure} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-measure}\\ \hline
@@
 
-As follows from Table~\ref{tab:result_boundary}, KNN performs best among all machine learning models in terms of precision and F1-measure, achieved 0.808 precision with 0.728 F1-measure when contextual features are used, and 0.820 precision with 0.779 F1-measure when a combination of contextual, topic, and punctuation mark features are used.
+As follows from Table~\ref{tab:result_boundary}, KNN performs best among all machine learning models in terms of precision and F1-measure, achieved 0.808 precision with 0.728 F1-measure when contextual features are used; and 0.820 precision with 0.779 F1-measure when a combination of contextual, punctuation mark and topic model-based features are used.
 However, NB demonstrates the lowest performance among all models in terms of all performance metrics.
 In this study, GRU appears as the second highest model, obtains 10.68\% higher recall with 3.78\%  higher F1-measure than LSTM.
-On the other hand, SVM exhibits highest recall 0.679 when only textual features are used.
-When textual features are used in combination with topic and punctuation mark features, recall increases by 0.6\%, 2.5\%, and 11.92\%; and F1-measure increases by 0\%, 3.39\%, and 7\% for NB, SVM, and KNN models, respectively.
+On the other hand, SVM exhibits highest recall 0.679 when only contextual features are used.
+When contextual features are used in combination with punctuation mark and topic model-based features, recall increases by 0.6\%, 2.5\%, and 11.92\%; and F1-measure increases by 0\%, 3.39\%, and 7\% for NB, SVM, and KNN models, respectively.
 Nevertheless, precision increases by 4.31\% and 1.49\% for SVM and KNN methods while decreases by 0.7\% in NB.
@@
   \hline
-   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{contextual features only}} & \multicolumn{3}{|c|}{\textbf{contextual + punctuation marks (+ topics except RNN)}} \\\cline{2-7}
+   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{contextual features only}} & \multicolumn{3}{|c|}{\textbf{contextual + punctuation marks (+ topic-based except RNN)}} \\\cline{2-7}
    & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-measure} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-measure}\\ \hline
@@
 We observed that performance is remarkably high in ``not boundary'' class compared to boundary detection which is expected because 96.71\% instances are from ``not boundary'' class.
-KNN achieves 22.40\%, 50.07\%, and 36.26\% higher precision, recall, and F1-measure for contextual features; and 20.85\%, 33.96\%, and 27.47\% higher precision, recall, and F1-measure for combined features compared to boundary class.
+KNN achieves 22.40\%, 50.07\%, and 36.26\% higher precision, recall, and F1-measure, respectively, for contextual features; and 20.85\%, 33.96\%, and 27.47\% higher precision, recall, and F1-measure, respectively, for combined features compared to boundary class.
 In contrast to boundary detection, RNN demonstrates the highest performance among all models.
-LSTM obtains 0.993 precision with 0.994 recall and F1-measure while LSTM exhibits 0.994 for all performance metrics.
-Impact of additional features is also consistent with ``not boundary'' classification.
+LSTM obtains 0.993 precision with 0.994 recall and F1-measure while GRU exhibits 0.994 for all performance metrics.
+Impact of punctuation mark and topic model-based features is also consistent with ``not boundary'' classification.
 Results show that F1-measure increases by 0\%, and 0.1\% for SVM and KNN models although decreases by 0.1\% for NB.
@@
  SVM & 0.981 & 0.982 & 0.981 & 0.983 & 0.983 & 0.983\\ \hline
- KNN & \textbf{0.983} & \textbf{0.984} & 0.983 & \textbf{0.986} & \textbf{0.986} & \textbf{0.986}\\ \hline
+ KNN & \textbf{0.983} & \textbf{0.984} & \textbf{0.983} & \textbf{0.986} & \textbf{0.986} & \textbf{0.986}\\ \hline
  LSTM & -- & -- & -- & \textbf{0.986} & 0.983 & 0.984 \\ \hline
@@
 GRU demonstrates the same result as KNN for precision and F1-measure except for recall.
-SVM shows moderate performance, obtains 0.981, 0.982, and 0.983 for precision, recall, and F1-measure, respectively when textual features are used.
-Influence of the additional features is also consistent as above, precision increases by 0.1\%, 0.2\%, and 0.3\%; recall increases by 0\%, 0.1\%, and 0.2\%; and F1-measure increases by 0\%, 0.2\%, and 0.3\% for NB, SVM, and KNN methods, respectively, when combined features are used.\\
+SVM shows moderate performance, obtains 0.981, 0.982, and 0.983 for precision, recall, and F1-measure, respectively, when contextual features are used.
+Influence of the additional features is also consistent as mentioned in Table~\ref{tab:result_boundary} and~\ref{tab:result_not_boundary}.
+Precision increases by 0.1\%, 0.2\%, and 0.3\%; recall increases by 0\%, 0.1\%, and 0.2\%; and F1-measure increases by 0\%, 0.2\%, and 0.3\% for NB, SVM, and KNN methods, respectively, when combined features are used.\\
 
@@
 
-The additional topic and punctuation mark feature made a significant improvement in performance of all machine learning methods.
-Nearly all cases, the model performs better when contextual features are used in combination with topic and punctuation mark features.
+Punctuation mark and topic model-based features made a significant improvement in performance of all machine learning methods.
+Nearly all cases, the model performs better when contextual features are used in combination with punctuation mark and topic model-based features.
 This results also mean that segmentation performance might be improved by adding more relevant features including human insight into the problem.
@@
     \includegraphics[width=1.0\textwidth]{figures/roc-curves.png}
-    \caption{\textbf{Receiver operating characteristic curves showing the performance of binary classifiers for the segmentation of e-coaching text when textual features (left) and combination of textual and other features (right) are used}.}
+    \caption{\textbf{Receiver operating characteristic curves showing the performance of binary classifiers for the segmentation of e-coaching text when contextual features (left) and combination of contextual and other features (right) are used}.}
     \label{fig:roc-curves}
@@
 NB shows the highest AUC values, achieved 0.978 for both cases while provides lowest classification results.
-On the other hand, KNN and SVM exhibit 0.972 and 0.835 AUCs when only textual features are used; and 0.959 and 0.844 AUCs when a combination of textual, topic and punctuation mark features are used.
+On the other hand, KNN and SVM exhibit 0.972 and 0.835 AUCs when only contextual features are used; and 0.959 and 0.844 AUCs when a combination of contextual, punctuation mark and topic model-based features are used.
 Finally, LSTM demonstrates lowest AUC values among all machine learning models, achieved 0.82 AUC while GRU achieves 4.15\% higher AUC than LSTM.
@@
 We believe that RNN will perform better if a large set of data is utilized.
-In this study, we employed 3,138 examples of boundary case which failed to achieve best results.
+In this study, we employed 3,138 examples of boundary case which limit to achieve the best performance.
 We also observed that GRU performs better than LSTM which was observed in other previous study\cite{chung2014empirical}.

### 2018-03-09 paper-amia-2018/manuscript.tex
Alex edited up to Methods

@@
 
-\title{Machine Learning Methods for Discourse Segmentation of E-Mail Based Clinical Communication}
+\title{Machine Learning Methods for Discourse Segmentation of Communications in E-Mail Based Behavioral Interventions}
 
@@
 \noindent{\bf Abstract}
-\textit{Communication science to understand clinical process like Motivational Interviews (MI) is limited by traditional qualitative coding methods.
-Qualitative coding of communication data is very resource-intensive and time-consuming process, which requires auto coding.
-This study utilized e-coaching data where email is used to deliver motivation-enhancing coaching to encourage healthy eating.
-A critical step toward automatic annotation of communication coding process is the segmentation of text data.
-In this study, we transformed segmentation task into a classification task and developed several state-of-the-art machine learning models including Support Vector Machine, Naive Bayes, K-Nearest Neighbor (KNN), Recurrent Neural Networks by utilizing contextual, topic and punctuation mark features.
-Results indicate that KNN is the best model and achieved 0.986 F1-measure in overall, 0.779 and 0.993 F1-measures for detecting ``boundary'' and ``not boundary'' classes, respectively.
-This study has a great implication to save money and accelerate the pace of identifying effective communication strategies.}
+\textit{Communication science approaches to developing effective behavior interventions, such as motivational interviewing (MI), are limited by traditionally manual qualitative coding of communication exchanges, which is a very resource-intensive and time-consuming process.
+This study focuses on the analysis of e-Coaching sessions, behavior interventions that are delivered via email and grounded in the principles of MI.
+A critical step towards automated annotation of e-Coaching communication exchanges is segmentation of emails into textual fragments that correspond to MI behaviors.
+In this work, we formulate this task as a classification problem and propose lexical, punctuation and topic features to address it.
+We experimented both with traditional machine learning methods, such as Support Vector Machine, Naive Bayes and K-Nearest Neighbor (KNN) classifiers and recurrent neural networks (RNNs).
+Experimental results indicate that KNN outperformed RNNs achieving 0.986 macro F1-score overall, and 0.779 and 0.993 macro F1-score for detecting ``new segment'' and ``same segment'' classes, respectively.}
 
@@
 
-In this study, we focus on the analysis of e-Coaching sessions, MI interventions that are delivered via email and grounded in the principles of motivational interviewing.
+In this study, we focus on the analysis of e-Coaching sessions, behavior interventions that are delivered via email and grounded in the principles of motivational interviewing.
 Specifically, the e-Coaches involved in this study used emails to communicate motivation-enhancing messages that encourage healthy eating among GenY adolescents.
@@
 The goal of this research study is to assess the applicability of machine learning methods for automated segmentation of e-Coaching emails into textual fragments corresponding to individual behaviors, which is the first step of the coding process of e-Coaching communications.
-In particular, we introduced contextual, topic and punctuation features and experimented with both traditional supervised machine learning methods, such as Support Vector Machine (SVM), Naive Bayes (NB) and $k$-Nearest Neighbor (KNN) classifiers, and deep learning methods, such as Long Short Term Memory (LSTM) and Gated Recurrent Unit (GRU), to find the best performing method and feature combination.
+In particular, we introduced lexical, topic and punctuation features and experimented with both traditional supervised machine learning methods, such as Support Vector Machine (SVM), Naive Bayes (NB) and K-Nearest Neighbor (KNN) classifiers, and deep learning methods, such as Long Short Term Memory (LSTM) and Gated Recurrent Unit (GRU), to find the best performing method and feature combination.
 
@@
 Each session represents an MI intervention delivered via email.
-To filter out noise from the dataset, non-ascii characters are removed and then applied stemming to obtain a general form of word from different word representations, such as ``eating'', ``eats'', and ``eat''.
-We formulate the text segmentation task into a binary classification, as shown in Figure~\ref{fig:classifier}.
-For NB, SVM and KNN models, clinical exchange is given as the input then it is partitioned as adjacent word pairs by sliding them.
-Each pair is classified into either ``boundary'' or ``not boundary'' class.
-The original text is segmented at the position, where an adjacent word pair classified into ``boundary'' class.
-If all adjacent pairs of a block of text classified into ``not boundary'', the whole block of text is then treated as one segment about a single MI behavior.
-Totally, we obtained 95,421 word pairs, which include 3,138 ``boundary'' and 92,283 ``not boundary'' instances.
-For RNN, a chunk of text was taken as input sequence which gets one hot encoding of word vector for the model input.
-A binary label was assigned as output sequence where ``boundary'' and `not boundary'' words were associated with label 1 and 0, respectively.
+During pre-processing, we removed all non-ASCII characters and applied stemming to normalize morphological variants of related concepts, such as ``eating'', ``eats'', and ``eat''.
+We formulate the segmentation task as a binary classification problem, as illustrated in Figure~\ref{fig:classifier} and experimented both with traditional machine learning methods, such as Support Vector Machine (SVM), Naive Bayes (NB) and K-Nearest Neighbor (KNN) classifiers and recurrent neural networks (RNNs), such as Long Short Term Memory (LSTM) and Gated Recurrent Unit (GRU).
+For NB, SVM and KNN models, e-Coaching email exchanges are partitioned into adjacent word pairs.
+Each pair is classified into either ``new segment'' or ``same segment'' class.
+The gold standard is manually segmented where adjacent word pairs are classified into ``new segment'' class.
+If all adjacent word pairs of a block of text are classified into ``same segment'' class, the entire block is treated as one textual fragment corresponding to a single MI behavior.
+In total, we obtained 95,421 word pairs, which include 3,138 ``new segment'' and 92,283 ``same segment'' instances.
+For RNNs, a block of text was taken as input sequence, such that one-hot encodings of each word or punctuation marks in a block were used as input into an RNN and binary labels (1 or 0) corresponding to ``new segment'' and ``same segment'' classification decision were considered as the output of RNN at each step.
+In the gold standard, words within the same segment were assigned the label of 0 and the last word or punctuation mark of a segment were assigned the label of 1.
 
@@
 
-For the experiment, we utilized three type of features including word (contextual feature), punctuation mark, and topic model-based features.
-In contextual features, each word represented in a binary format, where 1 indicates the appearance of the word and 0 for absence.
-Since topic models are very effective\cite{kotov2015interpretable,hashimoto2016topic,lu2016modeling} to represent text documents, we derived features from a topic model named Labeled LDA\cite{ramage2009labeled}.
-To obtain feature values, we utilize the experimental data which is segmented as MI code or class.
-First, we derived a class-specific multinomial $p(w|c)$ per each class $c$ from class-specific topics: $p(w|c, z)$ by marginalizing over topic z, where the number of topics is experimentally determined by the model performance\cite{kotov2015interpretable}.
+\subsection*{\textit{Features}}
 
-$$p(w|c) = {\sum_{z=1}^{K}{p(w|z, c)}}$$
+We utilized three types of features in conjunction with SVM, NB and KNN methods: lexical, punctuation and topic features.
+Lexical features represent each word in a pair as a binary vector, in which 1 corresponds to the word in question and 0 to all other words.
+Since topic models\cite{kotov2015interpretable,hashimoto2016topic,lu2016modeling} have been shown to be effective semantic abstractions of individual words, we derived topic features from Labeled LDA\cite{ramage2009labeled}, a topic model for annotated corpora.
+We considered each textual segment in the training set as a document and the MI code assigned to the segment as a label.
+First, we derived a label-specific multinomial $p(w|l)$ for each label $l$ from the word distributions $p(w|l,z)$ for topic $z$ specific to label $l$ by marginalizing over topics:
 
-Here, $K$ is the number of topics per class $c$. After that, word-specific class distributions $p(c|w)$ were estimated with class-specific multinomials $p(w|c)$ by using the following formula:
+$$p(w|l) = {\sum_{z=1}^{K}{p(w|z,l)}}$$
 
-$$p(c|w) = \frac{p(w|c)p(c)}{p(w)}$$
+where $K$ is the optimal number of topics experimentally determined to minimize perplexity.
+After that, distributions over labels, $p(l|w)$, for each word were obtained from label-specific multinomials $p(w|l)$ using Bayesian inversion:
 
-where $p(c)$ is the prior distribution of class $c$ in the training set and $p(w)$ is a probability of word $w$ in the collection language model estimated using maximum likelihood.
-In our experiment, we represent $p(c|w)$ as the feature showing that how indicative each word $w$ for a particular class $c$.
-Punctuation mark containing one of the symbols \{`.', `,', `!', `?', `:', `;', `-'\} is also employed as feature.
-This is one of the most important features as they indicate the boundary of a sentence, clause, and phrase.
+$$p(l|w) = \frac{p(w|l)p(l)}{p(w)}$$
 
-\subsection*{\textit{Segmentation classifiers}}
-Several state-of-the-art classifiers, including Naive Bayes (NB)\cite{pedregosa2011scikit}, Support Vector Machine (SVM)\cite{chang2011libsvm}, K-Nearest Neighbor (KNN)\cite{pedregosa2011scikit}, two variant of Recurrent Neural Networks (RNN)\cite{bengio1993problem}: Long Short Term Memory (LSTM)\cite{hochreiter1997long} and Gated Recurrent Unit (GRU)\cite{cho2014properties}, are employed to estimate the classification performance.
+where $p(l)$ is the prior probabilities of label $l$ in the training set and $p(w)$ is the prior probability of word $w$ in the language model estimated from the training set using maximum likelihood.
+We use the distribution $p(c|w)$ as the topical feature vector, which shows how indicative each word $w$ is for a particular behavior code $l$.
+Topical features allow to abstract away from individual words and detect segment boundaries by indirectly capturing transitions between behavior codes when only observing a pair of words that are specific to different behavior codes.
+Punctuation marks, which correspond to one of the symbols \{`.', `,', `!', `?', `:', `;', `-'\} between a pair of words, are also employed as a feature, since punctuation marks designate the boundary of a sentence, clause, and phrase and often also correspond to segment boundary.
 
-\textbf{Naive Bayes}: this model is constructed by using the training data and estimate the prior probability of classes, and each feature has given the class.
-Then, the posterior probability is computed to predict the class label by applying the Bayes theorem with the assumption that features are conditionally independent.
-This study utilized a specialized version of Naive Bayes called Multinomial Naive Bayes, which is best suitable for discrete features such as word.
+\subsection*{\textit{Classifiers}}
 
-\textbf{Support Vector Machine}: we used this model as one of the state-of-the-art classification technique proven to perform well in text categorisation\cite{joachims1998text} for its ability to cope with very high dimensional input feature space.
-SVM finds the best hyperplane in the feature space that maximizes the separation between the closest ``boundary'' and ``not boundary'' training examples.
-In this experiment, the polynomial kernel is employed to train the SVM model for the segmentation of e-coaching text.
+We experimented both with traditional classifiers, including Naive Bayes (NB)\cite{pedregosa2011scikit}, Support Vector Machine (SVM)\cite{chang2011libsvm}, K-Nearest Neighbor (KNN)\cite{pedregosa2011scikit} and two variants of Recurrent Neural Networks (RNNs)\cite{bengio1993problem}: Long Short Term Memory (LSTM)\cite{hochreiter1997long} and Gated Recurrent Unit (GRU)\cite{cho2014properties}.
 
-\textbf{K-Nearest Neighbor}: By this model, each training sample represented as a point in the input feature space.
-For a new test sample, Euclidean distance is calculated to find the k-nearest neighbors.
-Finally, the test sample is classified into majority class of the k-nearest neighbors.
-We experimentally determined that best performance was achieved with k = 3 for the classification of word pairs.
+\textbf{Naive Bayes}: this model estimates the prior probability of classes along with conditional probabilities of each feature given the class.
+Then, the posterior probability is computed to predict the class for each sample by applying the Bayes rule with the assumption that features are conditionally independent.
+We used Multinomial Naive Bayes in experiments.
 
-\textbf{Recurrent Neural Networks}: RNN is a neural network architecture designed to capture sequential patterns present in temporal sequence such as text data.
-When we predict the ``boundary'' point, adjacent word pair will help to understand the pattern of the sequence.
-Long Short Term Memory networks usually reffered as LSTMs\cite{hochreiter1997long}, are a special type of RNN capable of handling variable size input sequence, contains internal memory.
+\textbf{Support Vector Machine}: state-of-the-art supervised classification method proven to be effective for text categorization\cite{joachims1998text} for its ability to cope with high dimensional input feature space.
+SVM finds the hyperplane that maximizes the separation between the closest ``new segment'' and ``same segment'' training examples.
+We used SVM with polynomial kernel in experiments.
+
+\textbf{K-Nearest Neighbor}: this method considers each training sample as a point in the input feature space.
+For a new test sample, Euclidean distance is calculated to find the k-nearest classified neighbors from the training set.
+Finally, the test sample is classified into a majority class of its k-nearest neighbors.
+We experimentally determined that best performance was achieved when the number of neighbors is 3.
+
+\textbf{Recurrent Neural Networks}: RNN is a neural network architecture designed to capture sequential patterns.
+When we predict the ``new segment'' point using RNNs, particular combinations of words in a sequence will indicate the change of MI behavior.
+Long Short Term Memory (LSTM) networks\cite{hochreiter1997long} are a special type of RNN that are capable of handling variable size input sequence and have an internal memory that can be reset.
 GRU\cite{cho2014properties} is a variant of LSTM mathematically represented by the following formula:
@@
 The activation $h_t$ of the GRU at time $t$ is a linear combination of previous activation $h_{t-1}$ and the candidate activation $\tilde h_t$, which is represented by Eq.~(\ref{eq:lastgru}) and~(\ref{eq:candidategru}).
-We build our RNN model with one hidden layer, output layer, and input layer.
+Our GRU model includes one hidden layer, output layer, and input layer.
 We reset our model state after feeding each input sequence where input was given as one-hot encoding of word vector.
-Since one-hot vector is given in the input layer, results are reported with contextual and punctuation mark features only.
-We experimentally determined that the best performance is achieved when the number of hidden units = 25, batch size = 1, and optimizer = adam.
+Since LSTM and GRU use one-hot vector as input, results for these models are reported only when lexical and punctuation features are used.
+We experimentally determined that the best performance is achieved when the number of hidden units is 25, batch size is 1, and Adam is used for optimization.
 
 \subsection*{\textit{Evaluation metrics}}
-In this experiment, standard metrics: precision, recall, and F1-measure, are applied to evaluate the performance of binary classifiers\cite{aas1999text}.
+We report standard metrics for experiments (precision, recall, and F1-measure) to evaluate the performance of binary classifiers\cite{aas1999text}.
 However, accuracy is not reported as a performance metric because accuracy is highly sensitive to the prior class probabilities and does not fully describe the actual difficulty of the decision problem for an unbalanced dataset.
-We conduct the experiment with 5 folds cross-validation and weighted macro-averaging of these metrics over the folds.
-All models have trained on 80\% of the data and remaining 20\% of the data is used as a test set for reporting the performance of the model.
-We also estimate the area under the receiving operating characteristics (ROC) curve\cite{kumar2011receiver} (AUC) metric due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets\cite{hu2015kernelized}.
+The results are reported based on 5 folds cross-validation and weighted macro-averaging over the folds.
+We also estimate the area under the receiving operating characteristics (ROC) curve\cite{kumar2011receiver} (AUC) metric due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets \cite{hu2015kernelized}.
 
@@
   \hline
-   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{contextual features only}} & \multicolumn{3}{|c|}{\textbf{contextual + punctuation marks (+topic-based except RNN)}} \\\cline{2-7}
+   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{lexical features only}} & \multicolumn{3}{|c|}{\textbf{lexical + punctuation marks (+topic-based except RNN)}} \\\cline{2-7}
    & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-measure} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-measure}\\ \hline
@@
 
-As follows from Table~\ref{tab:result_boundary}, KNN performs best among all machine learning models in terms of precision and F1-measure, achieved 0.808 precision with 0.728 F1-measure when contextual features are used; and 0.820 precision with 0.779 F1-measure when a combination of contextual, punctuation mark and topic model-based features are used.
+As follows from Table~\ref{tab:result_boundary}, KNN performs best among all machine learning models in terms of precision and F1-measure, achieved 0.808 precision with 0.728 F1-measure when lexical features are used; and 0.820 precision with 0.779 F1-measure when a combination of lexical, punctuation mark and topic model-based features are used.
 However, NB demonstrates the lowest performance among all models in terms of all performance metrics.
 In this study, GRU appears as the second highest model, obtains 10.68\% higher recall with 3.78\%  higher F1-measure than LSTM.
-On the other hand, SVM exhibits highest recall 0.679 when only contextual features are used.
-When contextual features are used in combination with punctuation mark and topic model-based features, recall increases by 0.6\%, 2.5\%, and 11.92\%; and F1-measure increases by 0\%, 3.39\%, and 7\% for NB, SVM, and KNN models, respectively.
+On the other hand, SVM exhibits highest recall 0.679 when only lexical features are used.
+When lexical features are used in combination with punctuation mark and topic model-based features, recall increases by 0.6\%, 2.5\%, and 11.92\%; and F1-measure increases by 0\%, 3.39\%, and 7\% for NB, SVM, and KNN models, respectively.
 Nevertheless, precision increases by 4.31\% and 1.49\% for SVM and KNN methods while decreases by 0.7\% in NB.
@@
   \hline
-   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{contextual features only}} & \multicolumn{3}{|c|}{\textbf{contextual + punctuation marks (+ topic-based except RNN)}} \\\cline{2-7}
+   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{lexical features only}} & \multicolumn{3}{|c|}{\textbf{lexical + punctuation marks (+ topic-based except RNN)}} \\\cline{2-7}
    & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-measure} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-measure}\\ \hline
@@
 We observed that performance is remarkably high in ``not boundary'' class compared to boundary detection which is expected because 96.71\% instances are from ``not boundary'' class.
-KNN achieves 22.40\%, 50.07\%, and 36.26\% higher precision, recall, and F1-measure, respectively, for contextual features; and 20.85\%, 33.96\%, and 27.47\% higher precision, recall, and F1-measure, respectively, for combined features compared to boundary class.
+KNN achieves 22.40\%, 50.07\%, and 36.26\% higher precision, recall, and F1-measure, respectively, for lexical features; and 20.85\%, 33.96\%, and 27.47\% higher precision, recall, and F1-measure, respectively, for combined features compared to boundary class.
 In contrast to boundary detection, RNN demonstrates the highest performance among all models.
@@
 GRU demonstrates the same result as KNN for precision and F1-measure except for recall.
-SVM shows moderate performance, obtains 0.981, 0.982, and 0.983 for precision, recall, and F1-measure, respectively, when contextual features are used.
+SVM shows moderate performance, obtains 0.981, 0.982, and 0.983 for precision, recall, and F1-measure, respectively, when lexical features are used.
 Influence of the additional features is also consistent as mentioned in Table~\ref{tab:result_boundary} and~\ref{tab:result_not_boundary}.
@@
 Punctuation mark and topic model-based features made a significant improvement in performance of all machine learning methods.
-Nearly all cases, the model performs better when contextual features are used in combination with punctuation mark and topic model-based features.
+Nearly all cases, the model performs better when lexical features are used in combination with punctuation mark and topic model-based features.
 This results also mean that segmentation performance might be improved by adding more relevant features including human insight into the problem.
@@
     \includegraphics[width=1.0\textwidth]{figures/roc-curves.png}
-    \caption{\textbf{Receiver operating characteristic curves showing the performance of binary classifiers for the segmentation of e-coaching text when contextual features (left) and combination of contextual and other features (right) are used}.}
+    \caption{\textbf{Receiver operating characteristic curves showing the performance of binary classifiers for the segmentation of e-coaching text when lexical features (left) and combination of lexical and other features (right) are used}.}
     \label{fig:roc-curves}
@@
 NB shows the highest AUC values, achieved 0.978 for both cases while provides lowest classification results.
-On the other hand, KNN and SVM exhibit 0.972 and 0.835 AUCs when only contextual features are used; and 0.959 and 0.844 AUCs when a combination of contextual, punctuation mark and topic model-based features are used.
+On the other hand, KNN and SVM exhibit 0.972 and 0.835 AUCs when only lexical features are used; and 0.959 and 0.844 AUCs when a combination of lexical, punctuation mark and topic model-based features are used.
 Finally, LSTM demonstrates lowest AUC values among all machine learning models, achieved 0.82 AUC while GRU achieves 4.15\% higher AUC than LSTM.

### 2018-03-09 paper-amia-2018/manuscript.tex
Alex edited

@@
 \section*{Results}
-Experimental results are evaluated with ``boundary'' and ``not boundary'' classes as well as their weighted average, which are shown in Table~\ref{tab:result_boundary}, \ref{tab:result_not_boundary}, and \ref{tab:result_weighted_avg}, respectively.\\
+
+Experimental results are reported with respect to ``new segment'' and ``same segment'' classes as well as their weighted average in Tables~\ref{tab:result_boundary}, \ref{tab:result_not_boundary}, and \ref{tab:result_weighted_avg}, respectively.\\
 
@@
 \centering
-\caption{Performance of NB, SVM, KNN, and RNN methods for detecting segmentation boundary in e-coaching text. The highest value for each performance metric is highlighted in bold.}
+\caption{Performance of NB, SVM, KNN, and RNN methods for identification of ``new segment'' class in e-Coaching emails. The highest value for each performance metric is highlighted in bold.}
 \label{tab:result_boundary}
@@
   \hline
-   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{lexical features only}} & \multicolumn{3}{|c|}{\textbf{lexical + punctuation marks (+topic-based except RNN)}} \\\cline{2-7}
+   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{lexical features only}} & \multicolumn{3}{|c|}{\textbf{lexical + punctuation (+ topic features except RNN)}} \\\cline{2-7}
    & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-measure} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-measure}\\ \hline
@@
 
-As follows from Table~\ref{tab:result_boundary}, KNN performs best among all machine learning models in terms of precision and F1-measure, achieved 0.808 precision with 0.728 F1-measure when lexical features are used; and 0.820 precision with 0.779 F1-measure when a combination of lexical, punctuation mark and topic model-based features are used.
-However, NB demonstrates the lowest performance among all models in terms of all performance metrics.
-In this study, GRU appears as the second highest model, obtains 10.68\% higher recall with 3.78\%  higher F1-measure than LSTM.
-On the other hand, SVM exhibits highest recall 0.679 when only lexical features are used.
-When lexical features are used in combination with punctuation mark and topic model-based features, recall increases by 0.6\%, 2.5\%, and 11.92\%; and F1-measure increases by 0\%, 3.39\%, and 7\% for NB, SVM, and KNN models, respectively.
-Nevertheless, precision increases by 4.31\% and 1.49\% for SVM and KNN methods while decreases by 0.7\% in NB.
+As follows from Table~\ref{tab:result_boundary}, KNN outperforms all other models in terms of precision and F1-measure achieving 0.808 precision with 0.728 F1-measure when lexical features are used; 0.820 precision and 0.779 F1-measure when a combination of lexical, punctuation and topic features are used.
+NB demonstrates the lowest performance among all models in terms of all performance metrics.
+GRU has the second best results showing 10.68\% higher recall and 3.78\%  higher F1-measure than LSTM.
+On the other hand, SVM has the highest recall of 0.679 when only lexical features are used.
+When lexical features are used in combination with punctuation mark and topic model-based features, recall increases by 0.6\%, 2.5\%, and 11.92\%; and F1-measure increases by 0\%, 3.39\%, and 7\% for NB, SVM, and KNN models, respectively, while precision increases by 4.31\% and 1.49\% for SVM and KNN methods and decreases by 0.7\% for NB.
 \\
@@
 \centering
-\caption{Performance of NB, SVM, KNN, and RNN methods for the identification of ``not boundary'' class. The highest value for each performance metric is highlighted in bold.}
+\caption{Performance of NB, SVM, KNN, and RNN methods for identification of ``same segment'' class in e-Coaching emails. The highest value for each performance metric is highlighted in bold.}
 \label{tab:result_not_boundary}
@@
   \hline
-   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{lexical features only}} & \multicolumn{3}{|c|}{\textbf{lexical + punctuation marks (+ topic-based except RNN)}} \\\cline{2-7}
+   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{lexical features only}} & \multicolumn{3}{|c|}{\textbf{lexical + punctuation (+ topic features except RNN)}} \\\cline{2-7}
    & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-measure} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-measure}\\ \hline
@@
 
-Table~\ref{tab:result_not_boundary} summarizes the performance of NB, SVM, KNN, and RNN models for detecting ``not boundary'' class in e-coaching text.
-We observed that performance is remarkably high in ``not boundary'' class compared to boundary detection which is expected because 96.71\% instances are from ``not boundary'' class.
-KNN achieves 22.40\%, 50.07\%, and 36.26\% higher precision, recall, and F1-measure, respectively, for lexical features; and 20.85\%, 33.96\%, and 27.47\% higher precision, recall, and F1-measure, respectively, for combined features compared to boundary class.
+Table~\ref{tab:result_not_boundary} summarizes the performance of NB, SVM, KNN, and RNN models for detecting ``same segment'' class.
+We observed that performance is remarkably better for ``same segment'' class compared to new segment detection, which is expected, since 96.71\% instances belong to ``same segment'' class.
+KNN achieves 22.40\%, 50.07\%, and 36.26\% higher precision, recall, and F1-measure, respectively, for lexical features; and 20.85\%, 33.96\%, and 27.47\% higher precision, recall, and F1-measure, respectively, for combined features compared to new segment classification.
 In contrast to boundary detection, RNN demonstrates the highest performance among all models.
-LSTM obtains 0.993 precision with 0.994 recall and F1-measure while GRU exhibits 0.994 for all performance metrics.
-Impact of punctuation mark and topic model-based features is also consistent with ``not boundary'' classification.
-Results show that F1-measure increases by 0\%, and 0.1\% for SVM and KNN models although decreases by 0.1\% for NB.
+LSTM obtains 0.993 precision with 0.994 recall and F1-measure while GRU has 0.994 for all performance metrics.
+The impact of punctuation and topic model-based features is consistent with ``new segment'' classification.
+Results indicate that F1-measure increases by 0\%, and 0.1\% for SVM and KNN models and decreases by 0.1\% for NB.
 \\
@@
 \centering
-\caption{Weighted average performance of NB, SVM, KNN, and RNN methods for the segmentation of e-coaching text in detecting both ``boundary'' and ``not boundary'' classes.
+\caption{Performance of NB, SVM, KNN, and RNN methods for segmentation of e-Coaching emails as a weighted average over ``same segment'' and ``new segment'' classification results.
 The highest value for each performance metric is highlighted in bold.}
@@
 
-Table~\ref{tab:result_weighted_avg} outlines the weighted average results of the experiment on the models for the segmentation of e-coaching text by classifying them into ``boundary'' and ``not boundary'' classes.
-Overall, KNN obtains the best performance with all metrics and NB denotes the lowest performance among all methods.
-GRU demonstrates the same result as KNN for precision and F1-measure except for recall.
-SVM shows moderate performance, obtains 0.981, 0.982, and 0.983 for precision, recall, and F1-measure, respectively, when lexical features are used.
-Influence of the additional features is also consistent as mentioned in Table~\ref{tab:result_boundary} and~\ref{tab:result_not_boundary}.
-Precision increases by 0.1\%, 0.2\%, and 0.3\%; recall increases by 0\%, 0.1\%, and 0.2\%; and F1-measure increases by 0\%, 0.2\%, and 0.3\% for NB, SVM, and KNN methods, respectively, when combined features are used.\\
+Table~\ref{tab:result_weighted_avg} summarizes the weighted average results of the models for segmentation of e-Coaching emails.
+Overall, KNN has the best performance across all metrics and NB has the lowest performance among all methods.
+GRU demonstrates the same result as KNN for precision and F1-measure, but not for recall.
+SVM shows moderate performance, achieving 0.981, 0.982, and 0.983 for precision, recall, and F1-measure, respectively, when lexical features are used.
+Influence of the additional features is also consistent with the results in Tables~\ref{tab:result_boundary} and~\ref{tab:result_not_boundary}.
+Precision increases by 0.1\%, 0.2\%, and 0.3\%; recall increases by 0\%, 0.1\%, and 0.2\%; and F1-measure increases by 0\%, 0.2\%, and 0.3\% for NB, SVM, and KNN methods, respectively, when feature combination is used.\\
 
 \section*{Discussion}
-This study is the first efforts to evaluate the automatic segmentation of e-coaching text.
+This study is the first effort to evaluate the automatic segmentation of e-Coaching emails.
 Experimental results indicate that KNN is the best model among all machine learning methods considered for this study.
 KNN achieved 0.986 F1-measure in overall, 0.779 and 0.993 F1-measures for detecting ``boundary'' and ``not boundary'', respectively.
-The robust performance of KNN provides the evidence that machine learning models are capable to learn information from clinical exchange.
-Although the domain of this study was intentionally quite small, we believe that our study is not limited to the e-coaching domain, and it can be successfully applied to other domain as well.
+The robust performance of KNN provides the evidence that machine learning models are capable to learn conceptual information from clinical exchange.
+It also indicates that topic features are more important than lexical features, even when deep learning methods are employed.
+Although the domain of this study was intentionally quite small, we believe that our study is not limited to the e-coaching domain, and it can be successfully applied to other domains, in which discourse segmentation is an preliminary step for annotation.
 
 Punctuation mark and topic model-based features made a significant improvement in performance of all machine learning methods.
-Nearly all cases, the model performs better when lexical features are used in combination with punctuation mark and topic model-based features.
-This results also mean that segmentation performance might be improved by adding more relevant features including human insight into the problem.
+In nearly all cases, ML methods perform better, when lexical features are used in combination with punctuation and topic model-based features.
+This results also indicate that segmentation performance might be improved by adding additional relevant features.
 
@@
 
-In this paper, results are reported by each class to avoid confusion about the overall model performance.
-In addition, standard metrics: precision, recall, and F1-measure were used to eliminate doubt about the model performance because accuracy is misleading for imbalance dataset.
-AUC values are also outlined due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets\cite{hu2015kernelized}, which are demonstrated by the ROC curves in Figure~\ref{fig:roc-curves}.
-NB shows the highest AUC values, achieved 0.978 for both cases while provides lowest classification results.
+In this paper, results are reported by each class to avoid confusion regarding overall model performance due to severe class imbalance.
+In addition, standard metrics: precision, recall, and F1-measure were used to eliminate doubt about the model performance, since accuracy is misleading for imbalanced datasets.
+AUC is also utilized due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets\cite{hu2015kernelized}, which are demonstrated by the ROC curves in Figure~\ref{fig:roc-curves}.
+NB shows the highest AUC values, achieving 0.978 for both cases while provides lowest classification results.
 On the other hand, KNN and SVM exhibit 0.972 and 0.835 AUCs when only lexical features are used; and 0.959 and 0.844 AUCs when a combination of lexical, punctuation mark and topic model-based features are used.
-Finally, LSTM demonstrates lowest AUC values among all machine learning models, achieved 0.82 AUC while GRU achieves 4.15\% higher AUC than LSTM.
+Finally, LSTM demonstrates lowest AUC values among all machine learning models, achieving 0.82 AUC, while GRU achieves 4.15\% higher AUC, than LSTM.
 The conclusion drawn from the ROC curves also confirmed the robustness and superiority of KNN model for the segmentation of clinical exchange.
@@
 We observed the second highest performance of RNN, in particular, LSTM and GRU for the text segmentation.
-We believe that RNN will perform better if a large set of data is utilized.
+We believe that RNN will perform better if a larger dataset is utilized.
 In this study, we employed 3,138 examples of boundary case which limit to achieve the best performance.
-We also observed that GRU performs better than LSTM which was observed in other previous study\cite{chung2014empirical}.
+We also observed that GRU performs better than LSTM, which was observed in previous studies\cite{chung2014empirical}.
 
@@
 
-Our proposed approach is novel for the segmentation of e-coaching text because previous studies mainly focus on the segmentation of text into sections, headers, and sentences in other medical domain.
+Our proposed approach is novel for the segmentation of e-Coaching emails, since previous studies mainly focused on the segmentation of text into sections, headers, and sentences.
 However, this study segmented clinical exchange into groups of MI behaviors which will significantly reduce the amount of resource and time required to segment clinical exchange manually.
-Furthermore, these segmentation models could be integrated with auto coding classifiers to build a software package of automatic coding procedure of clinical exchange.
+Furthermore, these segmentation models could be integrated with auto coding classifiers to build a software pipeline for automated annotation of exchanges.
 
-The limitation of this study is that e-coaching text is collected from a single medical institute; formatting, style, and email segment can be different in other settings.
+The limitation of this study is that e-Coaching data is collected from a single medical institute; formatting, style, and email segment can be different in other settings.
 Therefore, there is a need to replicate the experiments with different data sets.
-As our future work, we plan to evaluate our approach on other datasets involves in discourse analysis.
-We also plan to use more relevant features to improve model performance.
-For example, part-of-speech tagging\cite{hasan2016feedback} and distance from the beginning and end of the current sentence might significantly enhance the classification performance.
+As our future work, we plan to evaluate our approach on other datasets for discourse analysis.
+
+%We also plan to use more relevant features to improve model performance.
+For example, part-of-speech tagging\cite{hasan2016feedback} and distance from %the beginning and end of the current sentence might significantly enhance the classification performance.
 
 \section*{Conclusion}
-Segmentation of e-coaching text is an integral part of developing an automated e-coaching intervention.
-Although several studies have done in clinical interventions, they are limited by the qualitative coding of clinical interactions.
-In addition, previous studies in the medical domain mainly segmented clinical text into sections and sentences, none of them are used for the segmentation of text into groups of MI behaviors in the setting of discourse analysis with email under the principle of motivational interviews.
+Segmentation of e-coaching emails is an integral part of developing e-Coaching interventions.
+Although several studies have focused on clinical interventions, they are limited by the qualitative coding of clinical interactions.
+In addition, previous studies in the medical domain mainly segmented clinical text into sections and sentences, none of them considered segmentation of text into groups of MI behaviors in the setting of discourse analysis with emails.
 In this paper, we compared the performance of machine learning models for the task of segmentation of e-coaching text.

### 2018-03-09 paper-amia-2018/manuscript.tex
submitted version of amia 2018

@@
 
-We experimented both with traditional classifiers, including Naive Bayes (NB)\cite{pedregosa2011scikit}, Support Vector Machine (SVM)\cite{chang2011libsvm}, K-Nearest Neighbor (KNN)\cite{pedregosa2011scikit} and two variants of Recurrent Neural Networks (RNNs)\cite{bengio1993problem}: Long Short Term Memory (LSTM)\cite{hochreiter1997long} and Gated Recurrent Unit (GRU)\cite{cho2014properties}.
+We experimented both with traditional classifiers, including Naive Bayes (NB)\cite{mccallum1998comparison}, Support Vector Machine (SVM)\cite{cortes1995support}, K-Nearest Neighbor (KNN)\cite{keller1985fuzzy} and two variants of Recurrent Neural Networks (RNNs)\cite{lipton2015critical}: Long Short Term Memory (LSTM)\cite{hochreiter1997long} and Gated Recurrent Unit (GRU)\cite{chung2014empirical}.
 
@@
 Long Short Term Memory (LSTM) networks\cite{hochreiter1997long} are a special type of RNN that are capable of handling variable size input sequence and have an internal memory that can be reset.
-GRU\cite{cho2014properties} is a variant of LSTM mathematically represented by the following formula:
+GRU\cite{chung2014empirical} is a variant of LSTM mathematically represented by the following formula:
 
@@
 Since LSTM and GRU use one-hot vector as input, results for these models are reported only when lexical and punctuation features are used.
-We experimentally determined that the best performance is achieved when the number of hidden units is 25, batch size is 1, and Adam is used for optimization.
+We experimentally determined that the best performance is achieved when the number of hidden units is 25, batch size is 1, and Adam\cite{kingma2014adam} is used for optimization.
 

### 2018-05-22 paper-amia-2018/manuscript.tex
update abstract, method, discussion and conclusion sections

@@
 In this work, we formulate this task as a classification problem and propose lexical, punctuation and topic features to address it.
-We experimented both with traditional machine learning methods, such as Support Vector Machine, Naive Bayes and K-Nearest Neighbor (KNN) classifiers and recurrent neural networks (RNNs).
-Experimental results indicate that KNN outperformed RNNs achieving 0.986 macro F1-score overall, and 0.779 and 0.993 macro F1-score for detecting ``new segment'' and ``same segment'' classes, respectively.}
+We experimented both with traditional machine learning methods, such as Support Vector Machine (SVM), Naive Bayes and K-Nearest Neighbor (KNN) classifiers and recurrent neural networks (RNNs).
+Results indicate that SVM outperformed KNN and RNNs achieving 0.990 macro F1-score overall, and 0.848 and 0.995 macro F1-score for detecting ``new segment'' and ``same segment'' classes, respectively.}
 
@@
 
-$$p(w|l) = {\sum_{z=1}^{K}{p(w|z,l)}}$$
+\begin{equation}
+\label{eq:pwl}
+p(w|l) = {\sum_{z=1}^{K}{p(w|z,l)}}
+\end{equation}
 
@@
 
-$$p(l|w) = \frac{p(w|l)p(l)}{p(w)}$$
+\begin{equation}
+\label{eq:plw}
+p(l|w) = \frac{p(w|l)p(l)}{p(w)}
+\end{equation}
 
 where $p(l)$ is the prior probabilities of label $l$ in the training set and $p(w)$ is the prior probability of word $w$ in the language model estimated from the training set using maximum likelihood.
-We use the distribution $p(c|w)$ as the topical feature vector, which shows how indicative each word $w$ is for a particular behavior code $l$.
-Topical features allow to abstract away from individual words and detect segment boundaries by indirectly capturing transitions between behavior codes when only observing a pair of words that are specific to different behavior codes.
+We use the distribution $p(l|w)$ as the topic feature vector, which shows how indicative each word $w$ is for a particular behavior code $l$.
+Topic features allow to abstract away from individual words and detect segment boundaries by indirectly capturing transitions between behavior codes when only observing a pair of words that are specific to different behavior codes.
+To find the optimal topic feature, we also derived the same topic feature $p(l|w)$ by using two recently proposed probabilistic generative latent variable models: Discriminative Labeled Latent Dirichlet Allocation (DL-LDA) and Latent Class Allocation (LCA)\cite{kotov2015interpretable}.
+Similar as L-LDA, DL-LDA-based topic feature is extracted by first deriving a label-specific language model $p(w|l)$ for each label $l$ from the word distributions $p(w|l,z)$ for topic $z$ specific to label $l$ by using the Eq.~\ref{eq:pwl} and \ref{eq:plw}.
+On the other hand, LCA directly provides label-specific multinomials $p(w|l)$, which can be used to derive topic feature by using Eq.~\ref{eq:plw}.
 Punctuation marks, which correspond to one of the symbols \{`.', `,', `!', `?', `:', `;', `-'\} between a pair of words, are also employed as a feature, since punctuation marks designate the boundary of a sentence, clause, and phrase and often also correspond to segment boundary.
@@
 SVM finds the hyperplane that maximizes the separation between the closest ``new segment'' and ``same segment'' training examples.
-We used SVM with polynomial kernel in experiments.
+We used L1-regularized linear SVM with squared error minimization in experiments.
 
@@
 \centering
-\caption{Performance of NB, SVM, KNN, and RNN methods for identification of ``new segment'' class in e-Coaching emails. The highest value for each performance metric is highlighted in bold.}
-\label{tab:result_boundary}
-  \begin{tabular}{|l|l|l|l|p{0.15\linewidth}|p{0.15\linewidth}|l|}
+\caption{Performance of NB, SVM, and KNN methods for identification of ``new segment'' and ``same segment'' classes as well as segmentation of e-Coaching emails as a weighted average over ``same segment'' and ``new segment'' in e-Coaching emails.
+The highest value for each performance metric is highlighted in bold.}
+\label{tab:result_base}
+  \begin{tabular}{|l|l|l|l|l|l|l|l|l|l|}
   \hline
-   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{lexical features only}} & \multicolumn{3}{|c|}{\textbf{lexical + punctuation (+ topic features except RNN)}} \\\cline{2-7}
-   & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-measure} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-measure}\\ \hline
-
- NB & 0.594 & 0.662 & 0.626 & 0.590 & 0.666 & 0.626 \\ \hline
- SVM & 0.742 & \textbf{0.679} & 0.709 & 0.774 & 0.696 & 0.733\\ \hline
- KNN & \textbf{0.808} & 0.663 & \textbf{0.728} & \textbf{0.820} & \textbf{0.742} & \textbf{0.779}\\ \hline
- LSTM & -- & -- & -- & 0.800 & 0.646 & 0.714  \\ \hline
- GRU & -- & -- & -- & 0.800 & 0.715 & 0.741 \\ \hline 
+   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{new segment}} & \multicolumn{3}{|c|}{\textbf{same segment}}  & \multicolumn{3}{|c|}{\textbf{overall}} \\\cline{2-10}
+   & \textbf{Precision}  & \textbf{Recall} & \textbf{F1} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1}\\ \hline    
+ \multicolumn{10}{|c|}{lexical features only} \\ \hline   
+ NB & 0.594 & 0.662 & 0.626 & 0.988 & 0.985 & 0.987 & 0.975 & 0.974 & 0.975  \\ \hline
+ SVM & 0.762 & \textbf{0.673} & 0.715 & \textbf{0.989} & 0.993 & 0.991  & 0.981 & 0.982 & 0.982 \\ \hline
+ KNN & \textbf{0.808} & 0.663 & \textbf{0.728} & \textbf{0.989} & \textbf{0.995} & \textbf{0.992} & \textbf{0.983} & \textbf{0.984} & \textbf{0.983} \\ \hline
+ \multicolumn{10}{|c|}{lexical and puctuation features} \\ \hline
+ LSTM & \textbf{0.800} & 0.646 & 0.714  & 0.993 & \textbf{0.994} & \textbf{0.994} & \textbf{0.986} & 0.983 & 0.984 \\ \hline
+ GRU & \textbf{0.800} & \textbf{0.715} & \textbf{0.741} & \textbf{0.994} & \textbf{0.994} & \textbf{0.994} & \textbf{0.986} & \textbf{0.985} & \textbf{0.986} \\ \hline
   \end{tabular}
@@
 \centering
-\caption{Performance of NB, SVM, KNN, and RNN methods for identification of ``same segment'' class in e-Coaching emails. The highest value for each performance metric is highlighted in bold.}
+\caption{Performance of NB, SVM, and KNN methods with different topic model's features for identification of ``new segment'' class in e-Coaching emails, when all three features are used together.
+The highest value for each performance metric is highlighted in bold.}
+\label{tab:result_boundary}
+  \begin{tabular}{|l|l|l|l|l|l|l|l|l|l|}
+  \hline
+   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{L-LDA-based topic features}} & \multicolumn{3}{|c|}{\textbf{DL-LDA-based topic features}}  & \multicolumn{3}{|c|}{\textbf{LCA-based topic features}} \\\cline{2-10} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1}\\ \hline
+
+ NB & 0.573 & 0.680 & 0.622 & 0.573 & 0.677 & 0.621 & 0.593 & 0.662 & 0.626 \\ \hline
+ SVM & 0.800 & 0.799 & 0.799 & \textbf{0.848} & \textbf{0.849} & \textbf{0.848} & 0.836 & 0.839 & 0.837  \\ \hline
+ KNN & 0.808 & 0.741 & 0.773 & 0.814 & 0.732 & 0.771 & 0.814 & 0.740 & 0.775 \\ \hline
+  \end{tabular}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\caption{Performance of NB, SVM, KNN, and RNN methods with different topic model's features for identification of ``same segment'' class in e-Coaching emails, when all three features are used together.
+The highest value for each performance metric is highlighted in bold.}
 \label{tab:result_not_boundary}
-  \begin{tabular}{|l|l|l|l|p{0.15\linewidth}|p{0.15\linewidth}|l|}
+  \begin{tabular}{|l|l|l|l|l|l|l|l|l|l|}
   \hline
-   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{lexical features only}} & \multicolumn{3}{|c|}{\textbf{lexical + punctuation (+ topic features except RNN)}} \\\cline{2-7}
-   & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-measure} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-measure}\\ \hline
+   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{L-LDA-based topic features}} & \multicolumn{3}{|c|}{\textbf{DL-LDA-based topic features}}  & \multicolumn{3}{|c|}{\textbf{LCA-based topic features}} \\\cline{2-10} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1}\\ \hline
 
- NB & 0.988 & 0.985 & 0.987 & 0.989 & 0.984 & 0.986 \\ \hline
- SVM & \textbf{0.989} & 0.992 & 0.991 & 0.990 & 0.993 & 0.991\\ \hline
- KNN & \textbf{0.989} & \textbf{0.995} & \textbf{0.992} & 0.991 & \textbf{0.994} & 0.993\\ \hline
- LSTM & -- & -- & -- & 0.993 & \textbf{0.994} & \textbf{0.994} \\ \hline
- GRU & -- & -- & -- & \textbf{0.994} & \textbf{0.994} & \textbf{0.994} \\ \hline 
+ NB & 0.989 & 0.983 & 0.986 & 0.989 & 0.983 & 0.986 & 0.988 & 0.985 & 0.987 \\ \hline
+ SVM & 0.993 & 0.993 & 0.993 & \textbf{0.995} & \textbf{0.995} & \textbf{0.995} & \textbf{0.995} & 0.994 & 0.994  \\ \hline
+ KNN & 0.991 & 0.994 & 0.993 & 0.991 & 0.994 & 0.993 & 0.991 & 0.994 & 0.993 \\ \hline
   \end{tabular}
@@
 KNN achieves 22.40\%, 50.07\%, and 36.26\% higher precision, recall, and F1-measure, respectively, for lexical features; and 20.85\%, 33.96\%, and 27.47\% higher precision, recall, and F1-measure, respectively, for combined features compared to new segment classification.
-In contrast to boundary detection, RNN demonstrates the highest performance among all models.
+In contrast to new segment detection, RNN demonstrates the highest performance among all models.
 LSTM obtains 0.993 precision with 0.994 recall and F1-measure while GRU has 0.994 for all performance metrics.
@@
 \centering
-\caption{Performance of NB, SVM, KNN, and RNN methods for segmentation of e-Coaching emails as a weighted average over ``same segment'' and ``new segment'' classification results.
+\caption{Performance of NB, SVM, KNN, and RNN methods with different topic model's features for segmentation of e-Coaching emails as a weighted average over ``same segment'' and ``new segment'' classification results, when all three features are used together.
 The highest value for each performance metric is highlighted in bold.}
 \label{tab:result_weighted_avg}
-  \begin{tabular}{|l|l|l|l|p{0.15\linewidth}|p{0.15\linewidth}|l|}
+  \begin{tabular}{|l|l|l|l|l|l|l|l|l|l|}
   \hline
-   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{lexical features only}} & \multicolumn{3}{|c|}{\textbf{lexical + punctuation (+ topic features except RNN)}} \\\cline{2-7}
-   & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-measure} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-measure}\\ \hline
+   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{L-LDA-based topic features}} & \multicolumn{3}{|c|}{\textbf{DL-LDA-based topic features}}  & \multicolumn{3}{|c|}{\textbf{LCA-based topic features}} \\\cline{2-10} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1}\\ \hline
 
- NB & 0.975 & 0.974 & 0.975 & 0.976 & 0.974 & 0.975 \\ \hline
- SVM & 0.981 & 0.982 & 0.981 & 0.983 & 0.983 & 0.983\\ \hline
- KNN & \textbf{0.983} & \textbf{0.984} & \textbf{0.983} & \textbf{0.986} & \textbf{0.986} & \textbf{0.986}\\ \hline
- LSTM & -- & -- & -- & \textbf{0.986} & 0.983 & 0.984 \\ \hline
- GRU & -- & -- & -- & \textbf{0.986} & 0.985 & \textbf{0.986} \\ \hline 
+ NB & 0.975 & 0.973 & 0.974 & 0.975 & 0.973 & 0.974 & 0.975 & 0.974 & 0.975 \\ \hline
+ SVM & 0.987 & 0.987 & 0.987 & \textbf{0.990} & \textbf{0.990} & \textbf{0.990} & 0.989 & 0.989 & 0.989 \\ \hline
+ KNN & 0.985 & 0.986 & 0.985 & 0.985 & 0.986 & 0.985 & 0.985 & 0.986 & 0.986 \\ \hline
   \end{tabular}
@@
 This study is the first effort to evaluate the automatic segmentation of e-Coaching emails.
-Experimental results indicate that KNN is the best model among all machine learning methods considered for this study.
-KNN achieved 0.986 F1-measure in overall, 0.779 and 0.993 F1-measures for detecting ``boundary'' and ``not boundary'', respectively.
-The robust performance of KNN provides the evidence that machine learning models are capable to learn conceptual information from clinical exchange.
+Experimental results indicate that SVM is the best model among all machine learning methods considered for this study.
+SVM achieved 0.990 F1-measure in overall, 0.848 and 0.995 F1-measures for detecting ``new segment'' and ``same segment'', respectively.
+The robust performance of SVM provides the evidence that machine learning models are capable to learn conceptual information from clinical exchange.
 It also indicates that topic features are more important than lexical features, even when deep learning methods are employed.
-Although the domain of this study was intentionally quite small, we believe that our study is not limited to the e-coaching domain, and it can be successfully applied to other domains, in which discourse segmentation is an preliminary step for annotation.
+Although the domain of this study was intentionally quite small, we believe that our study is not limited to the e-Coaching domain, and it can be successfully applied to other domains, in which discourse segmentation is an preliminary step for annotation.
 
 Punctuation mark and topic model-based features made a significant improvement in performance of all machine learning methods.
+Among all topic model-based feature, DL-LDA-based topic feature provides the highest performance in ``new segment'' and ``same segment'' classification as well as weighted average over ``new segment'' and ``same segment'' classification results.
 In nearly all cases, ML methods perform better, when lexical features are used in combination with punctuation and topic model-based features.
@@
     \includegraphics[width=1.0\textwidth]{figures/roc-curves.png}
-    \caption{\textbf{Receiver operating characteristic curves showing the performance of binary classifiers for the segmentation of e-coaching text when lexical features (left) and combination of lexical and other features (right) are used}.}
+    \caption{\textbf{Receiver operating characteristic curves showing the performance of binary classifiers for the segmentation of e-Coaching text when lexical features (left) and combination of lexical, DL-LDA-based topic, and punctuation features (right) are used}.}
     \label{fig:roc-curves}
@@
 NB shows the highest AUC values, achieving 0.978 for both cases while provides lowest classification results.
-On the other hand, KNN and SVM exhibit 0.972 and 0.835 AUCs when only lexical features are used; and 0.959 and 0.844 AUCs when a combination of lexical, punctuation mark and topic model-based features are used.
+On the other hand, KNN and SVM exhibit 0.972 and 0.833 AUCs when only lexical features are used; and 0.966 and 0.922 AUCs when a combination of lexical, punctuation mark and topic model-based features are used.
 Finally, LSTM demonstrates lowest AUC values among all machine learning models, achieving 0.82 AUC, while GRU achieves 4.15\% higher AUC, than LSTM.
@@
 
-We observed the second highest performance of RNN, in particular, LSTM and GRU for the text segmentation.
+We observed moderate performance of RNN, in particular, LSTM and GRU for the text segmentation.
 We believe that RNN will perform better if a larger dataset is utilized.
@@
 Similarly, additional information is the common reason for the classified original segment into multiple segments.
-For instance, the above text block can also be incorrectly segmented at position B3 and C4 because third sentence (C) only supports the MI code confirmed by first two sentences (A and B).
+For instance, the above text block can also be incorrectly segmented at position B3 and C4 if third sentence (C) only provides a supportive information of MI code associated with first two sentences (A and B).
 
@@
 \section*{Conclusion}
-Segmentation of e-coaching emails is an integral part of developing e-Coaching interventions.
+Segmentation of e-Coaching emails is an integral part of developing e-Coaching interventions.
 Although several studies have focused on clinical interventions, they are limited by the qualitative coding of clinical interactions.
 In addition, previous studies in the medical domain mainly segmented clinical text into sections and sentences, none of them considered segmentation of text into groups of MI behaviors in the setting of discourse analysis with emails.
-In this paper, we compared the performance of machine learning models for the task of segmentation of e-coaching text.
-We found out that k-nearest neighbor provides the best performance for the segmentation of text in terms of all performance metrics.
-Manual segmentation of e-coaching data is very resource-intensive and time-consuming task, which can significantly decrease the time and effort required to develop an effective behavioral intervention.
+In this paper, we compared the performance of machine learning models for the task of segmentation of e-Coaching text.
+We found out that SVM provides the best performance for the segmentation of text in terms of all performance metrics.
+Manual segmentation of e-Coaching data is very resource-intensive and time-consuming task, which can significantly decrease the time and effort required to develop an effective behavioral intervention.
 Our proposed methods can help to identify individual text segments, which can be annotated directly with a classification model.
-This approach will also help for developing fully automated e-coaching and accelerate the pace of identifying effective communication strategies.
+This approach will also help for developing fully automated e-Coaching and accelerate the pace of identifying effective communication strategies.
 

### 2018-05-23 paper-amia-2018/manuscript.tex
update roc curves

@@
     \centering
-    \includegraphics[width=1.0\textwidth]{figures/roc-curves.png}
+    \includegraphics[width=1.0\textwidth]{figures/roc-curves.eps}
     \caption{\textbf{Receiver operating characteristic curves showing the performance of binary classifiers for the segmentation of e-Coaching text when lexical features (left) and combination of lexical, DL-LDA-based topic, and punctuation features (right) are used}.}
@@
 AUC is also utilized due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets\cite{hu2015kernelized}, which are demonstrated by the ROC curves in Figure~\ref{fig:roc-curves}.
-NB shows the highest AUC values, achieving 0.978 for both cases while provides lowest classification results.
-On the other hand, KNN and SVM exhibit 0.972 and 0.833 AUCs when only lexical features are used; and 0.966 and 0.922 AUCs when a combination of lexical, punctuation mark and topic model-based features are used.
+Although NB provides lowest classification results, it shows the highest AUC values, achieving 0.978 AUC when only lexical features are used and 0.977 AUC when combination of lexical, punctuation and topic features are used.
+On the other hand, KNN and SVM exhibit 0.972 and 0.833 AUCs when only lexical features are used; and 0.966 and 0.922 AUCs when a combination of lexical, punctuation and topic model-based features are used.
 Finally, LSTM demonstrates lowest AUC values among all machine learning models, achieving 0.82 AUC, while GRU achieves 4.15\% higher AUC, than LSTM.

### 2018-05-23 paper-amia-2018/manuscript.tex
update results

@@
 \section*{Results}
-
-Experimental results are reported with respect to ``new segment'' and ``same segment'' classes as well as their weighted average in Tables~\ref{tab:result_boundary}, \ref{tab:result_not_boundary}, and \ref{tab:result_weighted_avg}, respectively.\\
+Our experimental results have two important directions.
+First, results are reported with respect to ``new segment'' and ``same segment'' classes as well as their weighted average in Table~\ref{tab:result_base}.
+Second, classification performance of different machine learning methods are outlined in Tables~\ref{tab:result_boundary}, \ref{tab:result_not_boundary}, and \ref{tab:result_weighted_avg} when topic features are extracted with different probabilistic generative latent variable models.\\
 
@@
 \centering
-\caption{Performance of NB, SVM, and KNN methods for identification of ``new segment'' and ``same segment'' classes as well as segmentation of e-Coaching emails as a weighted average over ``same segment'' and ``new segment'' in e-Coaching emails.
+\caption{Performance of NB, SVM, and KNN methods for identification of ``new segment'' and ``same segment'' classes as well as their weighted average.
 The highest value for each performance metric is highlighted in bold.}
@@
 
-As follows from Table~\ref{tab:result_boundary}, KNN outperforms all other models in terms of precision and F1-measure achieving 0.808 precision with 0.728 F1-measure when lexical features are used; 0.820 precision and 0.779 F1-measure when a combination of lexical, punctuation and topic features are used.
+As follows from Table~\ref{tab:result_base}, KNN outperforms all other models in terms of precision and F1-measure achieving 0.808 precision with 0.728 F1-measure for new segment detection.
+KNN also shows superior performance in all performance metrics for ``same segment'' class and wighted average over ``new segment'' and ``same segment'' classes.
 NB demonstrates the lowest performance among all models in terms of all performance metrics.
-GRU has the second best results showing 10.68\% higher recall and 3.78\%  higher F1-measure than LSTM.
-On the other hand, SVM has the highest recall of 0.679 when only lexical features are used.
-When lexical features are used in combination with punctuation mark and topic model-based features, recall increases by 0.6\%, 2.5\%, and 11.92\%; and F1-measure increases by 0\%, 3.39\%, and 7\% for NB, SVM, and KNN models, respectively, while precision increases by 4.31\% and 1.49\% for SVM and KNN methods and decreases by 0.7\% for NB.
+On the other hand, SVM has the highest recall of 0.673 when only lexical features are used to identify ``new segment''.
+Results indicate that performance of classifiers is remarkably higher for ``same segment'' class compared to ``new segment'' class, which is expected, since 96.71\% instances belong to ``same segment'' categories.
+For example, KNN achieves 22.40\%, 50.07\%, and 36.26\% higher precision, recall, and F1-measure, respectively, in same segment identification compared to new segment detection.
+When lexical features are used in combination with punctuation mark, GRU demonstrates the best performance in all cases showing 10.68\% higher recall and 3.78\%  higher F1-measure than LSTM for new segment identification, 0.1\% higher precision than LSTM for same segment identification and 0.2\% higher recall and 0.2\%  higher F1-measure than LSTM for overall classification.
+\\
+
+%When lexical features are used in combination with punctuation mark and topic model-based features, recall increases by 0.6\%, 2.5\%, and 11.92\%; and F1-measure increases by 0\%, 3.39\%, and 7\% for NB, SVM, and KNN models, respectively, while precision increases by 4.31\% and 1.49\% for SVM and KNN methods and decreases by 0.7\% for NB.
 \\
@@
 \centering
-\caption{Performance of NB, SVM, and KNN methods with different topic model's features for identification of ``new segment'' class in e-Coaching emails, when all three features are used together.
+\caption{Performance of NB, SVM, and KNN methods with different topic model-based features for identification of ``new segment'' class in e-Coaching emails, when all features are used together.
 The highest value for each performance metric is highlighted in bold.}
@@
 
+Table~\ref{tab:result_boundary} summarizes the performance of NB, SVM, KNN, and RNN models for detecting ``new segment'' class.
+We observed that performance is remarkably better for ``new segment'' class compared to new segment detection, which is expected, since 96.71\% instances belong to ``same segment'' class.
+GRU demonstrates the same result as KNN for precision and F1-measure, but not for recall.
+SVM shows moderate performance, achieving 0.981, 0.982, and 0.983 for precision, recall, and F1-measure, respectively, when lexical features are used.
+\\
+
 \begin{table}[ht]
 \centering
-\caption{Performance of NB, SVM, KNN, and RNN methods with different topic model's features for identification of ``same segment'' class in e-Coaching emails, when all three features are used together.
+\caption{Performance of NB, SVM, and KNN methods with different topic model-based features for identification of ``same segment'' class in e-Coaching emails, when all features are used together.
 The highest value for each performance metric is highlighted in bold.}
@@
 
-Table~\ref{tab:result_not_boundary} summarizes the performance of NB, SVM, KNN, and RNN models for detecting ``same segment'' class.
-We observed that performance is remarkably better for ``same segment'' class compared to new segment detection, which is expected, since 96.71\% instances belong to ``same segment'' class.
+As shown in Table~\ref{tab:result_not_boundary}, performance of classifiers is remarkably higher for ``same segment'' class compared to new segment class, which is expected, since 96.71\% instances belong to ``same segment'' categories.
 KNN achieves 22.40\%, 50.07\%, and 36.26\% higher precision, recall, and F1-measure, respectively, for lexical features; and 20.85\%, 33.96\%, and 27.47\% higher precision, recall, and F1-measure, respectively, for combined features compared to new segment classification.
 In contrast to new segment detection, RNN demonstrates the highest performance among all models.
-LSTM obtains 0.993 precision with 0.994 recall and F1-measure while GRU has 0.994 for all performance metrics.
 The impact of punctuation and topic model-based features is consistent with ``new segment'' classification.
@@
 \centering
-\caption{Performance of NB, SVM, KNN, and RNN methods with different topic model's features for segmentation of e-Coaching emails as a weighted average over ``same segment'' and ``new segment'' classification results, when all three features are used together.
+\caption{Performance of NB, SVM, and KNN methods with different topic model-based features for segmentation of e-Coaching emails as a weighted average over ``same segment'' and ``new segment'' classification results, when all features are used together.
 The highest value for each performance metric is highlighted in bold.}
@@
 Table~\ref{tab:result_weighted_avg} summarizes the weighted average results of the models for segmentation of e-Coaching emails.
-Overall, KNN has the best performance across all metrics and NB has the lowest performance among all methods.
-GRU demonstrates the same result as KNN for precision and F1-measure, but not for recall.
-SVM shows moderate performance, achieving 0.981, 0.982, and 0.983 for precision, recall, and F1-measure, respectively, when lexical features are used.
+Overall, SVM has the best performance across all topic model-based features in terms of all performance metrics, achieves 0.990 precision, recall and F1-measure when DL-LDA-based topic features are used.
+Similar to results in Tables~\ref{tab:result_base}, \ref{tab:result_boundary} and \ref{tab:result_not_boundary}, NB demonstrates the lowest performance among all methods while shows best performance with LCA-based topic features compared to other model-based topic features.
 Influence of the additional features is also consistent with the results in Tables~\ref{tab:result_boundary} and~\ref{tab:result_not_boundary}.
-Precision increases by 0.1\%, 0.2\%, and 0.3\%; recall increases by 0\%, 0.1\%, and 0.2\%; and F1-measure increases by 0\%, 0.2\%, and 0.3\% for NB, SVM, and KNN methods, respectively, when feature combination is used.\\
+Precision increases by 0\%, 1.43\%, and 0.2\%; recall increases by 0\%, 1.32\%, and 0.2\%; and F1-measure increases by 0\%, 1.32\%, and 0.3\% for NB, SVM, and KNN methods, respectively, when lexical feature is used in combination with punctuation and best model-based topic features.\\
 

### 2018-05-24 paper-amia-2018/manuscript.tex
update result section

@@
 For example, KNN achieves 22.40\%, 50.07\%, and 36.26\% higher precision, recall, and F1-measure, respectively, in same segment identification compared to new segment detection.
-When lexical features are used in combination with punctuation mark, GRU demonstrates the best performance in all cases showing 10.68\% higher recall and 3.78\%  higher F1-measure than LSTM for new segment identification, 0.1\% higher precision than LSTM for same segment identification and 0.2\% higher recall and 0.2\%  higher F1-measure than LSTM for overall classification.
+When lexical features are used in combination with punctuation mark, GRU demonstrates the best performance in all cases showing 10.68\% higher recall and 3.78\%  higher F1-measure than LSTM for new segment identification, 0.1\% higher precision than LSTM for same segment identification, and 0.2\% higher recall and 0.2\%  higher F1-measure than LSTM for overall classification.
 \\
@@
 
-Table~\ref{tab:result_boundary} summarizes the performance of NB, SVM, KNN, and RNN models for detecting ``new segment'' class.
-We observed that performance is remarkably better for ``new segment'' class compared to new segment detection, which is expected, since 96.71\% instances belong to ``same segment'' class.
-GRU demonstrates the same result as KNN for precision and F1-measure, but not for recall.
-SVM shows moderate performance, achieving 0.981, 0.982, and 0.983 for precision, recall, and F1-measure, respectively, when lexical features are used.
-\\
+Table~\ref{tab:result_boundary} summarizes the performance of NB, SVM and KNN models for detecting ``new segment'' class.
+We observed that classification performance significantly improved for ``new segment'' class when lexical features are used in combination with punctuation and model-based topic features.
+Precision increases by 4.99\% and 0\%, 11.29\% and 0.74\%, and 9.71\% and 0.74\% for SVM and KNN; recall increases by 18.72\% and 17.06\%, 26.15\% and 15.64\%, and 24.67\% and 16.90\% for SVM and KNN; and F1-measure increases by 11.75\% and 6.18\%, 18.60\% and 5.91\%, and 17.06\% and 6.46\% for SVM and KNN when L-LDA, DL-LDA and LCA-based topic features are used in conjunction combination with lexical and punctuation features, respectively.
+On the other hand, NB has very negligable impact of punctuation and model-based topic features.
+In contrast to results in Table~\ref{tab:result_base}, SVM demonstrates the best performance among all models achieving 0.848, 0.849, and 0.848 for precision, recall, and F1-measure, respectively, when DL-LDA-based topic fetaures are used in conjunction with lexical and punctuation features.
+NB and KNN provide their best performance when LCA-based topic features are used, achieving 0.626 and 0.775 F1-measure, respectively.\\
 
@@
 
-As shown in Table~\ref{tab:result_not_boundary}, performance of classifiers is remarkably higher for ``same segment'' class compared to new segment class, which is expected, since 96.71\% instances belong to ``same segment'' categories.
-KNN achieves 22.40\%, 50.07\%, and 36.26\% higher precision, recall, and F1-measure, respectively, for lexical features; and 20.85\%, 33.96\%, and 27.47\% higher precision, recall, and F1-measure, respectively, for combined features compared to new segment classification.
-In contrast to new segment detection, RNN demonstrates the highest performance among all models.
-The impact of punctuation and topic model-based features is consistent with ``new segment'' classification.
-Results indicate that F1-measure increases by 0\%, and 0.1\% for SVM and KNN models and decreases by 0.1\% for NB.
+Similar to baseline performance, classification performance is remarkably higher for same segment class compared to new segment class (Table~\ref{tab:result_not_boundary}).
+The impact of punctuation and topic model-based features is also consistent with ``new segment'' classification while improvements are not as significant as ``new segment'' class.
+We observed that precision increases by 4.99\% and 0\%, 11.29\% and 0.74\%, and 9.71\% and 0.74\% for SVM and KNN; recall increases by 18.72\% and 17.06\%, 26.15\% and 15.64\%, and 24.67\% and 16.90\% for SVM and KNN; and F1-measure increases by 11.75\% and 6.18\%, 18.60\% and 5.91\%, and 17.06\% and 6.46\% for SVM and KNN when L-LDA, DL-LDA and LCA-based topic features are used in conjunction combination with lexical and punctuation features, respectively.
+Similar as Table \ref{tab:result_boundary}, NB and KNN demonstrate the lowest and moderate performance on the task of identifying same segment, respectively.
 \\

### 2018-05-24 paper-amia-2018/manuscript.tex
update manuscript

--- 2018-05-23 paper-amia-2018/manuscript.tex
@@
 
-\author{Mehedi Hasan, BS$^{1}$\footnote[1]{Authors provided equal contribution.
+\author{Mehedi Hasan, BS$^{1}$\footnote[1]{Authors provided an equal contribution.
 \label{footnote1}}, Alexander Kotov, PhD$^{1}$\textsuperscript{\ref{footnote1}}, Sylvie Naar, PhD$^{2}$, Gwen L.
@@
 We experimented both with traditional machine learning methods, such as Support Vector Machine (SVM), Naive Bayes and K-Nearest Neighbor (KNN) classifiers and recurrent neural networks (RNNs).
-Results indicate that SVM outperformed KNN and RNNs achieving 0.990 macro F1-score overall, and 0.848 and 0.995 macro F1-score for detecting ``new segment'' and ``same segment'' classes, respectively.}
+Results indicate that SVM outperformed KNN and RNNs achieving 0.990 macro F1-score overall and 0.848 and 0.995 macro F1-score for detecting ``new segment'' and ``same segment'' classes, respectively.}
 
@@
 Specifically, we demonstrated that machine learning methods can be utilized for annotation of MI transcripts according to a simple communication code scheme with the accuracy comparable to human coders\cite{hasan2016study}.
-Experimental data utilized in these studies, however, were prepared by transcribing audio conversations, which were clearly segmented into utterances by a counselor, a patient, and, in some cases, a caregiver.
+Experimental data utilized in these studies, however, were prepared by transcribing audio conversations, which were clearly segmented into utterances by a counselor, a patient and a caregiver.
 
@@
 The goal of this research study is to assess the applicability of machine learning methods for automated segmentation of e-Coaching emails into textual fragments corresponding to individual behaviors, which is the first step of the coding process of e-Coaching communications.
-In particular, we introduced lexical, topic and punctuation features and experimented with both traditional supervised machine learning methods, such as Support Vector Machine (SVM), Naive Bayes (NB) and K-Nearest Neighbor (KNN) classifiers, and deep learning methods, such as Long Short Term Memory (LSTM) and Gated Recurrent Unit (GRU), to find the best performing method and feature combination.
+In particular, we introduced lexical, topic and punctuation features and experimented with both traditional supervised machine learning methods, such as Support Vector Machine (SVM), Naive Bayes (NB) and K-Nearest Neighbor (KNN) classifiers and deep learning methods, such as Long Short Term Memory (LSTM) and Gated Recurrent Unit (GRU), to find the best performing method and feature combination.
 
@@
 \cite{tepper2012statistical} trained Maximum Entropy models for section classification.
-Denny et al.\cite{denny2009evaluation} proposed a SecTag algorithm, which combined natural language processing techniques, terminology-based rules, and Naive Bayes classifier to identify the sections and headers that achieved 99\% recall with 95.6\% precision.
+Denny et al.\cite{denny2009evaluation} proposed a SecTag algorithm, which combined natural language processing techniques, terminology-based rules and Naive Bayes classifier to identify the sections and headers that achieved 99\% recall with 95.6\% precision.
 On the other hand, SVM based on prosodic and part of speech features \cite{kreuzthaler2015detection} and recurrent convolutional neural networks using word embeddings \cite{griffis2016quantitative} were utilized for detecting sentence boundaries.
-Segmentation of e-Coaching emails is different from traditional shallow discourse analysis of conversations \cite{galley2003discourse} in that the focus is on segmentation, rather than on determining the types of transitions between the utterances or assigning utterances to speakers.
+Segmentation of e-Coaching emails is different from a traditional shallow discourse analysis of conversations \cite{galley2003discourse} in that the focus is on segmentation, rather than on determining the types of transitions between the utterances or assigning utterances to speakers.
 
@@
 Each session represents an MI intervention delivered via email.
-During pre-processing, we removed all non-ASCII characters and applied stemming to normalize morphological variants of related concepts, such as ``eating'', ``eats'', and ``eat''.
+During pre-processing, we removed all non-ASCII characters and applied stemming to normalize morphological variants of related concepts, such as ``eating'', ``eats'' and ``eat''.
 We formulate the segmentation task as a binary classification problem, as illustrated in Figure~\ref{fig:classifier} and experimented both with traditional machine learning methods, such as Support Vector Machine (SVM), Naive Bayes (NB) and K-Nearest Neighbor (KNN) classifiers and recurrent neural networks (RNNs), such as Long Short Term Memory (LSTM) and Gated Recurrent Unit (GRU).
@@
 On the other hand, LCA directly provides label-specific multinomials $p(w|l)$, which can be used to derive topic feature by using Eq.~\ref{eq:plw}.
-Punctuation marks, which correspond to one of the symbols \{`.', `,', `!', `?', `:', `;', `-'\} between a pair of words, are also employed as a feature, since punctuation marks designate the boundary of a sentence, clause, and phrase and often also correspond to segment boundary.
+Punctuation marks, which correspond to one of the symbols \{`.', `,', `!', `?', `:', `;', `-'\} between a pair of words, are also employed as a feature since punctuation marks designate the boundary of a sentence, clause and phrase and often also correspond to segment boundary.
 
@@
 The activation $h_t$ of the GRU at time $t$ is a linear combination of previous activation $h_{t-1}$ and the candidate activation $\tilde h_t$, which is represented by Eq.~(\ref{eq:lastgru}) and~(\ref{eq:candidategru}).
-Our GRU model includes one hidden layer, output layer, and input layer.
+Our GRU model includes one hidden layer, output layer and input layer.
 We reset our model state after feeding each input sequence where input was given as one-hot encoding of word vector.
 Since LSTM and GRU use one-hot vector as input, results for these models are reported only when lexical and punctuation features are used.
-We experimentally determined that the best performance is achieved when the number of hidden units is 25, batch size is 1, and Adam\cite{kingma2014adam} is used for optimization.
+We experimentally determined that the best performance is achieved when the number of hidden units is 25, batch size is 1 and Adam\cite{kingma2014adam} is used for optimization.
 
 \subsection*{\textit{Evaluation metrics}}
-We report standard metrics for experiments (precision, recall, and F1-measure) to evaluate the performance of binary classifiers\cite{aas1999text}.
+We report standard metrics for experiments (precision, recall and F1-measure) to evaluate the performance of binary classifiers\cite{aas1999text}.
 However, accuracy is not reported as a performance metric because accuracy is highly sensitive to the prior class probabilities and does not fully describe the actual difficulty of the decision problem for an unbalanced dataset.
@@
 First, results are reported with respect to ``new segment'' and ``same segment'' classes as well as their weighted average in Table~\ref{tab:result_base}.
-Second, classification performance of different machine learning methods are outlined in Tables~\ref{tab:result_boundary}, \ref{tab:result_not_boundary}, and \ref{tab:result_weighted_avg} when topic features are extracted with different probabilistic generative latent variable models.\\
+Second, classification performance of different machine learning methods are outlined in Tables~\ref{tab:result_boundary}, \ref{tab:result_not_boundary} and \ref{tab:result_weighted_avg} when topic features are extracted with different probabilistic generative latent variable models.\\
 
@@
 \centering
-\caption{Performance of NB, SVM, and KNN methods for identification of ``new segment'' and ``same segment'' classes as well as their weighted average.
+\caption{Performance of NB, SVM and KNN methods for identification of ``new segment'' and ``same segment'' classes as well as their weighted average.
 The highest value for each performance metric is highlighted in bold.}
@@
 As follows from Table~\ref{tab:result_base}, KNN outperforms all other models in terms of precision and F1-measure achieving 0.808 precision with 0.728 F1-measure for new segment detection.
-KNN also shows superior performance in all performance metrics for ``same segment'' class and wighted average over ``new segment'' and ``same segment'' classes.
+KNN also shows superior performance in all performance metrics for ``same segment'' class and weighted average over ``new segment'' and ``same segment'' classes.
 NB demonstrates the lowest performance among all models in terms of all performance metrics.
 On the other hand, SVM has the highest recall of 0.673 when only lexical features are used to identify ``new segment''.
-Results indicate that performance of classifiers is remarkably higher for ``same segment'' class compared to ``new segment'' class, which is expected, since 96.71\% instances belong to ``same segment'' categories.
-For example, KNN achieves 22.40\%, 50.07\%, and 36.26\% higher precision, recall, and F1-measure, respectively, in same segment identification compared to new segment detection.
+Results indicate that performance of classifiers is remarkably higher for ``same segment'' class compared to ``new segment'' class, which is expected since 96.71\% instances belong to ``same segment'' categories.
+For example, KNN achieves 22.40\%, 50.07\% and 36.26\% higher precision, recall and F1-measure, respectively, in same segment identification compared to new segment detection.
 When lexical features are used in combination with punctuation mark, GRU demonstrates the best performance in all cases showing 10.68\% higher recall and 3.78\%  higher F1-measure than LSTM for new segment identification, 0.1\% higher precision than LSTM for same segment identification and 0.2\% higher recall and 0.2\%  higher F1-measure than LSTM for overall classification.
@@
 \centering
-\caption{Performance of NB, SVM, and KNN methods with different topic model-based features for identification of ``new segment'' class in e-Coaching emails, when all features are used together.
+\caption{Performance of NB, SVM and KNN methods with different topic model-based features for identification of ``new segment'' class in e-Coaching emails, when all features are used together.
 The highest value for each performance metric is highlighted in bold.}
@@
 
-Table~\ref{tab:result_boundary} summarizes the performance of NB, SVM, KNN, and RNN models for detecting ``new segment'' class.
-We observed that performance is remarkably better for ``new segment'' class compared to new segment detection, which is expected, since 96.71\% instances belong to ``same segment'' class.
-GRU demonstrates the same result as KNN for precision and F1-measure, but not for recall.
-SVM shows moderate performance, achieving 0.981, 0.982, and 0.983 for precision, recall, and F1-measure, respectively, when lexical features are used.
-\\
+Table~\ref{tab:result_boundary} summarizes the performance of NB, SVM and KNN models for detecting ``new segment'' class.
+We observed that classification performance significantly improved for ``new segment'' class when lexical features are used in combination with punctuation and model-based topic features.
+Precision increases by 4.99\% and 0\%, 11.29\% and 0.74\% and 9.71\% and 0.74\% for SVM and KNN; recall increases by 18.72\% and 17.06\%, 26.15\% and 15.64\% and 24.67\% and 16.90\% for SVM and KNN; and F1-measure increases by 11.75\% and 6.18\%, 18.60\% and 5.91\% and 17.06\% and 6.46\% for SVM and KNN when L-LDA, DL-LDA and LCA-based topic features are used in conjunction combination with lexical and punctuation features, respectively.
+On the other hand, NB has very negligible impact of punctuation and model-based topic features.
+In contrast to results in Table~\ref{tab:result_base}, SVM demonstrates the best performance among all models achieving 0.848, 0.849 and 0.848 for precision, recall and F1-measure, respectively, when DL-LDA-based topic features are used in conjunction with lexical and punctuation features.
+NB and KNN provide their best performance when LCA-based topic features are used, achieving 0.626 and 0.775 F1-measure, respectively.\\
 
@@
 \centering
-\caption{Performance of NB, SVM, and KNN methods with different topic model-based features for identification of ``same segment'' class in e-Coaching emails, when all features are used together.
+\caption{Performance of NB, SVM and KNN methods with different topic model-based features for identification of ``same segment'' class in e-Coaching emails, when all features are used together.
 The highest value for each performance metric is highlighted in bold.}
@@
 
-As shown in Table~\ref{tab:result_not_boundary}, performance of classifiers is remarkably higher for ``same segment'' class compared to new segment class, which is expected, since 96.71\% instances belong to ``same segment'' categories.
-KNN achieves 22.40\%, 50.07\%, and 36.26\% higher precision, recall, and F1-measure, respectively, for lexical features; and 20.85\%, 33.96\%, and 27.47\% higher precision, recall, and F1-measure, respectively, for combined features compared to new segment classification.
-In contrast to new segment detection, RNN demonstrates the highest performance among all models.
-The impact of punctuation and topic model-based features is consistent with ``new segment'' classification.
-Results indicate that F1-measure increases by 0\%, and 0.1\% for SVM and KNN models and decreases by 0.1\% for NB.
+Similar to baseline performance, classification performance is remarkably higher for same segment class compared to new segment class (Table~\ref{tab:result_not_boundary}).
+The impact of punctuation and topic model-based features is also consistent with ``new segment'' classification while improvements are not as significant as ``new segment'' class.
+We observed that precision increases by 0.4\% and 0.2\%, 0.6\% and 0.2\% and 0.6\% and 0.2\% for SVM and KNN; recall increases by 0\% and -0.1\%, 0.2\% and -0.1\% and 0.1\% and -0.1\% for SVM and KNN; and F1-measure increases by 0.2\% and 0.1\%, 0.4\% and 0.1\% and 0.3\% and 0.1\% for SVM and KNN when L-LDA, DL-LDA and LCA-based topic features are used in conjunction combination with lexical and punctuation features, respectively.
+Similar to results in Table \ref{tab:result_boundary}, NB and KNN demonstrate the lowest and moderate performance on the task of identifying same segment, respectively.
 \\
@@
 \centering
-\caption{Performance of NB, SVM, and KNN methods with different topic model-based features for segmentation of e-Coaching emails as a weighted average over ``same segment'' and ``new segment'' classification results, when all features are used together.
+\caption{Performance of NB, SVM and KNN methods with different topic model-based features for segmentation of e-Coaching emails as a weighted average over ``same segment'' and ``new segment'' classification results, when all features are used together.
 The highest value for each performance metric is highlighted in bold.}
@@
 Influence of the additional features is also consistent with the results in Tables~\ref{tab:result_boundary} and~\ref{tab:result_not_boundary}.
-Precision increases by 0\%, 1.43\%, and 0.2\%; recall increases by 0\%, 1.32\%, and 0.2\%; and F1-measure increases by 0\%, 1.32\%, and 0.3\% for NB, SVM, and KNN methods, respectively, when lexical feature is used in combination with punctuation and best model-based topic features.\\
+Precision increases by 0.9\% and 0.2\%; recall increases by 0.8\% and 0.2\%; and F1-measure increases by 0.8\% and 0.3\% for SVM and KNN methods, respectively, when lexical feature is used in combination with punctuation and best model-based topic features.\\
 
@@
 It also indicates that topic features are more important than lexical features, even when deep learning methods are employed.
-Although the domain of this study was intentionally quite small, we believe that our study is not limited to the e-Coaching domain, and it can be successfully applied to other domains, in which discourse segmentation is an preliminary step for annotation.
+Although the domain of this study was intentionally quite small, we believe that our study is not limited to the e-Coaching domain and it can be successfully applied to other domains, in which discourse segmentation is a preliminary step for annotation.
 
 Punctuation mark and topic model-based features made a significant improvement in performance of all machine learning methods.
-Among all topic model-based feature, DL-LDA-based topic feature provides the highest performance in ``new segment'' and ``same segment'' classification as well as weighted average over ``new segment'' and ``same segment'' classification results.
+Among all topic model-based feature, the DL-LDA-based topic feature provides the highest performance in ``new segment'' and ``same segment'' classification as well as weighted average over ``new segment'' and ``same segment'' classification results.
 In nearly all cases, ML methods perform better, when lexical features are used in combination with punctuation and topic model-based features.
@@
     \includegraphics[width=1.0\textwidth]{figures/roc-curves.eps}
-    \caption{\textbf{Receiver operating characteristic curves showing the performance of binary classifiers for the segmentation of e-Coaching text when lexical features (left) and combination of lexical, DL-LDA-based topic, and punctuation features (right) are used}.}
+    \caption{\textbf{Receiver operating characteristic curves showing the performance of binary classifiers for the segmentation of e-Coaching text when lexical features (left) and combination of lexical, DL-LDA-based topic (except LSTM and GRU) and punctuation features (right) are used}.}
     \label{fig:roc-curves}
@@
 In this paper, results are reported by each class to avoid confusion regarding overall model performance due to severe class imbalance.
-In addition, standard metrics: precision, recall, and F1-measure were used to eliminate doubt about the model performance, since accuracy is misleading for imbalanced datasets.
+In addition, standard metrics: precision, recall and F1-measure were used to eliminate doubt about the model performance since accuracy is misleading for imbalanced datasets.
 AUC is also utilized due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets\cite{hu2015kernelized}, which are demonstrated by the ROC curves in Figure~\ref{fig:roc-curves}.
-Although NB provides lowest classification results, it shows the highest AUC values, achieving 0.978 AUC when only lexical features are used and 0.977 AUC when combination of lexical, punctuation and topic features are used.
+Although NB provides lowest classification results, it shows the highest AUC values, achieving 0.978 AUC when only lexical features are used and 0.977 AUC when a combination of lexical, punctuation and topic features are used.
 On the other hand, KNN and SVM exhibit 0.972 and 0.833 AUCs when only lexical features are used; and 0.966 and 0.922 AUCs when a combination of lexical, punctuation and topic model-based features are used.
-Finally, LSTM demonstrates lowest AUC values among all machine learning models, achieving 0.82 AUC, while GRU achieves 4.15\% higher AUC, than LSTM.
+Finally, LSTM demonstrates lowest AUC values among all machine learning models, achieving 0.82 AUC, while GRU achieves 4.15\% higher AUC than LSTM.
 The conclusion drawn from the ROC curves also confirmed the robustness and superiority of KNN model for the segmentation of clinical exchange.
@@
 
-Although punctuation mark plays an important role in segmentation boundary detection, and large numbers of errors were encountered by the false positive of boundary identification.
+Although punctuation mark plays an important role in segmentation boundary detection and large numbers of errors were encountered by the false positive of boundary identification.
 For example, a text block ``A1 A2 A3.
 B1 B2 B3.
-C1 C2 C3 C4.'' can be incorrectly segmented at position A3, B3, and C4 where a punctuation mark was encountered.
+C1 C2 C3 C4.'' can be incorrectly segmented at position A3, B3 and C4 where a punctuation mark was encountered.
 Similarly, additional information is the common reason for the classified original segment into multiple segments.
-For instance, the above text block can also be incorrectly segmented at position B3 and C4 if third sentence (C) only provides a supportive information of MI code associated with first two sentences (A and B).
+For instance, the above text block can also be incorrectly segmented at position B3 and C4 if the third sentence (C) only provides a supportive information of MI code associated with first two sentences (A and B).
 
-Our proposed approach is novel for the segmentation of e-Coaching emails, since previous studies mainly focused on the segmentation of text into sections, headers, and sentences.
+Our proposed approach is novel for the segmentation of e-Coaching emails since previous studies mainly focused on the segmentation of text into sections, headers and sentences.
 However, this study segmented clinical exchange into groups of MI behaviors which will significantly reduce the amount of resource and time required to segment clinical exchange manually.
@@
 
-The limitation of this study is that e-Coaching data is collected from a single medical institute; formatting, style, and email segment can be different in other settings.
+The limitation of this study is that e-Coaching data is collected from a single medical institute; formatting, style and email segment can be different in other settings.
 Therefore, there is a need to replicate the experiments with different data sets.
-As our future work, we plan to evaluate our approach on other datasets for discourse analysis.
+As our future work, we plan to evaluate our approach to other datasets for discourse analysis.
 

### 2018-05-24 paper-amia-2018/manuscript.tex
update manuscript

@@
 \textbf{Recurrent Neural Networks}: RNN is a neural network architecture designed to capture sequential patterns.
-When we predict the ``new segment'' point using RNNs, particular combinations of words in a sequence will indicate the change of MI behavior.
+When we predict the new segment point using RNNs, particular combinations of words in a sequence will indicate the change of MI behavior.
 Long Short Term Memory (LSTM) networks\cite{hochreiter1997long} are a special type of RNN that are capable of handling variable size input sequence and have an internal memory that can be reset.
@@
 Results indicate that performance of classifiers is remarkably higher for ``same segment'' class compared to ``new segment'' class, which is expected since 96.71\% instances belong to ``same segment'' categories.
-For example, KNN achieves 22.40\%, 50.07\% and 36.26\% higher precision, recall and F1-measure, respectively, in same segment identification compared to new segment detection.
-When lexical features are used in combination with punctuation mark, GRU demonstrates the best performance in all cases showing 10.68\% higher recall and 3.78\%  higher F1-measure than LSTM for new segment identification, 0.1\% higher precision than LSTM for same segment identification and 0.2\% higher recall and 0.2\%  higher F1-measure than LSTM for overall classification.
+For example, KNN achieves 22.40\%, 50.07\% and 36.26\% higher precision, recall and F1-measure, respectively, in ``same segment'' identification compared to ``new segment'' detection.
+When lexical features are used in combination with punctuation mark, GRU demonstrates the best performance in all cases showing 10.68\% higher recall and 3.78\%  higher F1-measure than LSTM for ``new segment'' identification, 0.1\% higher precision than LSTM for ``same segment'' identification and 0.2\% higher recall and 0.2\%  higher F1-measure than LSTM for overall classification.
 \\
@@
 
-Similar to baseline performance, classification performance is remarkably higher for same segment class compared to new segment class (Table~\ref{tab:result_not_boundary}).
+Similar to baseline performance, classification performance is remarkably higher for ``same segment'' class compared to ``new segment'' class (Table~\ref{tab:result_not_boundary}).
 The impact of punctuation and topic model-based features is also consistent with ``new segment'' classification while improvements are not as significant as ``new segment'' class.
 We observed that precision increases by 0.4\% and 0.2\%, 0.6\% and 0.2\% and 0.6\% and 0.2\% for SVM and KNN; recall increases by 0\% and -0.1\%, 0.2\% and -0.1\% and 0.1\% and -0.1\% for SVM and KNN; and F1-measure increases by 0.2\% and 0.1\%, 0.4\% and 0.1\% and 0.3\% and 0.1\% for SVM and KNN when L-LDA, DL-LDA and LCA-based topic features are used in conjunction combination with lexical and punctuation features, respectively.
-Similar to results in Table \ref{tab:result_boundary}, NB and KNN demonstrate the lowest and moderate performance on the task of identifying same segment, respectively.
+Similar to results in Table \ref{tab:result_boundary}, NB and KNN demonstrate the lowest and moderate performance on the task of identifying ``same segment'', respectively.
 \\

### 2018-05-26 paper-amia-2018/manuscript.tex
update manuscript

@@
 Although punctuation mark plays an important role in segmentation boundary detection and large numbers of errors were encountered by the false positive of boundary identification.
-For example, a text block ``A1 A2 A3.
-B1 B2 B3.
-C1 C2 C3 C4.'' can be incorrectly segmented at position A3, B3 and C4 where a punctuation mark was encountered.
-Similarly, additional information is the common reason for the classified original segment into multiple segments.
-For instance, the above text block can also be incorrectly segmented at position B3 and C4 if the third sentence (C) only provides a supportive information of MI code associated with first two sentences (A and B).
+For example, a text segment from an e-Coaching email \textit{``A typical day in regards to fruit and vegetable has me eating about a serving at breakfast (our caf has cut up fruit) and then maybe a piece of fruit later in the day or as a snack.
+Vegetable tends to be a side serving at lunch and dinner and I get celery or carrot cuts with dressing for a snack a lot of times.
+I could probably add some sort of vegetable into my breakfast (like spinach in an omelet) and snack on another piece of fruit when I quotemark m hungry rather than the junk food I tend to eat.''} incorrectly segmented after the first sentence where a punctuation mark was encountered.
+Similarly, additional information is a common cause for misclassification of an email segment into multiple segments.
+For instance, the first sentence of the above email segment represents a positive commitment to adolescent's behavior change then next two sentences provide an additional information to support their commitment.
 

### 2018-05-26 paper-amia-2018/manuscript.tex
update ROC plots

@@
  LSTM & \textbf{0.800} & 0.646 & 0.714  & 0.993 & \textbf{0.994} & \textbf{0.994} & \textbf{0.986} & 0.983 & 0.984 \\ \hline
- GRU & \textbf{0.800} & \textbf{0.715} & \textbf{0.741} & \textbf{0.994} & \textbf{0.994} & \textbf{0.994} & \textbf{0.986} & \textbf{0.985} & \textbf{0.986} \\ \hline
+ GRU & \textbf{0.769} & \textbf{0.715} & \textbf{0.741} & \textbf{0.994} & \textbf{0.994} & \textbf{0.994} & \textbf{0.986} & \textbf{0.985} & \textbf{0.986} \\ \hline
   \end{tabular}
@@
 For example, KNN achieves 22.40\%, 50.07\% and 36.26\% higher precision, recall and F1-measure, respectively, in ``same segment'' identification compared to ``new segment'' detection.
-When lexical features are used in combination with punctuation mark, GRU demonstrates the best performance in all cases showing 10.68\% higher recall and 3.78\%  higher F1-measure than LSTM for ``new segment'' identification, 0.1\% higher precision than LSTM for ``same segment'' identification and 0.2\% higher recall and 0.2\%  higher F1-measure than LSTM for overall classification.
+When lexical features are used in combination with punctuation mark, GRU demonstrates the best performance in all cases except precision for ``new segment'' identification where LSTM shows 4.03\% higher precision than GRU.
+In contrast, GRU shows 10.68\% higher recall and 3.78\%  higher F1-measure than LSTM for ``new segment'' identification, 0.1\% higher precision than LSTM for ``same segment'' identification and 0.2\% higher recall and 0.2\%  higher F1-measure than LSTM for overall classification.
 \\
@@
 AUC is also utilized due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets\cite{hu2015kernelized}, which are demonstrated by the ROC curves in Figure~\ref{fig:roc-curves}.
-Although NB provides lowest classification results, it shows the highest AUC values, achieving 0.978 AUC when only lexical features are used and 0.977 AUC when a combination of lexical, punctuation and topic features are used.
+Although NB provides lowest classification results, it shows the second highest AUC values, achieving 0.978 AUC when only lexical features are used and 0.977 AUC when a combination of lexical, punctuation and topic features are used.
 On the other hand, KNN and SVM exhibit 0.972 and 0.833 AUCs when only lexical features are used; and 0.966 and 0.922 AUCs when a combination of lexical, punctuation and topic model-based features are used.
-Finally, LSTM demonstrates lowest AUC values among all machine learning models, achieving 0.82 AUC, while GRU achieves 4.15\% higher AUC than LSTM.
-The conclusion drawn from the ROC curves also confirmed the robustness and superiority of KNN model for the segmentation of clinical exchange.
+Finally, LSTM demonstrates the highest AUC value among all machine learning models, achieving 0.998 AUC, while GRU achieves only 0.2\% lower AUC than LSTM.
+
+%The conclusion drawn from the ROC curves also confirmed the robustness and superiority of KNN model for the segmentation of clinical exchange.
 

### 2018-05-26 results/roc/roc-curves-generator.R

@@ -16,16 +16,16 @@ plot(r1$fpr, r1$tpr, col='red', type="l", ylab = "Sensitivity", xlab = "1-Specif
 lines(r2$fpr, r2$tpr, col='blue', lty=1, lwd=1)
 lines(r3$fpr, r3$tpr, col='green', lty=1, lwd=1)
 lines(0:1, 0:1, col='black', lty=5, lwd=1)
-legend(0.75,0.36, c("KNN", "SVM", "NB"), 
+legend(0.75,0.33, c("KNN", "SVM", "NB"), 
        lty=c(1,1,1), col=c('red', 'blue', 'green'), cex=1, lwd=1)
 
 # print AUC values
-x1 <- read.csv("/home/mehedi/teana/projects/eCoaching-Text-Segmentation/results/roc/True/auc_RNN-LSTMTrue-ROC.csv")
-x2 <- read.csv("/home/mehedi/teana/projects/eCoaching-Text-Segmentation/results/roc/True/auc_RNN-GRUTrue-ROC.csv")
-roc_obj_lstm <- roc(x1$actual, x1$prediction)
+x1 <- read.csv("/home/mehedi/teana/projects/eCoaching-Text-Segmentation/results/roc/True/roc_LSTM_summit.csv")
+x2 <- read.csv("/home/mehedi/teana/projects/eCoaching-Text-Segmentation/results/roc/True/roc_GRU_summit.csv")
+roc_obj_lstm <- roc(x1$actual, x1$predict_prob)
 auc(roc_obj_lstm)
 
-roc_obj_gru <- roc(x2$actual, x2$prediction)
+roc_obj_gru <- roc(x2$actual, x2$predict_prob)
 auc(roc_obj_gru)
 
 # melt data
@@ -40,6 +40,6 @@ lines(r3$fpr, r3$tpr, col='green', lty=1, lwd=1)
 lines(1-roc_obj_lstm$specificities, roc_obj_lstm$sensitivities, col='purple', lty=1, lwd=1)
 lines(1-roc_obj_gru$specificities, roc_obj_gru$sensitivities, col='pink', lty=1, lwd=1)
 lines(0:1, 0:1, col='black', lty=5, lwd=1)
-legend(0.73,0.53, c("KNN", "SVM", "NB", "LSTM", "GRU"), 
+legend(0.73,0.47, c("KNN", "SVM", "NB", "LSTM", "GRU"), 
[…]

### 2018-05-29 paper-amia-2018/manuscript.tex
update ROC figures

@@
 
-%When lexical features are used in combination with punctuation mark and topic model-based features, recall increases by 0.6\%, 2.5\%, and 11.92\%; and F1-measure increases by 0\%, 3.39\%, and 7\% for NB, SVM, and KNN models, respectively, while precision increases by 4.31\% and 1.49\% for SVM and KNN methods and decreases by 0.7\% for NB.
-\\
-
 \begin{table}[ht]
@@
 On the other hand, KNN and SVM exhibit 0.972 and 0.833 AUCs when only lexical features are used; and 0.966 and 0.922 AUCs when a combination of lexical, punctuation and topic model-based features are used.
-Finally, LSTM demonstrates the highest AUC value among all machine learning models, achieving 0.998 AUC, while GRU achieves only 0.2\% lower AUC than LSTM.
-
-%The conclusion drawn from the ROC curves also confirmed the robustness and superiority of KNN model for the segmentation of clinical exchange.
+Finally, LSTM demonstrates the highest AUC value among all machine learning models, achieving 0.996 AUC, while GRU achieves only 0.4\% lower AUC than LSTM.
 
@@
 
-%We also plan to use more relevant features to improve model performance.
-For example, part-of-speech tagging\cite{hasan2016feedback} and distance from %the beginning and end of the current sentence might significantly enhance the classification performance.
-
 \section*{Conclusion}

### 2018-05-29 results/roc/preprocess-roc-data.py

@@ -8,7 +8,7 @@ import numpy
 def create_roc_csv(path):
     for file_name in os.listdir(path):
         file_name_writer = path + "/" + file_name.replace(".arff", "") + ".csv"
-        if 'RNN' not in file_name and 'csv' not in file_name:
+        if 'summit' not in file_name and 'csv' not in file_name:
             fw = open(file_name_writer, "w")
             fw.write("fpr,tpr\n")
             with open(path + "/" + file_name, "r") as file_stream:
@@ -18,40 +18,6 @@ def create_roc_csv(path):
                     current_line = line.split(",")
                     fw.write(current_line[5] + "," + current_line[6] + "\n")
             fw.close()
-
-
-# convert rnn results to roc.csv for figure
-def create_roc_csv_for_rnn(path):
-    for file_name in os.listdir(path):
-        if 'RNN' in file_name and 'csv' not in file_name:
-            actual = []
-            predicted = []
-            count = 1
-            with open(path + "/" + file_name, "r") as file_stream:
-                for line in file_stream:
-                    if count == 1:
-                        current_line = line.split(",")
-                        actual += current_line
-                        count = 0
-                    else:
-                        current_line = line.split(",")
-                        predicted += current_line
-                        count = 1
-
-            sorted_actual, sorted_predicted = [], []
-            for i in numpy.fliplr([numpy.argsort(predicted)])[0]:
-                sorted_actual.append(actual[i])
-                sorted_predicted.append(predicted[i])
-
-            # write prediction and actual values
-            auc_file_name_writer = path + "/auc_" + file_name + ".csv"
-            file_w = open(auc_file_name_writer, "w")
-            file_w.write("actual,prediction\n")
-            for i, j in zip(sorted_actual, sorted_predicted):
-                if len(i.strip()) > 0:
-                    file_w.write(i + "," + j + "\n")
-            file_w.close()
-
 
 # parse command line arguments
 parser = argparse.ArgumentParser(description='Prepare data for ROC curve.')
@@ -64,5 +30,4 @@ folder_path = args.path
 
 # call above functions for the creation of new files
 create_roc_csv(folder_path)
-create_roc_csv_for_rnn(folder_path)
 print('conversion is done!')

### 2018-05-29 results/roc/roc-curves-generator.R

@@ -16,7 +16,7 @@ plot(r1$fpr, r1$tpr, col='red', type="l", ylab = "Sensitivity", xlab = "1-Specif
 lines(r2$fpr, r2$tpr, col='blue', lty=1, lwd=1)
 lines(r3$fpr, r3$tpr, col='green', lty=1, lwd=1)
 lines(0:1, 0:1, col='black', lty=5, lwd=1)
-legend(0.75,0.33, c("KNN", "SVM", "NB"), 
+legend(0.79,0.30, c("KNN", "SVM", "NB"), 
        lty=c(1,1,1), col=c('red', 'blue', 'green'), cex=1, lwd=1)
 
 # print AUC values
@@ -37,9 +37,9 @@ r3 <- read.csv("/home/mehedi/teana/projects/eCoaching-Text-Segmentation/results/
 plot(r1$fpr, r1$tpr, col='red', type="l", ylab = "Sensitivity", xlab = "1-Specificity", cex.axis=1, cex.lab=1, lwd=1)
 lines(r2$fpr, r2$tpr, col='blue', lty=1, lwd=1)
 lines(r3$fpr, r3$tpr, col='green', lty=1, lwd=1)
-lines(1-roc_obj_lstm$specificities, roc_obj_lstm$sensitivities, col='purple', lty=1, lwd=1)
-lines(1-roc_obj_gru$specificities, roc_obj_gru$sensitivities, col='pink', lty=1, lwd=1)
+lines(1-roc_obj_lstm$specificities, roc_obj_lstm$sensitivities, col='cyan', lty=1, lwd=1)
+lines(1-roc_obj_gru$specificities, roc_obj_gru$sensitivities, col='magenta', lty=1, lwd=1)
 lines(0:1, 0:1, col='black', lty=5, lwd=1)
-legend(0.73,0.47, c("KNN", "SVM", "NB", "LSTM", "GRU"), 
-       lty=c(1,1,1,1,1), col=c('red', 'blue', 'green', 'purple', 'pink'), cex=1, lwd=1)
+legend(0.77,0.44, c("KNN", "SVM", "NB", "LSTM", "GRU"), 
+       lty=c(1,1,1,1,1), col=c('red', 'blue', 'green', 'cyan', 'magenta'), cex=1, lwd=1)
 

### 2018-05-29 rnn/gru_sequence.py

@@ -1,45 +1,0 @@
-from random import random
-from numpy import array
-from numpy import cumsum
-from keras.models import Sequential
-from keras.layers import LSTM
-from keras.layers import Dense
-from keras.layers import TimeDistributed
-
-
-# create a sequence classification instance
-def get_sequence(n_timesteps):
-    # create a sequence of random numbers in [0,1]
-    X = array([random() for _ in range(n_timesteps)])
-    # calculate cut-off value to change class values
-    limit = n_timesteps / 4.0
-    # determine the class outcome for each item in cumulative sequence
-    y = array([0 if x < limit else 1 for x in cumsum(X)])
-    # reshape input and output data to be suitable for LSTMs
-    X = X.reshape(1, n_timesteps, 1)
-    y = y.reshape(1, n_timesteps, 1)
-    return X, y
-
-
-# define problem properties
-n_timesteps = 10
-# define LSTM
-model = Sequential()
-model.add(LSTM(20, input_shape=(n_timesteps, 1), return_sequences=True))
-model.add(TimeDistributed(Dense(1, activation='sigmoid')))
-model.compile(loss='binary_crossentropy', optimizer='adam', metrics=['acc'])
-# train LSTM
-for epoch in range(1000):
-    # generate new random sequence
-    X, y = get_sequence(n_timesteps)
-    # fit model for one epoch on this sequence
-    model.fit(X, y, epochs=1, batch_size=1, verbose=2)
-# evaluate LSTM
-X, y = get_sequence(n_timesteps)
-
-print(X)
-print(y)
-
-yhat = model.predict_classes(X, verbose=0)
-for i in range(n_timesteps):
-    print('Expected:', y[0, i], 'Predicted', yhat[0, i])

### 2018-05-29 rnn/seg-gru-model-bidirect.py

@@ -1,83 +1,0 @@
-# RNN with variable length input sequences to one word output
-import numpy
-
-from keras.models import Sequential
-from keras.layers import Dense
-from keras.layers import GRU
-from keras.layers import LSTM
-from keras.layers import Bidirectional
-from keras.layers.embeddings import Embedding
-from keras.callbacks import EarlyStopping, ModelCheckpoint
-from keras.regularizers import l1_l2
-
-import utility
-
-from vocabulary import Vocabulary
-
-# fix random seed for reproducibility
-numpy.random.seed(42)
-
-# create e-coaching dictionary
-vocabulary_obj = Vocabulary()
-vocabulary_obj.create()
-
-# create mapping of characters to integers (0-25) and the reverse
-char_to_int = dict((c, i+1) for i, c in enumerate(vocabulary_obj.vocab))
-int_to_char = dict((i+1, c) for i, c in enumerate(vocabulary_obj.vocab))
-
-# system configuration
-max_len = 100
-folds = 5
-embed_dim = 30
-lstm_out = 32
-dropout_x = 0.5
-batch_size = 64
-
-# prepare the data set of input to output pairs encoded as integers
-data = utility.get_ecoaching_data()
-folds_data = utility.split_data(data, folds)
-results = []
-
-for fold_num in range(folds):
-
-    train_data, test_data = [], []
-    for i in range(folds):
-        if i == fold_num:
-            test_data = folds_data[i]
-        else:
-            train_data = train_data + folds_data[i]
-
-    # get encoded training data
-    X, y = utility.get_encoded_long_sequence(train_data, vocabulary_obj, max_len)
-
-    # print total number of instances
-    print("# of instances: ", len(y))
-
-    # create and fit the model
-    model = Sequential()
-    model.add(Embedding(vocabulary_obj.size(), embed_dim, input_length=X.shape[1]))
-    # model.add(GRU(lstm_out, recurrent_regularizer=l1_l2(l1=0.0, l2=0.033), dropout=dropout_x))
-    # model.add(LSTM(lstm_out, input_shape=(X.shape[1], 1)))
-    model.add(Bidirectional(GRU(lstm_out)))
-    model.add(Dense(y.shape[1], activation='softmax'))
-    model.compile(loss='categorical_crossentropy', optimizer='adam', metrics=['accuracy'])
[…]

### 2018-05-29 rnn/seg-gru-model-embed.py

@@ -1,86 +1,0 @@
-# RNN with variable length input sequences to one word output
-import numpy
-
-from keras.models import Sequential
-from keras.layers import Dense
-from keras.layers import GRU
-from keras.layers import LSTM
-from keras.layers.embeddings import Embedding
-from keras.callbacks import EarlyStopping, ModelCheckpoint
-from keras.regularizers import l1_l2
-
-import utility
-
-from vocabulary import Vocabulary
-
-# fix random seed for reproducibility
-numpy.random.seed(42)
-
-# create e-coaching dictionary
-vocabulary_obj = Vocabulary()
-vocabulary_obj.create()
-
-# create mapping of characters to integers (0-25) and the reverse
-char_to_int = dict((c, i+1) for i, c in enumerate(vocabulary_obj.vocab))
-int_to_char = dict((i+1, c) for i, c in enumerate(vocabulary_obj.vocab))
-
-# system configuration
-folds = 5
-embed_dim = 30
-lstm_out = 32
-dropout_x = 0.5
-batch_size = 64
-max_len = 3
-
-# prepare the data set of input to output pairs encoded as integers
-data = utility.get_ecoaching_data()
-folds_data = utility.split_data(data, folds)
-results = []
-
-# print maximum length of the sequence
-print("maximum length of the sequence: ", max_len)
-
-for fold_num in range(folds):
-
-    train_data, test_data = [], []
-    for i in range(folds):
-        if i == fold_num:
-            test_data = folds_data[i]
-        else:
-            train_data = train_data + folds_data[i]
-
-    # get encoded training data
-    X, y = utility.get_encoded_data_rnn_embed(train_data, vocabulary_obj, max_len)
-
-    # print total number of instances
-    print("# of instances: ", len(y))
-
-    # create and fit the model
-    model = Sequential()
-    model.add(Embedding(vocabulary_obj.size(), embed_dim, input_length=X.shape[1]))
-    # model.add(GRU(lstm_out, recurrent_regularizer=l1_l2(l1=0.0, l2=0.033), dropout=dropout_x))
-    # model.add(LSTM(lstm_out, input_shape=(X.shape[1], 1)))
-    # model.add(GRU(lstm_out, input_shape=(X.shape[1], 1)))
-    model.add(GRU(lstm_out))
-    model.add(Dense(y.shape[1], activation='softmax'))
[…]

### 2018-05-29 rnn/seg-gru-model-long.py

@@ -1,86 +1,0 @@
-# RNN with variable length input sequences to one word output
-import numpy
-
-from keras.models import Sequential
-from keras.layers import Dense
-from keras.layers import GRU
-from keras.layers import LSTM
-from keras.layers.embeddings import Embedding
-from keras.callbacks import EarlyStopping, ModelCheckpoint
-from keras.regularizers import l1_l2
-
-import utility
-
-from vocabulary import Vocabulary
-
-# fix random seed for reproducibility
-numpy.random.seed(42)
-
-# create e-coaching dictionary
-vocabulary_obj = Vocabulary()
-vocabulary_obj.create()
-
-# create mapping of characters to integers (0-25) and the reverse
-char_to_int = dict((c, i+1) for i, c in enumerate(vocabulary_obj.vocab))
-int_to_char = dict((i+1, c) for i, c in enumerate(vocabulary_obj.vocab))
-
-# system configuration
-folds = 5
-embed_dim = 30
-lstm_out = 32
-dropout_x = 0.5
-batch_size = 64
-max_len = 3
-
-# prepare the data set of input to output pairs encoded as integers
-data = utility.get_ecoaching_data()
-folds_data = utility.split_data(data, folds)
-results = []
-
-# print maximum length of the sequence
-print("maximum length of the sequence: ", max_len)
-
-for fold_num in range(folds):
-
-    train_data, test_data = [], []
-    for i in range(folds):
-        if i == fold_num:
-            test_data = folds_data[i]
-        else:
-            train_data = train_data + folds_data[i]
-
-    # get encoded training data
-    X, y = utility.get_encoded_data_rnn(train_data, vocabulary_obj, max_len)
-
-    # print total number of instances
-    print("# of instances: ", len(y))
-
-    # create and fit the model
-    model = Sequential()
-    # model.add(Embedding(vocabulary_obj.size(), embed_dim, input_length=X.shape[1]))
-    # model.add(GRU(lstm_out, recurrent_regularizer=l1_l2(l1=0.0, l2=0.033), dropout=dropout_x))
-    # model.add(LSTM(lstm_out, input_shape=(X.shape[1], 1)))
-    model.add(GRU(lstm_out, input_shape=(X.shape[1], 1)))
-    # model.add(GRU(lstm_out))
-    model.add(Dense(y.shape[1], activation='softmax'))
[…]

### 2018-05-29 rnn/seg-gru-model-stepwise.py

@@ -1,4 +1,3 @@
-# RNN with variable length input sequences to one word output
 import numpy
 
 from random import random
@@ -14,9 +13,6 @@ from keras.callbacks import EarlyStopping, ModelCheckpoint, Callback
 import utility_stepwise
 
 from vocabulary import Vocabulary
-
-# fix random seed for reproducibility
-numpy.random.seed(42)
 
 # system configuration
 max_len = 50
@@ -40,7 +36,7 @@ class ResetStatesCallback(Callback):
 vocabulary_obj = Vocabulary()
 vocabulary_obj.create()
 
-# create mapping of characters to integers (0-25) and the reverse
+# create mapping of words to integers (0-25) and the reverse
 char_to_int = dict((c, i+1) for i, c in enumerate(vocabulary_obj.vocab))
 int_to_char = dict((i+1, c) for i, c in enumerate(vocabulary_obj.vocab))
 

### 2018-05-29 rnn/seg-gru-model.py

@@ -1,82 +1,0 @@
-# RNN with variable length input sequences to one word output
-import numpy
-
-from keras.models import Sequential
-from keras.layers import Dense
-from keras.layers import GRU
-from keras.layers import LSTM
-from keras.layers.embeddings import Embedding
-from keras.callbacks import EarlyStopping, ModelCheckpoint
-from keras.regularizers import l1_l2
-
-import utility
-
-from vocabulary import Vocabulary
-
-# fix random seed for reproducibility
-numpy.random.seed(42)
-
-# create e-coaching dictionary
-vocabulary_obj = Vocabulary()
-vocabulary_obj.create()
-
-# create mapping of characters to integers (0-25) and the reverse
-char_to_int = dict((c, i+1) for i, c in enumerate(vocabulary_obj.vocab))
-int_to_char = dict((i+1, c) for i, c in enumerate(vocabulary_obj.vocab))
-
-# system configuration
-max_len = 3
-folds = 5
-embed_dim = 300
-lstm_out = 32
-dropout_x = 0.5
-batch_size = 8
-
-# prepare the data set of input to output pairs encoded as integers
-data = utility.get_ecoaching_data()
-folds_data = utility.split_data(data, folds)
-results = []
-
-for fold_num in range(folds):
-
-    train_data, test_data = [], []
-    for i in range(folds):
-        if i == fold_num:
-            test_data = folds_data[i]
-        else:
-            train_data = train_data + folds_data[i]
-
-    # get encoded training data
-    X, y = utility.get_encoded_data_rnn(train_data, vocabulary_obj, max_len)
-
-    # print total number of instances
-    print("# of instances: ", len(y))
-
-    # create and fit the model
-    model = Sequential()
-    # model.add(Embedding(vocabulary_obj.size()+1, embed_dim, input_length=X.shape[1]))
-    # model.add(GRU(lstm_out, recurrent_regularizer=l1_l2(l1=0.0, l2=0.033), dropout=dropout_x))
-    # model.add(LSTM(lstm_out, input_shape=(X.shape[1], 1)))
-    model.add(GRU(lstm_out, input_shape=(X.shape[1], 1)))
-    model.add(Dense(y.shape[1], activation='softmax'))
-    model.compile(loss='categorical_crossentropy', optimizer='adam', metrics=['accuracy'])
-    # early_stop = EarlyStopping(monitor='val_acc', min_delta=0.00005, patience=20, verbose=1, mode='auto')
[…]

### 2018-05-29 rnn/utility.py

@@ -1,359 +1,0 @@
-# include required packages
-import numpy
-import re
-import sys
-
-from keras.utils import np_utils
-from keras.preprocessing.sequence import pad_sequences
-
-
-# load e-coaching data from datafile
-def get_ecoaching_data(file_name='e-coaching_data_stem.txt'):
-    data = []
-    with open(file_name, "r") as file_stream:
-        for line in file_stream:
-            current_line = line.split(",")
-            if len(current_line) > 1:
-                data.append(clean_text(current_line[0]).strip())
-    return data
-
-
-# split data into 5 folds
-def split_data(data, folds=5):
-    fold_size = len(data)/folds
-    folds_data = []
-    for i in range(folds):
-        start, end = i*fold_size, i*fold_size + fold_size
-        folds_data.append(data[start: end])
-    return folds_data
-
-
-# encode text data into numeric values
-def get_encoded_data(data, vocabulary_obj, max_len):
-    x, y = [], []
-
-    for text in data:
-        input_text = vocabulary_obj.vectorize(text)
-        sample_len = len(input_text)
-
-        if sample_len <= max_len:
-            x.append(input_text)
-            y.append(1)
-        else:
-            for idx in range(sample_len-max_len, 0, -max_len):
-                if idx == sample_len-max_len:
-                    x.append(input_text[idx:sample_len])
-                    y.append(1)
-                else:
-                    x.append(input_text[idx:idx+max_len])
-                    y.append(0)
-
-                if idx - max_len < 0:
-                    x.append(input_text[0:idx])
-                    y.append(0)
-
-    data_x = numpy.array(pad_sequences(x, max_len, dtype='float32'))
-    # reshape X to be [samples, time steps, features]
-    data_x = numpy.reshape(data_x, (data_x.shape[0], max_len, 1))
-    # normalize
-    data_x = data_x / vocabulary_obj.size()
-    data_y = np_utils.to_categorical(y)
-    return data_x, data_y
-
-
-# encode text data into numeric values
-def get_encoded_long_sequence(data, vocabulary_obj, max_len):
-    x, y = [], []
-
-    for text in data:
-        input_text = vocabulary_obj.vectorize(text)
-        sample_len = len(input_text)
-        idx = 2
-
-        while idx < sample_len:
-            if idx == sample_len-1:
-                x.append(input_text[0:sample_len])
-                y.append(1)
-
[…]

### 2018-05-29 rnn/var_seq_length/utility_var_sequence.py

@@ -1,122 +1,0 @@
-# include required packages
-import numpy
-import re
-import sys
-
-from keras.utils import np_utils
-from keras.preprocessing.sequence import pad_sequences
-
-
-# load e-coaching data from datafile
-def get_ecoaching_data(file_name='e-coaching_data_stem.txt'):
-    data = []
-    with open(file_name, "r") as file_stream:
-        for line in file_stream:
-            current_line = line.split(",")
-            if len(current_line) > 1:
-                data.append(clean_text(current_line[0]).strip())
-    return data
-
-
-# split data into 5 folds
-def split_data(data, folds=5):
-    fold_size = len(data)/folds
-    folds_data = []
-    for i in range(folds):
-        start, end = i*fold_size, i*fold_size + fold_size
-        folds_data.append(data[start: end])
-    return folds_data
-
-
-# encode text data into numeric values
-def get_encoded_sequence(data, vocabulary_obj):
-    x, y = [], []
-
-    for text in data:
-        sample_x = vocabulary_obj.vectorize(text)
-        sample_y = [0]*(len(sample_x)-1)
-        sample_y.append(1)
-
-        sample_x = [a / float(vocabulary_obj.size()) for a in sample_x]
-        x.append(sample_x)
-        y.append(sample_y)
-
-    return x, y
-
-
-# clean text, removed weired chars
-def clean_text(text):
-    new_text = re.sub(r'[^\x00-\x7f]', r'', text)
-    new_text = new_text.replace("\n", "").strip()
-    new_text = re.sub(r'[?|`|>|<|$|.|!|;|:|=|&|\'|\"|\-|/|)|(|\]|[|]', r'', new_text)
-    return new_text
-
-
-# evaluate model with F1, Precision and Recall
-def get_macro_average_performance(actual, predicted):
-    f = open("/home/mehedi/Dropbox/results_sequence_GRU.txt", "a")
-    f.write("fold number \n")
-    f.write("actual: ")
-    f.writelines(["%s," % item for item in actual])
-    f.write("\npredicted: ")
-    f.writelines(["%s," % item for item in predicted])
-    f.write("\n")
-
-    precision = 0.0
-    recall = 0.0
-    f_measure = 0.0
-    accuracy = 0.0
-
-    labels = [1, 0, 0, 1]
-
-    for k in [0, 2]:
-
-        tp = 0
-        fp = 0
-        tn = 0
-        fn = 0
-
-        for i in range(0, len(actual)):
[…]

### 2018-05-29 rnn/var_seq_length/var-seq-model.py

@@ -1,116 +1,0 @@
-# RNN with variable length input sequences to one word output
-import numpy
-
-from random import random
-from numpy import array
-from numpy import cumsum
-from keras.models import Sequential
-from keras.layers import LSTM
-from keras.layers import GRU
-from keras.layers import Dense
-from keras.layers import TimeDistributed
-from keras.callbacks import EarlyStopping, ModelCheckpoint, Callback
-
-import utility_var_sequence
-
-from vocabulary import Vocabulary
-
-# fix random seed for reproducibility
-numpy.random.seed(42)
-
-# system configuration
-folds = 5
-lstm_out = 25
-batch_size = 1
-max_len = 50
-
-
-class ResetStatesCallback(Callback):
-    def __init__(self):
-        self.counter = 0
-
-    def on_batch_begin(self, batch, logs={}):
-        if self.counter % max_len == 0:
-            self.model.reset_states()
-            # print("Reset States..............", max_len)
-        self.counter += 1
-
-# create e-coaching dictionary
-vocabulary_obj = Vocabulary()
-vocabulary_obj.create()
-
-# create mapping of characters to integers (0-25) and the reverse
-char_to_int = dict((c, i+1) for i, c in enumerate(vocabulary_obj.vocab))
-int_to_char = dict((i+1, c) for i, c in enumerate(vocabulary_obj.vocab))
-
-# prepare the data set of input to output pairs encoded as integers
-data = utility_var_sequence.get_ecoaching_data()
-folds_data = utility_var_sequence.split_data(data, folds)
-results = []
-
-for fold_num in range(folds):
-
-    train_data, test_data = [], []
-    for i in range(folds):
-        if i == fold_num:
-            test_data = folds_data[i]
-        else:
-            train_data = train_data + folds_data[i]
-
-    # get encoded training data
-    X, y = utility_var_sequence.get_encoded_sequence(train_data, vocabulary_obj)
-
-    # print total number of instances
-    print("# of instances: ", len(y))
-
-    # create and fit the model
-    model = Sequential()
-    model.add(LSTM(lstm_out, input_shape=(None, 1), return_sequences=True))
-    model.add(TimeDistributed(Dense(1, activation='sigmoid')))
-    model.compile(loss='binary_crossentropy', optimizer='adam', metrics=['accuracy'])
-
-    for i in range(len(X)):
[…]

### 2018-05-29 rnn/var_seq_length/vocabulary.py

@@ -1,59 +1,0 @@
-import re
-import codecs
-import operator
-
-
-class Vocabulary:
-    def __init__(self, vocab_path='e-coaching.vocab'):
-        self.vocab = dict()
-        self.int_to_word = dict()
-        self.vocab_path = vocab_path
-
-    # build e-coaching dictionary
-    def build(self):
-        with codecs.open(self.vocab_path, 'r', 'UTF-8') as train_file:
-            words = [x.strip().rstrip('\n') for x in train_file.readlines()]
-            self.vocab = dict((c, i + 1) for i, c in enumerate(words))
-            self.int_to_word = dict((i+1, c) for i, c in enumerate(words))
-        return self
-
-    # read all the data from file and create vocabulary, also find # classes
-    def create(self, data_file='e-coaching_data_stem.txt'):
-        word_dict = {}
-
-        with open(data_file, "r") as file_stream:
-            for line in file_stream:
-                current_line = line.strip().split(",")
-                if len(current_line) < 2:
-                    continue
-
-                words = re.split('(\W+)', self.clean_text(current_line[0]).strip())
-                for word in words:
-                    if len(word.strip()) > 0:
-                        word_dict[word] = 1
-
-        sorted_map = sorted(word_dict.items(), key=operator.itemgetter(0))
-        file_pointer = open(self.vocab_path, "w")
-        for key, val in sorted_map:
-            file_pointer.write(key + "\n")
-        file_pointer.close()
-
-        self.build()
-        return self
-
-    def size(self):
-        return len(self.vocab)
-
-    def tokenize(self, text):
-        return [x.strip() for x in re.split('(\W+)', text) if x.strip()]
-
-    def vectorize(self, text):
-        text = text.lower()
-        words = filter(lambda x: x in self.vocab, self.tokenize(text))
-        return [self.vocab[w] for w in words]
-
-    def clean_text(self, text):
-        new_text = re.sub(r'[^\x00-\x7f]', r'', text)
-        new_text = new_text.replace("\n", "").strip()
-        new_text = re.sub(r'[?|`|>|<|$|.|!|;|:|=|&|\'|\"|\-|/|)|(|\]|[|]', r'', new_text)
-        return new_text

### 2018-05-30 paper-amia-2018/manuscript.tex
update manuscript
update figure 1 and results for RNNs

@@
 We experimentally determined that the best performance is achieved when the number of hidden units is 25, batch size is 1 and Adam\cite{kingma2014adam} is used for optimization.
+Source codes of this work will be available at https://github.com/teanalab/xxxxx upon acceptance of this paper for publication.
 
@@
  \multicolumn{10}{|c|}{lexical and puctuation features} \\ \hline
- LSTM & \textbf{0.800} & 0.646 & 0.714  & 0.993 & \textbf{0.994} & \textbf{0.994} & \textbf{0.986} & 0.983 & 0.984 \\ \hline
- GRU & \textbf{0.769} & \textbf{0.715} & \textbf{0.741} & \textbf{0.994} & \textbf{0.994} & \textbf{0.994} & \textbf{0.986} & \textbf{0.985} & \textbf{0.986} \\ \hline
+ LSTM & \textbf{0.810} & 0.720 & \textbf{0.762}  & \textbf{0.994} & \textbf{0.994} & \textbf{0.994} & \textbf{0.988} & \textbf{0.985} & \textbf{0.987} \\ \hline
+ GRU & 0.783 & \textbf{0.739} & 0.760 & \textbf{0.994} & \textbf{0.994} & \textbf{0.994} & \textbf{0.988} & \textbf{0.985} & \textbf{0.987} \\ \hline
   \end{tabular}
@@
 For example, KNN achieves 22.40\%, 50.07\% and 36.26\% higher precision, recall and F1-measure, respectively, in ``same segment'' identification compared to ``new segment'' detection.
-When lexical features are used in combination with punctuation mark, GRU demonstrates the best performance in all cases except precision for ``new segment'' identification where LSTM shows 4.03\% higher precision than GRU.
-In contrast, GRU shows 10.68\% higher recall and 3.78\%  higher F1-measure than LSTM for ``new segment'' identification, 0.1\% higher precision than LSTM for ``same segment'' identification and 0.2\% higher recall and 0.2\%  higher F1-measure than LSTM for overall classification.
+When lexical features are used in combination with punctuation mark, LSTM demonstrates the best performance in terms of precision and F1-measure for ``new segment'' identification while GRU shows 2.64\% higher recall than LSTM.
+GRU and LSTM show exactly same performance in terms of all performance metrics for ``same segement'' and overall classifications.
 \\
@@
 On the other hand, KNN and SVM exhibit 0.972 and 0.833 AUCs when only lexical features are used; and 0.966 and 0.922 AUCs when a combination of lexical, punctuation and topic model-based features are used.
-Finally, LSTM demonstrates the highest AUC value among all machine learning models, achieving 0.996 AUC, while GRU achieves only 0.4\% lower AUC than LSTM.
+Finally, LSTM demonstrates the highest AUC value among all machine learning models, achieving 0.996 AUC, while GRU achieves only 0.2\% lower AUC than LSTM.
 

### 2018-05-30 results/roc/roc-curves-generator.R

@@ -16,7 +16,7 @@ plot(r1$fpr, r1$tpr, col='red', type="l", ylab = "Sensitivity", xlab = "1-Specif
 lines(r2$fpr, r2$tpr, col='blue', lty=1, lwd=1)
 lines(r3$fpr, r3$tpr, col='green', lty=1, lwd=1)
 lines(0:1, 0:1, col='black', lty=5, lwd=1)
-legend(0.79,0.30, c("KNN", "SVM", "NB"), 
+legend(0.77,0.30, c("KNN", "SVM", "NB"), 
        lty=c(1,1,1), col=c('red', 'blue', 'green'), cex=1, lwd=1)
 
 # print AUC values
@@ -40,6 +40,6 @@ lines(r3$fpr, r3$tpr, col='green', lty=1, lwd=1)
 lines(1-roc_obj_lstm$specificities, roc_obj_lstm$sensitivities, col='cyan', lty=1, lwd=1)
 lines(1-roc_obj_gru$specificities, roc_obj_gru$sensitivities, col='magenta', lty=1, lwd=1)
 lines(0:1, 0:1, col='black', lty=5, lwd=1)
-legend(0.77,0.44, c("KNN", "SVM", "NB", "LSTM", "GRU"), 
+legend(0.75,0.44, c("KNN", "SVM", "NB", "LSTM", "GRU"), 
        lty=c(1,1,1,1,1), col=c('red', 'blue', 'green', 'cyan', 'magenta'), cex=1, lwd=1)
 

### 2018-05-30 rnn/seg-lstm-model-stepwise.py
update source codes for rnn

@@ -6,7 +6,7 @@ from keras.layers import Dense
 from keras.layers import TimeDistributed
 from keras.callbacks import Callback
 
-import utility_stepwise_gru
+import utility_stepwise
 
 from vocabulary import Vocabulary
 
@@ -37,8 +37,8 @@ char_to_int = dict((c, i+1) for i, c in enumerate(vocabulary_obj.vocab))
 int_to_char = dict((i+1, c) for i, c in enumerate(vocabulary_obj.vocab))
 
 # prepare the data set of input to output pairs encoded as integers
-data = utility_stepwise_gru.get_ecoaching_data()
-folds_data = utility_stepwise_gru.split_data(data, folds)
+data = utility_stepwise.get_ecoaching_data()
+folds_data = utility_stepwise.split_data(data, folds)
 results = []
 
 for fold_num in range(folds):
@@ -50,14 +50,14 @@ for fold_num in range(folds):
             train_data = train_data + folds_data[i]
 
     # get encoded training data
-    X, y = utility_stepwise_gru.get_encoded_sequence(train_data, vocabulary_obj, max_len)
+    X, y = utility_stepwise.get_encoded_sequence(train_data, vocabulary_obj, max_len)
 
     # print total number of instances
     print("# of instances: ", len(y))
 
     # create and fit the model
     model = Sequential()
-    model.add(GRU(lstm_out, input_shape=(max_len, 1), return_sequences=True))
+    model.add(LSTM(lstm_out, input_shape=(max_len, 1), return_sequences=True))
     model.add(TimeDistributed(Dense(1, activation='sigmoid')))
     model.compile(loss='binary_crossentropy', optimizer='adam', metrics=['accuracy'])
 
@@ -69,7 +69,7 @@ for fold_num in range(folds):
         model.fit(x_data, y_data, callbacks=[ResetStatesCallback()], shuffle=False, batch_size=batch_size, verbose=0)
 
     # get encoded test data
-    test_x, test_y = utility_stepwise_gru.get_encoded_sequence(test_data, vocabulary_obj, max_len)
+    test_x, test_y = utility_stepwise.get_encoded_sequence(test_data, vocabulary_obj, max_len)
 
     prediction_y = []
     prediction_prob_y = []
@@ -95,19 +95,19 @@ for fold_num in range(folds):
             original_y.append(orig_y[0, j][0])
 
 
-    f = open("results_GRU_summit.txt", "a")
[…]

### 2018-05-30 rnn/utility_stepwise_gru.py

@@ -60,7 +60,7 @@ def clean_text(text):
 
 # evaluate model with F1, Precision and Recall
 def get_macro_average_performance(actual, predicted):
-    f = open("results_LSTM_summit.txt", "a")
+    f = open("results_GRU_summit.txt", "a")
     f.write("fold number \n")
     f.write("actual: ")
     f.writelines(["%s," % item for item in actual])

### 2018-07-19 paper-amia-summit-2018/manuscript.tex
Added files for initial version of amia summit paper

--- 2018-05-30 paper-amia-2018/manuscript.tex
@@
 We experimented both with traditional machine learning methods, such as Support Vector Machine (SVM), Naive Bayes and K-Nearest Neighbor (KNN) classifiers and recurrent neural networks (RNNs).
-Results indicate that SVM outperformed KNN and RNNs achieving 0.990 macro F1-score overall and 0.848 and 0.995 macro F1-score for detecting ``new segment'' and ``same segment'' classes, respectively.}
+Results indicate that SVM outperformed KNN and RNNs achieving 0.990 macro F1-score overall and 0.845 and 0.995 macro F1-score for detecting ``new segment'' and ``same segment'' classes, respectively.}
 
@@
 SVM finds the hyperplane that maximizes the separation between the closest ``new segment'' and ``same segment'' training examples.
-We used L1-regularized linear SVM with squared error minimization in experiments.
+We scaled our dataset to utilize linear SVM with optimal cost 512 in experiments.
 
@@
  NB & 0.594 & 0.662 & 0.626 & 0.988 & 0.985 & 0.987 & 0.975 & 0.974 & 0.975  \\ \hline
- SVM & 0.762 & \textbf{0.673} & 0.715 & \textbf{0.989} & 0.993 & 0.991  & 0.981 & 0.982 & 0.982 \\ \hline
- KNN & \textbf{0.808} & 0.663 & \textbf{0.728} & \textbf{0.989} & \textbf{0.995} & \textbf{0.992} & \textbf{0.983} & \textbf{0.984} & \textbf{0.983} \\ \hline
+ SVM & 0.721 & \textbf{0.705} & 0.713 & \textbf{0.990} & 0.991 & 0.990  & 0.981 & 0.981 & 0.981 \\ \hline
+ KNN & \textbf{0.808} & 0.663 & \textbf{0.728} & 0.989 & \textbf{0.995} & \textbf{0.992} & \textbf{0.983} & \textbf{0.984} & \textbf{0.983} \\ \hline
  \multicolumn{10}{|c|}{lexical and puctuation features} \\ \hline
@@
 NB demonstrates the lowest performance among all models in terms of all performance metrics.
-On the other hand, SVM has the highest recall of 0.673 when only lexical features are used to identify ``new segment''.
+On the other hand, SVM has the highest recall of 0.705 when only lexical features are used to identify ``new segment''.
 Results indicate that performance of classifiers is remarkably higher for ``same segment'' class compared to ``new segment'' class, which is expected since 96.71\% instances belong to ``same segment'' categories.
@@
  NB & 0.573 & 0.680 & 0.622 & 0.573 & 0.677 & 0.621 & 0.593 & 0.662 & 0.626 \\ \hline
- SVM & 0.800 & 0.799 & 0.799 & \textbf{0.848} & \textbf{0.849} & \textbf{0.848} & 0.836 & 0.839 & 0.837  \\ \hline
+ SVM & 0.831 & 0.783 & 0.806 & \textbf{0.877} & \textbf{0.815} & \textbf{0.845} & 0.867 & 0.804 & 0.834  \\ \hline
  KNN & 0.808 & 0.741 & 0.773 & 0.814 & 0.732 & 0.771 & 0.814 & 0.740 & 0.775 \\ \hline
@@
 We observed that classification performance significantly improved for ``new segment'' class when lexical features are used in combination with punctuation and model-based topic features.
-Precision increases by 4.99\% and 0\%, 11.29\% and 0.74\% and 9.71\% and 0.74\% for SVM and KNN; recall increases by 18.72\% and 17.06\%, 26.15\% and 15.64\% and 24.67\% and 16.90\% for SVM and KNN; and F1-measure increases by 11.75\% and 6.18\%, 18.60\% and 5.91\% and 17.06\% and 6.46\% for SVM and KNN when L-LDA, DL-LDA and LCA-based topic features are used in conjunction combination with lexical and punctuation features, respectively.
+Precision increases by 15.26\% and 0\%, 21.64\% and 0.74\% and 20.25\% and 0.74\% for SVM and KNN; recall increases by 11.06\% and 17.06\%, 15.60\% and 15.64\% and 14.04\% and 16.90\% for SVM and KNN; and F1-measure increases by 13.04\% and 6.18\%, 18.51\% and 5.91\% and 16.97\% and 6.46\% for SVM and KNN when L-LDA, DL-LDA and LCA-based topic features are used in conjunction combination with lexical and punctuation features, respectively.
 On the other hand, NB has very negligible impact of punctuation and model-based topic features.
-In contrast to results in Table~\ref{tab:result_base}, SVM demonstrates the best performance among all models achieving 0.848, 0.849 and 0.848 for precision, recall and F1-measure, respectively, when DL-LDA-based topic features are used in conjunction with lexical and punctuation features.
+In contrast to results in Table~\ref{tab:result_base}, SVM demonstrates the best performance among all models achieving 0.877, 0.815 and 0.845 for precision, recall and F1-measure, respectively, when DL-LDA-based topic features are used in conjunction with lexical and punctuation features.
 NB and KNN provide their best performance when LCA-based topic features are used, achieving 0.626 and 0.775 F1-measure, respectively.\\
@@
  NB & 0.989 & 0.983 & 0.986 & 0.989 & 0.983 & 0.986 & 0.988 & 0.985 & 0.987 \\ \hline
- SVM & 0.993 & 0.993 & 0.993 & \textbf{0.995} & \textbf{0.995} & \textbf{0.995} & \textbf{0.995} & 0.994 & 0.994  \\ \hline
+ SVM & 0.993 & 0.995 & 0.994 & \textbf{0.994} & \textbf{0.996} & \textbf{0.995} & 0.993 & \textbf{0.996} & \textbf{0.995} \\ \hline
  KNN & 0.991 & 0.994 & 0.993 & 0.991 & 0.994 & 0.993 & 0.991 & 0.994 & 0.993 \\ \hline
@@
 The impact of punctuation and topic model-based features is also consistent with ``new segment'' classification while improvements are not as significant as ``new segment'' class.
-We observed that precision increases by 0.4\% and 0.2\%, 0.6\% and 0.2\% and 0.6\% and 0.2\% for SVM and KNN; recall increases by 0\% and -0.1\%, 0.2\% and -0.1\% and 0.1\% and -0.1\% for SVM and KNN; and F1-measure increases by 0.2\% and 0.1\%, 0.4\% and 0.1\% and 0.3\% and 0.1\% for SVM and KNN when L-LDA, DL-LDA and LCA-based topic features are used in conjunction combination with lexical and punctuation features, respectively.
+We observed that precision increases by 0.3\% and 0.2\%, 0.4\% and 0.2\% and 0.3\% and 0.2\% for SVM and KNN; recall increases by 0.4\% and -0.1\%, 0.5\% and -0.1\% and 0.5\% and -0.1\% for SVM and KNN; and F1-measure increases by 0.4\% and 0.1\%, 0.5\% and 0.1\% and 0.5\% and 0.1\% for SVM and KNN when L-LDA, DL-LDA and LCA-based topic features are used in conjunction combination with lexical and punctuation features, respectively.
 Similar to results in Table \ref{tab:result_boundary}, NB and KNN demonstrate the lowest and moderate performance on the task of identifying ``same segment'', respectively.
@@
  NB & 0.975 & 0.973 & 0.974 & 0.975 & 0.973 & 0.974 & 0.975 & 0.974 & 0.975 \\ \hline
- SVM & 0.987 & 0.987 & 0.987 & \textbf{0.990} & \textbf{0.990} & \textbf{0.990} & 0.989 & 0.989 & 0.989 \\ \hline
+ SVM & 0.987 & 0.988 & 0.987 & \textbf{0.990} & \textbf{0.990} & \textbf{0.990} & 0.989 & 0.989 & 0.989 \\ \hline
  KNN & 0.985 & 0.986 & 0.985 & 0.985 & 0.986 & 0.985 & 0.985 & 0.986 & 0.986 \\ \hline
@@
 Overall, SVM has the best performance across all topic model-based features in terms of all performance metrics, achieves 0.990 precision, recall and F1-measure when DL-LDA-based topic features are used.
-Similar to results in Tables~\ref{tab:result_base}, \ref{tab:result_boundary} and \ref{tab:result_not_boundary}, NB demonstrates the lowest performance among all methods while shows best performance with LCA-based topic features compared to other model-based topic features.
+Similar to results in Tables~\ref{tab:result_base}, \ref{tab:result_boundary} and \ref{tab:result_not_boundary}, NB demonstrates the lowest performance among all methods while best performance achieves with LCA-based topic features compared to other model-based topic features.
 Influence of the additional features is also consistent with the results in Tables~\ref{tab:result_boundary} and~\ref{tab:result_not_boundary}.
-Precision increases by 0.9\% and 0.2\%; recall increases by 0.8\% and 0.2\%; and F1-measure increases by 0.8\% and 0.3\% for SVM and KNN methods, respectively, when lexical feature is used in combination with punctuation and best model-based topic features.\\
+Precision increases by 0.9\% and 0.2\%; recall increases by 0.9\% and 0.2\%; and F1-measure increases by 0.9\% and 0.3\% for SVM and KNN methods, respectively, when lexical feature is used in combination with punctuation and best model-based topic features.\\
 
@@
 Experimental results indicate that SVM is the best model among all machine learning methods considered for this study.
-SVM achieved 0.990 F1-measure in overall, 0.848 and 0.995 F1-measures for detecting ``new segment'' and ``same segment'', respectively.
+SVM achieved 0.990 F1-measure in overall, 0.845 and 0.995 F1-measures for detecting ``new segment'' and ``same segment'', respectively.
 The robust performance of SVM provides the evidence that machine learning models are capable to learn conceptual information from clinical exchange.
@@
 AUC is also utilized due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets\cite{hu2015kernelized}, which are demonstrated by the ROC curves in Figure~\ref{fig:roc-curves}.
-Although NB provides lowest classification results, it shows the second highest AUC values, achieving 0.978 AUC when only lexical features are used and 0.977 AUC when a combination of lexical, punctuation and topic features are used.
-On the other hand, KNN and SVM exhibit 0.972 and 0.833 AUCs when only lexical features are used; and 0.966 and 0.922 AUCs when a combination of lexical, punctuation and topic model-based features are used.
+Although NB provides lowest classification results, it shows moderate AUC values, achieving 0.978 AUC when only lexical features are used and 0.977 AUC when a combination of lexical, punctuation and topic features are used.
+On the other hand, KNN and SVM exhibit 0.972 and 0.986 AUCs when only lexical features are used; and 0.966 and 0.980 AUCs when a combination of lexical, punctuation and topic model-based features are used.
 Finally, LSTM demonstrates the highest AUC value among all machine learning models, achieving 0.996 AUC, while GRU achieves only 0.2\% lower AUC than LSTM.

### 2018-07-21 paper-amia-summit-2018/manuscript.tex
update manuscript

@@
 \noindent{\bf Abstract}
-\textit{Communication science approaches to developing effective behavior interventions, such as motivational interviewing (MI), are limited by traditionally manual qualitative coding of communication exchanges, which is a very resource-intensive and time-consuming process.
+\textit{Communication science approaches to developing effective behavior interventions, such as motivational interviewing (MI), are limited by the traditionally manual qualitative coding of communication exchanges, which is a very resource-intensive and time-consuming process.
 This study focuses on the analysis of e-Coaching sessions, behavior interventions that are delivered via email and grounded in the principles of MI.
 A critical step towards automated annotation of e-Coaching communication exchanges is segmentation of emails into textual fragments that correspond to MI behaviors.
-In this work, we formulate this task as a classification problem and propose lexical, punctuation and topic features to address it.
-We experimented both with traditional machine learning methods, such as Support Vector Machine (SVM), Naive Bayes and K-Nearest Neighbor (KNN) classifiers and recurrent neural networks (RNNs).
-Results indicate that SVM outperformed KNN and RNNs achieving 0.990 macro F1-score overall and 0.845 and 0.995 macro F1-score for detecting ``new segment'' and ``same segment'' classes, respectively.}
+In this work, we formulate this task as a classification problem and propose lexical, punctuation and semantic features to address it.
+We experimented both with traditional machine learning method, Conditional Random Field (CRF) and deep learning methods, such as Multilayer Perceptrons (MLPs), Bidirectional Recurrent Neural Networks (BRNNs) and Convolutional Recurrent Neural Networks (CRNNs).
+Results indicate that CRNN outperformed CRF, MLP and BRNN achieving 0.990 macro F1-score overall and 0.821 macro F1-score for detecting new segment.}
 
@@
 The goal of this research study is to assess the applicability of machine learning methods for automated segmentation of e-Coaching emails into textual fragments corresponding to individual behaviors, which is the first step of the coding process of e-Coaching communications.
-In particular, we introduced lexical, topic and punctuation features and experimented with both traditional supervised machine learning methods, such as Support Vector Machine (SVM), Naive Bayes (NB) and K-Nearest Neighbor (KNN) classifiers and deep learning methods, such as Long Short Term Memory (LSTM) and Gated Recurrent Unit (GRU), to find the best performing method and feature combination.
+In particular, we introduced lexical, semantic and punctuation features and experimented with both traditional supervised machine learning method, linear-chain Conditional Random Field (CRF) and deep learning methods, such as Multilayer Perceptrons (MLPs), Bidirectional Recurrent Neural Networks (BRNNs) and Convolutional Recurrent Neural Networks (CRNNs), to find the best performing method and feature combination.
 
-Relevant previous work in the biomedical domain primarily focused on segmentation of text into sections and headers\cite{apostolova2009automatic,denny2009evaluation,tepper2012statistical,cho2002text} or sentence boundary detection\cite{griffis2016quantitative,kreuzthaler2015detection,treviso2016sentence}.
+Relevant previous work in the biomedical domain primarily focused on segmentation of text into sections and headers\cite{apostolova2009automatic,denny2009evaluation,tepper2012statistical,cho2002text} or sentence boundary detection\cite{griffis2016quantitative,kreuzthaler2015detection,treviso2016sentence,fraser2015sentence}.
 Apostolova et al.\cite{apostolova2009automatic} applied SVM along with word-vector cosine similarity metric combined with several heuristics to segment clinical reports into sections, such as demographics, history, procedure, finding and impression.
@@
 Each session represents an MI intervention delivered via email.
-During pre-processing, we removed all non-ASCII characters and applied stemming to normalize morphological variants of related concepts, such as ``eating'', ``eats'' and ``eat''.
-We formulate the segmentation task as a binary classification problem, as illustrated in Figure~\ref{fig:classifier} and experimented both with traditional machine learning methods, such as Support Vector Machine (SVM), Naive Bayes (NB) and K-Nearest Neighbor (KNN) classifiers and recurrent neural networks (RNNs), such as Long Short Term Memory (LSTM) and Gated Recurrent Unit (GRU).
-For NB, SVM and KNN models, e-Coaching email exchanges are partitioned into adjacent word pairs.
+We formulate the segmentation task as a binary classification problem where each word annotated with a label 1 or 0, to indicate whether it preceds a new segment or not.
+We experimented both with traditional machine learning method, linear-chain Conditional Random Field (CRF) and deep learning methods, such as Multilayer Perceptrons (MLPs), Bidirectional Recurrent Neural Networks (BRNNs) and Convolutional Recurrent Neural Networks (CRNNs).
+For MLP models, e-Coaching email exchanges are partitioned into adjacent word pairs.
 Each pair is classified into either ``new segment'' or ``same segment'' class.
@@
 In total, we obtained 95,421 word pairs, which include 3,138 ``new segment'' and 92,283 ``same segment'' instances.
-For RNNs, a block of text was taken as input sequence, such that one-hot encodings of each word or punctuation marks in a block were used as input into an RNN and binary labels (1 or 0) corresponding to ``new segment'' and ``same segment'' classification decision were considered as the output of RNN at each step.
+For deep learning models, an email was taken as input sequence, such that PoS tags and word embedding of each word or punctuation marks were used as input and binary labels (1 or 0) corresponding to ``new segment'' and ``same segment'' classification decision were considered as the output of the model at each step.
 In the gold standard, words within the same segment were assigned the label of 0 and the last word or punctuation mark of a segment were assigned the label of 1.
-
-\begin{figure}[!htb]
-    \centering
-    \includegraphics[width=0.80\textwidth]{figures/classifier.pdf}
-    \caption{\textbf{Transformation of text segmentation task into text classification task}.}
-    \label{fig:classifier}
-\end{figure}
 
@@
 
-We utilized three types of features in conjunction with SVM, NB and KNN methods: lexical, punctuation and topic features.
+We utilized three types of features in conjunction with CRF, MLP, BRNN and CRNN methods: lexical, punctuation and topic features.
 Lexical features represent each word in a pair as a binary vector, in which 1 corresponds to the word in question and 0 to all other words.
-Since topic models\cite{kotov2015interpretable,hashimoto2016topic,lu2016modeling} have been shown to be effective semantic abstractions of individual words, we derived topic features from Labeled LDA\cite{ramage2009labeled}, a topic model for annotated corpora.
-We considered each textual segment in the training set as a document and the MI code assigned to the segment as a label.
-First, we derived a label-specific multinomial $p(w|l)$ for each label $l$ from the word distributions $p(w|l,z)$ for topic $z$ specific to label $l$ by marginalizing over topics:
-
-\begin{equation}
-\label{eq:pwl}
-p(w|l) = {\sum_{z=1}^{K}{p(w|z,l)}}
-\end{equation}
-
-where $K$ is the optimal number of topics experimentally determined to minimize perplexity.
-After that, distributions over labels, $p(l|w)$, for each word were obtained from label-specific multinomials $p(w|l)$ using Bayesian inversion:
-
-\begin{equation}
-\label{eq:plw}
-p(l|w) = \frac{p(w|l)p(l)}{p(w)}
-\end{equation}
-
-where $p(l)$ is the prior probabilities of label $l$ in the training set and $p(w)$ is the prior probability of word $w$ in the language model estimated from the training set using maximum likelihood.
-We use the distribution $p(l|w)$ as the topic feature vector, which shows how indicative each word $w$ is for a particular behavior code $l$.
-Topic features allow to abstract away from individual words and detect segment boundaries by indirectly capturing transitions between behavior codes when only observing a pair of words that are specific to different behavior codes.
-To find the optimal topic feature, we also derived the same topic feature $p(l|w)$ by using two recently proposed probabilistic generative latent variable models: Discriminative Labeled Latent Dirichlet Allocation (DL-LDA) and Latent Class Allocation (LCA)\cite{kotov2015interpretable}.
-Similar as L-LDA, DL-LDA-based topic feature is extracted by first deriving a label-specific language model $p(w|l)$ for each label $l$ from the word distributions $p(w|l,z)$ for topic $z$ specific to label $l$ by using the Eq.~\ref{eq:pwl} and \ref{eq:plw}.
-On the other hand, LCA directly provides label-specific multinomials $p(w|l)$, which can be used to derive topic feature by using Eq.~\ref{eq:plw}.
+Since part-of-speech (PoS) tags \cite{kotov2015interpretable,hashimoto2016topic,lu2016modeling} have been shown to be effective semantic abstractions of individual words, we used topic features for our experiment.
 Punctuation marks, which correspond to one of the symbols \{`.', `,', `!', `?', `:', `;', `-'\} between a pair of words, are also employed as a feature since punctuation marks designate the boundary of a sentence, clause and phrase and often also correspond to segment boundary.
@@
 
-We experimented both with traditional classifiers, including Naive Bayes (NB)\cite{mccallum1998comparison}, Support Vector Machine (SVM)\cite{cortes1995support}, K-Nearest Neighbor (KNN)\cite{keller1985fuzzy} and two variants of Recurrent Neural Networks (RNNs)\cite{lipton2015critical}: Long Short Term Memory (LSTM)\cite{hochreiter1997long} and Gated Recurrent Unit (GRU)\cite{chung2014empirical}.
+We experimented both with traditional classifiers, including Conditional Random Field (CRF) and deep learning models, such as Multilayer Perceptrons (MLPs) Bidirectional Recurrent Neural Networks (BRNNs) and Convolutional Recurrent Neural Networks (CRNNs).
 
-\textbf{Naive Bayes}: this model estimates the prior probability of classes along with conditional probabilities of each feature given the class.
-Then, the posterior probability is computed to predict the class for each sample by applying the Bayes rule with the assumption that features are conditionally independent.
-We used Multinomial Naive Bayes in experiments.
-
-\textbf{Support Vector Machine}: state-of-the-art supervised classification method proven to be effective for text categorization\cite{joachims1998text} for its ability to cope with high dimensional input feature space.
+\textbf{Conditional Random Field}: state-of-the-art supervised classification method proven to be effective for text categorization\cite{joachims1998text} for its ability to cope with high dimensional input feature space.
 SVM finds the hyperplane that maximizes the separation between the closest ``new segment'' and ``same segment'' training examples.
@@
 
-\textbf{K-Nearest Neighbor}: this method considers each training sample as a point in the input feature space.
+\textbf{Multilayer Perceptrons}: this method considers each training sample as a point in the input feature space.
 For a new test sample, Euclidean distance is calculated to find the k-nearest classified neighbors from the training set.
@@
 
-\textbf{Recurrent Neural Networks}: RNN is a neural network architecture designed to capture sequential patterns.
+\textbf{Bidirectional Recurrent Neural Networks}: RNN is a neural network architecture designed to capture sequential patterns.
 When we predict the new segment point using RNNs, particular combinations of words in a sequence will indicate the change of MI behavior.
@@
 
-\begin{equation}
-z_t = \sigma(W_zx_t + U_zh_{t-1} + b_z)
-\label{eq:firstgru}
-\end{equation}
-\begin{equation}
-r_t = \sigma(W_rx_t + U_rh_{t-1} + b_r)
-\label{eq:resetgru}
-\end{equation}
-\begin{equation}
-\tilde h_t = tanh(W_hx_t + r_t \odot U_hh_{t-1} + b_h) 
-\label{eq:candidategru}
-\end{equation}
-\begin{equation}
-h_t = z_t \odot h_{t-1} + (1-z_t) \odot \tilde h_t
-\label{eq:lastgru}
-\end{equation}  
-In Eq.~\ref{eq:firstgru}-\ref{eq:lastgru}, $\sigma$ corresponds to sigmoid function and $\odot$ designates an element-wise product.
-The update gate $z_t$ and reset gate $r_t$ at time step $t$ are computed by the Eq.~(\ref{eq:firstgru}) and~(\ref{eq:resetgru}), where $W_z$, $W_r$, $W_h$, $U_z$, $U_r$, $U_h$ are the weight matrices and $b_z$, $b_h$ and $b_r$ are bias vectors.
-The activation $h_t$ of the GRU at time $t$ is a linear combination of previous activation $h_{t-1}$ and the candidate activation $\tilde h_t$, which is represented by Eq.~(\ref{eq:lastgru}) and~(\ref{eq:candidategru}).
+\textbf{Convolutional Recurrent Neural Networks}: RNN is a neural network architecture designed to capture sequential patterns.
+When we predict the new segment point using RNNs, particular combinations of words in a sequence will indicate the change of MI behavior.
+Long Short Term Memory (LSTM) networks\cite{hochreiter1997long} are a special type of RNN that are capable of handling variable size input sequence and have an internal memory that can be reset.
+GRU\cite{chung2014empirical} is a variant of LSTM mathematically represented by the following formula:
+
 Our GRU model includes one hidden layer, output layer and input layer.
@@
  KNN & \textbf{0.808} & 0.663 & \textbf{0.728} & 0.989 & \textbf{0.995} & \textbf{0.992} & \textbf{0.983} & \textbf{0.984} & \textbf{0.983} \\ \hline
- \multicolumn{10}{|c|}{lexical and puctuation features} \\ \hline
- LSTM & \textbf{0.810} & 0.720 & \textbf{0.762}  & \textbf{0.994} & \textbf{0.994} & \textbf{0.994} & \textbf{0.988} & \textbf{0.985} & \textbf{0.987} \\ \hline
- GRU & 0.783 & \textbf{0.739} & 0.760 & \textbf{0.994} & \textbf{0.994} & \textbf{0.994} & \textbf{0.988} & \textbf{0.985} & \textbf{0.987} \\ \hline
   \end{tabular}
@@
 GRU and LSTM show exactly same performance in terms of all performance metrics for ``same segement'' and overall classifications.
-\\
-
-\begin{table}[ht]
-\centering
-\caption{Performance of NB, SVM and KNN methods with different topic model-based features for identification of ``new segment'' class in e-Coaching emails, when all features are used together.
-The highest value for each performance metric is highlighted in bold.}
-\label{tab:result_boundary}
-  \begin{tabular}{|l|l|l|l|l|l|l|l|l|l|}
-  \hline
-   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{L-LDA-based topic features}} & \multicolumn{3}{|c|}{\textbf{DL-LDA-based topic features}}  & \multicolumn{3}{|c|}{\textbf{LCA-based topic features}} \\\cline{2-10} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1}\\ \hline
-
- NB & 0.573 & 0.680 & 0.622 & 0.573 & 0.677 & 0.621 & 0.593 & 0.662 & 0.626 \\ \hline
- SVM & 0.831 & 0.783 & 0.806 & \textbf{0.877} & \textbf{0.815} & \textbf{0.845} & 0.867 & 0.804 & 0.834  \\ \hline
- KNN & 0.808 & 0.741 & 0.773 & 0.814 & 0.732 & 0.771 & 0.814 & 0.740 & 0.775 \\ \hline
-  \end{tabular}
-\end{table}
-
-Table~\ref{tab:result_boundary} summarizes the performance of NB, SVM and KNN models for detecting ``new segment'' class.
-We observed that classification performance significantly improved for ``new segment'' class when lexical features are used in combination with punctuation and model-based topic features.
-Precision increases by 15.26\% and 0\%, 21.64\% and 0.74\% and 20.25\% and 0.74\% for SVM and KNN; recall increases by 11.06\% and 17.06\%, 15.60\% and 15.64\% and 14.04\% and 16.90\% for SVM and KNN; and F1-measure increases by 13.04\% and 6.18\%, 18.51\% and 5.91\% and 16.97\% and 6.46\% for SVM and KNN when L-LDA, DL-LDA and LCA-based topic features are used in conjunction combination with lexical and punctuation features, respectively.
-On the other hand, NB has very negligible impact of punctuation and model-based topic features.
-In contrast to results in Table~\ref{tab:result_base}, SVM demonstrates the best performance among all models achieving 0.877, 0.815 and 0.845 for precision, recall and F1-measure, respectively, when DL-LDA-based topic features are used in conjunction with lexical and punctuation features.
-NB and KNN provide their best performance when LCA-based topic features are used, achieving 0.626 and 0.775 F1-measure, respectively.\\
-
-\begin{table}[ht]
-\centering
-\caption{Performance of NB, SVM and KNN methods with different topic model-based features for identification of ``same segment'' class in e-Coaching emails, when all features are used together.
-The highest value for each performance metric is highlighted in bold.}
-\label{tab:result_not_boundary}
-  \begin{tabular}{|l|l|l|l|l|l|l|l|l|l|}
-  \hline
-   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{L-LDA-based topic features}} & \multicolumn{3}{|c|}{\textbf{DL-LDA-based topic features}}  & \multicolumn{3}{|c|}{\textbf{LCA-based topic features}} \\\cline{2-10} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1}\\ \hline
-
- NB & 0.989 & 0.983 & 0.986 & 0.989 & 0.983 & 0.986 & 0.988 & 0.985 & 0.987 \\ \hline
- SVM & 0.993 & 0.995 & 0.994 & \textbf{0.994} & \textbf{0.996} & \textbf{0.995} & 0.993 & \textbf{0.996} & \textbf{0.995} \\ \hline
- KNN & 0.991 & 0.994 & 0.993 & 0.991 & 0.994 & 0.993 & 0.991 & 0.994 & 0.993 \\ \hline
-  \end{tabular}
-\end{table}
-
-Similar to baseline performance, classification performance is remarkably higher for ``same segment'' class compared to ``new segment'' class (Table~\ref{tab:result_not_boundary}).
-The impact of punctuation and topic model-based features is also consistent with ``new segment'' classification while improvements are not as significant as ``new segment'' class.
-We observed that precision increases by 0.3\% and 0.2\%, 0.4\% and 0.2\% and 0.3\% and 0.2\% for SVM and KNN; recall increases by 0.4\% and -0.1\%, 0.5\% and -0.1\% and 0.5\% and -0.1\% for SVM and KNN; and F1-measure increases by 0.4\% and 0.1\%, 0.5\% and 0.1\% and 0.5\% and 0.1\% for SVM and KNN when L-LDA, DL-LDA and LCA-based topic features are used in conjunction combination with lexical and punctuation features, respectively.
-Similar to results in Table \ref{tab:result_boundary}, NB and KNN demonstrate the lowest and moderate performance on the task of identifying ``same segment'', respectively.
 \\
@@
 
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=1.0\textwidth]{figures/roc-curves.eps}
+    \caption{\textbf{Receiver operating characteristic curves showing the performance of binary classifiers for the segmentation of e-Coaching text when lexical features (left) and combination of lexical, DL-LDA-based topic (except LSTM and GRU) and punctuation features (right) are used}.}
+    \label{fig:roc-curves22}
+\end{figure}
+
 We observed moderate performance of RNN, in particular, LSTM and GRU for the text segmentation.
@@
 In this paper, we compared the performance of machine learning models for the task of segmentation of e-Coaching text.
-We found out that SVM provides the best performance for the segmentation of text in terms of all performance metrics.
+We found out that CRNN provides the best performance for the segmentation of text in terms of all performance metrics.
 Manual segmentation of e-Coaching data is very resource-intensive and time-consuming task, which can significantly decrease the time and effort required to develop an effective behavioral intervention.

### 2018-07-21 deepecoachts/deepecoach/train.py

@@ -127,7 +127,7 @@ def get_default_options():
 	return {
 		# required!
 		'id': 'default_example_ts',				# all models will be stored in data/models/:id:
-		'task': 'ts', 							# one of ts/dd_fillers/dd_editdisfs/ssdd
+		'task': 'ts', 							
 		'dataset': 'ecoach', 					# see loader.py (used only for training and error analysis)
 		'dataset-dir': None,
 
@@ -181,11 +181,6 @@ def configure(options={}):
 	file_handler = logging.FileHandler('{}.log'.format('data/log/'+options['id']))
 	file_handler.setFormatter(log_formatter)
 	root_logger.addHandler(file_handler)
-
-	# use GPU?
-	# if options['gpu']:
-	# 	import theano.sandbox.cuda
-	# 	theano.sandbox.cuda.use('gpu')
 
 	if not os.path.exists(options['model_dir']):
 		os.makedirs(options['model_dir'])

### 2018-07-21 text-segmentation-model/src/filemanager/Reader.java

@@ -186,7 +186,7 @@ public class Reader {
 	    for (String w : mapWordDictionary.keySet()) {
 	    	Map<Integer, Double> mapLabelByWord = new HashMap<Integer, Double>(); 
 	    	for (int label : mapLabelDist.keySet()) {
-	    		if (mapWordProbabilityForClass.get(label).containsKey(w)) {
+	    		if (mapWordProbabilityForClass.containsKey(label) && mapWordProbabilityForClass.get(label).containsKey(w)) {
 	    			double prob_w_by_l = mapWordProbabilityForClass.get(label).get(w);
 			    	double prob_w = (double)mapWordDictionary.get(w)/mapWordDictionary.size();
 			    	double prob_l_by_w = (prob_w_by_l*mapLabelDist.get(label))/prob_w;
@@ -240,7 +240,7 @@ public class Reader {
 	    for (String w : mapWordDictionary.keySet()) {
 	    	Map<Integer, Double> mapLabelByWord = new HashMap<Integer, Double>(); 
 	    	for (int label : mapLabelDist.keySet()) {
-	    		if (mapWordProbabilityForClass.get(label).containsKey(w)) {
+	    		if (mapWordProbabilityForClass.containsKey(label) && mapWordProbabilityForClass.get(label).containsKey(w)) {
 	    			double prob_w_by_l = mapWordProbabilityForClass.get(label).get(w);
 			    	double prob_w = (double)mapWordDictionary.get(w)/mapWordDictionary.size();
 			    	double prob_l_by_w = (prob_w_by_l*mapLabelDist.get(label))/prob_w;

### 2018-07-21 text-segmentation-model/src/filemanager/Writer.java

@@ -178,9 +178,13 @@ public class Writer {
 	
 	public static void createGoldStandard(String arffTextFile, String  arffFileStr2WordVector, 
 			String  goldStandardFile, String wordProbabilitiesFile, boolean withTopicDistribution, String model) {
+		
+		int invalidCount = 1;
 
 		// create word dictionary for the entire collections
 		HashMap<String, Integer> mapWordDictionary = Reader.getWordDictionary(arffFileStr2WordVector);
+		System.out.println(mapWordDictionary.size());
+		
 		Map<Integer, Double> mapLabelDist = Reader.getLabelDistribution(arffFileStr2WordVector);
 				
 		// create map for topic distribution
@@ -255,7 +259,8 @@ public class Writer {
 			        	}
 		        	}
 	        		else {
-	        			System.out.println(oneLineText[i]);
+	        			System.out.println(oneLineText[i] + " " + invalidCount);
+	        			invalidCount++;
 	        			continue;
 	        		}
 		        	

### 2018-07-21 text-segmentation-model/src/main/AppMain.java

@@ -48,7 +48,7 @@ public class AppMain {
 		// 5. add label as break (1) or not (0)
 		
 		FileManager.createGoldStandard(arffTextFile, arffFileStr2WordVector, goldStandardFile,
-				wordProbabilitiesFile, true, "lca");
+				wordProbabilitiesFile, true, "mg-lda");
 
 		// run model with arff file to evaluate the model performance
 		//SegmentationModel model = new SegmentationModel();

### 2018-07-24 paper-amia-summit-2018/manuscript.tex
Update manuscript

@@
 A critical step towards automated annotation of e-Coaching communication exchanges is segmentation of emails into textual fragments that correspond to MI behaviors.
-In this work, we formulate this task as a classification problem and propose lexical, punctuation and semantic features to address it.
+In this work, we formulate this task as a classification problem and propose lexical, punctuation and part-of-speech features to address it.
 We experimented both with traditional machine learning method, Conditional Random Field (CRF) and deep learning methods, such as Multilayer Perceptrons (MLPs), Bidirectional Recurrent Neural Networks (BRNNs) and Convolutional Recurrent Neural Networks (CRNNs).
@@
 Since part-of-speech (PoS) tags \cite{kotov2015interpretable,hashimoto2016topic,lu2016modeling} have been shown to be effective semantic abstractions of individual words, we used topic features for our experiment.
-Punctuation marks, which correspond to one of the symbols \{`.', `,', `!', `?', `:', `;', `-'\} between a pair of words, are also employed as a feature since punctuation marks designate the boundary of a sentence, clause and phrase and often also correspond to segment boundary.
+Punctuation marks, which correspond to one of the symbols \{`.', `,', `!', `?', `:', `;'\} between a pair of words, are also employed as a feature since punctuation marks designate the boundary of a sentence, clause and phrase and often also correspond to segment boundary.
 
@@
 
-\textbf{Multilayer Perceptrons}: this method considers each training sample as a point in the input feature space.
+\textbf{Deep Learning Methods}: this method considers each training sample as a point in the input feature space.
 For a new test sample, Euclidean distance is calculated to find the k-nearest classified neighbors from the training set.

### 2018-07-25 plots/preprocess-roc-data.py
added data for ROC plots

@@ -1,0 +1,37 @@
+# include required packages
+import os
+
+
+# convert roc.arff to roc.csv for figure
+def create_roc_csv(path):
+    for file_name in os.listdir(path):
+        file_name_writer = path + "/" + file_name.replace(".txt", ".csv")
+        if 'csv' not in file_name:
+            fw = open(file_name_writer, "w")
+            fw.write("actual,predict_prob\n")
+            truths = []
+            preds = []
+            with open(path + "/" + file_name, "r") as file_stream:
+                flag = 0
+                for line in file_stream:
+                    current_line = line.strip().split()
+                    if len(current_line) < 10:
+                        current_line = line.strip().split(',')
+                    if flag % 2 == 0:
+                        for x in current_line:
+                            truths.append(x)
+                    else:
+                        for x in current_line:
+                            preds.append(x)
+                    flag += 1
+
+                print(path + "/" + file_name)
+                for i in range(0, len(truths)):
+                    fw.write(truths[i] + "," + preds[i] + "\n")
+
+            fw.close()
+
+# call above functions for the creation of new files
+create_roc_csv('roc/base')
+create_roc_csv('roc/propose')
+print('conversion is done!')

### 2018-07-25 plots/roc-curves-generator.R

@@ -1,0 +1,64 @@
+
+#library
+library(ggplot2)
+library(reshape2)
+library(pROC)
+
+par(mfrow=c(1,2))
+wd = '/home/mehedi/teana/projects/eCoaching-Text-Segmentation/plots/'
+
+# melt data
+# print AUC values
+x1 <- read.csv(paste(wd,"roc/base/pred_brnn.csv", sep=''))
+x2 <- read.csv(paste(wd,"roc/base/pred_mlp.csv", sep=''))
+x3 <- read.csv(paste(wd,"roc/base/pred_crf.csv", sep=''))
+x4 <- read.csv(paste(wd,"roc/base/pred_crnn.csv", sep=''))
+
+roc_obj_brnn <- roc(x1$actual, x1$predict_prob)
+roc_obj_mlp <- roc(x2$actual, x2$predict_prob)
+roc_obj_crf <- roc(x3$actual, x3$predict_prob)
+roc_obj_crnn <- roc(x4$actual, x4$predict_prob)
+
+auc(roc_obj_brnn)
+auc(roc_obj_mlp)
+auc(roc_obj_crf)
+auc(roc_obj_crnn)
+
+# plot graph
+plot(1-roc_obj_brnn$specificities, roc_obj_brnn$sensitivities, col='red', 
+     type="l", ylab = "Sensitivity", xlab = "1-Specificity", cex.axis=1, cex.lab=1, lwd=1)
+lines(1-roc_obj_mlp$specificities, roc_obj_mlp$sensitivities, col='blue', lty=1, lwd=1)
+lines(1-roc_obj_crf$specificities, roc_obj_crf$sensitivities, col='green', lty=1, lwd=1)
+lines(1-roc_obj_crnn$specificities, roc_obj_crnn$sensitivities, col='magenta', lty=1, lwd=1)
+lines(0:1, 0:1, col='black', lty=5, lwd=1)
+legend(0.77,0.30, c("BRNN", "MLP", "CRF", "CRNN"), 
+       lty=c(1,1,1,1), col=c('red', 'blue', 'green', 'magenta'), cex=1, lwd=1)
+
+
+# print AUC values
+x1 <- read.csv(paste(wd,"roc/propose/pred_brnn.csv", sep=''))
+x2 <- read.csv(paste(wd,"roc/propose/pred_mlp.csv", sep=''))
+x3 <- read.csv(paste(wd,"roc/propose/pred_crf.csv", sep=''))
[…]

### 2018-07-25 paper-amia-summit-2018/manuscript.tex
update results

@@
 
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=1.0\textwidth]{figures/CRNN.eps}
+    \caption{\textbf{Receiver operating characteristic curves showing the performance of binary classifiers for the segmentation of e-Coaching text when lexical features (left) and combination of lexical, DL-LDA-based topic (except LSTM and GRU) and punctuation features (right) are used}.}
+    \label{fig:crnn}
+\end{figure}
+
 Our GRU model includes one hidden layer, output layer and input layer.
@@
 \label{tab:result_base}
-  \begin{tabular}{|l|l|l|l|l|l|l|l|l|l|}
+  \begin{tabular}{|l|l|l|l|l|l|l|}
   \hline
-   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{new segment}} & \multicolumn{3}{|c|}{\textbf{same segment}}  & \multicolumn{3}{|c|}{\textbf{overall}} \\\cline{2-10}
-   & \textbf{Precision}  & \textbf{Recall} & \textbf{F1} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1}\\ \hline    
- \multicolumn{10}{|c|}{lexical features only} \\ \hline   
- NB & 0.594 & 0.662 & 0.626 & 0.988 & 0.985 & 0.987 & 0.975 & 0.974 & 0.975  \\ \hline
- SVM & 0.721 & \textbf{0.705} & 0.713 & \textbf{0.990} & 0.991 & 0.990  & 0.981 & 0.981 & 0.981 \\ \hline
- KNN & \textbf{0.808} & 0.663 & \textbf{0.728} & 0.989 & \textbf{0.995} & \textbf{0.992} & \textbf{0.983} & \textbf{0.984} & \textbf{0.983} \\ \hline
+   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{new segment}} & \multicolumn{3}{|c|}{\textbf{weighted macro average}} \\\cline{2-7}
+   & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} \\ \hline    
+ CRF & 0.782 & 0.691 & 0.733 & 0.983 & 0.984 & 0.984 \\ \hline
+ MLP & \textbf{0.836} & 0.593 & 0.694 & 0.982 & 0.983 & 0.982 \\ \hline
+ BRNN & 0.606 & 0.680 & 0.641 & 0.977 & 0.976 & 0.976 \\ \hline
+ CRNN & 0.775 & \textbf{0.797} & \textbf{0.785} & \textbf{0.986} & \textbf{0.986} & \textbf{0.986} \\ \hline
   \end{tabular}
@@
 \label{tab:result_weighted_avg}
-  \begin{tabular}{|l|l|l|l|l|l|l|l|l|l|}
+ \begin{tabular}{|l|l|l|l|l|l|l|}
   \hline
-   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{L-LDA-based topic features}} & \multicolumn{3}{|c|}{\textbf{DL-LDA-based topic features}}  & \multicolumn{3}{|c|}{\textbf{LCA-based topic features}} \\\cline{2-10} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1}\\ \hline
-
- NB & 0.975 & 0.973 & 0.974 & 0.975 & 0.973 & 0.974 & 0.975 & 0.974 & 0.975 \\ \hline
- SVM & 0.987 & 0.988 & 0.987 & \textbf{0.990} & \textbf{0.990} & \textbf{0.990} & 0.989 & 0.989 & 0.989 \\ \hline
- KNN & 0.985 & 0.986 & 0.985 & 0.985 & 0.986 & 0.985 & 0.985 & 0.986 & 0.986 \\ \hline
+   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{new segment}} & \multicolumn{3}{|c|}{\textbf{weighted macro average}} \\\cline{2-7}
+   & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} \\ \hline    
+ CRF & 0.813 & 0.772 & 0.792 & 0.988 & 0.988 & 0.988 \\ \hline
+ MLP & \textbf{0.814} & 0.732 & 0.771 & 0.987 & 0.987 & 0.987 \\ \hline
+ BRNN & 0.683 & 0.820 & 0.745 & 0.985 & 0.983 & 0.984 \\ \hline
+ CRNN & 0.789 & \textbf{0.864} & \textbf{0.825} & \textbf{0.990} & \textbf{0.989} & \textbf{0.989} \\ \hline
   \end{tabular}
@@
 \end{figure}
+
+% Baseline ROC results: BRNN: 0.9198, MLP: 0.9507, CRF: 0.9662, CRNN: 0.9648
+% Proposed ROC Results: BRNN: 0.9807, MLP: 0.992, CRF: 0.9937, CRNN: 0.9859
 

### 2018-07-25 plots/roc-curves-generator.R

@@ -31,7 +31,7 @@ lines(1-roc_obj_mlp$specificities, roc_obj_mlp$sensitivities, col='blue', lty=1,
 lines(1-roc_obj_crf$specificities, roc_obj_crf$sensitivities, col='green', lty=1, lwd=1)
 lines(1-roc_obj_crnn$specificities, roc_obj_crnn$sensitivities, col='magenta', lty=1, lwd=1)
 lines(0:1, 0:1, col='black', lty=5, lwd=1)
-legend(0.77,0.30, c("BRNN", "MLP", "CRF", "CRNN"), 
+legend(0.77,0.33, c("BRNN", "MLP", "CRF", "CRNN"), 
        lty=c(1,1,1,1), col=c('red', 'blue', 'green', 'magenta'), cex=1, lwd=1)
 
 
@@ -58,7 +58,7 @@ lines(1-roc_obj_mlp$specificities, roc_obj_mlp$sensitivities, col='blue', lty=1,
 lines(1-roc_obj_crf$specificities, roc_obj_crf$sensitivities, col='green', lty=1, lwd=1)
 lines(1-roc_obj_crnn$specificities, roc_obj_crnn$sensitivities, col='magenta', lty=1, lwd=1)
 lines(0:1, 0:1, col='black', lty=5, lwd=1)
-legend(0.77,0.30, c("BRNN", "MLP", "CRF", "CRNN"), 
+legend(0.77,0.33, c("BRNN", "MLP", "CRF", "CRNN"), 
        lty=c(1,1,1,1), col=c('red', 'blue', 'green', 'magenta'), cex=1, lwd=1)
 
 print('done')

### 2018-07-25 paper-amia-summit-2018/manuscript.tex
update abstract and conclusion

@@
 In this work, we formulate this task as a classification problem and propose lexical, punctuation and part-of-speech features to address it.
-We experimented both with traditional machine learning method, Conditional Random Field (CRF) and deep learning methods, such as Multilayer Perceptrons (MLPs), Bidirectional Recurrent Neural Networks (BRNNs) and Convolutional Recurrent Neural Networks (CRNNs).
-Results indicate that CRNN outperformed CRF, MLP and BRNN achieving 0.990 macro F1-score overall and 0.821 macro F1-score for detecting new segment.}
+We experimented both with traditional machine learning method, Conditional Random Fields (CRF) and deep learning methods, such as Multilayer Perceptrons (MLP), Bidirectional Recurrent Neural Networks (BRNN) and Convolutional Recurrent Neural Networks (CRNN).
+Results indicate that CRNN outperformed CRF, MLP and BRNN achieving 0.989 macro F1-score overall and 0.825 macro F1-score for detecting new segment.}
 

### 2018-07-26 paper-amia-summit-2018/manuscript.tex
update Introduction

@@
 In this work, we formulate this task as a classification problem and propose lexical, punctuation and part-of-speech features to address it.
-We experimented both with traditional machine learning method, Conditional Random Fields (CRF) and deep learning methods, such as Multilayer Perceptrons (MLP), Bidirectional Recurrent Neural Networks (BRNN) and Convolutional Recurrent Neural Networks (CRNN).
+We experimented both with traditional machine learning method, conditional random fields (CRF) and deep learning methods, such as multilayer perceptrons (MLP), bidirectional recurrent neural networks (BRNN) and convolutional recurrent neural networks (CRNN).
 Results indicate that CRNN outperformed CRF, MLP and BRNN achieving 0.989 macro F1-score overall and 0.825 macro F1-score for detecting new segment.}
@@
 The goal of this research study is to assess the applicability of machine learning methods for automated segmentation of e-Coaching emails into textual fragments corresponding to individual behaviors, which is the first step of the coding process of e-Coaching communications.
-In particular, we introduced lexical, semantic and punctuation features and experimented with both traditional supervised machine learning method, linear-chain Conditional Random Field (CRF) and deep learning methods, such as Multilayer Perceptrons (MLPs), Bidirectional Recurrent Neural Networks (BRNNs) and Convolutional Recurrent Neural Networks (CRNNs), to find the best performing method and feature combination.
+In particular, we introduced lexical, punctuation and  part-of-speach (POS) features and experimented with both traditional supervised machine learning method, linear-chain conditional random fields (CRF) and deep learning methods, such as multilayer perceptrons (MLP), bidirectional recurrent neural networks (BRNN) and convolutional recurrent neural networks (CRNN), to find the best performing method and feature combination.
 
-Relevant previous work in the biomedical domain primarily focused on segmentation of text into sections and headers\cite{apostolova2009automatic,denny2009evaluation,tepper2012statistical,cho2002text} or sentence boundary detection\cite{griffis2016quantitative,kreuzthaler2015detection,treviso2016sentence,fraser2015sentence}.
+Relevant previous work in the biomedical domain primarily focused on segmentation of text into sections and headers\cite{apostolova2009automatic,denny2009evaluation,tepper2012statistical,cho2002text} or sentence boundary detection\cite{griffis2016quantitative,kreuzthaler2015detection,treviso2017sentence,fraser2015sentence}.
 Apostolova et al.\cite{apostolova2009automatic} applied SVM along with word-vector cosine similarity metric combined with several heuristics to segment clinical reports into sections, such as demographics, history, procedure, finding and impression.
@@
 On the other hand, SVM based on prosodic and part of speech features \cite{kreuzthaler2015detection} and recurrent convolutional neural networks using word embeddings \cite{griffis2016quantitative} were utilized for detecting sentence boundaries.
-Segmentation of e-Coaching emails is different from a traditional shallow discourse analysis of conversations \cite{galley2003discourse} in that the focus is on segmentation, rather than on determining the types of transitions between the utterances or assigning utterances to speakers.
+Liu et al.
+have shown that a linear chain CRF performed better than hidden markov model and Maxent on the NIST sentence boundary detection \cite{liu2005using}.
+Recently, Treviso et al.
+proposed a recurrent convolutional neural networks for automatic identification of sentence boundary in neuropsychological text in Portuguese language \cite{treviso2017sentence}.
+Their proposed model combined prosodic, part of speech and lexical features, which achieved 0.74 F1-score in sentence boundary detection in Brazilian constitution dataset.
+We adapted this model in our work, which was employed differently from the aforementioned studies.
+Our study focus on segmentation, rather than determining the sentence boudary.
+Another difference is that punctuation features were utilized in combination with part-of-speech features.
+Besides, our model trains using 300 dimension pre-trained word2Vec word vectors \cite{mikolov2013distributed}, rather than utilizing 50 dimension pre-trained GloVe word vectors \cite{pennington2014glove}.
+Our experiments show that these differences lead to significant improvement in the segmenteation of e-Coaching emails.
+Segmentation of e-Coaching emails is also different from a traditional shallow discourse analysis of conversations \cite{galley2003discourse} in that the focus is on segmentation, rather than on determining the types of transitions between the utterances or assigning utterances to speakers.
 
@@
 However, segmentation of clinical conversation in the context of electronically delivered interventions, in particular, segmentation of clinical interaction text into groups of MI behaviors, is still performed manually, which slows down qualitative analysis of these interventions.
-This study introduces a novel computational approach to address this problem and the authors are unaware of any other work that focused on the same problem.
+This study proposes several machine learning methods to address this problem and the authors are unaware of any other work that focused on the same problem.
 

### 2018-07-27 paper-amia-summit-2018/manuscript.tex
update data collection and Features sections

@@
 The goal of this research study is to assess the applicability of machine learning methods for automated segmentation of e-Coaching emails into textual fragments corresponding to individual behaviors, which is the first step of the coding process of e-Coaching communications.
-In particular, we introduced lexical, punctuation and  part-of-speach (POS) features and experimented with both traditional supervised machine learning method, linear-chain conditional random fields (CRF) and deep learning methods, such as multilayer perceptrons (MLP), bidirectional recurrent neural networks (BRNN) and convolutional recurrent neural networks (CRNN), to find the best performing method and feature combination.
+In particular, we introduced lexical, punctuation and  part-of-speech (POS) features and experimented with both traditional supervised machine learning method, linear-chain conditional random fields (CRF) and deep learning methods, such as multilayer perceptrons (MLP), bidirectional recurrent neural networks (BRNN) and convolutional recurrent neural networks (CRNN), to find the best performing method and feature combination.
 
@@
 Liu et al.
-have shown that a linear chain CRF performed better than hidden markov model and Maxent on the NIST sentence boundary detection \cite{liu2005using}.
+have shown that a linear chain CRF performed better than the hidden markov model and Maxent on the NIST sentence boundary detection \cite{liu2005using}.
 Recently, Treviso et al.
-proposed a recurrent convolutional neural networks for automatic identification of sentence boundary in neuropsychological text in Portuguese language \cite{treviso2017sentence}.
-Their proposed model combined prosodic, part of speech and lexical features, which achieved 0.74 F1-score in sentence boundary detection in Brazilian constitution dataset.
-We adapted this model in our work, which was employed differently from the aforementioned studies.
-Our study focus on segmentation, rather than determining the sentence boudary.
-Another difference is that punctuation features were utilized in combination with part-of-speech features.
-Besides, our model trains using 300 dimension pre-trained word2Vec word vectors \cite{mikolov2013distributed}, rather than utilizing 50 dimension pre-trained GloVe word vectors \cite{pennington2014glove}.
-Our experiments show that these differences lead to significant improvement in the segmenteation of e-Coaching emails.
+proposed a recurrent convolutional neural network for automatic identification of sentence boundary in a neuropsychological text in Portuguese language \cite{treviso2017sentence}.
+Their proposed model combined prosodic and part of speech features with pre-trained GloVe word vectors \cite{pennington2014glove}, which achieved 0.74 F1-score in sentence boundary detection in the Brazilian constitution dataset.
+
+In this study, we adopt the model proposed by Treviso et al.
+\cite{treviso2017sentence} and apply it differently from the aforementioned studies.
+The main difference is that our study focuses on segmentation of e-Coaching emails as distinct from determining the sentence boundary.
+Another important difference is that punctuation features were utilized in combination with word embeddings and part-of-speech features.
+Furthermore, we train our model with pre-trained word2vec word vectors \cite{mikolov2013distributed}, instead of using GloVe word vectors.
 Segmentation of e-Coaching emails is also different from a traditional shallow discourse analysis of conversations \cite{galley2003discourse} in that the focus is on segmentation, rather than on determining the types of transitions between the utterances or assigning utterances to speakers.
+
+%Our experiments show that these differences lead to significant improvement in the segmentation of e-Coaching emails.
 
@@
 \subsection*{\textit{Data collection}}
+
 The experimental dataset for this work was constructed from 49 e-coaching sessions, which include a total of 3,138 segmented and annotated MI behaviors.
 Each session represents an MI intervention delivered via email.
-We formulate the segmentation task as a binary classification problem where each word annotated with a label 1 or 0, to indicate whether it preceds a new segment or not.
-We experimented both with traditional machine learning method, linear-chain Conditional Random Field (CRF) and deep learning methods, such as Multilayer Perceptrons (MLPs), Bidirectional Recurrent Neural Networks (BRNNs) and Convolutional Recurrent Neural Networks (CRNNs).
-For MLP models, e-Coaching email exchanges are partitioned into adjacent word pairs.
-Each pair is classified into either ``new segment'' or ``same segment'' class.
-The gold standard is manually segmented where adjacent word pairs are classified into ``new segment'' class.
-If all adjacent word pairs of a block of text are classified into ``same segment'' class, the entire block is treated as one textual fragment corresponding to a single MI behavior.
-In total, we obtained 95,421 word pairs, which include 3,138 ``new segment'' and 92,283 ``same segment'' instances.
-For deep learning models, an email was taken as input sequence, such that PoS tags and word embedding of each word or punctuation marks were used as input and binary labels (1 or 0) corresponding to ``new segment'' and ``same segment'' classification decision were considered as the output of the model at each step.
+We formulate the segmentation task as a binary classification problem where each word or punctuation marks annotated with a label 1 or 0, to indicate whether it precedes a new segment or not.
+In total, we obtained 95,777 words and 7,140 punctuation marks, which include 3,138 ``new segment'' and 99,779 ``same segment'' instances.
+In this study, we experimented with traditional machine learning method, CRF and deep learning methods, such as MLP, BRNN and CRNN.
+For MLP models, samples were created with $2l$ words/punctuation marks at each step in a word sequence, where a sample contains next $l$ words/punctuation marks and prior $l$ words/punctuation marks including current word/punctuation mark.
+Each sample is classified into either ``new segment'' or ``same segment'' based on whether the current word/punctuation mark precedes a new segment or not.
+For CRF, BRNN and CRNN models, an email was taken as input sequence, such that POS tags and word embedding of each word or punctuation marks were used as input and binary labels (1 or 0) corresponding to ``new segment'' and ``same segment'' classification decision were considered as the output of the model at each step.
 In the gold standard, words within the same segment were assigned the label of 0 and the last word or punctuation mark of a segment were assigned the label of 1.
@@
 
-We utilized three types of features in conjunction with CRF, MLP, BRNN and CRNN methods: lexical, punctuation and topic features.
-Lexical features represent each word in a pair as a binary vector, in which 1 corresponds to the word in question and 0 to all other words.
-Since part-of-speech (PoS) tags \cite{kotov2015interpretable,hashimoto2016topic,lu2016modeling} have been shown to be effective semantic abstractions of individual words, we used topic features for our experiment.
-Punctuation marks, which correspond to one of the symbols \{`.', `,', `!', `?', `:', `;'\} between a pair of words, are also employed as a feature since punctuation marks designate the boundary of a sentence, clause and phrase and often also correspond to segment boundary.
+We utilized three types of features in conjunction with CRF, MLP, BRNN and CRNN methods: word embeddings, punctuation and POS features.
+Since POS tags have been shown to be effective semantic abstractions of individual words, we used topic features for our experiment \cite{liu2005using,treviso2017sentence}.
+To extract our POS features, we tag the e-Coaching emails using the NLTK POS tagger \cite{bird2009natural}.
+Punctuation marks, which correspond to one of the symbols \{`.', `,', `!', `?', `:', `;'\} between a pair of words, are also employed as a feature since punctuation marks designate the boundary of a sentence, clause and phrase and often also correspond to a segment boundary.
+For natural language processing (NLP) tasks, inputs are received as a text where words can be considered as the basic processing unit.
+Therefore, it is important to represent a word in such a way that it carries all relevant information.
+Word embeddings are one of this representation where each word represented as a real-valued vector in a high dimensional vector space.
+The distributed representation of word allows capturing semantic, syntactic and morphological information from large unannotated corpora  \cite{pennington2014glove, mikolov2013distributed}.
+In our experiment, we utilized pre-trained word2vec word vectors for words and trained word2vec word vectors for punctuation marks.
+
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=1.0\textwidth]{figures/roc-curves.eps}
+    \caption{\textbf{Receiver operating characteristic curves showing the performance of binary classifiers for the segmentation of e-Coaching text when lexical features (left) and combination of lexical, DL-LDA-based topic (except LSTM and GRU) and punctuation features (right) are used}.}
+    \label{fig:roc-curves22}
+\end{figure}
 
@@
 
-We experimented both with traditional classifiers, including Conditional Random Field (CRF) and deep learning models, such as Multilayer Perceptrons (MLPs) Bidirectional Recurrent Neural Networks (BRNNs) and Convolutional Recurrent Neural Networks (CRNNs).
+In the present study, we experimented with four different classifiers, including one traditional machine learning method, CRF and three neural networks models.
+Since neural networks provide a flexible architecture for constructing and synthesizing complex models, we take the advantage of this flexibility to test MLP, BRNN and CRNN models for the task of segmentation of e-Coaching emails.
+
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=1.0\textwidth]{figures/roc-curves.eps}
+    \caption{\textbf{Receiver operating characteristic curves showing the performance of binary classifiers for the segmentation of e-Coaching text when lexical features (left) and combination of lexical, DL-LDA-based topic (except LSTM and GRU) and punctuation features (right) are used}.}
+    \label{fig:roc-curves22}
+\end{figure}
 
@@
 
-\textbf{Deep Learning Methods}: this method considers each training sample as a point in the input feature space.
+\textbf{Multilayer Perceptrons}: this method considers each training sample as a point in the input feature space.
 For a new test sample, Euclidean distance is calculated to find the k-nearest classified neighbors from the training set.
@@
 \textbf{Bidirectional Recurrent Neural Networks}: RNN is a neural network architecture designed to capture sequential patterns.
-When we predict the new segment point using RNNs, particular combinations of words in a sequence will indicate the change of MI behavior.
-Long Short Term Memory (LSTM) networks\cite{hochreiter1997long} are a special type of RNN that are capable of handling variable size input sequence and have an internal memory that can be reset.
-GRU\cite{chung2014empirical} is a variant of LSTM mathematically represented by the following formula:
-
-\textbf{Convolutional Recurrent Neural Networks}: RNN is a neural network architecture designed to capture sequential patterns.
 When we predict the new segment point using RNNs, particular combinations of words in a sequence will indicate the change of MI behavior.
@@
 \end{figure}
+
+\textbf{Convolutional Recurrent Neural Networks}: RNN is a neural network architecture designed to capture sequential patterns.
+When we predict the new segment point using RNNs, particular combinations of words in a sequence will indicate the change of MI behavior.
+Long Short Term Memory (LSTM) networks\cite{hochreiter1997long} are a special type of RNN that are capable of handling variable size input sequence and have an internal memory that can be reset.
+GRU\cite{chung2014empirical} is a variant of LSTM mathematically represented by the following formula:
 
@@
 
-% Baseline ROC results: BRNN: 0.9198, MLP: 0.9507, CRF: 0.9662, CRNN: 0.9648
-% Proposed ROC Results: BRNN: 0.9807, MLP: 0.992, CRF: 0.9937, CRNN: 0.9859
-
 In this paper, results are reported by each class to avoid confusion regarding overall model performance due to severe class imbalance.
@@
 
-\begin{figure}[!htb]
+\begin{table}[ht]
 \centering
-    \includegraphics[width=1.0\textwidth]{figures/roc-curves.eps}
-    \caption{\textbf{Receiver operating characteristic curves showing the performance of binary classifiers for the segmentation of e-Coaching text when lexical features (left) and combination of lexical, DL-LDA-based topic (except LSTM and GRU) and punctuation features (right) are used}.}
-    \label{fig:roc-curves22}
-\end{figure}
+\caption{Performance of NB, SVM and KNN methods with different topic model-based features for segmentation of e-Coaching emails as a weighted average over ``same segment'' and ``new segment'' classification results, when all features are used together.
+The highest value for each performance metric is highlighted in bold.}
+\label{tab:result_weighted_avg}
+ \begin{tabular}{|l|l|l|l|l|}
+  \hline
+\multirow{2}{*}{\textbf{Features}} & \multicolumn{4}{|c|}{\textbf{Area Under the Curve (AUC)}} \\\cline{2-5}
+ & \textbf{CRF} & \textbf{MLP}  & \textbf{BRNN} & \textbf{CRNN} \\ \hline      
+ word embeddings & \textbf{0.966} & 0.951 & 0.920 & 0.965 \\ \hline
+ all features & \textbf{0.994} & 0.992 & 0.981 & 0.986 \\ \hline
+  \end{tabular}
+\end{table}
+
+% Baseline ROC results: BRNN: 0.9198, MLP: 0.9507, CRF: 0.9662, CRNN: 0.9648
+% Proposed ROC Results: BRNN: 0.9807, MLP: 0.992, CRF: 0.9937, CRNN: 0.9859
 

### 2018-07-27 paper-amia-summit-2018/manuscript.tex
added figures for parameters

@@
 The distributed representation of word allows capturing semantic, syntactic and morphological information from large unannotated corpora  \cite{pennington2014glove, mikolov2013distributed}.
-In our experiment, we utilized pre-trained word2vec word vectors for words and trained word2vec word vectors for punctuation marks.
+Since pre-trained word2vec word vectors provides the best results, in our experiment, we utilized pre-trained word2vec word vectors for words.
+Trained word vectors were also utilized when punctuation marks are not available in the pre-trained word2vec word vectors.
+Figure~\ref{fig:embedding-dimension} illustrates the performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and trained word embeddings with GloVe and word2vec models.
 
@@
     \centering
-    \includegraphics[width=1.0\textwidth]{figures/roc-curves.eps}
-    \caption{\textbf{Receiver operating characteristic curves showing the performance of binary classifiers for the segmentation of e-Coaching text when lexical features (left) and combination of lexical, DL-LDA-based topic (except LSTM and GRU) and punctuation features (right) are used}.}
-    \label{fig:roc-curves22}
+    \includegraphics[width=0.6\textwidth]{figures/embedding-dimension.eps}
+    \caption{\textbf{Performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and trained word embeddings with GloVe and word2vec models.}}
+    \label{fig:embedding-dimension}
 \end{figure}
@@
 In the present study, we experimented with four different classifiers, including one traditional machine learning method, CRF and three neural networks models.
-Since neural networks provide a flexible architecture for constructing and synthesizing complex models, we take the advantage of this flexibility to test MLP, BRNN and CRNN models for the task of segmentation of e-Coaching emails.
+Since neural networks provide a flexible architecture for constructing and synthesizing complex models, we take the advantage of this flexibility to test MLP, BRNN and CRNN models for the task of segmentation of e-Coaching emails.\\
 
@@
     \centering
-    \includegraphics[width=1.0\textwidth]{figures/roc-curves.eps}
-    \caption{\textbf{Receiver operating characteristic curves showing the performance of binary classifiers for the segmentation of e-Coaching text when lexical features (left) and combination of lexical, DL-LDA-based topic (except LSTM and GRU) and punctuation features (right) are used}.}
-    \label{fig:roc-curves22}
+    \includegraphics[width=0.60\textwidth]{figures/length-mlp.eps}
+    \caption{\textbf{Performance of MLP model on e-Coaching email segmentation by varying the length of sample size.}}
+    \label{fig:length-mlp}
 \end{figure}
@@
     \includegraphics[width=1.0\textwidth]{figures/CRNN.eps}
-    \caption{\textbf{Receiver operating characteristic curves showing the performance of binary classifiers for the segmentation of e-Coaching text when lexical features (left) and combination of lexical, DL-LDA-based topic (except LSTM and GRU) and punctuation features (right) are used}.}
+    \caption{\textbf{Architecture of convolutional recurrent neural networks for automatic segmentation of e-Coaching emails}.}
     \label{fig:crnn}
@@
 \centering
-\caption{Performance of NB, SVM and KNN methods for identification of ``new segment'' and ``same segment'' classes as well as their weighted average.
-The highest value for each performance metric is highlighted in bold.}
+\caption{Performance of CRF, MLP, BRNN and CRNN methods for identification of ``new segment'' class as well as weighted average over ``new segment'' and ``same segment'' classes when word embeddings or lexical features are used.
+The highest value for each performance metric
+is highlighted in bold.}
 \label{tab:result_base}
@@
 \centering
-\caption{Performance of NB, SVM and KNN methods with different topic model-based features for segmentation of e-Coaching emails as a weighted average over ``same segment'' and ``new segment'' classification results, when all features are used together.
-The highest value for each performance metric is highlighted in bold.}
+\caption{Performance of CRF, MLP, BRNN and CRNN methods for identification of ``new segment'' class as well as weighted average over ``new segment'' and ``same segment'' classes when all features are used together.
+The highest value for each performance metric
+is highlighted in bold.}
 \label{tab:result_weighted_avg}
@@
 This results also indicate that segmentation performance might be improved by adding additional relevant features.
+\\
 
@@
     \includegraphics[width=1.0\textwidth]{figures/roc-curves.eps}
-    \caption{\textbf{Receiver operating characteristic curves showing the performance of binary classifiers for the segmentation of e-Coaching text when lexical features (left) and combination of lexical, DL-LDA-based topic (except LSTM and GRU) and punctuation features (right) are used}.}
+    \caption{\textbf{Receiver operating characteristic curves showing the performance of binary classifiers for the segmentation of e-Coaching text when only word embeddings/lexical (left) and combination of word embeddings/lexical, POS and punctuation features (right) are used}.}
     \label{fig:roc-curves}
@@
 \centering
-\caption{Performance of NB, SVM and KNN methods with different topic model-based features for segmentation of e-Coaching emails as a weighted average over ``same segment'' and ``new segment'' classification results, when all features are used together.
-The highest value for each performance metric is highlighted in bold.}
-\label{tab:result_weighted_avg}
+\caption{AUC values of all classifiers when only word embeddings/lexical and combination of word embeddings/lexical, POS and punctuation features are used.
+Highest value for each experiment across all models is highlighted in boldface.}
+\label{tab:result_roc}
  \begin{tabular}{|l|l|l|l|l|}
   \hline
-\multirow{2}{*}{\textbf{Features}} & \multicolumn{4}{|c|}{\textbf{Area Under the Curve (AUC)}} \\\cline{2-5}
+\multirow{2}{*}{\textbf{Features}} & \multicolumn{4}{|c|}{\textbf{Area Under the ROC Curve (AUC)}} \\\cline{2-5}
  & \textbf{CRF} & \textbf{MLP}  & \textbf{BRNN} & \textbf{CRNN} \\ \hline      
- word embeddings & \textbf{0.966} & 0.951 & 0.920 & 0.965 \\ \hline
+ word embeddings or lexical features & \textbf{0.966} & 0.951 & 0.920 & 0.965 \\ \hline
  all features & \textbf{0.994} & 0.992 & 0.981 & 0.986 \\ \hline

### 2018-07-28 paper-amia-summit-2018/manuscript.tex
update methods section

@@
     \centering
-    \includegraphics[width=0.9\textwidth]{figures/segment-example.png}
+    \includegraphics[width=0.8\textwidth]{figures/segment-example.png}
     \caption{\textbf{Example of e-Coaching emails segmented into fragments that correspond to MI behaviors of an e-Coach and a patient}.}
@@
 In the present study, we experimented with four different classifiers, including one traditional machine learning method, CRF and three neural networks models.
-Since neural networks provide a flexible architecture for constructing and synthesizing complex models, we take the advantage of this flexibility to test MLP, BRNN and CRNN models for the task of segmentation of e-Coaching emails.\\
+Since neural networks provide a flexible architecture for constructing and synthesizing complex models, we take the advantage of this flexibility to test MLP, BRNN and CRNN models for the task of segmentation of e-Coaching emails.
+
+\textbf{Conditional Random Fields}: CRF have been widely used in various NLP tasks including part-of-speech tagging and segmentation tasks \cite{lafferty2001conditional, hirohata2008identifying}.
+Unlike a maximum entropy markov model which uses per-state exponential models for the conditional probabilities of next states given the current state, a CRF is a discriminative model of the conditional distribution of the entire sequence of labels given the observation sequence.
+A traditional linear-chain CRF model is defined as a conditional probability distribution $p(y|x)$ for output and input sequences, $y$ and $x$, by the following equation:
+
+\begin{equation}
+p(y|x) = \frac{1}{Z_x}\exp{\left(\sum_{t=1}^{T}\sum_{k}^{}\lambda_k f_k(y_{t-1}, y_t, x, t)\right)} \\
+\end{equation}
+
+Here $Z_x$ is a normalization factor over all possible labelings of $x$, and $f_k(y_{t-1}, y_t, x, t)$ is a feature function, and $\lambda_k$ is a learned weight associated with feature $f_k$.
+The optimal output sequence $y^*$ for an input sequence $x$, $y^* = {arg\,max}_y p(y|x)$, is obtained efficiently by the Viterbi algorithm.
+In our experiments, the following features were utilized with CRF models: i) current word/punctuation ii) next and previous 3 words/punctuations iii) whether the word/punctuation is special characters (';', '?', '.', ',', '!', ':', etc) or not iv) whether the word is title or not?
+(e.g.
+"The" is a title word but "the" is not) v) POS tags
+
+\textbf{Multilayer Perceptrons}: MLP is a feed forward artificial neural network model which maps set of input onto a set of appropriate output.
+This study utilized MLP with a nonlinear activation function (relu) and one hidden layer consisting of 128 hidden units.
+In order to prevent over-fitting, we implement dropout on fully connected layers, which randomly hide some neurons during the training phase \cite{srivastava2014dropout}.
+Dropout is also applied in the fully connected layers in BRNN and CRNN models.
+For a given sample with $2l$ words/punctuations, the first layer input of MLP was created by the summation of first $l$ word vectors and last $l$ word vectors.
 
@@
 
-\textbf{Conditional Random Field}: state-of-the-art supervised classification method proven to be effective for text categorization\cite{joachims1998text} for its ability to cope with high dimensional input feature space.
-SVM finds the hyperplane that maximizes the separation between the closest ``new segment'' and ``same segment'' training examples.
-We scaled our dataset to utilize linear SVM with optimal cost 512 in experiments.
+\textbf{Bidirectional Recurrent Neural Networks}: BRNN is a neural network designed to capture sequential patterns by considering both past and future inputs as well as complex relationships between input feature and output labels.
+The output of BRNN layer is computed as the summation of the forward RNN output with backward RNN output.
+Gated Recurrent Units (GRU) \cite{chung2014empirical} are a special type of RNN that are capable of handling variable size input sequence and have an internal memory that can be reset.
+Figure~\ref{fig:crnn} will represent an architecture of BRNN if convolution layer is removed.
 
-\textbf{Multilayer Perceptrons}: this method considers each training sample as a point in the input feature space.
-For a new test sample, Euclidean distance is calculated to find the k-nearest classified neighbors from the training set.
-Finally, the test sample is classified into a majority class of its k-nearest neighbors.
-We experimentally determined that best performance was achieved when the number of neighbors is 3.
-
-\textbf{Bidirectional Recurrent Neural Networks}: RNN is a neural network architecture designed to capture sequential patterns.
-When we predict the new segment point using RNNs, particular combinations of words in a sequence will indicate the change of MI behavior.
-Long Short Term Memory (LSTM) networks\cite{hochreiter1997long} are a special type of RNN that are capable of handling variable size input sequence and have an internal memory that can be reset.
-GRU\cite{chung2014empirical} is a variant of LSTM mathematically represented by the following formula:
+\textbf{Convolutional Recurrent Neural Networks}: CRNN is a deep neural network architecture consisting of 5 layers: 1) input layer 2) embedding layer 3) convolution layer with max pooling 4) BRNN layer 5) fully connected layer with dropout and sigmoid output.
 
@@
 
-\textbf{Convolutional Recurrent Neural Networks}: RNN is a neural network architecture designed to capture sequential patterns.
-When we predict the new segment point using RNNs, particular combinations of words in a sequence will indicate the change of MI behavior.
-Long Short Term Memory (LSTM) networks\cite{hochreiter1997long} are a special type of RNN that are capable of handling variable size input sequence and have an internal memory that can be reset.
-GRU\cite{chung2014empirical} is a variant of LSTM mathematically represented by the following formula:
-
-Our GRU model includes one hidden layer, output layer and input layer.
-We reset our model state after feeding each input sequence where input was given as one-hot encoding of word vector.
-Since LSTM and GRU use one-hot vector as input, results for these models are reported only when lexical and punctuation features are used.
-We experimentally determined that the best performance is achieved when the number of hidden units is 25, batch size is 1 and Adam\cite{kingma2014adam} is used for optimization.
-Source codes of this work will be available at https://github.com/teanalab/xxxxx upon acceptance of this paper for publication.
+A sequence of $m$ words is fed into the input layer which provides a $m \times n_e$ matrix after fetching the pre-trained word vectors.
+When POS tags are used in combination with word embedding, 10-dimensional POS vectors are concatenated with 300-dimensional word vectors to obtain new vectors $n_e = n_w + n_p$ of size 310.
+The 1D convolutional layer generates $m - h_c + 1$ new features with $h_c.n_e$ a filter of $h_c$ words and zero-padding on both sides of the text.
+A pooling operation was performed over time to find the most significant features in a region.
+After the convolutional layer, $m \times n_f$ are fed into bidirectional recurrent neural networks with 200 hidden units.
+Dropout is used to randomly drop 50\% of the neurons and passes them through a fully connected layer, where a sigmoid operation is calculated, giving us the probability of whether the word precedes a new segment or not.
+We experimentally determined that the best performance is achieved when filter length in convolution layer is 7, pool size is 3, batch size is 32, adam\cite{kingma2014adam} is used for optimization and the early stopping strategy is used with dropout to prevent over-fitting.
+Source codes of all experiments will be available at https://github.com/teanalab/xxxxx upon acceptance of this paper for publication.
 
@@
 Influence of the additional features is also consistent with the results in Tables~\ref{tab:result_boundary} and~\ref{tab:result_not_boundary}.
-Precision increases by 0.9\% and 0.2\%; recall increases by 0.9\% and 0.2\%; and F1-measure increases by 0.9\% and 0.3\% for SVM and KNN methods, respectively, when lexical feature is used in combination with punctuation and best model-based topic features.\\
+Precision increases by 0.9\% and 0.2\%; recall increases by 0.9\% and 0.2\%; and F1-measure increases by 0.9\% and 0.3\% for SVM and KNN methods, respectively, when lexical feature is used in combination with punctuation and best model-based topic features.
 
@@
 Finally, LSTM demonstrates the highest AUC value among all machine learning models, achieving 0.996 AUC, while GRU achieves only 0.2\% lower AUC than LSTM.
+\\
 

### 2018-07-28 plots/embedding-dimension.r

@@ -1,0 +1,29 @@
+#library
+library(ggplot2)
+library(reshape2)
+library(pROC)
+library(calibrate)
+
+# melt data
+wd = '/home/mehedi/teana/projects/eCoaching-Text-Segmentation/plots/'
+
+# melt data
+# print AUC values
+data <- read.csv(paste(wd,"parameters/embedding-dimension.csv", sep=''))
+
+# plot graph
+plot(range(50,320), range(0.795, 0.827), type="n", ylab = "F1-Measure", xlab = "vector size")
+lines(data$dim, data$pf1, lwd=1.5, lty=1, type='b', col='green', pch=18)
+points(x = 300, y = 0.825, col = "red", pch = 18, cex = 1, lwd=1.5)
+text(300, 0.825, labels = c('0.825'), pos = 3, cex = 0.7)
+textxy(data$dim, data$pf1, labs=data$pf1, pos=1,  cex = 0.7)
+
+lines(data$dim, data$tf1, lwd=1.5, lty=1, type='b', col='blue', pch=17)
+points(x = 300, y = 0.815, col = "black", pch = 17, cex = 1, lwd=1.5)
+text(300, 0.815, labels = c('0.815'), pos = 3, cex = 0.7)
+textxy(data$dim, data$tf1, labs=data$tf1, pos=1,  cex = 0.7)
+
+legend(222, 0.805, c("pre-trained GloVe", "trained GloVe", "pre-trained word2vec", "trained word2vec"), 
+       lty=c(1,1,1), col=c('blue', 'green', 'red', 'black'), cex=1, lwd=1.5)
+
+head(data)

### 2018-07-28 plots/length-mlp.r

@@ -1,0 +1,20 @@
+#library
+library(ggplot2)
+library(reshape2)
+library(pROC)
+library(calibrate)
+
+# melt data
+wd = '/home/mehedi/teana/projects/eCoaching-Text-Segmentation/plots/'
+
+# melt data
+# print AUC values
+data <- read.csv(paste(wd,"parameters/length.csv", sep=''))
+
+# plot graph
+plot(range(1,7), range(0.710, 0.775), type="n", ylab = "F1-Measure", xlab = "# of next or prior words/punctuation marks in a sample (l)")
+lines(data$n, data$fmeasure, lwd=1.5, lty=1, type='b', col='blue', pch=18)
+points(x = 2, y = 0.771, col = "red", pch = 1, cex = 2)
+textxy(data$n, data$fmeasure, labs=data$fmeasure, pos=3,  cex = 0.7)
+
+head(data)

### 2018-07-28 paper-amia-summit-2018/manuscript.tex
update methods section for spelling mistakes

@@
 The distributed representation of word allows capturing semantic, syntactic and morphological information from large unannotated corpora  \cite{pennington2014glove, mikolov2013distributed}.
-Since pre-trained word2vec word vectors provides the best results, in our experiment, we utilized pre-trained word2vec word vectors for words.
+Since pre-trained word2vec word vectors provide the best results, in our experiment, we utilized pre-trained word2vec word vectors for words.
 Trained word vectors were also utilized when punctuation marks are not available in the pre-trained word2vec word vectors.
@@
 
-\textbf{Multilayer Perceptrons}: MLP is a feed forward artificial neural network model which maps set of input onto a set of appropriate output.
+\textbf{Multilayer Perceptrons}: MLP is a feed-forward artificial neural network model which maps set of input onto a set of appropriate output.
 This study utilized MLP with a nonlinear activation function (relu) and one hidden layer consisting of 128 hidden units.
@@
 When POS tags are used in combination with word embedding, 10-dimensional POS vectors are concatenated with 300-dimensional word vectors to obtain new vectors $n_e = n_w + n_p$ of size 310.
-The 1D convolutional layer generates $m - h_c + 1$ new features with $h_c.n_e$ a filter of $h_c$ words and zero-padding on both sides of the text.
+The 1D convolutional layer generates $m - h_c + 1$ new features with $h_{c.n_e}$, a filter of $h_c$ words and zero-padding on both sides of the text.
 A pooling operation was performed over time to find the most significant features in a region.

### 2018-07-28 paper-amia-summit-2018/manuscript.tex
update methods section

@@
 Unlike a maximum entropy markov model which uses per-state exponential models for the conditional probabilities of next states given the current state, a CRF is a discriminative model of the conditional distribution of the entire sequence of labels given the observation sequence.
-A traditional linear-chain CRF model is defined as a conditional probability distribution $p(y|x)$ for output and input sequences, $y$ and $x$, by the following equation:
+A traditional linear-chain CRF model is defined as a conditional probability distribution $p(y|x)$ for output and input sequences, $y$ and $x$, by the Eq.
+\ref{Eq:crf}
 
 \begin{equation}
+\label{Eq:crf}
 p(y|x) = \frac{1}{Z_x}\exp{\left(\sum_{t=1}^{T}\sum_{k}^{}\lambda_k f_k(y_{t-1}, y_t, x, t)\right)} \\
@@
 For a given sample with $2l$ words/punctuations, the first layer input of MLP was created by the summation of first $l$ word vectors and last $l$ word vectors.
+Figure~\ref{fig:length-mlp} demostrates the performance of MLP model on e-Coaching email segmentation by varying value of $l$.
 
@@
 
-\textbf{Convolutional Recurrent Neural Networks}: CRNN is a deep neural network architecture consisting of 5 layers: 1) input layer 2) embedding layer 3) convolution layer with max pooling 4) BRNN layer 5) fully connected layer with dropout and sigmoid output.
+\textbf{Convolutional Recurrent Neural Networks}: CRNN is a deep neural networks architecture, shown in Figure~\ref{fig:crnn}, consisting of 5 layers: 1) input layer 2) embedding layer 3) convolution layer with max pooling 4) BRNN layer 5) fully connected layer with dropout and sigmoid output.
 

### 2018-07-28 paper-amia-summit-2018/manuscript.tex
update results section

@@
     \centering
-    \includegraphics[width=0.8\textwidth]{figures/segment-example.png}
+    \includegraphics[width=0.7\textwidth]{figures/segment-example.png}
     \caption{\textbf{Example of e-Coaching emails segmented into fragments that correspond to MI behaviors of an e-Coach and a patient}.}
@@
 A pooling operation was performed over time to find the most significant features in a region.
-After the convolutional layer, $m \times n_f$ are fed into bidirectional recurrent neural networks with 200 hidden units.
+After the convolutional layer, $m \times n_f$ inputs are fed into bidirectional recurrent neural networks with 200 hidden units.
 Dropout is used to randomly drop 50\% of the neurons and passes them through a fully connected layer, where a sigmoid operation is calculated, giving us the probability of whether the word precedes a new segment or not.
@@
 Our experimental results have two important directions.
-First, results are reported with respect to ``new segment'' and ``same segment'' classes as well as their weighted average in Table~\ref{tab:result_base}.
-Second, classification performance of different machine learning methods are outlined in Tables~\ref{tab:result_boundary}, \ref{tab:result_not_boundary} and \ref{tab:result_weighted_avg} when topic features are extracted with different probabilistic generative latent variable models.\\
+First, results are reported with respect to ``new segment'' class as well as weighted average over ``new segment'' and ``same segment'' classes in Table~\ref{tab:result_base}.
+Second, classification performance of different machine learning methods are outlined in Table~\ref{tab:result_weighted_avg} when word embeddings or lexical features are used in combination with punctuation and POS features.\\
 
@@
 
-As follows from Table~\ref{tab:result_base}, KNN outperforms all other models in terms of precision and F1-measure achieving 0.808 precision with 0.728 F1-measure for new segment detection.
-KNN also shows superior performance in all performance metrics for ``same segment'' class and weighted average over ``new segment'' and ``same segment'' classes.
-NB demonstrates the lowest performance among all models in terms of all performance metrics.
-On the other hand, SVM has the highest recall of 0.705 when only lexical features are used to identify ``new segment''.
-Results indicate that performance of classifiers is remarkably higher for ``same segment'' class compared to ``new segment'' class, which is expected since 96.71\% instances belong to ``same segment'' categories.
-For example, KNN achieves 22.40\%, 50.07\% and 36.26\% higher precision, recall and F1-measure, respectively, in ``same segment'' identification compared to ``new segment'' detection.
-When lexical features are used in combination with punctuation mark, LSTM demonstrates the best performance in terms of precision and F1-measure for ``new segment'' identification while GRU shows 2.64\% higher recall than LSTM.
-GRU and LSTM show exactly same performance in terms of all performance metrics for ``same segement'' and overall classifications.
-\\
+As follows from Table~\ref{tab:result_base}, CRNN outperforms all other models in terms of recall and F1-measure achieving 0.797 recall with 0.785 F1-measure for new segment detection.
+CRNN also shows superior performance in all performance metrics for weighted average over ``new segment'' and ``same segment'' classes.
+BRNN demonstrates the lowest performance among all models in terms of precision and F1-Measure.
+On the other hand, MLP has the highest precision of 0.836 when only word embeddings or lexical features are used to identify ``new segment''.
+CRF achieves 0.733 F1-Measure in new segment class and 0.984 F1-Measure in overall, which demonstrates the second highest performance to identify ``new segment'' as well as a weighted average over both classes.
+Experimental results indicate that performance of classifiers as a weighted average over both classes is remarkably higher compared to ``new segment'' class, which is expected since 96.95\% instances belong to the ``same segment'' categories and 99.3\% of them correctly classified.
+For example, CRNN achieves 27.23\%, 23.71\% and 25.61\% higher precision, recall and F1-measure, respectively, compared to ``new segment'' detection.
+
+Table~\ref{tab:result_weighted_avg} summarizes the results of all models for segmentation of e-Coaching emails when word embeddings or lexical features are used in combination with punctuation and POS features.
+Similar to results in Tables~\ref{tab:result_base}, CRNN demonstrates the highest performance among all methods achieving 0.864 recall with 0.825 F1-Measure for ``new segment'' class and 0.990 precision with 0.989 recall and F1-Measure for overall.
+BRNN shows the lowest and CRF shows the second highest performance for email segmentation among all methods.
+We observed that classification performance significantly improved for ``new segment'' class when word embeddings or lexical features are used in combination with punctuation and POS features.
+Precision increases by 3.96\%, -2.63\%, 12.71\% and 1.81\%; recall increases by 11.72\%, 23.44\%, 20.59\% and 8.41\%; and F1-measure increases by 8.05\%, 11.1\%, 16.22\% and 5.1\% for CRF, MLP, BRNN and CRNN methods, respectively, in new segment detection.
+Similarly, precision increases by 0.51\%, 0.51\%, 0.82\% and 0.41\%; recall increases by 0.41\%, 0.41\%, 0.72\% and 0.3\%; and F1-measure increases by 0.41\%, 0.51\%, 0.82\% and 0.3\% for CRF, MLP, BRNN and CRNN methods, respectively, in weighted average over ``new segment'' and ``same segment'' classes.\\
 
@@
 
-Table~\ref{tab:result_weighted_avg} summarizes the weighted average results of the models for segmentation of e-Coaching emails.
-Overall, SVM has the best performance across all topic model-based features in terms of all performance metrics, achieves 0.990 precision, recall and F1-measure when DL-LDA-based topic features are used.
-Similar to results in Tables~\ref{tab:result_base}, \ref{tab:result_boundary} and \ref{tab:result_not_boundary}, NB demonstrates the lowest performance among all methods while best performance achieves with LCA-based topic features compared to other model-based topic features.
-Influence of the additional features is also consistent with the results in Tables~\ref{tab:result_boundary} and~\ref{tab:result_not_boundary}.
-Precision increases by 0.9\% and 0.2\%; recall increases by 0.9\% and 0.2\%; and F1-measure increases by 0.9\% and 0.3\% for SVM and KNN methods, respectively, when lexical feature is used in combination with punctuation and best model-based topic features.
-
 \section*{Discussion}
@@
 This results also indicate that segmentation performance might be improved by adding additional relevant features.
-\\
 

### 2018-07-28 paper-amia-summit-2018/manuscript.tex
update discussion section

@@
 This study is the first effort to evaluate the automatic segmentation of e-Coaching emails.
-Experimental results indicate that SVM is the best model among all machine learning methods considered for this study.
-SVM achieved 0.990 F1-measure in overall, 0.845 and 0.995 F1-measures for detecting ``new segment'' and ``same segment'', respectively.
-The robust performance of SVM provides the evidence that machine learning models are capable to learn conceptual information from clinical exchange.
-It also indicates that topic features are more important than lexical features, even when deep learning methods are employed.
+Experimental results indicate that CRNN is the best model among all machine learning methods considered for this study.
+CRNN achieved 0.989 F1-measure in overall and 0.825 F1-measure for detecting ``new segment''.
+The robust performance of CRNN provides the evidence that machine learning models are capable to learn conceptual information from the clinical exchange.
+It also indicates that punctuation and POS features are more important than word embeddings or lexical features, even when deep learning methods are employed.
 Although the domain of this study was intentionally quite small, we believe that our study is not limited to the e-Coaching domain and it can be successfully applied to other domains, in which discourse segmentation is a preliminary step for annotation.
 
-Punctuation mark and topic model-based features made a significant improvement in performance of all machine learning methods.
-Among all topic model-based feature, the DL-LDA-based topic feature provides the highest performance in ``new segment'' and ``same segment'' classification as well as weighted average over ``new segment'' and ``same segment'' classification results.
-In nearly all cases, ML methods perform better, when lexical features are used in combination with punctuation and topic model-based features.
+Punctuation mark and POS features made a significant improvement in the performance of machine learning and all deep learning methods.
+In all cases, every method performs better, when word embeddings or lexical features are used in combination with punctuation and POS features.
 This results also indicate that segmentation performance might be improved by adding additional relevant features.
+\\
 
@@
 
-In this paper, results are reported by each class to avoid confusion regarding overall model performance due to severe class imbalance.
+In this paper, results are reported by ``new segment'' class as well as weighted average over both classes to avoid confusion regarding overall model performance due to severe class imbalance.
 In addition, standard metrics: precision, recall and F1-measure were used to eliminate doubt about the model performance since accuracy is misleading for imbalanced datasets.
-AUC is also utilized due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets\cite{hu2015kernelized}, which are demonstrated by the ROC curves in Figure~\ref{fig:roc-curves}.
-Although NB provides lowest classification results, it shows moderate AUC values, achieving 0.978 AUC when only lexical features are used and 0.977 AUC when a combination of lexical, punctuation and topic features are used.
-On the other hand, KNN and SVM exhibit 0.972 and 0.986 AUCs when only lexical features are used; and 0.966 and 0.980 AUCs when a combination of lexical, punctuation and topic model-based features are used.
-Finally, LSTM demonstrates the highest AUC value among all machine learning models, achieving 0.996 AUC, while GRU achieves only 0.2\% lower AUC than LSTM.
+AUC is also utilized due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets\cite{hu2015kernelized}, which are demonstrated by the Table~\ref{tab:result_roc} and ROC curves in Figure~\ref{fig:roc-curves}.
+Although CRF provides the second highest classification results, it shows best AUC values, achieving 0.966 AUC when only lexical features are used and 0.994 AUC when a combination of lexical, punctuation and POS features are used.
+Influence of the punctuation and POS features is also consistent in AUC values, which increase by 2.9\%, 4.31\%, 6.63\% and 2.18\% for CRF, MLP, BRNN and CRNN methods, respectively.
 \\
@@
 
-We observed moderate performance of RNN, in particular, LSTM and GRU for the text segmentation.
-We believe that RNN will perform better if a larger dataset is utilized.
-In this study, we employed 3,138 examples of boundary case which limit to achieve the best performance.
-We also observed that GRU performs better than LSTM, which was observed in previous studies\cite{chung2014empirical}.
-
-Although punctuation mark plays an important role in segmentation boundary detection and large numbers of errors were encountered by the false positive of boundary identification.
+Although punctuation mark plays an important role in segmentation boundary detection, large numbers of errors were encountered by the false positive of boundary identification.
 For example, a text segment from an e-Coaching email \textit{``A typical day in regards to fruit and vegetable has me eating about a serving at breakfast (our caf has cut up fruit) and then maybe a piece of fruit later in the day or as a snack.
@@
 However, this study segmented clinical exchange into groups of MI behaviors which will significantly reduce the amount of resource and time required to segment clinical exchange manually.
-Furthermore, these segmentation models could be integrated with auto coding classifiers to build a software pipeline for automated annotation of exchanges.
+Furthermore, these segmentation models could be integrated with auto coding classifiers to build a software pipeline for automated annotation of clinical exchanges.
 

### 2018-07-30 paper-amia-summit-2018/manuscript.tex
update paper

@@
 
-Qualitative coding of motivational interviews with pre-defined codes has been traditionally performed manually by trained annotators, which is a tedious and resource-intensive process that involves several iterations of reading, comprehension and interpretation of interview transcripts.
+Qualitative coding of motivational interviews with pre-defined codes, has been traditionally performed manually by trained annotators which is a tedious and resource-intensive process that involves several iterations of reading, comprehension and interpretation of interview transcripts.
 Rapidly developing computational technologies, specifically, machine learning methods, offer a unique opportunity to accelerate this process.
@@
 The unstructured nature of e-Coaching exchanges poses a unique set of challenges for their qualitative analysis.
+Usually, email segments are automatically code with the Minority Youth-Sequential Coding of Process Exchanges (MYSCOPE \cite{carcone2013provider}), a qualitative coding scheme to characterize client-coach communication during email exchange.
 A significant barrier to fully automating the behavior coding process of e-Coaching emails is their segmentation into textual fragments that correspond to distinct communication behaviors.
@@
 
-%Our experiments show that these differences lead to significant improvement in the segmentation of e-Coaching emails.
-
 Recently, an online clinical intervention called MENU GenY \cite{alexander2017motivations} (Making Effective Nutrition Choices for Generation Y) was proposed and evaluated.
@@
 The goal of MENU GenY was to develop a better coding dictionary among GenY to improve eating habits.
-However, segmentation of clinical conversation in the context of electronically delivered interventions, in particular, segmentation of clinical interaction text into groups of MI behaviors, is still performed manually, which slows down qualitative analysis of these interventions.
+However, segmentation of clinical exchange in the context of electronically delivered interventions, in particular, segmentation of clinical interaction text into groups of MI behaviors, is still performed manually, which slows down qualitative analysis of these interventions.
 This study proposes several machine learning methods to address this problem and the authors are unaware of any other work that focused on the same problem.
@@
 Each session represents an MI intervention delivered via email.
-We formulate the segmentation task as a binary classification problem where each word or punctuation marks annotated with a label 1 or 0, to indicate whether it precedes a new segment or not.
+We formulate the segmentation task as a binary classification problem where each word or punctuation mark annotated with a label 1 or 0, to indicate whether it precedes a new segment or not.
 In total, we obtained 95,777 words and 7,140 punctuation marks, which include 3,138 ``new segment'' and 99,779 ``same segment'' instances.
 In this study, we experimented with traditional machine learning method, CRF and deep learning methods, such as MLP, BRNN and CRNN.
-For MLP models, samples were created with $2l$ words/punctuation marks at each step in a word sequence, where a sample contains next $l$ words/punctuation marks and prior $l$ words/punctuation marks including current word/punctuation mark.
-Each sample is classified into either ``new segment'' or ``same segment'' based on whether the current word/punctuation mark precedes a new segment or not.
-For CRF, BRNN and CRNN models, an email was taken as input sequence, such that POS tags and word embedding of each word or punctuation marks were used as input and binary labels (1 or 0) corresponding to ``new segment'' and ``same segment'' classification decision were considered as the output of the model at each step.
-In the gold standard, words within the same segment were assigned the label of 0 and the last word or punctuation mark of a segment were assigned the label of 1.
+For MLP models, samples were created with $2l$ words/punctuations at each step in a word sequence, where a sample contains next $l$ words/punctuations and prior $l$ words/punctuations including current word/punctuation.
+Each sample is classified into either ``new segment'' or ``same segment'' based on whether the current word/punctuation precedes a new segment or not.
+For CRF, BRNN and CRNN models, an email was taken as input sequence, such that POS tags and word embedding of each word or punctuation mark were used as input and binary labels (1 or 0) corresponding to ``new segment'' and ``same segment'' classification decision were considered as the output of the model at each step.
+In the gold standard, words or punctuations within the same segment were assigned the label of 0 and the last word or punctuation mark of a segment were assigned the label of 1.
 
@@
 
-We utilized three types of features in conjunction with CRF, MLP, BRNN and CRNN methods: word embeddings, punctuation and POS features.
-Since POS tags have been shown to be effective semantic abstractions of individual words, we used topic features for our experiment \cite{liu2005using,treviso2017sentence}.
+We utilized three types of features in conjunction with CRF, MLP, BRNN and CRNN methods: word embeddings or lexical features, punctuation and POS features.
+Since POS tags have been shown to be effective semantic abstractions of individual words, we used POS features for our experiment \cite{liu2005using,treviso2017sentence}.
 To extract our POS features, we tag the e-Coaching emails using the NLTK POS tagger \cite{bird2009natural}.
@@
 Since pre-trained word2vec word vectors provide the best results, in our experiment, we utilized pre-trained word2vec word vectors for words.
-Trained word vectors were also utilized when punctuation marks are not available in the pre-trained word2vec word vectors.
+Trained word vectors were also utilized when rare words or punctuation marks are not available in the pre-trained word2vec word vectors.
 Figure~\ref{fig:embedding-dimension} illustrates the performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and trained word embeddings with GloVe and word2vec models.
+It was observed that best performance achieved with pre-trained 300 dimensional word2Vec word vectors when three types of features are used together.
+Therefore, we report our results with word2vec 300 dimensional word vectors for all deep learning models used in this study.
+CRF utilized lexical features which represent each word in an e-Coaching email.
 
@@
     \centering
-    \includegraphics[width=0.6\textwidth]{figures/embedding-dimension.eps}
+    \includegraphics[width=0.5\textwidth]{figures/embedding-dimension.eps}
     \caption{\textbf{Performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and trained word embeddings with GloVe and word2vec models.}}
@@
 
-In the present study, we experimented with four different classifiers, including one traditional machine learning method, CRF and three neural networks models.
-Since neural networks provide a flexible architecture for constructing and synthesizing complex models, we take the advantage of this flexibility to test MLP, BRNN and CRNN models for the task of segmentation of e-Coaching emails.
+In the present study, we experimented with four different classifiers, including one traditional machine learning method, CRF and three deep learning models.
+Since deep learning provides a flexible architecture for constructing and synthesizing complex models, we take the advantage of this flexibility to test MLP, BRNN and CRNN models for the task of segmentation of e-Coaching emails.
 
@@
 The optimal output sequence $y^*$ for an input sequence $x$, $y^* = {arg\,max}_y p(y|x)$, is obtained efficiently by the Viterbi algorithm.
-In our experiments, the following features were utilized with CRF models: i) current word/punctuation ii) next and previous 3 words/punctuations iii) whether the word/punctuation is special characters (';', '?', '.', ',', '!', ':', etc) or not iv) whether the word is title or not?
+In our experiments, the following features were utilized with CRF models: i) current word/punctuation ii) next and previous 3 words/punctuations iii) whether the word/punctuation is special characters (';', '?', '.', ',', '!', ':', etc) or not iv) whether the word is title word or not?
 (e.g.
-"The" is a title word but "the" is not) v) POS tags
+``The'' is a title word but ``the'' is not) v) POS tags
 
-\textbf{Multilayer Perceptrons}: MLP is a feed-forward artificial neural network model which maps set of input onto a set of appropriate output.
+\textbf{Multilayer Perceptrons}: MLP is a feed-forward artificial neural networks which map set of input onto a set of appropriate output.
 This study utilized MLP with a nonlinear activation function (relu) and one hidden layer consisting of 128 hidden units.
@@
 Figure~\ref{fig:length-mlp} demostrates the performance of MLP model on e-Coaching email segmentation by varying value of $l$.
+Based on the results in Figure~\ref{fig:length-mlp}, MLP results were reported with $l = 2$, since best performance of MLP models were observed when $l = 2$.
 
@@
     \centering
-    \includegraphics[width=0.60\textwidth]{figures/length-mlp.eps}
+    \includegraphics[width=0.5\textwidth]{figures/length-mlp.eps}
     \caption{\textbf{Performance of MLP model on e-Coaching email segmentation by varying the length of sample size.}}
@@
 The output of BRNN layer is computed as the summation of the forward RNN output with backward RNN output.
-Gated Recurrent Units (GRU) \cite{chung2014empirical} are a special type of RNN that are capable of handling variable size input sequence and have an internal memory that can be reset.
+Gated Recurrent Units (GRU) \cite{chung2014empirical} are utilized as RNN that are capable of handling variable size input sequence and have an internal memory that can be reset.
 Figure~\ref{fig:crnn} will represent an architecture of BRNN if convolution layer is removed.
@@
 A pooling operation was performed over time to find the most significant features in a region.
-After the convolutional layer, $m \times n_f$ inputs are fed into bidirectional recurrent neural networks with 200 hidden units.
-Dropout is used to randomly drop 50\% of the neurons and passes them through a fully connected layer, where a sigmoid operation is calculated, giving us the probability of whether the word precedes a new segment or not.
-We experimentally determined that the best performance is achieved when filter length in convolution layer is 7, pool size is 3, batch size is 32, adam\cite{kingma2014adam} is used for optimization and the early stopping strategy is used with dropout to prevent over-fitting.
-Source codes of all experiments will be available at https://github.com/teanalab/xxxxx upon acceptance of this paper for publication.
+After applying $n_f=100$ different filters on convolutional layer, $m \times n_f$ inputs are generated and fed into bidirectional recurrent neural networks, which contain 200 hidden units.
+Then, dropout is used to randomly drop 50\% of the neurons and passes them through a fully connected layer, where a sigmoid operation is calculated, giving us the probability of whether the word or punctuation precedes a new segment or not.
+We experimentally determined that the best performance is achieved with 5 folds corss-validation when filter length in convolution layer is 7, pool size is 3, batch size is 32, adam\cite{kingma2014adam} is used for optimization and the early stopping strategy is used with dropout to prevent over-fitting.
+To reuse of our source codes for other research studies, source codes of all experiments will be publicly available at https://github.com/teanalab/xxxxx upon acceptance of this paper for publication.
 
@@
 BRNN demonstrates the lowest performance among all models in terms of precision and F1-Measure.
-On the other hand, MLP has the highest precision of 0.836 when only word embeddings or lexical features are used to identify ``new segment''.
+On the other hand, MLP has the highest precision of 0.836 when word embeddings or lexical features are used to identify ``new segment''.
 CRF achieves 0.733 F1-Measure in new segment class and 0.984 F1-Measure in overall, which demonstrates the second highest performance to identify ``new segment'' as well as a weighted average over both classes.
-Experimental results indicate that performance of classifiers as a weighted average over both classes is remarkably higher compared to ``new segment'' class, which is expected since 96.95\% instances belong to the ``same segment'' categories and 99.3\% of them correctly classified.
-For example, CRNN achieves 27.23\%, 23.71\% and 25.61\% higher precision, recall and F1-measure, respectively, compared to ``new segment'' detection.
+Experimental results indicate that performance of all classifiers as a weighted average over both classes is remarkably higher compared to ``new segment'' class, which is expected since 96.95\% instances belong to the ``same segment'' categories and 99.3\% of them correctly classified.
+For example, as a weighted average over ``new segment'' and ``same segment'' classes, CRNN achieves 27.23\%, 23.71\% and 25.61\% higher precision, recall and F1-measure, respectively, compared to ``new segment'' detection.
 
@@
 Similar to results in Tables~\ref{tab:result_base}, CRNN demonstrates the highest performance among all methods achieving 0.864 recall with 0.825 F1-Measure for ``new segment'' class and 0.990 precision with 0.989 recall and F1-Measure for overall.
-BRNN shows the lowest and CRF shows the second highest performance for email segmentation among all methods.
+BRNN and CRF show the lowest and second highest performance, respectively, for email segmentation among all methods.
 We observed that classification performance significantly improved for ``new segment'' class when word embeddings or lexical features are used in combination with punctuation and POS features.
@@
 CRNN achieved 0.989 F1-measure in overall and 0.825 F1-measure for detecting ``new segment''.
-The robust performance of CRNN provides the evidence that machine learning models are capable to learn conceptual information from the clinical exchange.
-It also indicates that punctuation and POS features are more important than word embeddings or lexical features, even when deep learning methods are employed.
+The robust performance of CRNN provides the evidence that deep learning models are capable to learn MI behaviors from the clinical exchange.
+It also indicates that punctuation and POS features are important with word embeddings or lexical features, even when deep learning methods are employed.
 Although the domain of this study was intentionally quite small, we believe that our study is not limited to the e-Coaching domain and it can be successfully applied to other domains, in which discourse segmentation is a preliminary step for annotation.
@@
     \includegraphics[width=1.0\textwidth]{figures/roc-curves.eps}
-    \caption{\textbf{Receiver operating characteristic curves showing the performance of binary classifiers for the segmentation of e-Coaching text when only word embeddings/lexical (left) and combination of word embeddings/lexical, POS and punctuation features (right) are used}.}
+    \caption{\textbf{Receiver operating characteristic curves showing the performance of binary classifiers for the segmentation of e-Coaching text when only word embeddings/lexical features (left) and combination of word embeddings/lexical, punctuation and POS features (right) are used}.}
     \label{fig:roc-curves}
@@
 
-In this paper, results are reported by ``new segment'' class as well as weighted average over both classes to avoid confusion regarding overall model performance due to severe class imbalance.
-In addition, standard metrics: precision, recall and F1-measure were used to eliminate doubt about the model performance since accuracy is misleading for imbalanced datasets.
-AUC is also utilized due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets\cite{hu2015kernelized}, which are demonstrated by the Table~\ref{tab:result_roc} and ROC curves in Figure~\ref{fig:roc-curves}.
-Although CRF provides the second highest classification results, it shows best AUC values, achieving 0.966 AUC when only lexical features are used and 0.994 AUC when a combination of lexical, punctuation and POS features are used.
-Influence of the punctuation and POS features is also consistent in AUC values, which increase by 2.9\%, 4.31\%, 6.63\% and 2.18\% for CRF, MLP, BRNN and CRNN methods, respectively.
-\\
-
 \begin{table}[ht]
 \centering
-\caption{AUC values of all classifiers when only word embeddings/lexical and combination of word embeddings/lexical, POS and punctuation features are used.
+\caption{AUC values of all classifiers when only word embeddings/lexical features and combination of word embeddings/lexical, punctuation and POS features are used.
 Highest value for each experiment across all models is highlighted in boldface.}
@@
 
-Although punctuation mark plays an important role in segmentation boundary detection, large numbers of errors were encountered by the false positive of boundary identification.
+In this paper, results are reported by ``new segment'' class as well as weighted average over both classes to avoid confusion regarding overall model performance due to severe class imbalance.
+In addition, standard metrics: precision, recall and F1-measure were used to eliminate doubt about the model performance since accuracy is misleading for imbalanced datasets.
+AUC is also utilized due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets\cite{hu2015kernelized}, which are demonstrated by the Table~\ref{tab:result_roc} and ROC curves in Figure~\ref{fig:roc-curves}.
+Although CRF provides the second highest classification results, it shows best AUC values, achieving 0.966 AUC when only lexical features are used and 0.994 AUC when a combination of lexical, punctuation and POS features are used.
+Influence of the punctuation and POS features is also consistent in AUC values, which increase by 2.9\%, 4.31\%, 6.63\% and 2.18\% for CRF, MLP, BRNN and CRNN methods, respectively.
+
+Although punctuation mark plays an important role in segmentation boundary detection, a few errors were encountered by the presence of punctuation marks in boundary identification.
 For example, a text segment from an e-Coaching email \textit{``A typical day in regards to fruit and vegetable has me eating about a serving at breakfast (our caf has cut up fruit) and then maybe a piece of fruit later in the day or as a snack.
@@
 We found out that CRNN provides the best performance for the segmentation of text in terms of all performance metrics.
-Manual segmentation of e-Coaching data is very resource-intensive and time-consuming task, which can significantly decrease the time and effort required to develop an effective behavioral intervention.
+Manual segmentation of e-Coaching data is a very resource-intensive and time-consuming task, which can significantly decrease the time and effort required to develop an effective behavioral intervention.
 Our proposed methods can help to identify individual text segments, which can be annotated directly with a classification model.

### 2018-07-30 paper-amia-summit-2018/manuscript.tex
update paper

@@
 A critical step towards automated annotation of e-Coaching communication exchanges is segmentation of emails into textual fragments that correspond to MI behaviors.
-In this work, we formulate this task as a classification problem and propose lexical, punctuation and part-of-speech features to address it.
+In this work, we formulate this task as a classification problem and propose word embedding, punctuation and part-of-speech features to address it.
 We experimented both with traditional machine learning method, conditional random fields (CRF) and deep learning methods, such as multilayer perceptrons (MLP), bidirectional recurrent neural networks (BRNN) and convolutional recurrent neural networks (CRNN).
@@
 The emergence of e-Health technologies opened up new ways to deliver a variety of behavioral interventions to any demographic group of patients in any geographical location.
-Motivational interviewing (MI), an evidence-based communication technique to increase intrinsic motivation and self-efficacy for behavior change\cite{miller2012motivational,miller2009ten,miller2009toward}, is one type of these interventions.
+Motivational interviewing (MI), an evidence-based communication technique to increase intrinsic motivation and self-efficacy for behavior change,\cite{miller2012motivational,miller2009toward} is one type of these interventions.
 MI sessions are generally aimed at eliciting ``change talk'', or statements of intrinsic motivation about patients' own desire, ability, reasons, need for and commitment to behavior change, which have been established by previous research\cite{apodaca2009mechanisms} as a reliable mediator of health behavior change.
@@
 
-Qualitative coding of motivational interviews with pre-defined codes, has been traditionally performed manually by trained annotators which is a tedious and resource-intensive process that involves several iterations of reading, comprehension and interpretation of interview transcripts.
+Qualitative coding of motivational interviews with pre-defined codes has been traditionally performed manually by trained annotators, which is a tedious and resource-intensive process that involves several iterations of reading, comprehension and interpretation of interview transcripts.
 Rapidly developing computational technologies, specifically, machine learning methods, offer a unique opportunity to accelerate this process.
-In particular, machine learning methods have been successfully applied to a variety of analytical tasks involving textual data, such as classification\cite{nigam2000text} and sentiment analysis\cite{wang2012baselines}.
-In our previous work, we examined the utility of machine learning methods for automated annotation \cite{hasan2016study,kotov2015interpretable} and analysis \cite{hasan2018predicting} of in-person MI sessions.
-Specifically, we demonstrated that machine learning methods can be utilized for annotation of MI transcripts according to a simple communication code scheme with the accuracy comparable to human coders\cite{hasan2016study}.
-Experimental data utilized in these studies, however, were prepared by transcribing audio conversations, which were clearly segmented into utterances by a counselor, a patient and a caregiver.
+In particular, machine learning methods have been successfully applied to a variety of analytical tasks involving textual data, such as classification\cite{nigam2000text} and sentiment analysis.\cite{wang2012baselines} In our previous work, we examined the utility of machine learning methods for automated annotation \cite{hasan2016study,kotov2015interpretable} and analysis \cite{hasan2018predicting} of in-person MI sessions.
+Specifically, we demonstrated that machine learning methods can be utilized for annotation of MI transcripts according to a simple communication code scheme with the accuracy comparable to human coders.\cite{hasan2016study} Experimental data utilized in these studies, however, were prepared by transcribing audio conversations, which were clearly segmented into utterances by a counselor, a patient and a caregiver.
 
@@
 Specifically, the e-Coaches involved in this study used emails to communicate motivation-enhancing messages that encourage healthy eating among GenY adolescents.
-e-Coaching data is comprised of email responses, which are free-text documents, unlike more traditional dyadic clinical interviews that are naturally segmented into utterances due to their conversational nature.
+E-coaching data is comprised of email responses, which are free-text documents, unlike more traditional dyadic clinical interviews that are naturally segmented into utterances due to their conversational nature.
 
 The unstructured nature of e-Coaching exchanges poses a unique set of challenges for their qualitative analysis.
-Usually, email segments are automatically code with the Minority Youth-Sequential Coding of Process Exchanges (MYSCOPE \cite{carcone2013provider}), a qualitative coding scheme to characterize client-coach communication during email exchange.
+Usually, email segments are automatically coded with the Minority Youth-Sequential Coding of Process Exchanges (MYSCOPE), \cite{carcone2013provider} a qualitative coding scheme to characterize patient-coach communication during email exchanges.
 A significant barrier to fully automating the behavior coding process of e-Coaching emails is their segmentation into textual fragments that correspond to distinct communication behaviors.
@@
 The goal of this research study is to assess the applicability of machine learning methods for automated segmentation of e-Coaching emails into textual fragments corresponding to individual behaviors, which is the first step of the coding process of e-Coaching communications.
-In particular, we introduced lexical, punctuation and  part-of-speech (POS) features and experimented with both traditional supervised machine learning method, linear-chain conditional random fields (CRF) and deep learning methods, such as multilayer perceptrons (MLP), bidirectional recurrent neural networks (BRNN) and convolutional recurrent neural networks (CRNN), to find the best performing method and feature combination.
+In particular, we introduced word embedding or lexical, punctuation and  part-of-speech (POS) features and experimented with both traditional supervised machine learning method, linear-chain conditional random fields (CRF)\cite{lafferty2001conditional} and deep learning methods, such as multilayer perceptrons (MLP),\cite{rumelhart1986learning} bidirectional recurrent neural networks (BRNN)\cite{schuster1997bidirectional} and convolutional recurrent neural networks (CRNN),\cite{treviso2017sentence} to find the best performing method and feature combination.
 
 Relevant previous work in the biomedical domain primarily focused on segmentation of text into sections and headers\cite{apostolova2009automatic,denny2009evaluation,tepper2012statistical,cho2002text} or sentence boundary detection\cite{griffis2016quantitative,kreuzthaler2015detection,treviso2017sentence,fraser2015sentence}.
-Apostolova et al.\cite{apostolova2009automatic} applied SVM along with word-vector cosine similarity metric combined with several heuristics to segment clinical reports into sections, such as demographics, history, procedure, finding and impression.
-After identification of each line in the document, Tepper et al.
-\cite{tepper2012statistical} trained Maximum Entropy models for section classification.
-Denny et al.\cite{denny2009evaluation} proposed a SecTag algorithm, which combined natural language processing techniques, terminology-based rules and Naive Bayes classifier to identify the sections and headers that achieved 99\% recall with 95.6\% precision.
-On the other hand, SVM based on prosodic and part of speech features \cite{kreuzthaler2015detection} and recurrent convolutional neural networks using word embeddings \cite{griffis2016quantitative} were utilized for detecting sentence boundaries.
+Apostolova et al.
+applied SVM along with word-vector cosine similarity metric combined with several heuristics to segment clinical reports into sections, such as demographics, history, procedure, finding and impression.\cite{apostolova2009automatic} After identification of each line in the document, Tepper et al.
+trained Maximum Entropy models for section classification.\cite{tepper2012statistical} Denny et al.
+proposed a SecTag algorithm, which combined natural language processing techniques, terminology-based rules and a Naive Bayes classifier to identify the sections and headers that achieved 99\% recall with 95.6\% precision.
+\cite{denny2009evaluation} On the other hand, SVM based on prosodic and part of speech features \cite{kreuzthaler2015detection} and recurrent convolutional neural networks using word embeddings \cite{griffis2016quantitative} were utilized for detecting sentence boundaries.
 Liu et al.
-have shown that a linear chain CRF performed better than the hidden markov model and Maxent on the NIST sentence boundary detection \cite{liu2005using}.
-Recently, Treviso et al.
-proposed a recurrent convolutional neural network for automatic identification of sentence boundary in a neuropsychological text in Portuguese language \cite{treviso2017sentence}.
-Their proposed model combined prosodic and part of speech features with pre-trained GloVe word vectors \cite{pennington2014glove}, which achieved 0.74 F1-score in sentence boundary detection in the Brazilian constitution dataset.
+have shown that a linear chain CRF performed better than the hidden markov model and Maxent on the NIST sentence boundary detection.\cite{liu2005using} Recently, Treviso et al.
+proposed a recurrent convolutional neural network for automatic identification of sentence boundary in a neuropsychological text in Portuguese language.\cite{treviso2017sentence} Their proposed model combined prosodic and part of speech features with pre-trained GloVe word vectors,\cite{pennington2014glove} which achieved 0.74 F1-score in sentence boundary detection on the Brazilian constitution dataset.
 
 In this study, we adopt the model proposed by Treviso et al.
-\cite{treviso2017sentence} and apply it differently from the aforementioned studies.
-The main difference is that our study focuses on segmentation of e-Coaching emails as distinct from determining the sentence boundary.
+\cite{treviso2017sentence} and rename it as convolutional recurrent neural networks to apply the model differently from the aforementioned studies.
+The main difference is that our study focuses on the segmentation of e-Coaching emails as distinct from determining the sentence boundary.
 Another important difference is that punctuation features were utilized in combination with word embeddings and part-of-speech features.
-Furthermore, we train our model with pre-trained word2vec word vectors \cite{mikolov2013distributed}, instead of using GloVe word vectors.
+Furthermore, we train our model with pre-trained word2vec word vectors,\cite{mikolov2013distributed} instead of using GloVe word vectors.
 Segmentation of e-Coaching emails is also different from a traditional shallow discourse analysis of conversations \cite{galley2003discourse} in that the focus is on segmentation, rather than on determining the types of transitions between the utterances or assigning utterances to speakers.
 
-Recently, an online clinical intervention called MENU GenY \cite{alexander2017motivations} (Making Effective Nutrition Choices for Generation Y) was proposed and evaluated.
-MENU GenY is a technology-based public health intervention that relies on personalized e-coaching to encourage increased fruit and vegetable intake among young adults, aged 21-30.
+Recently, an online clinical intervention called MENU GenY (Making Effective Nutrition Choices for Generation Y) was proposed and evaluated.
+\cite{alexander2017motivations} MENU GenY is a technology-based public health intervention that relies on personalized e-coaching to encourage increased fruit and vegetable intake among young adults, aged 21-30.
 The goal of MENU GenY was to develop a better coding dictionary among GenY to improve eating habits.
@@
 In this study, we experimented with traditional machine learning method, CRF and deep learning methods, such as MLP, BRNN and CRNN.
-For MLP models, samples were created with $2l$ words/punctuations at each step in a word sequence, where a sample contains next $l$ words/punctuations and prior $l$ words/punctuations including current word/punctuation.
-Each sample is classified into either ``new segment'' or ``same segment'' based on whether the current word/punctuation precedes a new segment or not.
+For MLP models, samples were created with $2l$ words or punctuations at each step in a word sequence, where a sample contains next $l$ words or punctuations and prior $l$ words or punctuations including current wordor punctuation.
+Each sample is classified into either ``new segment'' or ``same segment'' based on whether the current word or punctuation precedes a new segment or not.
 For CRF, BRNN and CRNN models, an email was taken as input sequence, such that POS tags and word embedding of each word or punctuation mark were used as input and binary labels (1 or 0) corresponding to ``new segment'' and ``same segment'' classification decision were considered as the output of the model at each step.
@@
 We utilized three types of features in conjunction with CRF, MLP, BRNN and CRNN methods: word embeddings or lexical features, punctuation and POS features.
-Since POS tags have been shown to be effective semantic abstractions of individual words, we used POS features for our experiment \cite{liu2005using,treviso2017sentence}.
-To extract our POS features, we tag the e-Coaching emails using the NLTK POS tagger \cite{bird2009natural}.
-Punctuation marks, which correspond to one of the symbols \{`.', `,', `!', `?', `:', `;'\} between a pair of words, are also employed as a feature since punctuation marks designate the boundary of a sentence, clause and phrase and often also correspond to a segment boundary.
+Since POS tags have been shown to be effective semantic abstractions of individual words, we used POS features for our experiment.\cite{liu2005using,treviso2017sentence} To extract our POS features, we tag the e-Coaching emails using the NLTK POS tagger.
+\cite{bird2009natural} Punctuation marks, which correspond to one of the symbols \{`.', `,', `!', `?', `:', `;'\} between a pair of words, are also employed as a feature since punctuation marks designate the boundary of a sentence, clause and phrase and often also correspond to a segment boundary.
 For natural language processing (NLP) tasks, inputs are received as a text where words can be considered as the basic processing unit.
@@
 Word embeddings are one of this representation where each word represented as a real-valued vector in a high dimensional vector space.
-The distributed representation of word allows capturing semantic, syntactic and morphological information from large unannotated corpora  \cite{pennington2014glove, mikolov2013distributed}.
-Since pre-trained word2vec word vectors provide the best results, in our experiment, we utilized pre-trained word2vec word vectors for words.
-Trained word vectors were also utilized when rare words or punctuation marks are not available in the pre-trained word2vec word vectors.
+The distributed representation of word allows capturing semantic, syntactic and morphological information from large unannotated corpora.\cite{pennington2014glove, mikolov2013distributed} Since pre-trained word2vec word vectors provide the best results, in our experiment, we utilized pre-trained word2vec word vectors for words.
+Trained word vectors were also utilized when words or punctuation marks are not available in the pre-trained word2vec word vectors.
 Figure~\ref{fig:embedding-dimension} illustrates the performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and trained word embeddings with GloVe and word2vec models.
-It was observed that best performance achieved with pre-trained 300 dimensional word2Vec word vectors when three types of features are used together.
-Therefore, we report our results with word2vec 300 dimensional word vectors for all deep learning models used in this study.
+It was observed that best performance achieved with pre-trained 300-dimensional word2vec word vectors when three types of features are used together.
+Therefore, we report our results with word2vec 300-dimensional word vectors for all deep learning models used in this study.
 CRF utilized lexical features which represent each word in an e-Coaching email.
@@
 In the present study, we experimented with four different classifiers, including one traditional machine learning method, CRF and three deep learning models.
-Since deep learning provides a flexible architecture for constructing and synthesizing complex models, we take the advantage of this flexibility to test MLP, BRNN and CRNN models for the task of segmentation of e-Coaching emails.
+Since deep learning methods provide a flexible architecture for constructing and synthesizing complex models, we take the advantage of this flexibility to test MLP, BRNN and CRNN models for the task of segmentation of e-Coaching emails.
 
-\textbf{Conditional Random Fields}: CRF have been widely used in various NLP tasks including part-of-speech tagging and segmentation tasks \cite{lafferty2001conditional, hirohata2008identifying}.
-Unlike a maximum entropy markov model which uses per-state exponential models for the conditional probabilities of next states given the current state, a CRF is a discriminative model of the conditional distribution of the entire sequence of labels given the observation sequence.
+\textbf{Conditional Random Fields}: CRF has been widely used in various NLP tasks including part-of-speech tagging and segmentation tasks.\cite{lafferty2001conditional, hirohata2008identifying} Unlike a maximum entropy Markov model which uses per-state exponential models for the conditional probabilities of next states given the current state, a CRF is a discriminative model of the conditional distribution of the entire sequence of labels given the observation sequence.
 A traditional linear-chain CRF model is defined as a conditional probability distribution $p(y|x)$ for output and input sequences, $y$ and $x$, by the Eq.
@@
 The optimal output sequence $y^*$ for an input sequence $x$, $y^* = {arg\,max}_y p(y|x)$, is obtained efficiently by the Viterbi algorithm.
-In our experiments, the following features were utilized with CRF models: i) current word/punctuation ii) next and previous 3 words/punctuations iii) whether the word/punctuation is special characters (';', '?', '.', ',', '!', ':', etc) or not iv) whether the word is title word or not?
+In our experiments, the following features were utilized with CRF models: i) current word or punctuation ii) next and previous 3 words or punctuations iii) whether the word or punctuation is a special character (';', '?', '.', ',', '!', ':', etc) or not iv) whether the word is a title word or not?
 (e.g.
@@
 
-\textbf{Multilayer Perceptrons}: MLP is a feed-forward artificial neural networks which map set of input onto a set of appropriate output.
-This study utilized MLP with a nonlinear activation function (relu) and one hidden layer consisting of 128 hidden units.
-In order to prevent over-fitting, we implement dropout on fully connected layers, which randomly hide some neurons during the training phase \cite{srivastava2014dropout}.
-Dropout is also applied in the fully connected layers in BRNN and CRNN models.
-For a given sample with $2l$ words/punctuations, the first layer input of MLP was created by the summation of first $l$ word vectors and last $l$ word vectors.
-Figure~\ref{fig:length-mlp} demostrates the performance of MLP model on e-Coaching email segmentation by varying value of $l$.
+\textbf{Multilayer Perceptrons}: MLP is a feed-forward artificial neural network which maps set of input onto a set of appropriate output.\cite{rumelhart1986learning} This study utilized MLP with a nonlinear activation function (relu) and one hidden layer consisting of 128 hidden units.
+In order to prevent over-fitting, we implement dropout on fully connected layers, which randomly hide some neurons during the training phase.
+\cite{srivastava2014dropout} Dropout is also applied to the fully connected layers in BRNN and CRNN models.
+For a given sample with $2l$ words or punctuations, the first layer input of MLP was created by the summation of first $l$ word vectors and last $l$ word vectors.
+Figure~\ref{fig:length-mlp} demonstrates the performance of MLP model on e-Coaching email segmentation by varying value of $l$.
 Based on the results in Figure~\ref{fig:length-mlp}, MLP results were reported with $l = 2$, since best performance of MLP models were observed when $l = 2$.
@@
 
-\textbf{Bidirectional Recurrent Neural Networks}: BRNN is a neural network designed to capture sequential patterns by considering both past and future inputs as well as complex relationships between input feature and output labels.
-The output of BRNN layer is computed as the summation of the forward RNN output with backward RNN output.
-Gated Recurrent Units (GRU) \cite{chung2014empirical} are utilized as RNN that are capable of handling variable size input sequence and have an internal memory that can be reset.
+\textbf{Bidirectional Recurrent Neural Networks}: BRNN is a neural network designed to capture sequential patterns by considering both past \& future inputs as well as complex relationships between input features and output labels.\cite{schuster1997bidirectional} The output of BRNN layer is computed with the summation of the forward RNN output and backward RNN output.
+Gated Recurrent Units (GRU)\cite{chung2014empirical} are utilized as RNNs that are capable of handling variable size input sequence and have an internal memory that can be reset.
 Figure~\ref{fig:crnn} will represent an architecture of BRNN if convolution layer is removed.
 
-\textbf{Convolutional Recurrent Neural Networks}: CRNN is a deep neural networks architecture, shown in Figure~\ref{fig:crnn}, consisting of 5 layers: 1) input layer 2) embedding layer 3) convolution layer with max pooling 4) BRNN layer 5) fully connected layer with dropout and sigmoid output.
+\textbf{Convolutional Recurrent Neural Networks}: CRNN is a deep neural networks architecture,\cite{treviso2017sentence} shown in Figure~\ref{fig:crnn}, consisting of 5 layers: 1) input layer 2) embedding layer 3) convolution layer with max pooling 4) BRNN layer 5) fully connected layer with dropout and sigmoid output.
 
@@
 
-A sequence of $m$ words is fed into the input layer which provides a $m \times n_e$ matrix after fetching the pre-trained word vectors.
-When POS tags are used in combination with word embedding, 10-dimensional POS vectors are concatenated with 300-dimensional word vectors to obtain new vectors $n_e = n_w + n_p$ of size 310.
+E-coaching email exchanges, represented as a sequence of $m$ words and punctuations are fed into the input layer which provides an $m \times n_e$ matrix after fetching the pre-trained word vectors.
+When POS tags are used in combination with word embeddings, 10-dimensional POS vectors are concatenated with 300-dimensional word vectors to obtain new vectors $n_e = n_w + n_p$ of size 310.
 The 1D convolutional layer generates $m - h_c + 1$ new features with $h_{c.n_e}$, a filter of $h_c$ words and zero-padding on both sides of the text.
 A pooling operation was performed over time to find the most significant features in a region.
-After applying $n_f=100$ different filters on convolutional layer, $m \times n_f$ inputs are generated and fed into bidirectional recurrent neural networks, which contain 200 hidden units.
-Then, dropout is used to randomly drop 50\% of the neurons and passes them through a fully connected layer, where a sigmoid operation is calculated, giving us the probability of whether the word or punctuation precedes a new segment or not.
-We experimentally determined that the best performance is achieved with 5 folds corss-validation when filter length in convolution layer is 7, pool size is 3, batch size is 32, adam\cite{kingma2014adam} is used for optimization and the early stopping strategy is used with dropout to prevent over-fitting.
-To reuse of our source codes for other research studies, source codes of all experiments will be publicly available at https://github.com/teanalab/xxxxx upon acceptance of this paper for publication.
+After applying $n_f=100$ different filters on convolutional layer, $m \times n_f$ data are fed into bidirectional recurrent neural networks, which contain 200 hidden units.
+Then, dropout is used to randomly drop 50\% of the neurons and passes them through a fully connected layer, where a sigmoid operation is performed, provide the probability of whether the word or punctuation precedes a new segment or not.
+We experimentally determined that the best performance is achieved with 5 folds cross-validation when filter length in convolution layer is 7, pool size is 3, batch size is 32, adam\cite{kingma2014adam} is used for optimization and the early stopping strategy is used with dropout to prevent over-fitting.
+To utilize our source codes by other research studies, source codes of all experiments will be publicly available at https://github.com/teanalab/xxxxx upon acceptance of this paper for publication.
 
 \subsection*{\textit{Evaluation metrics}}
-We report standard metrics for experiments (precision, recall and F1-measure) to evaluate the performance of binary classifiers\cite{aas1999text}.
-However, accuracy is not reported as a performance metric because accuracy is highly sensitive to the prior class probabilities and does not fully describe the actual difficulty of the decision problem for an unbalanced dataset.
+We report standard metrics for experiments (precision, recall and F1-measure) to evaluate the performance of binary classifiers.\cite{aas1999text} However, accuracy is not reported as a performance metric because accuracy is highly sensitive to the prior class probabilities and does not fully describe the actual difficulty of the decision problem for an unbalanced dataset.
 The results are reported based on 5 folds cross-validation and weighted macro-averaging over the folds.
-We also estimate the area under the receiving operating characteristics (ROC) curve\cite{kumar2011receiver} (AUC) metric due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets \cite{hu2015kernelized}.
+We also estimate the area under the receiving operating characteristics (ROC) curve\cite{kumar2011receiver} (AUC) metric due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets.
+\cite{hu2015kernelized}
 
@@
   \hline
-   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{new segment}} & \multicolumn{3}{|c|}{\textbf{weighted macro average}} \\\cline{2-7}
+   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{New Segment}} & \multicolumn{3}{|c|}{\textbf{Weighted Macro Average}} \\\cline{2-7}
    & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} \\ \hline    
@@
 In addition, standard metrics: precision, recall and F1-measure were used to eliminate doubt about the model performance since accuracy is misleading for imbalanced datasets.
-AUC is also utilized due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets\cite{hu2015kernelized}, which are demonstrated by the Table~\ref{tab:result_roc} and ROC curves in Figure~\ref{fig:roc-curves}.
+AUC is also utilized due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets,\cite{hu2015kernelized} which are demonstrated by the Table~\ref{tab:result_roc} and ROC curves in Figure~\ref{fig:roc-curves}.
 Although CRF provides the second highest classification results, it shows best AUC values, achieving 0.966 AUC when only lexical features are used and 0.994 AUC when a combination of lexical, punctuation and POS features are used.

### 2018-07-31 paper-amia-summit-2018/manuscript.tex
finalize 1st draft for Alex

@@
     \includegraphics[width=0.7\textwidth]{figures/segment-example.png}
-    \caption{\textbf{Example of e-Coaching emails segmented into fragments that correspond to MI behaviors of an e-Coach and a patient}.}
+    \caption{\textbf{Example of e-Coaching emails segmented into fragments that correspond to MI behaviors of an e-Coach and a patient}}
     \label{fig:text-segment}
@@
 
-Relevant previous work in the biomedical domain primarily focused on segmentation of text into sections and headers\cite{apostolova2009automatic,denny2009evaluation,tepper2012statistical,cho2002text} or sentence boundary detection\cite{griffis2016quantitative,kreuzthaler2015detection,treviso2017sentence,fraser2015sentence}.
-Apostolova et al.
+Relevant previous work in the biomedical domain primarily focused on segmentation of text into sections and headers\cite{apostolova2009automatic,denny2009evaluation,tepper2012statistical,cho2002text} or sentence boundary detection.\cite{griffis2016quantitative,kreuzthaler2015detection,treviso2017sentence} Apostolova et al.
 applied SVM along with word-vector cosine similarity metric combined with several heuristics to segment clinical reports into sections, such as demographics, history, procedure, finding and impression.\cite{apostolova2009automatic} After identification of each line in the document, Tepper et al.
@@
 have shown that a linear chain CRF performed better than the hidden markov model and Maxent on the NIST sentence boundary detection.\cite{liu2005using} Recently, Treviso et al.
-proposed a recurrent convolutional neural network for automatic identification of sentence boundary in a neuropsychological text in Portuguese language.\cite{treviso2017sentence} Their proposed model combined prosodic and part of speech features with pre-trained GloVe word vectors,\cite{pennington2014glove} which achieved 0.74 F1-score in sentence boundary detection on the Brazilian constitution dataset.
+proposed a recurrent convolutional neural network for automatic identification of sentence boundary in a neuropsychological text in the Portuguese language.\cite{treviso2017sentence} Their proposed model combined prosodic and part of speech features with pre-trained GloVe word vectors,\cite{pennington2014glove} which achieved 0.74 F1-score in sentence boundary detection on the Brazilian constitution dataset.
 
 In this study, we adopt the model proposed by Treviso et al.
-\cite{treviso2017sentence} and rename it as convolutional recurrent neural networks to apply the model differently from the aforementioned studies.
+\cite{treviso2017sentence} and rename it as convolutional recurrent neural networks (CRNN) to apply it differently from the aforementioned studies.
 The main difference is that our study focuses on the segmentation of e-Coaching emails as distinct from determining the sentence boundary.
@@
 Since POS tags have been shown to be effective semantic abstractions of individual words, we used POS features for our experiment.\cite{liu2005using,treviso2017sentence} To extract our POS features, we tag the e-Coaching emails using the NLTK POS tagger.
-\cite{bird2009natural} Punctuation marks, which correspond to one of the symbols \{`.', `,', `!', `?', `:', `;'\} between a pair of words, are also employed as a feature since punctuation marks designate the boundary of a sentence, clause and phrase and often also correspond to a segment boundary.
+Punctuation marks, which correspond to one of the symbols \{`.', `,', `!', `?', `:', `;'\} between a pair of words, are also employed as a feature since punctuation marks designate the boundary of a sentence, clause and phrase and often also correspond to a segment boundary.
 For natural language processing (NLP) tasks, inputs are received as a text where words can be considered as the basic processing unit.
@@
     \includegraphics[width=0.5\textwidth]{figures/length-mlp.eps}
-    \caption{\textbf{Performance of MLP model on e-Coaching email segmentation by varying the length of sample size.}}
+    \caption{\textbf{Performance of MLP model on e-Coaching email segmentation by varying the length of sample size}}
     \label{fig:length-mlp}
@@
     \includegraphics[width=1.0\textwidth]{figures/CRNN.eps}
-    \caption{\textbf{Architecture of convolutional recurrent neural networks for automatic segmentation of e-Coaching emails}.}
+    \caption{\textbf{Architecture of convolutional recurrent neural networks for automatic segmentation of e-Coaching emails}}
     \label{fig:crnn}
@@
 
-E-coaching email exchanges, represented as a sequence of $m$ words and punctuations are fed into the input layer which provides an $m \times n_e$ matrix after fetching the pre-trained word vectors.
+E-coaching email exchanges are represented as a sequence of $m$ words and punctuations, which are fed into the input layer to produce a $m \times n_e$ matrix after fetching the pre-trained word vectors.
 When POS tags are used in combination with word embeddings, 10-dimensional POS vectors are concatenated with 300-dimensional word vectors to obtain new vectors $n_e = n_w + n_p$ of size 310.
@@
 After applying $n_f=100$ different filters on convolutional layer, $m \times n_f$ data are fed into bidirectional recurrent neural networks, which contain 200 hidden units.
-Then, dropout is used to randomly drop 50\% of the neurons and passes them through a fully connected layer, where a sigmoid operation is performed, provide the probability of whether the word or punctuation precedes a new segment or not.
+Then, dropout is used to randomly drop 50\% of the connections and passes them through a fully connected layer, where a sigmoid operation is performed, provide the probability of whether the word or punctuation precedes a new segment or not.
 We experimentally determined that the best performance is achieved with 5 folds cross-validation when filter length in convolution layer is 7, pool size is 3, batch size is 32, adam\cite{kingma2014adam} is used for optimization and the early stopping strategy is used with dropout to prevent over-fitting.
@@
   \hline
-   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{New Segment}} & \multicolumn{3}{|c|}{\textbf{Weighted Macro Average}} \\\cline{2-7}
+   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{New Segment}} & \multicolumn{3}{|c|}{\textbf{Overall}} \\\cline{2-7}
    & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} \\ \hline    
@@
 The robust performance of CRNN provides the evidence that deep learning models are capable to learn MI behaviors from the clinical exchange.
-It also indicates that punctuation and POS features are important with word embeddings or lexical features, even when deep learning methods are employed.
+It also indicates that punctuation and POS features are important with word embeddings or lexical features for all machine learning methods are employed.
 Although the domain of this study was intentionally quite small, we believe that our study is not limited to the e-Coaching domain and it can be successfully applied to other domains, in which discourse segmentation is a preliminary step for annotation.
@@
 This results also indicate that segmentation performance might be improved by adding additional relevant features.
+
+The convolutional layer made a significant difference between the performance of CRNN and BRNN on email segmentation.
+CRNN demonstrates 22.5\% and 10.7\% higher F1-Measure in ``new segment'' detection and 1\% and 0.5\% higher F1-Measure in overall compared to BRNN when word embeddings and all features are used, respectively.
 \\
@@
     \includegraphics[width=1.0\textwidth]{figures/roc-curves.eps}
-    \caption{\textbf{Receiver operating characteristic curves showing the performance of binary classifiers for the segmentation of e-Coaching text when only word embeddings/lexical features (left) and combination of word embeddings/lexical, punctuation and POS features (right) are used}.}
+    \caption{\textbf{Receiver operating characteristic curves showing the performance of binary classifiers for the segmentation of e-Coaching text when only word embeddings or lexical features (left) and combination of word embeddings or lexical, punctuation and POS features (right) are used}}
     \label{fig:roc-curves}
@@
 \centering
-\caption{AUC values of all classifiers when only word embeddings/lexical features and combination of word embeddings/lexical, punctuation and POS features are used.
+\caption{AUC values of all classifiers when only word embeddings or lexical features and combination of word embeddings or lexical, punctuation and POS features are used.
 Highest value for each experiment across all models is highlighted in boldface.}
@@
   \hline
-\multirow{2}{*}{\textbf{Features}} & \multicolumn{4}{|c|}{\textbf{Area Under the ROC Curve (AUC)}} \\\cline{2-5}
+\multirow{2}{*}{\textbf{Features}} & \multicolumn{4}{|c|}{\textbf{AUC}} \\\cline{2-5}
  & \textbf{CRF} & \textbf{MLP}  & \textbf{BRNN} & \textbf{CRNN} \\ \hline      
@@
 Our proposed methods can help to identify individual text segments, which can be annotated directly with a classification model.
-This approach will also help for developing fully automated e-Coaching and accelerate the pace of identifying effective communication strategies.
+This approach will also help in developing fully automated e-Coaching and accelerate the pace of identifying effective communication strategies.
 
@@
 This study was supported by a grant from the National Institutes of Health, NIDDK R21DK108071, Carcone and Kotov, MPIs.
-We would like to thank research staff and student assistants in the Department of Family Medicine and Public Health Sciences at Wayne State University School of Medicine for their help in preparing the training dataset.
+We would like to thank the research staff and student assistants in the Department of Family Medicine and Public Health Sciences at Wayne State University School of Medicine for their help in preparing the training dataset.
 

### 2018-07-31 paper-amia-summit-2018/manuscript.tex
finalize 1st draft for Alex

@@
     \includegraphics[width=0.5\textwidth]{figures/embedding-dimension.eps}
-    \caption{\textbf{Performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and trained word embeddings with GloVe and word2vec models.}}
+    \caption{\textbf{Performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and trained word embeddings with GloVe and word2vec models}}
     \label{fig:embedding-dimension}

### 2018-08-04 paper-amia-summit-2018/manuscript.tex
update manuscript

@@
 \usepackage{multirow}
-\renewcommand*{\thefootnote}{\fnsymbol{footnote}}
+\renewcommand*{\thefootnote}{\alph{footnote}}
 
@@
 
-\author{Mehedi Hasan, BS$^{1}$\footnote[1]{Authors provided an equal contribution.
+\author{Mehedi Hasan, BS$^{1}$\footnote{Authors provided an equal contribution.
 \label{footnote1}}, Alexander Kotov, PhD$^{1}$\textsuperscript{\ref{footnote1}}, Sylvie Naar, PhD$^{2}$, Gwen L.
@@
 \begin{enumerate}
-\item Emails are unstructured text that contains informal information exchange in a non-traditional format.
+\item Emails are unstructured text that contains informal information exchange in a non-traditional format. However, a traditional MI transcript corresponds to an MI interview
+session, contains a dialogue between a counselor and a patient.
 \item Discourse segments in e-Coaching emails do not necessarily correspond to sentences or collection of sentences.
@@
 
+Recently, an online clinical intervention called MENU GenY (Making Effective Nutrition Choices for Generation Y) was proposed and evaluated.
+\cite{alexander2017motivations} MENU GenY is a technology-based public health intervention that relies on personalized e-coaching to encourage increased fruit and vegetable intake among young adults, aged 21-30.
+The goal of MENU GenY was to develop a better coding dictionary among GenY to improve eating habits.
+However, segmentation of clinical exchange in the context of electronically delivered interventions, in particular, segmentation of clinical interaction text into groups of MI behaviors, is still performed manually, which slows down qualitative analysis of these interventions.
+This study proposes several machine learning methods to address this problem and the authors are unaware of any other work that focused on the same problem.
+
 The goal of this research study is to assess the applicability of machine learning methods for automated segmentation of e-Coaching emails into textual fragments corresponding to individual behaviors, which is the first step of the coding process of e-Coaching communications.
@@
 
-Relevant previous work in the biomedical domain primarily focused on segmentation of text into sections and headers\cite{apostolova2009automatic,denny2009evaluation,tepper2012statistical,cho2002text} or sentence boundary detection.\cite{griffis2016quantitative,kreuzthaler2015detection,treviso2017sentence} Apostolova et al.
-applied SVM along with word-vector cosine similarity metric combined with several heuristics to segment clinical reports into sections, such as demographics, history, procedure, finding and impression.\cite{apostolova2009automatic} After identification of each line in the document, Tepper et al.
-trained Maximum Entropy models for section classification.\cite{tepper2012statistical} Denny et al.
+Relevant previous work in the biomedical domain primarily focused on segmentation of text into sections and headers of electronic health records (EHR)\cite{apostolova2009automatic,denny2009evaluation,tepper2012statistical,cho2002text} or sentence boundary detection of EHR.\cite{griffis2016quantitative,kreuzthaler2015detection,treviso2017sentence} Apostolova et al.
+applied SVM along with word-vector cosine similarity metric combined with several heuristics to segment clinical reports in EHR into sections, such as demographics, history, procedure, finding and impression.\cite{apostolova2009automatic} After identification of each line in the document, Tepper et al.
+trained Maximum Entropy models to classify the document into pre-defined sections, such as general patient information, provider information, medical condition, medical history, clinical history, findings, impression, etc.\cite{tepper2012statistical} Denny et al.
 proposed a SecTag algorithm, which combined natural language processing techniques, terminology-based rules and a Naive Bayes classifier to identify the sections and headers that achieved 99\% recall with 95.6\% precision.
-\cite{denny2009evaluation} On the other hand, SVM based on prosodic and part of speech features \cite{kreuzthaler2015detection} and recurrent convolutional neural networks using word embeddings \cite{griffis2016quantitative} were utilized for detecting sentence boundaries.
+\cite{denny2009evaluation} On the other hand, SVM based on prosodic and part of speech features \cite{kreuzthaler2015detection} and recurrent convolutional neural networks using word embeddings \cite{griffis2016quantitative} were utilized for detecting sentence boundaries in EHR.
 Liu et al.
-have shown that a linear chain CRF performed better than the hidden markov model and Maxent on the NIST sentence boundary detection.\cite{liu2005using} Recently, Treviso et al.
-proposed a recurrent convolutional neural network for automatic identification of sentence boundary in a neuropsychological text in the Portuguese language.\cite{treviso2017sentence} Their proposed model combined prosodic and part of speech features with pre-trained GloVe word vectors,\cite{pennington2014glove} which achieved 0.74 F1-score in sentence boundary detection on the Brazilian constitution dataset.
+have shown that a linear chain CRF performed better than the hidden markov model and Maxent in sentence boundary detection, evaluated using the official
+NIST evaluation tools.\cite{liu2005using} Recently, Treviso et al.
+proposed a recurrent convolutional neural network for automatic identification of sentence boundary in a neuropsychological text in the Portuguese language.\cite{treviso2017sentence} Their proposed model combined prosodic and part of speech features with pre-trained GloVe word embeddings,\cite{pennington2014glove} which achieved 0.74 F1-score in sentence boundary detection on the Brazilian constitution dataset.
 
@@
 
-Recently, an online clinical intervention called MENU GenY (Making Effective Nutrition Choices for Generation Y) was proposed and evaluated.
-\cite{alexander2017motivations} MENU GenY is a technology-based public health intervention that relies on personalized e-coaching to encourage increased fruit and vegetable intake among young adults, aged 21-30.
-The goal of MENU GenY was to develop a better coding dictionary among GenY to improve eating habits.
-However, segmentation of clinical exchange in the context of electronically delivered interventions, in particular, segmentation of clinical interaction text into groups of MI behaviors, is still performed manually, which slows down qualitative analysis of these interventions.
-This study proposes several machine learning methods to address this problem and the authors are unaware of any other work that focused on the same problem.
+Our proposed approach is novel for the segmentation of e-Coaching emails since previous studies mainly focused on the segmentation of EHR documents into sections, headers and sentences.
+However, this study segmented clinical exchange into groups of MI behaviors which will significantly reduce the amount of resource and time required to segment clinical exchange manually.
+Furthermore, these segmentation models could be integrated with auto coding classifiers to build a software pipeline for automated annotation of clinical exchanges.
 
@@
 In total, we obtained 95,777 words and 7,140 punctuation marks, which include 3,138 ``new segment'' and 99,779 ``same segment'' instances.
-In this study, we experimented with traditional machine learning method, CRF and deep learning methods, such as MLP, BRNN and CRNN.
-For MLP models, samples were created with $2l$ words or punctuations at each step in a word sequence, where a sample contains next $l$ words or punctuations and prior $l$ words or punctuations including current wordor punctuation.
-Each sample is classified into either ``new segment'' or ``same segment'' based on whether the current word or punctuation precedes a new segment or not.
+In this study, we experimented with traditional machine learning method, conditional random fields (CRF)\cite{lafferty2001conditional} and deep learning methods, such as multilayer perceptrons (MLP),\cite{rumelhart1986learning} bidirectional recurrent neural networks (BRNN)\cite{schuster1997bidirectional} and convolutional recurrent neural networks (CRNN).\cite{treviso2017sentence} For MLP models, samples were created with $2l$ words or punctuations at each step in a word sequence, where a sample contains next $l$ words or punctuations and prior $l$ words or punctuations including current word or punctuation.
+Each sample is classified into either ``new segment'' or ``same segment'' based on whether the $l^{th}$ word or punctuation in the gold standard precedes a new segment or not.
 For CRF, BRNN and CRNN models, an email was taken as input sequence, such that POS tags and word embedding of each word or punctuation mark were used as input and binary labels (1 or 0) corresponding to ``new segment'' and ``same segment'' classification decision were considered as the output of the model at each step.
@@
 Since POS tags have been shown to be effective semantic abstractions of individual words, we used POS features for our experiment.\cite{liu2005using,treviso2017sentence} To extract our POS features, we tag the e-Coaching emails using the NLTK POS tagger.
-Punctuation marks, which correspond to one of the symbols \{`.', `,', `!', `?', `:', `;'\} between a pair of words, are also employed as a feature since punctuation marks designate the boundary of a sentence, clause and phrase and often also correspond to a segment boundary.
-For natural language processing (NLP) tasks, inputs are received as a text where words can be considered as the basic processing unit.
+Punctuation marks, which correspond to one of the symbols \{`.', `,', `!', `?', `:', `;'\} between a pair of words, are also employed as a feature since punctuation marks designate the boundary of a sentence, clause and phrase and often also correspond to a segment boundary.\cite{cho2002text} For natural language processing (NLP) tasks, inputs are received as a text where words can be considered as the basic processing unit.
 Therefore, it is important to represent a word in such a way that it carries all relevant information.
 Word embeddings are one of this representation where each word represented as a real-valued vector in a high dimensional vector space.
-The distributed representation of word allows capturing semantic, syntactic and morphological information from large unannotated corpora.\cite{pennington2014glove, mikolov2013distributed} Since pre-trained word2vec word vectors provide the best results, in our experiment, we utilized pre-trained word2vec word vectors for words.
-Trained word vectors were also utilized when words or punctuation marks are not available in the pre-trained word2vec word vectors.
+The distributed representation of word allows capturing semantic, syntactic and morphological information from large unannotated corpora.\cite{pennington2014glove, mikolov2013distributed} In our experiment, we utilized embeddings estimated on Google News corpus consisting of 1.6 billion words referred to pre-trained word2vec word embeddings.\footnote{https://code.google.com/p/word2vec/} When words or punctuation marks are not available in the pre-trained word vectors, we utilized word embeddings trained with our e-Coaching email corpus referred to trained word embeddings.
 Figure~\ref{fig:embedding-dimension} illustrates the performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and trained word embeddings with GloVe and word2vec models.
@@
 
-\textbf{Conditional Random Fields}: CRF has been widely used in various NLP tasks including part-of-speech tagging and segmentation tasks.\cite{lafferty2001conditional, hirohata2008identifying} Unlike a maximum entropy Markov model which uses per-state exponential models for the conditional probabilities of next states given the current state, a CRF is a discriminative model of the conditional distribution of the entire sequence of labels given the observation sequence.
+\textbf{Conditional Random Fields}: CRF has been widely used in various NLP tasks including part-of-speech tagging and segmentation tasks.\cite{lafferty2001conditional, hirohata2008identifying} Unlike a maximum entropy Markov model which uses per-state exponential models for the conditional probabilities of next states given the current state, a CRF is modeled with an exponential distribution provides the conditional distribution of the entire output sequence given the observation sequence.
 A traditional linear-chain CRF model is defined as a conditional probability distribution $p(y|x)$ for output and input sequences, $y$ and $x$, by the Eq.
@@
 The optimal output sequence $y^*$ for an input sequence $x$, $y^* = {arg\,max}_y p(y|x)$, is obtained efficiently by the Viterbi algorithm.
-In our experiments, the following features were utilized with CRF models: i) current word or punctuation ii) next and previous 3 words or punctuations iii) whether the word or punctuation is a special character (';', '?', '.', ',', '!', ':', etc) or not iv) whether the word is a title word or not?
+In our experiments, the following features were utilized with CRF models: i) current word or punctuation ii) next and previous 3 words or punctuations iii) binary feature indicating the word or punctuation is a special character (';', '?', '.', ',', '!', ':', etc) or not iv) binary feature indicating the word is a title word or not?
 (e.g.
@@
 
-\textbf{Multilayer Perceptrons}: MLP is a feed-forward artificial neural network which maps set of input onto a set of appropriate output.\cite{rumelhart1986learning} This study utilized MLP with a nonlinear activation function (relu) and one hidden layer consisting of 128 hidden units.
+\textbf{Multilayer Perceptrons}: MLP is a feed-forward artificial neural network which maps set of input onto a set of appropriate output.\cite{rumelhart1986learning} Figure~\ref{fig:mlp} shows a multi layer perceptron with a single hidden layer.
+In an MLP networks, there are no cycles or loops where the information moves forward only, from the input nodes through the hidden nodes and to the output nodes.
+This study utilized MLP with a nonlinear activation function (relu) and one hidden layer consisting of 128 hidden units.
 In order to prevent over-fitting, we implement dropout on fully connected layers, which randomly hide some neurons during the training phase.
 \cite{srivastava2014dropout} Dropout is also applied to the fully connected layers in BRNN and CRNN models.
-For a given sample with $2l$ words or punctuations, the first layer input of MLP was created by the summation of first $l$ word vectors and last $l$ word vectors.
-Figure~\ref{fig:length-mlp} demonstrates the performance of MLP model on e-Coaching email segmentation by varying value of $l$.
-Based on the results in Figure~\ref{fig:length-mlp}, MLP results were reported with $l = 2$, since best performance of MLP models were observed when $l = 2$.
 
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{figures/length-mlp.eps}
-    \caption{\textbf{Performance of MLP model on e-Coaching email segmentation by varying the length of sample size}}
-    \label{fig:length-mlp}
+    \includegraphics[width=0.5\textwidth]{figures/mlp.jpeg}
+    \caption{\textbf{A multi layer perceptron with a single hidden layer}}
+    \label{fig:mlp}
 \end{figure}
@@
 When POS tags are used in combination with word embeddings, 10-dimensional POS vectors are concatenated with 300-dimensional word vectors to obtain new vectors $n_e = n_w + n_p$ of size 310.
-The 1D convolutional layer generates $m - h_c + 1$ new features with $h_{c.n_e}$, a filter of $h_c$ words and zero-padding on both sides of the text.
-A pooling operation was performed over time to find the most significant features in a region.
-After applying $n_f=100$ different filters on convolutional layer, $m \times n_f$ data are fed into bidirectional recurrent neural networks, which contain 200 hidden units.
-Then, dropout is used to randomly drop 50\% of the connections and passes them through a fully connected layer, where a sigmoid operation is performed, provide the probability of whether the word or punctuation precedes a new segment or not.
-We experimentally determined that the best performance is achieved with 5 folds cross-validation when filter length in convolution layer is 7, pool size is 3, batch size is 32, adam\cite{kingma2014adam} is used for optimization and the early stopping strategy is used with dropout to prevent over-fitting.
-To utilize our source codes by other research studies, source codes of all experiments will be publicly available at https://github.com/teanalab/xxxxx upon acceptance of this paper for publication.
+The primary purpose of convolution is to extract new features depending on neighbor words.
+We used 1D convolution which simply moves one step at a time on vertically.
+In this layer, one filter is responsible for extraction of one feature.
+After applying $n_f$ different filters with zero-padding on both sides of the text, $n_f$ useful features are produced in the convolution layer for each word.
+A max pooling operation was then performed over time to find the most significant features in a text region.
+The bidirectional recurrent layer receives new features extracted in the convolution layer.
+RNN is usually used to capture the previous history.
+Moreover, bidirectional RNN is capable of capturing both previous and past histories because bidirectional RNN utilized the features by looking at forward and backward states.
+The purpose of the fully connected layer is to use the bidirectional RNN layer output for classifying each word or punctuation into ``new segment'' and ``same segment'' classes.
+Since a fully connected layer has a greater number of parameters, they are more likely excessively co-adapt these parameters causing over-fitting.
+To prevent this over-fitting problem, dropout is used to randomly drop 50\% of the connections on the fully connected layer.
+After that, a sigmoid operation is performed on the output layer to provide the probability of whether the word or punctuation precedes a new segment or not.
+We experimentally determined the optimal parameters and found that the best performance is achieved with 5 folds cross-validation when filter length in convolution layer is 7, number of filters is 100, max-pool size is 3, convolutional layer activation is $relu$, RNN layer activation is $tanh$ and number of recurrent units in RNN is 200.
+Adam\cite{kingma2014adam} is used for optimization and the early stopping strategy is applied with 50 epochs when the batch size is 32 and the learning rate is 0.001.
+To utilize our source codes by other research studies, source codes of all experiments will be publicly available\footnote{https://github.com/teanalab/xxxxx} upon acceptance of this paper for publication.
 
@@
 First, results are reported with respect to ``new segment'' class as well as weighted average over ``new segment'' and ``same segment'' classes in Table~\ref{tab:result_base}.
-Second, classification performance of different machine learning methods are outlined in Table~\ref{tab:result_weighted_avg} when word embeddings or lexical features are used in combination with punctuation and POS features.\\
+Second, classification performance of different machine learning methods are outlined in Table~\ref{tab:result_weighted_avg} when word embeddings or lexical features are used in combination with punctuation and POS features.
+
+For a given sample with $2l$ words or punctuations, the first layer input of MLP was created by the summation of first $l$ word vectors and last $l$ word vectors.
+Figure~\ref{fig:length-mlp} demonstrates the performance of MLP model on e-Coaching email segmentation by varying value of $l$.
+Based on the results in Figure~\ref{fig:length-mlp}, MLP results were reported with $l = 2$, since best performance of MLP models were observed when $l = 2$.
+
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=0.5\textwidth]{figures/length-mlp.eps}
+    \caption{\textbf{Performance of MLP model on e-Coaching email segmentation by varying the length of sample size}}
+    \label{fig:length-mlp}
+\end{figure}
+
+As follows from Table~\ref{tab:result_base}, CRNN outperforms all other models in terms of recall and F1-measure achieving 0.797 recall with 0.785 F1-measure for new segment detection.
+CRNN also shows superior performance in all performance metrics for weighted average over ``new segment'' and ``same segment'' classes.
+BRNN demonstrates the lowest performance among all models in terms of precision and F1-Measure.
+On the other hand, MLP has the highest precision of 0.836 when word embeddings or lexical features are used to identify ``new segment''.
+CRF achieves 0.733 F1-Measure in new segment class and 0.984 F1-Measure in overall, which demonstrates the second highest performance to identify ``new segment'' as well as a weighted average over both classes.
+Experimental results indicate that performance of all classifiers as a weighted average over both classes is remarkably higher compared to ``new segment'' class, which is expected since 96.95\% instances belong to the ``same segment'' categories and 99.3\% of them correctly classified.
+For example, as a weighted average over ``new segment'' and ``same segment'' classes, CRNN achieves 27.23\%, 23.71\% and 25.61\% higher precision, recall and F1-measure, respectively, compared to ``new segment'' detection.
+\\
 
@@
 \centering
-\caption{Performance of CRF, MLP, BRNN and CRNN methods for identification of ``new segment'' class as well as weighted average over ``new segment'' and ``same segment'' classes when word embeddings or lexical features are used.
-The highest value for each performance metric
-is highlighted in bold.}
+\caption{\textbf{Performance of CRF, MLP, BRNN and CRNN methods for identification of ``new segment'' class as well as weighted average over ``new segment'' and ``same segment'' classes when word embeddings or lexical features are used.
+The highest value for each performance metric is highlighted in bold.}}
 \label{tab:result_base}
@@
 
-As follows from Table~\ref{tab:result_base}, CRNN outperforms all other models in terms of recall and F1-measure achieving 0.797 recall with 0.785 F1-measure for new segment detection.
-CRNN also shows superior performance in all performance metrics for weighted average over ``new segment'' and ``same segment'' classes.
-BRNN demonstrates the lowest performance among all models in terms of precision and F1-Measure.
-On the other hand, MLP has the highest precision of 0.836 when word embeddings or lexical features are used to identify ``new segment''.
-CRF achieves 0.733 F1-Measure in new segment class and 0.984 F1-Measure in overall, which demonstrates the second highest performance to identify ``new segment'' as well as a weighted average over both classes.
-Experimental results indicate that performance of all classifiers as a weighted average over both classes is remarkably higher compared to ``new segment'' class, which is expected since 96.95\% instances belong to the ``same segment'' categories and 99.3\% of them correctly classified.
-For example, as a weighted average over ``new segment'' and ``same segment'' classes, CRNN achieves 27.23\%, 23.71\% and 25.61\% higher precision, recall and F1-measure, respectively, compared to ``new segment'' detection.
+Table~\ref{tab:result_roc} shows the impact of individual fetures as well as their combination on the segmentation task.
+Although CRF provides the second highest classification results, it shows best AUC values, achieving 0.966 AUC when only lexical features are used and 0.994 AUC when a combination of lexical, punctuation and POS features are used.
+Influence of the punctuation and POS features is also consistent in AUC values, which increase by 2.9\%, 4.31\%, 6.63\% and 2.18\% for CRF, MLP, BRNN and CRNN methods, respectively.\\
+
+\begin{table}[ht]
+\centering
+\caption{\textbf{AUC values of all classifiers when only word embeddings or lexical features and combination of word embeddings or lexical, punctuation and POS features are used.
+Highest value for each experiment across all models is highlighted in boldface.}}
+\label{tab:result_roc}
+ \begin{tabular}{|l|l|l|l|l|}
+  \hline
+\multirow{2}{*}{\textbf{Features}} & \multicolumn{4}{|c|}{\textbf{AUC}} \\\cline{2-5}
+ & \textbf{CRF} & \textbf{MLP}  & \textbf{BRNN} & \textbf{CRNN} \\ \hline      
+ word embeddings only & \textbf{0.966} & 0.951 & 0.920 & 0.965 \\ \hline
+ word embeddings + POS & \textbf{0.966} & 0.951 & 0.920 & 0.965 \\ \hline
+ word embeddings + punctuation & \textbf{0.966} & 0.951 & 0.920 & 0.965 \\ \hline
+ all features & \textbf{0.994} & 0.992 & 0.981 & 0.986 \\ \hline
+  \end{tabular}
+\end{table}
+
+% Baseline ROC results: BRNN: 0.9198, MLP: 0.9507, CRF: 0.9662, CRNN: 0.9648
+% Proposed ROC Results: BRNN: 0.9807, MLP: 0.992, CRF: 0.9937, CRNN: 0.9859
 
@@
 We observed that classification performance significantly improved for ``new segment'' class when word embeddings or lexical features are used in combination with punctuation and POS features.
-Precision increases by 3.96\%, -2.63\%, 12.71\% and 1.81\%; recall increases by 11.72\%, 23.44\%, 20.59\% and 8.41\%; and F1-measure increases by 8.05\%, 11.1\%, 16.22\% and 5.1\% for CRF, MLP, BRNN and CRNN methods, respectively, in new segment detection.
-Similarly, precision increases by 0.51\%, 0.51\%, 0.82\% and 0.41\%; recall increases by 0.41\%, 0.41\%, 0.72\% and 0.3\%; and F1-measure increases by 0.41\%, 0.51\%, 0.82\% and 0.3\% for CRF, MLP, BRNN and CRNN methods, respectively, in weighted average over ``new segment'' and ``same segment'' classes.\\
+Precision increases by 3.96\%, -2.63\%, 12.71\% and 1.81\%; recall increases by 11.72\%, 23.44\%, 20.59\% and 8.41\%; and F1-measure increases by 8.05\%, 11.1\%, 16.22\% and 5.1\% for CRF, MLP, BRNN and CRNN methods, respectively, in new segment detection when all features are utilized together.
+Similarly, precision increases by 0.51\%, 0.51\%, 0.82\% and 0.41\%; recall increases by 0.41\%, 0.41\%, 0.72\% and 0.3\%; and F1-measure increases by 0.41\%, 0.51\%, 0.82\% and 0.3\% for CRF, MLP, BRNN and CRNN methods, respectively, in weighted average over ``new segment'' and ``same segment'' classes when word embeddings or lexical features are used in combination with punctuation and POS features.\\
 
@@
 \centering
-\caption{Performance of CRF, MLP, BRNN and CRNN methods for identification of ``new segment'' class as well as weighted average over ``new segment'' and ``same segment'' classes when all features are used together.
+\caption{\textbf{Performance of CRF, MLP, BRNN and CRNN methods for identification of ``new segment'' class as well as weighted average over ``new segment'' and ``same segment'' classes when all features are used together.
 The highest value for each performance metric
-is highlighted in bold.}
+is highlighted in bold.}}
 \label{tab:result_weighted_avg}
@@
 CRNN achieved 0.989 F1-measure in overall and 0.825 F1-measure for detecting ``new segment''.
-The robust performance of CRNN provides the evidence that deep learning models are capable to learn MI behaviors from the clinical exchange.
+The robust performance of CRNN provides the evidence that deep learning models are capable to learn the hidden characteristics of MI behaviors from the clinical exchange.
 It also indicates that punctuation and POS features are important with word embeddings or lexical features for all machine learning methods are employed.
@@
 Punctuation mark and POS features made a significant improvement in the performance of machine learning and all deep learning methods.
+Specially, punctuation features have the higher individual impact on model performance compared to POS features.
 In all cases, every method performs better, when word embeddings or lexical features are used in combination with punctuation and POS features.
@@
 CRNN demonstrates 22.5\% and 10.7\% higher F1-Measure in ``new segment'' detection and 1\% and 0.5\% higher F1-Measure in overall compared to BRNN when word embeddings and all features are used, respectively.
-\\
+In CRNN, convolution layer performed a series of convolutions and pooling operations during which a number of high-level important features are extracted from the word embeddings of e-Coaching emails.
+These high-level features are then utilized by the bidirectional RNN of CRNN model, which provide an improved performance on email segmentation.
+On the other hand, traditional BRNN model received words as input features and their word embeddings are directly utilized in the input layer.
 
@@
 
-\begin{table}[ht]
-\centering
-\caption{AUC values of all classifiers when only word embeddings or lexical features and combination of word embeddings or lexical, punctuation and POS features are used.
-Highest value for each experiment across all models is highlighted in boldface.}
-\label{tab:result_roc}
- \begin{tabular}{|l|l|l|l|l|}
-  \hline
-\multirow{2}{*}{\textbf{Features}} & \multicolumn{4}{|c|}{\textbf{AUC}} \\\cline{2-5}
- & \textbf{CRF} & \textbf{MLP}  & \textbf{BRNN} & \textbf{CRNN} \\ \hline      
- word embeddings or lexical features & \textbf{0.966} & 0.951 & 0.920 & 0.965 \\ \hline
- all features & \textbf{0.994} & 0.992 & 0.981 & 0.986 \\ \hline
-  \end{tabular}
-\end{table}
-
-% Baseline ROC results: BRNN: 0.9198, MLP: 0.9507, CRF: 0.9662, CRNN: 0.9648
-% Proposed ROC Results: BRNN: 0.9807, MLP: 0.992, CRF: 0.9937, CRNN: 0.9859
-
 In this paper, results are reported by ``new segment'' class as well as weighted average over both classes to avoid confusion regarding overall model performance due to severe class imbalance.
 In addition, standard metrics: precision, recall and F1-measure were used to eliminate doubt about the model performance since accuracy is misleading for imbalanced datasets.
-AUC is also utilized due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets,\cite{hu2015kernelized} which are demonstrated by the Table~\ref{tab:result_roc} and ROC curves in Figure~\ref{fig:roc-curves}.
-Although CRF provides the second highest classification results, it shows best AUC values, achieving 0.966 AUC when only lexical features are used and 0.994 AUC when a combination of lexical, punctuation and POS features are used.
-Influence of the punctuation and POS features is also consistent in AUC values, which increase by 2.9\%, 4.31\%, 6.63\% and 2.18\% for CRF, MLP, BRNN and CRNN methods, respectively.
+AUC is also utilized due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets,\cite{hu2015kernelized} which are demonstrated by the ROC curves in Figure~\ref{fig:roc-curves}.
 
@@
 Vegetable tends to be a side serving at lunch and dinner and I get celery or carrot cuts with dressing for a snack a lot of times.
-I could probably add some sort of vegetable into my breakfast (like spinach in an omelet) and snack on another piece of fruit when I quotemark m hungry rather than the junk food I tend to eat.''} incorrectly segmented after the first sentence where a punctuation mark was encountered.
+I could probably add some sort of vegetable into my breakfast (like spinach in an omelet) and snack on another piece of fruit when I am hungry rather than the junk food I tend to eat.''} incorrectly segmented after the first sentence where a punctuation mark was encountered.
 Similarly, additional information is a common cause for misclassification of an email segment into multiple segments.
 For instance, the first sentence of the above email segment represents a positive commitment to adolescent's behavior change then next two sentences provide an additional information to support their commitment.
-
-Our proposed approach is novel for the segmentation of e-Coaching emails since previous studies mainly focused on the segmentation of text into sections, headers and sentences.
-However, this study segmented clinical exchange into groups of MI behaviors which will significantly reduce the amount of resource and time required to segment clinical exchange manually.
-Furthermore, these segmentation models could be integrated with auto coding classifiers to build a software pipeline for automated annotation of clinical exchanges.
 

### 2018-08-04 plots/embedding-dimension.r

@@ -12,14 +12,15 @@ wd = '/home/mehedi/teana/projects/eCoaching-Text-Segmentation/plots/'
 data <- read.csv(paste(wd,"parameters/embedding-dimension.csv", sep=''))
 
 # plot graph
-plot(range(50,320), range(0.798, 0.826), type="n", ylab = "F1-Measure", xlab = "vector size")
+plot(range(50,320), range(0.798, 0.826), type="n", ylab = "F1-Measure", xlab = "# of word embedding dimensions")
 lines(data$dim, data$pf1, lwd=1.5, lty=1, type='b', col='green', pch=18)
-points(x = 300, y = 0.825, col = "red", pch = 18, cex = 1, lwd=1.5)
+points(x = 300, y = 0.825, col = "red", pch = 18, cex = 1, lwd=1)
+points(x = 300, y = 0.825, col = "red", pch = 1, cex = 1.5)
 text(300, 0.825, labels = c('0.825'), pos = 1, cex = 0.7)
 textxy(data$dim, data$pf1, labs=data$pf1, pos=3,  cex = 0.7)
 
 lines(data$dim, data$tf1, lwd=1.5, lty=1, type='b', col='blue', pch=17)
-points(x = 300, y = 0.816, col = "blueviolet", pch = 17, cex = 1, lwd=1.5)
+points(x = 300, y = 0.816, col = "blueviolet", pch = 17, cex = 1, lwd=1)
 text(300, 0.816, labels = c('0.816'), pos = 1, cex = 0.7)
 textxy(data$dim, data$tf1, labs=data$tf1, pos=1,  cex = 0.7)
 

### 2018-08-04 paper-amia-summit-2018/manuscript.tex
update results section

@@
 proposed a recurrent convolutional neural network for automatic identification of sentence boundary in a neuropsychological text in the Portuguese language.\cite{treviso2017sentence} Their proposed model combined prosodic and part of speech features with pre-trained GloVe word embeddings,\cite{pennington2014glove} which achieved 0.74 F1-score in sentence boundary detection on the Brazilian constitution dataset.
-
-In this study, we adopt the model proposed by Treviso et al.
-\cite{treviso2017sentence} and rename it as convolutional recurrent neural networks (CRNN) to apply it differently from the aforementioned studies.
-The main difference is that our study focuses on the segmentation of e-Coaching emails as distinct from determining the sentence boundary.
-Another important difference is that punctuation features were utilized in combination with word embeddings and part-of-speech features.
-Furthermore, we train our model with pre-trained word2vec word vectors,\cite{mikolov2013distributed} instead of using GloVe word vectors.
-Segmentation of e-Coaching emails is also different from a traditional shallow discourse analysis of conversations \cite{galley2003discourse} in that the focus is on segmentation, rather than on determining the types of transitions between the utterances or assigning utterances to speakers.
+Segmentation of e-Coaching emails is different from a traditional shallow discourse analysis of conversations \cite{galley2003discourse} in that the focus is on segmentation, rather than on determining the types of transitions between the utterances or assigning utterances to speakers.
 
@@
 \section*{Results}
-Our experimental results have two important directions.
-First, results are reported with respect to ``new segment'' class as well as weighted average over ``new segment'' and ``same segment'' classes in Table~\ref{tab:result_base}.
-Second, classification performance of different machine learning methods are outlined in Table~\ref{tab:result_weighted_avg} when word embeddings or lexical features are used in combination with punctuation and POS features.
-
-For a given sample with $2l$ words or punctuations, the first layer input of MLP was created by the summation of first $l$ word vectors and last $l$ word vectors.
-Figure~\ref{fig:length-mlp} demonstrates the performance of MLP model on e-Coaching email segmentation by varying value of $l$.
-Based on the results in Figure~\ref{fig:length-mlp}, MLP results were reported with $l = 2$, since best performance of MLP models were observed when $l = 2$.
+Our experimental results have four important directions.
+First, optimal sample size is identified by reporting MLP results with different sample size.
+Second, results are reported with respect to ``new segment'' class as well as weighted average over ``new segment'' and ``same segment'' classes in Table~\ref{tab:result_base}.
+Third, classification performance of different machine learning methods are outlined in Table~\ref{tab:result_weighted_avg} when word embeddings or lexical features are used in combination with punctuation and POS features.
+Fourth, the impact of individual fetures as well as their combination on the segmentation task are reported.
 
@@
 \end{figure}
+
+For a given sample with $2l$ words or punctuations, the first layer input of MLP was created by the summation of first $l$ word vectors and last $l$ word vectors.
+Figure~\ref{fig:length-mlp} demonstrates the performance of MLP model on e-Coaching email segmentation by varying value of $l$.
+Based on the results in Figure~\ref{fig:length-mlp}, MLP results were reported with $l = 2$, since best performance of MLP models were observed when $l = 2$.
 
@@
 
+\begin{table}[ht]
+\centering
+\caption{\textbf{Performance of CRF, MLP, BRNN and CRNN methods for identification of ``new segment'' class as well as weighted average over ``new segment'' and ``same segment'' classes when all features are used together.
+The highest value for each performance metric
+is highlighted in bold.}}
+\label{tab:result_weighted_avg}
+ \begin{tabular}{|l|l|l|l|l|l|l|}
+  \hline
+   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{New Segment}} & \multicolumn{3}{|c|}{\textbf{Overall}} \\\cline{2-7}
+   & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} \\ \hline    
+ CRF & 0.813 & 0.772 & 0.792 & 0.988 & 0.988 & 0.988 \\ \hline
+ MLP & \textbf{0.814} & 0.732 & 0.771 & 0.987 & 0.987 & 0.987 \\ \hline
+ BRNN & 0.683 & 0.820 & 0.745 & 0.985 & 0.983 & 0.984 \\ \hline
+ CRNN & 0.789 & \textbf{0.864} & \textbf{0.825} & \textbf{0.990} & \textbf{0.989} & \textbf{0.989} \\ \hline
+  \end{tabular}
+\end{table}
+
+Table~\ref{tab:result_weighted_avg} summarizes the results of all models for segmentation of e-Coaching emails when word embeddings or lexical features are used in combination with punctuation and POS features.
+Similar to results in Tables~\ref{tab:result_base}, CRNN demonstrates the highest performance among all methods achieving 0.864 recall with 0.825 F1-Measure for ``new segment'' class and 0.990 precision with 0.989 recall and F1-Measure for overall.
+BRNN and CRF show the lowest and second highest performance, respectively, for email segmentation among all methods.
+We observed that classification performance significantly improved for ``new segment'' class when word embeddings or lexical features are used in combination with punctuation and POS features.
+Precision increases by 3.96\%, -2.63\%, 12.71\% and 1.81\%; recall increases by 11.72\%, 23.44\%, 20.59\% and 8.41\%; and F1-measure increases by 8.05\%, 11.1\%, 16.22\% and 5.1\% for CRF, MLP, BRNN and CRNN methods, respectively, in new segment detection when all features are utilized together.
+Similarly, precision increases by 0.51\%, 0.51\%, 0.82\% and 0.41\%; recall increases by 0.41\%, 0.41\%, 0.72\% and 0.3\%; and F1-measure increases by 0.41\%, 0.51\%, 0.82\% and 0.3\% for CRF, MLP, BRNN and CRNN methods, respectively, in weighted average over ``new segment'' and ``same segment'' classes when word embeddings or lexical features are used in combination with punctuation and POS features.
+
 Table~\ref{tab:result_roc} shows the impact of individual fetures as well as their combination on the segmentation task.
@@
 % Proposed ROC Results: BRNN: 0.9807, MLP: 0.992, CRF: 0.9937, CRNN: 0.9859
-
-Table~\ref{tab:result_weighted_avg} summarizes the results of all models for segmentation of e-Coaching emails when word embeddings or lexical features are used in combination with punctuation and POS features.
-Similar to results in Tables~\ref{tab:result_base}, CRNN demonstrates the highest performance among all methods achieving 0.864 recall with 0.825 F1-Measure for ``new segment'' class and 0.990 precision with 0.989 recall and F1-Measure for overall.
-BRNN and CRF show the lowest and second highest performance, respectively, for email segmentation among all methods.
-We observed that classification performance significantly improved for ``new segment'' class when word embeddings or lexical features are used in combination with punctuation and POS features.
-Precision increases by 3.96\%, -2.63\%, 12.71\% and 1.81\%; recall increases by 11.72\%, 23.44\%, 20.59\% and 8.41\%; and F1-measure increases by 8.05\%, 11.1\%, 16.22\% and 5.1\% for CRF, MLP, BRNN and CRNN methods, respectively, in new segment detection when all features are utilized together.
-Similarly, precision increases by 0.51\%, 0.51\%, 0.82\% and 0.41\%; recall increases by 0.41\%, 0.41\%, 0.72\% and 0.3\%; and F1-measure increases by 0.41\%, 0.51\%, 0.82\% and 0.3\% for CRF, MLP, BRNN and CRNN methods, respectively, in weighted average over ``new segment'' and ``same segment'' classes when word embeddings or lexical features are used in combination with punctuation and POS features.\\
-
-\begin{table}[ht]
-\centering
-\caption{\textbf{Performance of CRF, MLP, BRNN and CRNN methods for identification of ``new segment'' class as well as weighted average over ``new segment'' and ``same segment'' classes when all features are used together.
-The highest value for each performance metric
-is highlighted in bold.}}
-\label{tab:result_weighted_avg}
- \begin{tabular}{|l|l|l|l|l|l|l|}
-  \hline
-   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{New Segment}} & \multicolumn{3}{|c|}{\textbf{Overall}} \\\cline{2-7}
-   & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} \\ \hline    
- CRF & 0.813 & 0.772 & 0.792 & 0.988 & 0.988 & 0.988 \\ \hline
- MLP & \textbf{0.814} & 0.732 & 0.771 & 0.987 & 0.987 & 0.987 \\ \hline
- BRNN & 0.683 & 0.820 & 0.745 & 0.985 & 0.983 & 0.984 \\ \hline
- CRNN & 0.789 & \textbf{0.864} & \textbf{0.825} & \textbf{0.990} & \textbf{0.989} & \textbf{0.989} \\ \hline
-  \end{tabular}
-\end{table}
 

### 2018-08-04 paper-amia-summit-2018/manuscript.tex
update results section

@@
 Figure~\ref{fig:length-mlp} demonstrates the performance of MLP model on e-Coaching email segmentation by varying value of $l$.
-Based on the results in Figure~\ref{fig:length-mlp}, MLP results were reported with $l = 2$, since best performance of MLP models were observed when $l = 2$.
+From the results illustrated in Figure~\ref{fig:length-mlp}, we observed that best performance is achieved when $l = 2$.
+Therefore, all MLP results are reported with $l = 2$ in the remaining part of this section.
 
@@
  CRF & 0.813 & 0.772 & 0.792 & 0.988 & 0.988 & 0.988 \\ \hline
- MLP & \textbf{0.814} & 0.732 & 0.771 & 0.987 & 0.987 & 0.987 \\ \hline
+ MLP & \textbf{0.817} & 0.710 & 0.760 & 0.986 & 0.987 & 0.986 \\ \hline
  BRNN & 0.683 & 0.820 & 0.745 & 0.985 & 0.983 & 0.984 \\ \hline
@@
 We observed that classification performance significantly improved for ``new segment'' class when word embeddings or lexical features are used in combination with punctuation and POS features.
-Precision increases by 3.96\%, -2.63\%, 12.71\% and 1.81\%; recall increases by 11.72\%, 23.44\%, 20.59\% and 8.41\%; and F1-measure increases by 8.05\%, 11.1\%, 16.22\% and 5.1\% for CRF, MLP, BRNN and CRNN methods, respectively, in new segment detection when all features are utilized together.
-Similarly, precision increases by 0.51\%, 0.51\%, 0.82\% and 0.41\%; recall increases by 0.41\%, 0.41\%, 0.72\% and 0.3\%; and F1-measure increases by 0.41\%, 0.51\%, 0.82\% and 0.3\% for CRF, MLP, BRNN and CRNN methods, respectively, in weighted average over ``new segment'' and ``same segment'' classes when word embeddings or lexical features are used in combination with punctuation and POS features.
+Precision increases by 3.96\%, -2.27\%, 12.71\% and 1.81\%; recall increases by 11.72\%, 19.73\%, 20.59\% and 8.41\%; and F1-measure increases by 8.05\%, 9.51\%, 16.22\% and 5.1\% for CRF, MLP, BRNN and CRNN methods, respectively, in new segment detection when all features are utilized together.
+Similarly, precision increases by 0.51\%, 0.41\%, 0.82\% and 0.41\%; recall increases by 0.41\%, 0.41\%, 0.72\% and 0.3\%; and F1-measure increases by 0.41\%, 0.41\%, 0.82\% and 0.3\% for CRF, MLP, BRNN and CRNN methods, respectively, in weighted average over ``new segment'' and ``same segment'' classes when word embeddings or lexical features are used in combination with punctuation and POS features.
 
-Table~\ref{tab:result_roc} shows the impact of individual fetures as well as their combination on the segmentation task.
+Table~\ref{tab:result_roc} shows the impact of individual features as well as their combination on the segmentation task.
 Although CRF provides the second highest classification results, it shows best AUC values, achieving 0.966 AUC when only lexical features are used and 0.994 AUC when a combination of lexical, punctuation and POS features are used.
-Influence of the punctuation and POS features is also consistent in AUC values, which increase by 2.9\%, 4.31\%, 6.63\% and 2.18\% for CRF, MLP, BRNN and CRNN methods, respectively.\\
+Influence of the punctuation and POS features is also consistent in AUC values, which increase by 2.9\%, 4.31\%, 6.63\% and 2.18\% for CRF, MLP, BRNN and CRNN methods, respectively.
+Individually, POS and punctuation features improve the performance of all classifiers except CRNN when word embeddings are used with POS features.
+CRF and MLP achieve their highest AUCs when all features are used together.
+On the other hand, BRNN and CRNN demonstrate the highest AUCs when word embeddings are combined with punctuation features.
+\\
 
@@
  word embeddings only & \textbf{0.966} & 0.951 & 0.920 & 0.965 \\ \hline
- word embeddings + POS & \textbf{0.966} & 0.951 & 0.920 & 0.965 \\ \hline
- word embeddings + punctuation & \textbf{0.966} & 0.951 & 0.920 & 0.965 \\ \hline
+ word embeddings + POS & \textbf{0.973} & 0.953 & 0.920 & 0.962 \\ \hline
+ word embeddings + punctuation & \textbf{0.993} & 0.989 & 0.982 & 0.988 \\ \hline
  all features & \textbf{0.994} & 0.992 & 0.981 & 0.986 \\ \hline

### 2018-08-04 paper-amia-summit-2018/manuscript.tex
Fixed all issues reviewed by Alex on the first draft

@@
 A critical step towards automated annotation of e-Coaching communication exchanges is segmentation of emails into textual fragments that correspond to MI behaviors.
-In this work, we formulate this task as a classification problem and propose word embedding, punctuation and part-of-speech features to address it.
+In this work, we formulate this task as a classification problem and propose word embeddings, punctuation and part-of-speech features to address it.
 We experimented both with traditional machine learning method, conditional random fields (CRF) and deep learning methods, such as multilayer perceptrons (MLP), bidirectional recurrent neural networks (BRNN) and convolutional recurrent neural networks (CRNN).
@@
 \begin{enumerate}
-\item Emails are unstructured text that contains informal information exchange in a non-traditional format. However, a traditional MI transcript corresponds to an MI interview
-session, contains a dialogue between a counselor and a patient.
+\item Emails are unstructured text that contains informal information exchange in a non-traditional format.
+For example, an e-Coach usually responds to several queries of a patient in one email.
+However, a provider utterance of MI transcript denotes a particular response for a query of a patient during a traditional MI interview session, which involve a dialogue between a provider and a patient.
 \item Discourse segments in e-Coaching emails do not necessarily correspond to sentences or collection of sentences.
@@
 The goal of this research study is to assess the applicability of machine learning methods for automated segmentation of e-Coaching emails into textual fragments corresponding to individual behaviors, which is the first step of the coding process of e-Coaching communications.
-In particular, we introduced word embedding or lexical, punctuation and  part-of-speech (POS) features and experimented with both traditional supervised machine learning method, linear-chain conditional random fields (CRF)\cite{lafferty2001conditional} and deep learning methods, such as multilayer perceptrons (MLP),\cite{rumelhart1986learning} bidirectional recurrent neural networks (BRNN)\cite{schuster1997bidirectional} and convolutional recurrent neural networks (CRNN),\cite{treviso2017sentence} to find the best performing method and feature combination.
+In particular, we introduced word embeddings or lexical, punctuation and  part-of-speech (POS) features and experimented with both traditional supervised machine learning method, linear-chain conditional random fields (CRF)\cite{lafferty2001conditional} and deep learning methods, such as multilayer perceptrons (MLP),\cite{rumelhart1986learning} bidirectional recurrent neural networks (BRNN)\cite{schuster1997bidirectional} and convolutional recurrent neural networks (CRNN),\cite{treviso2017sentence} to find the best performing method and feature combination.
 
@@
 In total, we obtained 95,777 words and 7,140 punctuation marks, which include 3,138 ``new segment'' and 99,779 ``same segment'' instances.
-In this study, we experimented with traditional machine learning method, conditional random fields (CRF)\cite{lafferty2001conditional} and deep learning methods, such as multilayer perceptrons (MLP),\cite{rumelhart1986learning} bidirectional recurrent neural networks (BRNN)\cite{schuster1997bidirectional} and convolutional recurrent neural networks (CRNN).\cite{treviso2017sentence} For MLP models, samples were created with $2l$ words or punctuations at each step in a word sequence, where a sample contains next $l$ words or punctuations and prior $l$ words or punctuations including current word or punctuation.
+In this study, we experimented with traditional machine learning method, conditional random fields (CRF)\cite{lafferty2001conditional} and deep learning methods, such as multilayer perceptrons (MLP),\cite{rumelhart1986learning} bidirectional recurrent neural networks (bidirectional RNN or simply BRNN)\cite{schuster1997bidirectional} and convolutional recurrent neural networks (CRNN).\cite{treviso2017sentence} For MLP models, samples were created with $2l$ words or punctuations at each step in a word sequence, where a sample contains next $l$ words or punctuations and prior $l$ words or punctuations including current word or punctuation.
 Each sample is classified into either ``new segment'' or ``same segment'' based on whether the $l^{th}$ word or punctuation in the gold standard precedes a new segment or not.
-For CRF, BRNN and CRNN models, an email was taken as input sequence, such that POS tags and word embedding of each word or punctuation mark were used as input and binary labels (1 or 0) corresponding to ``new segment'' and ``same segment'' classification decision were considered as the output of the model at each step.
+For CRF, BRNN and CRNN models, an email was taken as input sequence, such that POS tags and word embeddings of each word or punctuation mark were used as input and binary labels (1 or 0) corresponding to ``new segment'' and ``same segment'' classification decision were considered as the output of the model at each step.
 In the gold standard, words or punctuations within the same segment were assigned the label of 0 and the last word or punctuation mark of a segment were assigned the label of 1.
@@
 Word embeddings are one of this representation where each word represented as a real-valued vector in a high dimensional vector space.
-The distributed representation of word allows capturing semantic, syntactic and morphological information from large unannotated corpora.\cite{pennington2014glove, mikolov2013distributed} In our experiment, we utilized embeddings estimated on Google News corpus consisting of 1.6 billion words referred to pre-trained word2vec word embeddings.\footnote{https://code.google.com/p/word2vec/} When words or punctuation marks are not available in the pre-trained word vectors, we utilized word embeddings trained with our e-Coaching email corpus referred to trained word embeddings.
-Figure~\ref{fig:embedding-dimension} illustrates the performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and trained word embeddings with GloVe and word2vec models.
-It was observed that best performance achieved with pre-trained 300-dimensional word2vec word vectors when three types of features are used together.
-Therefore, we report our results with word2vec 300-dimensional word vectors for all deep learning models used in this study.
+The distributed representation of word allows capturing semantic, syntactic and morphological information from large unannotated corpora.\cite{pennington2014glove, mikolov2013distributed} In our experiment, we utilized embeddings estimated on Google News corpus consisting of 1.6 billion words referred to pre-trained word2vec word embeddings.\footnote{https://code.google.com/p/word2vec/} When words or punctuation marks are not found in the pre-trained word vectors, we utilized word embeddings trained with our e-Coaching email corpus referred to trained word embeddings.
 CRF utilized lexical features which represent each word in an e-Coaching email.
-
-\begin{figure}[!htb]
-    \centering
-    \includegraphics[width=0.5\textwidth]{figures/embedding-dimension.eps}
-    \caption{\textbf{Performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and trained word embeddings with GloVe and word2vec models}}
-    \label{fig:embedding-dimension}
-\end{figure}
 
@@
 
-\textbf{Multilayer Perceptrons}: MLP is a feed-forward artificial neural network which maps set of input onto a set of appropriate output.\cite{rumelhart1986learning} Figure~\ref{fig:mlp} shows a multi layer perceptron with a single hidden layer.
-In an MLP networks, there are no cycles or loops where the information moves forward only, from the input nodes through the hidden nodes and to the output nodes.
+\textbf{Multilayer Perceptrons}: MLP is a feed-forward artificial neural network which maps set of input onto a set of appropriate output.\cite{rumelhart1986learning} Figure~\ref{fig:mlp} shows a multilayer perceptron with a single hidden layer.
+In MLP networks, there is no cycles or loops and information moves forward only, from the input nodes through the hidden nodes and to the output nodes.
 This study utilized MLP with a nonlinear activation function (relu) and one hidden layer consisting of 128 hidden units.
@@
 Our experimental results have four important directions.
-First, optimal sample size is identified by reporting MLP results with different sample size.
+First, the optimal dimension of word embeddings and sample size of MLP models are reported.
 Second, results are reported with respect to ``new segment'' class as well as weighted average over ``new segment'' and ``same segment'' classes in Table~\ref{tab:result_base}.
 Third, classification performance of different machine learning methods are outlined in Table~\ref{tab:result_weighted_avg} when word embeddings or lexical features are used in combination with punctuation and POS features.
-Fourth, the impact of individual fetures as well as their combination on the segmentation task are reported.
+Fourth, the impact of the individual, as well as their combination on the segmentation task, are reported.
+
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=0.5\textwidth]{figures/embedding-dimension.eps}
+    \caption{\textbf{Performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and trained word embeddings with GloVe and word2vec models}}
+    \label{fig:embedding-dimension}
+\end{figure}
+
+Figure~\ref{fig:embedding-dimension} illustrates the performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and trained word embeddings with GloVe and word2vec models.
+It was observed that best performance achieved with pre-trained 300-dimensional word2vec word vectors when three types of features are used together.
+Therefore, we report our results with word2vec 300-dimensional word vectors for all deep learning models used in this study.
+For MLP models, the first layer input was prepared by the summation of first $l$ word vectors and last $l$ word vectors when a sample contains $2l$ words or punctuations.
+Figure~\ref{fig:length-mlp} demonstrates the performance of MLP model on e-Coaching email segmentation by varying the value of $l$.
+It was observed that the best performance of MLP is achieved when $l = 2$.
+Therefore, results of all MLP are reported with $l = 2$ in the remaining part of this section.
 
@@
 \end{figure}
-
-For a given sample with $2l$ words or punctuations, the first layer input of MLP was created by the summation of first $l$ word vectors and last $l$ word vectors.
-Figure~\ref{fig:length-mlp} demonstrates the performance of MLP model on e-Coaching email segmentation by varying value of $l$.
-From the results illustrated in Figure~\ref{fig:length-mlp}, we observed that best performance is achieved when $l = 2$.
-Therefore, all MLP results are reported with $l = 2$ in the remaining part of this section.
 
@@
 Precision increases by 3.96\%, -2.27\%, 12.71\% and 1.81\%; recall increases by 11.72\%, 19.73\%, 20.59\% and 8.41\%; and F1-measure increases by 8.05\%, 9.51\%, 16.22\% and 5.1\% for CRF, MLP, BRNN and CRNN methods, respectively, in new segment detection when all features are utilized together.
-Similarly, precision increases by 0.51\%, 0.41\%, 0.82\% and 0.41\%; recall increases by 0.41\%, 0.41\%, 0.72\% and 0.3\%; and F1-measure increases by 0.41\%, 0.41\%, 0.82\% and 0.3\% for CRF, MLP, BRNN and CRNN methods, respectively, in weighted average over ``new segment'' and ``same segment'' classes when word embeddings or lexical features are used in combination with punctuation and POS features.
-
-Table~\ref{tab:result_roc} shows the impact of individual features as well as their combination on the segmentation task.
-Although CRF provides the second highest classification results, it shows best AUC values, achieving 0.966 AUC when only lexical features are used and 0.994 AUC when a combination of lexical, punctuation and POS features are used.
-Influence of the punctuation and POS features is also consistent in AUC values, which increase by 2.9\%, 4.31\%, 6.63\% and 2.18\% for CRF, MLP, BRNN and CRNN methods, respectively.
-Individually, POS and punctuation features improve the performance of all classifiers except CRNN when word embeddings are used with POS features.
-CRF and MLP achieve their highest AUCs when all features are used together.
-On the other hand, BRNN and CRNN demonstrate the highest AUCs when word embeddings are combined with punctuation features.
-\\
+Similarly, precision increases by 0.51\%, 0.41\%, 0.82\% and 0.41\%; recall increases by 0.41\%, 0.41\%, 0.72\% and 0.3\%; and F1-measure increases by 0.41\%, 0.41\%, 0.82\% and 0.3\% for CRF, MLP, BRNN and CRNN methods, respectively, in weighted average over ``new segment'' and ``same segment'' classes when word embeddings or lexical features are used in combination with punctuation and POS features.\\
 
@@
 \centering
-\caption{\textbf{AUC values of all classifiers when only word embeddings or lexical features and combination of word embeddings or lexical, punctuation and POS features are used.
-Highest value for each experiment across all models is highlighted in boldface.}}
+\caption{\textbf{AUC values of all classifiers demonstrating the impact of word embeddings, punctuation and POS features on email segmentation.
+Highest AUC value for each feature set across all models is highlighted in boldface.}}
 \label{tab:result_roc}
@@
 % Proposed ROC Results: BRNN: 0.9807, MLP: 0.992, CRF: 0.9937, CRNN: 0.9859
+
+Table~\ref{tab:result_roc} shows the impact of individual features as well as their combination on the segmentation task.
+Although CRF provides the second highest classification results, it shows best AUC values, achieving 0.966 AUC when only lexical features are used and 0.994 AUC when a combination of lexical, punctuation and POS features are used.
+Influence of the punctuation and POS features is also consistent in AUC values, which increase by 2.9\%, 4.31\%, 6.63\% and 2.18\% for CRF, MLP, BRNN and CRNN methods, respectively.
+Individually, POS and punctuation features improve the performance of all classifiers except CRNN when word embeddings are used with POS features.
+CRF and MLP achieve their highest AUCs when all features are used together.
+On the other hand, BRNN and CRNN demonstrate the highest AUCs when word embeddings are combined with punctuation features.
 

### 2018-08-04 paper-amia-summit-2018/manuscript.tex
Fixed all issues reviewed by Alex on the first draft

@@
 In addition, standard metrics: precision, recall and F1-measure were used to eliminate doubt about the model performance since accuracy is misleading for imbalanced datasets.
-AUC is also utilized due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets,\cite{hu2015kernelized} which are demonstrated by the ROC curves in Figure~\ref{fig:roc-curves}.
+AUC is also utilized due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets,\cite{hu2015kernelized} which are also demonstrated by the ROC curves in Figure~\ref{fig:roc-curves}.
 

### 2018-08-05 paper-amia-summit-2018/manuscript.tex
update manuscript to make 10 pages

@@
 
-Figure~\ref{fig:text-segment} illustrates a segmentation of an e-Coaching email exchange, in which the first sentence is segmented into 2 MI behavior fragments, while the fourth and fifth segments correspond to one and three sentences, respectively.
-Segmentation of e-Coaching emails corresponds to a special type of discourse analysis \cite{webber2012discourse} aimed at better understanding the effective e-Coaching communication strategies and revealing the unique socio-psychological characteristics of a patient.
-
-\begin{figure}[!htb]
-    \centering
-    \includegraphics[width=0.7\textwidth]{figures/segment-example.png}
-    \caption{\textbf{Example of e-Coaching emails segmented into fragments that correspond to MI behaviors of an e-Coach and a patient}}
-    \label{fig:text-segment}
-\end{figure}
-
 Recently, an online clinical intervention called MENU GenY (Making Effective Nutrition Choices for Generation Y) was proposed and evaluated.
@@
 Segmentation of e-Coaching emails is different from a traditional shallow discourse analysis of conversations \cite{galley2003discourse} in that the focus is on segmentation, rather than on determining the types of transitions between the utterances or assigning utterances to speakers.
+Segmentation of e-Coaching emails corresponds to a special type of discourse analysis \cite{webber2012discourse} aimed at better understanding the effective e-Coaching communication strategies and revealing the unique socio-psychological characteristics of a patient.
 
@@
 This study utilized MLP with a nonlinear activation function (relu) and one hidden layer consisting of 128 hidden units.
-In order to prevent over-fitting, we implement dropout on fully connected layers, which randomly hide some neurons during the training phase.
+In order to prevent over-fitting, we implement dropout on last layer, which randomly hide some neurons during the training phase.
 \cite{srivastava2014dropout} Dropout is also applied to the fully connected layers in BRNN and CRNN models.
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{figures/mlp.jpeg}
-    \caption{\textbf{A multi layer perceptron with a single hidden layer}}
+    \includegraphics[width=0.45\textwidth]{figures/mlp.eps}
+    \caption{\textbf{A multilayer perceptron with a single hidden layer}}
     \label{fig:mlp}
@@
 \textbf{Convolutional Recurrent Neural Networks}: CRNN is a deep neural networks architecture,\cite{treviso2017sentence} shown in Figure~\ref{fig:crnn}, consisting of 5 layers: 1) input layer 2) embedding layer 3) convolution layer with max pooling 4) BRNN layer 5) fully connected layer with dropout and sigmoid output.
-
-\begin{figure}[!htb]
-    \centering
-    \includegraphics[width=1.0\textwidth]{figures/CRNN.eps}
-    \caption{\textbf{Architecture of convolutional recurrent neural networks for automatic segmentation of e-Coaching emails}}
-    \label{fig:crnn}
-\end{figure}
-
 E-coaching email exchanges are represented as a sequence of $m$ words and punctuations, which are fed into the input layer to produce a $m \times n_e$ matrix after fetching the pre-trained word vectors.
@@
 
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=1.0\textwidth]{figures/CRNN.eps}
+    \caption{\textbf{Architecture of convolutional recurrent neural networks for automatic segmentation of e-Coaching emails}}
+    \label{fig:crnn}
+\end{figure}
+
 \subsection*{\textit{Evaluation metrics}}
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{figures/embedding-dimension.eps}
-    \caption{\textbf{Performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and trained word embeddings with GloVe and word2vec models}}
-    \label{fig:embedding-dimension}
+    \includegraphics[width=1.0\textwidth]{figures/mlp-and-vector.eps}
+    \caption{\textbf{Performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and trained word embeddings with GloVe and word2vec models (left).
+Performance of MLP model on e-Coaching email segmentation by varying the length of sample size (right).}}
+    \label{fig:embedding-dimension-mlp}
 \end{figure}
 
-Figure~\ref{fig:embedding-dimension} illustrates the performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and trained word embeddings with GloVe and word2vec models.
+Figure~\ref{fig:embedding-dimension-mlp} (left) illustrates the performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and trained word embeddings with GloVe and word2vec models.
 It was observed that best performance achieved with pre-trained 300-dimensional word2vec word vectors when three types of features are used together.
@@
 For MLP models, the first layer input was prepared by the summation of first $l$ word vectors and last $l$ word vectors when a sample contains $2l$ words or punctuations.
-Figure~\ref{fig:length-mlp} demonstrates the performance of MLP model on e-Coaching email segmentation by varying the value of $l$.
-It was observed that the best performance of MLP is achieved when $l = 2$.
-Therefore, results of all MLP are reported with $l = 2$ in the remaining part of this section.
-
-\begin{figure}[!htb]
-    \centering
-    \includegraphics[width=0.5\textwidth]{figures/length-mlp.eps}
-    \caption{\textbf{Performance of MLP model on e-Coaching email segmentation by varying the length of sample size}}
-    \label{fig:length-mlp}
-\end{figure}
+Figure~\ref{fig:embedding-dimension-mlp} (right) demonstrates the performance of MLP model on e-Coaching email segmentation by varying the value of $l$.
+It was observed that the best performance of MLP classifier is achieved when $l = 2$.
+Therefore, results of all MLP models are reported with $l = 2$.
 
@@
   \hline
-\multirow{2}{*}{\textbf{Features}} & \multicolumn{4}{|c|}{\textbf{AUC}} \\\cline{2-5}
- & \textbf{CRF} & \textbf{MLP}  & \textbf{BRNN} & \textbf{CRNN} \\ \hline      
+\textbf{Features} & \textbf{CRF} & \textbf{MLP}  & \textbf{BRNN} & \textbf{CRNN} \\ \hline      
  word embeddings only & \textbf{0.966} & 0.951 & 0.920 & 0.965 \\ \hline

### 2018-08-05 plots/preprocess-roc-data.py

@@ -34,4 +34,7 @@ def create_roc_csv(path):
 # call above functions for the creation of new files
 create_roc_csv('roc/base')
 create_roc_csv('roc/propose')
+create_roc_csv('roc/pos')
+create_roc_csv('roc/punctuation')
+
 print('conversion is done!')

### 2018-08-05 paper-amia-summit-2018/manuscript.tex
added footnote for GloVe word embeddings

@@
 
-Figure~\ref{fig:embedding-dimension-mlp} (left) illustrates the performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and trained word embeddings with GloVe and word2vec models.
+Figure~\ref{fig:embedding-dimension-mlp} (left) illustrates the performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and trained word embeddings with GloVe\footnote{https://nlp.stanford.edu/projects/glove/} and word2vec models.
 It was observed that best performance achieved with pre-trained 300-dimensional word2vec word vectors when three types of features are used together.

### 2018-08-07 paper-amia-summit-2018/manuscript.tex
finalize abstract with 150 words

@@
 This study focuses on the analysis of e-Coaching sessions, behavior interventions that are delivered via email and grounded in the principles of MI.
-A critical step towards automated annotation of e-Coaching communication exchanges is segmentation of emails into textual fragments that correspond to MI behaviors.
-In this work, we formulate this task as a classification problem and propose word embeddings, punctuation and part-of-speech features to address it.
+A critical step towards automated annotation of e-Coaching communication exchanges is segmentation of emails into fragments that correspond to MI behaviors.
+This study formulates this task as a classification problem and proposes word embeddings, punctuation and part-of-speech features to address it.
 We experimented both with traditional machine learning method, conditional random fields (CRF) and deep learning methods, such as multilayer perceptrons (MLP), bidirectional recurrent neural networks (BRNN) and convolutional recurrent neural networks (CRNN).

### 2018-08-07 paper-amia-summit-2018/manuscript.tex
fixed small issues of writing such as gramatical mistakes and spelling errors

@@
 The unstructured nature of e-Coaching exchanges poses a unique set of challenges for their qualitative analysis.
-Usually, email segments are automatically coded with the Minority Youth-Sequential Coding of Process Exchanges (MYSCOPE), \cite{carcone2013provider} a qualitative coding scheme to characterize patient-coach communication during email exchanges.
+Usually, email segments are automatically coded with the Minority Youth-Sequential Coding of Process Exchanges (MYSCOPE), \cite{carcone2013provider} a qualitative coding scheme to characterize patient-provider communication during MI session.
 A significant barrier to fully automating the behavior coding process of e-Coaching emails is their segmentation into textual fragments that correspond to distinct communication behaviors.
@@
 applied SVM along with word-vector cosine similarity metric combined with several heuristics to segment clinical reports in EHR into sections, such as demographics, history, procedure, finding and impression.\cite{apostolova2009automatic} After identification of each line in the document, Tepper et al.
-trained Maximum Entropy models to classify the document into pre-defined sections, such as general patient information, provider information, medical condition, medical history, clinical history, findings, impression, etc.\cite{tepper2012statistical} Denny et al.
+trained maximum entropy models to classify the document into pre-defined sections, such as general patient information, provider information, medical condition, medical history, clinical history, findings, impression, etc.\cite{tepper2012statistical} Denny et al.
 proposed a SecTag algorithm, which combined natural language processing techniques, terminology-based rules and a Naive Bayes classifier to identify the sections and headers that achieved 99\% recall with 95.6\% precision.
@@
 Liu et al.
-have shown that a linear chain CRF performed better than the hidden markov model and Maxent in sentence boundary detection, evaluated using the official
-NIST evaluation tools.\cite{liu2005using} Recently, Treviso et al.
+have shown that a linear chain CRF performed better than the hidden markov and maximum entropy model in sentence boundary detection, which was evaluated using the official NIST evaluation tools.\cite{liu2005using} Recently, Treviso et al.
 proposed a recurrent convolutional neural network for automatic identification of sentence boundary in a neuropsychological text in the Portuguese language.\cite{treviso2017sentence} Their proposed model combined prosodic and part of speech features with pre-trained GloVe word embeddings,\cite{pennington2014glove} which achieved 0.74 F1-score in sentence boundary detection on the Brazilian constitution dataset.
@@
 
-The experimental dataset for this work was constructed from 49 e-coaching sessions, which include a total of 3,138 segmented and annotated MI behaviors.
+The experimental dataset for this work was constructed from 49 e-Coaching sessions, which include a total of 3,138 segmented and annotated MI behaviors.
 Each session represents an MI intervention delivered via email.
@@
 In total, we obtained 95,777 words and 7,140 punctuation marks, which include 3,138 ``new segment'' and 99,779 ``same segment'' instances.
-In this study, we experimented with traditional machine learning method, conditional random fields (CRF)\cite{lafferty2001conditional} and deep learning methods, such as multilayer perceptrons (MLP),\cite{rumelhart1986learning} bidirectional recurrent neural networks (bidirectional RNN or simply BRNN)\cite{schuster1997bidirectional} and convolutional recurrent neural networks (CRNN).\cite{treviso2017sentence} For MLP models, samples were created with $2l$ words or punctuations at each step in a word sequence, where a sample contains next $l$ words or punctuations and prior $l$ words or punctuations including current word or punctuation.
-Each sample is classified into either ``new segment'' or ``same segment'' based on whether the $l^{th}$ word or punctuation in the gold standard precedes a new segment or not.
+In this study, we experimented with traditional machine learning method, conditional random fields (CRF)\cite{lafferty2001conditional} and deep learning methods, such as multilayer perceptrons (MLP),\cite{rumelhart1986learning} bidirectional recurrent neural networks (bidirectional RNN or simply BRNN)\cite{schuster1997bidirectional} and convolutional recurrent neural networks (CRNN).\cite{treviso2017sentence} For MLP models, instances were created with $2l$ words or punctuations at each step in a word sequence, where an instance contains next $l$ words or punctuations and prior $l$ words or punctuations including current word or punctuation.
+Each instance is classified into either ``new segment'' or ``same segment'' based on whether the $l^{th}$ word or punctuation in the gold standard precedes a new segment or not.
 For CRF, BRNN and CRNN models, an email was taken as input sequence, such that POS tags and word embeddings of each word or punctuation mark were used as input and binary labels (1 or 0) corresponding to ``new segment'' and ``same segment'' classification decision were considered as the output of the model at each step.
@@
 Word embeddings are one of this representation where each word represented as a real-valued vector in a high dimensional vector space.
-The distributed representation of word allows capturing semantic, syntactic and morphological information from large unannotated corpora.\cite{pennington2014glove, mikolov2013distributed} In our experiment, we utilized embeddings estimated on Google News corpus consisting of 1.6 billion words referred to pre-trained word2vec word embeddings.\footnote{https://code.google.com/p/word2vec/} When words or punctuation marks are not found in the pre-trained word vectors, we utilized word embeddings trained with our e-Coaching email corpus referred to trained word embeddings.
+The distributed representation of word allows capturing semantic, syntactic and morphological information from large unannotated corpora.\cite{pennington2014glove, mikolov2013distributed} This study utilized embeddings estimated on Google News corpus consisting of 1.6 billion words referred to pre-trained word2vec word embeddings.\footnote{\label{fn:word2vec}https://code.google.com/p/word2vec/} When words or punctuation marks are not found in the pre-trained word vectors, word embeddings trained with e-Coaching email corpus referred to trained word embeddings were utilized.
 CRF utilized lexical features which represent each word in an e-Coaching email.
@@
 
-\textbf{Conditional Random Fields}: CRF has been widely used in various NLP tasks including part-of-speech tagging and segmentation tasks.\cite{lafferty2001conditional, hirohata2008identifying} Unlike a maximum entropy Markov model which uses per-state exponential models for the conditional probabilities of next states given the current state, a CRF is modeled with an exponential distribution provides the conditional distribution of the entire output sequence given the observation sequence.
+\textbf{Conditional Random Fields}: CRF has been widely used in various NLP tasks including part-of-speech tagging and segmentation tasks.\cite{lafferty2001conditional, hirohata2008identifying} Unlike a maximum entropy markov model which uses per-state exponential models for the conditional probabilities of next states given the current state, a CRF is modeled with an exponential distribution provides the conditional distribution of the entire output sequence given the observation sequence.
 A traditional linear-chain CRF model is defined as a conditional probability distribution $p(y|x)$ for output and input sequences, $y$ and $x$, by the Eq.
@@
 \textbf{Multilayer Perceptrons}: MLP is a feed-forward artificial neural network which maps set of input onto a set of appropriate output.\cite{rumelhart1986learning} Figure~\ref{fig:mlp} shows a multilayer perceptron with a single hidden layer.
-In MLP networks, there is no cycles or loops and information moves forward only, from the input nodes through the hidden nodes and to the output nodes.
-This study utilized MLP with a nonlinear activation function (relu) and one hidden layer consisting of 128 hidden units.
-In order to prevent over-fitting, we implement dropout on last layer, which randomly hide some neurons during the training phase.
+In MLP networks, there is no cycle or loop and information moves forward only, from the input nodes through the hidden nodes and to the output nodes.
+This study utilized MLP with a nonlinear activation function ($relu$) and one hidden layer consisting of 128 hidden units.
+In order to prevent over-fitting, we implement dropout on the last layer, which randomly hides some neurons during the training phase.
 \cite{srivastava2014dropout} Dropout is also applied to the fully connected layers in BRNN and CRNN models.
@@
 \textbf{Convolutional Recurrent Neural Networks}: CRNN is a deep neural networks architecture,\cite{treviso2017sentence} shown in Figure~\ref{fig:crnn}, consisting of 5 layers: 1) input layer 2) embedding layer 3) convolution layer with max pooling 4) BRNN layer 5) fully connected layer with dropout and sigmoid output.
-E-coaching email exchanges are represented as a sequence of $m$ words and punctuations, which are fed into the input layer to produce a $m \times n_e$ matrix after fetching the pre-trained word vectors.
+E-coaching email exchanges are represented as a sequence of $m$ words and punctuations, which are fed into the input layer.
+The embedding layer provides a distributed representation of words, gives us $m \times n_e$ matrix after fetching the pre-trained word vectors.
 When POS tags are used in combination with word embeddings, 10-dimensional POS vectors are concatenated with 300-dimensional word vectors to obtain new vectors $n_e = n_w + n_p$ of size 310.
-The primary purpose of convolution is to extract new features depending on neighbor words.
-We used 1D convolution which simply moves one step at a time on vertically.
-In this layer, one filter is responsible for extraction of one feature.
+The convolution layer used this word embeddings to extract new features depending on neighbor words.
+This study used 1D convolution which simply moves one step at a time on vertically.
+In this layer, one filter is responsible for the extraction of one feature.
 After applying $n_f$ different filters with zero-padding on both sides of the text, $n_f$ useful features are produced in the convolution layer for each word.
@@
 Moreover, bidirectional RNN is capable of capturing both previous and past histories because bidirectional RNN utilized the features by looking at forward and backward states.
-The purpose of the fully connected layer is to use the bidirectional RNN layer output for classifying each word or punctuation into ``new segment'' and ``same segment'' classes.
+The fully connected layer used the bidirectional RNN output for classifying each word or punctuation into ``new segment'' and ``same segment'' classes.
 Since a fully connected layer has a greater number of parameters, they are more likely excessively co-adapt these parameters causing over-fitting.
@@
 Our experimental results have four important directions.
-First, the optimal dimension of word embeddings and sample size of MLP models are reported.
+First, the optimal word embedding dimension and instance size of MLP models are illustrated in Figure~\ref{fig:embedding-dimension-mlp}.
 Second, results are reported with respect to ``new segment'' class as well as weighted average over ``new segment'' and ``same segment'' classes in Table~\ref{tab:result_base}.
 Third, classification performance of different machine learning methods are outlined in Table~\ref{tab:result_weighted_avg} when word embeddings or lexical features are used in combination with punctuation and POS features.
-Fourth, the impact of the individual, as well as their combination on the segmentation task, are reported.
+Fourth, the impact of individual features, as well as their combination on the task of email segmentation, are demonstrated in Table~\ref{tab:result_roc}.
 
@@
     \caption{\textbf{Performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and trained word embeddings with GloVe and word2vec models (left).
-Performance of MLP model on e-Coaching email segmentation by varying the length of sample size (right).}}
+Performance of MLP model on e-Coaching email segmentation by varying the length of instance (right).}}
     \label{fig:embedding-dimension-mlp}
@@
 
-Figure~\ref{fig:embedding-dimension-mlp} (left) illustrates the performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and trained word embeddings with GloVe\footnote{https://nlp.stanford.edu/projects/glove/} and word2vec models.
-It was observed that best performance achieved with pre-trained 300-dimensional word2vec word vectors when three types of features are used together.
-Therefore, we report our results with word2vec 300-dimensional word vectors for all deep learning models used in this study.
-For MLP models, the first layer input was prepared by the summation of first $l$ word vectors and last $l$ word vectors when a sample contains $2l$ words or punctuations.
-Figure~\ref{fig:embedding-dimension-mlp} (right) demonstrates the performance of MLP model on e-Coaching email segmentation by varying the value of $l$.
-It was observed that the best performance of MLP classifier is achieved when $l = 2$.
+Figure~\ref{fig:embedding-dimension-mlp} (left) shows the performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and trained word embeddings with GloVe\footnote{https://nlp.stanford.edu/projects/glove/} and word2vec models.
+It was observed that the best performance of email segmentation was achieved with pre-trained 300-dimensional word2vec word vectors when three types of features are used together.
+Therefore, results of all deep learning models are reported with pre-trained word2vec 300-dimensional word vectors.
+In MLP, inputs were prepared by computing the summation of first $l$ word vectors and last $l$ word vectors where each instance contains $2l$ words or punctuations.
+Figure~\ref{fig:embedding-dimension-mlp} (right) demonstrates the performance of MLP models on e-Coaching email segmentation by varying the value of $l$.
+It was observed that the best performance of MLP classifiers is achieved when $l = 2$.
 Therefore, results of all MLP models are reported with $l = 2$.
@@
 Table~\ref{tab:result_weighted_avg} summarizes the results of all models for segmentation of e-Coaching emails when word embeddings or lexical features are used in combination with punctuation and POS features.
-Similar to results in Tables~\ref{tab:result_base}, CRNN demonstrates the highest performance among all methods achieving 0.864 recall with 0.825 F1-Measure for ``new segment'' class and 0.990 precision with 0.989 recall and F1-Measure for overall.
+Similar to results in Tables~\ref{tab:result_base}, CRNN demonstrates the highest performance among all methods achieving 0.864 recall with 0.825 F1-Measure for ``new segment'' class and 0.990 precision with 0.989 recall and 0.989 F1-Measure for overall.
 BRNN and CRF show the lowest and second highest performance, respectively, for email segmentation among all methods.
@@
 \centering
-\caption{\textbf{AUC values of all classifiers demonstrating the impact of word embeddings, punctuation and POS features on email segmentation.
+\caption{\textbf{AUC values of all classifiers demonstrating the impact of word embeddings, punctuation and POS features on e-Coaching email segmentation.
 Highest AUC value for each feature set across all models is highlighted in boldface.}}
@@
 Table~\ref{tab:result_roc} shows the impact of individual features as well as their combination on the segmentation task.
-Although CRF provides the second highest classification results, it shows best AUC values, achieving 0.966 AUC when only lexical features are used and 0.994 AUC when a combination of lexical, punctuation and POS features are used.
+Although CRF provides the second highest classification results, it shows best AUC values, achieving AUC 0.966 when only lexical features are used and AUC 0.994 when a combination of lexical, punctuation and POS features are used.
 Influence of the punctuation and POS features is also consistent in AUC values, which increase by 2.9\%, 4.31\%, 6.63\% and 2.18\% for CRF, MLP, BRNN and CRNN methods, respectively.
@@
 
-Punctuation mark and POS features made a significant improvement in the performance of machine learning and all deep learning methods.
+Punctuation mark and POS features made a significant improvement in the performance of traditional machine learning and deep learning methods.
 Specially, punctuation features have the higher individual impact on model performance compared to POS features.
@@
 These high-level features are then utilized by the bidirectional RNN of CRNN model, which provide an improved performance on email segmentation.
-On the other hand, traditional BRNN model received words as input features and their word embeddings are directly utilized in the input layer.
+On the other hand, traditional BRNN model directly utilized word embeddings at the bidirectional RNN layer.
 
@@
 Similarly, additional information is a common cause for misclassification of an email segment into multiple segments.
-For instance, the first sentence of the above email segment represents a positive commitment to adolescent's behavior change then next two sentences provide an additional information to support their commitment.
+For example, the first sentence of the above email segment represents a positive commitment to adolescent's behavior change then next two sentences provide an additional information to support their commitment.
 

### 2018-08-07 plots/length-mlp.r
update figure 3

@@ -12,7 +12,7 @@ wd = '/home/mehedi/teana/projects/eCoaching-Text-Segmentation/plots/'
 data <- read.csv(paste(wd,"parameters/length.csv", sep=''))
 
 # plot graph
-plot(range(1,7), range(0.605, 0.765), type="n", ylab = "F1-Measure", xlab = "# of next or prior words/punctuation marks in a sample (l)")
+plot(range(1,7), range(0.605, 0.765), type="n", ylab = "F1-Measure", xlab = "# of next or prior words/punctuation marks per instance (l)")
 lines(data$n, data$fmeasure, lwd=1.5, lty=1, type='b', col='blue', pch=18)
 points(x = 2, y = 0.760, col = "red", pch = 1, cex = 2)
 textxy(data$n, data$fmeasure, labs=data$fmeasure, pos=3,  cex = 0.7)

### 2018-08-07 paper-amia-summit-2018/manuscript.tex
update introduction

@@
 \item Emails are unstructured text that contains informal information exchange in a non-traditional format.
-For example, an e-Coach usually responds to several queries of a patient in one email.
-However, a provider utterance of MI transcript denotes a particular response for a query of a patient during a traditional MI interview session, which involve a dialogue between a provider and a patient.
+For example, an e-Coach usually responds to a patient with their various concerns in one email.
+On the other hand, a traditional communication exchange involves a dialogue between a provider and a patient, where a provider addresses a particular concern of a patient in one utterance during the MI interview session.
 \item Discourse segments in e-Coaching emails do not necessarily correspond to sentences or collection of sentences.

### 2018-08-10 paper-amia-summit-2018/manuscript.tex
submitted to amia summit 2019

--- 2018-08-07 paper-amia-summit-2018/manuscript.tex
@@
 
-\title{Machine Learning Methods for Discourse Segmentation of Communications in E-Mail Based Behavioral Interventions}
+\title{Deep Neural Architectures for Discourse Segmentation in E-Mail Based Behavioral Interventions}
 
@@
 
-Qualitative coding of motivational interviews with pre-defined codes has been traditionally performed manually by trained annotators, which is a tedious and resource-intensive process that involves several iterations of reading, comprehension and interpretation of interview transcripts.
+Qualitative analysis of motivational interviews has been traditionally performed manually by MI researchers, which is a tedious and resource-intensive process involving several iterations of reading, comprehension and interpretation of interview transcripts.
 Rapidly developing computational technologies, specifically, machine learning methods, offer a unique opportunity to accelerate this process.
-In particular, machine learning methods have been successfully applied to a variety of analytical tasks involving textual data, such as classification\cite{nigam2000text} and sentiment analysis.\cite{wang2012baselines} In our previous work, we examined the utility of machine learning methods for automated annotation \cite{hasan2016study,kotov2015interpretable} and analysis \cite{hasan2018predicting} of in-person MI sessions.
-Specifically, we demonstrated that machine learning methods can be utilized for annotation of MI transcripts according to a simple communication code scheme with the accuracy comparable to human coders.\cite{hasan2016study} Experimental data utilized in these studies, however, were prepared by transcribing audio conversations, which were clearly segmented into utterances by a counselor, a patient and a caregiver.
+In particular, machine learning methods have been successfully applied to a variety of analytical tasks involving textual data, such as classification\cite{nigam2000text} and sentiment analysis.\cite{wang2012baselines} In our previous work, we examined the utility of machine learning methods for automated annotation \cite{hasan2016study,kotov2015interpretable} and predicting the outcome \cite{hasan2018predicting} of in-person MI sessions.
+Experimental data utilized in these studies, however, were transcribed auto recordings of in-person MI sessions with a counselor, which have a clear discourse structure and were clearly segmented into the utterances by a counselor, a patient or a caregiver.
 
-In this study, we focus on the analysis of e-Coaching sessions, behavior interventions that are delivered via email and grounded in the principles of motivational interviewing.
+In this study, we focus on the analysis of e-Coaching sessions, behavioral interventions that are delivered via email and are grounded in the principles of MI.
 Specifically, the e-Coaches involved in this study used emails to communicate motivation-enhancing messages that encourage healthy eating among GenY adolescents.
@@
 Usually, email segments are automatically coded with the Minority Youth-Sequential Coding of Process Exchanges (MYSCOPE), \cite{carcone2013provider} a qualitative coding scheme to characterize patient-provider communication during MI session.
-A significant barrier to fully automating the behavior coding process of e-Coaching emails is their segmentation into textual fragments that correspond to distinct communication behaviors.
-Automating this task is a unique and challenging problem due to the following major reasons:
+A significant barrier to qualitative analysis of e-Coaching exchanges is their segmentation into textual fragments that correspond to a distinct e-Coach and patient communication behaviors.
+Automating this task is a unique and challenging problem due to the following reasons:
 
 \begin{enumerate}
-\item Emails are unstructured text that contains informal information exchange in a non-traditional format.
-For example, an e-Coach usually responds to several queries of a patient in one email.
-However, a provider utterance of MI transcript denotes a particular response for a query of a patient during a traditional MI interview session, which involve a dialogue between a provider and a patient.
-\item Discourse segments in e-Coaching emails do not necessarily correspond to sentences or collection of sentences.
-One sentence can be segmented into multiple MI behavior fragments.
+\item Emails are unstructured text containing informal information exchange in a non-traditional format.
+For example, an e-Coach usually responds to several previous patient statements in one email.
+However, in a traditional MI interview session involving a dialogue between a provider and a patient, a provider utterance is a response to an immediately preceding inquiry or statement by a patient.
+\item Discourse segments in e-Coaching emails do not necessarily correspond to sentences or paragraphs.
+One sentence can be divided into fragments corresponding to multiple MI behaviors.
 On the other hand, an MI behavior may comprise several sentences.
@@
 
-Recently, an online clinical intervention called MENU GenY (Making Effective Nutrition Choices for Generation Y) was proposed and evaluated.
-\cite{alexander2017motivations} MENU GenY is a technology-based public health intervention that relies on personalized e-coaching to encourage increased fruit and vegetable intake among young adults, aged 21-30.
-The goal of MENU GenY was to develop a better coding dictionary among GenY to improve eating habits.
-However, segmentation of clinical exchange in the context of electronically delivered interventions, in particular, segmentation of clinical interaction text into groups of MI behaviors, is still performed manually, which slows down qualitative analysis of these interventions.
-This study proposes several machine learning methods to address this problem and the authors are unaware of any other work that focused on the same problem.
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=0.7\textwidth]{figures/segment-example.png}
+    \caption{\textbf{Example of e-Coaching emails segmented into fragments that correspond to MI behaviors of an e-Coach and a patient}}
+    \label{fig:text-segment}
+\end{figure}
 
-The goal of this research study is to assess the applicability of machine learning methods for automated segmentation of e-Coaching emails into textual fragments corresponding to individual behaviors, which is the first step of the coding process of e-Coaching communications.
-In particular, we introduced word embeddings or lexical, punctuation and  part-of-speech (POS) features and experimented with both traditional supervised machine learning method, linear-chain conditional random fields (CRF)\cite{lafferty2001conditional} and deep learning methods, such as multilayer perceptrons (MLP),\cite{rumelhart1986learning} bidirectional recurrent neural networks (BRNN)\cite{schuster1997bidirectional} and convolutional recurrent neural networks (CRNN),\cite{treviso2017sentence} to find the best performing method and feature combination.
+Figure~\ref{fig:text-segment} illustrates a segmentation of an e-Coaching email exchange, in which the first sentence is segmented into 2 MI behavior fragments, while the fourth and fifth segments correspond to one and three sentences, respectively.
+Segmentation of e-Coaching email exchanges between patients and e-Coaches corresponds to a special case of discourse analysis \cite{webber2012discourse} aimed at better understanding the effective communication strategies specific to this type of behavioral interventions and revealing the unique socio-psychological characteristics of patients.
 
-Relevant previous work in the biomedical domain primarily focused on segmentation of text into sections and headers of electronic health records (EHR)\cite{apostolova2009automatic,denny2009evaluation,tepper2012statistical,cho2002text} or sentence boundary detection of EHR.\cite{griffis2016quantitative,kreuzthaler2015detection,treviso2017sentence} Apostolova et al.
-applied SVM along with word-vector cosine similarity metric combined with several heuristics to segment clinical reports in EHR into sections, such as demographics, history, procedure, finding and impression.\cite{apostolova2009automatic} After identification of each line in the document, Tepper et al.
-trained maximum entropy models to classify the document into pre-defined sections, such as general patient information, provider information, medical condition, medical history, clinical history, findings, impression, etc.\cite{tepper2012statistical} Denny et al.
-proposed a SecTag algorithm, which combined natural language processing techniques, terminology-based rules and a Naive Bayes classifier to identify the sections and headers that achieved 99\% recall with 95.6\% precision.
-\cite{denny2009evaluation} On the other hand, SVM based on prosodic and part of speech features \cite{kreuzthaler2015detection} and recurrent convolutional neural networks using word embeddings \cite{griffis2016quantitative} were utilized for detecting sentence boundaries in EHR.
+The goal of this study is to assess the effectiveness of deep learning methods for the task of automated segmentation of e-Coaching emails into textual fragments corresponding to individual behaviors, which is the first step of qualitative analysis of this type of clinical communications.
+Specifically, for this study, we utilized the data from MENU GenY (Making Effective Nutrition Choices for Generation Y) \cite{alexander2017motivations}, a recent technology-based public health intervention that relies on personalized e-Coaching to encourage increased fruit and vegetable intake among young adults, aged 21--30.
+The goal of MENU GenY was to develop a better coding scheme for e-Coaching communications aimed at improving GenY eating habits.
+Segmentation of clinical conversation in the context of electronically delivered interventions into groups of MI behaviors, is traditionally performed manually by MI researchers, which significantly slows down their qualitative analysis.
+This paper is the first work to evaluate the empirical effectiveness of deep learning architectures in addressing \textit{the problem of discourse segmentation in the context of e-mail based behavioral interventions}.
+
+Specifically, we introduce and evaluate the effectiveness of word embedding or lexical, punctuation and part-of-speech (POS) features in conjunction with both traditional supervised machine learning methods, such as linear-chain Conditional Random Field (CRF)\cite{lafferty2001conditional} and deep learning architectures, such as Multi-Layer Perceptron (MLP),\cite{rumelhart1986learning} Bidirectional Recurrent Neural Network (BRNN)\cite{schuster1997bidirectional} and Convolutional Recurrent Neural Network (CRNN),\cite{treviso2017sentence} to find the best performing method and feature combination.
+
+\section*{Relevant work}
+Previous relevant work in the biomedical domain primarily focused on segmentation of text in electronic health records (EHR) into sections and headers \cite{apostolova2009automatic,denny2009evaluation,tepper2012statistical,cho2002text} or sentence boundary detection.\cite{griffis2016quantitative,kreuzthaler2015detection,treviso2017sentence} In particular, Maximum Entropy models \cite{tepper2012statistical} and Support Vector Machine (SVM) along with word-vector cosine similarity metric and several heuristics \cite{apostolova2009automatic} have been applied to segment or classify clinical documents in EHR into pre-defined sections, such as general patient information, medical history, procedures, findings, etc.
+Denny et al.
+\cite{denny2009evaluation} proposed a SecTag algorithm, which combined natural language processing techniques, terminology-based rules and a Naive Bayes classifier to identify sections and headers in EHR.
+Segmentation of e-Coaching emails, however, is different from segmentation of other clinical documents in that the focus is on clinical conversation.
+
+SVM in conjunction with prosodic and part of speech features \cite{kreuzthaler2015detection} and Recurrent Convolutional Neural Networks \cite{griffis2016quantitative} have also been utilized for \textit{sentence boundary detection} in general text.
 Liu et al.
-have shown that a linear chain CRF performed better than the hidden markov and maximum entropy model in sentence boundary detection, which was evaluated using the official NIST evaluation tools.\cite{liu2005using} Recently, Treviso et al.
-proposed a recurrent convolutional neural network for automatic identification of sentence boundary in a neuropsychological text in the Portuguese language.\cite{treviso2017sentence} Their proposed model combined prosodic and part of speech features with pre-trained GloVe word embeddings,\cite{pennington2014glove} which achieved 0.74 F1-score in sentence boundary detection on the Brazilian constitution dataset.
+\cite{liu2005using} demonstrated that a linear-chain CRF outperforms Hidden Markov and Maximum Entropy models for this task.
+
 Segmentation of e-Coaching emails is different from a traditional shallow discourse analysis of conversations \cite{galley2003discourse} in that the focus is on segmentation, rather than on determining the types of transitions between the utterances or assigning utterances to speakers.
-Segmentation of e-Coaching emails corresponds to a special type of discourse analysis \cite{webber2012discourse} aimed at better understanding the effective e-Coaching communication strategies and revealing the unique socio-psychological characteristics of a patient.
-
-Our proposed approach is novel for the segmentation of e-Coaching emails since previous studies mainly focused on the segmentation of EHR documents into sections, headers and sentences.
-However, this study segmented clinical exchange into groups of MI behaviors which will significantly reduce the amount of resource and time required to segment clinical exchange manually.
-Furthermore, these segmentation models could be integrated with auto coding classifiers to build a software pipeline for automated annotation of clinical exchanges.
+The proposed methods will automate the process of segmenting clinical exchanges into groups of MI behaviors, which will significantly reduce the resources and time required to perform this task manually.
+Furthermore, these methods can be integrated with auto coding methods \cite{hasan2016study,kotov2015interpretable} to create a software pipeline for automated analysis of clinical conversation.
 
@@
 \subsection*{\textit{Data collection}}
-
 The experimental dataset for this work was constructed from 49 e-Coaching sessions, which include a total of 3,138 segmented and annotated MI behaviors.
 Each session represents an MI intervention delivered via email.
-We formulate the segmentation task as a binary classification problem where each word or punctuation mark annotated with a label 1 or 0, to indicate whether it precedes a new segment or not.
-In total, we obtained 95,777 words and 7,140 punctuation marks, which include 3,138 ``new segment'' and 99,779 ``same segment'' instances.
-In this study, we experimented with traditional machine learning method, conditional random fields (CRF)\cite{lafferty2001conditional} and deep learning methods, such as multilayer perceptrons (MLP),\cite{rumelhart1986learning} bidirectional recurrent neural networks (bidirectional RNN or simply BRNN)\cite{schuster1997bidirectional} and convolutional recurrent neural networks (CRNN).\cite{treviso2017sentence} For MLP models, instances were created with $2l$ words or punctuations at each step in a word sequence, where an instance contains next $l$ words or punctuations and prior $l$ words or punctuations including current word or punctuation.
-Each instance is classified into either ``new segment'' or ``same segment'' based on whether the $l^{th}$ word or punctuation in the gold standard precedes a new segment or not.
-For CRF, BRNN and CRNN models, an email was taken as input sequence, such that POS tags and word embeddings of each word or punctuation mark were used as input and binary labels (1 or 0) corresponding to ``new segment'' and ``same segment'' classification decision were considered as the output of the model at each step.
+We consider segmentation as a binary classification problem, in which each word or punctuation mark is annotated with a label of 1 or 0 to indicate whether it precedes a new segment or not.
+In total, the dataset consists of 95,777 words and 7,140 punctuation marks and includes 3,138 ``new segment'' and 99,779 ``same segment'' instances.
+In this study, we experimented with traditional machine learning methods, such as Conditional Random Field (CRF)\cite{lafferty2001conditional} and deep learning methods, such as Multi-Layer Perceptron (MLP),\cite{rumelhart1986learning} Bidirectional Recurrent Neural Network (bidirectional RNN or BRNN)\cite{schuster1997bidirectional} and Convolutional Recurrent Neural Network (CRNN).\cite{treviso2017sentence} For an MLP model, training and testing instances were created based on a sliding window of $2n$ words or punctuation marks over each position in a given input sequence, such that each instance consists of a pair of the next $n$ words or punctuation marks and prior $n$ words or punctuation marks, including the current word or punctuation mark.
+Each sample is classified into either a ``new segment'' or a ``same segment'' based on whether there should be a segment break after the current word or not.
+In the case of CRF, BRNN and CRNN models, an email was taken as an input sequence, POS tags and word embeddings of each word or punctuation mark were used as input and binary labels (1 or 0) corresponding to ``new segment'' and ``same segment'' classification decisions were considered as the model output of a model.
 In the gold standard, words or punctuations within the same segment were assigned the label of 0 and the last word or punctuation mark of a segment were assigned the label of 1.
@@
 We utilized three types of features in conjunction with CRF, MLP, BRNN and CRNN methods: word embeddings or lexical features, punctuation and POS features.
-Since POS tags have been shown to be effective semantic abstractions of individual words, we used POS features for our experiment.\cite{liu2005using,treviso2017sentence} To extract our POS features, we tag the e-Coaching emails using the NLTK POS tagger.
-Punctuation marks, which correspond to one of the symbols \{`.', `,', `!', `?', `:', `;'\} between a pair of words, are also employed as a feature since punctuation marks designate the boundary of a sentence, clause and phrase and often also correspond to a segment boundary.\cite{cho2002text} For natural language processing (NLP) tasks, inputs are received as a text where words can be considered as the basic processing unit.
+Since POS tags have been shown to be effective semantic abstractions of individual words, we used POS features for our experiment.\cite{liu2005using,treviso2017sentence} To extract POS features, we pre-processed e-Coaching emails using the NLTK POS tagger.
+Punctuation marks, which correspond to one of the symbols \{`.', `,', `!', `?', `:', `;'\} between a pair of words, were also used as a feature, since punctuation marks designate the boundary of a sentence, clause and phrase and often also correspond to a segment boundary.\cite{cho2002text} For natural language processing (NLP) tasks, inputs are received as a text, in which individual words are as the basic units of semantics.
 Therefore, it is important to represent a word in such a way that it carries all relevant information.
-Word embeddings are one of this representation where each word represented as a real-valued vector in a high dimensional vector space.
-The distributed representation of word allows capturing semantic, syntactic and morphological information from large unannotated corpora.\cite{pennington2014glove, mikolov2013distributed} This study utilized embeddings estimated on Google News corpus consisting of 1.6 billion words referred to pre-trained word2vec word embeddings.\footnote{\label{fn:word2vec}https://code.google.com/p/word2vec/} When words or punctuation marks are not found in the pre-trained word vectors, word embeddings trained with e-Coaching email corpus referred to trained word embeddings were utilized.
-CRF utilized lexical features which represent each word in an e-Coaching email.
+Word embedding is one such representation, when each word is associated with a real-valued vector in a high dimensional vector space.
+Distributed representations of words have been shown to capture semantic, syntactic and morphological properties of words.\cite{pennington2014glove, mikolov2013distributed} For experiments reported in this paper, we utilized embeddings estimated on Google News corpus consisting of 1.6 billion words pre-trained using word2vec word embedding method.\footnote{\label{fn:word2vec}https://code.google.com/p/word2vec/} When words or punctuation marks are not found in the pre-trained word vectors, we utilized word embeddings trained with our e-Coaching email corpus, which we refer to as corpus-based embeddings.
+CRF utilized lexical features, which correspond to words in e-Coaching emails.
 
@@
 
-In the present study, we experimented with four different classifiers, including one traditional machine learning method, CRF and three deep learning models.
-Since deep learning methods provide a flexible architecture for constructing and synthesizing complex models, we take the advantage of this flexibility to test MLP, BRNN and CRNN models for the task of segmentation of e-Coaching emails.
+We experimented with 4 different classifiers, including one traditional machine learning method, CRF and three deep learning models.
+Since deep learning architectures provide a flexible mechanism for constructing complex models, we take the advantage of this flexibility to test MLP, BRNN and CRNN models for the task of segmentation of e-Coaching emails.
 
-\textbf{Conditional Random Fields}: CRF has been widely used in various NLP tasks including part-of-speech tagging and segmentation tasks.\cite{lafferty2001conditional, hirohata2008identifying} Unlike a maximum entropy markov model which uses per-state exponential models for the conditional probabilities of next states given the current state, a CRF is modeled with an exponential distribution provides the conditional distribution of the entire output sequence given the observation sequence.
-A traditional linear-chain CRF model is defined as a conditional probability distribution $p(y|x)$ for output and input sequences, $y$ and $x$, by the Eq.
-\ref{Eq:crf}
+\textbf{Conditional Random Field}: CRF has been widely used in various NLP tasks, such as part-of-speech tagging.\cite{lafferty2001conditional, hirohata2008identifying} Unlike a maximum entropy Markov model, which uses per-state exponential models for conditional probability of the next state given a current state, CRF model directly estimates conditional distribution of the entire output sequence, given the observation sequence.
+A traditional linear-chain CRF model is defined as a conditional probability distribution $p(y|x)$ for output and input sequences, $y$ and $x$:
 
@@
 
-Here $Z_x$ is a normalization factor over all possible labelings of $x$, and $f_k(y_{t-1}, y_t, x, t)$ is a feature function, and $\lambda_k$ is a learned weight associated with feature $f_k$.
-The optimal output sequence $y^*$ for an input sequence $x$, $y^* = {arg\,max}_y p(y|x)$, is obtained efficiently by the Viterbi algorithm.
-In our experiments, the following features were utilized with CRF models: i) current word or punctuation ii) next and previous 3 words or punctuations iii) binary feature indicating the word or punctuation is a special character (';', '?', '.', ',', '!', ':', etc) or not iv) binary feature indicating the word is a title word or not?
-(e.g.
+where $Z_x$ is a normalization factor, $f_k(y_{t-1}, y_t, x, t)$ is a feature function, and $\lambda_k$ is a learned weight associated with feature $f_k$.
+The optimal output sequence $y^*$ for an input sequence $x$, $y^* = {arg\,max}_y p(y|x)$, is obtained efficiently using the Viterbi algorithm.
+In our experiments, the following features were utilized in conjunction with CRF models: i) current word or punctuation ii) next and previous 3 words or punctuations iii) binary feature indicating whether a word or punctuation is a special character (';', '?', '.', ',', '!', ':', etc.) or not iv) binary feature indicating whether a word is a title word or not (e.g.
 ``The'' is a title word but ``the'' is not) v) POS tags
-
-\textbf{Multilayer Perceptrons}: MLP is a feed-forward artificial neural network which maps set of input onto a set of appropriate output.\cite{rumelhart1986learning} Figure~\ref{fig:mlp} shows a multilayer perceptron with a single hidden layer.
-In MLP networks, there is no cycle or loop and information moves forward only, from the input nodes through the hidden nodes and to the output nodes.
-This study utilized MLP with a nonlinear activation function ($relu$) and one hidden layer consisting of 128 hidden units.
-In order to prevent over-fitting, we implement dropout on the last layer, which randomly hides some neurons during the training phase.
-\cite{srivastava2014dropout} Dropout is also applied to the fully connected layers in BRNN and CRNN models.
 
@@
 
-\textbf{Bidirectional Recurrent Neural Networks}: BRNN is a neural network designed to capture sequential patterns by considering both past \& future inputs as well as complex relationships between input features and output labels.\cite{schuster1997bidirectional} The output of BRNN layer is computed with the summation of the forward RNN output and backward RNN output.
-Gated Recurrent Units (GRU)\cite{chung2014empirical} are utilized as RNNs that are capable of handling variable size input sequence and have an internal memory that can be reset.
-Figure~\ref{fig:crnn} will represent an architecture of BRNN if convolution layer is removed.
+\textbf{Multi-Layer Perceptron}: MLP is a feed-forward artificial neural network, which maps an input onto one or several outputs.\cite{rumelhart1986learning} Figure~\ref{fig:mlp} shows a multilayer perceptron with a single hidden layer.
+In MLP networks, there is no cycle or loop and information moves forward only, from the input nodes through the hidden nodes and to the output nodes.
+This study utilized MLP with a nonlinear activation function (rectified linear unit) and one hidden layer consisting of 128 hidden units.
+In order to prevent over-fitting, we utilized dropout in fully connected layers by randomly hiding neurons during training.
+\cite{srivastava2014dropout} Dropout is also applied to the fully connected layers in BRNN and CRNN models.
 
-\textbf{Convolutional Recurrent Neural Networks}: CRNN is a deep neural networks architecture,\cite{treviso2017sentence} shown in Figure~\ref{fig:crnn}, consisting of 5 layers: 1) input layer 2) embedding layer 3) convolution layer with max pooling 4) BRNN layer 5) fully connected layer with dropout and sigmoid output.
-E-coaching email exchanges are represented as a sequence of $m$ words and punctuations, which are fed into the input layer.
-The embedding layer provides a distributed representation of words, gives us $m \times n_e$ matrix after fetching the pre-trained word vectors.
-When POS tags are used in combination with word embeddings, 10-dimensional POS vectors are concatenated with 300-dimensional word vectors to obtain new vectors $n_e = n_w + n_p$ of size 310.
-The convolution layer used this word embeddings to extract new features depending on neighbor words.
-This study used 1D convolution which simply moves one step at a time on vertically.
-In this layer, one filter is responsible for the extraction of one feature.
-After applying $n_f$ different filters with zero-padding on both sides of the text, $n_f$ useful features are produced in the convolution layer for each word.
-A max pooling operation was then performed over time to find the most significant features in a text region.
+\textbf{Bidirectional Recurrent Neural Network}: BRNN is a neural network designed to capture sequential patterns by considering both past \& future inputs as well as complex relationships between input features and output labels.\cite{schuster1997bidirectional} The output of BRNN layer is computed as an aggregation of outputs of the forward and backward RNNs.
+Gated Recurrent Units (GRU)\cite{chung2014empirical} capable of handling variable size input sequence and having internal memory, which can be reset, were utilized as an RNN.
+Figure~\ref{fig:crnn} represents the architecture of BRNN, in the case when a convolution layer is removed.
+
+\textbf{Convolutional Recurrent Neural Network}: CRNN is a deep neural network architecture,\cite{treviso2017sentence} shown in Figure~\ref{fig:crnn}, which consists of 5 layers: 1) input layer 2) embedding layer 3) convolution layer with max pooling 4) BRNN layer 5) fully connected layer with dropout and sigmoid output.
+E-coaching email exchanges are represented as a sequence of $m$ words and punctuations, which are fed into the input layer to produce a $m \times n_e$ matrix after fetching the pre-trained word vectors.
+When POS tags are used in combination with word embeddings, 10-dimensional POS vectors are concatenated with 300-dimensional word vectors to obtain new vectors $n_e = [n_w;n_p]$ of size 310.
+The primary purpose of convolution is to extract new features depending on the neighboring words.
+We used 1D convolution.
+In this layer, one filter is responsible for extraction of one feature.
+After applying $n_f$ different filters with zero-padding on both sides of the input text, $n_f$ useful features are produced in the convolution layer for each word.
+A max pooling operation was then performed over time to find the most significant features in a textual fragment.
 The bidirectional recurrent layer receives new features extracted in the convolution layer.
-RNN is usually used to capture the previous history.
-Moreover, bidirectional RNN is capable of capturing both previous and past histories because bidirectional RNN utilized the features by looking at forward and backward states.
-The fully connected layer used the bidirectional RNN output for classifying each word or punctuation into ``new segment'' and ``same segment'' classes.
-Since a fully connected layer has a greater number of parameters, they are more likely excessively co-adapt these parameters causing over-fitting.
-To prevent this over-fitting problem, dropout is used to randomly drop 50\% of the connections on the fully connected layer.
-After that, a sigmoid operation is performed on the output layer to provide the probability of whether the word or punctuation precedes a new segment or not.
-We experimentally determined the optimal parameters and found that the best performance is achieved with 5 folds cross-validation when filter length in convolution layer is 7, number of filters is 100, max-pool size is 3, convolutional layer activation is $relu$, RNN layer activation is $tanh$ and number of recurrent units in RNN is 200.
-Adam\cite{kingma2014adam} is used for optimization and the early stopping strategy is applied with 50 epochs when the batch size is 32 and the learning rate is 0.001.
-To utilize our source codes by other research studies, source codes of all experiments will be publicly available\footnote{https://github.com/teanalab/xxxxx} upon acceptance of this paper for publication.
+RNN is usually used to capture long-range dependencies in a sequence of observations.
+Moreover, bidirectional RNNs are capable of capturing both past and future contexts through forward and backward traversals of a sequence.
+The purpose of the fully connected layer is to use the output of bidirectional RNN layer for classifying each word or punctuation into ``new segment'' and ``same segment'' classes.
+Since a fully connected layer has a greater number of parameters, they are more likely to excessively co-adapt with these parameters, which may cause over-fitting.
+To prevent this, we utilized dropout by randomly disregarding 50\% of the connections to the fully connected layer.
+After that, logistic sigmoid is applied to the output layer to provide the probability of whether the word or punctuation precedes a new segment or not.
+We experimentally determined the optimal parameters and found that the best performance is achieved with 5-fold cross-validation when filter length in convolutional layer is 7, number of filters is 100, max-pool size is 3, ReLU is used as a convolutional layer activation function, RNN layer activation is hyperbolic tangent and the number of recurrent units in RNN is 200.
+Adam\cite{kingma2014adam} with 50 epochs, the batch size of 32 and learning rate of 0.001 was used for optimization and the early stopping strategy was applied.\footnote{source code will be available at https://github.com/teanalab/xxxxx upon acceptance of this paper for publication}
 
@@
     \centering
-    \includegraphics[width=1.0\textwidth]{figures/CRNN.eps}
+    \includegraphics[width=0.95\textwidth]{figures/CRNN.eps}
     \caption{\textbf{Architecture of convolutional recurrent neural networks for automatic segmentation of e-Coaching emails}}
@@
 \subsection*{\textit{Evaluation metrics}}
-We report standard metrics for experiments (precision, recall and F1-measure) to evaluate the performance of binary classifiers.\cite{aas1999text} However, accuracy is not reported as a performance metric because accuracy is highly sensitive to the prior class probabilities and does not fully describe the actual difficulty of the decision problem for an unbalanced dataset.
-The results are reported based on 5 folds cross-validation and weighted macro-averaging over the folds.
+We report standard metrics of precision, recall and F1-measure to evaluate the performance of the classifiers.\cite{aas1999text} Accuracy is not reported as a performance metric because it is highly sensitive to the prior class probabilities and does not fully describe the actual difficulty of the decision problem, when highly unbalanced datasets are involved.
+The results are reported based on 5 fold cross-validation and weighted macro-averaging over the folds.
 We also estimate the area under the receiving operating characteristics (ROC) curve\cite{kumar2011receiver} (AUC) metric due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets.
@@
 \section*{Results}
-Our experimental results have four important directions.
-First, the optimal word embedding dimension and instance size of MLP models are illustrated in Figure~\ref{fig:embedding-dimension-mlp}.
+Our experiments span four dimensions.
+First, we report the optimal size of word embeddings and sliding window size of the MLP model.
 Second, results are reported with respect to ``new segment'' class as well as weighted average over ``new segment'' and ``same segment'' classes in Table~\ref{tab:result_base}.
-Third, classification performance of different machine learning methods are outlined in Table~\ref{tab:result_weighted_avg} when word embeddings or lexical features are used in combination with punctuation and POS features.
-Fourth, the impact of individual features, as well as their combination on the task of email segmentation, are demonstrated in Table~\ref{tab:result_roc}.
+Third, classification performance of different machine learning methods are summarized in Table~\ref{tab:result_weighted_avg} when word embeddings or lexical features are used in combination with punctuation and POS features.
+Fourth, the impact of the individual features, as well as their combination on the segmentation task, is reported.
 
@@
     \centering
-    \includegraphics[width=1.0\textwidth]{figures/mlp-and-vector.eps}
-    \caption{\textbf{Performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and trained word embeddings with GloVe and word2vec models (left).
-Performance of MLP model on e-Coaching email segmentation by varying the length of instance (right).}}
+    \includegraphics[width=0.95\textwidth]{figures/mlp-and-vector.eps}
+    \caption{\textbf{Performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and corpus word embeddings with GloVe and word2vec models (left).
+MLP model on e-Coaching email segmentation by varying the sliding window size $2n$ (right).}}
     \label{fig:embedding-dimension-mlp}
@@
 
-Figure~\ref{fig:embedding-dimension-mlp} (left) shows the performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and trained word embeddings with GloVe\footnote{https://nlp.stanford.edu/projects/glove/} and word2vec models.
-It was observed that the best performance of email segmentation was achieved with pre-trained 300-dimensional word2vec word vectors when three types of features are used together.
-Therefore, results of all deep learning models are reported with pre-trained word2vec 300-dimensional word vectors.
-In MLP, inputs were prepared by computing the summation of first $l$ word vectors and last $l$ word vectors where each instance contains $2l$ words or punctuations.
-Figure~\ref{fig:embedding-dimension-mlp} (right) demonstrates the performance of MLP models on e-Coaching email segmentation by varying the value of $l$.
-It was observed that the best performance of MLP classifiers is achieved when $l = 2$.
-Therefore, results of all MLP models are reported with $l = 2$.
+Figure~\ref{fig:embedding-dimension-mlp} (left) shows the performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and corpus word embeddings with GloVe \footnote{https://nlp.stanford.edu/projects/glove/} and word2vec models.
+It was observed that best performance is achieved with pre-trained 300-dimensional word2vec word vectors when three types of features are used together.
+Therefore, we report our results with word2vec 300-dimensional word vectors for all deep learning models used in this study.
+For MLP models, the first layer input was prepared by the summation of the first $n$ word vectors in the sliding window and the last $n$ word vectors when a sliding window contains $2n$ words or punctuations.
+Figure~\ref{fig:embedding-dimension-mlp} (right) demonstrates the performance of the MLP model on e-Coaching email segmentation by varying the size of the sliding window.
+It can be observed that the best performance of MLP is achieved when $n$ is 2.
+Therefore, results of MLP in the remaining experiments are reported with $n$ set to 2.
 
@@
 On the other hand, MLP has the highest precision of 0.836 when word embeddings or lexical features are used to identify ``new segment''.
-CRF achieves 0.733 F1-Measure in new segment class and 0.984 F1-Measure in overall, which demonstrates the second highest performance to identify ``new segment'' as well as a weighted average over both classes.
-Experimental results indicate that performance of all classifiers as a weighted average over both classes is remarkably higher compared to ``new segment'' class, which is expected since 96.95\% instances belong to the ``same segment'' categories and 99.3\% of them correctly classified.
-For example, as a weighted average over ``new segment'' and ``same segment'' classes, CRNN achieves 27.23\%, 23.71\% and 25.61\% higher precision, recall and F1-measure, respectively, compared to ``new segment'' detection.
-\\
+CRF achieves 0.733 F1-Measure in new segment class and 0.984 F1-Measure overall, which corresponds to the second highest performance in identifying ``new segment'' as well as a weighted average over both classes.
+Experimental results indicate that performance of all classifiers as a weighted average over both classes is remarkably higher compared to ``new segment'' class, which is expected since 96.95\% of instances belong to the ``same segment'' class and 99.3\% of them are correctly classified.
+For example, as a weighted average over ``new segment'' and ``same segment'' classes, CRNN achieves 27.23\%, 23.71\% and 25.61\% higher precision, recall and F1-measure, respectively, compared to the ``new segment'' detection.
 
@@
 \end{table}
+
+Table~\ref{tab:result_weighted_avg} summarizes the results of all models for segmentation of e-Coaching emails when word embeddings or lexical features are used in combination with punctuation and POS features.
+Similar to results in Tables~\ref{tab:result_base}, CRNN demonstrates the highest performance among all methods achieving 0.864 recall with 0.825 F1-Measure for ``new segment'' detection and 0.990 precision with 0.989 recall and F1-Measure overall.
+BRNN and CRF show the lowest and second highest performance, respectively, for email segmentation among all methods.
+We observed that classification performance significantly improved for ``new segment'' class when word embeddings or lexical features are used in combination with punctuation and POS features.
+Precision increases by 3.96\%, -2.27\%, 12.71\% and 1.81\%; recall increases by 11.72\%, 19.73\%, 20.59\% and 8.41\%; and F1-measure increases by 8.05\%, 9.51\%, 16.22\% and 5.1\% for CRF, MLP, BRNN and CRNN methods, respectively, in new segment detection when all features are utilized together.
+Similarly, precision increases by 0.51\%, 0.41\%, 0.82\% and 0.41\%; recall increases by 0.41\%, 0.41\%, 0.72\% and 0.3\%; and F1-measure increases by 0.41\%, 0.41\%, 0.82\% and 0.3\% for CRF, MLP, BRNN and CRNN methods, respectively, in weighted average over ``new segment'' and ``same segment'' classes when word embeddings or lexical features are used in combination with punctuation and POS features.\\
 
@@
 
-Table~\ref{tab:result_weighted_avg} summarizes the results of all models for segmentation of e-Coaching emails when word embeddings or lexical features are used in combination with punctuation and POS features.
-Similar to results in Tables~\ref{tab:result_base}, CRNN demonstrates the highest performance among all methods achieving 0.864 recall with 0.825 F1-Measure for ``new segment'' class and 0.990 precision with 0.989 recall and 0.989 F1-Measure for overall.
-BRNN and CRF show the lowest and second highest performance, respectively, for email segmentation among all methods.
-We observed that classification performance significantly improved for ``new segment'' class when word embeddings or lexical features are used in combination with punctuation and POS features.
-Precision increases by 3.96\%, -2.27\%, 12.71\% and 1.81\%; recall increases by 11.72\%, 19.73\%, 20.59\% and 8.41\%; and F1-measure increases by 8.05\%, 9.51\%, 16.22\% and 5.1\% for CRF, MLP, BRNN and CRNN methods, respectively, in new segment detection when all features are utilized together.
-Similarly, precision increases by 0.51\%, 0.41\%, 0.82\% and 0.41\%; recall increases by 0.41\%, 0.41\%, 0.72\% and 0.3\%; and F1-measure increases by 0.41\%, 0.41\%, 0.82\% and 0.3\% for CRF, MLP, BRNN and CRNN methods, respectively, in weighted average over ``new segment'' and ``same segment'' classes when word embeddings or lexical features are used in combination with punctuation and POS features.\\
+Table~\ref{tab:result_roc} shows the impact of individual features as well as their combination on the segmentation task.
+Although CRF provides the second highest classification results, it shows best AUC values, achieving AUC 0.966 when only lexical features are used and AUC 0.994 when a combination of lexical, punctuation and POS features are used.
+Influence of the punctuation and POS features is also consistent in AUC values, which increase by 2.9\%, 4.31\%, 6.63\% and 2.18\% for CRF, MLP, BRNN and CRNN methods, respectively, when all features are used together.
+Individually, POS and punctuation features also improve the performance of all classifiers except CRNN when word embeddings are used with POS features.
+CRF and MLP achieve their highest AUCs when all features are used together.
+On the other hand, BRNN and CRNN demonstrate the highest AUCs when word embeddings are combined with punctuation features.
+\\
 
@@
 
-% Baseline ROC results: BRNN: 0.9198, MLP: 0.9507, CRF: 0.9662, CRNN: 0.9648
-% Proposed ROC Results: BRNN: 0.9807, MLP: 0.992, CRF: 0.9937, CRNN: 0.9859
-
-Table~\ref{tab:result_roc} shows the impact of individual features as well as their combination on the segmentation task.
-Although CRF provides the second highest classification results, it shows best AUC values, achieving AUC 0.966 when only lexical features are used and AUC 0.994 when a combination of lexical, punctuation and POS features are used.
-Influence of the punctuation and POS features is also consistent in AUC values, which increase by 2.9\%, 4.31\%, 6.63\% and 2.18\% for CRF, MLP, BRNN and CRNN methods, respectively.
-Individually, POS and punctuation features improve the performance of all classifiers except CRNN when word embeddings are used with POS features.
-CRF and MLP achieve their highest AUCs when all features are used together.
-On the other hand, BRNN and CRNN demonstrate the highest AUCs when word embeddings are combined with punctuation features.
-
 \section*{Discussion}
@@
 Experimental results indicate that CRNN is the best model among all machine learning methods considered for this study.
-CRNN achieved 0.989 F1-measure in overall and 0.825 F1-measure for detecting ``new segment''.
-The robust performance of CRNN provides the evidence that deep learning models are capable to learn the hidden characteristics of MI behaviors from the clinical exchange.
-It also indicates that punctuation and POS features are important with word embeddings or lexical features for all machine learning methods are employed.
-Although the domain of this study was intentionally quite small, we believe that our study is not limited to the e-Coaching domain and it can be successfully applied to other domains, in which discourse segmentation is a preliminary step for annotation.
+CRNN achieved 0.989 F1-measure overall and 0.825 F1-measure for detecting ``new segment''.
+The robust performance of CRNN provides the evidence that deep learning models are capable of learning the transitions between MI behaviors from clinical exchanges.
+It also indicates that punctuation and POS features are important along with word embeddings for all machine learning methods employed.
+Although the domain of this study was intentionally quite small, we believe that our study is not limited to the e-Coaching domain and our conclusions can be extended to other domains, which require discourse segmentation.
 
-Punctuation mark and POS features made a significant improvement in the performance of traditional machine learning and deep learning methods.
-Specially, punctuation features have the higher individual impact on model performance compared to POS features.
-In all cases, every method performs better, when word embeddings or lexical features are used in combination with punctuation and POS features.
+Punctuation mark and POS features resulted in significant improvement in the performance of machine learning and all deep learning methods.
+Especially, punctuation features have the higher individual impact on model performance compared to POS features.
+In almost all cases, every method performs better, when word embeddings are used in conjunction with punctuation and POS features.
 This results also indicate that segmentation performance might be improved by adding additional relevant features.
 
-The convolutional layer made a significant difference between the performance of CRNN and BRNN on email segmentation.
-CRNN demonstrates 22.5\% and 10.7\% higher F1-Measure in ``new segment'' detection and 1\% and 0.5\% higher F1-Measure in overall compared to BRNN when word embeddings and all features are used, respectively.
-In CRNN, convolution layer performed a series of convolutions and pooling operations during which a number of high-level important features are extracted from the word embeddings of e-Coaching emails.
-These high-level features are then utilized by the bidirectional RNN of CRNN model, which provide an improved performance on email segmentation.
-On the other hand, traditional BRNN model directly utilized word embeddings at the bidirectional RNN layer.
-
-\begin{figure}[!htb]
-    \centering
-    \includegraphics[width=1.0\textwidth]{figures/roc-curves.eps}
-    \caption{\textbf{Receiver operating characteristic curves showing the performance of binary classifiers for the segmentation of e-Coaching text when only word embeddings or lexical features (left) and combination of word embeddings or lexical, punctuation and POS features (right) are used}}
-    \label{fig:roc-curves}
-\end{figure}
-
-In this paper, results are reported by ``new segment'' class as well as weighted average over both classes to avoid confusion regarding overall model performance due to severe class imbalance.
-In addition, standard metrics: precision, recall and F1-measure were used to eliminate doubt about the model performance since accuracy is misleading for imbalanced datasets.
-AUC is also utilized due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets,\cite{hu2015kernelized} which are also demonstrated by the ROC curves in Figure~\ref{fig:roc-curves}.
+The convolutional layer made a significant difference between the performance of CRNN and BRNN in MI session discourse segmentation.
+CRNN results in 22.46\% and 10.74\% higher F1-Measure in ``new segment'' detection and 1.02\% and 0.51\% higher F1-Measure overall compared to BRNN, when word embeddings and all other features are used, respectively.
+In CRNN, a convolution layer performs a series of convolutions and pooling operations, which produce a number of high-level important features from word embeddings.
+These high-level features are then utilized by the bidirectional RNN of the CRNN model, which results in a significant increase in performance.
+On the other hand, traditional BRNN model received words as input features and their word embeddings are directly utilized in the input layer.
 
 Although punctuation mark plays an important role in segmentation boundary detection, a few errors were encountered by the presence of punctuation marks in boundary identification.
-For example, a text segment from an e-Coaching email \textit{``A typical day in regards to fruit and vegetable has me eating about a serving at breakfast (our caf has cut up fruit) and then maybe a piece of fruit later in the day or as a snack.
+For example, a text segment from an e-Coaching email \textit{``A typical day in regards to fruit and vegetable has me eating about a serving at breakfast (our cafe has cut up fruit) and then maybe a piece of fruit later in the day or as a snack.
 Vegetable tends to be a side serving at lunch and dinner and I get celery or carrot cuts with dressing for a snack a lot of times.
-I could probably add some sort of vegetable into my breakfast (like spinach in an omelet) and snack on another piece of fruit when I am hungry rather than the junk food I tend to eat.''} incorrectly segmented after the first sentence where a punctuation mark was encountered.
+I could probably add some sort of vegetable into my breakfast (like spinach in an omelet) and snack on another piece of fruit when I am hungry rather than the junk food I tend to eat.''} was incorrectly segmented after the first sentence, when a punctuation mark was encountered.
 Similarly, additional information is a common cause for misclassification of an email segment into multiple segments.
-For example, the first sentence of the above email segment represents a positive commitment to adolescent's behavior change then next two sentences provide an additional information to support their commitment.
+For instance, although the first sentence of the above email segment represents a positive commitment to behavior change, the next two sentences provide an additional information to support the patient's commitment.
 
@@
 Therefore, there is a need to replicate the experiments with different data sets.
-As our future work, we plan to evaluate our approach to other datasets for discourse analysis.
+As our future work, we plan to evaluate our approach on the datasets from other behavioral interventions.
 
 \section*{Conclusion}
-Segmentation of e-Coaching emails is an integral part of developing e-Coaching interventions.
-Although several studies have focused on clinical interventions, they are limited by the qualitative coding of clinical interactions.
-In addition, previous studies in the medical domain mainly segmented clinical text into sections and sentences, none of them considered segmentation of text into groups of MI behaviors in the setting of discourse analysis with emails.
-In this paper, we compared the performance of machine learning models for the task of segmentation of e-Coaching text.
-We found out that CRNN provides the best performance for the segmentation of text in terms of all performance metrics.
+Segmentation of e-Coaching emails is an integral part of developing and analyzing e-Coaching behavioral interventions.
+Although several studies have focused on segmentation problem in biomedical context, they are limited to segmenting clinical text in EHR into sections and sentences, with none of them considering segmentation of text into groups of MI behaviors in the context of MI session discourse analysis.
+By comparing the performance of machine learning methods for the task of segmentation of e-Coaching emails, we found out that CRNN provides the best performance in terms of all performance metrics.
 Manual segmentation of e-Coaching data is a very resource-intensive and time-consuming task, which can significantly decrease the time and effort required to develop an effective behavioral intervention.
-Our proposed methods can help to identify individual text segments, which can be annotated directly with a classification model.
-This approach will also help in developing fully automated e-Coaching and accelerate the pace of identifying effective communication strategies.
+Our proposed methods can help to identify textual segments corresponding to MI behaviors in unstructured clinical dialog, which can then be annotated with a classification model in a pipeline setting.
+Automated segmentation and analysis of e-Coaching emails can significantly decrease the time to identify effective communication strategies in e-mail based motivational interviewing.
 

### 2018-08-10 plots/embedding-dimension.r

@@ -24,7 +24,7 @@ points(x = 300, y = 0.816, col = "blueviolet", pch = 17, cex = 1, lwd=1)
 text(300, 0.816, labels = c('0.816'), pos = 1, cex = 0.7)
 textxy(data$dim, data$tf1, labs=data$tf1, pos=1,  cex = 0.7)
 
-legend(222, 0.807, c("trained GloVe", "pre-trained GloVe", "trained word2vec", "pre-trained word2vec"), 
+legend(211, 0.807, c("corpus-based GloVe", "pre-trained GloVe", "corpus-based word2vec", "pre-trained word2vec"), 
        lty=c(1,1,1), col=c('blue', 'green', 'blueviolet', 'red'), cex=1, lwd=1.5)
 
 head(data)

### 2018-08-10 plots/length-mlp.r

@@ -12,9 +12,9 @@ wd = '/home/mehedi/teana/projects/eCoaching-Text-Segmentation/plots/'
 data <- read.csv(paste(wd,"parameters/length.csv", sep=''))
 
 # plot graph
-plot(range(1,7), range(0.605, 0.765), type="n", ylab = "F1-Measure", xlab = "# of next or prior words/punctuation marks per instance (l)")
+plot(range(2,14), range(0.605, 0.765), type="n", ylab = "F1-Measure", xlab = "sliding window size (2n)")
 lines(data$n, data$fmeasure, lwd=1.5, lty=1, type='b', col='blue', pch=18)
-points(x = 2, y = 0.760, col = "red", pch = 1, cex = 2)
+points(x = 4, y = 0.760, col = "red", pch = 1, cex = 2)
 textxy(data$n, data$fmeasure, labs=data$fmeasure, pos=3,  cex = 0.7)
 
 head(data)

### 2018-11-13 bin/run-stem-lda.sh
clean project directory

@@ -1,19 +1,0 @@
-#!/bin/bash
-echo "Running LDA model..."
-echo "Splitting data..."
-./Splitter  ../config/stem/split.cfg
-echo "Training LDA model..."
-./LabLDA ../config/stem/lda-1.cfg
-./LabLDA ../config/stem/lda-2.cfg
-./LabLDA ../config/stem/lda-3.cfg
-./LabLDA ../config/stem/lda-4.cfg
-./LabLDA ../config/stem/lda-5.cfg
-echo "Classifying test data with trained LDA model..."
-./LDAClass ../config/stem/lda-cls-1.cfg
-./LDAClass ../config/stem/lda-cls-2.cfg
-./LDAClass ../config/stem/lda-cls-3.cfg
-./LDAClass ../config/stem/lda-cls-4.cfg
-./LDAClass ../config/stem/lda-cls-5.cfg
-echo "Evaluating LDA model..."
-./Eval ../config/stem/lda-eval.cfg
-echo "Finished execution of LDA model"

### 2018-11-13 bin/run-stem-lta.sh

@@ -1,19 +1,0 @@
-#!/bin/bash
-echo "Running LTA model..."
-echo "Splitting data..."
-./Splitter  ../config/stem/split.cfg
-echo "Training LTA model..."
-./LTA ../config/stem/lta-1.cfg
-./LTA ../config/stem/lta-2.cfg
-./LTA ../config/stem/lta-3.cfg
-./LTA ../config/stem/lta-4.cfg
-./LTA ../config/stem/lta-5.cfg
-echo "Classifying test data with trained LTA model..."
-./LTAClass ../config/stem/lta-cls-1.cfg
-./LTAClass ../config/stem/lta-cls-2.cfg
-./LTAClass ../config/stem/lta-cls-3.cfg
-./LTAClass ../config/stem/lta-cls-4.cfg
-./LTAClass ../config/stem/lta-cls-5.cfg
-echo "Evaluating LTA model..."
-./Eval ../config/stem/lta-eval.cfg
-echo "Finished execution of LTA model"

### 2018-11-13 bin/run-stem-mglda.sh

@@ -1,19 +1,0 @@
-#!/bin/bash
-echo "Running MGLabLDA model..."
-echo "Splitting data..."
-./Splitter  ../config/stem/split.cfg
-echo "Training MGLabLDA model..."
-./MGLabLDA ../config/stem/mg-lda-1.cfg
-./MGLabLDA ../config/stem/mg-lda-2.cfg
-./MGLabLDA ../config/stem/mg-lda-3.cfg
-./MGLabLDA ../config/stem/mg-lda-4.cfg
-./MGLabLDA ../config/stem/mg-lda-5.cfg
-echo "Classifying test data with trained MGLDA model..."
-./LDAClass ../config/stem/mg-lda-cls-1.cfg
-./LDAClass ../config/stem/mg-lda-cls-2.cfg
-./LDAClass ../config/stem/mg-lda-cls-3.cfg
-./LDAClass ../config/stem/mg-lda-cls-4.cfg
-./LDAClass ../config/stem/mg-lda-cls-5.cfg
-echo "Evaluating MGLabLDA model..."
-./Eval ../config/stem/mg-lda-eval.cfg
-echo "Finished execution of MGLabLDA model"

### 2018-11-13 results/roc/roc-curves-generator.R

@@ -1,45 +1,0 @@
-
-#library
-library(ggplot2)
-library(reshape2)
-library(pROC)
-
-par(mfrow=c(1,2))
-
-# melt data
-r1 <- read.csv("/home/mehedi/teana/projects/eCoaching-Text-Segmentation/results/roc/False/KNN3False-ROC.csv")
-r2 <- read.csv("/home/mehedi/teana/projects/eCoaching-Text-Segmentation/results/roc/False/SVMFalse-ROC.csv")
-r3 <- read.csv("/home/mehedi/teana/projects/eCoaching-Text-Segmentation/results/roc/False/NBMFalse-ROC.csv")
-
-# plot graph
-plot(r1$fpr, r1$tpr, col='red', type="l", ylab = "Sensitivity", xlab = "1-Specificity", cex.axis=1, cex.lab=1, lwd=1)
-lines(r2$fpr, r2$tpr, col='blue', lty=1, lwd=1)
-lines(r3$fpr, r3$tpr, col='green', lty=1, lwd=1)
-lines(0:1, 0:1, col='black', lty=5, lwd=1)
-legend(0.77,0.30, c("KNN", "SVM", "NB"), 
-       lty=c(1,1,1), col=c('red', 'blue', 'green'), cex=1, lwd=1)
-
-# print AUC values
-x1 <- read.csv("/home/mehedi/teana/projects/eCoaching-Text-Segmentation/results/roc/True/roc_LSTM_summit.csv")
-x2 <- read.csv("/home/mehedi/teana/projects/eCoaching-Text-Segmentation/results/roc/True/roc_GRU_summit.csv")
-roc_obj_lstm <- roc(x1$actual, x1$predict_prob)
-auc(roc_obj_lstm)
-
-roc_obj_gru <- roc(x2$actual, x2$predict_prob)
-auc(roc_obj_gru)
-
-# melt data
-r1 <- read.csv("/home/mehedi/teana/projects/eCoaching-Text-Segmentation/results/roc/True/KNN3True-ROC.csv")
-r2 <- read.csv("/home/mehedi/teana/projects/eCoaching-Text-Segmentation/results/roc/True/SVMTrue-ROC.csv")
-r3 <- read.csv("/home/mehedi/teana/projects/eCoaching-Text-Segmentation/results/roc/True/NBMTrue-ROC.csv")
-
-# plot graph
[…]

### 2018-11-13 rnn/seg-gru-model-stepwise.py

@@ -1,116 +1,0 @@
-import numpy
-
-from keras.models import Sequential
-from keras.layers import GRU, LSTM
-from keras.layers import Dense
-from keras.layers import TimeDistributed
-from keras.callbacks import Callback
-
-import utility_stepwise_gru
-
-from vocabulary import Vocabulary
-
-# system configuration
-max_len = 50
-folds = 5
-lstm_out = 25
-batch_size = 1
-
-
-class ResetStatesCallback(Callback):
-    def __init__(self):
-        self.counter = 0
-
-    def on_batch_begin(self, batch, logs={}):
-        if self.counter % max_len == 0:
-            self.model.reset_states()
-            # print("Reset States..............")
-        self.counter += 1
-
-
-# create e-coaching dictionary
-vocabulary_obj = Vocabulary()
-vocabulary_obj.create()
-
-# create mapping of word to integers and the reverse
-char_to_int = dict((c, i+1) for i, c in enumerate(vocabulary_obj.vocab))
-int_to_char = dict((i+1, c) for i, c in enumerate(vocabulary_obj.vocab))
-
-# prepare the data set of input to output pairs encoded as integers
-data = utility_stepwise_gru.get_ecoaching_data()
-folds_data = utility_stepwise_gru.split_data(data, folds)
-results = []
-
-for fold_num in range(folds):
-    train_data, test_data = [], []
-    for i in range(folds):
-        if i == fold_num:
-            test_data = folds_data[i]
-        else:
-            train_data = train_data + folds_data[i]
-
-    # get encoded training data
-    X, y = utility_stepwise_gru.get_encoded_sequence(train_data, vocabulary_obj, max_len)
-
-    # print total number of instances
-    print("# of instances: ", len(y))
-
-    # create and fit the model
-    model = Sequential()
-    model.add(GRU(lstm_out, input_shape=(max_len, 1), return_sequences=True))
-    model.add(TimeDistributed(Dense(1, activation='sigmoid')))
-    model.compile(loss='binary_crossentropy', optimizer='adam', metrics=['accuracy'])
-
-    for i in range(X.shape[0]):
-        x_data = X[i]
-        y_data = y[i]
-        if i % 100 == 0:
-            print("Records processed: ", i)
-        model.fit(x_data, y_data, callbacks=[ResetStatesCallback()], shuffle=False, batch_size=batch_size, verbose=0)
-
-    # get encoded test data
[…]

### 2018-11-13 rnn/seg-lstm-model-stepwise.py

@@ -1,116 +1,0 @@
-import numpy
-
-from keras.models import Sequential
-from keras.layers import GRU, LSTM
-from keras.layers import Dense
-from keras.layers import TimeDistributed
-from keras.callbacks import Callback
-
-import utility_stepwise
-
-from vocabulary import Vocabulary
-
-# system configuration
-max_len = 50
-folds = 5
-lstm_out = 25
-batch_size = 1
-
-
-class ResetStatesCallback(Callback):
-    def __init__(self):
-        self.counter = 0
-
-    def on_batch_begin(self, batch, logs={}):
-        if self.counter % max_len == 0:
-            self.model.reset_states()
-            # print("Reset States..............")
-        self.counter += 1
-
-
-# create e-coaching dictionary
-vocabulary_obj = Vocabulary()
-vocabulary_obj.create()
-
-# create mapping of word to integers and the reverse
-char_to_int = dict((c, i+1) for i, c in enumerate(vocabulary_obj.vocab))
-int_to_char = dict((i+1, c) for i, c in enumerate(vocabulary_obj.vocab))
-
-# prepare the data set of input to output pairs encoded as integers
-data = utility_stepwise.get_ecoaching_data()
-folds_data = utility_stepwise.split_data(data, folds)
-results = []
-
-for fold_num in range(folds):
-    train_data, test_data = [], []
-    for i in range(folds):
-        if i == fold_num:
-            test_data = folds_data[i]
-        else:
-            train_data = train_data + folds_data[i]
-
-    # get encoded training data
-    X, y = utility_stepwise.get_encoded_sequence(train_data, vocabulary_obj, max_len)
-
-    # print total number of instances
-    print("# of instances: ", len(y))
-
-    # create and fit the model
-    model = Sequential()
-    model.add(LSTM(lstm_out, input_shape=(max_len, 1), return_sequences=True))
-    model.add(TimeDistributed(Dense(1, activation='sigmoid')))
-    model.compile(loss='binary_crossentropy', optimizer='adam', metrics=['accuracy'])
-
-    for i in range(X.shape[0]):
-        x_data = X[i]
-        y_data = y[i]
-        if i % 100 == 0:
-            print("Records processed: ", i)
-        model.fit(x_data, y_data, callbacks=[ResetStatesCallback()], shuffle=False, batch_size=batch_size, verbose=0)
-
-    # get encoded test data
[…]

### 2018-11-13 rnn/utility_stepwise.py

@@ -1,128 +1,0 @@
-# include required packages
-import numpy
-import re
-import sys
-
-from keras.utils import np_utils
-from keras.preprocessing.sequence import pad_sequences
-
-
-# load e-coaching data from datafile
-def get_ecoaching_data(file_name='e-coaching_data_stem.txt'):
-    data = []
-    with open(file_name, "r") as file_stream:
-        for line in file_stream:
-            current_line = line.split(",")
-            if len(current_line) > 1:
-                data.append(clean_text(current_line[0]).strip())
-    return data
-
-
-# split data into 5 folds
-def split_data(data, folds=5):
-    fold_size = len(data)/folds
-    folds_data = []
-    for i in range(folds):
-        start, end = i*fold_size, i*fold_size + fold_size
-        folds_data.append(data[start: end])
-    return folds_data
-
-
-# encode text data into numeric values
-def get_encoded_sequence(data, vocabulary_obj, max_len):
-    x, y = [], []
-
-    for text in data:
-        sample_x = vocabulary_obj.vectorize(text)
-        sample_y = [0]*(len(sample_x)-1)
-        sample_y.append(1)
-
-        x.append(sample_x)
-        y.append(sample_y)
-
-    data_x = numpy.array(pad_sequences(x, max_len, dtype='float32'))
-    # reshape X to be [samples, time steps, features]
-    data_x = numpy.reshape(data_x, (data_x.shape[0], 1, max_len, 1))
-    # normalize
-    data_x = data_x / vocabulary_obj.size()
-    data_y = numpy.array(pad_sequences(y, max_len))
-    data_y = numpy.reshape(data_y, (data_y.shape[0], 1, max_len, 1))
-    return data_x, data_y
-
-
-# clean text, removed weired chars
-def clean_text(text):
-    new_text = re.sub(r'[^\x00-\x7f]', r'', text)
-    new_text = new_text.replace("\n", "").strip()
-    new_text = re.sub(r'[?|`|>|<|$|.|!|;|:|=|&|\'|\"|\-|/|)|(|\]|[|]', r'', new_text)
-    return new_text
-
-
-# evaluate model with F1, Precision and Recall
-def get_macro_average_performance(actual, predicted):
-    f = open("results_LSTM_summit.txt", "a")
-    f.write("fold number \n")
-    f.write("actual: ")
-    f.writelines(["%s," % item for item in actual])
-    f.write("\npredicted: ")
-    f.writelines(["%s," % item for item in predicted])
[…]

### 2018-11-13 rnn/utility_stepwise_gru.py

@@ -1,128 +1,0 @@
-# include required packages
-import numpy
-import re
-import sys
-
-from keras.utils import np_utils
-from keras.preprocessing.sequence import pad_sequences
-
-
-# load e-coaching data from datafile
-def get_ecoaching_data(file_name='e-coaching_data_stem.txt'):
-    data = []
-    with open(file_name, "r") as file_stream:
-        for line in file_stream:
-            current_line = line.split(",")
-            if len(current_line) > 1:
-                data.append(clean_text(current_line[0]).strip())
-    return data
-
-
-# split data into 5 folds
-def split_data(data, folds=5):
-    fold_size = len(data)/folds
-    folds_data = []
-    for i in range(folds):
-        start, end = i*fold_size, i*fold_size + fold_size
-        folds_data.append(data[start: end])
-    return folds_data
-
-
-# encode text data into numeric values
-def get_encoded_sequence(data, vocabulary_obj, max_len):
-    x, y = [], []
-
-    for text in data:
-        sample_x = vocabulary_obj.vectorize(text)
-        sample_y = [0]*(len(sample_x)-1)
-        sample_y.append(1)
-
-        x.append(sample_x)
-        y.append(sample_y)
-
-    data_x = numpy.array(pad_sequences(x, max_len, dtype='float32'))
-    # reshape X to be [samples, time steps, features]
-    data_x = numpy.reshape(data_x, (data_x.shape[0], 1, max_len, 1))
-    # normalize
-    data_x = data_x / vocabulary_obj.size()
-    data_y = numpy.array(pad_sequences(y, max_len))
-    data_y = numpy.reshape(data_y, (data_y.shape[0], 1, max_len, 1))
-    return data_x, data_y
-
-
-# clean text, removed weired chars
-def clean_text(text):
-    new_text = re.sub(r'[^\x00-\x7f]', r'', text)
-    new_text = new_text.replace("\n", "").strip()
-    new_text = re.sub(r'[?|`|>|<|$|.|!|;|:|=|&|\'|\"|\-|/|)|(|\]|[|]', r'', new_text)
-    return new_text
-
-
-# evaluate model with F1, Precision and Recall
-def get_macro_average_performance(actual, predicted):
-    f = open("results_GRU_summit.txt", "a")
-    f.write("fold number \n")
-    f.write("actual: ")
-    f.writelines(["%s," % item for item in actual])
-    f.write("\npredicted: ")
-    f.writelines(["%s," % item for item in predicted])
[…]

### 2018-11-13 src/ColLTA.cpp

@@ -1,160 +1,0 @@
-// $Id: ColLTA.cpp,v 1.2 2014/06/18 01:16:15 fn6418 Exp $
-
-// Author: Alexander Kotov (kotov@wayne.edu), Wayne State University, 2013
-
-#include "Utils.hpp"
-#include "LTA.hpp"
-
-void usage(const char* progName) {
-  printf("Usage: %s arg_file \n", progName);
-  printf("Arguments:\n");
-  printf("  sids_path - path to the file with sample IDs\n");
-  printf("  col_path - path to the file with classified sample collection\n");
-  printf("  beta - concentration hyper-parameter of Dirichlet prior for language models\n");
-  printf("  gamma_bkg - Beta prior for the background component in document background-class LMs mixture\n");
-  printf("  gamma_cls - Beta prior for the class-specific component in document background-class LMs mixture\n");
-  printf("  num_iters - number of iterations\n");
-  printf("  rnd_seed - seed for random number generator (initial state of the Markov chain)\n");
-  printf("  bkg_lm_file - file with background LM\n");
-  printf("  cls_lms_file - file with class-specific LMs\n");
-  printf("  verbose - print debug information\n");
-}
-
-int main(int argc, char* argv[]) {
-  std::string sidsPath, colPath;
-  IDStrMap id2class, id2term;
-  bool verbose = false;
-  unsigned int numIters = 100, rndSeed = 0;
-  float beta = BETA_DEF, gammaBkg = GAMMA_BKG_DEF, gammaCls = GAMMA_CLS_DEF;
-  FILE *blmFD = stdout, *clmFD = stdout;
-  StrSet *sampIDs = NULL;
-  ClsSampCol *sampCol = NULL;
-  LTA *model = NULL;
-
-  if(argc != 2) {
-    usage(argv[0]);
-    exit(1);
-  }
-
-  try {
-    StrStrMapIt pit;
-
-    StrStrMap* params = readConfFile(argv[1]);
-
-    if((pit = params->find("sids_path")) != params->end()) {
-      sidsPath = pit->second;
-    } else {
-      fprintf(stderr, "Error: path to the file with sample IDs is not in the configuration file (parameter sids_path)\n");
-      exit(1);
-    }
-
-    if((pit = params->find("col_path")) != params->end()) {
-      colPath = pit->second;
-    } else {
[…]

### 2018-11-13 src/ColLabLDA.cpp

@@ -1,145 +1,0 @@
-/* -*- C++ -*- */
-
-// Author: Alexander Kotov (kotov@wayne.edu), Wayne State University, 2014
-// $Id: ColLabLDA.cpp,v 1.3 2014/06/22 02:30:48 fn6418 Exp $
-
-#include <cstdio>
-#include <cstdlib>
-#include "Utils.hpp"
-#include "LabLDA.hpp"
-
-void usage(const char* progName) {
-  printf("Usage: %s arg_file\n", progName);
-  printf("Arguments:\n");
-  printf("  sids_path - path to the file with sample IDs\n");
-  printf("  col_path - path to the file with tagged document collection\n");
-  printf("  alpha - concentration hyper-parameter of Dirichlet prior for document-topic distributions\n");
-  printf("  beta - concentration hyper-parameter of Dirichlet prior for word-topic-class distributions\n");
-  printf("  num_cls_tops - number of topics per class\n");
-  printf("  num_iters - number of iterations of the Gibbs sampler\n");
-  printf("  rnd_seed - seed for random number generator (initial state of the Markov chain)\n");
-  printf("  cls_tops_file - file with topics for each class\n");
-  printf("  verbose - print debug information\n");
-}
-
-int main(int argc, char* argv[]) {
-  std::string sidsPath, colPath;
-  IDStrMap id2class, id2term;
-  bool verbose = false;
-  float alpha = ALPHA_DEF, beta = BETA_DEF;
-  unsigned int numClsTops = 0, numIters = 100, rndSeed = 0;
-  FILE *wtFD = stdout;
-  StrSet *sampIDs = NULL;
-  ClsSampCol *sampCol = NULL;
-  LabLDA *model = NULL;
-
-  if(argc != 2) {
-    usage(argv[0]);
-    exit(1);
-  }
-
-  try {
-    StrStrMapIt pit;
-
-    StrStrMap* params = readConfFile(argv[1]);
-
-    if((pit = params->find("sids_path")) != params->end()) {
-      sidsPath = pit->second;
-    } else {
-      fprintf(stderr, "Error: path to the file with sample IDs is not in the configuration file (parameter sids_path)\n");
-      exit(1);
-    }
-
-    if((pit = params->find("col_path")) != params->end()) {
-      colPath = pit->second;
-    } else {
[…]

### 2018-11-13 src/ColMGLabLDA.cpp

@@ -1,169 +1,0 @@
-/* -*- C++ -*- */
-
-// Author: Alexander Kotov (kotov@wayne.edu), Wayne State University, 2014
-// $Id: ColMGLabLDA.cpp,v 1.3 2014/06/23 03:59:38 fn6418 Exp $
-
-#include <cstdio>
-#include <cstdlib>
-#include "Utils.hpp"
-#include "MGLabLDA.hpp"
-
-void usage(const char* progName) {
-  printf("Usage: %s arg_file\n", progName);
-  printf("Arguments:\n");
-  printf("  sids_path - path to the file with sample IDs\n");
-  printf("  col_path - path to the file with classified sample collection\n");
-  printf("  alpha - concentration hyper-parameter of Dirichlet prior for document-topic distributions\n");
-  printf("  beta - concentration hyper-parameter of Dirichlet prior for language models\n");
-  printf("  gamma_bkg - Beta prior for the background component in document background-class LMs mixture\n");
-  printf("  gamma_cls - Beta prior for the class-specific component in document background-class LMs mixture\n");
-  printf("  num_cls_tops - number of topics per class\n");
-  printf("  num_iters - number of iterations\n");
-  printf("  rnd_seed - seed for random number generator (initial state of the Markov chain)\n");
-  printf("  bkg_top_file - file with background topic\n");
-  printf("  cls_tops_file - file with class-specific topics\n");
-  printf("  verbose - print debug information\n");
-}
-
-int main(int argc, char* argv[]) {
-  std::string sidsPath, colPath;
-  IDStrMap id2class, id2term;
-  bool verbose = false;
-  unsigned int numClsTops = 0, numIters = 100, rndSeed = 0;
-  float alpha = ALPHA_DEF, beta = BETA_DEF, gammaBkg = GAMMA_BKG_DEF, gammaCls = GAMMA_CLS_DEF;
-  FILE *bwtFD = stdout, *cwtFD = stdout;
-  StrSet *sampIDs = NULL;
-  ClsSampCol *sampCol = NULL;
-  MGLabLDA *model = NULL;
-
-  if(argc != 2) {
-    usage(argv[0]);
-    exit(1);
-  }
-
-  try {
-    StrStrMapIt pit;
-
-    StrStrMap* params = readConfFile(argv[1]);
-
-    if((pit = params->find("sids_path")) != params->end()) {
-      sidsPath = pit->second;
-    } else {
[…]

### 2018-11-13 src/DirReader.cpp

@@ -1,151 +1,0 @@
-/* -*- C++ -*- */
-
-// Author: Alexander Kotov (Emory University) alex.s.kotov@gmail.com, 2011
-// $Rev$
-
-#include "DirReader.hpp"
-
-DirReader::~DirReader() {
-  while(!_dirs.empty()) {
-    while((closedir(_dirs.top()) == -1) && (errno = EINTR));
-    _dirs.pop();
-  }
-}
-
-bool DirReader::pathExists(const char *path) {
-  struct stat statbuf;
-  return stat(path, &statbuf) != -1;
-}
-
-bool DirReader::fileExists(const char *file, const char *dir) {
-  DIR* dirp = NULL;
-  struct dirent *dirent;
-
-  if(dir == NULL) {
-    dirp = opendir(".");
-  } else {
-    if(!DirReader::isDir(dir)) {
-      return false;
-    } else {
-      dirp = opendir(dir);
-    }
-  }
-
-  while((dirent = readdir(dirp)) != NULL) {
-    if(strcmp(dirent->d_name, file) == 0) {
-      return true;
-    }
-  }
-
-  return false;
-}
-
-bool DirReader::isDir(const char *path) {
-  struct stat statbuf;
-  if(stat(path, &statbuf) == -1) {
-    return false;
-  } else {
-    return S_ISDIR(statbuf.st_mode);
-  }
-}
-
-bool DirReader::isFile(const char *path) {
-  struct stat statbuf;
-  if(stat(path, &statbuf) == -1) {
-    return false;
-  } else {
-    return S_ISREG(statbuf.st_mode);
-  }
-}
-
-void DirReader::open() throw(Exception) {
-  struct stat statbuf;
-
-  if(stat(_path.c_str(), &statbuf) == -1) {
-    throwException(std::string("DirReader::open()"), _path + std::string(" does not exist"));
-  } else {
-    if(!S_ISDIR(statbuf.st_mode)) {
-      throwException(std::string("DirReader::open()"), _path + std::string(" is not a directory"));
-    }
-  }
-
-  _rel_path = "";
-  addDir(_path.c_str());
-}
-
-void DirReader::open(const std::string& path) throw(Exception) {
-  _path = path;
-  open();
-}
-
-void DirReader::addDir(const char* path) throw(Exception) {
-  DIR* dir = opendir(path);
-  if(dir == NULL) {
-    if(errno == EACCES) {
-      throwException(std::string("DirReader::open()"), "Access denied");
-    } else {
[…]

### 2018-11-13 src/DirReader.hpp

@@ -1,70 +1,0 @@
-/* -*- C++ -*- */
-
-// Author: Alexander Kotov (Emory University) alex.s.kotov@gmail.com, 2011
-// $Rev$
-
-#ifndef DIRREADER_HPP_
-#define DIRREADER_HPP_
-
-#include <stdio.h>
-#include <dirent.h>
-#include <errno.h>
-#include <string.h>
-#include <sys/stat.h>
-#include <unistd.h>
-#include <limits.h>
-#include <exception>
-#include <string>
-#include <stack>
-#include "Exception.hpp"
-
-#ifndef PATH_MAX
-#define PATH_MAX 512
-#endif
-
-/* Class provides sequential access to the contents of a directory, including its sub-directories */
-
-class DirReader {
-public:
-  DirReader()
-  { }
-  DirReader(const std::string& dirPath) : _path(dirPath)
-  { }
-  DirReader(const char* dirPath) : _path(dirPath)
-  { }
-  ~DirReader();
-  // check if path is a directory
-  static bool isDir(const char *path);
-  // check if path is a regular file
-  static bool isFile(const char *path);
-  // check if path exists
-  static bool pathExists(const char *path);
-  // check if file exists in directory
-  static bool fileExists(const char *file, const char *dir = NULL);
-  // open directory for reading
-  void open() throw(Exception);
-  // open directory for reading
-  void open(const std::string& path) throw(Exception);
-  // get the name of the next file in the directory
-  std::string getFileName();
-  // get the name of the next file in the directory with the full path to it
-  std::string getFullFilePath();
-  // get the name of the next file with the path relative to the top directory
-  std::string getRelFileName();
-  // go to the next file in the directory
-  bool goToNext();
-  // get path to the directory
-  inline std::string getPath() const {
-    return _path;
-  }
-private:
-  void addDir(const char* path) throw(Exception);
-private:
-  std::stack<DIR* > _dirs;
-  std::string _rel_path;
-  std::string _path;
-  char _cwd[PATH_MAX];
-  dirent *_dire;
-};
-
-#endif // DIRREADER_HPP_

### 2018-11-13 src/Exception.hpp

@@ -1,107 +1,0 @@
-/* -*- C++ -*- */
-
-// Author: Alexander Kotov (alex.s.kotov@gmail.com), Emory University, 2011
-// $Id: Exception.hpp,v 1.1 2014/01/10 06:08:21 fn6418 Exp $
-
-#ifndef EXCEPTION_HPP_
-#define EXCEPTION_HPP_
-
-#include <string>
-#include <exception>
-#include <stdarg.h>
-#include <stdio.h>
-
-#define MAX_MSG_LEN 512
-
-class Exception : public std::exception {
-public:
-  Exception(const char *loc, const char *msg) :  _loc(loc), _msg(msg) {
-    update();
-  }
-
-  Exception(const char *msg) : _loc(""), _msg(msg) {
-    update();
-  }
-
-  Exception(const char *loc, const char *msg, ...) {
-    va_list args;
-    _loc = loc;
-    va_start(args, msg);
-    create_msg(msg, args);
-    va_end(args);
-  }
-
-  Exception(const char *msg, ...) {
-    va_list args;
-    _loc = "";
-    va_start(args, msg);
-    create_msg(msg, args);
-    va_end(args);
-  }
-
-  Exception(const std::string &loc, const std::string &msg) : _loc(loc), _msg(msg) {
-    update();
-  }
-
-  Exception(const std::string &msg) : _msg(msg) {
-    update();
-  }
-
-  Exception(const Exception& e) {
-    _loc = e._loc;
-    _msg = e._msg;
-    update();
-  }
-
-  Exception& operator=(Exception& e) {
-    _loc = e._loc;
-    _msg = e._msg;
-    update();
-    return *this;
-  }
-
-  virtual ~Exception() throw() {}
-
-  const char* get_loc() const { return _loc.c_str(); }
-  const char* get_msg() const { return _msg.c_str(); }
-
-  virtual const char *what() const throw() {
-    return _desc.c_str();
-  }
-
-protected:
-  void update() {
-    _desc = _loc.empty() ? _msg : "(" + _loc + ") " + _msg;
-  }
-
-  void create_msg(const char *msg, va_list args) {
-    char buf[MAX_MSG_LEN];
-    vsnprintf(buf, MAX_MSG_LEN, msg, args);
-    _msg = buf;
-  }
-
-protected:
-  std::string _loc;
-  std::string _msg;
-  std::string _desc;
-};
-
-#if defined(_DEBUG) || defined(_VERBOSE_DEBUG)
-#define throwException(loc, msg)                                                      \
[…]

### 2018-11-13 src/Formatter.cpp

@@ -1,415 +1,0 @@
-// $Id: Formatter.cpp,v 1.6 2014/06/23 20:44:00 fn6418 Exp $
-
-// Author: Alexander Kotov (kotov@wayne.edu), Wayne State University, 2013
-
-#include "Utils.hpp"
-#include "DirReader.hpp"
-
-
-typedef struct SampInfo_ {
-  SampInfo_() : shortCode(""), longCode("")
-  { }
-
-  SampInfo_(const std::string& shortCode, const std::string& longCode, const StrVec& sampTerms) : shortCode(shortCode), longCode(longCode), sampTerms(sampTerms)
-  { }
-
-  std::string shortCode;
-  std::string longCode;
-  StrVec sampTerms;
-} SampInfo;
-
-typedef std::vector<SampInfo> SampVec;
-typedef SampVec::const_iterator cSampVecIt;
-
-typedef struct TermStats_ {
-  TermStats_() : term(""), sampFrac(0)
-  { }
-
-  TermStats_(const std::string& term, float sampFrac) : term(term), sampFrac(sampFrac)
-  { }
-
-  bool operator<(const TermStats_& termStats) const {
-    return termStats.sampFrac <= sampFrac;
-  }
-
-  std::string term;
-  float sampFrac;
-} TermStats;
-
-typedef std::set<TermStats> TermStatsSet;
-typedef TermStatsSet::iterator TermStatsSetIt;
-
-void usage(const char* progName) {
-  printf("Usage: %s arg_file\n", progName);
-  printf("Arguments:\n");
-  printf("  in_dir - path to the input directory\n");
-  printf("  code_map_file - file with mapping of classification codes\n");
-  printf("  code_stats_file - file with statistics of classification codes\n");
-  printf("  out_xml_file - path to the output file in XML format\n");
-  printf("  out_arff_file - path to the output file in XML format\n");
-  printf("  term_stats_file - path to the file with term statistics\n");
-  printf("  max_samp_frac - maximum fraction of samples a term can occur\n");
-  printf("  col_lm_file - path to the file with collection LM\n");
-  printf("  remove_stopwords - remove stopwords\n");
-  printf("  stop_words - path to the file with a list of stopwords\n");
[…]

### 2018-11-13 src/LDAClass.cpp

@@ -1,199 +1,0 @@
-// $Id: LDAClass.cpp,v 1.3 2014/06/23 03:59:38 fn6418 Exp $
-
-// Copyright: Alexander Kotov (kotov@wayne.edu), Wayne State University, 2014
-
-#include "Utils.hpp"
-
-void usage(const char* progName) {
-  printf("Usage: %s arg_file\n", progName);
-  printf("Arguments:\n");
-  printf("  sids_path - path to the file with sample IDs\n");
-  printf("  col_path - path to the file with sample collection\n");
-  printf("  col_lm_path - path to the file with background topic\n");
-  printf("  cls_tops_path - path to the file with class-specific topics\n");
-  printf("  out_file - file with classification results\n");
-  printf("  verbose - print debug information\n");
-}
-
-ClsTopics* loadClsTopsFile(const char *filePath, StrSet** clsSet) throw(Exception) {
-  unsigned int topID;
-  std::string line, curCls;
-  std::string::size_type lpos, rpos;
-  ClsTopics* clsTopics = NULL;
-  IDLangModMap* curTopMap = NULL;
-  LangMod* curTopic = NULL;
-  std::ifstream fs(filePath);
-
-  if(!fs.is_open()) {
-    throwException("parsing file with geo-specific topic-word distributions", std::string("can't open file ") + filePath);
-  }
-
-  clsTopics = new ClsTopics();
-  *clsSet = new StrSet();
-
-  while(std::getline(fs, line)) {
-    StrVec toks;
-    stripChars(line);
-
-    if(!line.length()) {
-      continue;
-    }
-
-    splitToks(line, toks, "\t");
-
-    if(toks.size() == 1) {
-      if((lpos = toks[0].find('<')) != std::string::npos) {
-        if((rpos = toks[0].find('>')) != std::string::npos) {
-          curCls = toks[0].substr(lpos+1, rpos-lpos-1);
-          (*clsSet)->insert(curCls);
-          clsTopics->insert(std::make_pair(curCls, IDLangModMap()));
-          curTopMap = &((*clsTopics)[curCls]);
-        } else {
-          fprintf(stderr, "Warning: malformed geo-location\n");
-        }
-      } else {
-        topID = atoi(toks[0].c_str());
-        curTopMap->insert(std::make_pair(topID, LangMod()));
[…]

### 2018-11-13 src/LTA.cpp

@@ -1,307 +1,0 @@
-// $Id: LTA.cpp,v 1.3 2014/06/18 01:16:15 fn6418 Exp $
-
-// Author: Alexander Kotov (kotov@wayne.edu), Emory University, 2014
-
-#include "LTA.hpp"
-
-LTA::LTA() {
-  setDefaults();
-  _betaBkg = BETA_BKG_DEF; _betaCls = BETA_CLS_DEF; _gammaBkg = GAMMA_BKG_DEF; _gammaCls = GAMMA_CLS_DEF;
-}
-
-LTA::LTA(float betaBkg, float betaCls, float gammaBkg, float gammaCls) {
-  setDefaults();
-  _betaBkg = betaBkg; _betaCls = betaCls; _gammaBkg = gammaBkg; _gammaCls = gammaCls;
-}
-
-void LTA::setDefaults() {
-  _C = 0; _S = 0; _W = 0;
-  _betaBkg = 0; _betaCls = 0; _gammaBkg = 0; _gammaCls = 0;
-  _wc = NULL; _c = NULL; _sb = NULL; _sc = NULL; _wb = NULL; _b = 0; _Nd = NULL; _wt = NULL;
-  _bphi = NULL; _cphi = NULL;
-}
-
-LTA::~LTA() {
-  delTrainData();
-  delLangModData();
-}
-
-void LTA::trainInit() {
-  unsigned int w, c, d;
-
-  _wc = new unsigned int*[_W];
-  for(w = 0; w < _W; w++) {
-    _wc[w] = new unsigned int[_C];
-    for(c = 0; c < _C; c++) {
-      _wc[w][c] = 0;
-    }
-  }
-
-  _c = new unsigned int[_C];
-  for(c = 0; c < _C; c++) {
-    _c[c] = 0;
-  }
-
-  _sb = new unsigned int[_S];
-  _sc = new unsigned int[_S];
-  for(d = 0; d < _S; d++) {
-    _sb[d] = 0;
-    _sc[d] = 0;
-  }
-
-  _wb = new unsigned int[_W];
-  for(w = 0; w < _W; w++) {
-    _wb[w] = 0;
-  }
-
-  _Nd = new unsigned int[_S];
-  _wt = new bool*[_S];
-}
-
-void LTA::langModInit() {
-  unsigned int w, c;
-
-  _bphi = new float[_W];
-  for(w = 0; w < _W; w++) {
-    _bphi[w] = 0;
-  }
-
[…]

### 2018-11-13 src/LTA.hpp

@@ -1,90 +1,0 @@
-// $Id: LTA.hpp,v 1.2 2014/06/18 01:16:15 fn6418 Exp $
-
-/*
- ******************************************************************************
- *      Latent Term Allocation Model with Background and Class-specific terms
- ******************************************************************************
- *  Author: Alexander Kotov (kotov@wayne.edu), Wayne State University, 2014
- */
-
-#ifndef LTA_HPP_
-#define LTA_HPP_
-
-#include <time.h>
-#include <math.h>
-#include <stdio.h>
-#include <stdlib.h>
-#include "Utils.hpp"
-
-#define BETA_DEF 0.01
-#define BETA_BKG_DEF 0.01
-#define BETA_CLS_DEF 0.01
-#define GAMMA_BKG_DEF 1
-#define GAMMA_CLS_DEF 1
-
-class LTA {
-public:
-  LTA();
-  LTA(float betaBkg, float betaCls, float gammaBkg, float gammaCls);
-  virtual ~LTA();
-  void runSampler(const ClsSampCol* sampCol, unsigned int numCls, unsigned int vocSize, unsigned int numIters, unsigned int rndSeed, bool verbose);
-  void computePhi();
-  void storeBkgLM(FILE* fd, IDStrMap& id2term);
-  void storeClassLMs(FILE* fd, IDStrMap& id2class, IDStrMap& id2term);
-
-protected:
-  // set default values
-  void setDefaults();
-  // allocate memory and initialize training data
-  void trainInit();
-  // allocate memory and initialize language models
-  void langModInit();
-  // deallocate memory for model training data
-  void delTrainData();
-  // deallocate memory for language model data
-  void delLangModData();
-  // sample a type for a word in a document
-  void sampleTermType(unsigned int sampID, unsigned int sampCls, unsigned int termInd, unsigned int termID, bool& termType);
-protected:
-  ////////////////////////////////////////////////////////////////////////////////////
-  // model parameters
-  ///////////////////////////////////////////////////////////////////////////////////
-  // number of classes
-  unsigned int _C;
-  // number of samples
-  unsigned int _S;
-  // number of unique words in collection vocabulary
-  unsigned int _W;
-  // hyper-parameters
-  float _betaBkg, _betaCls, _gammaBkg, _gammaCls;
-  ////////////////////////////////////////////////////////////////////////////////////
-  // training data
-  ///////////////////////////////////////////////////////////////////////////////////
-  // number of times a word has been assigned to each class (size W x C)
-  unsigned int **_wc;
[…]

### 2018-11-13 src/LTAClass.cpp

@@ -1,124 +1,0 @@
-// $Id: LTAClass.cpp,v 1.5 2014/06/23 03:59:38 fn6418 Exp $
-
-// Author: Alexander Kotov (kotov@wayne.edu), Wayne State University, 2014
-
-#include "Utils.hpp"
-
-void usage(const char* progName) {
-  printf("Usage: %s arg_file\n", progName);
-  printf("Arguments:\n");
-  printf("  sids_path - path to the file with sample IDs\n");
-  printf("  col_path - path to the file with sample collection\n");
-  printf("  col_lm_path - path to the file with background LM\n");
-  printf("  cls_lms_path - path to the file with class-specific LMs\n");
-  printf("  out_file - file with classification results\n");
-  printf("  verbose - print debug information\n");
-}
-
-int main(int argc, char* argv[]) {
-  bool verbose = false;
-  std::string sidsPath, colPath, colLMPath, clsLMsPath;
-  FILE *outFD = stdout;
-  StrSet *sampIDs = NULL, *clsIDs = NULL;
-  StrFloatMap *clsStats = NULL;
-  SampCol *sampCol = NULL;
-  LangMod *colLM = NULL;
-  StrLangModMap *clsLMs = NULL, *termClsProbs = NULL;
-
-  if(argc != 2) {
-    usage(argv[0]);
-    exit(1);
-  }
-
-  try {
-    StrStrMapIt pit;
-
-    StrStrMap* params = readConfFile(argv[1]);
-
-    if((pit = params->find("sids_path")) != params->end()) {
-      sidsPath = pit->second;
-    } else {
-      fprintf(stderr, "Error: path to the file with sample IDs is not in the configuration file (parameter sids_path)\n");
-      exit(1);
-    }
-
-    if((pit = params->find("col_path")) != params->end()) {
-      colPath = pit->second;
-    } else {
-      fprintf(stderr, "Error: path to the collection file is not in the configuration file (parameter col_path)\n");
-      exit(1);
-    }
-
-    if((pit = params->find("col_lm_path")) != params->end()) {
-      colLMPath = pit->second;
-    } else {
-      fprintf(stderr, "Error: path to the file with background LM is not in the configuration file (parameter col_lm_path)\n");
-      exit(1);
-    }
-
-    if((pit = params->find("cls_lms_path")) != params->end()) {
[…]

### 2018-11-13 src/LabLDA.cpp

@@ -1,305 +1,0 @@
-/* -*- C++ -*- */
-
-// Author: Alexander Kotov (alex.s.kotov@gmail.com), Emory University, 2013
-
-#include "LabLDA.hpp"
-
-LabLDA::LabLDA(unsigned int numClsTops) {
-  setDefaults();
-  _K = numClsTops; _alpha = ALPHA_DEF; _beta = BETA_DEF;
-}
-
-LabLDA::LabLDA(unsigned int numClsTops, float alpha, float beta) {
-  setDefaults();
-  _K = numClsTops; _alpha = alpha; _beta = beta;
-}
-
-LabLDA::~LabLDA() {
-  delTrainData();
-  delModelData();
-}
-
-void LabLDA::runSampler(const ClsSampCol* sampCol, unsigned int numCls, unsigned int vocSize, unsigned int numIters,
-                            unsigned int rndSeed, bool verbose) {
-  unsigned int *ind2tid = NULL;
-  unsigned int sampInd, termInd, termID, curIter, topic;
-  ClsSampCol& C = *((ClsSampCol* ) sampCol);
-
-  _C = numCls;
-  _W = vocSize;
-  _M = C.size();
-
-  if(verbose) {
-    fprintf(stderr, "Initializing training data...");
-  }
-
-  trainInit();
-
-  if(rndSeed == 0) {
-    srandom(time(NULL));
-  } else {
-    srandom(rndSeed);
-  }
-
-  for(sampInd = 0; sampInd < C.size(); sampInd++) {
-    _z[sampInd] = new unsigned int[C[sampInd].termIDs.size()];
-
-    for(termInd = 0; termInd < C[sampInd].termIDs.size(); termInd++) {
-      termID = C[sampInd].termIDs[termInd];
-      topic = (int)(random() / (RAND_MAX + 1.0) * _K);
-      _z[sampInd][termInd] = topic;
-      _wcz[termID][C[sampInd].sampCls][topic]++;
-      _cz[C[sampInd].sampCls][topic]++;
-      _dz[sampInd][topic]++;
-    }
-
-    delete ind2tid;
-  }
-
-  if(verbose) {
-    fprintf(stderr, "OK\n");
-    fprintf(stderr, "Number of document tags: %u\n", _C);
-    fprintf(stderr, "Number of documents: %u\n", _M);
[…]

### 2018-11-13 src/LabLDA.hpp

@@ -1,80 +1,0 @@
-/* -*- C++ -*- */
-
-// Author: Alexander Kotov (kotov@wayne.edu), Wayne State University, 2014
-// $Id: LabLDA.hpp,v 1.4 2014/06/22 02:30:48 fn6418 Exp $
-
-#ifndef LABLDA_HPP_
-#define LABLDA_HPP_
-
-#include <ctime>
-#include <cmath>
-#include <cstdio>
-#include <cstdlib>
-#include "Utils.hpp"
-
-#define ALPHA_DEF 0.01
-#define BETA_DEF 0.1
-
-class LabLDA {
-public:
-  LabLDA(unsigned int numClsTops);
-  LabLDA(unsigned int numClsTops, float alpha, float beta);
-  virtual ~LabLDA();
-  // estimate the parameters of the model using Gibbs sampling
-  void runSampler(const ClsSampCol* docCol, unsigned int numCls, unsigned int vocSize, unsigned int numIters,
-                    unsigned int rndSeed=0, bool verbose=false);
-  // compute word-topic-class distributions
-  void computePhi();
-  // save global topics for each tag into a file
-  void storeClassTopics(FILE* fd, const IDStrMap& id2tag, const IDStrMap& id2term, unsigned int maxTerms=0, float probThresh=0);
-
-protected:
-  void setDefaults();
-  // allocate memory and initialize training data
-  void trainInit();
-  // allocate memory and initialize class-specific topics
-  void classTopicsInit();
-  // deallocate memory for training data
-  void delTrainData();
-  // deallocate memory for word-tag-topic distributions
-  void delModelData();
-  // sample a topic for word n of document m during training
-  unsigned int sampleTopic(unsigned int docID, unsigned int docCls, unsigned int docLen, unsigned int termID, unsigned int termInd);
-
-protected:
-  ////////////////////////////////////////////////////////////////////////////////////
-  // model parameters
-  ///////////////////////////////////////////////////////////////////////////////////
-  // number of distinct classes
-  unsigned int _C;
-  // number of topics per class
-  unsigned int _K;
-  // number of documents
-  unsigned int _M;
-  // vocabulary size
-  unsigned int _W;
-  // hyper-parameters
-  float _alpha, _beta;
-  ////////////////////////////////////////////////////////////////////////////////////
-  // training data
-  ///////////////////////////////////////////////////////////////////////////////////
-  // number of times each word has been labeled with different topics for each class (size W x C x K)
-  unsigned int ***_wcz;
[…]

### 2018-11-13 src/MGLabLDA.cpp

@@ -1,391 +1,0 @@
-/* -*- C++ -*- */
-
-// Author: Alexander Kotov (kotov@wayne.edu), Wayne State University, 2014
-// $Id: MGLabLDA.cpp,v 1.4 2014/06/23 20:44:00 fn6418 Exp $
-
-#include "MGLabLDA.hpp"
-
-MGLabLDA::MGLabLDA(unsigned int numClsTops) {
-  setDefaults();
-  _Kcls = numClsTops; _alpha = ALPHA_DEF; _betaBkg = BETA_BKG_DEF; _betaCls = BETA_CLS_DEF; _gammaBkg = GAMMA_BKG_DEF; _gammaCls = GAMMA_CLS_DEF;
-}
-
-MGLabLDA::MGLabLDA(unsigned int numClsTops, float alpha, float betaBkg, float betaCls, float gammaBkg, float gammaCls) {
-  setDefaults();
-  _Kcls = numClsTops; _alpha = alpha; _betaBkg = betaBkg; _betaCls = betaCls; _gammaBkg = gammaBkg; _gammaCls = gammaCls;
-}
-
-void MGLabLDA::setDefaults() {
-  _C = 0; _Kcls = 0; _M = 0; _W = 0;
-  _alpha = 0; _betaBkg = 0; _betaCls = 0; _gammaBkg = 0; _gammaCls = 0;
-  _cwz = NULL; _cz = NULL; _dcz = NULL; _db = NULL; _dc = NULL;
-  _wb = NULL; _b = 0; _Nd = NULL; _pz = NULL; _z = NULL; _ct = NULL;
-  _bphi = NULL; _cphi = NULL;
-}
-
-MGLabLDA::~MGLabLDA() {
-  delTrainData();
-  delModelData();
-}
-
-void MGLabLDA::trainInit() {
-  unsigned int c, d, w, z;
-
-  _cwz = new unsigned int**[_C];
-  _cz = new unsigned int*[_C];
-
-  for(c = 0; c < _C; c++) {
-    _cz[c] = new unsigned int[_Kcls];
-    for(z = 0; z < _Kcls; z++) {
-      _cz[c][z] = 0;
-    }
-    _cwz[c] = new unsigned int*[_W];
-    for(w = 0; w < _W; w++) {
-      _cwz[c][w] = new unsigned int[_Kcls];
-      for(z = 0; z < _Kcls; z++) {
-        _cwz[c][w][z] = 0;
-      }
[…]

### 2018-11-13 src/MGLabLDA.hpp

@@ -1,97 +1,0 @@
-/* -*- C++ -*- */
-
-// Author: Alexander Kotov (kotov@wayne.edu), Wayne State University, 2014
-// $Id: MGLabLDA.hpp,v 1.3 2014/06/22 02:30:48 fn6418 Exp $
-
-
-#ifndef MGLABLDA_HPP_
-#define MGLABLDA_HPP_
-
-#include <ctime>
-#include <cmath>
-#include <cstdio>
-#include <cstdlib>
-#include "Utils.hpp"
-
-#define ALPHA_DEF 0.1
-#define BETA_DEF 0.01
-#define BETA_BKG_DEF 0.01
-#define BETA_CLS_DEF 0.01
-#define GAMMA_BKG_DEF 1
-#define GAMMA_CLS_DEF 1
-
-class MGLabLDA {
-public:
-  MGLabLDA(unsigned int numClsTops);
-  MGLabLDA(unsigned int numClsTops, float alpha, float betaBkg, float betaCls, float gammaBkg, float gammaCls);
-  virtual ~MGLabLDA();
-  // estimate the parameters of the model using Gibbs sampling
-  void runSampler(const ClsSampCol* docCol, unsigned int numCls, unsigned int vocSize, unsigned int numIters,
-                    unsigned int rndSeed=0, bool verbose=false);
-  // compute word-topic-class distributions
-  void computePhi();
-  void storeBkgTopic(FILE* fd, IDStrMap& id2term);
-  void storeClassTopics(FILE* fd, IDStrMap& id2class, IDStrMap& id2term);
-
-protected:
-  // sample word type and topic for word n of document m
-  void sampleHiddenVars(unsigned int sampID, unsigned int sampCls, unsigned int termID, unsigned int termInd, bool& isClsTop, unsigned int& topic);
-  // set model parameters to default values
-  void setDefaults();
-  // allocate memory and initialize training data
-  void trainInit();
-  // allocate memory and initialize topic-tag-word distributions
-  void topicsInit();
-  // deallocate memory for training data
-  void delTrainData();
-  // deallocate memory for word-tag-topic distributions
-  void delModelData();
-protected:
-  ////////////////////////////////////////////////////////////////////////////////////
-  // model parameters
-  ///////////////////////////////////////////////////////////////////////////////////
-  // number of classes
-  unsigned int _C;
-  // number of location-specific topics
-  unsigned int _Kcls;
-  // number of documents
[…]

### 2018-11-13 src/Splitter.cpp

@@ -1,292 +1,0 @@
-// $Id: Splitter.cpp,v 1.2 2014/02/04 08:47:14 fn6418 Exp $
-
-#include "Utils.hpp"
-
-#define NUM_SPLITS_DEF 10
-
-typedef struct SampCodes_ {
-  SampCodes_() : codeOne(""), codeTwo("")
-  { }
-
-  SampCodes_(std::string& codeOne, std::string& codeTwo) : codeOne(codeOne), codeTwo(codeTwo)
-  { }
-
-  std::string codeOne;
-  std::string codeTwo;
-} SampCodes;
-
-typedef std::map<std::string, SampCodes> SampCodeMap;
-typedef SampCodeMap::iterator SampCodeMapIt;
-typedef SampCodeMap::const_iterator cSampCodeMapIt;
-
-void usage(const char* progName) {
-  printf("Usage: %s arg_file\n", progName);
-  printf("Arguments:\n");
-  printf("  samp_file - input file with sample IDs\n");
-  printf("  train_dir - path to the directory with training splits\n");
-  printf("  test_dir - path to the directory with test splits\n");
-  printf("  num_splits - number of splits\n");
-  printf("  stats_file - file with statistics per split\n");
-  printf("  rnd_seed - seed for random shuffling of sample IDs\n");
-}
-
-SampCodeMap* loadSampCodeMap(const char *filePath) throw(Exception) {
-  bool inSample = false;
-  std::string curLine, curSampID;
-  std::string::size_type lpos, rpos;
-  SampCodes* sampCodes = NULL;
-  SampCodeMap* sampMap = NULL;
-
-  std::ifstream fs(filePath);
-
-  if(!fs.is_open()) {
-    throwException(std::string("parsing sample file ") + filePath, "can't open file for reading");
-  } else {
-    sampMap = new SampCodeMap();
-  }
-
-  while(getline(fs, curLine)) {
-    stripChars(curLine, SPACE_CHARS);
-
-    if(!curLine.length()) {
-      continue;
-    }
-
-    if(inSample) {
-      if((lpos = curLine.find("<ID>")) != std::string::npos) {
-        lpos += 4;
-        if((rpos = curLine.find("</ID>", lpos)) != std::string::npos) {
-          curSampID = curLine.substr(lpos, rpos-lpos);
-        } else {
[…]

### 2018-11-13 src/Types.hpp

@@ -1,139 +1,0 @@
-// $Id: Types.hpp,v 1.5 2014/06/22 02:30:48 fn6418 Exp $
-
-#ifndef TYPES_HPP_
-#define TYPES_HPP_
-
-#include <set>
-#include <string>
-#include <vector>
-#include <map>
-
-typedef std::vector<std::string> StrVec;
-typedef std::vector<StrVec> StrVecs;
-typedef std::set<std::string> StrSet;
-typedef StrSet::iterator StrSetIt;
-typedef StrSet::const_iterator cStrSetIt;
-
-typedef std::set<unsigned int> IDSet;
-typedef IDSet::iterator IDSetIt;
-typedef IDSet::const_iterator cIDSetIt;
-
-typedef std::vector<unsigned int> IDVec;
-
-typedef std::map<unsigned int, std::string> IDStrMap;
-typedef IDStrMap::iterator IDStrMapIt;
-typedef IDStrMap::const_iterator cIDStrMapIt;
-
-typedef std::map<std::string, unsigned int> StrIDMap, StrCntMap;
-typedef StrIDMap::iterator StrIDMapIt;
-typedef StrCntMap::iterator StrCntMapIt;
-typedef StrIDMap::const_iterator cStrIDMapIt;
-typedef StrCntMap::const_iterator cStrCntMapIt;
-
-typedef std::map<std::string, float> StrFloatMap;
-typedef StrFloatMap::iterator StrFloatMapIt;
-typedef StrFloatMap::iterator StrFloatMapIt;
-
-typedef std::map<unsigned int, float> MultDist;
-typedef MultDist::iterator MultDistIt;
-typedef MultDist::const_iterator cMultDistIt;
-
-typedef std::map<unsigned int, unsigned int> IDIDMap, CntMap;
-typedef IDIDMap::iterator IDIDMapIt;
-typedef IDIDMap::const_iterator cIDIDMapIt;
-typedef CntMap::iterator CntMapIt;
-typedef CntMap::const_iterator cCntMapIt;
-
-typedef struct TermIDWeight_ {
-  TermIDWeight_() : id(0), weight(0)
-  { }
-
-  TermIDWeight_(unsigned int id, float weight) : id(id), weight(weight)
-  { }
-
-  bool operator<(const TermIDWeight_& tidWgt) const {
-    return tidWgt.weight <= weight;
-  }
-
-  unsigned int id;
-  float weight;
-} TermIDWeight;
-
[…]

### 2018-11-13 src/Utils.cpp

@@ -1,636 +1,0 @@
-// $Id: Utils.cpp,v 1.5 2014/06/22 02:30:48 fn6418 Exp $
-
-#include "Utils.hpp"
-
-void toLower(std::string& token) {
-  for(std::string::size_type pos = 0; pos < token.size(); pos++) {
-    if(isupper(token[pos])) {
-      token[pos] = tolower(token[pos]);
-    }
-  }
-}
-
-void removeNonAlphaSpace(std::string& newToken, const std::string& oldToken) {
-  newToken = "";
-  for(std::string::const_iterator sit = oldToken.begin(); sit != oldToken.end(); sit++) {
-    if(isalpha(*sit) || isspace(*sit)) {
-      newToken += *sit;
-    } else {
-      newToken += " ";
-    }
-  }
-}
-
-void stripChars(std::string& token, const char* chars) {
-  rightStrip(token, chars);
-  leftStrip(token, chars);
-}
-
-void rightStrip(std::string& token, const char* chars) {
-  std::string::size_type pos = token.find_first_not_of(chars);
-  token.erase(0, pos);
-}
-
-void leftStrip(std::string& token, const char* chars) {
-  std::string::size_type pos = token.find_last_not_of(chars);
-  token.erase(pos + 1);
-}
-
-bool isAllAlpha(const std::string& token) {
-  for(std::string::const_iterator sit = token.begin(); sit != token.end(); sit++) {
-    if(!isalpha(*sit)) {
-      return false;
-    }
-  }
-
-  return true;
-}
-
-void stemWord(std::string& token, lemur::parse::PorterStemmer *stemmer) {
-  std::string::size_type pos;
-  char *buf = new char[token.size()+1];
-
-  if((pos = token.find("'")) != std::string::npos) {
-    token = token.substr(0, pos);
-  }
-
-  memset((void *) buf, 0, sizeof(char)*(token.size()+1));
-  token.copy(buf, token.size()+1);
-  token = stemmer->stemWord(buf);
-  delete[] buf;
-}
-
-void splitToks(const std::string& str, StrVec& toks, const char *seps, const std::string::size_type lpos,
-               const std::string::size_type rpos) {
-  std::string::size_type prev = lpos; // previous delimiter position
-  std::string::size_type cur = lpos; // current position of the delimiter
-  std::string::size_type rbound = rpos == std::string::npos ? str.length()-1 : rpos;
-
[…]

### 2018-11-13 src/Utils.hpp

@@ -1,43 +1,0 @@
-// $Id: Utils.hpp,v 1.4 2014/06/22 02:30:48 fn6418 Exp $
-
-#ifndef UTILS_HPP_
-#define UTILS_HPP_
-
-#include <fstream>
-#include <stdexcept>
-#include <utility>
-#include <math.h>
-#include <stdlib.h>
-#include "Types.hpp"
-#include "Exception.hpp"
-#include "PorterStemmer.hpp"
-
-#define SPACE_CHARS " \t\r\n"
-#define PUNCT_CHARS ",.;:!?()-'[]`{}\"$"
-
-void toLower(std::string& token);
-void removeNonAlphaSpace(std::string& newToken, const std::string& oldToken);
-void stripChars(std::string& token, const char *chars=SPACE_CHARS);
-void rightStrip(std::string& token, const char *chars=SPACE_CHARS);
-void leftStrip(std::string& token, const char *chars=SPACE_CHARS);
-bool isAllAlpha(const std::string& token);
-void stemWord(std::string& token, lemur::parse::PorterStemmer *stemmer);
-void splitToks(const std::string& str, StrVec& toks, const char *seps=" ", const std::string::size_type lpos = 0,
-               const std::string::size_type rpos = std::string::npos);
-void sortByWeight(IDVec& ord, const MultDist& topic);
-void parseFileName(const std::string& filePath, std::string& fileName);
-StrSet* loadOneColFile(const char *filePath) throw(Exception);
-StrSet* loadColFromFile(const char *filePath, unsigned int colNum) throw(Exception);
-StrStrMap* loadTwoColMapFile(const char *filePath) throw(Exception);
-StrStrMap* readConfFile(const char *filePath) throw(Exception);
-ClsSampCol* loadClsSampSubCol(const char *filePath, const StrSet& sampIDs, IDStrMap& id2class, IDStrMap& id2term) throw(Exception);
-SampCol* loadSampSubCol(const char *filePath, const StrSet& sampIDs) throw(Exception);
-LangMod* loadLangMod(const char *filePath) throw(Exception);
-StrLangModMap* loadClsLangMods(const char *filePath, StrSet** clsSet) throw(Exception);
-StrSet* loadTrainSampIDs(const char *filePath, StrFloatMap** clsStats) throw(Exception);
-void normClsStats(StrFloatMap* clsStats);
[…]

### 2018-11-13 text-segmentation-model/src/filemanager/FileManager.java

@@ -1,27 +1,0 @@
-package filemanager;
-
-public class FileManager {
-	public static void preprocessRawData(String inputDir, String outputDir, String codeMapFile) {
-		try {
-			Writer.prepareDataAsCodeAndText(inputDir, outputDir, codeMapFile);
-		} catch (Exception e) {
-			e.printStackTrace();
-		}
-	}
-	
-	public static void formatInstancesForLDA(String inputDataDirectory, String formatInputDataDirectory) {
-		Writer.removeFilesFromDirectory(formatInputDataDirectory);
-		Writer.createInputForLDAFormat(inputDataDirectory, formatInputDataDirectory);
-	}
-
-	public static void createGoldStandard(String arffTextFile, String  arffFileStr2WordVector, 
-			String  goldStandardFile, String wordProbabilitiesFile, boolean withTopicDistribution, String model) {
-		Writer.createGoldStandard(arffTextFile, arffFileStr2WordVector, goldStandardFile,
-				wordProbabilitiesFile, withTopicDistribution, model);
-	}
-	
-	public static void createArffFileWithTextAndCode(
-			String inputDataDirectory, String arffTextFile, String codeMapFile) {
-		Writer.createArffFileWithTextAndCode(inputDataDirectory, arffTextFile, codeMapFile);
-	}
-}

### 2018-11-13 text-segmentation-model/src/filemanager/Reader.java

@@ -1,346 +1,0 @@
-package filemanager;
-
-import java.io.BufferedReader;
-import java.io.File;
-import java.io.FileReader;
-import java.io.IOException;
-import java.util.HashMap;
-import java.util.Map;
-import java.util.TreeMap;
-import java.util.Vector;
-
-import main.Entry;
-import main.Helper;
-
-public class Reader {
-    
-	public static Vector<Entry> readProcessedData(String inputFolder, String delimiter) throws Exception {
-		
-		Vector<Entry> out = new Vector<Entry>();
-		
-		try {	
-			File rawDataFolder = new File(inputFolder);
-			File []rawFiles = rawDataFolder.listFiles();
-			
-			for (int i = 0; i < rawFiles.length; ++i) {
-				BufferedReader br = new BufferedReader(new FileReader(rawFiles[i]));			
-				String line;
-				while ((line = br.readLine()) != null) {
-					String[] lineData = line.split(delimiter);
-					String label = lineData[0].trim();
-					Entry entry = new Entry();
-					entry.label = label;
-					entry.text = line.substring(lineData[0].trim().length()).trim();
-					out.add(entry);
-				}
-				
-				br.close();	
-			}
-		}
-		catch(Exception e){
-			System.out.print("Error in reading file: "+ inputFolder);
-		}
-		
-		return out;
-	}
-	
-	public static Map<String, Entry> readCodeBook(String fileName) throws IOException {
-		
-		Map<String, Entry> codeBook = new HashMap<String, Entry>();  
-	    BufferedReader br = new BufferedReader(new FileReader(fileName));
-	    
-	    try {	    	
-	        
-	        String line = br.readLine();
-
-	        while (line != null) {
-	        	String[] oneLine = line.split("\t");
-	        	Entry code = new Entry();
-	            code.labelCode = oneLine[0].trim();
-	            code.label = oneLine[1].trim();
-	            code.text = oneLine[2].trim();
-	            codeBook.put(code.labelCode.trim(), code);
-	            line = br.readLine();	            
-	        }
-	        
-	        br.close();
-	        return codeBook;
-	        
-	    } catch(Exception e) {
-	        br.close();
-	        System.out.print("Error in reading file: "+ fileName);
-	        return codeBook;
-	    }
-	}
-	
[…]

### 2018-11-13 text-segmentation-model/src/filemanager/Writer.java

@@ -1,363 +1,0 @@
-package filemanager;
-
-import java.io.BufferedReader;
-import java.io.BufferedWriter;
-import java.io.File;
-import java.io.FileReader;
-import java.io.FileWriter;
-import java.io.IOException;
-import java.io.PrintWriter;
-import java.util.HashMap;
-import java.util.Map;
-import java.util.TreeMap;
-
-import main.Helper;
-
-public class Writer {
-	
-	public static boolean removeDirectory(File directory) {
-		if (directory == null)
-			return false;
-		if (!directory.exists())
-			return true;
-		if (!directory.isDirectory())
-			return false;
-		
-		System.out.println("Remove directory " + directory.getName());
-
-		String[] list = directory.list();
-
-		if (list != null) {
-			for (int i = 0; i < list.length; i++) {
-				File entry = new File(directory, list[i]);
-
-				if (entry.isDirectory())
-				{
-					if (!removeDirectory(entry))
-						return false;
-				}
-				else
-				{
-					if (!entry.delete())
-						return false;
-				}
-			}
-		}		
-		return directory.delete();
-	}
-	
-	public static void removeFilesFromDirectory(String directory) {
-		System.out.println("Remove files from " + directory);
-
-		File rawDataFolder = new File(directory);
-		File []rawFiles = rawDataFolder.listFiles();
-		
-		for (int i = 0; i < rawFiles.length; ++i) {		
-			rawFiles[i].deleteOnExit();
-		}
-	}
-		
-	public static void prepareDataAsCodeAndText(String inputDir, String outputDir, String classMapFile) throws Exception {		
-		PrintWriter classPw = new PrintWriter(classMapFile);	
-		TreeMap<String, String> classMap = new TreeMap<String, String>();
-		
-		File rawDataFolder = new File(inputDir);
-		File []rawFiles = rawDataFolder.listFiles();
-		
-		for (int i = 0; i < rawFiles.length; ++i) {			
-			PrintWriter pw = new PrintWriter(outputDir + "/" + rawFiles[i].getName());			
-			BufferedReader br = new BufferedReader(new FileReader(rawFiles[i]));
-			String line, text, code;
-			
-			while ((line = br.readLine()) != null) {	
-				if (line.trim().length() > 1) {
-					if (line.indexOf("}", 0) == -1) {
-						code = line.split("\\s+")[0].trim();
[…]

### 2018-11-13 text-segmentation-model/src/main/AppMain.java

@@ -1,60 +1,0 @@
-package main;
-
-import java.time.LocalDateTime;
-
-import filemanager.FileManager;
-import model.SegmentationModel;
-
-public class AppMain {
-
-	public static void main(String[] args) {
-		// start execution
-		System.out.println("Start execution at: " + LocalDateTime.now());
-
-		// configuration:
-		String rawDataDirectory = "../../../data-source/ecoaching-text-segmentation-data/input-data-raw";
-		String inputDataDirectory = "../../../data-source/ecoaching-text-segmentation-data/input-data";
-		String formatInputDataDirectory = "../../../data-source/ecoaching-text-segmentation-data/input-data-format";
-		String codeMapFile = "../config/stem/code-map-ecoaching.cfg";
-		String arffTextFile = "../../../data-source/ecoaching-text-segmentation-data/data/text_stem.arff";
-		String arffFileStr2WordVector = "../../../data-source/ecoaching-text-segmentation-data/data/stem_Str2WordVector.arff";
-		String goldStandardFile = "../../../data-source/ecoaching-text-segmentation-data/data/gold";
-		
-		// use the correct topic models 
-		String wordProbabilitiesFile = "../../../data-source/ecoaching-text-segmentation-data/topic/lta/stem/cls-lms.txt"; 
-		//String wordProbabilitiesFile = "../../../data-source/ecoaching-text-segmentation-data/topic/mg-lda/stem/cls-tops.txt"; 
-		//String wordProbabilitiesFile = "../../../data-source/ecoaching-text-segmentation-data/topic/lda/stem/cls-tops.txt"; 
-
-		// pre-process data to run LDA model for getting the features values
-		//FileManager.preprocessRawData(rawDataDirectory, inputDataDirectory, codeMapFile);
-		
-		// create one file for all documents after removing non-ascii characters
-		//FileManager.createArffFileWithTextAndCode(inputDataDirectory, arffTextFile, codeMapFile);
-		
-		// prepare data for Formatter of LDA model, [all instances in one file named with class]
-		//FileManager.formatInstancesForLDA(inputDataDirectory, formatInputDataDirectory);
-
-		// run C++ code from parent directory for LDA model; generate topic distribution
-		// !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
-		
-		// by using terminal run stringToVector from data directory; generate arffFileStr2WordVector file
[…]

### 2018-11-13 text-segmentation-model/src/main/Entry.java

@@ -1,17 +1,0 @@
-package main;
-
-public class Entry {	
-	public String labelCode;
-	public String label;
-	public String text;
-	
-	public Entry(){
-		
-	}
-		
-	public Entry(String _labelCode, String _label, String _text){
-		labelCode = _labelCode;
-		label = _label;
-		text = _text;
-	}
-}

### 2018-11-13 text-segmentation-model/src/main/Helper.java

@@ -1,94 +1,0 @@
-package main;
-
-import java.util.Comparator;
-import java.util.HashMap;
-import java.util.Map;
-import java.util.TreeMap;
-import java.util.Vector;
-
-public class Helper {
-	public static boolean isNumeric(String s) {  
-	    return s != null && s.matches("[-+]?\\d*\\.?\\d+");  
-	} 
-
-	public static String stemByPorterStemmer(String s) {
-		Stemmer st = new Stemmer();
-		StringBuilder sb = new StringBuilder();
-	    String[] words = s.split("\\s+");
-	    
-	    for (int i = 0; i < words.length; i++){
-			st.add(words[i]);
-			sb.append(st.stem());
-			sb.append(" ");
-			st.clear();
-	    }
-	    return sb.toString();
-	}
-	
-	public static Map<String, Integer> getDataDistribution(Vector<Entry> data){
-		Map<String, Integer> dist = new HashMap<String, Integer>();
-		for(Entry e: data){
-			if(dist.containsKey(e.label)){
-				dist.put(e.label, dist.get(e.label)+1);
-			}
-			else{
-				dist.put(e.label, 1);
-			}
-		}		
-		return dist;
-	}
-	
-	public static void printMap(Map<String, Integer> data){
-		TreeMap<String, Integer> tmap = new TreeMap<String, Integer>(data);
-		for(String key: tmap.keySet()){
-			System.out.println(key + "," + key);
-		}		
-	}
-	
-	public static boolean isSpecialChar(String data){
-		if (data.contains("exclam")){
-			return true;
-		}
-		else if (data.contains("fullstop")){
-			return true;
-		}
-		else if (data.contains("questionmark")){
-			return true;
-		}
-		else if (data.contains("comma")){
-			return true;
-		}
-		else if (data.contains("hyphen")){
-			return true;
-		}
-		else if (data.contains("clone")){
-			return true;
-		}
-		else if (data.contains("semiclone")){
-			return true;
-		}
-		else
-			return false;
-	}
-	
-	public static TreeMap<String, Integer> sortMapByValue(HashMap<String, Integer> map){
-		Comparator<String> comparator = new ValueComparator<String, Integer>(map);
-		TreeMap<String, Integer> result = new TreeMap<String, Integer>(comparator);
-		result.putAll(map);
-		return result;
-	}
-}
-
-//a comparator using generic type
[…]

### 2018-11-13 text-segmentation-model/src/main/Stemmer.java

@@ -1,632 +1,0 @@
-/*
- *  This program is free software: you can redistribute it and/or modify
-    it under the terms of the GNU General Public License as published by
-    the Free Software Foundation, either version 3 of the License, or
-    (at your option) any later version.
-
-    This program is distributed in the hope that it will be useful,
-    but WITHOUT ANY WARRANTY; without even the implied warranty of
-    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-    GNU General Public License for more details.
-
-    You should have received a copy of the GNU General Public License
-    along with this program.  If not, see <http://www.gnu.org/licenses/>. 
- */
-package main;
-
-/*
-
- Porter stemmer in Java. The original paper is in
-
- Porter, 1980, An algorithm for suffix stripping, Program, Vol. 14,
- no. 3, pp 130-137,
-
- See also http://www.tartarus.org/~martin/PorterStemmer
-
- History:
-
- Release 1
-
- Bug 1 (reported by Gonzalo Parra 16/10/99) fixed as marked below.
- The words 'aed', 'eed', 'oed' leave k at 'a' for step 3, and b[k-1]
- is then out outside the bounds of b.
-
- Release 2
-
- Similarly,
-
- Bug 2 (reported by Steve Dyrdahl 22/2/00) fixed as marked below.
- 'ion' by itself leaves j = -1 in the test for 'ion' in step 5, and
- b[j] is then outside the bounds of b.
-
- Release 3
-
- Considerably revised 4/9/00 in the light of many helpful suggestions
- from Brian Goetz of Quiotix Corporation (brian@quiotix.com).
-
- Release 4
-
- */
-
-
-
-/**
- * @author Jeff Byrd <sjbyrd2@crimson.ua.edu>
- * Stemmer, implementing the Porter Stemming Algorithm
- * 
- * The Stemmer class transforms a word into its root form. The input word is provided from the
- * add() methods. The stem() method will return the stem (as will toString() after stem() has been
- * called). The clear() method will wipe the Stemmer buffer and allow a new word to be input.
- * 
- * This version extends Martin Porter's original stemming algorithm by allowing capital letters
[…]

### 2018-11-13 text-segmentation-model/src/model/SegmentationModel.java

@@ -1,8 +1,0 @@
-package model;
-
-public class SegmentationModel {
-
-	public void runSVM(String arffFile) {
-		
-	}
-}

### 2018-11-13 plots/aupr.r
clean project directory

@@ -1,0 +1,78 @@
+
+#library
+library(ggplot2)
+library(reshape2)
+library(PRROC)
+
+par(mfrow=c(1,2))
+wd = '/home/mehedi/teana/projects/eCoaching-Text-Segmentation/plots/'
+
+# melt data
+x1 <- read.csv(paste(wd,"roc/base/pred_brnn.csv", sep=''))
+x2 <- read.csv(paste(wd,"roc/base/pred_mlp.csv", sep=''))
+x3 <- read.csv(paste(wd,"roc/base/pred_crf.csv", sep=''))
+x4 <- read.csv(paste(wd,"roc/base/pred_crnn.csv", sep=''))
+
+curve_obj_brnn <- pr.curve(scores.class0 = x1[x1$actual == 1,]$predict_prob, scores.class1 = x1[x1$actual == 0,]$predict_prob, curve = T)
+curve.points.brnn<-curve_obj_brnn$curve
+
+curve_obj_mlp <- pr.curve(scores.class0 = x2[x2$actual == 1,]$predict_prob, scores.class1 = x2[x2$actual == 0,]$predict_prob, curve = T)
+curve.points.mlp<-curve_obj_mlp$curve
+
+curve_obj_crf <- pr.curve(scores.class0 = x3[x3$actual == 1,]$predict_prob, scores.class1 = x3[x3$actual == 0,]$predict_prob, curve = T)
+curve.points.crf<-curve_obj_crf$curve
+
+curve_obj_crnn <- pr.curve(scores.class0 = x4[x4$actual == 1,]$predict_prob, scores.class1 = x4[x4$actual == 0,]$predict_prob, curve = T)
+curve.points.crnn<-curve_obj_crnn$curve
+
+# print AUPR values
+print(curve_obj_brnn)
+print(curve_obj_mlp)
+print(curve_obj_crf)
+print(curve_obj_crnn)
+
+# plot graph
+plot(curve.points.brnn[,1], curve.points.brnn[,2], col='red', xlim=c(0, 1), ylim=c(0, 1),  
+     type="l", ylab = "Precision", xlab = "Recall", cex.axis=1, cex.lab=1, lwd=1)
+lines(curve.points.mlp[,1], curve.points.mlp[,2], col='blue', lty=1, lwd=1)
+lines(curve.points.crf[,1], curve.points.crf[,2], col='green', lty=1, lwd=1)
[…]

### 2018-11-14 paper-amia-summit-2019/manuscript_marked.tex
duplicate manuscript for highlighing the changes

--- 2018-08-10 paper-amia-summit-2018/manuscript.tex
@@
 \usepackage{color}
+\usepackage{soul}
 \usepackage{multirow}
@@
 MI sessions are generally aimed at eliciting ``change talk'', or statements of intrinsic motivation about patients' own desire, ability, reasons, need for and commitment to behavior change, which have been established by previous research\cite{apodaca2009mechanisms} as a reliable mediator of health behavior change.
-However, communication science approaches to understanding the efficacy of MI are inherently limited by traditional qualitative coding methods.
+\hl{However, communication science approaches to understanding the efficacy of MI are inherently limited by traditional qualitative coding methods}.
 
@@
 The goal of this study is to assess the effectiveness of deep learning methods for the task of automated segmentation of e-Coaching emails into textual fragments corresponding to individual behaviors, which is the first step of qualitative analysis of this type of clinical communications.
-Specifically, for this study, we utilized the data from MENU GenY (Making Effective Nutrition Choices for Generation Y) \cite{alexander2017motivations}, a recent technology-based public health intervention that relies on personalized e-Coaching to encourage increased fruit and vegetable intake among young adults, aged 21--30.
+\hl {Specifically, for this study, we utilized the data from MENU GenY (Making Effective Nutrition Choices for Generation Y) }\cite{alexander2017motivations}, a recent technology-based public health intervention that relies on personalized e-Coaching to encourage increased fruit and vegetable intake among young adults, aged 21--30.
 The goal of MENU GenY was to develop a better coding scheme for e-Coaching communications aimed at improving GenY eating habits.
@@
 \subsection*{\textit{Data collection}}
-The experimental dataset for this work was constructed from 49 e-Coaching sessions, which include a total of 3,138 segmented and annotated MI behaviors.
+\hl{The experimental dataset for this work was constructed from 49 e-Coaching sessions, which include a total of 3,138 segmented and annotated MI behaviors.
 Each session represents an MI intervention delivered via email.
 We consider segmentation as a binary classification problem, in which each word or punctuation mark is annotated with a label of 1 or 0 to indicate whether it precedes a new segment or not.
-In total, the dataset consists of 95,777 words and 7,140 punctuation marks and includes 3,138 ``new segment'' and 99,779 ``same segment'' instances.
+In total, the dataset consists of 95,777 words and 7,140 punctuation marks and includes 3,138 ``new segment'' and 99,779 ``same segment'' instances}.
 In this study, we experimented with traditional machine learning methods, such as Conditional Random Field (CRF)\cite{lafferty2001conditional} and deep learning methods, such as Multi-Layer Perceptron (MLP),\cite{rumelhart1986learning} Bidirectional Recurrent Neural Network (bidirectional RNN or BRNN)\cite{schuster1997bidirectional} and Convolutional Recurrent Neural Network (CRNN).\cite{treviso2017sentence} For an MLP model, training and testing instances were created based on a sliding window of $2n$ words or punctuation marks over each position in a given input sequence, such that each instance consists of a pair of the next $n$ words or punctuation marks and prior $n$ words or punctuation marks, including the current word or punctuation mark.
@@
 
-\textbf{Convolutional Recurrent Neural Network}: CRNN is a deep neural network architecture,\cite{treviso2017sentence} shown in Figure~\ref{fig:crnn}, which consists of 5 layers: 1) input layer 2) embedding layer 3) convolution layer with max pooling 4) BRNN layer 5) fully connected layer with dropout and sigmoid output.
+\textbf{Convolutional Recurrent Neural Network}: CRNN is a deep neural network architecture,\cite{treviso2017sentence} shown in Figure~\ref{fig:crnn}, \hl {which consists of 5 layers: 1) input layer 2) embedding layer 3) convolution layer with max pooling 4) BRNN layer 5) fully connected layer with dropout and sigmoid output.
 E-coaching email exchanges are represented as a sequence of $m$ words and punctuations, which are fed into the input layer to produce a $m \times n_e$ matrix after fetching the pre-trained word vectors.
@@
 To prevent this, we utilized dropout by randomly disregarding 50\% of the connections to the fully connected layer.
-After that, logistic sigmoid is applied to the output layer to provide the probability of whether the word or punctuation precedes a new segment or not.
+After that, logistic sigmoid is applied to the output layer to provide the probability of whether the word or punctuation precedes a new segment or not}.
 We experimentally determined the optimal parameters and found that the best performance is achieved with 5-fold cross-validation when filter length in convolutional layer is 7, number of filters is 100, max-pool size is 3, ReLU is used as a convolutional layer activation function, RNN layer activation is hyperbolic tangent and the number of recurrent units in RNN is 200.
-Adam\cite{kingma2014adam} with 50 epochs, the batch size of 32 and learning rate of 0.001 was used for optimization and the early stopping strategy was applied.\footnote{source code will be available at https://github.com/teanalab/xxxxx upon acceptance of this paper for publication}
+Adam\cite{kingma2014adam} with 50 epochs, the batch size of 32 and learning rate of 0.001 was used for optimization and the early stopping strategy was applied.\footnote{\hl {source code will be available at https://github.com/teanalab/xxxxx upon acceptance of this paper for publication}}
 
@@
 The results are reported based on 5 fold cross-validation and weighted macro-averaging over the folds.
-We also estimate the area under the receiving operating characteristics (ROC) curve\cite{kumar2011receiver} (AUC) metric due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets.
+\hl{We also estimate the area under the receiving operating characteristics (ROC) curve} \cite{kumar2011receiver} \hl{(AUC) metric due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets}.
 \cite{hu2015kernelized}
@@
     \caption{\textbf{Performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and corpus word embeddings with GloVe and word2vec models (left).
-MLP model on e-Coaching email segmentation by varying the sliding window size $2n$ (right).}}
+MLP model on e-Coaching email segmentation by varying the sliding window size 2n (right).}}
     \label{fig:embedding-dimension-mlp}
@@
 
-Table~\ref{tab:result_roc} shows the impact of individual features as well as their combination on the segmentation task.
+Table~\ref{tab:result_roc} \hl {shows the impact of individual features as well as their combination on the segmentation task.
 Although CRF provides the second highest classification results, it shows best AUC values, achieving AUC 0.966 when only lexical features are used and AUC 0.994 when a combination of lexical, punctuation and POS features are used.
@@
 CRF and MLP achieve their highest AUCs when all features are used together.
-On the other hand, BRNN and CRNN demonstrate the highest AUCs when word embeddings are combined with punctuation features.
+On the other hand, BRNN and CRNN demonstrate the highest AUCs when word embeddings are combined with punctuation features}.
 \\
@@
 Especially, punctuation features have the higher individual impact on model performance compared to POS features.
-In almost all cases, every method performs better, when word embeddings are used in conjunction with punctuation and POS features.
+\hl {In almost all cases, every method performs better, when word embeddings are used in conjunction with punctuation and POS features}.
 This results also indicate that segmentation performance might be improved by adding additional relevant features.
@@
 \bibliographystyle{vancouver}
-\bibliography{manuscript}
+\bibliography{manuscript_marked}
 

### 2018-11-14 paper-amia-summit-2019/manuscript.tex
added new table for dataset

--- 2018-08-10 paper-amia-summit-2018/manuscript.tex
@@
 \subsection*{\textit{Data collection}}
-The experimental dataset for this work was constructed from 49 e-Coaching sessions, which include a total of 3,138 segmented and annotated MI behaviors.
+The experimental dataset for this work was constructed from 49 e-Coaching sessions, which include a total of 3,138 segmented and annotated MI behaviors \hl{shown in Table}~\ref{tab:datastat}.
 Each session represents an MI intervention delivered via email.
@@
 In the gold standard, words or punctuations within the same segment were assigned the label of 0 and the last word or punctuation mark of a segment were assigned the label of 1.
+
+\begin{table}[ht]
+\centering
+\caption{\textbf{\hl{Summary of the experimental dataset}}}
+\label{tab:datastat}
+ \begin{tabular}{|l|l|l|l|l|l|l|l|l|l|}
+  \hline
+   \multirow{2}{*}{\textbf{Sessions}} & \multirow{2}{*}{\textbf{Instances}} & \multicolumn{2}{|c|}{\textbf{Class Segments}} & \multicolumn{2}{|c|}{\textbf{Tokens}} & \multicolumn{2}{|c|}{\textbf{Emails}} & \multicolumn{2}{|c|}{\textbf{Annotation}} \\\cline{3-10}
+   &  & \textbf{new}  & \textbf{same} & \textbf{word} & \textbf{punctuation}  & \textbf{patient} & \textbf{provider} & \textbf{method}  & \textbf{codes} \\ \hline    
+ 49 & 102,917 & 3,138 & 99,779 & 95,777 & 7,140 & 281 & 330 & MYSCOPE & 115 \\ \hline
+  \end{tabular}
+\end{table}
 
@@
 We experimentally determined the optimal parameters and found that the best performance is achieved with 5-fold cross-validation when filter length in convolutional layer is 7, number of filters is 100, max-pool size is 3, ReLU is used as a convolutional layer activation function, RNN layer activation is hyperbolic tangent and the number of recurrent units in RNN is 200.
-Adam\cite{kingma2014adam} with 50 epochs, the batch size of 32 and learning rate of 0.001 was used for optimization and the early stopping strategy was applied.\footnote{source code will be available at https://github.com/teanalab/xxxxx upon acceptance of this paper for publication}
+Adam\cite{kingma2014adam} with 50 epochs, the batch size of 32 and learning rate of 0.001 was used for optimization and the early stopping strategy was applied.\footnote{\hl{https://github.com/teanalab/eCoaching-Text-Segmentation}}
 

### 2018-11-15 paper-amia-summit-2019/manuscript.tex
update area under precision-recall values and update manuscript

@@
 The goal of this study is to assess the effectiveness of deep learning methods for the task of automated segmentation of e-Coaching emails into textual fragments corresponding to individual behaviors, which is the first step of qualitative analysis of this type of clinical communications.
-Specifically, for this study, we utilized the data from MENU GenY (Making Effective Nutrition Choices for Generation Y) \cite{alexander2017motivations}, a recent technology-based public health intervention that relies on personalized e-Coaching to encourage increased fruit and vegetable intake among young adults, aged 21--30.
+Specifically, for this study, we utilized the data from MENU GenY (Making Effective Nutrition Choices for Generation Y) \cite{alexander2017motivations}, a recent \hl{email-based} public health intervention that relies on personalized e-Coaching to encourage increased fruit and vegetable intake among young adults, aged 21--30.
 The goal of MENU GenY was to develop a better coding scheme for e-Coaching communications aimed at improving GenY eating habits.
@@
 \subsection*{\textit{Data collection}}
-The experimental dataset for this work was constructed from 49 e-Coaching sessions, which include a total of 3,138 segmented and annotated MI behaviors \hl{shown in Table}~\ref{tab:datastat}.
-Each session represents an MI intervention delivered via email.
-We consider segmentation as a binary classification problem, in which each word or punctuation mark is annotated with a label of 1 or 0 to indicate whether it precedes a new segment or not.
-In total, the dataset consists of 95,777 words and 7,140 punctuation marks and includes 3,138 ``new segment'' and 99,779 ``same segment'' instances.
-In this study, we experimented with traditional machine learning methods, such as Conditional Random Field (CRF)\cite{lafferty2001conditional} and deep learning methods, such as Multi-Layer Perceptron (MLP),\cite{rumelhart1986learning} Bidirectional Recurrent Neural Network (bidirectional RNN or BRNN)\cite{schuster1997bidirectional} and Convolutional Recurrent Neural Network (CRNN).\cite{treviso2017sentence} For an MLP model, training and testing instances were created based on a sliding window of $2n$ words or punctuation marks over each position in a given input sequence, such that each instance consists of a pair of the next $n$ words or punctuation marks and prior $n$ words or punctuation marks, including the current word or punctuation mark.
+The experimental dataset for this work was constructed from 49 e-Coaching sessions, which include \hl{330 and 281 emails from ecoaches and patients, respectively }(Table~\ref{tab:datastat}).
+\hl{Each session represents an MI intervention delivered via email.
+Emails were segmented into 3,138 text fragments and annotated using 115 MYSCOPE}~\cite{carcone2013provider} \hl{behavior codes.
+We consider email segmentation as a binary classification problem, in which each word or punctuation mark is annotated with a label of 1 or 0 to indicate whether it precedes a new segment or not.
+In total, the dataset consists of 95,777 words and 7,140 punctuation marks and includes 3,138 ``new segment'' and 99,779 ``same segment'' instances.} In this study, we experimented with traditional machine learning methods, such as Conditional Random Field (CRF)\cite{lafferty2001conditional} and deep learning methods, such as Multi-Layer Perceptron (MLP),\cite{rumelhart1986learning} Bidirectional Recurrent Neural Network (bidirectional RNN or BRNN)\cite{schuster1997bidirectional} and Convolutional Recurrent Neural Network (CRNN).\cite{treviso2017sentence} For an MLP model, training and testing instances were created based on a sliding window of $2n$ words or punctuation marks over each position in a given input sequence, such that each instance consists of a pair of the next $n$ words or punctuation marks and prior $n$ words or punctuation marks, including the current word or punctuation mark.
 Each sample is classified into either a ``new segment'' or a ``same segment'' based on whether there should be a segment break after the current word or not.
@@
 In the gold standard, words or punctuations within the same segment were assigned the label of 0 and the last word or punctuation mark of a segment were assigned the label of 1.
+\\
 
@@
 \subsection*{\textit{Features}}
-
 We utilized three types of features in conjunction with CRF, MLP, BRNN and CRNN methods: word embeddings or lexical features, punctuation and POS features.
@@
 In our experiments, the following features were utilized in conjunction with CRF models: i) current word or punctuation ii) next and previous 3 words or punctuations iii) binary feature indicating whether a word or punctuation is a special character (';', '?', '.', ',', '!', ':', etc.) or not iv) binary feature indicating whether a word is a title word or not (e.g.
-``The'' is a title word but ``the'' is not) v) POS tags
+``The'' is a title word but ``the'' is not) v) POS tags.
+
+\textbf{Multi-Layer Perceptron}: MLP is a feed-forward artificial neural network, which maps an input onto one or several outputs.\cite{rumelhart1986learning} Figure~\ref{fig:mlp} shows a multilayer perceptron with a single hidden layer.
+In MLP networks, there is no cycle or loop and information moves forward only, from the input nodes through the hidden nodes and to the output nodes.
+This study utilized MLP with a nonlinear activation function (rectified linear unit) and one hidden layer consisting of 128 hidden units.
+In order to prevent over-fitting, we utilized dropout in fully connected layers by randomly hiding neurons during training.
+\cite{srivastava2014dropout} Dropout is also applied to the fully connected layers in BRNN and CRNN models.
 
@@
 \end{figure}
-
-\textbf{Multi-Layer Perceptron}: MLP is a feed-forward artificial neural network, which maps an input onto one or several outputs.\cite{rumelhart1986learning} Figure~\ref{fig:mlp} shows a multilayer perceptron with a single hidden layer.
-In MLP networks, there is no cycle or loop and information moves forward only, from the input nodes through the hidden nodes and to the output nodes.
-This study utilized MLP with a nonlinear activation function (rectified linear unit) and one hidden layer consisting of 128 hidden units.
-In order to prevent over-fitting, we utilized dropout in fully connected layers by randomly hiding neurons during training.
-\cite{srivastava2014dropout} Dropout is also applied to the fully connected layers in BRNN and CRNN models.
 
@@
 The results are reported based on 5 fold cross-validation and weighted macro-averaging over the folds.
-We also estimate the area under the receiving operating characteristics (ROC) curve\cite{kumar2011receiver} (AUC) metric due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets.
-\cite{hu2015kernelized}
+We also estimate the area under \hl{the precision-recall curve (AUPR) metric due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets.} \cite{davis2006relationship}
 
@@
 
-As follows from Table~\ref{tab:result_base}, CRNN outperforms all other models in terms of recall and F1-measure achieving 0.797 recall with 0.785 F1-measure for new segment detection.
-CRNN also shows superior performance in all performance metrics for weighted average over ``new segment'' and ``same segment'' classes.
-BRNN demonstrates the lowest performance among all models in terms of precision and F1-Measure.
-On the other hand, MLP has the highest precision of 0.836 when word embeddings or lexical features are used to identify ``new segment''.
-CRF achieves 0.733 F1-Measure in new segment class and 0.984 F1-Measure overall, which corresponds to the second highest performance in identifying ``new segment'' as well as a weighted average over both classes.
-Experimental results indicate that performance of all classifiers as a weighted average over both classes is remarkably higher compared to ``new segment'' class, which is expected since 96.95\% of instances belong to the ``same segment'' class and 99.3\% of them are correctly classified.
-For example, as a weighted average over ``new segment'' and ``same segment'' classes, CRNN achieves 27.23\%, 23.71\% and 25.61\% higher precision, recall and F1-measure, respectively, compared to the ``new segment'' detection.
-
 \begin{table}[ht]
@@
 
-Table~\ref{tab:result_weighted_avg} summarizes the results of all models for segmentation of e-Coaching emails when word embeddings or lexical features are used in combination with punctuation and POS features.
-Similar to results in Tables~\ref{tab:result_base}, CRNN demonstrates the highest performance among all methods achieving 0.864 recall with 0.825 F1-Measure for ``new segment'' detection and 0.990 precision with 0.989 recall and F1-Measure overall.
-BRNN and CRF show the lowest and second highest performance, respectively, for email segmentation among all methods.
-We observed that classification performance significantly improved for ``new segment'' class when word embeddings or lexical features are used in combination with punctuation and POS features.
-Precision increases by 3.96\%, -2.27\%, 12.71\% and 1.81\%; recall increases by 11.72\%, 19.73\%, 20.59\% and 8.41\%; and F1-measure increases by 8.05\%, 9.51\%, 16.22\% and 5.1\% for CRF, MLP, BRNN and CRNN methods, respectively, in new segment detection when all features are utilized together.
-Similarly, precision increases by 0.51\%, 0.41\%, 0.82\% and 0.41\%; recall increases by 0.41\%, 0.41\%, 0.72\% and 0.3\%; and F1-measure increases by 0.41\%, 0.41\%, 0.82\% and 0.3\% for CRF, MLP, BRNN and CRNN methods, respectively, in weighted average over ``new segment'' and ``same segment'' classes when word embeddings or lexical features are used in combination with punctuation and POS features.\\
+As follows from Table~\ref{tab:result_base}, CRNN outperforms all other models in terms of recall and F1-measure achieving 0.797 recall with 0.785 F1-measure for new segment detection.
+CRNN also shows superior performance in all performance metrics for weighted average over ``new segment'' and ``same segment'' classes.
+BRNN demonstrates the lowest performance among all models in terms of precision and F1-Measure.
+On the other hand, MLP has the highest precision of 0.836 when word embeddings or lexical features are used to identify ``new segment''.
+CRF achieves 0.733 F1-Measure in new segment class and 0.984 F1-Measure overall, which corresponds to the second highest performance in identifying ``new segment'' as well as a weighted average over both classes.
+Experimental results indicate that performance of all classifiers as a weighted average over both classes is remarkably higher compared to ``new segment'' class, which is expected since 96.95\% of instances belong to the ``same segment'' class and 99.3\% of them are correctly classified.
+For example, as a weighted average over ``new segment'' and ``same segment'' classes, CRNN achieves 27.23\%, 23.71\% and 25.61\% higher precision, recall and F1-measure, respectively, compared to the ``new segment'' detection.
+\\
 
@@
 
-Table~\ref{tab:result_roc} shows the impact of individual features as well as their combination on the segmentation task.
-Although CRF provides the second highest classification results, it shows best AUC values, achieving AUC 0.966 when only lexical features are used and AUC 0.994 when a combination of lexical, punctuation and POS features are used.
-Influence of the punctuation and POS features is also consistent in AUC values, which increase by 2.9\%, 4.31\%, 6.63\% and 2.18\% for CRF, MLP, BRNN and CRNN methods, respectively, when all features are used together.
-Individually, POS and punctuation features also improve the performance of all classifiers except CRNN when word embeddings are used with POS features.
-CRF and MLP achieve their highest AUCs when all features are used together.
-On the other hand, BRNN and CRNN demonstrate the highest AUCs when word embeddings are combined with punctuation features.
-\\
+Table~\ref{tab:result_weighted_avg} summarizes the results of all models for segmentation of e-Coaching emails when word embeddings or lexical features are used in combination with punctuation and POS features.
+Similar to results in Tables~\ref{tab:result_base}, CRNN demonstrates the highest performance among all methods achieving 0.864 recall with 0.825 F1-Measure for ``new segment'' detection and 0.990 precision with 0.989 recall and F1-Measure overall.
+BRNN and CRF show the lowest and second highest performance, respectively, for email segmentation among all methods.
+We observed that classification performance significantly improved for ``new segment'' class when word embeddings or lexical features are used in combination with punctuation and POS features.
+Precision increases by 3.96\%, -2.27\%, 12.71\% and 1.81\%; recall increases by 11.72\%, 19.73\%, 20.59\% and 8.41\%; and F1-measure increases by 8.05\%, 9.51\%, 16.22\% and 5.1\% for CRF, MLP, BRNN and CRNN methods, respectively, in new segment detection when all features are utilized together.
+Similarly, precision increases by 0.51\%, 0.41\%, 0.82\% and 0.41\%; recall increases by 0.41\%, 0.41\%, 0.72\% and 0.3\%; and F1-measure increases by 0.41\%, 0.41\%, 0.82\% and 0.3\% for CRF, MLP, BRNN and CRNN methods, respectively, in weighted average over ``new segment'' and ``same segment'' classes when word embeddings or lexical features are used in combination with punctuation and POS features.\\
 
@@
 \centering
-\caption{\textbf{AUC values of all classifiers demonstrating the impact of word embeddings, punctuation and POS features on e-Coaching email segmentation.
-Highest AUC value for each feature set across all models is highlighted in boldface.}}
-\label{tab:result_roc}
+\caption{\textbf{\hl{Area under the precision-recall curve (AUPR) values of all classifiers demonstrating the impact of word embeddings, punctuation and POS features on e-Coaching email segmentation.
+Highest AUPR value for each feature set across all models is highlighted in boldface.}}}
+\label{tab:result_aupr}
  \begin{tabular}{|l|l|l|l|l|}
@@
 \textbf{Features} & \textbf{CRF} & \textbf{MLP}  & \textbf{BRNN} & \textbf{CRNN} \\ \hline      
- word embeddings only & \textbf{0.966} & 0.951 & 0.920 & 0.965 \\ \hline
- word embeddings + POS & \textbf{0.973} & 0.953 & 0.920 & 0.962 \\ \hline
- word embeddings + punctuation & \textbf{0.993} & 0.989 & 0.982 & 0.988 \\ \hline
- all features & \textbf{0.994} & 0.992 & 0.981 & 0.986 \\ \hline
+ word embeddings only & 0.780 & 0.736 & 0.655 & \textbf{0.818} \\ \hline
+ word embeddings + POS & 0.797 & 0.746 & 0.647 & \textbf{0.798} \\ \hline
+ word embeddings + punctuation & \textbf{0.876} & 0.835 & 0.774 & 0.874 \\ \hline
+ all features & \textbf{0.877} & 0.842 & 0.770 & 0.867 \\ \hline
   \end{tabular}
 \end{table}
+
+Table~\ref{tab:result_aupr} shows the impact of individual features as well as their combination on the segmentation task.
+\hl{Influence of the punctuation and POS features is also consistent in AUPR values, which increase by 12.44\%, 14.4\%, 17.56\% and 5.99\% for CRF, MLP, BRNN and CRNN methods, respectively, when all features are used together.
+Individually, POS and punctuation features also improve the performance of all classifiers except BRNN and CRNN when POS features are used.
+CRF achieved the highest AUPR when all features are used together.
+On the other hand, CRNN demonstrates the highest AUPR when word embeddings are used individually as well as combined with POS features.}
 
@@
 Especially, punctuation features have the higher individual impact on model performance compared to POS features.
-In almost all cases, every method performs better, when word embeddings are used in conjunction with punctuation and POS features.
-This results also indicate that segmentation performance might be improved by adding additional relevant features.
+\hl{In all cases, CRF and MLP methods performed better, when word embeddings are used in conjunction with punctuation and POS features.
+Punctuation features improved the performance of BRNN and CRNN while POS features lowered their AUPR values.
+We believe that the BRNN and CRNN performed poorly with POS features because POS tagging is a supervised learning solution that uses features like the previous word and next word.
+Since we already considered neighbor words by utilizing bi-directional RNN, it failed to achieve good results with redundant information.
+We also observed that MLP achieved highest precision which may be related to the fact that MLP poorly learned new segment class and misclassiified segment words as non-segment words.}
 

### 2018-11-15 paper-amia-summit-2019/manuscript.tex
update manuscript

@@
 \hl{Each session represents an MI intervention delivered via email.
-Emails were segmented into 3,138 text fragments and annotated using 115 MYSCOPE}~\cite{carcone2013provider} \hl{behavior codes.
+Emails were segmented into 3,138 text fragments and annotated using 115 distinct MYSCOPE}~\cite{carcone2013provider} \hl{behavior codes.
 We consider email segmentation as a binary classification problem, in which each word or punctuation mark is annotated with a label of 1 or 0 to indicate whether it precedes a new segment or not.
@@
 \centering
-\caption{\textbf{\hl{Summary of the experimental dataset}}}
+\caption{\textbf{\hl{Summary of experimental dataset}}}
 \label{tab:datastat}
@@
 Individually, POS and punctuation features also improve the performance of all classifiers except BRNN and CRNN when POS features are used.
-CRF achieved the highest AUPR when all features are used together.
-On the other hand, CRNN demonstrates the highest AUPR when word embeddings are used individually as well as combined with POS features.}
+CRF achieved the highest AUPR when punctuation or all features are used together.
+On the other hand, CRNN demonstrates the highest AUPR when word embeddings are used individually or combined with POS features.}
 
@@
 Since we already considered neighbor words by utilizing bi-directional RNN, it failed to achieve good results with redundant information.
-We also observed that MLP achieved highest precision which may be related to the fact that MLP poorly learned new segment class and misclassiified segment words as non-segment words.}
+We observed that MLP achieved highest precision which may be related to the fact that MLP poorly learned ``new segment'' and misclassiified new segment words as same segment in 30\%-40\% of the time.}
 

### 2018-11-15 paper-amia-summit-2019/manuscript.tex
update CRNN

@@
 Emails were segmented into 3,138 text fragments and annotated using 115 distinct MYSCOPE}~\cite{carcone2013provider} \hl{behavior codes.
-We consider email segmentation as a binary classification problem, in which each word or punctuation mark is annotated with a label of 1 or 0 to indicate whether it precedes a new segment or not.
+We consider email segmentation as a binary classification problem, in which each word or punctuation mark is annotated with a class label ``new segment'' or ``same segment'' to indicate whether it precedes a new segment or not.
 In total, the dataset consists of 95,777 words and 7,140 punctuation marks and includes 3,138 ``new segment'' and 99,779 ``same segment'' instances.} In this study, we experimented with traditional machine learning methods, such as Conditional Random Field (CRF)\cite{lafferty2001conditional} and deep learning methods, such as Multi-Layer Perceptron (MLP),\cite{rumelhart1986learning} Bidirectional Recurrent Neural Network (bidirectional RNN or BRNN)\cite{schuster1997bidirectional} and Convolutional Recurrent Neural Network (CRNN).\cite{treviso2017sentence} For an MLP model, training and testing instances were created based on a sliding window of $2n$ words or punctuation marks over each position in a given input sequence, such that each instance consists of a pair of the next $n$ words or punctuation marks and prior $n$ words or punctuation marks, including the current word or punctuation mark.
@@
 E-coaching email exchanges are represented as a sequence of $m$ words and punctuations, which are fed into the input layer to produce a $m \times n_e$ matrix after fetching the pre-trained word vectors.
-When POS tags are used in combination with word embeddings, 10-dimensional POS vectors are concatenated with 300-dimensional word vectors to obtain new vectors $n_e = [n_w;n_p]$ of size 310.
+\hl{This matrix represents embedding output, contain sufficient morpho-syntactic information for segmentation of email exchanges}.
+When POS tags are used in combination with word embeddings, \hl{POS tags are embedded with 10-dimensional POS vectors and concatenated with 300-dimensional word vectors to obtain new embedding} vectors $n_e = [n_w;n_p]$ of size 310.
 The primary purpose of convolution is to extract new features depending on the neighboring words.
@@
 To prevent this, we utilized dropout by randomly disregarding 50\% of the connections to the fully connected layer.
-After that, logistic sigmoid is applied to the output layer to provide the probability of whether the word or punctuation precedes a new segment or not.
+After that, logistic sigmoid \hl{outputs the probability of each class}.
 We experimentally determined the optimal parameters and found that the best performance is achieved with 5-fold cross-validation when filter length in convolutional layer is 7, number of filters is 100, max-pool size is 3, ReLU is used as a convolutional layer activation function, RNN layer activation is hyperbolic tangent and the number of recurrent units in RNN is 200.
-Adam\cite{kingma2014adam} with 50 epochs, the batch size of 32 and learning rate of 0.001 was used for optimization and the early stopping strategy was applied.\footnote{\hl{https://github.com/teanalab/eCoaching-Text-Segmentation}}
+Adam\cite{kingma2014adam} with 50 epochs, the batch size of 32 and learning rate of 0.001 was used for optimization and the early stopping strategy was applied.\footnote{\hl{source codes are available at https://github.com/teanalab/eCoaching-Text-Segmentation}}
 

### 2018-11-16 paper-amia-summit-2019/manuscript.tex
update response to reviewers

@@
 We report standard metrics of precision, recall and F1-measure to evaluate the performance of the classifiers.\cite{aas1999text} Accuracy is not reported as a performance metric because it is highly sensitive to the prior class probabilities and does not fully describe the actual difficulty of the decision problem, when highly unbalanced datasets are involved.
-The results are reported based on 5 fold cross-validation and weighted macro-averaging over the folds.
+The results are reported based on 5 fold cross-validation and weighted macro-averaging over the folds, \hl{in which each fold was used as a test set and remaining 4 folds was utilized as a train set}.
 We also estimate the area under \hl{the precision-recall curve (AUPR) metric due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets.} \cite{davis2006relationship}

### 2018-11-16 paper-amia-summit-2019/manuscript.tex
update response to reviewers and manuscript

@@
 \centering
-\caption{\textbf{\hl{Summary of experimental dataset}}}
+\caption{\textbf{\hl{Summary of the experimental dataset}}}
 \label{tab:datastat}
@@
 Punctuation features improved the performance of BRNN and CRNN while POS features lowered their AUPR values.
-We believe that the BRNN and CRNN performed poorly with POS features because POS tagging is a supervised learning solution that uses features like the previous word and next word.
+We believe that the BRNN and CRNN performed poorly with POS features because POS tagging is a supervised learning solution that uses features like the previous and next word.
 Since we already considered neighbor words by utilizing bi-directional RNN, it failed to achieve good results with redundant information.
-We observed that MLP achieved highest precision which may be related to the fact that MLP poorly learned ``new segment'' and misclassiified new segment words as same segment in 30\%-40\% of the time.}
+We observed that MLP achieved the highest precision which may be related to the fact that MLP poorly learned ``new segment'' and misclassified new segment words to same segments in 30\%-40\% of the time.}
 

### 2018-11-18 paper-amia-summit-2019/manuscript.tex
update manuscript and response to reviewers comments

@@
 In the case of CRF, BRNN and CRNN models, an email was taken as an input sequence, POS tags and word embeddings of each word or punctuation mark were used as input and binary labels (1 or 0) corresponding to ``new segment'' and ``same segment'' classification decisions were considered as the model output of a model.
-In the gold standard, words or punctuations within the same segment were assigned the label of 0 and the last word or punctuation mark of a segment were assigned the label of 1.
+In the gold standard, words or punctuations within the same segment were assigned the label of 0 and the last word or punctuation mark of a segment were assigned the label of 1 (an example is given at $2^{nd}$ row in Table~\ref{tab:datastat}).
 \\
@@
    \multirow{2}{*}{\textbf{Sessions}} & \multirow{2}{*}{\textbf{Instances}} & \multicolumn{2}{|c|}{\textbf{Class Segments}} & \multicolumn{2}{|c|}{\textbf{Tokens}} & \multicolumn{2}{|c|}{\textbf{Emails}} & \multicolumn{2}{|c|}{\textbf{Annotation}} \\\cline{3-10}
-   &  & \textbf{new}  & \textbf{same} & \textbf{word} & \textbf{punctuation}  & \textbf{patient} & \textbf{provider} & \textbf{method}  & \textbf{codes} \\ \hline    
+   &  & \textbf{new}  & \textbf{same} & \textbf{words} & \textbf{punctuations}  & \textbf{patient} & \textbf{provider} & \textbf{method}  & \textbf{codes} \\ \hline    
  49 & 102,917 & 3,138 & 99,779 & 95,777 & 7,140 & 281 & 330 & MYSCOPE & 115 \\ \hline
+ \multicolumn{10}{|c|}{segmentation example: Hi[0] XXX[0] ,[0] its[0] good[0] to[0] hear[0] from[0] you[0] .[1] it[0] sounds[0] like[0]...} \\ \hline
   \end{tabular}
@@
  word embeddings only & 0.780 & 0.736 & 0.655 & \textbf{0.818} \\ \hline
- word embeddings + POS & 0.797 & 0.746 & 0.647 & \textbf{0.798} \\ \hline
- word embeddings + punctuation & \textbf{0.876} & 0.835 & 0.774 & 0.874 \\ \hline
- all features & \textbf{0.877} & 0.842 & 0.770 & 0.867 \\ \hline
+ word embeddings + POS & 0.797 (+2.18\%) & 0.746 (+1.36\%) & 0.647 (-1.22\%) & \textbf{0.798 (-2.44\%)} \\ \hline
+ word embeddings + punctuation & \textbf{0.876 (+12.31\%)} & 0.835 (+13.45\%) & 0.774 (+18.17\%) & 0.874 (+6.85\%) \\ \hline
+ all features & \textbf{0.877 (+12.44\%)} & 0.842 (+14.4\%) & 0.770 (+17.56\%) & 0.867 (+6\%) \\ \hline
   \end{tabular}
@@
 Table~\ref{tab:result_aupr} shows the impact of individual features as well as their combination on the segmentation task.
-\hl{Influence of the punctuation and POS features is also consistent in AUPR values, which increase by 12.44\%, 14.4\%, 17.56\% and 5.99\% for CRF, MLP, BRNN and CRNN methods, respectively, when all features are used together.
+\hl{Influence of the punctuation and POS features is also consistent in AUPR values, which increase by 12.44\%, 14.4\%, 17.56\% and 6\% for CRF, MLP, BRNN and CRNN methods, respectively, when all features are used together.
 Individually, POS and punctuation features also improve the performance of all classifiers except BRNN and CRNN when POS features are used.

### 2018-11-18 paper-amia-summit-2019/manuscript.tex
update manuscript and response to reviewers comments

@@
 In the case of CRF, BRNN and CRNN models, an email was taken as an input sequence, POS tags and word embeddings of each word or punctuation mark were used as input and binary labels (1 or 0) corresponding to ``new segment'' and ``same segment'' classification decisions were considered as the model output of a model.
-In the gold standard, words or punctuations within the same segment were assigned the label of 0 and the last word or punctuation mark of a segment were assigned the label of 1 (an example is given at $2^{nd}$ row in Table~\ref{tab:datastat}).
+In the gold standard, words or punctuations within the same segment were assigned the label of 0 and the last word or punctuation mark of a segment were assigned the label of 1 (\hl{an example is given at $2^{nd}$ row in Table}~\ref{tab:datastat}).
 \\

### 2018-11-21 paper-amia-summit-2019/manuscript.tex
edited by April and Alex up to page 5

--- 2018-11-18 paper-amia-summit-2019/manuscript.tex
@@
 \usepackage{multirow}
+\usepackage{url}
 \renewcommand*{\thefootnote}{\alph{footnote}}
@@
 \noindent{\bf Abstract}
-\textit{Communication science approaches to developing effective behavior interventions, such as motivational interviewing (MI), are limited by the traditionally manual qualitative coding of communication exchanges, which is a very resource-intensive and time-consuming process.
+\textit{Communication science approaches to develop effective behavior interventions, such as motivational interviewing (MI), are limited by traditional qualitative coding of communication exchanges, a very resource-intensive and time-consuming process.
 This study focuses on the analysis of e-Coaching sessions, behavior interventions that are delivered via email and grounded in the principles of MI.
-A critical step towards automated annotation of e-Coaching communication exchanges is segmentation of emails into fragments that correspond to MI behaviors.
-This study formulates this task as a classification problem and proposes word embeddings, punctuation and part-of-speech features to address it.
-We experimented both with traditional machine learning method, conditional random fields (CRF) and deep learning methods, such as multilayer perceptrons (MLP), bidirectional recurrent neural networks (BRNN) and convolutional recurrent neural networks (CRNN).
-Results indicate that CRNN outperformed CRF, MLP and BRNN achieving 0.989 macro F1-score overall and 0.825 macro F1-score for detecting new segment.}
+A critical step towards automated qualitative coding of e-Coaching communication exchanges is segmentation of emails into fragments that correspond to MI behaviors.
+This study frames email segmentation task as a classification problem and utilizes word embeddings in conjunction with punctuation and part-of-speech features to address it.
+We experimentally evaluated conditional random fields (CRF), a traditional machine learning method, as well as deep learning methods, such as multilayer perceptron (MLP), bidirectional recurrent neural network (BRNN) and convolutional recurrent neural network (CRNN).
+Our results indicate that CRNN outperform CRF, MLP and BRNN for the task of email segmentation achieving 0.989 macro F1-score overall and 0.825 macro F1-score for new segment detection.}
 
 \section*{Introduction}
-The emergence of e-Health technologies opened up new ways to deliver a variety of behavioral interventions to any demographic group of patients in any geographical location.
-Motivational interviewing (MI), an evidence-based communication technique to increase intrinsic motivation and self-efficacy for behavior change,\cite{miller2012motivational,miller2009toward} is one type of these interventions.
-MI sessions are generally aimed at eliciting ``change talk'', or statements of intrinsic motivation about patients' own desire, ability, reasons, need for and commitment to behavior change, which have been established by previous research\cite{apodaca2009mechanisms} as a reliable mediator of health behavior change.
+The emergence of e-Health technologies has greatly expanded the reach of behavioral interventions.
+One such intervention is email-delivered Motivational Interviewing (MI).
+MI is an evidence-based communication technique to increase intrinsic motivation and self-efficacy for behavior change\cite{miller2012motivational,miller2009toward}.
+MI is linked to behavior change through the elicitation of patient ``change talk'', or statements of intrinsic motivation expressing patients’ own desires, abilities, reasons, need for, and commitment to behavior change\cite{apodaca2009mechanisms}.
 However, communication science approaches to understanding the efficacy of MI are inherently limited by traditional qualitative coding methods.
+The specific provider behaviors responsible for the elicitation of change talk, however, are less clear and may vary by treatment context.
+Thus, a current focus of MI research is understanding which provider behaviors in which contexts lead to patient change talk.
+To date, no study has examined this process in email-delivered MI.
 
-Qualitative analysis of motivational interviews has been traditionally performed manually by MI researchers, which is a tedious and resource-intensive process involving several iterations of reading, comprehension and interpretation of interview transcripts.
-Rapidly developing computational technologies, specifically, machine learning methods, offer a unique opportunity to accelerate this process.
+The study of patient-provider communication in MI relies on the behavioral coding of intervention transcripts, an iterative, resource-intensive, and cognitively demanding process of qualitative analysis involving several iterations of reading, comprehension and interpretation of interview transcripts.
+Rapidly developing computational technologies, specifically, machine learning methods, offer an opportunity to dramatically accelerate this process.
 In particular, machine learning methods have been successfully applied to a variety of analytical tasks involving textual data, such as classification\cite{nigam2000text} and sentiment analysis.\cite{wang2012baselines} In our previous work, we examined the utility of machine learning methods for automated annotation \cite{hasan2016study,kotov2015interpretable} and predicting the outcome \cite{hasan2018predicting} of in-person MI sessions.
-Experimental data utilized in these studies, however, were transcribed auto recordings of in-person MI sessions with a counselor, which have a clear discourse structure and were clearly segmented into the utterances by a counselor, a patient or a caregiver.
+Experimental data utilized in these studies were transcribed auto recordings of in-person MI sessions with a counselor, which were segmented into counselor and client utterances during the transcription process.
 
-In this study, we focus on the analysis of e-Coaching sessions, behavioral interventions that are delivered via email and are grounded in the principles of MI.
-Specifically, the e-Coaches involved in this study used emails to communicate motivation-enhancing messages that encourage healthy eating among GenY adolescents.
-E-coaching data is comprised of email responses, which are free-text documents, unlike more traditional dyadic clinical interviews that are naturally segmented into utterances due to their conversational nature.
-
-The unstructured nature of e-Coaching exchanges poses a unique set of challenges for their qualitative analysis.
-Usually, email segments are automatically coded with the Minority Youth-Sequential Coding of Process Exchanges (MYSCOPE), \cite{carcone2013provider} a qualitative coding scheme to characterize patient-provider communication during MI session.
-A significant barrier to qualitative analysis of e-Coaching exchanges is their segmentation into textual fragments that correspond to a distinct e-Coach and patient communication behaviors.
+In this study, we focus on the analysis of email-delivered MI, or e-Coaching, to promote healthy eating among young adults.
+The e-Coaching dataset is composed of email correspondence between a MI counselor and the young adult patient.
+Unlike transcribed in-person discourse, email correspondence is not clearly segmented into codable speech acts (i.e., utterances).
+Thus, the unstructured nature of e-Coaching exchanges poses a unique set of analytic challenges.
+Segmentation of e-Coaching exchanges into textual fragments that correspond to distinct e-Coach and patient communication behaviors is a significant barrier to qualitative analysis of this type of clinical conversation.
 Automating this task is a unique and challenging problem due to the following reasons:
@@
 For example, an e-Coach usually responds to several previous patient statements in one email.
-However, in a traditional MI interview session involving a dialogue between a provider and a patient, a provider utterance is a response to an immediately preceding inquiry or statement by a patient.
-\item Discourse segments in e-Coaching emails do not necessarily correspond to sentences or paragraphs.
-One sentence can be divided into fragments corresponding to multiple MI behaviors.
+In contrast, in a traditional, in-person MI session, an utterance is assumed to be a response to the immediately preceding utterance.
+\item Discourse segments in e-Coaching do not have a clear breakpoint, such as the end of a sentence or paragraph.
+One sentence might be divided into fragments corresponding to multiple MI behaviors.
 On the other hand, an MI behavior may comprise several sentences.
@@
     \includegraphics[width=0.7\textwidth]{figures/segment-example.png}
-    \caption{\textbf{Example of e-Coaching emails segmented into fragments that correspond to MI behaviors of an e-Coach and a patient}}
+    \caption{\textbf{Example of an e-Coaching exchange segmented into fragments corresponding to MI behaviors of an e-Coach and a patient}}
     \label{fig:text-segment}
@@
 
-Figure~\ref{fig:text-segment} illustrates a segmentation of an e-Coaching email exchange, in which the first sentence is segmented into 2 MI behavior fragments, while the fourth and fifth segments correspond to one and three sentences, respectively.
-Segmentation of e-Coaching email exchanges between patients and e-Coaches corresponds to a special case of discourse analysis \cite{webber2012discourse} aimed at better understanding the effective communication strategies specific to this type of behavioral interventions and revealing the unique socio-psychological characteristics of patients.
+Figure~\ref{fig:text-segment} illustrates a segmentation of an e-Coaching exchange, in which the first sentence is segmented into 2 MI behavior fragments, while the fourth and fifth segments correspond to one and three sentences, respectively.
+Segmentation of e-Coaching exchanges constitutes a special case of clinical discourse analysis \cite{webber2012discourse} aimed at better understanding the effective communication strategies specific to this type of behavioral interventions.
 
-The goal of this study is to assess the effectiveness of deep learning methods for the task of automated segmentation of e-Coaching emails into textual fragments corresponding to individual behaviors, which is the first step of qualitative analysis of this type of clinical communications.
-Specifically, for this study, we utilized the data from MENU GenY (Making Effective Nutrition Choices for Generation Y) \cite{alexander2017motivations}, a recent \hl{email-based} public health intervention that relies on personalized e-Coaching to encourage increased fruit and vegetable intake among young adults, aged 21--30.
-The goal of MENU GenY was to develop a better coding scheme for e-Coaching communications aimed at improving GenY eating habits.
+The goal of this study is to assess the effectiveness of deep learning methods for the task of automated segmentation of e-Coaching emails into textual fragments corresponding to individual patient and provider behaviors.
+For this study, we utilized the data from MENU GenY (Making Effective Nutrition Choices for Generation Y) \cite{alexander2017motivations}, a web-delivered public health intervention with email-based coaching to encourage increased fruit and vegetable intake among young adults, aged 21–-30.
+A secondary goal of the MENU GenY project was to identify the specific communication strategies used by e-Coaches to elicit change talk for healthier eating among young adult patients.
 Segmentation of clinical conversation in the context of electronically delivered interventions into groups of MI behaviors, is traditionally performed manually by MI researchers, which significantly slows down their qualitative analysis.
-This paper is the first work to evaluate the empirical effectiveness of deep learning architectures in addressing \textit{the problem of discourse segmentation in the context of e-mail based behavioral interventions}.
+This paper is the first work to evaluate the empirical effectiveness of deep learning architectures in addressing \textit{the problem of discourse segmentation in the context of e-mail-based behavioral interventions}.
 
-Specifically, we introduce and evaluate the effectiveness of word embedding or lexical, punctuation and part-of-speech (POS) features in conjunction with both traditional supervised machine learning methods, such as linear-chain Conditional Random Field (CRF)\cite{lafferty2001conditional} and deep learning architectures, such as Multi-Layer Perceptron (MLP),\cite{rumelhart1986learning} Bidirectional Recurrent Neural Network (BRNN)\cite{schuster1997bidirectional} and Convolutional Recurrent Neural Network (CRNN),\cite{treviso2017sentence} to find the best performing method and feature combination.
+Specifically, we introduce and evaluate the effectiveness of distributed word representations (i.e.
+word embeddings) as well as lexical, punctuation and part-of-speech (POS) features in conjunction with both traditional supervised machine learning methods, such as linear-chain Conditional Random Fields (CRF)\cite{lafferty2001conditional} and deep learning architectures, such as Multi-Layer Perceptron (MLP),\cite{rumelhart1986learning} Bidirectional Recurrent Neural Network (BRNN)\cite{schuster1997bidirectional} and Convolutional Recurrent Neural Network (CRNN),\cite{treviso2017sentence} to find the best performing method and feature combination.
 
-\section*{Relevant work}
-Previous relevant work in the biomedical domain primarily focused on segmentation of text in electronic health records (EHR) into sections and headers \cite{apostolova2009automatic,denny2009evaluation,tepper2012statistical,cho2002text} or sentence boundary detection.\cite{griffis2016quantitative,kreuzthaler2015detection,treviso2017sentence} In particular, Maximum Entropy models \cite{tepper2012statistical} and Support Vector Machine (SVM) along with word-vector cosine similarity metric and several heuristics \cite{apostolova2009automatic} have been applied to segment or classify clinical documents in EHR into pre-defined sections, such as general patient information, medical history, procedures, findings, etc.
+\section*{Related work}
+
+Prior work on textual segmentation in the biomedical domain primarily focused on sentence boundary detection \cite{griffis2016quantitative,kreuzthaler2015detection,treviso2017sentence} and segmentation of clinical documents in patients' electronic health records (EHR) into sections and headers.
+\cite{apostolova2009automatic,denny2009evaluation,tepper2012statistical,cho2002text} In particular, Maximum Entropy models \cite{tepper2012statistical} and Support Vector Machine (SVM) along with word-vector cosine similarity metric and several heuristics \cite{apostolova2009automatic} have been applied to identify specific sections in EHR, such as general patient information, medical history, procedures, findings, etc.
 Denny et al.
-\cite{denny2009evaluation} proposed a SecTag algorithm, which combined natural language processing techniques, terminology-based rules and a Naive Bayes classifier to identify sections and headers in EHR.
-Segmentation of e-Coaching emails, however, is different from segmentation of other clinical documents in that the focus is on clinical conversation.
+\cite{denny2009evaluation} proposed SecTag algorithm, which combined natural language processing techniques, terminology-based rules and a Na\"{i}ve Bayes classifier to identify sections and headers in EHR.
+Segmentation of e-Coaching emails, however, is different from segmentation of other clinical documents, since the focus is on dialog acts in clinical conversation.
 
-SVM in conjunction with prosodic and part of speech features \cite{kreuzthaler2015detection} and Recurrent Convolutional Neural Networks \cite{griffis2016quantitative} have also been utilized for \textit{sentence boundary detection} in general text.
+SVM in conjunction with prosodic and part-of-speech features \cite{kreuzthaler2015detection} and Recurrent Convolutional Neural Networks \cite{griffis2016quantitative} have also been utilized for \textit{sentence boundary detection} in general text.
 Liu et al.
@@
 
-Segmentation of e-Coaching emails is different from a traditional shallow discourse analysis of conversations \cite{galley2003discourse} in that the focus is on segmentation, rather than on determining the types of transitions between the utterances or assigning utterances to speakers.
-The proposed methods will automate the process of segmenting clinical exchanges into groups of MI behaviors, which will significantly reduce the resources and time required to perform this task manually.
-Furthermore, these methods can be integrated with auto coding methods \cite{hasan2016study,kotov2015interpretable} to create a software pipeline for automated analysis of clinical conversation.
+Segmentation of e-Coaching emails is also different from traditional shallow discourse analysis \cite{galley2003discourse}, which besides identification of speech acts, also aims to determine the types of transitions between speech acts and labels speech acts with speakers who performed them in a multi-speaker conversation.
+The proposed methods will automate the process of segmenting clinical exchanges into MI behaviors, which will significantly reduce the time and resources required to perform such segmentation manually.
+Furthermore, these methods can be integrated with auto coding methods \cite{hasan2016study,kotov2015interpretable} to create a software pipeline for fully automated analysis of email-delivered behavioral interventions.
 
-\section*{Methods}
+\section*{Materials and methods}
 \subsection*{\textit{Data collection}}
-The experimental dataset for this work was constructed from 49 e-Coaching sessions, which include \hl{330 and 281 emails from ecoaches and patients, respectively }(Table~\ref{tab:datastat}).
-\hl{Each session represents an MI intervention delivered via email.
-Emails were segmented into 3,138 text fragments and annotated using 115 distinct MYSCOPE}~\cite{carcone2013provider} \hl{behavior codes.
+The experimental dataset for this work was constructed from 49 e-Coaching sessions, which include 330 and 281 emails by e-Coaches and patients, respectively.
+Various statistics of experimental dataset are provided in Table~\ref{tab:datastat}.
+Each session represents an MI intervention delivered via email.
+Emails were segmented into 3,138 text fragments and and annotated using the Minority Youth-Sequential Coding of Process Exchanges (MYSCOPE),~\cite{carcone2013provider} a qualitative coding scheme to characterize patient-provider communication during MI sessions with 115 distinct behavior codes.
 We consider email segmentation as a binary classification problem, in which each word or punctuation mark is annotated with a class label ``new segment'' or ``same segment'' to indicate whether it precedes a new segment or not.
-In total, the dataset consists of 95,777 words and 7,140 punctuation marks and includes 3,138 ``new segment'' and 99,779 ``same segment'' instances.} In this study, we experimented with traditional machine learning methods, such as Conditional Random Field (CRF)\cite{lafferty2001conditional} and deep learning methods, such as Multi-Layer Perceptron (MLP),\cite{rumelhart1986learning} Bidirectional Recurrent Neural Network (bidirectional RNN or BRNN)\cite{schuster1997bidirectional} and Convolutional Recurrent Neural Network (CRNN).\cite{treviso2017sentence} For an MLP model, training and testing instances were created based on a sliding window of $2n$ words or punctuation marks over each position in a given input sequence, such that each instance consists of a pair of the next $n$ words or punctuation marks and prior $n$ words or punctuation marks, including the current word or punctuation mark.
-Each sample is classified into either a ``new segment'' or a ``same segment'' based on whether there should be a segment break after the current word or not.
-In the case of CRF, BRNN and CRNN models, an email was taken as an input sequence, POS tags and word embeddings of each word or punctuation mark were used as input and binary labels (1 or 0) corresponding to ``new segment'' and ``same segment'' classification decisions were considered as the model output of a model.
-In the gold standard, words or punctuations within the same segment were assigned the label of 0 and the last word or punctuation mark of a segment were assigned the label of 1 (an example is given at $2^{nd}$ row in Table~\ref{tab:datastat}).
+In total, the dataset consists of 95,777 words and 7,140 punctuation marks and includes 3,138 ``new segment'' and 99,779 ``same segment'' instances.
+In this study, we experimented with traditional machine learning methods, such as Conditional Random Fields (CRF)\cite{lafferty2001conditional} and deep learning methods, such as Multi-Layer Perceptron (MLP),\cite{rumelhart1986learning} Bidirectional Recurrent Neural Network (bidirectional RNN or BRNN)\cite{schuster1997bidirectional} and Convolutional Recurrent Neural Network (CRNN).\cite{treviso2017sentence} For an MLP model, training and testing samples were created based on a sliding window of $2n$ words or punctuation marks over each position (which could be a word or a punctuation mark) in a given input sequence, such that each instance consists of the $n$ words or punctuation marks after this position and $n$ words or punctuation marks prior to this position, including the position itself.
+Each sample is assigned a ``new segment'' or ``same segment'' label based on whether there should be a segment break after the current position or not.
+In the case of CRF, BRNN and CRNN models, an e-Coaching email was taken as an input sequence, POS tags and word embeddings of each word or punctuation mark were used as input and binary labels corresponding to ``new segment'' and ``same segment'' classification decisions were considered as the model output.
+In the gold standard, words or punctuations within the same segment were assigned the label of 0 and the last word or punctuation mark of a segment were assigned the label of 1 (\hl{an example of a sequence with assigned labels is provided in the $2^{nd}$ row of Table}~\ref{tab:datastat}).
 \\
@@
 \centering
-\caption{\textbf{\hl{Summary of the experimental dataset}}}
+\caption{\textbf{Summary of statistics of experimental dataset.}}
 \label{tab:datastat}
@@
 We utilized three types of features in conjunction with CRF, MLP, BRNN and CRNN methods: word embeddings or lexical features, punctuation and POS features.
-Since POS tags have been shown to be effective semantic abstractions of individual words, we used POS features for our experiment.\cite{liu2005using,treviso2017sentence} To extract POS features, we pre-processed e-Coaching emails using the NLTK POS tagger.
-Punctuation marks, which correspond to one of the symbols \{`.', `,', `!', `?', `:', `;'\} between a pair of words, were also used as a feature, since punctuation marks designate the boundary of a sentence, clause and phrase and often also correspond to a segment boundary.\cite{cho2002text} For natural language processing (NLP) tasks, inputs are received as a text, in which individual words are as the basic units of semantics.
-Therefore, it is important to represent a word in such a way that it carries all relevant information.
-Word embedding is one such representation, when each word is associated with a real-valued vector in a high dimensional vector space.
-Distributed representations of words have been shown to capture semantic, syntactic and morphological properties of words.\cite{pennington2014glove, mikolov2013distributed} For experiments reported in this paper, we utilized embeddings estimated on Google News corpus consisting of 1.6 billion words pre-trained using word2vec word embedding method.\footnote{\label{fn:word2vec}https://code.google.com/p/word2vec/} When words or punctuation marks are not found in the pre-trained word vectors, we utilized word embeddings trained with our e-Coaching email corpus, which we refer to as corpus-based embeddings.
-CRF utilized lexical features, which correspond to words in e-Coaching emails.
+Syntactic abstractions of individual words, such as POS tags, have been previously shown to be effective features for similar natural language processing tasks.\cite{liu2005using,treviso2017sentence} To extract POS features, we pre-processed e-Coaching emails using the NLTK POS tagger\footnote{\url{https://www.nltk.org/}}.
+Punctuation marks, which correspond to one of the symbols \{`.', `,', `!', `?', `:', `;'\} between a pair of words, were also used as a feature, since punctuation marks designate the boundary of a sentence, clause or phrase and often also correspond to a segment boundary.\cite{cho2002text} For natural language processing (NLP) tasks, inputs are received as a text, in which individual words are as the basic lexico-semantic units.
+Therefore, it is important to represent a word in such a way that preserves all relevant lexical and semantic information.
+Dense real-valued vector (i.e.
+embedding) is a form of distributed word representation, in which each word is associated with a vector in a low-dimensional vector space.
+Embeddings have been previously shown to effectively capture semantic, syntactic and morphological properties of words.\cite{pennington2014glove, mikolov2013distributed} For experiments reported in this paper, we utilized word embeddings pre-trained on Google News corpus consisting of 1.6 billion words using word2vec software package.\footnote{\label{fn:word2vec}https://code.google.com/p/word2vec/} We skipped the words or punctuation marks, for which pre-trained embeddings could not be found.
+CRF utilized lexical features and POS tags.
 
@@
 
-We experimented with 4 different classifiers, including one traditional machine learning method, CRF and three deep learning models.
-Since deep learning architectures provide a flexible mechanism for constructing complex models, we take the advantage of this flexibility to test MLP, BRNN and CRNN models for the task of segmentation of e-Coaching emails.
+We experimented with 4 different classifiers, including one traditional machine learning model (CRF) and three deep learning methods (MLP, BRNN and CRNN).
+Since deep learning architectures provide a flexible mechanism for constructing complex models, we take advantage of this flexibility to test different variations of MLP, BRNN and CRNN models for the task of segmentation of e-Coaching emails.
 
-\textbf{Conditional Random Field}: CRF has been widely used in various NLP tasks, such as part-of-speech tagging.\cite{lafferty2001conditional, hirohata2008identifying} Unlike a maximum entropy Markov model, which uses per-state exponential models for conditional probability of the next state given a current state, CRF model directly estimates conditional distribution of the entire output sequence, given the observation sequence.
-A traditional linear-chain CRF model is defined as a conditional probability distribution $p(y|x)$ for output and input sequences, $y$ and $x$:
+\textbf{Conditional Random Field}: CRF has been widely used in various sequence annotation NLP tasks, such as part-of-speech tagging.\cite{lafferty2001conditional, hirohata2008identifying} Unlike a maximum entropy Markov model, which uses per-state exponential models for conditional probability of the next state given a current state, CRF model directly estimates a distribution of the entire output sequence conditioned on the observation sequence.
+A traditional linear-chain CRF model is defined as a conditional probability distribution $p(y|x)$ of output sequence $y$, given input sequence $x$:
 
@@
 where $Z_x$ is a normalization factor, $f_k(y_{t-1}, y_t, x, t)$ is a feature function, and $\lambda_k$ is a learned weight associated with feature $f_k$.
-The optimal output sequence $y^*$ for an input sequence $x$, $y^* = {arg\,max}_y p(y|x)$, is obtained efficiently using the Viterbi algorithm.
+The optimal output sequence $y^*$ for input sequence $x$, $y^* = {arg\,max}_y p(y|x)$, is obtained efficiently using the Viterbi algorithm.
 In our experiments, the following features were utilized in conjunction with CRF models: i) current word or punctuation ii) next and previous 3 words or punctuations iii) binary feature indicating whether a word or punctuation is a special character (';', '?', '.', ',', '!', ':', etc.) or not iv) binary feature indicating whether a word is a title word or not (e.g.
@@
 
-\textbf{Multi-Layer Perceptron}: MLP is a feed-forward artificial neural network, which maps an input onto one or several outputs.\cite{rumelhart1986learning} Figure~\ref{fig:mlp} shows a multilayer perceptron with a single hidden layer.
-In MLP networks, there is no cycle or loop and information moves forward only, from the input nodes through the hidden nodes and to the output nodes.
+\textbf{Multi-Layer Perceptron}: MLP is a neural network, which consists of multiple fully connected layers map an input onto one or several outputs.\cite{rumelhart1986learning} Figure~\ref{fig:mlp} shows a multi-layer perceptron with a single hidden layer.
+In MLP networks, there is no cycles or loops and information flows only forward, from the input layer through the hidden layer(s) to the output layer.
 This study utilized MLP with a nonlinear activation function (rectified linear unit) and one hidden layer consisting of 128 hidden units.
@@
 \section*{Conclusion}
+
+%Segmentation is the first step of qualitative analysis of unstructured clinical communications.
+
 Segmentation of e-Coaching emails is an integral part of developing and analyzing e-Coaching behavioral interventions.

### 2018-11-26 paper-amia-summit-2019/manuscript.tex
initial final version by Alex

@@
 Specifically, we introduce and evaluate the effectiveness of distributed word representations (i.e.
-word embeddings) as well as lexical, punctuation and part-of-speech (POS) features in conjunction with both traditional supervised machine learning methods, such as linear-chain Conditional Random Fields (CRF)\cite{lafferty2001conditional} and deep learning architectures, such as Multi-Layer Perceptron (MLP),\cite{rumelhart1986learning} Bidirectional Recurrent Neural Network (BRNN)\cite{schuster1997bidirectional} and Convolutional Recurrent Neural Network (CRNN),\cite{treviso2017sentence} to find the best performing method and feature combination.
+word embeddings) as well as lexical, punctuation and part-of-speech (POS) features in conjunction with both traditional supervised machine learning methods, such as linear-chain Conditional Random Fields (CRF)\cite{lafferty2001conditional} and deep learning methods, such as Multi-Layer Perceptron (MLP),\cite{rumelhart1986learning} Bidirectional Recurrent Neural Network (BRNN)\cite{schuster1997bidirectional} and Convolutional Recurrent Neural Network (CRNN),\cite{treviso2017sentence} to find the best performing method and feature combination.
 
@@
 In the case of CRF, BRNN and CRNN models, an e-Coaching email was taken as an input sequence, POS tags and word embeddings of each word or punctuation mark were used as input and binary labels corresponding to ``new segment'' and ``same segment'' classification decisions were considered as the model output.
-In the gold standard, words or punctuations within the same segment were assigned the label of 0 and the last word or punctuation mark of a segment were assigned the label of 1 (\hl{an example of a sequence with assigned labels is provided in the $2^{nd}$ row of Table}~\ref{tab:datastat}).
+In the gold standard, words or punctuations within the same segment were assigned the label of 0 and the last word or punctuation mark of a segment were assigned the label of 1 (an example of a sequence with assigned labels is provided in the $2^{nd}$ row of Table~\ref{tab:datastat}).
 \\
@@
 \subsection*{\textit{Features}}
-We utilized three types of features in conjunction with CRF, MLP, BRNN and CRNN methods: word embeddings or lexical features, punctuation and POS features.
+We utilized three types of features in conjunction with CRF, MLP, BRNN and CRNN methods: word embeddings as lexical features, punctuation and POS features.
 Syntactic abstractions of individual words, such as POS tags, have been previously shown to be effective features for similar natural language processing tasks.\cite{liu2005using,treviso2017sentence} To extract POS features, we pre-processed e-Coaching emails using the NLTK POS tagger\footnote{\url{https://www.nltk.org/}}.
@@
 embedding) is a form of distributed word representation, in which each word is associated with a vector in a low-dimensional vector space.
-Embeddings have been previously shown to effectively capture semantic, syntactic and morphological properties of words.\cite{pennington2014glove, mikolov2013distributed} For experiments reported in this paper, we utilized word embeddings pre-trained on Google News corpus consisting of 1.6 billion words using word2vec software package.\footnote{\label{fn:word2vec}https://code.google.com/p/word2vec/} We skipped the words or punctuation marks, for which pre-trained embeddings could not be found.
-CRF utilized lexical features and POS tags.
+Embeddings have been previously shown to effectively capture semantic, syntactic and morphological properties of words.\cite{pennington2014glove, mikolov2013distributed} For experiments reported in this paper, we utilized word embeddings pre-trained on Google News corpus consisting of 1.6 billion words using word2vec software package.\footnote{\label{fn:word2vec}https://code.google.com/p/word2vec/} For words or punctuation marks, which do not have pre-trained embeddings, we utilized the embeddings of the same dimensionality trained on the experimental dataset.
+CRF utilized lexical features, POS tags and the preceding label.
 
-\subsection*{\textit{Classifiers}}
+\subsection*{\textit{Methods}}
 
@@
 
-\textbf{Multi-Layer Perceptron}: MLP is a neural network, which consists of multiple fully connected layers map an input onto one or several outputs.\cite{rumelhart1986learning} Figure~\ref{fig:mlp} shows a multi-layer perceptron with a single hidden layer.
-In MLP networks, there is no cycles or loops and information flows only forward, from the input layer through the hidden layer(s) to the output layer.
-This study utilized MLP with a nonlinear activation function (rectified linear unit) and one hidden layer consisting of 128 hidden units.
-In order to prevent over-fitting, we utilized dropout in fully connected layers by randomly hiding neurons during training.
-\cite{srivastava2014dropout} Dropout is also applied to the fully connected layers in BRNN and CRNN models.
+\textbf{Multi-Layer Perceptron}: MLP is a neural network, which consists of multiple fully connected layers that map an input onto one or several outputs.\cite{rumelhart1986learning} Figure~\ref{fig:mlp} illustrates a multi-layer perceptron with a single hidden layer.
+MLPs have no cycles or loops.
+Information in them flows only forward, from the input layer through the hidden layer(s) to the output layer.
+The MLP in this study utilizes one hidden layer consisting of 128 neurons and rectified linear unit (ReLU) as a nonlinear activation function.
+In order to prevent over-fitting, we applied dropout (random masking of neurons) \cite{srivastava2014dropout} to fully connected layers during training.
+Dropout was also applied to fully connected layers in BRNN and CRNN.
 
@@
     \includegraphics[width=0.45\textwidth]{figures/mlp.eps}
-    \caption{\textbf{A multilayer perceptron with a single hidden layer}}
+    \caption{\textbf{Multi-layer perceptron with a single hidden layer}}
     \label{fig:mlp}
@@
 
-\textbf{Bidirectional Recurrent Neural Network}: BRNN is a neural network designed to capture sequential patterns by considering both past \& future inputs as well as complex relationships between input features and output labels.\cite{schuster1997bidirectional} The output of BRNN layer is computed as an aggregation of outputs of the forward and backward RNNs.
+\textbf{Bidirectional Recurrent Neural Network}: BRNN is a neural network designed to capture sequential patterns by considering both past  and future inputs as well as complex relationships between input features and output labels.\cite{schuster1997bidirectional} The hidden state of BRNN is an aggregation of the hidden states of a forward and backward recurrent neural networks (RNNs).
 Gated Recurrent Units (GRU)\cite{chung2014empirical} capable of handling variable size input sequence and having internal memory, which can be reset, were utilized as an RNN.
-Figure~\ref{fig:crnn} represents the architecture of BRNN, in the case when a convolution layer is removed.
+%Figure~\ref{fig:crnn} represents the architecture of BRNN, in the case when a convolution layer is removed.
 
-\textbf{Convolutional Recurrent Neural Network}: CRNN is a deep neural network architecture,\cite{treviso2017sentence} shown in Figure~\ref{fig:crnn}, which consists of 5 layers: 1) input layer 2) embedding layer 3) convolution layer with max pooling 4) BRNN layer 5) fully connected layer with dropout and sigmoid output.
-E-coaching email exchanges are represented as a sequence of $m$ words and punctuations, which are fed into the input layer to produce a $m \times n_e$ matrix after fetching the pre-trained word vectors.
-\hl{This matrix represents embedding output, contain sufficient morpho-syntactic information for segmentation of email exchanges}.
-When POS tags are used in combination with word embeddings, \hl{POS tags are embedded with 10-dimensional POS vectors and concatenated with 300-dimensional word vectors to obtain new embedding} vectors $n_e = [n_w;n_p]$ of size 310.
-The primary purpose of convolution is to extract new features depending on the neighboring words.
-We used 1D convolution.
-In this layer, one filter is responsible for extraction of one feature.
-After applying $n_f$ different filters with zero-padding on both sides of the input text, $n_f$ useful features are produced in the convolution layer for each word.
-A max pooling operation was then performed over time to find the most significant features in a textual fragment.
-The bidirectional recurrent layer receives new features extracted in the convolution layer.
-RNN is usually used to capture long-range dependencies in a sequence of observations.
-Moreover, bidirectional RNNs are capable of capturing both past and future contexts through forward and backward traversals of a sequence.
-The purpose of the fully connected layer is to use the output of bidirectional RNN layer for classifying each word or punctuation into ``new segment'' and ``same segment'' classes.
-Since a fully connected layer has a greater number of parameters, they are more likely to excessively co-adapt with these parameters, which may cause over-fitting.
-To prevent this, we utilized dropout by randomly disregarding 50\% of the connections to the fully connected layer.
-After that, logistic sigmoid \hl{outputs the probability of each class}.
-We experimentally determined the optimal parameters and found that the best performance is achieved with 5-fold cross-validation when filter length in convolutional layer is 7, number of filters is 100, max-pool size is 3, ReLU is used as a convolutional layer activation function, RNN layer activation is hyperbolic tangent and the number of recurrent units in RNN is 200.
-Adam\cite{kingma2014adam} with 50 epochs, the batch size of 32 and learning rate of 0.001 was used for optimization and the early stopping strategy was applied.\footnote{\hl{source codes are available at https://github.com/teanalab/eCoaching-Text-Segmentation}}
+\textbf{Convolutional Recurrent Neural Network}: CRNN \cite{treviso2017sentence} shown in Figure~\ref{fig:crnn} is a deep neural network architecture, which combines convolutional and recurrent layers.
+Our implementation of CRNN consists of 5 layers: 1) input layer 2) embedding layer 3) convolution layer with max pooling 4) BRNN layer 5) fully connected layer with dropout and sigmoid output.
+E-coaching email exchanges are represented as a sequence of $m$ words and punctuations, which are fed into the input and embedding layers to produce a $m \times n_e$ matrix after fetching the embeddings for words and punctuations in the input sequence.
+This matrix corresponds to the distributed representation of an input email exchange, which contains rich morpho-syntactic information that can be utilized for its segmentation.
+When POS tags are utilized along with word embeddings, they are represented with a 10-dimensional vector, which is concatenated with 300-dimensional word embeddings to obtain new embedding vectors $n_e = [n_w;n_p]$ of size 310.
+The primary purpose of a convolution layer is to extract new features for each word or punctuation mark based on the neighboring words or punctuation marks.
+A one-dimensional (1D) convolution operation is utilized in this layer in our implementation of BRNN.
+In 1D convolution, one filter is responsible for extraction of one feature.
+After applying $n_f$ different filters with zero-padding on both sides of the input text, $n_f$ features are produced by the convolution layer for each word.
+A max pooling over time operation was then performed to find the most significant features in a textual fragment.
+The bidirectional recurrent layer receives new features extracted by the convolution layer.
+Unidirectional RNNs are usually used to capture long-range dependencies in a sequence of observations.
+Bidirectional RNNs, on the other hand, are capable of capturing both past and future contexts through forward and backward traversals of a sequence.
+The purpose of the fully connected layer is to use the output of the bidirectional RNN layer for classifying each word or punctuation into ``new segment'' or ``same segment'' classes.
+Since a fully connected layer has a larger number of parameters, they are more likely to excessively co-adapt to  other parameters, which may result in over-fitting.
+To prevent this, we utilized dropout by randomly ignoring 50\% of the connections in the fully connected layer.
+Finally, logistic sigmoid outputs the probability of classifying or labeling each word or punctuation mark with ``same segment'' class.
+We experimentally determined the optimal parameters using 5-fold cross-validation and found out that the best performance is achieved when filter length in the convolution layer is 7, number of filters is 100, max-pool size is 3, ReLU is used as an activation function in the convolution layer, hyperbolic tangent is used as an activation function in the bi-directional RNN layer and the number of dimensions in the hidden state of RNNs is 200.
+Adam\cite{kingma2014adam} with 50 epochs, the batch size of 32 and learning rate of 0.001 was used for optimization and the early stopping strategy was applied.\footnote{source code of all methods is available at \url{https://github.com/teanalab/eCoaching-Text-Segmentation}}
 
@@
     \includegraphics[width=0.95\textwidth]{figures/CRNN.eps}
-    \caption{\textbf{Architecture of convolutional recurrent neural networks for automatic segmentation of e-Coaching emails}}
+    \caption{\textbf{Architecture of a convolutional recurrent neural network for automated segmentation of e-Coaching emails}}
     \label{fig:crnn}
@@
 \subsection*{\textit{Evaluation metrics}}
-We report standard metrics of precision, recall and F1-measure to evaluate the performance of the classifiers.\cite{aas1999text} Accuracy is not reported as a performance metric because it is highly sensitive to the prior class probabilities and does not fully describe the actual difficulty of the decision problem, when highly unbalanced datasets are involved.
-The results are reported based on 5 fold cross-validation and weighted macro-averaging over the folds, \hl{in which each fold was used as a test set and remaining 4 folds was utilized as a train set}.
-We also estimate the area under \hl{the precision-recall curve (AUPR) metric due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets.} \cite{davis2006relationship}
+We report standard metrics of precision, recall and F1-measure to evaluate the performance of the classifiers.\cite{aas1999text} Accuracy is not reported as a performance metric, since it is highly sensitive to the prior class probabilities and does not fully describe the actual difficulty of the decision problem, when highly unbalanced datasets are involved.
+The results are reported based on 5-fold cross-validation (one fold was used as a test set and the remaining 4 folds were utilized as a training set) and weighted macro-averaging over the folds.
+We also report the area under the precision-recall curve (AUPR), due to its effectiveness in measuring the performance of binary classifiers in the case of the datasets with imbalanced class distribution.\cite{davis2006relationship}
 
 \section*{Results}
-Our experiments span four dimensions.
-First, we report the optimal size of word embeddings and sliding window size of the MLP model.
-Second, results are reported with respect to ``new segment'' class as well as weighted average over ``new segment'' and ``same segment'' classes in Table~\ref{tab:result_base}.
-Third, classification performance of different machine learning methods are summarized in Table~\ref{tab:result_weighted_avg} when word embeddings or lexical features are used in combination with punctuation and POS features.
-Fourth, the impact of the individual features, as well as their combination on the segmentation task, is reported.
+Our experiments spanned three dimensions.
+First, we determined the optimal size of word embedding vectors and the sliding window of MLP.
+Second, we evaluated the performance of different methods with respect to detecting ``new segment'' as well as the weighted average over ``new segment'' and ``same segment'' classes.
+Third, we assessed the impact of different types of features as well as their combination on the performance of different machine learning methods on the e-Coaching email segmentation task.
 
@@
     \includegraphics[width=0.95\textwidth]{figures/mlp-and-vector.eps}
-    \caption{\textbf{Performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and corpus word embeddings with GloVe and word2vec models (left).
-MLP model on e-Coaching email segmentation by varying the sliding window size 2n (right).}}
+    \caption{\textbf{F1-measure of CRNN on the task of e-Coaching email segmentation by varying the number of dimensions in pre-trained and corpus-based GloVe and word2vec embeddings (left).
+F1-measure of MLP on the task of e-Coaching email segmentation by varying the size of the sliding window (right).}}
     \label{fig:embedding-dimension-mlp}
@@
 
-Figure~\ref{fig:embedding-dimension-mlp} (left) shows the performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and corpus word embeddings with GloVe \footnote{https://nlp.stanford.edu/projects/glove/} and word2vec models.
-It was observed that best performance is achieved with pre-trained 300-dimensional word2vec word vectors when three types of features are used together.
-Therefore, we report our results with word2vec 300-dimensional word vectors for all deep learning models used in this study.
-For MLP models, the first layer input was prepared by the summation of the first $n$ word vectors in the sliding window and the last $n$ word vectors when a sliding window contains $2n$ words or punctuations.
-Figure~\ref{fig:embedding-dimension-mlp} (right) demonstrates the performance of the MLP model on e-Coaching email segmentation by varying the size of the sliding window.
-It can be observed that the best performance of MLP is achieved when $n$ is 2.
-Therefore, results of MLP in the remaining experiments are reported with $n$ set to 2.
+Figure~\ref{fig:embedding-dimension-mlp} (left) illustrates the performance of CRNN on the task of e-Coaching email segmentation by varying the number of dimensions in pre-trained and corpus-based GloVe \footnote{https://nlp.stanford.edu/projects/glove/} and word2vec embeddings.
+We observed that the best performance is achieved with pre-trained 300-dimensional word2vec word vectors, when three types of features are used together.
+Therefore, we report the results for other deep learning models used in this study when 300-dimensional word2vec embedding vectors are utilized.
+The input layer of MLP consists of a sum of embeddings of $n$ words or punctuation marks before and the sum of embeddings of $n$ words or punctuation marks after the word or punctuation mark, which is the center of a sliding window of $2n$ words or punctuation marks.
+Figure~\ref{fig:embedding-dimension-mlp} (right) demonstrates the performance of MLP on e-Coaching email segmentation by varying the size of the sliding window.
+It can be observed that the best performance of MLP is achieved when the size of the sliding window is 4 (or $n=2$).
+Therefore, MLP results in the remaining experiments are reported when $n$ is set to 2.
 
@@
 \centering
-\caption{\textbf{Performance of CRF, MLP, BRNN and CRNN methods for identification of ``new segment'' class as well as weighted average over ``new segment'' and ``same segment'' classes when word embeddings or lexical features are used.
-The highest value for each performance metric is highlighted in bold.}}
+\caption{\textbf{Performance of CRF, MLP, BRNN and CRNN on ``new segment'' detection as well as the weighted average over ``new segment'' and ``same segment'' classes when only lexical features are used.
+The highest value for each performance metric is highlighted in boldface.}}
 \label{tab:result_base}
@@
 
-As follows from Table~\ref{tab:result_base}, CRNN outperforms all other models in terms of recall and F1-measure achieving 0.797 recall with 0.785 F1-measure for new segment detection.
-CRNN also shows superior performance in all performance metrics for weighted average over ``new segment'' and ``same segment'' classes.
-BRNN demonstrates the lowest performance among all models in terms of precision and F1-Measure.
-On the other hand, MLP has the highest precision of 0.836 when word embeddings or lexical features are used to identify ``new segment''.
-CRF achieves 0.733 F1-Measure in new segment class and 0.984 F1-Measure overall, which corresponds to the second highest performance in identifying ``new segment'' as well as a weighted average over both classes.
-Experimental results indicate that performance of all classifiers as a weighted average over both classes is remarkably higher compared to ``new segment'' class, which is expected since 96.95\% of instances belong to the ``same segment'' class and 99.3\% of them are correctly classified.
-For example, as a weighted average over ``new segment'' and ``same segment'' classes, CRNN achieves 27.23\%, 23.71\% and 25.61\% higher precision, recall and F1-measure, respectively, compared to the ``new segment'' detection.
+As follows from Table~\ref{tab:result_base}, CRNN outperforms all other methods in terms of recall and F1-measure achieving 0.797 recall and 0.785 F1-measure for new segment detection.
+CRNN also shows superior performance according to all performance metrics calculated as a weighted average over ``new segment'' and ``same segment'' classes.
+BRNN demonstrates the lowest performance among all models in terms of precision and F1-measure.
+On the other hand, MLP demonstrated the highest precision of 0.836 when word embeddings or lexical features are used to identify ``new segment''.
+CRF achieves 0.733 F1-measure on new segment detection and 0.984 F1-measure overall, which corresponds to the second highest performance in identifying ``new segment''.
+CRF also demonstrated the second highest performance according to all metrics calculated as a weighted average over both classes.
+Experimental results indicate that the performance of all classifiers according to all metrics calculated as a weighted average over both classes is significantly higher than their performance on ``new segment'' detection, which is expected since 96.95\% of instances belong to the ``same segment'' class and 99.3\% of them are correctly classified.
+For example, CRNN achieves 27.23\%, 23.71\% and 25.61\% higher precision, recall and F1-measure calculated as a weighted average over ``new segment'' and ``same segment'' classes, compared to the ``new segment'' detection.
 \\
@@
 \centering
-\caption{\textbf{Performance of CRF, MLP, BRNN and CRNN methods for identification of ``new segment'' class as well as weighted average over ``new segment'' and ``same segment'' classes when all features are used together.
-The highest value for each performance metric
-is highlighted in bold.}}
+\caption{\textbf{Performance of CRF, MLP, BRNN and CRNN on ``new segment'' detection as well as the weighted average over ``new segment'' and ``same segment'' classes when all types of features are used together.
+The highest value for each performance metric is highlighted in boldface.}}
 \label{tab:result_weighted_avg}
@@
 
-Table~\ref{tab:result_weighted_avg} summarizes the results of all models for segmentation of e-Coaching emails when word embeddings or lexical features are used in combination with punctuation and POS features.
-Similar to results in Tables~\ref{tab:result_base}, CRNN demonstrates the highest performance among all methods achieving 0.864 recall with 0.825 F1-Measure for ``new segment'' detection and 0.990 precision with 0.989 recall and F1-Measure overall.
-BRNN and CRF show the lowest and second highest performance, respectively, for email segmentation among all methods.
-We observed that classification performance significantly improved for ``new segment'' class when word embeddings or lexical features are used in combination with punctuation and POS features.
-Precision increases by 3.96\%, -2.27\%, 12.71\% and 1.81\%; recall increases by 11.72\%, 19.73\%, 20.59\% and 8.41\%; and F1-measure increases by 8.05\%, 9.51\%, 16.22\% and 5.1\% for CRF, MLP, BRNN and CRNN methods, respectively, in new segment detection when all features are utilized together.
-Similarly, precision increases by 0.51\%, 0.41\%, 0.82\% and 0.41\%; recall increases by 0.41\%, 0.41\%, 0.72\% and 0.3\%; and F1-measure increases by 0.41\%, 0.41\%, 0.82\% and 0.3\% for CRF, MLP, BRNN and CRNN methods, respectively, in weighted average over ``new segment'' and ``same segment'' classes when word embeddings or lexical features are used in combination with punctuation and POS features.\\
+Table~\ref{tab:result_weighted_avg} summarizes the results of all models on the task of segmentation of e-Coaching emails when word embeddings or lexical features are used in combination with punctuation and POS features.
+Similar to results in Tables~\ref{tab:result_base}, CRNN demonstrates the best performance among all methods achieving 0.864 recall with 0.825 F1-measure for ``new segment'' detection and 0.990 precision with 0.989 recall and F1-measure overall.
+BRNN and CRF demonstrated the lowest and second highest performance on the task of email segmentation among all methods, respectively.
+We observed that classification performance significantly improved for ``new segment'' detection when lexical features are used in combination with punctuation and POS features.
+Specifically, precision increases by 3.96\%, -2.27\%, 12.71\% and 1.81\%; recall increases by 11.72\%, 19.73\%, 20.59\% and 8.41\%; and F1-measure increases by 8.05\%, 9.51\%, 16.22\% and 5.1\% for CRF, MLP, BRNN and CRNN methods, respectively, on new segment detection when all types of features are utilized together.
+Similarly, precision increases by 0.51\%, 0.41\%, 0.82\% and 0.41\%; recall increases by 0.41\%, 0.41\%, 0.72\% and 0.3\%; and F1-measure increases by 0.41\%, 0.41\%, 0.82\% and 0.3\% for CRF, MLP, BRNN and CRNN methods, respectively, as a weighted average over ``new segment'' and ``same segment'' classes when lexical features are used in combination with punctuation and POS features.\\
 
@@
 \centering
-\caption{\textbf{\hl{Area under the precision-recall curve (AUPR) values of all classifiers demonstrating the impact of word embeddings, punctuation and POS features on e-Coaching email segmentation.
-Highest AUPR value for each feature set across all models is highlighted in boldface.}}}
+\caption{\textbf{Area under the precision-recall curve (AUPR) values of all classifiers demonstrating the impact of different types of features on e-Coaching email segmentation performance.
+Highest AUPR value for each feature set across all models is highlighted in boldface.}}
 \label{tab:result_aupr}
@@
 
-Table~\ref{tab:result_aupr} shows the impact of individual features as well as their combination on the segmentation task.
-\hl{Influence of the punctuation and POS features is also consistent in AUPR values, which increase by 12.44\%, 14.4\%, 17.56\% and 6\% for CRF, MLP, BRNN and CRNN methods, respectively, when all features are used together.
-Individually, POS and punctuation features also improve the performance of all classifiers except BRNN and CRNN when POS features are used.
-CRF achieved the highest AUPR when punctuation or all features are used together.
-On the other hand, CRNN demonstrates the highest AUPR when word embeddings are used individually or combined with POS features.}
+Table~\ref{tab:result_aupr} shows the impact of different types of features as well as their combination on e-Coaching email segmentation performance.
+Punctuation and POS features have similar effect measured by the AUPR, which increases by 12.44\%, 14.4\%, 17.56\% and 6\% for CRF, MLP, BRNN and CRNN, respectively, when all features are used together.
+Individually, although punctuation features improve the performance of all classifiers, POS features improve the performance of only CRF and MLP.
+CRF achieved the highest AUPR when all types of features are used together.
+On the other hand, POS features degraded the AUPR of CRNN.
 
 \section*{Discussion}
-This study is the first effort to evaluate the automatic segmentation of e-Coaching emails.
+This study is the first effort to design and evaluate machine learning methods for automated segmentation of e-Coaching sessions.
 Experimental results indicate that CRNN is the best model among all machine learning methods considered for this study.
 CRNN achieved 0.989 F1-measure overall and 0.825 F1-measure for detecting ``new segment''.
-The robust performance of CRNN provides the evidence that deep learning models are capable of learning the transitions between MI behaviors from clinical exchanges.
-It also indicates that punctuation and POS features are important along with word embeddings for all machine learning methods employed.
-Although the domain of this study was intentionally quite small, we believe that our study is not limited to the e-Coaching domain and our conclusions can be extended to other domains, which require discourse segmentation.
+Robust performance of CRNN provides an evidence that deep learning models are capable of detecting the boundaries of patient and provider behaviors in email delivered behavioral interventions.
+Our experiments also highlight the importance of punctuation and POS features along with word embeddings for all machine learning methods employed this study.
+Although the domain of this study was intentionally focused, we believe that the proposed methods are not limited to e-Coaching and our conclusions can be generalized to other domains, which require discourse segmentation.
 
-Punctuation mark and POS features resulted in significant improvement in the performance of machine learning and all deep learning methods.
-Especially, punctuation features have the higher individual impact on model performance compared to POS features.
-\hl{In all cases, CRF and MLP methods performed better, when word embeddings are used in conjunction with punctuation and POS features.
-Punctuation features improved the performance of BRNN and CRNN while POS features lowered their AUPR values.
-We believe that the BRNN and CRNN performed poorly with POS features because POS tagging is a supervised learning solution that uses features like the previous and next word.
-Since we already considered neighbor words by utilizing bi-directional RNN, it failed to achieve good results with redundant information.
-We observed that MLP achieved the highest precision which may be related to the fact that MLP poorly learned ``new segment'' and misclassified new segment words to same segments in 30\%-40\% of the time.}
+Punctuation marks and POS features resulted in significant improvement in the performance of traditional machine learning and deep learning methods.
+Punctuation features had stronger individual impact on model performance than POS features.
+In all cases, CRF and MLP performed better, when word embeddings were used in conjunction with punctuation embeddings and POS features.
+Punctuation embeddings improved the performance of BRNN and CRNN measured by precision, recall and F1-measure, while POS features lowered their AUPR.
+%We believe that BRNN and CRNN performed poorly with POS features, since POS taggers already utilize previous and following word as features.
+Since neighboring words are also considered by bi-directional RNN, such redundancy did not contribute to better performance.
+%We observed that MLP achieved the highest precision which may be related to the fact that MLP poorly learned ``new segment'' and misclassified new segment words to same segments in 30\%-40\% of the time.
 
-The convolutional layer made a significant difference between the performance of CRNN and BRNN in MI session discourse segmentation.
-CRNN results in 22.46\% and 10.74\% higher F1-Measure in ``new segment'' detection and 1.02\% and 0.51\% higher F1-Measure overall compared to BRNN, when word embeddings and all other features are used, respectively.
-In CRNN, a convolution layer performs a series of convolutions and pooling operations, which produce a number of high-level important features from word embeddings.
-These high-level features are then utilized by the bidirectional RNN of the CRNN model, which results in a significant increase in performance.
-On the other hand, traditional BRNN model received words as input features and their word embeddings are directly utilized in the input layer.
+The convolution layer made a significant difference between the performance of CRNN and BRNN in MI session discourse segmentation.
+CRNN had 22.46\% and 10.74\% higher F1-measure in ``new segment'' detection and 1.02\% and 0.51\% higher F1-measure overall compared to BRNN, when word embeddings and all other features were used, respectively.
+In CRNN, a convolution layer performs a series of convolution and pooling operations, which produce a number of important high-level features from word and punctuation mark embeddings.
+These high-level features are then utilized by the bidirectional RNN layer in CRNN, which translates to a significant increase in performance.
+In contrast, BRNN utilizes the input word embeddings directly as features.
 
-Although punctuation mark plays an important role in segmentation boundary detection, a few errors were encountered by the presence of punctuation marks in boundary identification.
+Although punctuation marks play an important role in segmentation boundary detection, a few errors were triggered by the presence of punctuation marks.
 For example, a text segment from an e-Coaching email \textit{``A typical day in regards to fruit and vegetable has me eating about a serving at breakfast (our cafe has cut up fruit) and then maybe a piece of fruit later in the day or as a snack.
 Vegetable tends to be a side serving at lunch and dinner and I get celery or carrot cuts with dressing for a snack a lot of times.
-I could probably add some sort of vegetable into my breakfast (like spinach in an omelet) and snack on another piece of fruit when I am hungry rather than the junk food I tend to eat.''} was incorrectly segmented after the first sentence, when a punctuation mark was encountered.
+I could probably add some sort of vegetable into my breakfast (like spinach in an omelet) and snack on another piece of fruit when I am hungry rather than the junk food I tend to eat.''} was incorrectly segmented after the first sentence, when period was encountered.
 Similarly, additional information is a common cause for misclassification of an email segment into multiple segments.
-For instance, although the first sentence of the above email segment represents a positive commitment to behavior change, the next two sentences provide an additional information to support the patient's commitment.
+For instance, although the first sentence in the above email segment represents a positive commitment to behavior change, the next two sentences provide additional information to support the patient's commitment.
 
 The limitation of this study is that e-Coaching data is collected from a single medical institute; formatting, style and email segment can be different in other settings.
-Therefore, there is a need to replicate the experiments with different data sets.
-As our future work, we plan to evaluate our approach on the datasets from other behavioral interventions.
+Therefore, there is a need to replicate the experiments in this study with different data sets.
+As future work, we are planning to evaluate the methods presented in this work on the datasets from other behavioral interventions.
 
@@
 
-%Segmentation is the first step of qualitative analysis of unstructured clinical communications.
-
-Segmentation of e-Coaching emails is an integral part of developing and analyzing e-Coaching behavioral interventions.
-Although several studies have focused on segmentation problem in biomedical context, they are limited to segmenting clinical text in EHR into sections and sentences, with none of them considering segmentation of text into groups of MI behaviors in the context of MI session discourse analysis.
-By comparing the performance of machine learning methods for the task of segmentation of e-Coaching emails, we found out that CRNN provides the best performance in terms of all performance metrics.
+Segmentation is the first step of qualitative analysis of unstructured clinical communications, such as e-Coaching.
+Although several studies have focused on segmentation problem in biomedical context, they are limited to segmenting clinical text in EHR into sections and sentences.
+No previous studies considered the task of automated segmentation of clinical communications into groups of MI behaviors in the context of unstructured MI sessions.
+By comparing the performance of machine learning methods for the task of segmentation of e-Coaching emails, we found out that convolutional recurrent neural networks provides the best performance in terms of all performance metrics.
 Manual segmentation of e-Coaching data is a very resource-intensive and time-consuming task, which can significantly decrease the time and effort required to develop an effective behavioral intervention.
-Our proposed methods can help to identify textual segments corresponding to MI behaviors in unstructured clinical dialog, which can then be annotated with a classification model in a pipeline setting.
-Automated segmentation and analysis of e-Coaching emails can significantly decrease the time to identify effective communication strategies in e-mail based motivational interviewing.
+Our proposed methods can help to identify textual segments corresponding to MI behaviors in unstructured clinical dialog, which can then be annotated with an MI behavior annotation model in a pipeline setting.
+Automated segmentation and annotation of e-Coaching emails can significantly decrease the time to identify effective communication strategies in e-mail based motivational interviewing.
 

### 2018-11-26 paper-amia-summit-2019/manuscript.tex
initial final version by Alex

@@
 \textit{Communication science approaches to develop effective behavior interventions, such as motivational interviewing (MI), are limited by traditional qualitative coding of communication exchanges, a very resource-intensive and time-consuming process.
-This study focuses on the analysis of e-Coaching sessions, behavior interventions that are delivered via email and grounded in the principles of MI.
-A critical step towards automated qualitative coding of e-Coaching communication exchanges is segmentation of emails into fragments that correspond to MI behaviors.
-This study frames email segmentation task as a classification problem and utilizes word embeddings in conjunction with punctuation and part-of-speech features to address it.
-We experimentally evaluated conditional random fields (CRF), a traditional machine learning method, as well as deep learning methods, such as multilayer perceptron (MLP), bidirectional recurrent neural network (BRNN) and convolutional recurrent neural network (CRNN).
-Our results indicate that CRNN outperform CRF, MLP and BRNN for the task of email segmentation achieving 0.989 macro F1-score overall and 0.825 macro F1-score for new segment detection.}
+This study focuses on the analysis of e-Coaching sessions, behavior interventions delivered via email and grounded in the principles of MI.
+A critical step towards automated qualitative coding of e-Coaching sessions is segmentation of emails into fragments that correspond to MI behaviors.
+This study frames email segmentation task as a classification problem and utilizes word and punctuation mark embeddings in conjunction with part-of-speech features to address it.
+We evaluated the performance of conditional random fields (CRF) as well as multi-layer perceptron (MLP), bi-directional recurrent neural network (BRNN) and convolutional recurrent neural network (CRNN) for the task of email segmentation.
+Our results indicate that CRNN outperforms CRF, MLP and BRNN achieving 0.989 macro F1-score overall and 0.825 macro F1-score for new segment detection.}
 
@@
 
-The limitation of this study is that e-Coaching data is collected from a single medical institute; formatting, style and email segment can be different in other settings.
+%The limitation of this study is that e-Coaching data is collected from a single medical institute; formatting, style and email segment can be different in other settings.
 Therefore, there is a need to replicate the experiments in this study with different data sets.
@@
 No previous studies considered the task of automated segmentation of clinical communications into groups of MI behaviors in the context of unstructured MI sessions.
-By comparing the performance of machine learning methods for the task of segmentation of e-Coaching emails, we found out that convolutional recurrent neural networks provides the best performance in terms of all performance metrics.
-Manual segmentation of e-Coaching data is a very resource-intensive and time-consuming task, which can significantly decrease the time and effort required to develop an effective behavioral intervention.
+By comparing the performance of machine learning methods for the task of segmentation of e-Coaching emails, we found out that convolutional recurrent neural networks demonstrate the best performance in terms of all performance metrics.
+Manual segmentation of e-Coaching sessions is a very resource-intensive and time-consuming task, which can significantly decrease the time and effort required to develop an effective behavioral intervention.
 Our proposed methods can help to identify textual segments corresponding to MI behaviors in unstructured clinical dialog, which can then be annotated with an MI behavior annotation model in a pipeline setting.
-Automated segmentation and annotation of e-Coaching emails can significantly decrease the time to identify effective communication strategies in e-mail based motivational interviewing.
+%Automated segmentation and annotation of e-Coaching emails can significantly decrease the time to identify effective communication strategies in e-mail based motivational interviewing.
+As future work, we are planning to evaluate the methods presented in this work on the datasets from other behavioral interventions.
 
@@
 This study was supported by a grant from the National Institutes of Health, NIDDK R21DK108071, Carcone and Kotov, MPIs.
-We would like to thank the research staff and student assistants in the Department of Family Medicine and Public Health Sciences at Wayne State University School of Medicine for their help in preparing the training dataset.
+We would like to thank the student assistants in the Department of Family Medicine and Public Health Sciences at Wayne State University School of Medicine for their help in preparing the training dataset.
 

### 2018-11-26 paper-amia-summit-2019/manuscript.tex
checked typos before submission of the final manuscript

@@
 In this study, we focus on the analysis of email-delivered MI, or e-Coaching, to promote healthy eating among young adults.
-The e-Coaching dataset is composed of email correspondence between a MI counselor and the young adult patient.
+The e-Coaching dataset is composed of email correspondence between an MI counselor and the young adult patient.
 Unlike transcribed in-person discourse, email correspondence is not clearly segmented into codable speech acts (i.e., utterances).
@@
 The experimental dataset for this work was constructed from 49 e-Coaching sessions, which include 330 and 281 emails by e-Coaches and patients, respectively.
-Various statistics of experimental dataset are provided in Table~\ref{tab:datastat}.
+Various statistics of the experimental dataset are provided in Table~\ref{tab:datastat}.
 Each session represents an MI intervention delivered via email.
-Emails were segmented into 3,138 text fragments and and annotated using the Minority Youth-Sequential Coding of Process Exchanges (MYSCOPE),~\cite{carcone2013provider} a qualitative coding scheme to characterize patient-provider communication during MI sessions with 115 distinct behavior codes.
+Emails were segmented into 3,138 text fragments and annotated using the Minority Youth-Sequential Coding of Process Exchanges (MYSCOPE),~\cite{carcone2013provider} a qualitative coding scheme to characterize patient-provider communication during MI sessions with 115 distinct behavior codes.
 We consider email segmentation as a binary classification problem, in which each word or punctuation mark is annotated with a class label ``new segment'' or ``same segment'' to indicate whether it precedes a new segment or not.
@@
 \centering
-\caption{\textbf{Summary of statistics of experimental dataset.}}
+\caption{\textbf{Summary of statistics of the experimental dataset}}
 \label{tab:datastat}
@@
 embedding) is a form of distributed word representation, in which each word is associated with a vector in a low-dimensional vector space.
-Embeddings have been previously shown to effectively capture semantic, syntactic and morphological properties of words.\cite{pennington2014glove, mikolov2013distributed} For experiments reported in this paper, we utilized word embeddings pre-trained on Google News corpus consisting of 1.6 billion words using word2vec software package.\footnote{\label{fn:word2vec}https://code.google.com/p/word2vec/} For words or punctuation marks, which do not have pre-trained embeddings, we utilized the embeddings of the same dimensionality trained on the experimental dataset.
+Embeddings have been previously shown to effectively capture semantic, syntactic and morphological properties of words.\cite{pennington2014glove, mikolov2013distributed} For experiments reported in this paper, we utilized word embeddings pre-trained on Google News corpus consisting of 1.6 billion words using word2vec software package.\footnote{\label{fn:word2vec}\url{https://code.google.com/p/word2vec/}} For words or punctuation marks, which do not have pre-trained embeddings, we utilized the embeddings of the same dimensionality trained on the experimental dataset.
 CRF utilized lexical features, POS tags and the preceding label.
@@
 
-Figure~\ref{fig:embedding-dimension-mlp} (left) illustrates the performance of CRNN on the task of e-Coaching email segmentation by varying the number of dimensions in pre-trained and corpus-based GloVe \footnote{https://nlp.stanford.edu/projects/glove/} and word2vec embeddings.
+Figure~\ref{fig:embedding-dimension-mlp} (left) illustrates the performance of CRNN on the task of e-Coaching email segmentation by varying the number of dimensions in pre-trained and corpus-based GloVe \footnote{\url{https://nlp.stanford.edu/projects/glove/}} and word2vec embeddings.
 We observed that the best performance is achieved with pre-trained 300-dimensional word2vec word vectors, when three types of features are used together.
@@
 CRNN achieved 0.989 F1-measure overall and 0.825 F1-measure for detecting ``new segment''.
-Robust performance of CRNN provides an evidence that deep learning models are capable of detecting the boundaries of patient and provider behaviors in email delivered behavioral interventions.
+The robust performance of CRNN provides an evidence that deep learning models are capable of detecting the boundaries of patient and provider behaviors in email delivered behavioral interventions.
 Our experiments also highlight the importance of punctuation and POS features along with word embeddings for all machine learning methods employed this study.
@@
 Punctuation marks and POS features resulted in significant improvement in the performance of traditional machine learning and deep learning methods.
-Punctuation features had stronger individual impact on model performance than POS features.
+Punctuation features had a stronger individual impact on model performance than POS features.
 In all cases, CRF and MLP performed better, when word embeddings were used in conjunction with punctuation embeddings and POS features.
@@
 Segmentation is the first step of qualitative analysis of unstructured clinical communications, such as e-Coaching.
-Although several studies have focused on segmentation problem in biomedical context, they are limited to segmenting clinical text in EHR into sections and sentences.
+Although several studies have focused on the segmentation problem in biomedical context, they are limited to segmenting clinical text in EHR into sections and sentences.
 No previous studies considered the task of automated segmentation of clinical communications into groups of MI behaviors in the context of unstructured MI sessions.

### 2018-11-26 paper-amia-summit-2019/manuscript.tex
final version by Alex

--- 2018-11-26 paper-amia-summit-2019/manuscript.tex
@@
 
-\author{Mehedi Hasan, BS$^{1}$\footnote{Authors provided an equal contribution.
+\author{Mehedi Hasan, BS$^{1}$\footnote{These authors provided an equal contribution.
 \label{footnote1}}, Alexander Kotov, PhD$^{1}$\textsuperscript{\ref{footnote1}}, Sylvie Naar, PhD$^{2}$, Gwen L.
@@
 We evaluated the performance of conditional random fields (CRF) as well as multi-layer perceptron (MLP), bi-directional recurrent neural network (BRNN) and convolutional recurrent neural network (CRNN) for the task of email segmentation.
-Our results indicate that CRNN outperforms CRF, MLP and BRNN achieving 0.989 macro F1-score overall and 0.825 macro F1-score for new segment detection.}
+Our results indicate that CRNN outperforms CRF, MLP and BRNN achieving 0.989 weighted macro-averaged F1-measure and 0.825 F1-measure for new segment detection.}
 
@@
 The specific provider behaviors responsible for the elicitation of change talk, however, are less clear and may vary by treatment context.
-Thus, a current focus of MI research is understanding which provider behaviors in which contexts lead to patient change talk.
+Thus, current MI research is focused on understanding which provider behaviors and in which contexts lead to patient change talk.
 To date, no study has examined this process in email-delivered MI.
@@
 The e-Coaching dataset is composed of email correspondence between a MI counselor and the young adult patient.
-Unlike transcribed in-person discourse, email correspondence is not clearly segmented into codable speech acts (i.e., utterances).
+Unlike transcribed in-person exchanges, email correspondence is not clearly segmented into codable speech acts (i.e., utterances).
 Thus, the unstructured nature of e-Coaching exchanges poses a unique set of analytic challenges.
@@
 For example, an e-Coach usually responds to several previous patient statements in one email.
-In contrast, in a traditional, in-person MI session, an utterance is assumed to be a response to the immediately preceding utterance.
-\item Discourse segments in e-Coaching do not have a clear breakpoint, such as the end of a sentence or paragraph.
-One sentence might be divided into fragments corresponding to multiple MI behaviors.
+In contrast, in a traditional, in-person MI session, each utterance is assumed to be a response to an immediately preceding utterance.
+\item Discourse segments in e-Coaching do not have a clear breakpoint, such as the end of a sentence or a paragraph.
+One sentence may be divided into fragments corresponding to multiple MI behaviors.
 On the other hand, an MI behavior may comprise several sentences.
@@
 
-Figure~\ref{fig:text-segment} illustrates a segmentation of an e-Coaching exchange, in which the first sentence is segmented into 2 MI behavior fragments, while the fourth and fifth segments correspond to one and three sentences, respectively.
+Figure~\ref{fig:text-segment} illustrates a segmentation of an e-Coaching exchange, in which the first sentence is segmented into 2 MI behavior fragments, while the fourth and fifth MI behavior fragments comprise one and three sentences, respectively.
 Segmentation of e-Coaching exchanges constitutes a special case of clinical discourse analysis \cite{webber2012discourse} aimed at better understanding the effective communication strategies specific to this type of behavioral interventions.
@@
 The goal of this study is to assess the effectiveness of deep learning methods for the task of automated segmentation of e-Coaching emails into textual fragments corresponding to individual patient and provider behaviors.
-For this study, we utilized the data from MENU GenY (Making Effective Nutrition Choices for Generation Y) \cite{alexander2017motivations}, a web-delivered public health intervention with email-based coaching to encourage increased fruit and vegetable intake among young adults, aged 21–-30.
+For this study, we utilized the data from MENU GenY (Making Effective Nutrition Choices for Generation Y) \cite{alexander2017motivations}, a web-delivered public health intervention with email-based coaching to encourage increased fruit and vegetable intake among young adults, aged 21–30.
 A secondary goal of the MENU GenY project was to identify the specific communication strategies used by e-Coaches to elicit change talk for healthier eating among young adult patients.
-Segmentation of clinical conversation in the context of electronically delivered interventions into groups of MI behaviors, is traditionally performed manually by MI researchers, which significantly slows down their qualitative analysis.
-This paper is the first work to evaluate the empirical effectiveness of deep learning architectures in addressing \textit{the problem of discourse segmentation in the context of e-mail-based behavioral interventions}.
+Segmentation of clinical conversation in the context of electronically delivered interventions into groups of MI behaviors is traditionally performed manually by MI researchers, which significantly slows down its qualitative analysis.
+\textit{This paper is the first work to evaluate the empirical effectiveness of deep learning architectures in addressing the problem of discourse segmentation in the context of email-based behavioral interventions}.
 
-Specifically, we introduce and evaluate the effectiveness of distributed word representations (i.e.
-word embeddings) as well as lexical, punctuation and part-of-speech (POS) features in conjunction with both traditional supervised machine learning methods, such as linear-chain Conditional Random Fields (CRF)\cite{lafferty2001conditional} and deep learning methods, such as Multi-Layer Perceptron (MLP),\cite{rumelhart1986learning} Bidirectional Recurrent Neural Network (BRNN)\cite{schuster1997bidirectional} and Convolutional Recurrent Neural Network (CRNN),\cite{treviso2017sentence} to find the best performing method and feature combination.
+Specifically, we evaluate the effectiveness of distributed representations (i.e.
+embeddings) of words and punctuation marks as well as part-of-speech (POS) features in conjunction with both traditional supervised machine learning methods, such as linear-chain Conditional Random Fields (CRF)\cite{lafferty2001conditional} and deep learning methods, such as Multi-Layer Perceptron (MLP),\cite{rumelhart1986learning} Bidirectional Recurrent Neural Network (BRNN)\cite{schuster1997bidirectional} and Convolutional Recurrent Neural Network (CRNN),\cite{treviso2017sentence} to determine the best performing method and feature combination for the task of segmentation of e-Coaching emails into MI behaviors.
 
@@
 Prior work on textual segmentation in the biomedical domain primarily focused on sentence boundary detection \cite{griffis2016quantitative,kreuzthaler2015detection,treviso2017sentence} and segmentation of clinical documents in patients' electronic health records (EHR) into sections and headers.
-\cite{apostolova2009automatic,denny2009evaluation,tepper2012statistical,cho2002text} In particular, Maximum Entropy models \cite{tepper2012statistical} and Support Vector Machine (SVM) along with word-vector cosine similarity metric and several heuristics \cite{apostolova2009automatic} have been applied to identify specific sections in EHR, such as general patient information, medical history, procedures, findings, etc.
+\cite{apostolova2009automatic,denny2009evaluation,tepper2012statistical,cho2002text} In particular, maximum entropy models \cite{tepper2012statistical} and Support Vector Machine (SVM) along with word vector similarity metrics and several heuristics \cite{apostolova2009automatic} have been applied to identify specific sections in EHR, such as general patient information, medical history, procedures, findings, etc.
 Denny et al.
@@
 
-SVM in conjunction with prosodic and part-of-speech features \cite{kreuzthaler2015detection} and Recurrent Convolutional Neural Networks \cite{griffis2016quantitative} have also been utilized for \textit{sentence boundary detection} in general text.
+SVM in conjunction with prosodic and part-of-speech features \cite{kreuzthaler2015detection} and recurrent convolutional neural networks \cite{griffis2016quantitative} have also been utilized for \textit{sentence boundary detection} in general text.
 Liu et al.
-\cite{liu2005using} demonstrated that a linear-chain CRF outperforms Hidden Markov and Maximum Entropy models for this task.
+\cite{liu2005using} demonstrated that a linear-chain CRF outperforms Hidden Markov and maximum entropy models for this task.
 
-Segmentation of e-Coaching emails is also different from traditional shallow discourse analysis \cite{galley2003discourse}, which besides identification of speech acts, also aims to determine the types of transitions between speech acts and labels speech acts with speakers who performed them in a multi-speaker conversation.
+Segmentation of e-Coaching emails is also different from traditional shallow discourse analysis \cite{galley2003discourse}, which besides identification of speech acts, also aims to determine the types of transitions between speech acts and label speech acts with the speakers who performed them in a multi-speaker conversation.
 The proposed methods will automate the process of segmenting clinical exchanges into MI behaviors, which will significantly reduce the time and resources required to perform such segmentation manually.
-Furthermore, these methods can be integrated with auto coding methods \cite{hasan2016study,kotov2015interpretable} to create a software pipeline for fully automated analysis of email-delivered behavioral interventions.
+Furthermore, these methods can be integrated with the automated MI behavior coding methods \cite{hasan2016study,kotov2015interpretable} to create a software pipeline for fully automated analysis of email-delivered behavioral interventions.
 
 \section*{Materials and methods}
-\subsection*{\textit{Data collection}}
+\subsection*{\textit{Dataset}}
 The experimental dataset for this work was constructed from 49 e-Coaching sessions, which include 330 and 281 emails by e-Coaches and patients, respectively.
-Various statistics of experimental dataset are provided in Table~\ref{tab:datastat}.
-Each session represents an MI intervention delivered via email.
-Emails were segmented into 3,138 text fragments and and annotated using the Minority Youth-Sequential Coding of Process Exchanges (MYSCOPE),~\cite{carcone2013provider} a qualitative coding scheme to characterize patient-provider communication during MI sessions with 115 distinct behavior codes.
-We consider email segmentation as a binary classification problem, in which each word or punctuation mark is annotated with a class label ``new segment'' or ``same segment'' to indicate whether it precedes a new segment or not.
+Various statistics of the experimental dataset are provided in Table~\ref{tab:datastat}.
+Each e-Coaching session represents a MI intervention delivered via email.
+Emails were segmented into 3,138 text fragments and annotated with the Minority Youth-Sequential Coding of Process Exchanges (MY-SCOPE),~\cite{carcone2013provider} a qualitative coding scheme to characterize patient-provider communication during MI sessions with 115 distinct behavior codes.
+Email segmentation can be considered as sequence tagging, which can be framed as a binary classification problem, in which each word or punctuation mark is annotated with one of the two class labels (``new segment'' or ``same segment'') to indicate whether it is a beginning of a new MI behavior segment or not.
 In total, the dataset consists of 95,777 words and 7,140 punctuation marks and includes 3,138 ``new segment'' and 99,779 ``same segment'' instances.
-In this study, we experimented with traditional machine learning methods, such as Conditional Random Fields (CRF)\cite{lafferty2001conditional} and deep learning methods, such as Multi-Layer Perceptron (MLP),\cite{rumelhart1986learning} Bidirectional Recurrent Neural Network (bidirectional RNN or BRNN)\cite{schuster1997bidirectional} and Convolutional Recurrent Neural Network (CRNN).\cite{treviso2017sentence} For an MLP model, training and testing samples were created based on a sliding window of $2n$ words or punctuation marks over each position (which could be a word or a punctuation mark) in a given input sequence, such that each instance consists of the $n$ words or punctuation marks after this position and $n$ words or punctuation marks prior to this position, including the position itself.
-Each sample is assigned a ``new segment'' or ``same segment'' label based on whether there should be a segment break after the current position or not.
-In the case of CRF, BRNN and CRNN models, an e-Coaching email was taken as an input sequence, POS tags and word embeddings of each word or punctuation mark were used as input and binary labels corresponding to ``new segment'' and ``same segment'' classification decisions were considered as the model output.
-In the gold standard, words or punctuations within the same segment were assigned the label of 0 and the last word or punctuation mark of a segment were assigned the label of 1 (an example of a sequence with assigned labels is provided in the $2^{nd}$ row of Table~\ref{tab:datastat}).
-\\
+In this study, we experimented with traditional machine learning methods, such as Conditional Random Fields (CRF)\cite{lafferty2001conditional} and deep learning methods, such as Multi-Layer Perceptron (MLP),\cite{rumelhart1986learning} Bi-directional Recurrent Neural Network (BRNN)\cite{schuster1997bidirectional} and Convolutional Recurrent Neural Network (CRNN).\cite{treviso2017sentence} In the case of MLP, training and testing samples were created based on a sliding window of $2n$ words or punctuation marks over each position (which could be a word or a punctuation mark) in a given input sequence, such that each sample consists of the $n$ words or punctuation marks after the current position and $n$ words or punctuation marks prior to the current position, including the position itself.
+In the case of CRF, BRNN and CRNN models, an e-Coaching email was taken as an input sequence, POS tags and embeddings of each word or punctuation mark were used as input and binary labels corresponding to ``new segment'' and ``same segment'' classification decisions were considered as the model output.
+In the gold standard, words or punctuations within the same segment were assigned the label of 0 and the last word or punctuation mark of a segment were assigned the label of 1 (an example of an input with the assigned labels is provided in the $2^{nd}$ row of Table~\ref{tab:datastat}).
 
@@
 \centering
-\caption{\textbf{Summary of statistics of experimental dataset.}}
+\caption{\textbf{Summary of statistics of the experimental dataset and example of segmented sequence.}}
 \label{tab:datastat}
@@
   \hline
-   \multirow{2}{*}{\textbf{Sessions}} & \multirow{2}{*}{\textbf{Instances}} & \multicolumn{2}{|c|}{\textbf{Class Segments}} & \multicolumn{2}{|c|}{\textbf{Tokens}} & \multicolumn{2}{|c|}{\textbf{Emails}} & \multicolumn{2}{|c|}{\textbf{Annotation}} \\\cline{3-10}
+   \multirow{2}{*}{\textbf{Sessions}} & \multirow{2}{*}{\textbf{Instances}} & \multicolumn{2}{|c|}{\textbf{Class labels}} & \multicolumn{2}{|c|}{\textbf{Tokens}} & \multicolumn{2}{|c|}{\textbf{Emails}} & \multicolumn{2}{|c|}{\textbf{Annotation}} \\\cline{3-10}
    &  & \textbf{new}  & \textbf{same} & \textbf{words} & \textbf{punctuations}  & \textbf{patient} & \textbf{provider} & \textbf{method}  & \textbf{codes} \\ \hline    
@@
 \subsection*{\textit{Features}}
-We utilized three types of features in conjunction with CRF, MLP, BRNN and CRNN methods: word embeddings as lexical features, punctuation and POS features.
+We utilized three types of features in conjunction with CRF, MLP, BRNN and CRNN: word embeddings as lexical features, punctuation and POS features.
 Syntactic abstractions of individual words, such as POS tags, have been previously shown to be effective features for similar natural language processing tasks.\cite{liu2005using,treviso2017sentence} To extract POS features, we pre-processed e-Coaching emails using the NLTK POS tagger\footnote{\url{https://www.nltk.org/}}.
-Punctuation marks, which correspond to one of the symbols \{`.', `,', `!', `?', `:', `;'\} between a pair of words, were also used as a feature, since punctuation marks designate the boundary of a sentence, clause or phrase and often also correspond to a segment boundary.\cite{cho2002text} For natural language processing (NLP) tasks, inputs are received as a text, in which individual words are as the basic lexico-semantic units.
+Punctuation marks, which correspond to one of the symbols \{`.', `,', `!', `?', `:', `;'\} between a pair of words, were also used as a feature, since punctuation marks designate the boundary of a sentence, clause or a phrase and often also correspond to a segment boundary.\cite{cho2002text} For natural language processing (NLP) tasks, inputs are received as textual fragments, in which individual words are as the basic lexico-semantic units.
 Therefore, it is important to represent a word in such a way that preserves all relevant lexical and semantic information.
-Dense real-valued vector (i.e.
-embedding) is a form of distributed word representation, in which each word is associated with a vector in a low-dimensional vector space.
-Embeddings have been previously shown to effectively capture semantic, syntactic and morphological properties of words.\cite{pennington2014glove, mikolov2013distributed} For experiments reported in this paper, we utilized word embeddings pre-trained on Google News corpus consisting of 1.6 billion words using word2vec software package.\footnote{\label{fn:word2vec}https://code.google.com/p/word2vec/} For words or punctuation marks, which do not have pre-trained embeddings, we utilized the embeddings of the same dimensionality trained on the experimental dataset.
+Embedding is a form of distributed representation, when each word is associated with a dense real-valued vector in low-dimensional space.
+Embeddings have been previously shown to effectively capture semantic, syntactic and morphological properties of words.\cite{pennington2014glove, mikolov2013distributed} For experiments reported in this paper, we utilized word embeddings pre-trained on Google News corpus consisting of 1.6 billion words using word2vec software package.\footnote{\label{fn:word2vec}\url{https://code.google.com/p/word2vec/}} For words or punctuation marks, which do not have pre-trained embeddings, we utilized the embeddings of the same dimensionality trained on the experimental dataset.
 CRF utilized lexical features, POS tags and the preceding label.
@@
 
-\textbf{Conditional Random Field}: CRF has been widely used in various sequence annotation NLP tasks, such as part-of-speech tagging.\cite{lafferty2001conditional, hirohata2008identifying} Unlike a maximum entropy Markov model, which uses per-state exponential models for conditional probability of the next state given a current state, CRF model directly estimates a distribution of the entire output sequence conditioned on the observation sequence.
-A traditional linear-chain CRF model is defined as a conditional probability distribution $p(y|x)$ of output sequence $y$, given input sequence $x$:
+\textbf{Conditional Random Fields (CRF)}: CRF has been widely used in various NLP tasks that involve sequence annotation, such as part-of-speech tagging.\cite{lafferty2001conditional, hirohata2008identifying} Unlike the maximum-entropy Markov model, which uses per-state exponential models for conditional probability of the next state given a current state, CRF model directly estimates a distribution of the entire output sequence conditioned on the observation sequence.
+A linear-chain CRF model is defined as a conditional probability distribution $p(y|x)$ of output sequence $y$, given input sequence $x$:
 
@@
 The optimal output sequence $y^*$ for input sequence $x$, $y^* = {arg\,max}_y p(y|x)$, is obtained efficiently using the Viterbi algorithm.
-In our experiments, the following features were utilized in conjunction with CRF models: i) current word or punctuation ii) next and previous 3 words or punctuations iii) binary feature indicating whether a word or punctuation is a special character (';', '?', '.', ',', '!', ':', etc.) or not iv) binary feature indicating whether a word is a title word or not (e.g.
+In our experiments, the following features were utilized in conjunction with CRF: i) current word or punctuation ii) next and previous 3 words or punctuations iii) binary feature indicating whether a word or punctuation is a special character (';', '?', '.', ',', '!', ':', etc.) or not iv) binary feature indicating whether a word is a title word or not (e.g.
 ``The'' is a title word but ``the'' is not) v) POS tags.
 
-\textbf{Multi-Layer Perceptron}: MLP is a neural network, which consists of multiple fully connected layers that map an input onto one or several outputs.\cite{rumelhart1986learning} Figure~\ref{fig:mlp} illustrates a multi-layer perceptron with a single hidden layer.
+\textbf{Multi-Layer Perceptron (MLP)}: MLP is a neural network, which consists of multiple fully connected layers that map an input to one or several outputs.\cite{rumelhart1986learning} Figure~\ref{fig:mlp} illustrates a multi-layer perceptron with a single hidden layer.
 MLPs have no cycles or loops.
@@
 In order to prevent over-fitting, we applied dropout (random masking of neurons) \cite{srivastava2014dropout} to fully connected layers during training.
-Dropout was also applied to fully connected layers in BRNN and CRNN.
+Dropout was also applied to a fully connected layer in CRNN.
 
@@
 
-\textbf{Bidirectional Recurrent Neural Network}: BRNN is a neural network designed to capture sequential patterns by considering both past  and future inputs as well as complex relationships between input features and output labels.\cite{schuster1997bidirectional} The hidden state of BRNN is an aggregation of the hidden states of a forward and backward recurrent neural networks (RNNs).
-Gated Recurrent Units (GRU)\cite{chung2014empirical} capable of handling variable size input sequence and having internal memory, which can be reset, were utilized as an RNN.
-%Figure~\ref{fig:crnn} represents the architecture of BRNN, in the case when a convolution layer is removed.
+\textbf{Bi-directional Recurrent Neural Network (BRNN)}: BRNN is a neural network designed to capture sequential patterns by considering both past  and future inputs as well as complex relationships between input features and output labels.\cite{schuster1997bidirectional} The hidden state of BRNN is an aggregation of the hidden states of a forward and backward recurrent neural networks (RNNs).
+Gated Recurrent Units (GRU)\cite{chung2014empirical} capable of handling variable size input sequence and having internal memory, which can be reset, were utilized as an RNN in this work.
 
-\textbf{Convolutional Recurrent Neural Network}: CRNN \cite{treviso2017sentence} shown in Figure~\ref{fig:crnn} is a deep neural network architecture, which combines convolutional and recurrent layers.
+\textbf{Convolutional Recurrent Neural Network (CRNN)}: CRNN \cite{treviso2017sentence} shown in Figure~\ref{fig:crnn} is a deep neural architecture, which combines convolutional and recurrent layers.
 Our implementation of CRNN consists of 5 layers: 1) input layer 2) embedding layer 3) convolution layer with max pooling 4) BRNN layer 5) fully connected layer with dropout and sigmoid output.
 E-coaching email exchanges are represented as a sequence of $m$ words and punctuations, which are fed into the input and embedding layers to produce a $m \times n_e$ matrix after fetching the embeddings for words and punctuations in the input sequence.
-This matrix corresponds to the distributed representation of an input email exchange, which contains rich morpho-syntactic information that can be utilized for its segmentation.
+This matrix is a distributed representation of an input email exchange, which contains rich morpho-syntactic information that can be utilized for its segmentation.
 When POS tags are utilized along with word embeddings, they are represented with a 10-dimensional vector, which is concatenated with 300-dimensional word embeddings to obtain new embedding vectors $n_e = [n_w;n_p]$ of size 310.
@@
 After applying $n_f$ different filters with zero-padding on both sides of the input text, $n_f$ features are produced by the convolution layer for each word.
-A max pooling over time operation was then performed to find the most significant features in a textual fragment.
-The bidirectional recurrent layer receives new features extracted by the convolution layer.
-Unidirectional RNNs are usually used to capture long-range dependencies in a sequence of observations.
-Bidirectional RNNs, on the other hand, are capable of capturing both past and future contexts through forward and backward traversals of a sequence.
-The purpose of the fully connected layer is to use the output of the bidirectional RNN layer for classifying each word or punctuation into ``new segment'' or ``same segment'' classes.
-Since a fully connected layer has a larger number of parameters, they are more likely to excessively co-adapt to  other parameters, which may result in over-fitting.
-To prevent this, we utilized dropout by randomly ignoring 50\% of the connections in the fully connected layer.
+A max pooling over time operation is then applied to find the most significant features in a textual fragment.
+The bi-directional recurrent layer receives new features extracted by the convolution layer.
+Unidirectional RNNs are typically utilized to capture long-range dependencies in a sequence of observations.
+Bi-directional RNNs, on the other hand, are capable of capturing both past and future contexts through forward and backward traversals of a sequence.
+The purpose of the fully connected layer in CRNN is to use the output of the bidirectional RNN layer for classifying each word or punctuation into ``new segment'' or ``same segment'' classes.
+Since a fully connected layer has a larger number of parameters, they are more likely to excessively co-adapt to other parameters in the network and result in over-fitting.
+To prevent this, we utilized dropout by randomly ignoring 50\% of the connections in the fully connected layer of CRNN.
 Finally, logistic sigmoid outputs the probability of classifying or labeling each word or punctuation mark with ``same segment'' class.
-We experimentally determined the optimal parameters using 5-fold cross-validation and found out that the best performance is achieved when filter length in the convolution layer is 7, number of filters is 100, max-pool size is 3, ReLU is used as an activation function in the convolution layer, hyperbolic tangent is used as an activation function in the bi-directional RNN layer and the number of dimensions in the hidden state of RNNs is 200.
+We experimentally determined the optimal parameters using 5-fold cross-validation and found out that the best performance is achieved when filter length in the convolution layer is 7, number of filters is 100, max pooling size is 3, ReLU is used as an activation function in the convolution layer, hyperbolic tangent is used as an activation function in the bi-directional RNN layer and the number of dimensions in the hidden state of RNNs is 200.
 Adam\cite{kingma2014adam} with 50 epochs, the batch size of 32 and learning rate of 0.001 was used for optimization and the early stopping strategy was applied.\footnote{source code of all methods is available at \url{https://github.com/teanalab/eCoaching-Text-Segmentation}}
@@
     \includegraphics[width=0.95\textwidth]{figures/CRNN.eps}
-    \caption{\textbf{Architecture of a convolutional recurrent neural network for automated segmentation of e-Coaching emails}}
+    \caption{\textbf{Architecture of a convolutional recurrent neural network for automated segmentation of e-Coaching emails into fragments corresponding to MI behaviors}}
     \label{fig:crnn}
@@
 \subsection*{\textit{Evaluation metrics}}
-We report standard metrics of precision, recall and F1-measure to evaluate the performance of the classifiers.\cite{aas1999text} Accuracy is not reported as a performance metric, since it is highly sensitive to the prior class probabilities and does not fully describe the actual difficulty of the decision problem, when highly unbalanced datasets are involved.
-The results are reported based on 5-fold cross-validation (one fold was used as a test set and the remaining 4 folds were utilized as a training set) and weighted macro-averaging over the folds.
+We report standard metrics of precision, recall and F1-measure to evaluate the performance of the classifiers.\cite{aas1999text} Accuracy is not reported as a performance metric, since it is highly sensitive to the distribution of prior class probabilities, which is skewed when datasets with unbalanced classes are involved.
+The results are reported based on 5-fold cross-validation (one fold was used as a test set and the remaining 4 folds were used as a training set) and weighted macro-averaging over the folds.
 We also report the area under the precision-recall curve (AUPR), due to its effectiveness in measuring the performance of binary classifiers in the case of the datasets with imbalanced class distribution.\cite{davis2006relationship}
@@
 Our experiments spanned three dimensions.
-First, we determined the optimal size of word embedding vectors and the sliding window of MLP.
+First, we determined the optimal sizes of word embedding vectors and the sliding window of MLP.
 Second, we evaluated the performance of different methods with respect to detecting ``new segment'' as well as the weighted average over ``new segment'' and ``same segment'' classes.
@@
 CRNN also shows superior performance according to all performance metrics calculated as a weighted average over ``new segment'' and ``same segment'' classes.
-BRNN demonstrates the lowest performance among all models in terms of precision and F1-measure.
-On the other hand, MLP demonstrated the highest precision of 0.836 when word embeddings or lexical features are used to identify ``new segment''.
-CRF achieves 0.733 F1-measure on new segment detection and 0.984 F1-measure overall, which corresponds to the second highest performance in identifying ``new segment''.
-CRF also demonstrated the second highest performance according to all metrics calculated as a weighted average over both classes.
+BRNN had the lowest performance among all models in terms of precision and F1-measure.
+On the other hand, MLP had the highest precision of 0.836 when lexical features are used to identify ``new segment''.
+CRF achieves 0.733 F1-measure, the second highest in identifying ``new segment''.
+CRF also demonstrated the second best performance among all models according to all metrics calculated as a weighted average over both classes.
 Experimental results indicate that the performance of all classifiers according to all metrics calculated as a weighted average over both classes is significantly higher than their performance on ``new segment'' detection, which is expected since 96.95\% of instances belong to the ``same segment'' class and 99.3\% of them are correctly classified.
 For example, CRNN achieves 27.23\%, 23.71\% and 25.61\% higher precision, recall and F1-measure calculated as a weighted average over ``new segment'' and ``same segment'' classes, compared to the ``new segment'' detection.
-\\
 
@@
 
-Table~\ref{tab:result_aupr} shows the impact of different types of features as well as their combination on e-Coaching email segmentation performance.
+Table~\ref{tab:result_aupr} illustrates the impact of different types of features as well as their combination on e-Coaching email segmentation performance.
 Punctuation and POS features have similar effect measured by the AUPR, which increases by 12.44\%, 14.4\%, 17.56\% and 6\% for CRF, MLP, BRNN and CRNN, respectively, when all features are used together.
@@
 CRF achieved the highest AUPR when all types of features are used together.
-On the other hand, POS features degraded the AUPR of CRNN.
+On the other hand, POS features degraded the AUPR of BRNN and CRNN.
 
@@
 Our experiments also highlight the importance of punctuation and POS features along with word embeddings for all machine learning methods employed this study.
-Although the domain of this study was intentionally focused, we believe that the proposed methods are not limited to e-Coaching and our conclusions can be generalized to other domains, which require discourse segmentation.
+Although the domain of this study was intentionally focused, we believe that the proposed methods are not limited to e-Coaching sessions and our conclusions can be generalized to other domains, which require discourse segmentation.
 
@@
 Punctuation features had stronger individual impact on model performance than POS features.
-In all cases, CRF and MLP performed better, when word embeddings were used in conjunction with punctuation embeddings and POS features.
-Punctuation embeddings improved the performance of BRNN and CRNN measured by precision, recall and F1-measure, while POS features lowered their AUPR.
+In all cases, CRF and MLP performed better, when word embeddings were used in conjunction with punctuations and POS features.
+Considering punctuations improved the performance of BRNN and CRNN measured by precision, recall and F1-measure, while POS features lowered their AUPR.
 %We believe that BRNN and CRNN performed poorly with POS features, since POS taggers already utilize previous and following word as features.
@@
 CRNN had 22.46\% and 10.74\% higher F1-measure in ``new segment'' detection and 1.02\% and 0.51\% higher F1-measure overall compared to BRNN, when word embeddings and all other features were used, respectively.
-In CRNN, a convolution layer performs a series of convolution and pooling operations, which produce a number of important high-level features from word and punctuation mark embeddings.
+In CRNN, a convolution layer performs a series of convolution and pooling operations, which produce a number of important high-level features from input embeddings.
 These high-level features are then utilized by the bidirectional RNN layer in CRNN, which translates to a significant increase in performance.
-In contrast, BRNN utilizes the input word embeddings directly as features.
+In contrast, BRNN utilizes the input embeddings directly as features.
 
@@
 No previous studies considered the task of automated segmentation of clinical communications into groups of MI behaviors in the context of unstructured MI sessions.
-By comparing the performance of machine learning methods for the task of segmentation of e-Coaching emails, we found out that convolutional recurrent neural networks demonstrate the best performance in terms of all performance metrics.
-Manual segmentation of e-Coaching sessions is a very resource-intensive and time-consuming task, which can significantly decrease the time and effort required to develop an effective behavioral intervention.
-Our proposed methods can help to identify textual segments corresponding to MI behaviors in unstructured clinical dialog, which can then be annotated with an MI behavior annotation model in a pipeline setting.
-%Automated segmentation and annotation of e-Coaching emails can significantly decrease the time to identify effective communication strategies in e-mail based motivational interviewing.
-As future work, we are planning to evaluate the methods presented in this work on the datasets from other behavioral interventions.
+By comparing the performance of machine learning methods for the task of segmentation of e-Coaching emails, we found out that convolutional recurrent neural networks demonstrate the best performance in terms of most performance metrics.
+Manual segmentation of e-Coaching sessions is a very resource-intensive and time-consuming task, which can significantly decrease the time and effort required to develop effective behavioral interventions.
+Our proposed methods can help to identify textual segments corresponding to MI behaviors in unstructured clinical dialog, which can then be annotated with MI behavior annotation methods in a pipeline setting.
+Automated segmentation and annotation of e-Coaching emails can significantly decrease the time to identify effective communication strategies in email-based MI.
+As future work, we are planning to evaluate the methods presented in this work on the datasets from other types of behavioral interventions.
 

### 2018-11-26 paper-amia-summit-2019/manuscript.tex
finalize camera ready version of the submitted manuscript

@@
 In this study, we focus on the analysis of email-delivered MI, or e-Coaching, to promote healthy eating among young adults.
-The e-Coaching dataset is composed of email correspondence between a MI counselor and the young adult patient.
+The e-Coaching dataset is composed of email correspondence between an MI counselor and the young adult patient.
 Unlike transcribed in-person exchanges, email correspondence is not clearly segmented into codable speech acts (i.e., utterances).
@@
 Various statistics of the experimental dataset are provided in Table~\ref{tab:datastat}.
-Each e-Coaching session represents a MI intervention delivered via email.
+Each e-Coaching session represents an MI intervention delivered via email.
 Emails were segmented into 3,138 text fragments and annotated with the Minority Youth-Sequential Coding of Process Exchanges (MY-SCOPE),~\cite{carcone2013provider} a qualitative coding scheme to characterize patient-provider communication during MI sessions with 115 distinct behavior codes.
@@
 \centering
-\caption{\textbf{Summary of statistics of the experimental dataset and example of segmented sequence.}}
+\caption{\textbf{Summary of statistics of the experimental dataset and example of a segmented sequence}}
 \label{tab:datastat}
@@
 A one-dimensional (1D) convolution operation is utilized in this layer in our implementation of BRNN.
-In 1D convolution, one filter is responsible for extraction of one feature.
+In 1D convolution, one filter is responsible for the extraction of one feature.
 After applying $n_f$ different filters with zero-padding on both sides of the input text, $n_f$ features are produced by the convolution layer for each word.
 A max pooling over time operation is then applied to find the most significant features in a textual fragment.
-The bi-directional recurrent layer receives new features extracted by the convolution layer.
+The bi-directional recurrent layer receives new features extracted from the convolution layer.
 Unidirectional RNNs are typically utilized to capture long-range dependencies in a sequence of observations.
@@
 \label{tab:result_base}
-  \begin{tabular}{|l|l|l|l|l|l|l|}
+  \begin{tabular}{|l|l|l|l|l|l|l|l|}
   \hline
-   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{New Segment}} & \multicolumn{3}{|c|}{\textbf{Overall}} \\\cline{2-7}
-   & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} \\ \hline    
- CRF & 0.782 & 0.691 & 0.733 & 0.983 & 0.984 & 0.984 \\ \hline
- MLP & \textbf{0.836} & 0.593 & 0.694 & 0.982 & 0.983 & 0.982 \\ \hline
- BRNN & 0.606 & 0.680 & 0.641 & 0.977 & 0.976 & 0.976 \\ \hline
- CRNN & 0.775 & \textbf{0.797} & \textbf{0.785} & \textbf{0.986} & \textbf{0.986} & \textbf{0.986} \\ \hline
+   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{New Segment}} & \multicolumn{3}{|c|}{\textbf{Overall}} & \multirow{2}{*}{\textbf{AUPR}} \\\cline{2-7}
+   & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} & \\ \hline    
+ CRF & 0.782 & 0.691 & 0.733 & 0.983 & 0.984 & 0.984 & 0.780\\ \hline
+ MLP & \textbf{0.836} & 0.593 & 0.694 & 0.982 & 0.983 & 0.982 & 0.736 \\ \hline
+ BRNN & 0.606 & 0.680 & 0.641 & 0.977 & 0.976 & 0.976 & 0.655\\ \hline
+ CRNN & 0.775 & \textbf{0.797} & \textbf{0.785} & \textbf{0.986} & \textbf{0.986} & \textbf{0.986} & \textbf{0.818} \\ \hline
   \end{tabular}
@@
 \label{tab:result_weighted_avg}
- \begin{tabular}{|l|l|l|l|l|l|l|}
+ \begin{tabular}{|l|l|l|l|l|l|l|l|}
   \hline
-   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{New Segment}} & \multicolumn{3}{|c|}{\textbf{Overall}} \\\cline{2-7}
-   & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} \\ \hline    
- CRF & 0.813 & 0.772 & 0.792 & 0.988 & 0.988 & 0.988 \\ \hline
- MLP & \textbf{0.817} & 0.710 & 0.760 & 0.986 & 0.987 & 0.986 \\ \hline
- BRNN & 0.683 & 0.820 & 0.745 & 0.985 & 0.983 & 0.984 \\ \hline
- CRNN & 0.789 & \textbf{0.864} & \textbf{0.825} & \textbf{0.990} & \textbf{0.989} & \textbf{0.989} \\ \hline
+   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{New Segment}} & \multicolumn{3}{|c|}{\textbf{Overall}} & \multirow{2}{*}{\textbf{AUPR}} \\\cline{2-7}
+   & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} & \\ \hline    
+ CRF & 0.813 & 0.772 & 0.792 & 0.988 & 0.988 & 0.988 & \textbf{0.877} \\ \hline
+ MLP & \textbf{0.817} & 0.710 & 0.760 & 0.986 & 0.987 & 0.986 & 0.842 \\ \hline
+ BRNN & 0.683 & 0.820 & 0.745 & 0.985 & 0.983 & 0.984 & 0.770 \\ \hline
+ CRNN & 0.789 & \textbf{0.864} & \textbf{0.825} & \textbf{0.990} & \textbf{0.989} & \textbf{0.989} & 0.867 \\ \hline
   \end{tabular}
@@
 Specifically, precision increases by 3.96\%, -2.27\%, 12.71\% and 1.81\%; recall increases by 11.72\%, 19.73\%, 20.59\% and 8.41\%; and F1-measure increases by 8.05\%, 9.51\%, 16.22\% and 5.1\% for CRF, MLP, BRNN and CRNN methods, respectively, on new segment detection when all types of features are utilized together.
-Similarly, precision increases by 0.51\%, 0.41\%, 0.82\% and 0.41\%; recall increases by 0.41\%, 0.41\%, 0.72\% and 0.3\%; and F1-measure increases by 0.41\%, 0.41\%, 0.82\% and 0.3\% for CRF, MLP, BRNN and CRNN methods, respectively, as a weighted average over ``new segment'' and ``same segment'' classes when lexical features are used in combination with punctuation and POS features.\\
+Similarly, precision increases by 0.51\%, 0.41\%, 0.82\% and 0.41\%; recall increases by 0.41\%, 0.41\%, 0.72\% and 0.3\%; and F1-measure increases by 0.41\%, 0.41\%, 0.82\% and 0.3\% for CRF, MLP, BRNN and CRNN methods, respectively, as a weighted average over ``new segment'' and ``same segment'' classes when lexical features are used in combination with punctuation and POS features.
 
@@
 Punctuation marks and POS features resulted in significant improvement in the performance of traditional machine learning and deep learning methods.
-Punctuation features had stronger individual impact on model performance than POS features.
+Punctuation features had a stronger individual impact on model performance than POS features.
 In all cases, CRF and MLP performed better, when word embeddings were used in conjunction with punctuations and POS features.
